### 2021-03-18 paper/paper.tex
Add basic structure for paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath}
\usepackage{url}
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{
  Synthesizing Verified Components for Cyber Assured Systems Engineering 
  \thanks{
    Funded by DARPA Cyber Assured Systems Engineering (CASE)
  }
}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{
Konrad Slind, 
Junaid Babar, and 
Isaac Amundson}
\IEEEauthorblockA{
Trusted Systems Group   \\
Collins Aerospace       \\
Minneapolis, Minnesota  \\
}
\and
\IEEEauthorblockN{Eric Mercer}
\IEEEauthorblockA{Department of Computer Science\\
Brigham Young University\\
Provo, Utah}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\input{intro}

\section{Simple Example}
\input{example}

\section{Specification Language}
\input{specification}

\section{Verified Synthesis}
\input{synthesis}

\section{OSATE Integration for AADL}
\input{aadl-agree-osate}

\section{Conclusion}
\input{conclusion}

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section
\bibliographystyle{IEEEtran}
\bibliography{paper}
\end{document}

### 2021-03-18 paper/synthesis.tex
Add basic structure for paper

\subsection{Filters}
\subsection{Monitors}
\subsection{Gates}

### 2021-03-18 paper/example.tex
Add notes on content

Show a filter, monitor, and gate form Phase II and indicate each of their intended roles.

### 2021-03-18 paper/intro.tex
Add notes on content

Set the context for this work: define system level theorem that we intend to prove with the added components? Big picture of the overall goal of the project.

Components (complexity (and expressiveness) increases along the way):
* Filter: takes a stream of data and delivers a new stream of data where the property is enforced--reasons over an infinite stream of date. Connects to the system correctness over traces. Predicate on a single piece of data, paying no attention to previous history--no temporal awareness, deciting if it passes or not. Can be very secure and very efficient.
* Monitor: Captures a relation over time on the data. Is able to reason about temporal properties. Supports multiple inputs and is able to do arbitrary computation for complex output over the inputs.

The attestation gate is an aside to show how the tools we have are rich enough to create other complex things that blend the two: filter and monitor.

This paper addresses only the verified components. Here is what we verify:

* The resulting assembly code exactly implements the specification--preserves the meaning of the original specification in the actual assembly code.

-- Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately. In the later section we give an example of a system that uses seL4 with a pacer to schedule over different domains to provide memory isolation. Scheduling atd the relation to the trace semantics is not part of this paper.

\cite{kalhauge2018sound}

### 2021-03-31 paper/related-work.tex
Add some related work on synthesis

Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, composition1, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}. Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}. Algorithms for actual component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}. These synthesis algorithms generate code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.

### 2021-04-01 paper/related-work.tex
Add contract work

@@
 These synthesis algorithms generate code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.
+
+Contracts are similar to assume-guarantee reasoning but are targeted to programming languages.
+Contracts are often more expressive than assume-guarantee reasoning and can not only be stateful but higher-order \cite{10.1145/583852.581484}.
+As contracts are often written in the target language, synthesis is not a problem for monitoring but comes with significant overhead \cite{10.1007/978-3-642-28869-2_11}.
+A monitor for a contract can be removed when it can be statically proved that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.

### 2021-04-01 paper/filters.tex
Add contig types

\newcommand{\konst}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\imp}{\Rightarrow}
\newcommand{\lval}{\ensuremath{\mathit{lval}}}
\newcommand{\set}[1]{\ensuremath{\{ {#1} \}}}
\newcommand{\kstar}[1]{\ensuremath{{#1}^{*}}}
\newcommand{\Lang}[1]{\ensuremath{{\mathcal L}({#1})}}
\newcommand{\LangTheta}[1]{\ensuremath{{\mathcal L}_{\theta}({#1})}}
\newcommand{\itelse}[3]{\mbox{$\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}$}}

% Latex trickery for infix div operator, from stackexchange

\makeatletter
\newcommand*{\bdiv}{%
  \nonscript\mskip-\medmuskip\mkern5mu%
  \mathbin{\operator@font div}\penalty900\mkern5mu%
  \nonscript\mskip-\medmuskip
}
\makeatother

A \emph{string} is an array of bytes. A \emph{language} is a set of strings. A \emph{recognizer} for a language determines if a string belongs to the language.  A language \emph{filter} uses a recognizer to pass an input string unchanged to the output if the string belongs to the given language; otherwise, it outputs an empty string.

Languages in this work are formally defined by \emph{contiguity types}. A contiguity type specifies data as a laid out side-by-side array of elements in a string. Every contiguity type has an associated recognizer that precisely recognizes strings in the language. Filter synthesis translates the contiguity type recognizer into CakeML in a way that preserves the precision of the recognizer to guarantees that the CakeML recognizer recognizes the same language as the contiguity type recognizer from which it is derived. This statement is the main result of the paper for filters.

\subsection{Contiguity Types}

See \cite{contiguity} for a complete discussion. Contiguity types assume that a string is the \emph{result} of a
map from structured data. Figure \ref{contig-types} is the syntax. Contiguity types start with common base types (booleans,
characters, signed and unsigned integers, \etc) and are closed under
the construction of records, arrays, and unions. \footnote{We use
the terms ``\emph{contiguity type}, \konst{contig}, and $\tau$
interchangeably.}

\begin{figure}
\[
\begin{array}{rcl}
 \mathit{base} & = & \konst{bool} \mid \konst{char} \mid \konst{u8} \mid \konst{u16} \mid \konst{u32} \\
               & \mid & \konst{u64}  \mid \konst{i16} \mid \konst{i32} \mid \konst{i64} \mid \konst{float} \mid \konst{double} \\ \\

 \tau & = & \mathit{base} \\
      & \mid & \konst{Recd}\; (f_1 : \tau_1) \ldots (f_n : \tau_n) \\
      & \mid & \konst{Array}\; \tau \; \mathit{exp} \\
      & \mid & \konst{Alt}\; \mathit{bexp}\; \tau_1 \; \tau_2 \\ \\

\mathit{lval} & = & \mathit{varname} \mid
                    \mathit{lval} \, [ \mathit{exp} ] \mid
                    \mathit{lval} . \mathit{fieldname} \\ \\

\mathit{exp} & = & \konst{Loc}\; \mathit{lval}
              \mid \konst{nLit}\; \konst{nat}
              \mid \mathit{constname} \\
             & \mid & \mathit{exp} + \mathit{exp}
                   \mid \mathit{exp} * \mathit{exp} \\ \\

\mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
              \mid  \konst{bLit}\; \konst{bool}
              \mid  \neg \mathit{bexp}
              \mid  \mathit{bexp} \land \mathit{bexp} \\
              & \mid & \mathit{exp} = \mathit{exp} 
              \mid  \mathit{exp} < \mathit{exp}
\end{array}
\]
\caption{Syntax for contiguity types.}
\label{contig-types}
\end{figure}

Arithmetic and Boolean expressions define dependent types. The expressions rely on the notion of an \emph{L-value} to reference values in previous fields.

Theorem statement: the traces accepted by the specification match the traces accepted by the implementation. If we have have uninterpreted functions, then it may not be precise as the agree analysis as something may be rejected by agree but accepted by the implementation. Need to think about that carefully.

\subsection{Matcher Synthesis}

### 2021-04-07 paper/contig.tex
Add sections and contig content

A \emph{string} is an array of bytes. A \emph{language} is a set of strings. A \emph{recognizer} for a language determines if a string belongs to the language.  A language \emph{filter} uses a recognizer to pass an input string unchanged to the output if the string belongs to the given language; otherwise, it outputs an empty string.

Languages in this work are formally defined by \emph{contiguity types}. A contiguity type specifies data as a laid out side-by-side array of elements in a string. Every contiguity type has an associated recognizer that precisely recognizes strings in the language.

\subsection{Contiguity Types}

See \cite{contiguity} for a complete discussion. Contiguity types assume that a string is the \emph{result} of a
map from structured data. Figure \ref{contig-types} is the syntax. Contiguity types start with common base types (booleans,
characters, signed and unsigned integers, \etc) and are closed under
the construction of records, arrays, and unions. \footnote{We use
the terms ``\emph{contiguity type}, \konst{contig}, and $\tau$
interchangeably.}

\begin{figure}
\[
\begin{array}{rcl}
 \mathit{base} & = & \konst{bool} \mid \konst{char} \mid \konst{u8} \mid \konst{u16} \mid \konst{u32} \\
               & \mid & \konst{u64}  \mid \konst{i16} \mid \konst{i32} \mid \konst{i64} \mid \konst{float} \mid \konst{double} \\ \\

 \tau & = & \mathit{base} \\
      & \mid & \konst{Recd}\; [ (f_1 : \tau_1) \ldots (f_n : \tau_n) ] \\
      & \mid & \konst{Array}\; \tau \; \mathit{exp} \\
      & \mid & \konst{Alt}\; \mathit{bexp}\; \tau_1 \; \tau_2 \\ \\

\mathit{lval} & = & f \mid
                    \mathit{lval} \, [ \mathit{exp} ] \mid
                    \mathit{lval} . f \\ \\
f            & = & \mathit{varName} \\ \\
\mathit{exp} & = & \konst{Loc}\; \mathit{lval}
              \mid \konst{nLit}\; \konst{nat}
              \mid \mathit{constname} \\
             & \mid & \mathit{exp} + \mathit{exp}
                   \mid \mathit{exp} * \mathit{exp} \\ \\

\mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
              \mid  \konst{bLit}\; \konst{bool}
              \mid  \neg \mathit{bexp}
              \mid  \mathit{bexp} \land \mathit{bexp} \\
              & \mid & \mathit{exp} = \mathit{exp} 
              \mid  \mathit{exp} < \mathit{exp}
\end{array}
\]
\caption{Syntax for contiguity types.}
\label{contig-types}
\end{figure}

Arithmetic and Boolean expressions define dependent types. The expressions rely on the notion of an \emph{L-value} to reference values in previous fields.

Theorem statement: the traces accepted by the specification match the traces accepted by the implementation. If we have have uninterpreted functions, then it may not be precise as the agree analysis as something may be rejected by agree but accepted by the implementation. Need to think about that carefully.

An L-value denotes an \emph{offset} from the beginning of a
data structure, plus a \emph{width}. In an R-value, an occurrence of an
L-value is mapped to the value of the patch of memory between
$\mathit{offset}$ and $\mathit{offset} + \mathit{width}$. For the
purpose of specifying message formats, it may not be immediately
obvious that a notation supporting assignment in imperative languages
can help, but there is indeed a form of assignment lurking.

The above explanation of L-values centers on indices into a byte
buffer; in the following we will give a mild variant of this: instead
of indices into the buffer, we lift out the designated slices. Thus,
given environments $\theta: \mathit{lval} \mapsto \konst{string}$ (binding
L-values to strings), $\Delta : \konst{string} \to \mathbb{N}$
(binding constant names to numbers) and functions
$\konst{toN}:\konst{string}\to\mathbb{N}$ and
$\konst{toB}:\konst{string}\to\konst{bool}$ (which interpret byte
sequences to numbers and booleans, respectively), expression
evaluation and boolean expression evaluation have conventional
definitions:

\[
\begin{array}{l}
\konst{eval} \; e =
\mathtt{case}\; e\
 \left\{
 \begin{array}{lcl}
    \konst{Loc}\; \lval & \Rightarrow & \konst{toN}(\theta(\lval)) \\
    \konst{nLit}\; n & \Rightarrow & n  \\
    \mathit{constname} & \Rightarrow & \Delta(\mathit{constname})  \\
    e_1 + e_2 & \Rightarrow & \konst{eval}\; e_1 + \konst{eval}\; e_2  \\
    e_1 * e_2 & \Rightarrow & \konst{eval}\; e_1 * \konst{eval}\; e_2  \\
  \end{array}
 \right.
 \\ \\
\konst{evalB} \; b =
\mathtt{case}\; b\
 \left\{
 \begin{array}{lcl}
    \konst{bLoc}\; \lval & \Rightarrow & \konst{toB}(\theta(\lval)) \\
    \konst{bLit}\; b & \Rightarrow & b \\
    \neg b & \Rightarrow & \neg(\konst{evalB} \; b)  \\
    b_1 \lor b_2 & \Rightarrow & \konst{evalB} \;b_1 \lor \konst{evalB} \;b_2   \\
    b_1 \land b_2 & \Rightarrow & \konst{evalB} \;b_1 \land \konst{evalB} \;b_2   \\
    e_1 = e_2 & \Rightarrow & \konst{eval} \;e_1 = \konst{eval} \;e_2   \\
    e_1 < e_2 & \Rightarrow & \konst{eval} \;e_1 < \konst{eval} \;e_2
  \end{array}
 \right.

\end{array}
\]

\begin{remark}[Partiality]
Expression evaluation is partial because there is no guarantee that
$\theta(\lval)$ is defined: an \lval{} being looked-up may not be in the
map $\theta$. Failure in evaluation is modelled by the \konst{option}
type, and must be handled in the semantics and the matching
algorithm. However error handling is omitted in the presentation since
it hampers readability. See the HOL4 formalization\footnote{
\texttt{https://github.com/loonwerks/splat/blob/monitors/contig/abscontigScript.sml}}
for full details.
\end{remark}

\subsection{Semantics}

 We now confess to misleading the reader: in spite of the notational
 similarity, a contiguity type is \emph{not} a type: it is a formal
 language. A type is usually understood to represent a set, or domain,
 of values, \eg, the type \konst{int32} represents a set of
 integers. In contrast, the contiguity type \konst{i32} represents the
 set of strings of width 32 bits. An element of a contiguity type can
 be turned into an element of a type by providing interpretations for
 all the strings at the leaves and interpreting the \konst{Recd} and
 \konst{Array} constructors into the corresponding type constructs. (A
 base contiguity type therefore serves mainly as a \emph{tag} to be
 interpreted as a width and also as an intended target type.) Thus,
 contiguity types sit---conveniently---between the types in a
 programming language and the strings used to make messages.

The semantics definition depends on a few basic notions familiar from
language theory: language concatenation, and iterated language
concatenation.

\begin{align*}
L_1 \cdot L_2 &= \set{w_1 w_2 \mid w_1 \in L_1 \land w_2 \in  L_2} \\
L^0 &= \varepsilon \\
L^{n+1} &= L \cdot L^n
\end{align*}

\begin{definition}[Semantics of contiguity types]

 In the following, we assume given an assignment $\theta$ adequate for
 evaluating expressions. If an expression evaluation fails, the
 language being constructed will be empty (e.g., $\emptyset$). $\LangTheta{\tau}$ is defined recursively as

\[
% \begin{array}{l}
%\LangTheta{\tau} =
\mathtt{case}\; \tau\
% \hspace*{3mm}
 \left\{
 \begin{array}{l}
 \mathit{base} \Rightarrow \set{s \mid \konst{len}(s) = \konst{width}(base)} \\
 \konst{Recd}\; [(f_1 : \tau_1) \ldots (f_n : \tau_n)]
      \Rightarrow \\
      \hspace*{5mm} \LangTheta{\tau_1} \cdot \ldots \cdot \LangTheta{\tau_n}
\\
 \konst{Array}\; \tau \; \mathit{exp} \Rightarrow  \\
  \hspace*{5mm}
 \left\{
 \begin{array}{ll}
    \LangTheta{\tau}^{\konst{eval}\;\theta\;\mathit{exp}} &
       \mathrm{if}\ \konst{eval}\;\theta\;\mathit{exp}\ \mathrm{succeeds} \\
    \emptyset & \text{otherwise}
 \end{array}
 \right.
\\
 \konst{Alt}\; \mathit{bexp}\;\tau_1\; \tau_2 \Rightarrow \\
  \hspace*{5mm}
 \left\{
 \begin{array}{ll}
    \LangTheta{\tau_1} & \mathrm{if}\ \konst{evalB}\;\theta\;\mathit{bexp} = \konst{true} \\
    \LangTheta{\tau_2} & \mathrm{if}\ \konst{evalB}\;\theta\;\mathit{bexp} = \konst{false} \\
    \emptyset          & \text{otherwise}
 \end{array}
 \right.
 \\
\end{array}
 \right.
%\end{array}
\]
\end{definition}

\subsection{Recognizer Synthesis}

A key result for contiguity types is that every type has a corresponding recognizer that recognizes strings belonging to the language defined by the type. In this work, the recognizer is a function \konst{match} that recursively traverses the contiguity type $\tau$ consuming the given input string $s$ as it constructs the environment $\theta$. The functions uses the environment $\theta$ to dynamically evaluate and resolve the expressions in $\tau$ during the traversal. The \konst{match} function returns a option type where $\konst{match}\; \tau\; s_1s_2\; \theta_o = \konst{SOME}(s_2, \theta)$ is a successful match having consumed $s_1$ leaving the postfix $s_2$ untouched and built the environment $\theta$ for $s_1$ according to $\tau$ from an empty initial environment $\theta_0$.

\begin{theorem}[Matcher soundness] 
$\konst{match}\; \tau\; s_1s_2\; \theta_0 = \konst{SOME}(s_2, \theta) \imp s_1 \in \LangTheta{\tau}$
\end{theorem}

### 2021-04-07 paper/filters.tex
Add sections and contig content

@@
-\newcommand{\konst}[1]{\ensuremath{\mathsf{#1}}}
-\newcommand{\imp}{\Rightarrow}
-\newcommand{\lval}{\ensuremath{\mathit{lval}}}
-\newcommand{\set}[1]{\ensuremath{\{ {#1} \}}}
-\newcommand{\kstar}[1]{\ensuremath{{#1}^{*}}}
-\newcommand{\Lang}[1]{\ensuremath{{\mathcal L}({#1})}}
-\newcommand{\LangTheta}[1]{\ensuremath{{\mathcal L}_{\theta}({#1})}}
-\newcommand{\itelse}[3]{\mbox{$\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}$}}
-
-% Latex trickery for infix div operator, from stackexchange
-
-\makeatletter
-\newcommand*{\bdiv}{%
-  \nonscript\mskip-\medmuskip\mkern5mu%
-  \mathbin{\operator@font div}\penalty900\mkern5mu%
-  \nonscript\mskip-\medmuskip
-}
-\makeatother
-
-A \emph{string} is an array of bytes.
-A \emph{language} is a set of strings.
-A \emph{recognizer} for a language determines if a string belongs to the language.
-A language \emph{filter} uses a recognizer to pass an input string unchanged to the output if the string belongs to the given language; otherwise, it outputs an empty string.
-
-Languages in this work are formally defined by \emph{contiguity types}.
-A contiguity type specifies data as a laid out side-by-side array of elements in a string.
-Every contiguity type has an associated recognizer that precisely recognizes strings in the language.
-Filter synthesis translates the contiguity type recognizer into CakeML in a way that preserves the precision of the recognizer to guarantees that the CakeML recognizer recognizes the same language as the contiguity type recognizer from which it is derived.
+Filter synthesis translates the contiguity type recognizer into CakeML in a way that preserves the precision of the recognizer to guarantee that the CakeML recognizer recognizes the same language as the contiguity type recognizer from which it is derived.
 This statement is the main result of the paper for filters.
-
-\subsection{Contiguity Types}
-
-See \cite{contiguity} for a complete discussion. Contiguity types assume that a string is the \emph{result} of a
-map from structured data. Figure \ref{contig-types} is the syntax. Contiguity types start with common base types (booleans,
-characters, signed and unsigned integers, \etc) and are closed under
-the construction of records, arrays, and unions. \footnote{We use
-the terms ``\emph{contiguity type}, \konst{contig}, and $\tau$
-interchangeably.}
-
-\begin{figure}
-\[
-\begin{array}{rcl}
- \mathit{base} & = & \konst{bool} \mid \konst{char} \mid \konst{u8} \mid \konst{u16} \mid \konst{u32} \\
-               & \mid & \konst{u64}  \mid \konst{i16} \mid \konst{i32} \mid \konst{i64} \mid \konst{float} \mid \konst{double} \\ \\
-
- \tau & = & \mathit{base} \\
-      & \mid & \konst{Recd}\; (f_1 : \tau_1) \ldots (f_n : \tau_n) \\
-      & \mid & \konst{Array}\; \tau \; \mathit{exp} \\
-      & \mid & \konst{Alt}\; \mathit{bexp}\; \tau_1 \; \tau_2 \\ \\
-
-\mathit{lval} & = & \mathit{varname} \mid
-                    \mathit{lval} \, [ \mathit{exp} ] \mid
-                    \mathit{lval} . \mathit{fieldname} \\ \\
-
-\mathit{exp} & = & \konst{Loc}\; \mathit{lval}
-              \mid \konst{nLit}\; \konst{nat}
-              \mid \mathit{constname} \\
-             & \mid & \mathit{exp} + \mathit{exp}
-                   \mid \mathit{exp} * \mathit{exp} \\ \\
-
-\mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
-              \mid  \konst{bLit}\; \konst{bool}
-              \mid  \neg \mathit{bexp}
-              \mid  \mathit{bexp} \land \mathit{bexp} \\
-              & \mid & \mathit{exp} = \mathit{exp} 
-              \mid  \mathit{exp} < \mathit{exp}
-\end{array}
-\]
-\caption{Syntax for contiguity types.}
-\label{contig-types}
-\end{figure}
-
-Arithmetic and Boolean expressions define dependent types. The expressions rely on the notion of an \emph{L-value} to reference values in previous fields.
-
-Theorem statement: the traces accepted by the specification match the traces accepted by the implementation.
-If we have have uninterpreted functions, then it may not be precise as the agree analysis as something may be rejected by agree but accepted by the implementation.
-Need to think about that carefully.
-
-\subsection{Matcher Synthesis}

### 2021-04-07 paper/system.tex
Add sections and contig content

A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs. The computation for the component in this work is defined entirely by a step function:
\[
\konst{stepFn} : \mathit{inports} \times \mathit{stateVars} \to \mathit{stateVars}
\]
The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
\[
\begin{array}{ll}
 \mathit{inportVals} & = \konst{readInputs}(); \\
 (v_1,\ldots,v_k) & = \konst{readState}() ; \\
 ({v_1}',\ldots,{v_k}') & = \konst{stepFn} (\mathit{inportVals},\mathit{stateVars}) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots,{v_k}');} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({v_1}',\ldots,{v_k}');} \\
\end{array}
\]
It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.

### 2021-04-23 paper/intro.tex
Add some related work on synthesis
Add draft of component specification

@@
+Here is the basic process:
+\begin{enumerate}
+  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
+  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
+  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
+  \item Contiguity types include dependent types and the ability to define when a type is malformed
+  \item Define a contract language that makes assumptions on input and guarantees properties on output
+  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
+  \item Show how the contract can be expressed in terms of a step function 
+  \item The step-function is lifted to traces
+  \item A contract defines a language over traces
+  \item A step function recognizes membership in that language (prefix closed)
+  \item Follow the same pattern used in contiguity types
+  \item The CakeML comes for free once the HOL4 theories are shown
+\end{enumerate}
+
 Set the context for this work: define system level theorem that we intend to prove with the added components? Big picture of the overall goal of the project.
@@
 
--- Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately.
+Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately.
 In the later section we give an example of a system that uses seL4 with a pacer to schedule over different domains to provide memory isolation.
@@
 
-\cite{kalhauge2018sound}
+Revised outline:
+
+Describe the system level context with the AGREE analysis that proves system level properties of streams.
+A system is a collection of components and connections.
+We assume a finite set of defined datatypes for the connections.
+A well-formed system is acyclic and provides some suitable set of input streams to consume.
+
+Every component is defined by its inputs, outputs, and a contract.
+The contract language is simplified AGREE: requires, guarantees, eq, and expressions that include \emph{pre}.
+Requires are pre-conditions assumed on the input.
+Guarantees are post-conditions the output must meet.
+A component consumes a stream of input and produces and stream of output.
+A well-formed component is also acyclic.
+
+For a component to by synthesized, its contract must be strong enough to define the meaning of every output (e.g., the value assumed by the output and under what conditions it assumes that value).
+
+The AGREE analysis reasons about the system with its component composition proving that all contracts are respected.
+Failures are counter-examples where some input requirement is violated.
+The analysis provides a system level guarantee.
+
+Component synthesis compiles the contract to imperative code that implements the contract.
+The claim is that the final compiled code preserves the meaning of the contract: it produces the same output stream for any input stream that conforms to its pre-condition.
+The synthesis is accomplished with the CakeML framework that provides the additional guarantee that the post-condition holds over an infinite stream (e.g., that it the component does not terminate and always satisfies the post-condition if the pre-condition holds---Hoare logic extended to statement that diverge).
+
+The semantics is synchronous data-flow. Computation takes zero time as does transport. The system must be well-formed.
+
+Motivate with some example: watchdog example on DNS monitoring for DOS attacks?
+Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.

### 2021-04-23 paper/related-work.tex
Add draft of component specification

@@
 A monitor for a contract can be removed when it can be statically proved that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+
+Lustre langugage definition and semantics \cite{10.1145/41625.41641,97300}. Classical clock directed compilation for synchronous dataflow \cite{10.1145/1379023.1375674,10.1145/2345141.2248426}.
+
+A denotational Kahn semantics for synchronous dataflow is in \cite{10.1007/978-3-540-45212-6_10}.
+
+State machine semantics can be added to synchronous dataflow \cite{10.1145/1086228.1086261}.
+The idea is to translate imperative constructs into equivalent synchronous dataflow constructs.
+The resulting Lustre can them be compiled to target platform.
+The goal is to provide a seamless connection between pure datalflow and pure control design.
+
+There is a fully verified compiler that takes Lustre and turns it into a binary executable \cite{10.1145/3140587.3062358}.
+The first compilation pass turns synchronous dataflow in Lustre to assembly.
+The compilation is specified and verified in Coq.
+The key is in combining infinite sequences of dataflow models with incremental manipulation of memories akin to an imperative model.
+CompCert is used on the backend to create the final rendered executable.

### 2021-04-23 paper/system.tex
Add draft of component specification

@@
-A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs.
-The computation for the component in this work is defined entirely by a step function:
+\begin{figure}
   \[
-\konst{stepFn} : \mathit{inports} \times \mathit{stateVars} \to \mathit{stateVars}
-\]
-The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
-\[
-\begin{array}{ll}
- \mathit{inportVals} & = \konst{readInputs}(); \\
- (v_1,\ldots,v_k) & = \konst{readState}() ; \\
- ({v_1}',\ldots,{v_k}') & = \konst{stepFn} (\mathit{inportVals},\mathit{stateVars}) ; \\
- \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots,{v_k}');} \\
- \multicolumn{2}{l}{\konst{writeOutputs}({v_1}',\ldots,{v_k}');} \\
+    \begin{array}{rcl}
+      \mathit{c}    & = & \konst{input}\ [(f : \tau)\ldots] \\
+                    &   & \konst{output}\ [(f : \tau)\ldots] \\
+                    &   & \konst{components}\ [(f : c)\ldots]\\ 
+                    &   & \konst{eq}\ [(\lval : \tau := exp) \dots] \\
+                    &   & \konst{assume}\ [bexp\ldots] \\
+                    &   & \konst{guarantee}\ [bexp\ldots] \\ \\
+
+      \mathit{lval} & = & f \mid \mathit{lval} \, [ \mathit{exp} ]
+                          \mid \mathit{lval} . f \\ \\
+
+      f             & = & \mathit{varName} \\ \\
+
+      \mathit{exp}  & = & \konst{Loc}\; \mathit{lval}
+                          \mid \konst{nLit}\; \konst{nat}
+                          \mid \mathit{constname} \\
+                    & | & \mathit{exp} + \mathit{exp}
+                          \mid \mathit{exp} * \mathit{exp} \\
+                    & | & (exp\ \rightarrow\ exp) \\
+                    & | & (\konst{pre}\ exp) \\
+                    & | & (\konst{ite}\ bexp\ exp\ exp)\\
+                    & | & bexp \\ \\
+
+      \mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
+                          \mid  \konst{bLit}\; \konst{bool}
+                          \mid  \neg \mathit{bexp}
+                          \mid  \mathit{bexp} \land \mathit{bexp} \\
+                    & | & \mathit{exp} = \mathit{exp} 
+                    \mid  \mathit{exp} < \mathit{exp}
 \end{array}
 \]
-It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+\caption{Syntax for component specifications.}
+\label{fig:syntax}
+\end{figure}
+
+A \emph{system} in this model of computation is a \emph{component} that is defined by a specification (see \figref{fig:syntax}).
+A specification defines inputs, outputs, sub-components, local variables ($\konst{eq}$), assumptions for pre-conditions, and guarantees for post-conditions.
+A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation).
+An $\mathit{lval}$ is a reference to an \emph{L-value} from compiler concepts and is an expression that can appear on the right hand side of an assignment.
+Alpha renaming is assumed so that every \emph{lval} is unique.
+
+An environment, $\theta: \mathit{lval} \mapsto \konst{string}$ binds L-values to strings.
+$\Delta : \konst{string} \to \mathbb{N}$ binds constant names to numbers.
+Functions $\konst{toN}:\konst{string}\to\mathbb{N}$ and $\konst{toB}:\konst{string}\to\konst{bool}$ interpret byte sequences to numbers and booleans, respectively.
+
+The semantics are synchronous data-flow defined over a sequence of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream.
+Expression evaluation is defined in the context of the environment stream in \figref{fig:eval}.
+\begin{figure*}
+\[
+\begin{array}{l}
+\konst{eval}\; i\; e =
+\mathtt{case}\; e\
+ \left\{
+ \begin{array}{lcl}
+    \konst{Loc}\; \lval & \Rightarrow & \konst{toN}(\theta^i(\lval)) \\
+    \konst{nLit}\; n & \Rightarrow & n  \\
+    \mathit{constname} & \Rightarrow & \Delta(\mathit{constname})  \\
+    e_1 + e_2 & \Rightarrow & \konst{eval}\; i \; e_1 + \konst{eval}\; i \; e_2  \\
+    e_1 * e_2 & \Rightarrow & \konst{eval}\; i \; e_1 * \konst{eval}\; i \; e_2  \\
+    e_1 \rightarrow e_2 & \Rightarrow &  \mathbf{if}\; i = 0\; \mathbf{then}\; \konst{eval}\; i \; e_1\; 
+                                         \mathbf{else}\; \konst{eval}\; i \; e_2 \\
+    (\konst{pre}\; e) & \Rightarrow &  \konst{eval}\; i-1 \; e
+  \end{array}
+ \right.
+ \\ \\
+\konst{evalB}\; i \; b =
+\mathtt{case}\; b\
+ \left\{
+ \begin{array}{lcl}
+    \konst{bLoc}\; \lval & \Rightarrow & \konst{toB}(\theta^i(\lval)) \\
+    \konst{bLit}\; b & \Rightarrow & b \\
+    \neg b & \Rightarrow & \neg(\konst{evalB} \; b)  \\
+    b_1 \lor b_2 & \Rightarrow & \konst{evalB}\; i \;b_1 \lor \konst{evalB}\; i \;b_2   \\
+    b_1 \land b_2 & \Rightarrow & \konst{evalB}\; i \;b_1 \land \konst{evalB}\; i \;b_2   \\
+    e_1 = e_2 & \Rightarrow & \konst{eval} \;e_1 = \konst{eval}\; i \;e_2   \\
+    e_1 < e_2 & \Rightarrow & \konst{eval} \;e_1 < \konst{eval}\; i \;e_2
+  \end{array}
+ \right.
+\end{array}
+\]
+\caption{Expression evaluation in the context of a stream on environments.}
+\label{fig:eval}
+\end{figure*}
+
+The initial environment stream only contains mappings for primary inputs along the entire stream.
+Stepping the component updates the current environment.
+In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of each component evaluation.
+Stepping a component is recursively defined similar to the eval functions but track the path for the environment binding.

### 2021-04-24 paper/system.tex
Refine system syntax

@@
                     &   & \konst{output}\ [(f : \tau)\ldots] \\
-                    &   & \konst{components}\ [(f : c)\ldots]\\ 
-                    &   & \konst{eq}\ [(\lval : \tau := exp) \dots] \\
-                    &   & \konst{assume}\ [bexp\ldots] \\
-                    &   & \konst{guarantee}\ [bexp\ldots] \\ \\
+                    &   & [\mathit{stmts}\ldots] \\
+                    &   & \konst{assume}\ [\mathit{bexp}\ldots] \\
+                    &   & \konst{guarantee}\ [\mathit{bexp}\ldots] \\ \\
 
-      \mathit{lval} & = & f \mid \mathit{lval} \, [ \mathit{exp} ]
-                          \mid \mathit{lval} . f \\ \\
+      \mathit{stmts}& = & \konst{subcomp}\ \lval\dots\ c \\
+                    & | & \konst{eq}\; f : \tau := \mathit{exp} \\ \\
+
+      \lval & = & f \mid \lval \, [ \mathit{exp} ]
+                          \mid \lval . f \\ \\
 
@@
 
-      \mathit{exp}  & = & \konst{Loc}\; \mathit{lval}
+      \mathit{exp}  & = & \konst{Loc}\; \lval
                           \mid \konst{nLit}\; \konst{nat}
@@
                           \mid \mathit{exp} * \mathit{exp} \\
-                    & | & (exp\ \rightarrow\ exp) \\
-                    & | & (\konst{pre}\ exp) \\
-                    & | & (\konst{ite}\ bexp\ exp\ exp)\\
-                    & | & bexp \\ \\
+                    & | & (\mathit{exp}\ \rightarrow\ \mathit{exp}) \\
+                    & | & (\konst{pre}\ \mathit{exp}) \\
+                    & | & (\konst{ite}\ \mathit{bexp}\ \mathit{exp}\ \mathit{exp})\\
+                    & | & \mathit{bexp} \\ \\
 
-      \mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
+      \mathit{bexp} & = & \konst{bLoc}\; \lval
                           \mid  \konst{bLit}\; \konst{bool}
@@
 A \emph{system} in this model of computation is a \emph{component} that is defined by a specification (see \figref{fig:syntax}).
-A specification defines inputs, outputs, sub-components, local variables ($\konst{eq}$), assumptions for pre-conditions, and guarantees for post-conditions.
+A specification defines inputs, outputs, statements, assumptions for pre-conditions, and guarantees for post-conditions.
+A statement can be a subcomponent definition with its input binding or a local variable definition.
+
 A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation).
-An $\mathit{lval}$ is a reference to an \emph{L-value} from compiler concepts and is an expression that can appear on the right hand side of an assignment.
-Alpha renaming is assumed so that every \emph{lval} is unique.
+An $\lval$ is a reference to an \emph{L-value} from compiler concepts and is an expression that can appear on the left hand side of an assignment.
+Alpha renaming is assumed so that every $\lval$ is unique.
 
-An environment, $\theta: \mathit{lval} \mapsto \konst{string}$ binds L-values to strings.
+An environment, $\theta: \lval \mapsto \konst{string}$ binds L-values to strings.
 $\Delta : \konst{string} \to \mathbb{N}$ binds constant names to numbers.
@@
 
-The semantics are synchronous data-flow defined over a sequence of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream.
+The semantics are synchronous data-flow on a single clock defined over a sequence of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream.
 Expression evaluation is defined in the context of the environment stream in \figref{fig:eval}.
+
 \begin{figure*}
@@
 In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of each component evaluation.
-Stepping a component is recursively defined similar to the eval functions but track the path for the environment binding.
+A component step traverses the component sequentially to update the environment recursing into subcomponents after binding inputs.
+Assume the component is written in dependency order.

### 2021-04-30 paper/abstract.tex
Add abstract

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in this cyber-resiliency in a Model-Based Systems Engineering environment. This paper discusses the automated model transformations that add in high-assurance components. These components are filters and monitors that prevent malicious input and detect supply chain attacks respectively. A formal specification defines each component and is used to verify that the component addresses system level cyber-requirements. Implementations are synthesized from the specification, and that synthesis is proven to preserve the exact meaning of the specification. The transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports a case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-03 paper/paper.tex
Add some related work on synthesis
Add contig types
Add sections and contig content
Add draft of component specification
Add abstract
Add acknowledgement section

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{amsfonts}
 \usepackage{url}
+\usepackage{comment}
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\newcommand{\konst}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\imp}{\Rightarrow}
+\newcommand{\lval}{\ensuremath{\mathit{lval}}}
+\newcommand{\set}[1]{\ensuremath{\{ {#1} \}}}
+\newcommand{\kstar}[1]{\ensuremath{{#1}^{*}}}
+\newcommand{\Lang}[1]{\ensuremath{{\mathcal L}({#1})}}
+\newcommand{\LangTheta}[1]{\ensuremath{{\mathcal L}_{\theta}({#1})}}
+\newcommand{\itelse}[3]{\mbox{$\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}$}}
+
+\newcommand{\figref}[1]{Fig.~\ref{#1}}
+
+% Latex trickery for infix div operator, from stackexchange
+
+\makeatletter
+\newcommand*{\bdiv}{%
+  \nonscript\mskip-\medmuskip\mkern5mu%
+  \mathbin{\operator@font div}\penalty900\mkern5mu%
+  \nonscript\mskip-\medmuskip
+}
+\makeatother
+
+\newcommand{\ie}{\textit{i.e.}}
+\newcommand{\eg}{\textit{e.g.}}
+\newcommand{\etal}{\textit{et al.}}
+\newcommand{\etc}{\textit{etc.}}
+\newcommand{\adhoc}{\textit{ad hoc}}
+
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
 
@@
 Konrad Slind, 
-Junaid Babar, and 
-Isaac Amundson}
+Junaid Babar, 
+Isaac Amundson, and
+David Hardin}
 \IEEEauthorblockA{
@@
 
-\section{Introduction}
-\input{intro}
+% \section{Introduction}
+% \input{intro}
 
-\section{Simple Example}
-\input{example}
+% \section{Simple Example}
+% \input{example}
 
-\section{Specification Language}
-\input{specification}
+\section{System Model and Semantics}
+\input{system}
 
-\section{Verified Synthesis}
-\input{synthesis}
+% \section{Contiguity Types}
+% \input{contig}
 
-\section{OSATE Integration for AADL}
-\input{aadl-agree-osate}
+% \section{Filters}
+% \input{filters}
 
-\section{Conclusion}
-\input{conclusion}
+% \section{Monitors}
+% \input{monitors}
+
+% \section{OSATE Integration for AADL}
+% \input{aadl-agree-osate}
+
+% \section{Related Work}
+% \input{related-work}
+
+% \section{Conclusion}
+% \input{conclusion}
+
+\section*{Acknowledgments}
+This work is funded by DARPA contract HR00111890001. The views, opinions and/or findings expressed are those of the authors and
+should not be interpreted as representing the official views or policies of the Department of Defense or the U.S. Government.
 
@@
 % references section
-\bibliographystyle{IEEEtran}
-\bibliography{paper}
+% \bibliographystyle{IEEEtran}
+% \bibliography{paper}
 \end{document}

### 2021-05-03 paper/system.tex
Add comment with old system model

@@
 Assume the component is written in dependency order.
+
+\begin{comment}
+  A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs.
+The computation for the component in this work is defined entirely by a step function:
+\[
+\konst{stepFn} : \mathit{inports} \times \mathit{stateVars} \to \mathit{stateVars}
+\]
+The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
+\[
+\begin{array}{ll}
+ \mathit{inportVals} & = \konst{readInputs}(); \\
+ (v_1,\ldots,v_k) & = \konst{readState}() ; \\
+ ({v_1}',\ldots,{v_k}') & = \konst{stepFn} (\mathit{inportVals},\mathit{stateVars}) ; \\
+ \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots,{v_k}');} \\
+ \multicolumn{2}{l}{\konst{writeOutputs}({v_1}',\ldots,{v_k}');} \\
+\end{array}
+\]
+It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+\end{comment}

### 2021-05-04 paper/example.tex
Add example description.

@@
-Show a filter, monitor, and gate form Phase II and indicate each of their intended roles.
+\begin{figure*}
+\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
+\label{fig:example}
+\end{figure*}
+
+\figref{fig:example} is an AADL model of an implementation of a software system for route planning and automated control for a UAV.
+It is loosely based on the system in the case study presented later.
+The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
+The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight control as the UAV location changes.
+It is a legacy component that is not subject to change.
+The alert manager never raises the alert as there is no capacity in the initial implementation to know when an alert is needed.
+
+The goal is to cyber-harden the system so that it is resilient to cyber-attack.
+For this simple example, two attacks are considered, both originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight control; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
+
+\begin{figure}
+  \caption{The SW component contract.}
+  \label{fig:sw}
+\end{figure}
+
+This work uses an assume guarantee modeling paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
+The contracts constrain the inputs and state properties of the output for component models in the system.
+In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component.
+The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See (cite repository) for the full model with contracts.}
+
+The contract to model the SW component with its primary inputs and outputs is in \figref{fig:sw}.
+The contract language is a first order predicate calculus.
+The contract semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams.
+The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic.
+Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream.
+
+The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract.
+For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression.
+An \texttt{event} expression is true if there is data on the named AADL event port.
+The system contract assumes well-formed input and guarantees a set of properties about the output.
+
+There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
+This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.
+
+The guarantee is an invariant on the expression \texttt{not alrt => policy} meaning that if there is not an alert output then the policy must hold (and if there is an alert, then the policy does not matter).
+The \texttt{policy} is satisfied if there is no pending request (\texttt{nothing}), a request with a response right now (\texttt{today}), or a request yesterday with its response today.
+
+The expression for \texttt{yesterday} is helpful to understanding data streams and the synchronous data-flow semantics.
+It uses the \texttt{->} operator for initialization.
+The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.
+The right operand is the value of \texttt{yesterday} after the initial step.
+Here the stream nature of expressions is apparent with the \texttt{pre} operator referring to the value of \texttt{(req and not rsp)} in the prior time step.
+Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response.
+The \emph{Alert if start is not bounded relative to a request} defines the behavior of the correctly implemented system which is that it alerts if the bounded response policy does not hold.
+In this way, the contracts model the expected input and output of the system as a whole and between components that implement the system.
+
+Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}.
+The results of the model checker are in \figref{fig:example}(b).
+No unexpectedly, since the model of the route planner leaves the output behavior unconstrained, the component can do anything since it is not trusted, the model checker proves that it does not implement the system requirements.
+
+\begin{figure*}
+  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
+  \label{fig:hardened}
+\end{figure*}
+
+\begin{figure}
+  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
+  \label{fig:assurance}
+\end{figure}
+
+The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
+The contract model for the filter only passes well-formed output from the route planner (see \figref{fig:assurance}(a)).
+The monitor only forwards responses that meet the bounded response property (see \figref{fig:assurance}(b)).
+The contract models for both components are based on the contract in \figref{fig:sw}.
+The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b).
+The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.
+
+The high-assurance components are automatically synthesized from the contract models for the final deployed system.
+Notice that the contracts in \figref{fig:assurance} completely specify the components' behavior stating both the conditions for output events and the values of outputs at those events.
+Synthesizing a component first turns it into its equivalent CakeML code.
+Most expressions have a direct CakeML counterpart. The \texttt{pre} expressions are handled by creating state variables to capture the values in the current step to use in the next step.

### 2021-05-04 paper/example.tex
Add figures

@@
 \begin{figure*}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.4]{example.png} \\
+      (a) \\
+      \includegraphics[scale=0.4]{example-certificate.png} \\
+      (b)
+    \end{tabular}
+  \end{center}
 \caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
@@
 
-\figref{fig:example} is an AADL model of an implementation of a software system for route planning and automated control for a UAV.
+\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV.
 It is loosely based on the system in the case study presented later.
@@
 It is a legacy component that is not subject to change.
-The alert manager never raises the alert as there is no capacity in the initial implementation to know when an alert is needed.
+The alert manager never raises the alert as there is no capacity in the initial design to know when an alert is needed.
 
@@
 \begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.4]{sw.png}
+  \end{center}
   \caption{The SW component contract.}
@@
 
-This work uses an assume guarantee modeling paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
-The contracts constrain the inputs and state properties of the output for component models in the system.
+This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
+The contracts constrain the inputs and state properties output for component models.
 In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component.
@@
 The \emph{Alert if start is not bounded relative to a request} defines the behavior of the correctly implemented system which is that it alerts if the bounded response policy does not hold.
-In this way, the contracts model the expected input and output of the system as a whole and between components that implement the system.
+In this way, the contracts model the expected input and output of the system as a whole and also between components that implement the system.
 
@@
 The results of the model checker are in \figref{fig:example}(b).
-No unexpectedly, since the model of the route planner leaves the output behavior unconstrained, the component can do anything since it is not trusted, the model checker proves that it does not implement the system requirements.
+Not unexpectedly, since the model of the route planner leaves the output behavior unconstrained (i.e., the component can do anything since it is not trusted) the model checker proves that the component composition does not implement the system requirements.
 
 \begin{figure*}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.3]{hardened.png} \\
+      (a) \\
+      \includegraphics[scale=0.4]{hardened-certificate.png} \\
+      (b)
+    \end{tabular}
+  \end{center}
   \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
@@
 \begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.45]{filter.png} \\
+      (a) \\
+      \includegraphics[scale=0.4]{monitor.png} \\
+      (b)
+    \end{tabular}
+  \end{center}
   \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
@@
 The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
-The contract model for the filter only passes well-formed output from the route planner (see \figref{fig:assurance}(a)).
-The monitor only forwards responses that meet the bounded response property (see \figref{fig:assurance}(b)).
-The contract models for both components are based on the contract in \figref{fig:sw}.
+A filter takes a stream of data and delivers a new stream of data where some property is enforced.
+Its stylized form generated by the CASE tools is shown in \figref{fig:assurance}(a).
+The user must provide the meaning of the property, a predicate, while the tool generates everything else.
+In this example, the property is the assumption made by the waypoint manager about the automation response being well-formed.
+The guarantee is such that it has a trivial direct mapping to synthesized code by completely specifying the behavior of the output.
+
+A monitor captures a relation over time on data and is able to reason about temporal properties.
+In this example, the monitor only forwards responses that meet the bounded response property as shown in \figref{fig:assurance}(b).
+Like the filter, the generated monitor, with its contract, is stylized for synthesis only requiring the user to indicate the policy and some extra information for the output.
+The policy is taken from the contract in \figref{fig:sw} since it is the monitor that needs to alert violations of the bounded response.
+
 The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b).
@@
 The high-assurance components are automatically synthesized from the contract models for the final deployed system.
-Notice that the contracts in \figref{fig:assurance} completely specify the components' behavior stating both the conditions for output events and the values of outputs at those events.
-Synthesizing a component first turns it into its equivalent CakeML code.
-Most expressions have a direct CakeML counterpart. The \texttt{pre} expressions are handled by creating state variables to capture the values in the current step to use in the next step.
+Synthesis assumes the existence of the communication fabric and a scheduler that implements the synchronous data-flow semantics by dispatching components in dependency order.
+For this example, and the later case study, the HAMR framework provides both requirements.
+
+\newsavebox{\boxa}
+\begin{lrbox}{\boxa}
+\begin{lstlisting}
+fun filter_step () =
+let val () = Utils.clear_buf filterBuffer
+    val () = API.callFFI "get_input" "" filterBuffer
+    val string = Utils.buf2string filterBuffer
+in
+    if WELL_FORMED_AUTOMATION_RESPONSE string
+    then
+        let val string = Word8Array.substring filterBuffer 1
+                         (Word8Array.length filterBuffer - 1)
+        in API.callFFI "put_output" string Utils.emptybuf
+        end
+    else print"Filter rejects message.\n"
+end
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\boxa}}
+    \end{tabular}
+  \end{center}
+  \caption{Synthesized CakeML for the filter.}
+  \label{fig:cakeml}
+\end{figure}
+
+Synthesis targets CakeML, and CakeML provides verified compilation to binaries for several different platforms.
+The generated code for the filter is shown in \figref{fig:cakeml}.
+The code is called at dispatch by the scheduler.
+The \texttt{API.callFFI} is the link to the communication fabric to capture input.
+The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate.
+The synthesis is proven to exactly implement the meaning of the contract in the synchronous data-flow semantics in the CakeML assuming perfect communication and scheduling.

### 2021-05-04 paper/paper.tex
Add example description.
Add figures

@@
 \usepackage{comment}
+\usepackage{paralist}
+\usepackage{graphicx}
+\graphicspath{{./figs/}} 
+\usepackage{listings}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+\definecolor{light-gray}{gray}{0.88}
+
+\lstset{frame=none,
+  basicstyle=\ttfamily,
+  language=ML,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=false,
+  breakatwhitespace=true,
+  tabsize=2,
+  morekeywords={function, method, then, reads, modifies, requires, ensures, var, assume, assert}
+}
+
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
   \thanks{
-    Funded by DARPA Cyber Assured Systems Engineering (CASE)
+    Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001.
   }
@@
 
-% \section{Simple Example}
-% \input{example}
+\section{Simple Example}
+\input{example}
 

### 2021-05-05 paper/example.tex
Revise example and use text for the contracts

@@
 The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
-The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight control as the UAV location changes.
-It is a legacy component that is not subject to change.
-The alert manager never raises the alert as there is no capacity in the initial design to know when an alert is needed.
+The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
+It is an \emph{as is} legacy component.
+The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.
 
 The goal is to cyber-harden the system so that it is resilient to cyber-attack.
-For this simple example, two attacks are considered, both originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight control; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
+For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight control; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
+
+\newsavebox{\sw}
+\begin{lrbox}{\sw}
+\begin{lstlisting}[style=agree]
+eq req : bool = event(AutomationRequest);
+eq avl : bool = event(AirVehicleLocation);
+eq wp : bool = event(Waypoint);
+eq rsp: bool = event(Start);
+eq alrt : bool = event(Alert);
+
+assume "Automation request is well-formed" :
+    req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
+assume "Air vehicle location is well-formed" :
+    avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);
+
+eq nothing : bool = (not req) -> (not pre (req)) and (not req);
+eq today : bool = req and rsp;    
+eq yesterday : bool = false -> pre (req and not rsp);
+eq policy : bool = nothing
+                or (today and not yesterday)
+                or (not today and rsp and yesterday);
+eq since : bool = (alrt and (false -> pre(since))) 
+               or (not policy and alrt);
+
+guarantee "Start includes a waypoint" :
+    rsp => wp;
+guarantee "Locations required after the start waypoint" :
+    (wp and not rsp) => avl;
+guarantee "Waypoint is well-formed" :
+    wp => WELL_FORMED_WAYPOINT(Waypoint);
+guarantee "Alert if start is not bounded relative to a request" :
+    (policy and not alrt) or since;
+\end{lstlisting}
+\end{lrbox}
 
@@
   \begin{center}
-    \includegraphics[scale=0.4]{sw.png}
+    \scalebox{0.60}{\usebox{\sw}}
   \end{center}
@@
 This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
-The contracts constrain the inputs and state properties output for component models.
+The contracts constrain the inputs and state properties of output for component models.
 In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component.
-The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See (cite repository) for the full model with contracts.}
+The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See \cite{repo} for the full model.}
 
@@
 Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream.
+The semantics do not model computation or communication delay.
+The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
 
@@
 
-There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
+There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
 This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.
 
-The guarantee is an invariant on the expression \texttt{not alrt => policy} meaning that if there is not an alert output then the policy must hold (and if there is an alert, then the policy does not matter).
+The guarantee is an invariant on the expression \texttt{(policy and not alrt) or since} meaning that either the policy holds and there is no alert or the alert has been sounding since the policy was first violated.
 The \texttt{policy} is satisfied if there is no pending request (\texttt{nothing}), a request with a response right now (\texttt{today}), or a request yesterday with its response today.
-
-The expression for \texttt{yesterday} is helpful to understanding data streams and the synchronous data-flow semantics.
-It uses the \texttt{->} operator for initialization.
+The value of \texttt{nothing} and \texttt{yesterday} in the current time step rely on values from the previous time step.
+The \texttt{yesterday} uses the \texttt{->} operator for initialization.
 The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.
 The right operand is the value of \texttt{yesterday} after the initial step.
-Here the stream nature of expressions is apparent with the \texttt{pre} operator referring to the value of \texttt{(req and not rsp)} in the prior time step.
+Here the \texttt{pre} operator refers to the value of \texttt{(req and not rsp)} in the prior time step.
 Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response.
-The \emph{Alert if start is not bounded relative to a request} defines the behavior of the correctly implemented system which is that it alerts if the bounded response policy does not hold.
-In this way, the contracts model the expected input and output of the system as a whole and also between components that implement the system.
+
+The value of \texttt{since} relies on its own value in the previous time step.
+The intuitive reading of the expression is that the alert has been true since the time when the policy first failed.
+The first failure sets \texttt{since} to true (i.e., \texttt{(not policy and alrt)}), and once \texttt{since} true, it persists as long as \texttt{alrt} holds.
+As such, the \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system.
+In this way, the contracts model the expected input and output of the system as a whole, and they also model the expected behavior of each component used to implement the system.
 
@@
 
+\newsavebox{\flt}
+\begin{lrbox}{\flt}
+\begin{lstlisting}[style=agree]
+property CASE_Filter_policy = 
+  WELL_FORMED_AUTOMATION_RESPONSE(Input);
+guarantee CASE_Filter_Output "Filter output is well-formed" :
+  if event(Input) and CASE_Filter_policy then
+    event(Output) and Output = Input
+  else not event(Output);
+\end{lstlisting}
+\end{lrbox}
+
+\newsavebox{\mntr}
+\begin{lrbox}{\mntr}
+\begin{lstlisting}[style=agree]
+const is_latched : bool = 
+  Get_Property(this, CASE_Properties::Monitor_Latched);
+eq rsp : bool = event(Response);
+eq req : bool = event(Request);
+eq nothing : bool = (not req) -> (not pre (req)) and (not req);
+eq today : bool = req and rsp;    
+eq yesterday : bool = false -> pre (req and not rsp);
+property policy = nothing
+                or (today and not yesterday)
+                or (not today and rsp and yesterday);
+eq alert : bool = (not policy) 
+               -> ((is_latched and pre(alert)) or not policy);
+
+guarantee "Alert port tracks alert variable" :
+    event(Alert) <=> alert;
+guarantee "Output if not alerted" :
+    if event(Alert) then (not event(Output)) else
+    if event(Response) then (event(Output) and (Output = Response))
+    else (not event(Output));
+\end{lstlisting}
+\end{lrbox}
+
 \begin{figure}
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.45]{filter.png} \\
+      \scalebox{0.60}{\usebox{\flt}} \\
       (a) \\
-      \includegraphics[scale=0.4]{monitor.png} \\
+      \scalebox{0.60}{\usebox{\mntr}} \\
       (b)
@@
 The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
-A filter takes a stream of data and delivers a new stream of data where some property is enforced.
-Its stylized form generated by the CASE tools is shown in \figref{fig:assurance}(a).
-The user must provide the meaning of the property, a predicate, while the tool generates everything else.
+A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
+Its stylized form generated the CASE tools is shown in \figref{fig:assurance}(a).
+The user must provide the meaning of the property, which is just a predicate, while the tool generates everything else.
 In this example, the property is the assumption made by the waypoint manager about the automation response being well-formed.
@@
 In this example, the monitor only forwards responses that meet the bounded response property as shown in \figref{fig:assurance}(b).
-Like the filter, the generated monitor, with its contract, is stylized for synthesis only requiring the user to indicate the policy and some extra information for the output.
-The policy is taken from the contract in \figref{fig:sw} since it is the monitor that needs to alert violations of the bounded response.
+Like the filter, the generated monitor, with its contract, is stylized for synthesis only requiring the user to indicate the policy and some extra information for the outputs.
+The policy is taken from the contract in \figref{fig:sw} since it is the monitor that needs to alert violations of the bounded response policy.
 
@@
 
-The high-assurance components are automatically synthesized from the contract models for the final deployed system.
-Synthesis assumes the existence of the communication fabric and a scheduler that implements the synchronous data-flow semantics by dispatching components in dependency order.
-For this example, and the later case study, the HAMR framework provides both requirements.
-
-\newsavebox{\boxa}
-\begin{lrbox}{\boxa}
-\begin{lstlisting}
+\newsavebox{\cml}
+\begin{lrbox}{\cml}
+\begin{lstlisting}[style=myML]
 fun filter_step () =
-let val () = Utils.clear_buf filterBuffer
-    val () = API.callFFI "get_input" "" filterBuffer
-    val string = Utils.buf2string filterBuffer
+let val () = Utils.clear_buf buffer
+    val () = API.callFFI "get_input" "" buffer
+    val string = Utils.buf2string buffer
 in
@@
     then
-        let val string = Word8Array.substring filterBuffer 1
-                         (Word8Array.length filterBuffer - 1)
+        let val string = Word8Array.substring buffer 1
+                         (Word8Array.length buffer - 1)
         in API.callFFI "put_output" string Utils.emptybuf
@@
     \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\boxa}}
+      \scalebox{0.60}{\usebox{\cml}}
     \end{tabular}
@@
 
-Synthesis targets CakeML, and CakeML provides verified compilation to binaries for several different platforms.
-The generated code for the filter is shown in \figref{fig:cakeml}.
+The high-assurance components are automatically synthesized from the contract models to equivalent models in the CakeML language.
+The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}.
 The code is called at dispatch by the scheduler.
-The \texttt{API.callFFI} is the link to the communication fabric to capture input.
-The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate.
-The synthesis is proven to exactly implement the meaning of the contract in the synchronous data-flow semantics in the CakeML assuming perfect communication and scheduling.
+The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
+The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
+
+CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.
+A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML.
+A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML.
+In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.
+
+Preserving the input/output relationship of streams between the two models lifts the contract verification results to the deployed system.
+If the contract model verifies, then the meaning of those results hold for the deployed system if the other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.

### 2021-05-05 paper/paper.tex
Revise example and use text for the contracts

@@
 
-\lstset{frame=none,
+\lstdefinestyle{myML}
+{frame=none,
   basicstyle=\ttfamily,
@@
   tabsize=2,
-  morekeywords={function, method, then, reads, modifies, requires, ensures, var, assume, assert}
+  linewidth=2\linewidth
+}
+
+\lstdefinestyle{agree}
+{frame=none,
+  basicstyle=\ttfamily,
+  language=ML,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=false,
+  breakatwhitespace=true,
+  tabsize=2,
+  linewidth=2\linewidth,
+  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const}
 }
@@
 
-% \section{Related Work}
-% \input{related-work}
+\section{Related Work}
+\input{related-work}
 
-% \section{Conclusion}
-% \input{conclusion}
+\section{Conclusion}
+\input{conclusion}
 
@@
 % references section
-% \bibliographystyle{IEEEtran}
-% \bibliography{paper}
+\bibliographystyle{IEEEtran}
+\bibliography{paper}
 \end{document}

### 2021-05-06 paper/example.tex
Update system model

@@
 \begin{lstlisting}[style=agree]
-property CASE_Filter_policy = 
-  WELL_FORMED_AUTOMATION_RESPONSE(Input);
-guarantee CASE_Filter_Output "Filter output is well-formed" :
-  if event(Input) and CASE_Filter_policy then
+eq policy : bool = WELL_FORMED_AUTOMATION_RESPONSE(Input);
+property Policy_Output = 
+      if event(Input) and policy then 
         event(Output) and Output = Input
       else not event(Output);
+guarantee "Filter output is well-formed" :
+    Policy_Output;
 \end{lstlisting}
@@
 eq yesterday : bool = false -> pre (req and not rsp);
-property policy = nothing
+eq policy: bool = nothing
                 or (today and not yesterday)
@@
                 -> ((is_latched and pre(alert)) or not policy);
-
-guarantee "Alert port tracks alert variable" :
-    event(Alert) <=> alert;
-guarantee "Output if not alerted" :
+property Policy_Output = 
   if event(Alert) then (not event(Output)) else
@@
   else (not event(Output));
+property Policy_Alert =  event(Alert) = alert;
+
+guarantee "Alert port tracks alert variable" :
+    Policy_Alert;
+guarantee "Output if not alerted" :
+    Policy_Output;
 \end{lstlisting}
@@
 
-The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
+The system implementation is cyber-hardened by using the CASE tools to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
 A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
-Its stylized form generated the CASE tools is shown in \figref{fig:assurance}(a).
-The user must provide the meaning of the property, which is just a predicate, while the tool generates everything else.
-In this example, the property is the assumption made by the waypoint manager about the automation response being well-formed.
-The guarantee is such that it has a trivial direct mapping to synthesized code by completely specifying the behavior of the output.
+Its stylized form as generated by the tool is shown in \figref{fig:assurance}(a).
+The user must declare the meaning of the \texttt{policy} value as part of the transform on the model.
+In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed.
+The output policies and guarantees are automatically generated by the tool.
 
-A monitor captures a relation over time on data and is able to reason about temporal properties.
-In this example, the monitor only forwards responses that meet the bounded response property as shown in \figref{fig:assurance}(b).
-Like the filter, the generated monitor, with its contract, is stylized for synthesis only requiring the user to indicate the policy and some extra information for the outputs.
-The policy is taken from the contract in \figref{fig:sw} since it is the monitor that needs to alert violations of the bounded response policy.
+A monitor captures a relation over time on input data and is able to reason about temporal properties of that input.
+By definition, it raises an alert output when the relation is violated.
+The generated monitor contract is shown in \figref{fig:assurance}(b).
+The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised.
+The definition for \texttt{policy} is taken from the contract in \figref{fig:sw} since this high-assurance component is the monitor that needs to alert violations of the bounded response policy.
+As before the policies for the outputs and the guarantees are autogenerated by the tool.
 

### 2021-05-06 paper/system.tex
Update system model

@@
                     &   & \konst{output}\ [(f : \tau)\ldots] \\
-                    &   & [\mathit{stmts}\ldots] \\
-                    &   & \konst{assume}\ [\mathit{bexp}\ldots] \\
-                    &   & \konst{guarantee}\ [\mathit{bexp}\ldots] \\ \\
-
-      \mathit{stmts}& = & \konst{subcomp}\ \lval\dots\ c \\
-                    & | & \konst{eq}\; f : \tau := \mathit{exp} \\ \\
+                    &   & \konst{eq}\; f : \tau := \mathit{exp} \\
+                    &   & \konst{policies}\ [\mathit{bexp}\ldots]\\
 
@@
 \]
-\caption{Syntax for component specifications.}
+\caption{Syntax for high-assurance component specifications.}
 \label{fig:syntax}
@@
 
-A \emph{system} in this model of computation is a \emph{component} that is defined by a specification (see \figref{fig:syntax}).
-A specification defines inputs, outputs, statements, assumptions for pre-conditions, and guarantees for post-conditions.
-A statement can be a subcomponent definition with its input binding or a local variable definition.
-
+The specification language for high-assurance components is in (see \figref{fig:syntax}).
+It is stylized contracts that defines the inputs, outputs, local values, and policies for each output.
+Each output policy must be invariant over the input stream and completely define the meaning of its corresponding output.
 A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation).
@@
 
-The initial environment stream only contains mappings for primary inputs along the entire stream.
-Stepping the component updates the current environment.
-In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of each component evaluation.
-A component step traverses the component sequentially to update the environment recursing into subcomponents after binding inputs.
-Assume the component is written in dependency order.
+The initial environment stream only contains mappings for the inputs along the entire stream.
+Stepping the component updates the current environment and checks the policies.
+In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification, which are assumed to be in dependency order, and then the policies are checked for invariance.
 

### 2021-05-06 paper/example.tex
Update with transform patterns

@@
 \begin{lstlisting}[style=agree]
-eq policy : bool = WELL_FORMED_AUTOMATION_RESPONSE(Input);
-property Policy_Output = 
+eq policy : bool = 
+  WELL_FORMED_AUTOMATION_RESPONSE(Input);       
+guarantee Filter_Output "Filter output is well-formed" :
   if event(Input) and policy then 
@@
   else not event(Output);
-guarantee "Filter output is well-formed" :
-    Policy_Output;
 \end{lstlisting}
@@
                -> ((is_latched and pre(alert)) or not policy);
-property Policy_Output = 
+guarantee Monitor_Alert 
+  "Alert port tracks alert variable" :
+  event(Alert) = alert;
+guarantee Monitor_Output
+  "Output if not alerted" :
   if event(Alert) then (not event(Output)) else
@@
   else (not event(Output));
-property Policy_Alert =  event(Alert) = alert;
-
-guarantee "Alert port tracks alert variable" :
-    Policy_Alert;
-guarantee "Output if not alerted" :
-    Policy_Output;
 \end{lstlisting}
@@
 In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed.
-The output policies and guarantees are automatically generated by the tool.
+The output guarantees are automatically generated by the tool.
 
@@
 The definition for \texttt{policy} is taken from the contract in \figref{fig:sw} since this high-assurance component is the monitor that needs to alert violations of the bounded response policy.
-As before the policies for the outputs and the guarantees are autogenerated by the tool.
+As before the guarantees for the outputs are autogenerated by the tool.
 

### 2021-05-06 paper/system.tex
Update with transform patterns

@@
                     &   & \konst{eq}\; f : \tau := \mathit{exp} \\
-                    &   & \konst{policies}\ [\mathit{bexp}\ldots]\\
+                    &   & \konst{guarantees}\ [\mathit{bexp}\ldots]\\
 
@@
 The specification language for high-assurance components is in (see \figref{fig:syntax}).
-It is stylized contracts that defines the inputs, outputs, local values, and policies for each output.
-Each output policy must be invariant over the input stream and completely define the meaning of its corresponding output.
+It is stylized contracts that defines the inputs, outputs, local values, and guarantees for each output.
+Each guarantee must be invariant over the input stream and define the meaning of its corresponding output under all conditions.
 A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation).
@@
 Stepping the component updates the current environment and checks the policies.
-In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification, which are assumed to be in dependency order, and then the policies are checked for invariance.
+In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification, which are assumed to be in dependency order, and then the guarantees are checked for invariance.
+
+The meaning of the specification is a set of finite environment streams where the guarantees hold at each step of the streams and the updated mappings or the \konst{eq}-statements are consistent with the statements.
 

### 2021-05-06 paper/example.tex
Fix typo and update ML

@@
     val string = Utils.buf2string buffer
+    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
 in
-    if WELL_FORMED_AUTOMATION_RESPONSE string
+    if policy
     then

### 2021-05-06 paper/system.tex
Fix typo and update ML

@@
                     &   & \konst{output}\ [(f : \tau)\ldots] \\
-                    &   & \konst{eq}\; f : \tau := \mathit{exp} \\
+                    &   & \konst{eq}\; [(f : \tau := \mathit{exp})\ldots] \\
                     &   & \konst{guarantees}\ [\mathit{bexp}\ldots]\\

### 2021-05-06 paper/example.tex
Simplify example (fix bug)

@@
 The goal is to cyber-harden the system so that it is resilient to cyber-attack.
-For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight control; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
+For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
 
@@
 
-eq nothing : bool = (not req) -> (not pre (req)) and (not req);
-eq today : bool = req and rsp;    
-eq yesterday : bool = false -> pre (req and not rsp);
-eq policy : bool = nothing
-                or (today and not yesterday)
-                or (not today and rsp and yesterday);
-eq since : bool = (alrt and (false -> pre(since))) 
-               or (not policy and alrt);
+eq today : bool = (req = rsp);
+eq yesterday : bool = (req and not rsp) ->
+                      pre(req and not rsp) and (not req and rsp);
+eq policy : bool = today or yesterday;
+eq since : bool =  alrt or (alrt and (false -> pre(since)));
 
@@
 guarantee "Alert if start is not bounded relative to a request" :
-    (policy and not alrt) or since;
+    policy or since;
 \end{lstlisting}
@@
 For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression.
-An \texttt{event} expression is true if there is data on the named AADL event port.
+There is an implicit \emph{event} input (or output) associated with every named event port in a component.
+The semantics used here do not not buffer events so the implicit input (or output) is a boolean value.
+An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port.
 The system contract assumes well-formed input and guarantees a set of properties about the output.
@@
 
-The guarantee is an invariant on the expression \texttt{(policy and not alrt) or since} meaning that either the policy holds and there is no alert or the alert has been sounding since the policy was first violated.
-The \texttt{policy} is satisfied if there is no pending request (\texttt{nothing}), a request with a response right now (\texttt{today}), or a request yesterday with its response today.
-The value of \texttt{nothing} and \texttt{yesterday} in the current time step rely on values from the previous time step.
-The \texttt{yesterday} uses the \texttt{->} operator for initialization.
-The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.
+The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
+The \texttt{policy} is satisfied if there is a request with a response or nothing in the current time step (\texttt{today}) or a request yesterday with its response today.
+The value of \texttt{yesterday} in the current time step relies on values from the previous time step.
+The \texttt{yesterday} value uses the \texttt{->} operator for initialization.
+The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{(req and not rsp)} because seeing a request with no response is inconclusive at the start.
 The right operand is the value of \texttt{yesterday} after the initial step.
-Here the \texttt{pre} operator refers to the value of \texttt{(req and not rsp)} in the prior time step.
-Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response.
+Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step.
+Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
 
 The value of \texttt{since} relies on its own value in the previous time step.
-The intuitive reading of the expression is that the alert has been true since the time when the policy first failed.
-The first failure sets \texttt{since} to true (i.e., \texttt{(not policy and alrt)}), and once \texttt{since} true, it persists as long as \texttt{alrt} holds.
-As such, the \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system.
-In this way, the contracts model the expected input and output of the system as a whole, and they also model the expected behavior of each component used to implement the system.
+The intuitive reading of the expression is that the alert has been true since the time when it first sounded.
+The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds.
+The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system.
+Together with the other guarantees, the contract models the expected input and output of the system as a whole.
+Contracts also model the expected behavior of each component used to implement the system.
 
@@
 The results of the model checker are in \figref{fig:example}(b).
-Not unexpectedly, since the model of the route planner leaves the output behavior unconstrained (i.e., the component can do anything since it is not trusted) the model checker proves that the component composition does not implement the system requirements.
+The red exclamation points are properties that do not hold.
+Each comes with a counter-example.
+The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything).
+The counter-example showing each vio with a counterexample that violates the policy and does not raise the alert.
 
@@
 eq req : bool = event(Request);
-eq nothing : bool = (not req) -> (not pre (req)) and (not req);
-eq today : bool = req and rsp;    
-eq yesterday : bool = false -> pre (req and not rsp);
-eq policy: bool = nothing
-               or (today and not yesterday)
-               or (not today and rsp and yesterday);
+eq today : bool = (req = rsp);
+eq yesterday : bool = (req and not rsp) ->
+                      pre(req and not rsp) and (not req and rsp);
+eq policy : bool = today or yesterday;
 eq alert : bool = (not policy) 

### 2021-05-06 paper/contig.tex
Add contig types

@@
-A \emph{string} is an array of bytes.
-A \emph{language} is a set of strings.
-A \emph{recognizer} for a language determines if a string belongs to the language.
-A language \emph{filter} uses a recognizer to pass an input string unchanged to the output if the string belongs to the given language; otherwise, it outputs an empty string.
+The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types} (cite contiguity).
+A contiguity type is a self-describing specification for messages.
+Its formalism has basis in formal languages.
+Similar to how a regular expression implies a set of words that form its language, so does a contiguity type specification imply a set of messages for its language where a message is a finite sequence of contiguous bytes (e.g., a string).
 
-Languages in this work are formally defined by \emph{contiguity types}.
-A contiguity type specifies data as a laid out side-by-side array of elements in a string.
-Every contiguity type has an associated recognizer that precisely recognizes strings in the language.
+What makes contiguity type specification more expressive than regular expressions is that it is self-describing meaning that the contents of the message itself may determine the rest of the message.
+An example is the \texttt{AutomationResponse} from the system in the previous section with its contiguity type specification.
+{\small
+\begin{verbatim}
+  {TaskID : i64
+   Length : u8
+   Waypoints : Waypoint[Length]
+  }
+\end{verbatim}
+}
+\noindent The \texttt{Waypoints} array size depends on the value of \texttt{Length} so the actual number of bytes in the message depends on the contents of the message itself.
 
-\subsection{Contiguity Types}
+The type specifications may also carry meta-information about the contents of the message.
+{\small
+\begin{verbatim}
+  {Latitude : float
+   lt-rng : Assert (-90 <= Latitude <= 90) 
+   Longitude : float
+   lng-rng : Assert (-180 <= Longitute <= 180)
+   Altitude : float
+   a-rng : Assert (10000 <= Altitude <= 15000)
+  }
+\end{verbatim}
+}
+\noindent Here the specification encodes the allowed ranges for each field of the waypoint. These can be checked while constructing a message from a sequence of bytes.
 
-See \cite{contiguity} for a complete discussion. Contiguity types assume that a string is the \emph{result} of a
-map from structured data. Figure \ref{contig-types} is the syntax. Contiguity types start with common base types (booleans,
-characters, signed and unsigned integers, \etc) and are closed under
-the construction of records, arrays, and unions. \footnote{We use
-the terms ``\emph{contiguity type}, \konst{contig}, and $\tau$
-interchangeably.}
+Every contiguity type specification has a corresponding \emph{matcher} that when given a message string returns true or false if that message belongs to the language of the specification.
+If the message does belong to the language, an \emph{environment} is provided to access each part of the message.
+An environment, $\theta: \lval \mapsto \konst{string}$ binds \emph{L-values} to strings.
+An L-value is an expression that can appear on the left hand side of an assignment.
+Alpha renaming is assumed so that every $\lval$ is unique.
 
-\begin{figure}
-\[
-\begin{array}{rcl}
- \mathit{base} & = & \konst{bool} \mid \konst{char} \mid \konst{u8} \mid \konst{u16} \mid \konst{u32} \\
-               & \mid & \konst{u64}  \mid \konst{i16} \mid \konst{i32} \mid \konst{i64} \mid \konst{float} \mid \konst{double} \\ \\
-
- \tau & = & \mathit{base} \\
-      & \mid & \konst{Recd}\; [ (f_1 : \tau_1) \ldots (f_n : \tau_n) ] \\
-      & \mid & \konst{Array}\; \tau \; \mathit{exp} \\
-      & \mid & \konst{Alt}\; \mathit{bexp}\; \tau_1 \; \tau_2 \\ \\
-
-\mathit{lval} & = & f \mid
-                    \mathit{lval} \, [ \mathit{exp} ] \mid
-                    \mathit{lval} . f \\ \\
-f            & = & \mathit{varName} \\ \\
-\mathit{exp} & = & \konst{Loc}\; \mathit{lval}
-              \mid \konst{nLit}\; \konst{nat}
-              \mid \mathit{constname} \\
-             & \mid & \mathit{exp} + \mathit{exp}
-                   \mid \mathit{exp} * \mathit{exp} \\ \\
-
-\mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
-              \mid  \konst{bLit}\; \konst{bool}
-              \mid  \neg \mathit{bexp}
-              \mid  \mathit{bexp} \land \mathit{bexp} \\
-              & \mid & \mathit{exp} = \mathit{exp} 
-              \mid  \mathit{exp} < \mathit{exp}
-\end{array}
-\]
-\caption{Syntax for contiguity types.}
-\label{contig-types}
-\end{figure}
-
-Arithmetic and Boolean expressions define dependent types. The expressions rely on the notion of an \emph{L-value} to reference values in previous fields.
-
-Theorem statement: the traces accepted by the specification match the traces accepted by the implementation.
-If we have have uninterpreted functions, then it may not be precise as the agree analysis as something may be rejected by agree but accepted by the implementation.
-Need to think about that carefully.
-
-An L-value denotes an \emph{offset} from the beginning of a
-data structure, plus a \emph{width}. In an R-value, an occurrence of an
-L-value is mapped to the value of the patch of memory between
-$\mathit{offset}$ and $\mathit{offset} + \mathit{width}$. For the
-purpose of specifying message formats, it may not be immediately
-obvious that a notation supporting assignment in imperative languages
-can help, but there is indeed a form of assignment lurking.
-
-The above explanation of L-values centers on indices into a byte
-buffer; in the following we will give a mild variant of this: instead
-of indices into the buffer, we lift out the designated slices. Thus,
-given environments $\theta: \mathit{lval} \mapsto \konst{string}$ (binding
-L-values to strings), $\Delta : \konst{string} \to \mathbb{N}$
-(binding constant names to numbers) and functions
-$\konst{toN}:\konst{string}\to\mathbb{N}$ and
-$\konst{toB}:\konst{string}\to\konst{bool}$ (which interpret byte
-sequences to numbers and booleans, respectively), expression
-evaluation and boolean expression evaluation have conventional
-definitions:
-
-\[
-\begin{array}{l}
-\konst{eval} \; e =
-\mathtt{case}\; e\
- \left\{
- \begin{array}{lcl}
-    \konst{Loc}\; \lval & \Rightarrow & \konst{toN}(\theta(\lval)) \\
-    \konst{nLit}\; n & \Rightarrow & n  \\
-    \mathit{constname} & \Rightarrow & \Delta(\mathit{constname})  \\
-    e_1 + e_2 & \Rightarrow & \konst{eval}\; e_1 + \konst{eval}\; e_2  \\
-    e_1 * e_2 & \Rightarrow & \konst{eval}\; e_1 * \konst{eval}\; e_2  \\
-  \end{array}
- \right.
- \\ \\
-\konst{evalB} \; b =
-\mathtt{case}\; b\
- \left\{
- \begin{array}{lcl}
-    \konst{bLoc}\; \lval & \Rightarrow & \konst{toB}(\theta(\lval)) \\
-    \konst{bLit}\; b & \Rightarrow & b \\
-    \neg b & \Rightarrow & \neg(\konst{evalB} \; b)  \\
-    b_1 \lor b_2 & \Rightarrow & \konst{evalB} \;b_1 \lor \konst{evalB} \;b_2   \\
-    b_1 \land b_2 & \Rightarrow & \konst{evalB} \;b_1 \land \konst{evalB} \;b_2   \\
-    e_1 = e_2 & \Rightarrow & \konst{eval} \;e_1 = \konst{eval} \;e_2   \\
-    e_1 < e_2 & \Rightarrow & \konst{eval} \;e_1 < \konst{eval} \;e_2
-  \end{array}
- \right.
-
-\end{array}
-\]
-
-\begin{remark}[Partiality]
-Expression evaluation is partial because there is no guarantee that
-$\theta(\lval)$ is defined: an \lval{} being looked-up may not be in the
-map $\theta$. Failure in evaluation is modelled by the \konst{option}
-type, and must be handled in the semantics and the matching
-algorithm. However error handling is omitted in the presentation since
-it hampers readability. See the HOL4 formalization\footnote{
-\texttt{https://github.com/loonwerks/splat/blob/monitors/contig/abscontigScript.sml}}
-for full details.
-\end{remark}
-
-\subsection{Semantics}
-
- We now confess to misleading the reader: in spite of the notational
- similarity, a contiguity type is \emph{not} a type: it is a formal
- language. A type is usually understood to represent a set, or domain,
- of values, \eg, the type \konst{int32} represents a set of
- integers. In contrast, the contiguity type \konst{i32} represents the
- set of strings of width 32 bits. An element of a contiguity type can
- be turned into an element of a type by providing interpretations for
- all the strings at the leaves and interpreting the \konst{Recd} and
- \konst{Array} constructors into the corresponding type constructs. (A
- base contiguity type therefore serves mainly as a \emph{tag} to be
- interpreted as a width and also as an intended target type.) Thus,
- contiguity types sit---conveniently---between the types in a
- programming language and the strings used to make messages.
-
-The semantics definition depends on a few basic notions familiar from
-language theory: language concatenation, and iterated language
-concatenation.
-
-\begin{align*}
-L_1 \cdot L_2 &= \set{w_1 w_2 \mid w_1 \in L_1 \land w_2 \in  L_2} \\
-L^0 &= \varepsilon \\
-L^{n+1} &= L \cdot L^n
-\end{align*}
-
-\begin{definition}[Semantics of contiguity types]
-
- In the following, we assume given an assignment $\theta$ adequate for
- evaluating expressions. If an expression evaluation fails, the
- language being constructed will be empty (e.g., $\emptyset$). $\LangTheta{\tau}$ is defined recursively as
-
-\[
-% \begin{array}{l}
-%\LangTheta{\tau} =
-\mathtt{case}\; \tau\
-% \hspace*{3mm}
- \left\{
- \begin{array}{l}
- \mathit{base} \Rightarrow \set{s \mid \konst{len}(s) = \konst{width}(base)} \\
- \konst{Recd}\; [(f_1 : \tau_1) \ldots (f_n : \tau_n)]
-      \Rightarrow \\
-      \hspace*{5mm} \LangTheta{\tau_1} \cdot \ldots \cdot \LangTheta{\tau_n}
-\\
- \konst{Array}\; \tau \; \mathit{exp} \Rightarrow  \\
-  \hspace*{5mm}
- \left\{
- \begin{array}{ll}
-    \LangTheta{\tau}^{\konst{eval}\;\theta\;\mathit{exp}} &
-       \mathrm{if}\ \konst{eval}\;\theta\;\mathit{exp}\ \mathrm{succeeds} \\
-    \emptyset & \text{otherwise}
- \end{array}
- \right.
-\\
- \konst{Alt}\; \mathit{bexp}\;\tau_1\; \tau_2 \Rightarrow \\
-  \hspace*{5mm}
- \left\{
- \begin{array}{ll}
-    \LangTheta{\tau_1} & \mathrm{if}\ \konst{evalB}\;\theta\;\mathit{bexp} = \konst{true} \\
-    \LangTheta{\tau_2} & \mathrm{if}\ \konst{evalB}\;\theta\;\mathit{bexp} = \konst{false} \\
-    \emptyset          & \text{otherwise}
- \end{array}
- \right.
- \\
-\end{array}
- \right.
-%\end{array}
-\]
-\end{definition}
-
-\subsection{Recognizer Synthesis}
-
-A key result for contiguity types is that every type has a corresponding recognizer that recognizes strings belonging to the language defined by the type.
-In this work, the recognizer is a function \konst{match} that recursively traverses the contiguity type $\tau$ consuming the given input string $s$ as it constructs the environment $\theta$.
-The functions uses the environment $\theta$ to dynamically evaluate and resolve the expressions in $\tau$ during the traversal.
-The \konst{match} function returns a option type where $\konst{match}\; \tau\; s_1s_2\; \theta_o = \konst{SOME}(s_2, \theta)$ is a successful match having consumed $s_1$ leaving the postfix $s_2$ untouched and built the environment $\theta$ for $s_1$ according to $\tau$ from an empty initial environment $\theta_0$.
-
-\begin{theorem}[Matcher soundness] 
-$\konst{match}\; \tau\; s_1s_2\; \theta_0 = \konst{SOME}(s_2, \theta) \imp s_1 \in \LangTheta{\tau}$
-\end{theorem}
+The matcher itself is synthesized from the specification to CakeML.
+The synthesis includes a corresponding proof that the matcher recognizes the language of the specification, and the resulting environment, $\theta$, from a matched message produces the same message as the one matched when serialized.

### 2021-05-06 paper/contig.tex
Add details on semantics

@@
 An L-value is an expression that can appear on the left hand side of an assignment.
-Alpha renaming is assumed so that every $\lval$ is unique.
+The syntax for $\lval$ is given in the next section.
 
@@
 The synthesis includes a corresponding proof that the matcher recognizes the language of the specification, and the resulting environment, $\theta$, from a matched message produces the same message as the one matched when serialized.
+
+The inputs and outputs for a high-assurance component are defined by contiguity type specifications.
+The synthesis from the component specification relies on the corresponding matchers from the contiguity type specifications.
+Details of which are in the next section.

### 2021-05-06 paper/system.tex
Add details on semantics

--- 2021-05-06 paper/system.tex
@@
                     &   & \konst{output}\ [(f : \tau)\ldots] \\
-                    &   & \konst{eq}\; f : \tau := \mathit{exp} \\
-                    &   & \konst{guarantees}\ [\mathit{bexp}\ldots]\\
+                    &   & \konst{eq}\; [(f : \tau := \mathit{exp})\ldots] \\
+                    &   & \konst{guarantee}\ [\mathit{bexp}\ldots]\\\\
 
@@
 
-The specification language for high-assurance components is in (see \figref{fig:syntax}).
-It is stylized contracts that defines the inputs, outputs, local values, and guarantees for each output.
-Each guarantee must be invariant over the input stream and define the meaning of its corresponding output under all conditions.
-A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation).
-An $\lval$ is a reference to an \emph{L-value} from compiler concepts and is an expression that can appear on the left hand side of an assignment.
-Alpha renaming is assumed so that every $\lval$ is unique.
+The specification language for a high-assurance component is in \figref{fig:syntax}.
+A specification defines the inputs, outputs, local values, and guarantees for each output.
+A type $\tau$ is a contiguity type, and $(f : \tau)\ldots$ means zero or more repetition (e.g., Kleene star).
+An $\lval$ must eventually resolve to something that can be assigned.
+The expression language divides out Boolean expressions to simplify the semantics but is otherwise typical.
+The \konst{Loc} and \konst{bLoc} refer to the value of an $\lval$, while \konst{nLit} and \konst{bLit} indicate a literal.
+The language includes the initialization ($\rightarrow$), \konst{pre}, and if-then-else (\konst{ite}) operators.
+
+The semantics are only defined for \emph{well-formed} specifications. A specification is well-formed if and only if
+\begin{compactenum}
+\item Every $\lval$ is unique;
+\item the \konst{eq} list is in dependency order and the expressions are acyclic;
+\item the associated $\lval$ with each \konst{Loc} and \konst{bLoc} expression is a valid reference in the environment;
+\item the associated literal with each \konst{nLit} and \konst{bLit} has the correct type;
+\item \konst{pre} expressions do not refer past the beginning of the associated streams;
+\item the expression list from \konst{guarantee} exactly corresponds in size and order to the list from \konst{output}; and
+\item every expression in the list from \konst{guarantee} defines its corresponding output value under all input combinations.
+\end{compactenum}
+These checks are part of the synthesis but omitted to simplify the presentation.
 
@@
 
-The semantics are synchronous data-flow on a single clock defined over a sequence of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream.
-Expression evaluation is defined in the context of the environment stream in \figref{fig:eval}.
+The well-formed assumption enables the use of a single global environment for the semantics.
+The semantics are synchronous data-flow on a single clock defined over a sequences of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream.
+Expression evaluation is defined in the context of this environment stream as shown is \figref{fig:eval}.
+Here, $\konst{eval}\; i\; e$ carries with it the index of the environment to be used for the expression.
 
@@
 
-The initial environment stream only contains mappings for the inputs along the entire stream.
-Stepping the component updates the current environment and checks the policies.
-In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification, which are assumed to be in dependency order, and then the guarantees are checked for invariance.
+Each environment in the stream is initially partial meaning that it only contains mappings for the inputs.
+\emph{Stepping} the specification updates the current environment and checks the invariance of the guarantees.
+In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification and then the guarantees are checked for invariance.
 
 The meaning of the specification is a set of finite environment streams where the guarantees hold at each step of the streams and the updated mappings or the \konst{eq}-statements are consistent with the statements.
+
+At the $i^\mathrm{th}$ step of the stream, $\theta^i$
+
+Each \emph{step} of the specification updates a corresponding environment in the stream making it complete at the end of the step.
+
+The environments in the stream are partial containing only mappings for inputs.
 

### 2021-05-07 paper/example.tex
Add minor revisions to example

@@
 \figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV.
-It is loosely based on the system in the case study presented later.
+It is loosely based on the system in the case study introduced later.
 The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
@@
 The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract.
-For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression.
+For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
 There is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not not buffer events so the implicit input (or output) is a boolean value.
+The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value.
 An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port.
@@
 The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
-The \texttt{policy} is satisfied if there is a request with a response or nothing in the current time step (\texttt{today}) or a request yesterday with its response today.
-The value of \texttt{yesterday} in the current time step relies on values from the previous time step.
-The \texttt{yesterday} value uses the \texttt{->} operator for initialization.
-The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{(req and not rsp)} because seeing a request with no response is inconclusive at the start.
+The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}.
+The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
+The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system.
+The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
 The right operand is the value of \texttt{yesterday} after the initial step.
@@
 
-The value of \texttt{since} relies on its own value in the previous time step.
+The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step.
 The intuitive reading of the expression is that the alert has been true since the time when it first sounded.
@@
 Together with the other guarantees, the contract models the expected input and output of the system as a whole.
-Contracts also model the expected behavior of each component used to implement the system.
 
@@
 The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything).
-The counter-example showing each vio with a counterexample that violates the policy and does not raise the alert.
+The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.
 
@@
 The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised.
-The definition for \texttt{policy} is taken from the contract in \figref{fig:sw} since this high-assurance component is the monitor that needs to alert violations of the bounded response policy.
+The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}.
 As before the guarantees for the outputs are autogenerated by the tool.

### 2021-05-07 paper/contig.tex
Add love to contig types

--- 2021-05-06 paper/contig.tex
@@
-The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types} (cite contiguity).
+The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types} to define the input and output data (cite contiguity).
 A contiguity type is a self-describing specification for messages.
@@
 }
-\noindent Here the specification encodes the allowed ranges for each field of the waypoint. These can be checked while constructing a message from a sequence of bytes.
+\noindent Here the specification encodes the allowed ranges for each field of the waypoint.
+These assumptions restrict the resulting language to include only conforming messages and can be checked while constructing a message from a sequence of bytes.
+The notation $\LangTheta{\tau}$ denotes the language defined by the specification $\tau$ using the environment $\theta$ for expression evaluation.
 
-Every contiguity type specification has a corresponding \emph{matcher} that when given a message string returns true or false if that message belongs to the language of the specification.
+Every contiguity type specification has a corresponding CakeML \emph{matcher} that when given a message string returns true or false if that message belongs to the language of the specification.
 If the message does belong to the language, an \emph{environment} is provided to access each part of the message.
-An environment, $\theta: \lval \mapsto \konst{string}$ binds \emph{L-values} to strings.
-An L-value is an expression that can appear on the left hand side of an assignment.
-Alpha renaming is assumed so that every $\lval$ is unique.
+An environment, $\theta: \lval \mapsto \konst{string}$ binds \emph{L-values} to strings, where an L-value is an expression that can appear on the left hand side of an assignment (e.g., \texttt{AutomationRequest.Waypoints[0].Latitude}).
 
-The matcher itself is synthesized from the specification to CakeML.
-The synthesis includes a corresponding proof that the matcher recognizes the language of the specification, and the resulting environment, $\theta$, from a matched message produces the same message as the one matched when serialized.
+The main result of contiguity types is the proof of the relationship between the language of the specification and the synthesized matcher from the specification that is summarized below.
+\[
+  \konst{match}\; s_1s_2 = \konst{SOME}(\theta, s_2)
+  \imp \theta(\tau) \cdot s_2 = s_1s_2 \wedge s_1 \in \LangTheta{\tau}
+\]
+If there is a match on the substring $s_1$, then reconstituting the string from the resulting environment and concatenating it with $s_2$ yields the original string, and the matched string $s_1$ is in the language of the type specification.

### 2021-05-07 paper/synthesis.tex
Restructure tex files

@@
-\subsection{Filters}
-\subsection{Monitors}
-\subsection{Gates}
+A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs.
+The computation for the component in this work is defined entirely by a step function:
+\[
+\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
+\]
+The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
+\[
+\begin{array}{ll}
+ \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
+ (v_1,\ldots) & = \konst{readState}() ; \\
+ ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
+ \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
+ \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
+\end{array}
+\]
+It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.

### 2021-05-07 paper/system.tex
Restructure tex files

@@
 An environment, $\theta: \lval \mapsto \konst{string}$ binds L-values to strings.
-$\Delta : \konst{string} \to \mathbb{N}$ binds constant names to numbers.
-Functions $\konst{toN}:\konst{string}\to\mathbb{N}$ and $\konst{toB}:\konst{string}\to\konst{bool}$ interpret byte sequences to numbers and booleans, respectively.
 
@@
 Here, $\konst{eval}\; i\; e$ carries with it the index of the environment to be used for the expression.
+$\Delta : \konst{string} \to \mathbb{N}$ binds constant names to numbers.
+Functions $\konst{toN}:\konst{string}\to\mathbb{N}$ and $\konst{toB}:\konst{string}\to\konst{bool}$ interpret byte sequences to numbers and booleans, respectively.
 
@@
 In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification and then the guarantees are checked for invariance.
-
-The meaning of the specification is a set of finite environment streams where the guarantees hold at each step of the streams and the updated mappings or the \konst{eq}-statements are consistent with the statements.
-
-At the $i^\mathrm{th}$ step of the stream, $\theta^i$
-
-Each \emph{step} of the specification updates a corresponding environment in the stream making it complete at the end of the step.
-
-The environments in the stream are partial containing only mappings for inputs.
-
-\begin{comment}
-  A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs.
-The computation for the component in this work is defined entirely by a step function:
-\[
-\konst{stepFn} : \mathit{inports} \times \mathit{stateVars} \to \mathit{stateVars}
-\]
-The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
-\[
-\begin{array}{ll}
- \mathit{inportVals} & = \konst{readInputs}(); \\
- (v_1,\ldots,v_k) & = \konst{readState}() ; \\
- ({v_1}',\ldots,{v_k}') & = \konst{stepFn} (\mathit{inportVals},\mathit{stateVars}) ; \\
- \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots,{v_k}');} \\
- \multicolumn{2}{l}{\konst{writeOutputs}({v_1}',\ldots,{v_k}');} \\
-\end{array}
-\]
-It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
-\end{comment}

### 2021-05-07 paper/system.tex
Add summary of semantics

@@
 In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of the sequential evaluation of the \konst{eq}-statements in the specification and then the guarantees are checked for invariance.
+
+The notation $(\lval \mapsto \mathit{slice}) \bullet \theta$ denotes the addition of binding $\lval \mapsto \mathit{slice}$ to $\theta$.
+Create an eval function for \konst{eq}-statement.
+Map it to the list of statements.
+Create a similar function for the guarantees with its map that fails if invariance does not hold.
+Define the language of the specification using the contiguity notation.
+Need to munge all the types into a single $\tau$, but the gist is the language is any finite stream possible that conform to the input specification are the result of the eq-statements, and are invariant.
+The set is prefix closed (trace theory).

### 2021-05-07 paper/system.tex
Add notes on core language normal form

@@
                     &   & \konst{eq}\; [(f : \tau := \mathit{exp})\ldots] \\
-                    &   & \konst{guarantee}\ [\mathit{bexp}\ldots]\\\\
+                    &   & \konst{guarantee}\ [(\mathit{bexp})\ldots]\\\\
 
@@
 The language includes the initialization ($\rightarrow$), \konst{pre}, and if-then-else (\konst{ite}) operators.
+
+Change the following paragraph and preceding paragraphs to define the specification as the core language with a well defined normal form where expressions are flat, lvals are unique etc.
+It takes care of all the normal semantic checks related to types, dependency order, etc.
+Anything in the core language has normal form and is perfect.
 

### 2021-05-11 paper/agree.tex
Update with Hardin feedback

--- 2021-05-07 paper/example.tex
@@
-\begin{figure*}
-  \begin{center}
-    \begin{tabular}{c}
-      \includegraphics[scale=0.4]{example.png} \\
-      (a) \\
-      \includegraphics[scale=0.4]{example-certificate.png} \\
-      (b)
-    \end{tabular}
-  \end{center}
-\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
-\label{fig:example}
-\end{figure*}
-
-\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV.
-It is loosely based on the system in the case study introduced later.
-The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
-The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
-It is an \emph{as is} legacy component.
-The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.
-
-The goal is to cyber-harden the system so that it is resilient to cyber-attack.
-For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
-
 \newsavebox{\sw}
@@
 
-eq today : bool = (req = rsp);
-eq yesterday : bool = (req and not rsp) ->
+eq current : bool = (req = rsp);
+eq previous : bool = (req and not rsp) ->
                       pre(req and not rsp) and (not req and rsp);
-eq policy : bool = today or yesterday;
+eq policy : bool = current or previous;
 eq since : bool =  alrt or (alrt and (false -> pre(since)));
@@
 
-This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
-The contracts constrain the inputs and state properties of output for component models.
-In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component.
-The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See \cite{repo} for the full model.}
-
-The contract to model the SW component with its primary inputs and outputs is in \figref{fig:sw}.
-The contract language is a first order predicate calculus.
-The contract semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams.
+The revised AGREE specification for the SW component with its primary inputs and outputs is in \figref{fig:sw}.
+Without going into too much detail, the language is a first order predicate calculus.
+The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams.
 The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic.
@@
 
-There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
-This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.
-
 The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
-The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}.
-The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
-The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system.
-The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
-The right operand is the value of \texttt{yesterday} after the initial step.
+The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
+The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
+The value of \texttt{previous} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system.
+The left operand is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
+The right operand is the value of \texttt{previous} after the initial step.
 Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step.
-Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
+Intuitively, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
 
@@
 Together with the other guarantees, the contract models the expected input and output of the system as a whole.
-
-Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}.
-The results of the model checker are in \figref{fig:example}(b).
-The red exclamation points are properties that do not hold.
-Each comes with a counter-example.
-The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything).
-The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.
-
-\begin{figure*}
-  \begin{center}
-    \begin{tabular}{c}
-      \includegraphics[scale=0.3]{hardened.png} \\
-      (a) \\
-      \includegraphics[scale=0.4]{hardened-certificate.png} \\
-      (b)
-    \end{tabular}
-  \end{center}
-  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
-  \label{fig:hardened}
-\end{figure*}
 
@@
 eq req : bool = event(Request);
-eq today : bool = (req = rsp);
-eq yesterday : bool = (req and not rsp) ->
+eq current : bool = (req = rsp);
+eq previous : bool = (req and not rsp) ->
                       pre(req and not rsp) and (not req and rsp);
-eq policy : bool = today or yesterday;
+eq policy : bool = current or previous;
 eq alert : bool = (not policy) 
@@
 
-The system implementation is cyber-hardened by using the CASE tools to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
-A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
-Its stylized form as generated by the tool is shown in \figref{fig:assurance}(a).
+The hardened system is implemented by inserting the high-assurance components.
+The AGREE specifications for both are in \figref{fig:hardened}.
+Both are stylized for synthesis.
+The stylized filter form as generated by the tool is shown in \figref{fig:assurance}(a).
 The user must declare the meaning of the \texttt{policy} value as part of the transform on the model.
@@
 The output guarantees are automatically generated by the tool.
-
-A monitor captures a relation over time on input data and is able to reason about temporal properties of that input.
-By definition, it raises an alert output when the relation is violated.
-The generated monitor contract is shown in \figref{fig:assurance}(b).
+The monitor is shown in \figref{fig:assurance}(b).
 The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised.
@@
 
-The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b).
-The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.
-
-\newsavebox{\cml}
-\begin{lrbox}{\cml}
-\begin{lstlisting}[style=myML]
-fun filter_step () =
-let val () = Utils.clear_buf buffer
-    val () = API.callFFI "get_input" "" buffer
-    val string = Utils.buf2string buffer
-    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
-in
-    if policy
-    then
-        let val string = Word8Array.substring buffer 1
-                         (Word8Array.length buffer - 1)
-        in API.callFFI "put_output" string Utils.emptybuf
-        end
-    else print"Filter rejects message.\n"
-end
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\cml}}
-    \end{tabular}
-  \end{center}
-  \caption{Synthesized CakeML for the filter.}
-  \label{fig:cakeml}
-\end{figure}
-
-The high-assurance components are automatically synthesized from the contract models to equivalent models in the CakeML language.
-The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}.
-The code is called at dispatch by the scheduler.
-The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
-The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
-
-CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.
-A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML.
-A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML.
-In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.
-
-Preserving the input/output relationship of streams between the two models lifts the contract verification results to the deployed system.
-If the contract model verifies, then the meaning of those results hold for the deployed system if the other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.
+Add text to this section that defines the meaning of the contract over streams.
+State the prove guarantee from AGREE is terms of the streams.
+The statement will be used in the synthesis section to argue correctness.

### 2021-05-11 paper/example.tex
Update with Hardin feedback

@@
 
-\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV.
+\figref{fig:example} is an diagram form an AADL model of an implementation of a software system  (SW) for route planning and automated control for a UAV.
 It is loosely based on the system in the case study introduced later.
@@
 It is an \emph{as is} legacy component.
-The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.
 
-The goal is to cyber-harden the system so that it is resilient to cyber-attack.
-For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.
+The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contracts.
+The contracts constrain input and state properties of output for component models.
+AGREE model checks this assume-guarantee system to hierarchically prove the composite system obeys all contract obligations under all possible finite input streams.
 
-\newsavebox{\sw}
-\begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree]
-eq req : bool = event(AutomationRequest);
-eq avl : bool = event(AirVehicleLocation);
-eq wp : bool = event(Waypoint);
-eq rsp: bool = event(Start);
-eq alrt : bool = event(Alert);
+The initial AGREE specification for the components and SW system make no assumptions about the integrity of the inputs and outputs.
+The cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
+As such, the AGREE specification for the AI route planner is updated to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.
 
-assume "Automation request is well-formed" :
-    req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
-assume "Air vehicle location is well-formed" :
-    avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);
+The contracts for the other components are also updated with the threat analysis information.
+For example, the specification for the legacy waypoint manager adds assumptions about about its input being \emph{well-formed} since that is no longer known \emph{a priori} as AI route planner is now untrusted.
 
-eq today : bool = (req = rsp);
-eq yesterday : bool = (req and not rsp) ->
-                      pre(req and not rsp) and (not req and rsp);
-eq policy : bool = today or yesterday;
-eq since : bool =  alrt or (alrt and (false -> pre(since)));
+There are two requirements added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
+Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.
 
-guarantee "Start includes a waypoint" :
-    rsp => wp;
-guarantee "Locations required after the start waypoint" :
-    (wp and not rsp) => avl;
-guarantee "Waypoint is well-formed" :
-    wp => WELL_FORMED_WAYPOINT(Waypoint);
-guarantee "Alert if start is not bounded relative to a request" :
-    policy or since;
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}
-  \begin{center}
-    \scalebox{0.60}{\usebox{\sw}}
-  \end{center}
-  \caption{The SW component contract.}
-  \label{fig:sw}
-\end{figure}
-
-This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE).
-The contracts constrain the inputs and state properties of output for component models.
-In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component.
-The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See \cite{repo} for the full model.}
-
-The contract to model the SW component with its primary inputs and outputs is in \figref{fig:sw}.
-The contract language is a first order predicate calculus.
-The contract semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams.
-The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic.
-Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream.
-The semantics do not model computation or communication delay.
-The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
-
-The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract.
-For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
-There is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value.
-An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port.
-The system contract assumes well-formed input and guarantees a set of properties about the output.
-
-There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
-This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.
-
-The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
-The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}.
-The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
-The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system.
-The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
-The right operand is the value of \texttt{yesterday} after the initial step.
-Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step.
-Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
-
-The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step.
-The intuitive reading of the expression is that the alert has been true since the time when it first sounded.
-The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds.
-The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system.
-Together with the other guarantees, the contract models the expected input and output of the system as a whole.
-
-Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}.
+AGREE model checks the composition of the new specifications for the implementation in \figref{fig:example}(a) to see if it complies with the added cyber-resilience properties.
 The results of the model checker are in \figref{fig:example}(b).
@@
 Each comes with a counter-example.
-The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything).
+The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager.
 The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.
+A clear malicious behavior from the untrusted component.
 
@@
 
-\newsavebox{\flt}
-\begin{lrbox}{\flt}
-\begin{lstlisting}[style=agree]
-eq policy : bool = 
-  WELL_FORMED_AUTOMATION_RESPONSE(Input);       
-guarantee Filter_Output "Filter output is well-formed" :
-  if event(Input) and policy then 
-    event(Output) and Output = Input
-  else not event(Output);
-\end{lstlisting}
-\end{lrbox}
-
-\newsavebox{\mntr}
-\begin{lrbox}{\mntr}
-\begin{lstlisting}[style=agree]
-const is_latched : bool = 
-  Get_Property(this, CASE_Properties::Monitor_Latched);
-eq rsp : bool = event(Response);
-eq req : bool = event(Request);
-eq today : bool = (req = rsp);
-eq yesterday : bool = (req and not rsp) ->
-                      pre(req and not rsp) and (not req and rsp);
-eq policy : bool = today or yesterday;
-eq alert : bool = (not policy) 
-               -> ((is_latched and pre(alert)) or not policy);
-guarantee Monitor_Alert 
-  "Alert port tracks alert variable" :
-  event(Alert) = alert;
-guarantee Monitor_Output
-  "Output if not alerted" :
-  if event(Alert) then (not event(Output)) else
-  if event(Response) then (event(Output) and (Output = Response))
-  else (not event(Output));
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\flt}} \\
-      (a) \\
-      \scalebox{0.60}{\usebox{\mntr}} \\
-      (b)
-    \end{tabular}
-  \end{center}
-  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
-  \label{fig:assurance}
-\end{figure}
-
-The system implementation is cyber-hardened by using the CASE tools to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
+The system implementation is cyber-hardened by automatically transforming the model to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
 A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
-Its stylized form as generated by the tool is shown in \figref{fig:assurance}(a).
-The user must declare the meaning of the \texttt{policy} value as part of the transform on the model.
-In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed.
-The output guarantees are automatically generated by the tool.
+The auto-generated AGREE specification states that only well-formed inputs are passed to the output.
+The user must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
 
 A monitor captures a relation over time on input data and is able to reason about temporal properties of that input.
-By definition, it raises an alert output when the relation is violated.
-The generated monitor contract is shown in \figref{fig:assurance}(b).
-The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised.
-The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}.
-As before the guarantees for the outputs are autogenerated by the tool.
+It raises an alert if ever the specified temporal properties are violated.
+The auto-generated AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and that response must come with the request or in the next step after the request.
+As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.
 
-The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b).
-The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.
+The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}(b).
+Here AGREE gives a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber vulnerability from the untrusted AI route planner.
 
-\newsavebox{\cml}
-\begin{lrbox}{\cml}
-\begin{lstlisting}[style=myML]
-fun filter_step () =
-let val () = Utils.clear_buf buffer
-    val () = API.callFFI "get_input" "" buffer
-    val string = Utils.buf2string buffer
-    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
-in
-    if policy
-    then
-        let val string = Word8Array.substring buffer 1
-                         (Word8Array.length buffer - 1)
-        in API.callFFI "put_output" string Utils.emptybuf
-        end
-    else print"Filter rejects message.\n"
-end
-\end{lstlisting}
-\end{lrbox}
+The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language.
+CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.
 
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\cml}}
-    \end{tabular}
-  \end{center}
-  \caption{Synthesized CakeML for the filter.}
-  \label{fig:cakeml}
-\end{figure}
-
-The high-assurance components are automatically synthesized from the contract models to equivalent models in the CakeML language.
-The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}.
-The code is called at dispatch by the scheduler.
-The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
-The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
-
-CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.
 A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML.

### 2021-05-11 paper/intro.tex
Update with Hardin feedback

@@
+In recent years, aerospace stakeholders have become aware that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
+In addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
+Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
+
+Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.
+Cyber vulnerabilities are often discovered during penetration testing late in the development process.
+Worse yet, they may be discovered after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
+This is not a sustainable development model.
+
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.
+
+BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.
+A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system.
+Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020}.
+Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
+Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
+If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.
+
+A novel aspect of the approach is the ability to automatically transformation the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities.
+The behavior of the model is given by contracts on components expressed in the AGREE modeling language.
+These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system.
+In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.
+
+The two automatic transforms discussed in this paper are the insertion of filters to prevent malformed data from malicious actors from being propagated to downstream components, and monitors to detect and alert unexpected behaviors arising from untrusted components.
+These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language.
+Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.
+
+The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML.
+This paper agues that synthesis to CakeML to be correct.
+CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
+Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system in that in detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
+Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.
+
+The approach is motivated, and illustrated, in a simple example.
+A full case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is briefly reported.
+Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
+The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.
+
+\begin{comment}
 Here is the basic process:
@@
 Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.
+\end{comment}

### 2021-05-11 paper/synthesis.tex
Update with Hardin feedback

@@
 It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+
+\newsavebox{\cml}
+\begin{lrbox}{\cml}
+\begin{lstlisting}[style=myML]
+fun filter_step () =
+let val () = Utils.clear_buf buffer
+    val () = API.callFFI "get_input" "" buffer
+    val string = Utils.buf2string buffer
+    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
+in
+    if policy
+    then
+        let val string = Word8Array.substring buffer 1
+                          (Word8Array.length buffer - 1)
+        in API.callFFI "put_output" string Utils.emptybuf
+        end
+    else print"Filter rejects message.\n"
+end
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\cml}}
+    \end{tabular}
+  \end{center}
+  \caption{Synthesized CakeML for the filter.}
+  \label{fig:cakeml}
+\end{figure}
+
+The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}.
+The code is called at dispatch by the scheduler.
+The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
+The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
+The auto-generated AGREE specification raises an alert output when the relation is violated.

### 2021-05-11 paper/abstract.tex
Fix typos
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

@@
-Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they recover or safely shut down as requirements dictate.
-The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in this cyber-resiliency in a Model-Based Systems Engineering environment.
-This paper discusses the automated model transformations that add in high-assurance components.
-These components are filters and monitors that prevent malicious input and detect supply chain attacks respectively.
-A formal specification defines each component and is used to verify that the component addresses system level cyber-requirements.
-Implementations are synthesized from the specification, and that synthesis is proven to preserve the exact meaning of the specification.
+Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either recover or safely shut down as requirements dictate.
+The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment.
+This paper describes automated model transformations that introduce  high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively.
+A formal specification defines each component and is used to verify that the component addresses system level cyber requirements.
+Implementations are synthesized from the specifications, and that synthesis is proven to preserve the exact meaning of the specifications.
 The transformations are integrated into the Open Source AADL Tool Environment (OSATE).
 The paper further reports a case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
-Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
+In the case study, the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-11 paper/contig.tex
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

@@
-The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types} to define the input and output data (cite contiguity).
+The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types}~\cite{contiguity-types} to define the input and output data (cite contiguity).
 A contiguity type is a self-describing specification for messages.

### 2021-05-11 paper/example.tex
Fix fig and add link to repo
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

@@
 
-\figref{fig:example} is an diagram form an AADL model of an implementation of a software system  (SW) for route planning and automated control for a UAV.
+\figref{fig:example} is an AADL model of an implementation of a software system (SW) for route planning and automated control for a UAV.
 It is loosely based on the system in the case study introduced later.
+The source for the entire model is found at \cite{repo}.
 The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
@@
 
-There are two requirements added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
+Two requirements are added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
 Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.
 
-AGREE model checks the composition of the new specifications for the implementation in \figref{fig:example}(a) to see if it complies with the added cyber-resilience properties.
+AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber-resilience properties, by way of model checking.
 The results of the model checker are in \figref{fig:example}(b).

### 2021-05-11 paper/intro.tex
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
+  views expressed are those of the authors and do not reflect the
+  official policy or position of the Defense Advanced
+  Research Projects Agency (DARPA) or the U.S. Government.}
 The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
@@
 A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system.
-Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020}.
+Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
 Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.

### 2021-05-11 paper/paper.tex
Add contig types
Restructure tex files
Update with Hardin feedback
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

@@
   Synthesizing Verified Components for Cyber Assured Systems Engineering 
-  \thanks{
-    Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001.
+  \thanks{{DISTRIBUTION STATEMENT A.  Approved for public release.}
   }
@@
 \author{
+\IEEEauthorblockN{Eric Mercer}
+\IEEEauthorblockA{Department of Computer Science\\
+Brigham Young University\\
+Provo, Utah}
+\and
 \IEEEauthorblockN{
-Konrad Slind, 
-Junaid Babar, 
-Isaac Amundson, and
+Isaac Amundson and
+Konrad Slind}
+\IEEEauthorblockA{
+Trusted Systems Group   \\
+Collins Aerospace       \\
+Minneapolis, Minnesota}
+\and
+\IEEEauthorblockN{
+Junaid Babar and 
 David Hardin}
@@
 Collins Aerospace       \\
-Minneapolis, Minnesota  \\
+Cedar Rapids, Iowa}
 }
-\and
-\IEEEauthorblockN{Eric Mercer}
-\IEEEauthorblockA{Department of Computer Science\\
-Brigham Young University\\
-Provo, Utah}
-}
-
 % conference papers do not typically use \thanks and this command
@@
 
-% \section{Introduction}
-% \input{intro}
+\section{Introduction}
+\input{intro}
 
@@
 
-\section{System Model and Semantics}
-\input{system}
+\section{AGREE Specification}
+\input{agree}
 
-% \section{Contiguity Types}
-% \input{contig}
+\section{Synthesis}
+\input{synthesis}
 
-% \section{Filters}
-% \input{filters}
-
-% \section{Monitors}
-% \input{monitors}
-
-% \section{OSATE Integration for AADL}
-% \input{aadl-agree-osate}
+\section{UxAS Case Study}
+\input{case-study}
 
@@
 \section*{Acknowledgments}
-This work is funded by DARPA contract HR00111890001. The views, opinions and/or findings expressed are those of the authors and
-should not be interpreted as representing the official views or policies of the Department of Defense or the U.S. Government.
+This work was sponsored in part by the Defense Advanced Research
+Projects Agency (DARPA).
 
@@
 \bibliography{paper}
+
+\appendix
+\section{Contiguity Types}
+\input{contig}
+
+\section{System Model and Semantics}
+\input{system}
+
 \end{document}

### 2021-05-12 paper/agree.tex
Add text on synthesis

@@
 
-Add text to this section that defines the meaning of the contract over streams.
+\emph{Add text to this section that defines the meaning of the contract over streams.
 State the prove guarantee from AGREE is terms of the streams.
-The statement will be used in the synthesis section to argue correctness.
+The statement will be used in the synthesis section to argue correctness.}

### 2021-05-12 paper/example.tex
Add text on synthesis

@@
 It raises an alert if ever the specified temporal properties are violated.
-The auto-generated AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and that response must come with the request or in the next step after the request.
+The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request.
 As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.
@@
 The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language.
-CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.
+CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.
 
 A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML.
-A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML.
-In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.
+A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the generated code.
+In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.
 

### 2021-05-12 paper/intro.tex
Add text on synthesis

@@
 
-A novel aspect of the approach is the ability to automatically transformation the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities.
+A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities.
 The behavior of the model is given by contracts on components expressed in the AGREE modeling language.
@@
 
-The two automatic transforms discussed in this paper are the insertion of filters to prevent malformed data from malicious actors from being propagated to downstream components, and monitors to detect and alert unexpected behaviors arising from untrusted components.
+The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components.
 These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language.
@@
 The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML.
-This paper agues that synthesis to CakeML to be correct.
+This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
 CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
-Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system in that in detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
+Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
 Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

### 2021-05-12 paper/synthesis.tex
Add text on synthesis

@@
-A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs.
-The computation for the component in this work is defined entirely by a step function:
-\[
-\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
-\]
-The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
-\[
-\begin{array}{ll}
- \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
- (v_1,\ldots) & = \konst{readState}() ; \\
- ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
- \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
- \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
-\end{array}
-\]
-It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+Synthesis maps from model and specifications to code. The synthesis algorithm traverses the system architecture looking
+for occurrences of filter and monitor specifications. For each such 
+occurrence it generates a CakeML program. In the following, we examine both filter and monitor synthesis. The latter is typically much more involved, and we will devote more attention to it.
+
+\subsection{Filter Generation}
+
+A filter is intended to be very simple; it is expected to have one
+input port and one output; messages on the input that the filter
+policy admits pass unchanged to the output port; all others are not
+passed on. We have explored in our work two kinds of filter. In the
+first, a relatively shallow scan of the input buffer can enforce the
+policy. For example, we have used the expressive power of contiguity
+types \cite{contig} to enforce \emph{lightweight} bounds constraints
+on GPS coordinates in UxAS messages. On the other hand, the second
+kind of filter will parse the input buffer into a data structure
+specified in AGREE and apply an user-defined \emph{well-formedness}
+property, also specified in AGREE, to the data. This allows
+arbitrarily complex well-formedness checking.
+
+The decision of a filter is made and performed within one thread
+invocation. Thus, in its given time slice, the filter does the following:
+
+\begin{enumerate}
+
+\item checks to see if there is any input available; if there is none
+then it yields control; otherwise,
+
+\item the input is read (and parsed if need be),
+
+\item the wellformedness predicate is evaluated,
+
+\item if the predicate returns \konst{true} then the input buffer
+ is copied to the output; otherwise no action is taken, and
+
+\item the filter yields control
+\end{enumerate}
+
+\begin{remark}[Partiality]
+
+The role of partiality should be emphasized: steps 2 and 3 can fail;
+the data might not be parseable or the wellformedness computation
+could be badly written and fail at run time. In such cases, the filter
+should recover and yield control without passing the input onwards. In
+these cases, the filter is behaving as it should, but there are also
+cases in which a correctly specified filter would not meet its
+specification at runtime. This situation arises when the
+filter \emph{ought} to accept a message, but lack of resources result
+in the filter failing to do so. Examples of this would be, for
+example, if the parse of a message needed more space than allocated;
+another example would be if the timeslice provided by the scheduler is
+too small for the wellformedness computation to finish.
+
+\end{remark}
+
+\emph{Need some discussion of filters and their step-wise propertie s in relation to their infinitary properties. Reference to Johannes' work.}
 
@@
     val () = API.callFFI "get_input" "" buffer
-    val string = Utils.buf2string buffer
-    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
 in
-    if policy
+    if WELL_FORMED_AUTOMATION_RESPONSE buffer
     then
-        let val string = Word8Array.substring buffer 1
-                          (Word8Array.length buffer - 1)
-        in API.callFFI "put_output" string Utils.emptybuf
-        end
+      API.callFFI "put_output" buffer Utils.emptybuf
     else print"Filter rejects message.\n"
@@
   \caption{Synthesized CakeML for the filter.}
-  \label{fig:cakeml}
+  \label{fig:filter-cakeml}
 \end{figure}
 
-The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}.
+The synthesized CakeML code for the filter is shown in \figref{fig:filter-cakeml}.
 The code is called at dispatch by the scheduler.
@@
 The auto-generated AGREE specification raises an alert output when the relation is violated.
+
+% A \emph{system} is a collection of \emph{components}, \emph{connections}
+% between components, a \emph{scheduler} to order execution, and a
+% \emph{system environment} for primary inputs.
+
+\subsection{Monitor Generation}
+
+Monitors, since they are intended to track and analyze the externally visible behavior of system components through time, require more computational features than filters.
+In particular, our conception of a monitor is a predicate on its input (and output) streams, being able to access the value of a stream at any earlier point in time,  if necessary.
+Thus monitors commonly use state to keep track of earlier values, unlike filters which are stateless.
+This reasoning leads us to specify the  computation for a monitor component  by a step function of the following form:
+\[
+\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
+\]
+Thus, in its given timeslice, a monitor evaluates the \konst{stepFn} on its inputs and the current values of the state variables. In detail, it takes the following steps:
+
+\begin{enumerate}
+
+\item each available input is parsed into data of the type specified
+by the port type;
+
+\item the stateful variables in $\mathit{stateVars}$  are evaluated in dependency order.
+
+\item values of outputs are computed
+
+\item outputs are written and the new state is written
+
+\item control is yielded
+\end{enumerate}
+
+Our earlier remarks on partiality apply here too, of course.
+
+The scheduler \emph{activates} components in some order.It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+ Activating a monitor component takes the form of the following pseudo-code: 
+\[
+\begin{array}{ll}
+ \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
+ (v_1,\ldots) & = \konst{readState}() ; \\
+ ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
+ \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
+ \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
+\end{array}
+\]
+
+\paragraph{Initialization}
+A monitor may need to make a certain minimum number of observations before being able to make a meaningful assessment of behavior.
+Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization}.
+In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases.
+For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at time $t$ and $t+1$ is such that the distance between the two locations is unusually large.
+Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made.
+The specification for this monitor has to explicitly say what the correct first output is.

### 2021-05-12 paper/example.tex
Revise intro and example

@@
 
-The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contracts.
+The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contract specifications.
 The contracts constrain input and state properties of output for component models.
@@
 
-The initial AGREE specification for the components and SW system make no assumptions about the integrity of the inputs and outputs.
-The cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
-As such, the AGREE specification for the AI route planner is updated to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.
+The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs.
+A resulting cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
+As such, the designer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.
 
-The contracts for the other components are also updated with the threat analysis information.
-For example, the specification for the legacy waypoint manager adds assumptions about about its input being \emph{well-formed} since that is no longer known \emph{a priori} as AI route planner is now untrusted.
+The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities.
+For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner is now untrusted.
+The systems designer is also responsible for defining the meaning of well-formed, and in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude with an additional check on IDs in the message for integrity.
 
-Two requirements are added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request.
-Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.
+Two other requirements are added to the SW AGREE specification by the designer related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream); and the second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request).
+Here the goal of the requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.
 
-AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber-resilience properties, by way of model checking.
+AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber requirements, by way of model checking.
 The results of the model checker are in \figref{fig:example}(b).
@@
 The auto-generated AGREE specification states that only well-formed inputs are passed to the output.
-The user must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
+The system designer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
 
@@
 The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request.
-As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.
+As with the filter, the system designer provides the policy, and that policy is based on the existing AGREE specification in the SW system.
 

### 2021-05-12 paper/intro.tex
Revise intro and example

@@
 
-A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities.
-The behavior of the model is given by contracts on components expressed in the AGREE modeling language.
-These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system.
-In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.
+A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to satisfy requirements added to mitigate cyber vulnerabilities.
+In this work, the behavior of the system is modeled by contracts on components expressed in the AGREE modeling language.
+These contracts are modified as a result of the cyber-vulnerability analysis with the new requirements for mitigation.
+The requirements assume properties on the input and guarantee properties of the output that reflect a hardened system.
+In the absence of any transformation to implement such requirements, AGREE fails to prove the system meets these new requirements.
 
@@
 These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language.
-Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.
+Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.
 
-The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML.
+The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML which is a verified implementation of ML \cite{cakeml}.
 This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
 CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
-Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
+Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
 Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.
@@
 The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.
+
+All the aforementioned tools are publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4, demonstration videos, etc.
+\cite{phase2, camkes} ---ADD LINK TO VIDEO.
 

### 2021-05-12 paper/agree.tex
Add formalism to AGREE semantics

--- 2021-05-11 paper/agree.tex
@@
+The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
+The AGREE specifications generated by the transforms for the high-assurance components are explained.
+The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component.
+This meaning is what must be preserved by synthesis.
+
 \newsavebox{\sw}
@@
 
-The revised AGREE specification for the SW component with its primary inputs and outputs is in \figref{fig:sw}.
+The AGREE specification for the SW component in the example with the added cyber requirements is given in \figref{fig:sw}.
 Without going into too much detail, the language is a first order predicate calculus.
-The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams.
-The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic.
+The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions.
+The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.
 Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream.
@@
 
-The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract.
+The AGREE model checker attempts to prove several properties of the top-level model being verified.
+The first is that the output guarantees of each component implementing the system are strong enough to guarantee the input assumptions of any downstream component and strong enough to satisfy the guarantees of the top-level component's output that is being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output).
+These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
+The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself).
+These are the unexpanded results to the same two figures.
+
+Returning back to the contract in \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
 For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
-There is an implicit \emph{event} input (or output) associated with every named event port in a component.
+In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
 The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value.
 An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port.
-The system contract assumes well-formed input and guarantees a set of properties about the output.
+The system contract here states assumptions on well-formed input, and it then states guarantees on properties about the output.
 
-The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
+The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
 The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
 The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
-The value of \texttt{previous} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system.
-The left operand is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
+
+The value of \texttt{previous} in the current time step relies on values from the previous time step.
+The \texttt{->} operator is for initialization as the previous time step is undefined in the first step of the system.
+The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
 The right operand is the value of \texttt{previous} after the initial step.
@@
 The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds.
-The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system.
+The \emph{Alert if start is not bounded relative to a request} guarantee defines one requirement of a cyber-hardened system implementation.
 Together with the other guarantees, the contract models the expected input and output of the system as a whole.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.3]{dialogue.png}
+    \end{tabular}
+  \end{center}
+  \caption{Dialogue for automatically transforming the model with a filter.}
+  \label{fig:dialogue}
+\end{figure}
 
@@
 
-The hardened system is implemented by inserting the high-assurance components.
-The AGREE specifications for both are in \figref{fig:hardened}.
-Both are stylized for synthesis.
-The stylized filter form as generated by the tool is shown in \figref{fig:assurance}(a).
-The user must declare the meaning of the \texttt{policy} value as part of the transform on the model.
-In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed.
-The output guarantees are automatically generated by the tool.
-The monitor is shown in \figref{fig:assurance}(b).
-The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised.
-The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}.
-As before the guarantees for the outputs are autogenerated by the tool.
+As noted previously, the original system fails to guarantee the cyber requirement.
+BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor.
+The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation.
+The dialogue or a filter transformation is shown in \figref{fig:dialogue}.
+The system designer can state the policy directly in the dialogue, or leave it  blank.
+In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}.
+Additionally the filter can be attached to any requirement the vulnerability analysis.
+This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021].
+The \emph{OK} button creates the component and inserts it accordingly into the implementation
 
-Add text to this section that defines the meaning of the contract over streams.
-State the prove guarantee from AGREE is terms of the streams.
-The statement will be used in the synthesis section to argue correctness.
+The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a).
+The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input.
+As similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed.
+The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).
+The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised and is one of the several options available in the dialogue.
+The definition for \texttt{policy} is taken by the system designer from the contract in \figref{fig:sw}.
+As before the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
+
+At this point it is important to state the meaning of a high-assurance component more formally.
+Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component such as what might be sent over a wire by the communication fabric.
+An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL.
+As such, it binds \emph{L-values} to strings where an L-value is a concept from compilers and represents anything that can appear on the left-hand side of an assignment.
+In the context of this presentation, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc.
+The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.
+
+Let $s$ be the AGREE contract specification for a high-assurance component.
+The function, $\konst{eval}\ s\ \theta\dots$, evaluates $s$ on the given stream of total environments, $\theta\dots$, and returns \konst{true} if the specification is invariant along the entire stream and \konst{false} otherwise.
+The meaning of invariance for the specification merits some discussion.
+A guarantee has its intuitive meaning: the truth value of the associated expression at every step of the stream must be preserved.
+An eq-statement, however, is treated as a guarantee rather than a computation that adds a new binding to the environment.
+It asserts an equivalence between the named value and its associated expression.
+That equivalence must be preserved at every step of the stream.
+
+The set $\Theta_s$ is the universe of environment bindings for the given specification $s$. 
+The meaning of an AGREE contract specification is now defined as follows.
+%
+\[
+  \mathcal{L}(s) = \{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots\}  
+\]
+%
+The meaning is the prefix closed set of finite environment streams on which the specification is invariant.
+Each environment in the stream is total meaning that it binds all L-values in the scope if the specification.
+
+This paper claims that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely.
+Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings.
+Let $\Theta_I$ be the universe of environments that only bind the input for the specification.
+This paper claims the following.
+%
+\[
+  \forall \theta \in \Theta_I^*\ (\konst{eval}\ \konst{synth}(s)\ \theta\ldots) \in \mathcal{L}(s)) 
+\]
+%
+As noted previously, the semantics of the AGREE specification language are understood in terms of streams of data.

### 2021-05-12 paper/synthesis.tex
Fixed Contiguity Types citation

@@
 first, a relatively shallow scan of the input buffer can enforce the
-policy. For example, we have used the expressive power of contiguity
-types \cite{contig} to enforce \emph{lightweight} bounds constraints
+policy. For example, we have used the expressive power of Contiguity
+Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints
 on GPS coordinates in UxAS messages. On the other hand, the second

### 2021-05-13 paper/related-work.tex
Bibliography fixes

@@
-Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, composition1, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}.
+Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, agree2013, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}.
 Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}.

### 2021-05-13 paper/abstract.tex
Abstract Wordsmithing

@@
 This paper describes automated model transformations that introduce  high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively.
-A formal specification defines each component and is used to verify that the component addresses system level cyber requirements.
-Implementations are synthesized from the specifications, and that synthesis is proven to preserve the exact meaning of the specifications.
-The transformations are integrated into the Open Source AADL Tool Environment (OSATE).
-The paper further reports a case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
-In the case study, the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
+A formal specification defines each component, and is used to verify that the component addresses system level cyber requirements.
+Implementations are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level.
+The model transformations are integrated into the Open Source AADL Tool Environment (OSATE).
+The paper further reports on a case study applying security-enhancing model transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
+In the case study, the model transformations add filters to guard against malformed input, as well as monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-13 paper/paper.tex
Add formalism to AGREE semantics
Remove appendix
Abstract Wordsmithing

@@
 \IEEEauthorblockA{
-Trusted Systems Group   \\
-Collins Aerospace       \\
+Applied Research and Technology\\
+Collins Aerospace\\
 Minneapolis, Minnesota}
@@
 \IEEEauthorblockA{
-Trusted Systems Group   \\
-Collins Aerospace       \\
+Applied Research and Technology\\
+Collins Aerospace\\
 Cedar Rapids, Iowa}
@@
 
-\section{AGREE Specification}
+\section{AGREE Contract Specification}
 \input{agree}
@@
 
-\appendix
-\section{Contiguity Types}
-\input{contig}
+% \appendix
+% \section{Contiguity Types}
+% \input{contig}
 
-\section{System Model and Semantics}
-\input{system}
+% \section{System Model and Semantics}
+% \input{system}
 

### 2021-05-13 paper/example.tex
Updates from Overleaf

@@
 
-The system implementation is cyber-hardened by automatically transforming the model to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
+The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
 A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
@@
 
-The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language.
+The high-assurance components are automatically synthesized by SPLAT from the AGREE specifications to equivalent models in the CakeML language.
 CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.

### 2021-05-13 paper/synthesis.tex
Updates from Overleaf

@@
 \paragraph{Initialization}
-A monitor may need to make a certain minimum number of observations before being able to make a meaningful assessment of behavior.
-Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization}.
+A monitor may need to accumulate a certain minimum number of observations before being able to make a meaningful assessment of behavior.
+Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization} phase.
 In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases.
-For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at time $t$ and $t+1$ is such that the distance between the two locations is unusually large.
+For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at times $t$ and $t+1$ is such that the distance between the two locations is unusually large.
 Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made.
-The specification for this monitor has to explicitly say what the correct first output is.
+The specification has to explicitly say what the correct first output is.

### 2021-05-13 paper/agree.tex

@@
 The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component.
-This meaning is what must be preserved by synthesis.
+This meaning is what must be preserved by the synthesis.
 
@@
 The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.
-Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream.
+Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
 The semantics do not model computation or communication delay.
@@
 The AGREE model checker attempts to prove several properties of the top-level model being verified.
-The first is that the output guarantees of each component implementing the system are strong enough to guarantee the input assumptions of any downstream component and strong enough to satisfy the guarantees of the top-level component's output that is being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output).
+The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component and strong enough to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output).
 These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
 The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself).
-These are the unexpanded results to the same two figures.
+These are the unexpanded results at the bottom of the figures.
 
@@
 In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value.
-An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port.
-The system contract here states assumptions on well-formed input, and it then states guarantees on properties about the output.
+The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
+An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
+The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.
 
@@
   \end{center}
-  \caption{Dialogue for automatically transforming the model with a filter.}
+  \caption{Wizard for automatically transforming the model with a filter.}
   \label{fig:dialogue}
@@
 The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation.
-The dialogue or a filter transformation is shown in \figref{fig:dialogue}.
-The system designer can state the policy directly in the dialogue, or leave it  blank.
+The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}.
+The policy of the high assurance component can be stated directly in the wizard, or leave it  blank.
 In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}.
-Additionally the filter can be attached to any requirement the vulnerability analysis.
+Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.
+Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run anytime throughout the design phase to determine whether those artifacts are valid.
+%Additionally, the filter can be attached to any requirement the vulnerability analysis.
 This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021].
@@
 The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input.
-As similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed.
+%A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. 
 The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).

### 2021-05-13 paper/intro.tex

@@
 The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components.
-These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language.
+These transformations not only change the architecture of the model by adding in new components, but they generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
 Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.
 
-The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML which is a verified implementation of ML \cite{cakeml}.
+The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, which is a verified implementation of ML \cite{cakeml}.
 This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
@@
 The approach is motivated, and illustrated, in a simple example.
-A full case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is briefly reported.
+A case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented.
 Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-14 paper/intro.tex
Wordsmiting; added CakeML infinitary proof papers

@@
-In recent years, aerospace stakeholders have become aware that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
-In addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
+In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
+Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
 Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
@@
 Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.
-Cyber vulnerabilities are often discovered during penetration testing late in the development process.
-Worse yet, they may be discovered after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
+Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
 This is not a sustainable development model.
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
+The DARPA Cyber Assured Systems Engineering (CASE) program is focused on developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
   views expressed are those of the authors and do not reflect the
@@
   Research Projects Agency (DARPA) or the U.S. Government.}
-The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE}, based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
 BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.
@@
 Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
-Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
+The AGREE analysis \emph{assumes}  properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
+Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
 If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.
 
-A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to satisfy requirements added to mitigate cyber vulnerabilities.
-In this work, the behavior of the system is modeled by contracts on components expressed in the AGREE modeling language.
-These contracts are modified as a result of the cyber-vulnerability analysis with the new requirements for mitigation.
-The requirements assume properties on the input and guarantee properties of the output that reflect a hardened system.
-In the absence of any transformation to implement such requirements, AGREE fails to prove the system meets these new requirements.
+%% DSH  This paragraph seems mostly duplicative of the paragraph above; attempted to merge the two.
+%A novel aspect of the BriefCASE approach is the ability to automatically transform the model to insert high-assurance %components to satisfy requirements added to mitigate cyber vulnerabilities.
+In this work, the behavior of the system is %modeled by contracts on components expressed in the AGREE modeling language.
+These contracts are modified as a result %of the cyber-vulnerability analysis with the new requirements for mitigation.
+The requirements assume properties on the %input and guarantee properties of the output that reflect a hardened system.
+In the absence of any transformation to %implement such requirements, AGREE fails to prove the system meets these new requirements.
 
 The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components.
-These transformations not only change the architecture of the model by adding in new components, but they generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
+These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
 Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.
 
-The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, which is a verified implementation of ML \cite{cakeml}.
+Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}.
 This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
@@
 Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
-Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.
+Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
 
@@
 \begin{comment}
-Here is the basic process:
+The basic process proceeds as follows:
 \begin{enumerate}

### 2021-05-14 paper/intro.tex
Defined UAV acronmym

@@
 The approach is motivated, and illustrated, in a simple example.
-A case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented.
+A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented.
 Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-14 paper/intro.tex
Added Section labels

--- 2021-05-14 paper/intro.tex
@@
 
-The approach is motivated, and illustrated, in a simple example.
-A case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented.
+The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}.
+A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}.
 Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-14 paper/synthesis.tex
adding contig type info for filter

@@
-Synthesis maps from model and specifications to code. The synthesis algorithm traverses the system architecture looking
-for occurrences of filter and monitor specifications. For each such 
-occurrence it generates a CakeML program. In the following, we examine both filter and monitor synthesis. The latter is typically much more involved, and we will devote more attention to it.
+Synthesis maps from model and specifications to code. The synthesis
+algorithm traverses the system architecture looking for occurrences of
+filter and monitor specifications;  for each such occurrence it
+generates a CakeML program. In the following, we examine both filter
+and monitor synthesis. The latter is typically much more involved, and
+we will therefore devote more attention to it.
 
@@
 input port and one output; messages on the input that the filter
-policy admits pass unchanged to the output port; all others are not
-passed on. We have explored in our work two kinds of filter. In the
-first, a relatively shallow scan of the input buffer can enforce the
-policy. For example, we have used the expressive power of Contiguity
-Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints
-on GPS coordinates in UxAS messages. On the other hand, the second
-kind of filter will parse the input buffer into a data structure
-specified in AGREE and apply an user-defined \emph{well-formedness}
-property, also specified in AGREE, to the data. This allows
-arbitrarily complex well-formedness checking.
+policy admits pass unchanged to the output port; all others are
+dropped (not passed on). We have explored in our work two kinds of
+filter. In the first, a relatively shallow scan of the input buffer
+can enforce the policy. For example, we have used the expressive power
+of Contiguity Types \cite{contiguity-types} to enforce
+\emph{lightweight} bounds constraints on GPS coordinates in UxAS
+messages. On the other hand, the second kind of filter will parse the
+input buffer into a data structure specified in AGREE and apply an
+user-defined \emph{well-formedness} property, also specified in AGREE,
+to the data. This allows arbitrarily complex well-formedness checking.
 
@@
 
-\emph{Need some discussion of filters and their step-wise propertie s in relation to their infinitary properties. Reference to Johannes' work.}
+{\emph{Need some discussion of filters and their step-wise properties
+  in relation to their infinitary properties. Reference to Johannes'
+  work.}
+
+\newsavebox{\contig}
+\begin{lrbox}{\contig}
+\begin{lstlisting}[style=myML]
+  Waypoint =
+    {Latitude  : f64
+     Longitude : f64
+     Altitude  : f32
+     Check     : Assert
+      (~90.0 <= Latitude and Latitude <= 90.0 and
+       ~180.0 <= Longitude and Longitude <= 180.0 and
+       1000.0 <= Altitude and Altitude <= 15000.0)}
+
+  AutomationResponse =
+    {TaskID : i64
+     Length : u8
+     Waypoints : Waypoint [3]}
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\contig}}
+    \end{tabular}
+  \end{center}
+  \caption{Contiguity type specification for filter.}
+  \label{fig:filter-contig}
+\end{figure}
 
@@
 
-The synthesized CakeML code for the filter is shown in \figref{fig:filter-cakeml}.
-The code is called at dispatch by the scheduler.
-The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
-The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
-The auto-generated AGREE specification raises an alert output when the relation is violated.
+The contiguity type specification for the filter is shown in
+\figref{fig:filter-contig}. The synthesized CakeML code for the filter is shown in
+\figref{fig:filter-cakeml}. The code is called at dispatch by the
+scheduler. The \texttt{API.callFFI} is the link to the communication
+fabric to capture input and provide output. The body of the function
+restates the filter contract to make the appropriate assignments in a
+way that matches the truth value of the predicate in the filter
+guarantee.  The auto-generated AGREE specification raises an alert
+output when the relation is violated.
 
@@
 
-Monitors, since they are intended to track and analyze the externally visible behavior of system components through time, require more computational features than filters.
-In particular, our conception of a monitor is a predicate on its input (and output) streams, being able to access the value of a stream at any earlier point in time,  if necessary.
-Thus monitors commonly use state to keep track of earlier values, unlike filters which are stateless.
-This reasoning leads us to specify the  computation for a monitor component  by a step function of the following form:
+Monitors, since they are intended to track and analyze the externally
+visible behavior of system components through time, require more
+computational features than filters. In particular, our conception of
+a monitor is a predicate on its input (and output) streams, being able
+to access the value of a stream at any earlier point in time, if
+necessary. Thus monitors commonly use state to keep track of earlier
+values, unlike filters which are stateless. This reasoning leads us to
+specify the computation for a monitor component by a step function of
+the following form:
 \[
@@
 
-The scheduler \emph{activates} components in some order.It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
- Activating a monitor component takes the form of the following pseudo-code: 
+The scheduler \emph{activates} components in some order.It is an
+obligation on the system that the scheduler follows some sensible
+partial order of component activation and allows each component
+sufficient time for its computation.  Activating a monitor component
+takes the form of the following pseudo-code:
 \[
@@
 \paragraph{Initialization}
-A monitor may need to accumulate a certain minimum number of observations before being able to make a meaningful assessment of behavior.
-Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization} phase.
-In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases.
-For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at times $t$ and $t+1$ is such that the distance between the two locations is unusually large.
-Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made.
-The specification has to explicitly say what the correct first output is.
+A monitor may need to accumulate a certain minimum number of
+observations before being able to make a meaningful assessment of
+behavior. Until that threshold is attained, the monitor is essentially
+in a kind of \emph{initialization} phase. In order for correct code to
+be generated, monitor specifications need to spell out the values of
+output ports when in an initialization phase. For example, suppose a
+monitor does some kind of differential assessment of inputs at
+adjacent timeslices, alerting when (say) the measured location of a
+UAV at times $t$ and $t+1$ is such that the distance between the two
+locations is unusually large. Such a monitor needs two measurements
+before making its first judgement, but at the time of its first
+output, only one measurement will have been made. The specification
+for must explicitly state what the correct first output is.

### 2021-05-14 paper/agree.tex

@@
 Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings.
-Let $\Theta_I$ be the universe of environments that only bind the input for the specification.
+Let $I_s(\theta\ldots)$ be a function that maps an environment stream to a new environment stream that only retains the input bindings of each environment in the stream.
 This paper claims the following.
@@
 \[
-  \forall \theta \in \Theta_I^*\ (\konst{eval}\ \konst{synth}(s)\ \theta\ldots) \in \mathcal{L}(s)) 
+  \forall \theta\ldots \in \mathcal{L}(s)\ (\konst{eval}\ \konst{synth}(s)\ I_s(\theta\ldots) = \theta\ldots) \\
 \]
 %
-As noted previously, the semantics of the AGREE specification language are understood in terms of streams of data.

### 2021-05-14 paper/case-study.tex

In this section, we apply the BriefCASE tool to the development of a UAV surveillance system.  The system includes  UAV receives commands from a ground station to conduct
surveillance along a geographical feature such as a river. The on-board mission computer then generates a flight plan consisting ofa series of waypoints that the UAV must traverse to complete it smission. The UAV is also given a set of keep-in and keep-out zones that may constrain its flight path.

The initial software architecture

### 2021-05-14 paper/example.tex
Wordsmithing Example section

@@
 
-\figref{fig:example} is an AADL model of an implementation of a software system (SW) for route planning and automated control for a UAV.
-It is loosely based on the system in the case study introduced later.
+\figref{fig:example} is an AADL architectural model of a software system (SW) for route planning and automated control of a UAV.
+It is loosely based on the system in the case study introduced in Section~\ref{sec:case-study}.
 The source for the entire model is found at \cite{repo}.
-The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task.
+The system receives an automation request that is forwarded to an untrusted third-party route planner (AI).
+The route planner decides the flight path of the UAV based on its current position and the requested task.
 The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
-It is an \emph{as is} legacy component.
+The waypoint manager is an \emph{as is} legacy component.
 
@@
 The contracts constrain input and state properties of output for component models.
-AGREE model checks this assume-guarantee system to hierarchically prove the composite system obeys all contract obligations under all possible finite input streams.
+AGREE performs model checking on this assume-guarantee system to hierarchically prove that the composite system obeys all contract obligations under all possible finite input streams.
 
 The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs.
-A resulting cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
-As such, the designer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.
+A cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
+In response, the developer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output.
+In other words, the AI output is unconstrained in the AGREE specification, allowing it to take on any value and allowing the AI component to send that value at any time.
 
 The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities.
-For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner is now untrusted.
-The systems designer is also responsible for defining the meaning of well-formed, and in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude with an additional check on IDs in the message for integrity.
+For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner output is unconstrained.
+The systems designer is also responsible for defining the meaning of ``well-formed''; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.
 
-Two other requirements are added to the SW AGREE specification by the designer related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream); and the second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request).
-Here the goal of the requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.
+The developer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.
+The first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream).
+The second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request).
+The goal of these two requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.
 
 AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber requirements, by way of model checking.
-The results of the model checker are in \figref{fig:example}(b).
-The red exclamation points are properties that do not hold.
-Each comes with a counter-example.
-The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager.
-The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.
-A clear malicious behavior from the untrusted component.
+The output from the model checker is shown in \figref{fig:example}(b).
+The red exclamation points designate properties that do not hold.
+Each of these failures comes with a counterexample.
+The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about the wellformedness of the input to the legacy waypoint manager.
+The counterexample for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request, a  clear malicious behavior from the untrusted component.
 
@@
 The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
-A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating.
+A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input violates the filter invariant.
 The auto-generated AGREE specification states that only well-formed inputs are passed to the output.
-The system designer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
+The system developer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
 
-A monitor captures a relation over time on input data and is able to reason about temporal properties of that input.
-It raises an alert if ever the specified temporal properties are violated.
-The AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and, also,  that response must come with the request or in the next step after the request.
+A monitor captures a relation on input data over time and is thus able to reason about temporal properties of that input.
+A monitor raises an alert if the specified temporal properties are ever violated.
+The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further,   that response must come with the request or in the next step after the request.
 As with the filter, the system designer provides the policy, and that policy is based on the existing AGREE specification in the SW system.
@@
 The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}(b).
-Here AGREE gives a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber vulnerability from the untrusted AI route planner.
+Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber vulnerabilities from the untrusted AI route planner.
 
@@
 Preserving the input/output relationship of streams between the two models lifts the contract verification results to the deployed system.
-If the contract model verifies, then the meaning of those results hold for the deployed system if the other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.
+If the contract model verification succeeds, then the meaning of those results hold for the deployed system, given that all other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.

### 2021-05-14 paper/synthesis.tex
Worsmithing the Synthesis section

--- 2021-05-13 paper/synthesis.tex
@@
-Synthesis maps from model and specifications to code. The synthesis algorithm traverses the system architecture looking
+Synthesis maps from model and specifications to code. Our synthesis algorithm traverses the system architecture looking
 for occurrences of filter and monitor specifications. For each such 
-occurrence it generates a CakeML program. In the following, we examine both filter and monitor synthesis. The latter is typically much more involved, and we will devote more attention to it.
+occurrence, the synthesis algorithm generates a CakeML program. In the
+following, we examine both filter and monitor synthesis. The latter is
+typically much more involved, and we will thus devote more attention to it.
 
@@
 
-A filter is intended to be very simple; it is expected to have one
+A filter is intended to be very simple,  It is expected to have one
 input port and one output; messages on the input that the filter
 policy admits pass unchanged to the output port; all others are not
-passed on. We have explored in our work two kinds of filter. In the
+passed on. We have explored two kinds of filter in our work to date. In the
 first, a relatively shallow scan of the input buffer can enforce the
-policy. For example, we have used the expressive power of Contiguity
+desired policy. For example, we have used the expressive power of Contiguity
 Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints
-on GPS coordinates in UxAS messages. On the other hand, the second
-kind of filter will parse the input buffer into a data structure
-specified in AGREE and apply an user-defined \emph{well-formedness}
+on GPS coordinates in UxAS messages.  The second
+kind of filter parses the input buffer into a data structure
+specified in AGREE and applies a user-defined \emph{well-formedness}
 property, also specified in AGREE, to the data. This allows
-arbitrarily complex well-formedness checking.
+arbitrarily complex wellformedness checking.
 
-The decision of a filter is made and performed within one thread
+A filter performs its processing and makes its decision all within one thread
 invocation. Thus, in its given time slice, the filter does the following:
@@
 
-The role of partiality should be emphasized: steps 2 and 3 can fail;
+The role of partiality should be emphasized.  Steps 2 and 3 can fail;
 the data might not be parseable or the wellformedness computation
-could be badly written and fail at run time. In such cases, the filter
-should recover and yield control without passing the input onwards. In
+could fail at run time. In such cases, the filter
+should recover and yield control without passing the input to the output. In
 these cases, the filter is behaving as it should, but there are also
@@
 another example would be if the timeslice provided by the scheduler is
-too small for the wellformedness computation to finish.
+too short for the wellformedness computation to complete.
 
@@
 
-Monitors, since they are intended to track and analyze the externally visible behavior of system components through time, require more computational features than filters.
-In particular, our conception of a monitor is a predicate on its input (and output) streams, being able to access the value of a stream at any earlier point in time,  if necessary.
+Monitors require more computational features than filters, as they are
+intended to track and analyze the externally visible behavior of
+system components through time.  Our conception of
+a monitor is a predicate on its input (and output) streams, where the
+monitor is able to access the value of a stream at any earlier point in time,  if necessary.
 Thus monitors commonly use state to keep track of earlier values, unlike filters which are stateless.
@@
 
-The scheduler \emph{activates} components in some order.It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+The scheduler \emph{activates} components in some order.
+It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
  Activating a monitor component takes the form of the following pseudo-code: 
@@
 \paragraph{Initialization}
-A monitor may need to accumulate a certain minimum number of observations before being able to make a meaningful assessment of behavior.
-Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization} phase.
-In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases.
-For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at times $t$ and $t+1$ is such that the distance between the two locations is unusually large.
-Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made.
-The specification has to explicitly say what the correct first output is.
+A monitor may need to accumulate a certain minimum number of
+observations before being able to make a meaningful assessment of
+behavior. Until that threshold is attained, the monitor is essentially
+in a kind of \emph{initialization} phase. In order for correct code to
+be generated, monitor specifications need to spell out the values of
+output ports when in such initialization phases. For example, suppose
+a monitor does some sort of differential assessment of inputs at
+adjacent timeslices, alerting when (say) the measured location of a
+UAV at times $t$ and $t+1$ is such that the distance between the two
+locations is unusually large. Such a monitor needs two measurements
+before making its first judgement, but at the time of its first
+output, only one measurement will have been made. In this case, the
+specification has to explicitly say what the correct first output is.

### 2021-05-14 paper/synthesis.tex
First pass of monitor generation

--- 2021-05-14 paper/synthesis.tex
@@
 
-A filter is intended to be very simple; it is expected to have one
-input port and one output; messages on the input that the filter
-policy admits pass unchanged to the output port; all others are
-dropped (not passed on). We have explored in our work two kinds of
-filter. In the first, a relatively shallow scan of the input buffer
-can enforce the policy. For example, we have used the expressive power
-of Contiguity Types \cite{contiguity-types} to enforce
-\emph{lightweight} bounds constraints on GPS coordinates in UxAS
-messages. On the other hand, the second kind of filter will parse the
-input buffer into a data structure specified in AGREE and apply an
+A filter is intended to be simple, although it may make deep semantic
+checks. A filter has one input port and one output; messages on the
+input that the filter policy admits pass unchanged to the output port;
+all others are dropped (not passed on). We have investigated two kinds
+of filter. In the first, a relatively shallow scan of the input
+suffices to enforce the policy. For example, we have used the
+expressive power of Contiguity Types \cite{contiguity-types} to
+enforce \emph{lightweight} bounds constraints on GPS coordinates in
+UxAS messages. On the other hand, a filter may need to parse the input
+buffer into a data structure specified in AGREE and apply a
 user-defined \emph{well-formedness} property, also specified in AGREE,
-to the data. This allows arbitrarily complex well-formedness checking.
+to the data. Arbitrarily complex well-formedness checks can be made in
+this way. \figref{fig:filter-spec} shows a combination where the
+checking specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+}
+depends on an underlying check specified by the contiguity type
+checking bounds on waypoints.
 
-The decision of a filter is made and performed within one thread
-invocation. Thus, in its given time slice, the filter does the following:
+The verdict of a filter is made and performed within one thread
+invocation. Thus, in its given time slice, the filter must complete the
+following steps:
 
@@
 
-\item the wellformedness predicate is evaluated,
+\item the wellformedness predicate is evaluated on the input,
 
@@
 
-\item the filter yields control
+\item the filter yields control.
 \end{enumerate}
@@
 
-The role of partiality should be emphasized: steps 2 and 3 can fail;
-the data might not be parseable or the wellformedness computation
-could be badly written and fail at run time. In such cases, the filter
-should recover and yield control without passing the input onwards. In
-these cases, the filter is behaving as it should, but there are also
-cases in which a correctly specified filter would not meet its
-specification at runtime. This situation arises when the
-filter \emph{ought} to accept a message, but lack of resources result
-in the filter failing to do so. Examples of this would be, for
-example, if the parse of a message needed more space than allocated;
-another example would be if the timeslice provided by the scheduler is
-too small for the wellformedness computation to finish.
+Partiality is an important consideration: steps 2 and 3 above can
+fail; the data might not be parseable or the wellformedness
+computation could be badly written and fail at runtime. In such cases,
+the filter should recover and yield control without passing the input
+onwards. In these cases, the filter is behaving as it should, but we
+must also guard against situations in which a \emph{correctly
+  specified} filter fails at runtime. This kind of defect arises when
+the filter \emph{ought} to accept a message, but lack of resources
+results in the filter failing to do so. For example, the parse of a
+message might need more space than has been allocated; another example
+could be if the timeslice provided by the scheduler is too short for
+the wellformedness computation to finish. Thus resource bounds need to
+be included in the correctness argument.
 
@@
 
-{\emph{Need some discussion of filters and their step-wise properties
-  in relation to their infinitary properties. Reference to Johannes'
-  work.}
+%% \emph{Need some discussion of filters and their step-wise properties
+%%   in relation to their infinitary properties. Reference to Johannes'
+%%   work.}
 
@@
      Check     : Assert
-      (~90.0 <= Latitude and Latitude <= 90.0 and
+      (~90.0 <= Latitude and Latitude <= 90.0 andp
        ~180.0 <= Longitude and Longitude <= 180.0 and
@@
      Waypoints : Waypoint [3]}
+
+ fun WELL_FORMED_AUTOMATION_RESPONSE(aresp) =
+   (forall wpt in aresp.Waypoints, WELL_FORMED_WAYPOINT(wpt))
+   and ... ;
 \end{lstlisting}
@@
   \end{center}
-  \caption{Contiguity type specification for filter.}
-  \label{fig:filter-contig}
+  \caption{Filter specification.}
+  \label{fig:filter-spec}
 \end{figure}
@@
 
-The contiguity type specification for the filter is shown in
-\figref{fig:filter-contig}. The synthesized CakeML code for the filter is shown in
-\figref{fig:filter-cakeml}. The code is called at dispatch by the
-scheduler. The \texttt{API.callFFI} is the link to the communication
-fabric to capture input and provide output. The body of the function
-restates the filter contract to make the appropriate assignments in a
-way that matches the truth value of the predicate in the filter
-guarantee.  The auto-generated AGREE specification raises an alert
-output when the relation is violated.
+The contiguity type specification and well-formedness predicate for
+the filter are shown in \figref{fig:filter-spec} and the synthesized
+CakeML code is in \figref{fig:filter-cakeml}. The code is called at
+dispatch by the scheduler. The \texttt{API.callFFI} is the link to the
+communication fabric to capture input and provide output. The body of
+the function restates the filter contract to make the appropriate
+assignments in a way that matches the truth value of the predicate in
+the filter guarantee.  The auto-generated AGREE specification raises
+an alert output when the relation is violated.
 
@@
 
-Monitors, since they are intended to track and analyze the externally
-visible behavior of system components through time, require more
-computational features than filters. In particular, our conception of
-a monitor is a predicate on its input (and output) streams, being able
-to access the value of a stream at any earlier point in time, if
-necessary. Thus monitors commonly use state to keep track of earlier
-values, unlike filters which are stateless. This reasoning leads us to
-specify the computation for a monitor component by a step function of
-the following form:
+Monitors are intended to track and analyze the externally visible
+behavior of system components through time. Therefore, they require
+more extensive computational ability than filters. In particular, our
+basic notion of a monitor is that it embodies a predicate over its
+input and output streams, able to access the value of a stream at any
+earlier point in time, if necessary. Monitors commonly use state to
+keep track of earlier values, unlike filters which, for us, are
+typically stateless components. (However, there is nothing in our
+approach that forbids stateful filters: they can be realized by
+monitors.) A monitor specification is mapped by code generation to a
+state transformation function of the following abstract type:
 \[
@@
 \]
-Thus, in its given timeslice, a monitor evaluates the \konst{stepFn} on its inputs and the current values of the state variables. In detail, it takes the following steps:
 
-\begin{enumerate}
-
-\item each available input is parsed into data of the type specified
-by the port type;
-
-\item the stateful variables in $\mathit{stateVars}$  are evaluated in dependency order.
-
-\item values of outputs are computed
-
-\item outputs are written and the new state is written
-
-\item control is yielded
-\end{enumerate}
-
-Our earlier remarks on partiality apply here too, of course.
-
-The scheduler \emph{activates} components in some order.It is an
-obligation on the system that the scheduler follows some sensible
+The system scheduler \emph{activates} components in some order. It is
+an obligation on the system that the scheduler follows some sensible
 partial order of component activation and allows each component
 sufficient time for its computation.  Activating a monitor component
-takes the form of the following pseudo-code:
+takes the form of the following pseudo-code, in which the monitor
+evaluates the \konst{stepFn} on its current inputs and the current
+values of the state variables, returning the new state and the output
+values.
 \[
@@
 
-\paragraph{Initialization}
+\subsubsection{Initialization}
+
 A monitor may need to accumulate a certain minimum number of
@@
 output, only one measurement will have been made. The specification
-for must explicitly state what the correct first output is.
+must then explicitly state what the correct first output is.
+
+\subsubsection{Step function}
+
+The \konst{stepFn} works as follows:
+
+\begin{enumerate}
+
+\item Each input is parsed into data of the type specified by the port
+  type.
+
+\item New values for the state variables are computed, in dependency
+  order. The discussion above on initialization now comes into
+  play. Suppose the variable declarations have the following form:
+\[
+\begin{array}{l}
+  v_1 = i_1 \longrightarrow e_1 \\
+  \cdots \\
+  v_n = i_n \longrightarrow e_n \\
+\end{array}
+\]
+In the generated code, for the first invocation of \konst{stepFn} only,
+the initializations are executed in order:
+\[
+\begin{array}{l}
+  v_1 = i_1; \\
+  \cdots \\
+  v_n = i_n; \\
+\end{array}
+\]
+In all subsequent steps, the \emph{non-initialization} assignments are performed:
+\[
+\begin{array}{l}
+  v_1 = e_1; \\
+  \cdots \\
+  v_n = e_n; \\
+\end{array}
+\]
+
+\item Values of the outputs are computed;
+
+\item Outputs are written and the new state is written;
+
+\item Control is yielded.
+\end{enumerate}
+
+The \konst{stepFn} for the monitor of the example is displayed in \figref{fig:monitor-cakeml}.
+
+\newsavebox{\monFn}
+\begin{lrbox}{\monFn}
+\begin{lstlisting}[style=myML]
+stepFn (Request,Response)
+       (req,rsp,current,previous,policy,alert) =
+let val stateVars' =
+     if !initStep then
+        let val req = event(Request)
+            val rsp = event(Response)
+            val current = (req = rsp)
+            val previous = req and not(rsp)
+            val policy = current or previous
+            alert = not policy
+        in (req,rsp,current,previous,policy,alert)
+        end
+     else
+        let val req = event(Request)
+            val rsp = event(Response)
+            val current = (req = rsp)
+            val previous = pre(req and not rsp) and (not req and rsp)
+            val policy = current or previous
+            alert = (is_latched and pre(alert)) or not(policy)
+        in (req,rsp,current,previous,policy,alert)
+        end
+    val (_,rsp',_,_,_,alert') = stateVars'
+    val Alert = if alert' then Some () else None
+    val Output =
+       if alert' then None else
+       if rsp'   then Some Response
+       else None
+in
+   (stateVars', (Alert,Output))
+end
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\monFn}}
+    \end{tabular}
+  \end{center}
+  \caption{Synthesized CakeML for the filter.}
+  \label{fig:monitor-cakeml}
+\end{figure}
+
+\subsection{Component Behavior}
+
+Possibly some material to add here ...

### 2021-05-14 paper/agree.tex

@@
 The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
-The AGREE specifications generated by the transforms for the high-assurance components are explained.
+The AGREE specifications generated by the transforms for the high-assurance components are then explained.
 The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component.
@@
 
-The AGREE specification for the SW component in the example with the added cyber requirements is given in \figref{fig:sw}.
-Without going into too much detail, the language is a first order predicate calculus.
+The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
+The AGREE specification language is a first-order predicate calculus.
 The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions.
@@
 The AGREE model checker attempts to prove several properties of the top-level model being verified.
-The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component and strong enough to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output).
+The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input as well as the guarantees on the system output).
 These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
@@
 
-The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding.
+The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding.
 The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
@@
 The value of \texttt{previous} in the current time step relies on values from the previous time step.
-The \texttt{->} operator is for initialization as the previous time step is undefined in the first step of the system.
+The \texttt{->} operator designates initialization, as the previous time step is undefined in the first step of the system.
 The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
 The right operand is the value of \texttt{previous} after the initial step.
-Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step.
-Intuitively, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
+Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
 
@@
 
-As noted previously, the original system fails to guarantee the cyber requirement.
+As noted previously, the original system fails to guarantee the cyber requirements.
 BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor.
@@
 The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}.
-The policy of the high assurance component can be stated directly in the wizard, or leave it  blank.
+The policy of the high-assurance component can be stated directly in the wizard, or it can be left  blank.
 In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}.
 Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.
-Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run anytime throughout the design phase to determine whether those artifacts are valid.
+Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.
 %Additionally, the filter can be attached to any requirement the vulnerability analysis.
@@
 The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a).
-The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input.
+The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
 %A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. 
 The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).
-The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised and is one of the several options available in the dialogue.
-The definition for \texttt{policy} is taken by the system designer from the contract in \figref{fig:sw}.
-As before the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
+The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
+This behavior is one of the several options available in the dialogue.
+The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
+As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
 
 At this point it is important to state the meaning of a high-assurance component more formally.
-Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component such as what might be sent over a wire by the communication fabric.
+Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component, such as what might be sent over a wire by the communication fabric.
 An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL.
-As such, it binds \emph{L-values} to strings where an L-value is a concept from compilers and represents anything that can appear on the left-hand side of an assignment.
-In the context of this presentation, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc.
+As such, it binds \emph{L-values} to strings where an L-value (a concept from compilers) represents anything that can appear on the left-hand side of an assignment.
+In the context of this paper, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc.
 The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.
@@
 The meaning is the prefix closed set of finite environment streams on which the specification is invariant.
-Each environment in the stream is total meaning that it binds all L-values in the scope if the specification.
+Each environment in the stream is total, meaning that it binds all L-values in the scope if the specification.
 
-This paper claims that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely.
+We claim that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely.
 Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings.
 Let $I_s(\theta\ldots)$ be a function that maps an environment stream to a new environment stream that only retains the input bindings of each environment in the stream.
-This paper claims the following.
+We then claim the following:
 %

### 2021-05-14 paper/agree.tex
Added Meaning subsection and posed a few questions therein

@@
-The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
-The AGREE specifications generated by the transforms for the high-assurance components are then explained.
-The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component.
-This meaning is what must be preserved by the synthesis.
+The goal of this section is to illustrate in more detail the process a
+system designer follows to add cyber requirements to the AGREE
+specifications and then automatically transform the system to insert
+the high-assurance components. The AGREE specifications generated by
+the transforms for the high-assurance components are then
+explained. The section ends with a concise formal statement of the
+meaning of an AGREE specification for a high-assurance component. This
+meaning is what must be preserved by the synthesis.
 
@@
 
-At this point it is important to state the meaning of a high-assurance component more formally.
-Assume that the inputs to an AGREE contract specification are raw strings (e.g., contiguous sequences of bytes) for each named port of the component, such as what might be sent over a wire by the communication fabric.
-An \emph{environment}, $\theta: \lval \mapsto \konst{string}$, divides the raw string into its constituent parts as given by its data definition in AADL.
-As such, it binds \emph{L-values} to strings where an L-value (a concept from compilers) represents anything that can appear on the left-hand side of an assignment.
-In the context of this paper, an $\lval$ is any reference to a named port, field in a record, entry in an array, local value defined by an eq-statement etc.
-The environment maps any L-value in the scope of the AGREE contract specification to its associated raw representation as a subset of bytes in a contiguous string.
+\subsection{Meaning of AGREE specifications}
 
-Let $s$ be the AGREE contract specification for a high-assurance component.
-The function, $\konst{eval}\ s\ \theta\dots$, evaluates $s$ on the given stream of total environments, $\theta\dots$, and returns \konst{true} if the specification is invariant along the entire stream and \konst{false} otherwise.
-The meaning of invariance for the specification merits some discussion.
-A guarantee has its intuitive meaning: the truth value of the associated expression at every step of the stream must be preserved.
-An eq-statement, however, is treated as a guarantee rather than a computation that adds a new binding to the environment.
-It asserts an equivalence between the named value and its associated expression.
-That equivalence must be preserved at every step of the stream.
+At this point it is important to state the meaning of a high-assurance
+component more formally. Assume that the inputs to an AGREE contract
+specification are raw strings (e.g., contiguous sequences of bytes)
+for each named port of the component, such as what might be sent over
+a wire by the communication fabric. An \emph{environment},
+$\theta: \lval \mapsto \konst{string}$, divides the raw string into
+its constituent parts as given by its data definition in AADL. As
+such, it binds \emph{L-values} to strings where an L-value (a concept
+from compilers) represents anything that can appear on the left-hand
+side of an assignment. In the context of this paper, an $\lval$ is any
+reference to a named port, field in a record, entry in an array, local
+value defined by an eq-statement etc. The environment $\theta$ maps
+any L-value in the scope of the AGREE contract specification to its
+associated raw representation as a subset of bytes in a contiguous
+string.\footnote{Might want to discuss why we want to stay down at the
+representation level.}
 
-The set $\Theta_s$ is the universe of environment bindings for the given specification $s$. 
-The meaning of an AGREE contract specification is now defined as follows.
+[\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
+generate a contig type for the inputs plus state vars.} ]
+
+[Question: what is a \emph{total} environment? One that supplies at
+least the bindings in $\theta$ needed to evaluate the spec?]
+
+Let $s$ be the AGREE contract specification for a high-assurance
+component. The function $\konst{eval}\; s\; \theta\dots$ evaluates $s$
+on the given stream of total environments, $\theta\dots$, and
+returns \konst{true} if the specification is invariant along the
+entire stream and \konst{false} otherwise. The meaning of invariance
+for the specification merits some discussion. A guarantee has its
+intuitive meaning: the truth value of the associated expression at
+every step of the stream must be preserved. An eq-statement, however,
+is treated as a guarantee rather than a computation that adds a new
+binding to the environment. It asserts an equivalence between the
+named value and its associated expression. That equivalence must be
+preserved at every step of the stream.
+
+The set $\Theta_s$ is the universe of environment bindings for the
+given specification $s$.  The meaning of an AGREE contract
+specification is now defined as follows.
 %
 \[
-  \mathcal{L}(s) = \{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots\}  
+  \mathcal{L}(s) = \set{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots}
 \]
 %
-The meaning is the prefix closed set of finite environment streams on which the specification is invariant.
-Each environment in the stream is total, meaning that it binds all L-values in the scope if the specification.
+The meaning is the prefix closed set of finite environment streams on
+which the specification is invariant. Each environment in the stream
+is total, meaning that it binds all L-values in the scope if the
+specification.
 
-We claim that a synthesized high-assurance component generates $\mathcal{L}(s)$ precisely.
-Let $\konst{synth}(s)$ be the CakeML synthesized component, and let $\theta^\prime = \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that evaluates the synthesized component on the input stream to produce a corresponding output stream that is total in its bindings.
-Let $I_s(\theta\ldots)$ be a function that maps an environment stream to a new environment stream that only retains the input bindings of each environment in the stream.
-We then claim the following:
+We claim that a synthesized high-assurance component generates
+$\mathcal{L}(s)$ precisely. Let $\konst{synth}(s)$ be the CakeML
+synthesized component, and let $\theta^\prime
+= \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that
+evaluates the synthesized component on the input stream to produce a
+corresponding output stream that is total in its bindings. Let
+$I_s(\theta\ldots)$ be a function that maps an environment stream to a
+new environment stream that only retains the input bindings of each
+environment in the stream. We then claim the following:
 %

### 2021-05-14 paper/intro.tex
Updates from Overleaf

--- 2021-05-14 paper/intro.tex
@@
-In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
-Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
+In recent years, aerospace stakeholders have become aware that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
+In addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
 Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
@@
 Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.
-Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
+Cyber vulnerabilities are often discovered during penetration testing late in the development process.
+Worse yet, they may be discovered after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
 This is not a sustainable development model.
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is focused on developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
-  views expressed are those of the authors and do not reflect the
-  official policy or position of the Defense Advanced
-  Research Projects Agency (DARPA) or the U.S. Government.}
-The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE}, based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+%\footnote{The
+%  views expressed are those of the authors and do not reflect the
+%  official policy or position of the Defense Advanced
+%  Research Projects Agency (DARPA) or the U.S. Government.}
+The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
 BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.
@@
 Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
-The AGREE analysis \emph{assumes}  properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
-Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
+Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
 If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.
 
-%% DSH  This paragraph seems mostly duplicative of the paragraph above; attempted to merge the two.
-%A novel aspect of the BriefCASE approach is the ability to automatically transform the model to insert high-assurance %components to satisfy requirements added to mitigate cyber vulnerabilities.
-In this work, the behavior of the system is %modeled by contracts on components expressed in the AGREE modeling language.
-These contracts are modified as a result %of the cyber-vulnerability analysis with the new requirements for mitigation.
-The requirements assume properties on the %input and guarantee properties of the output that reflect a hardened system.
-In the absence of any transformation to %implement such requirements, AGREE fails to prove the system meets these new requirements.
+A novel aspect of the approach is the ability to automatically transform the model to insert high-assurance components to satisfy requirements added to mitigate cyber vulnerabilities.
+In this work, the behavior of the system is modeled by contracts on components expressed in the AGREE modeling language.
+These contracts are modified as a result of the cyber-vulnerability analysis with the new requirements for mitigation.
+The requirements assume properties on the input and guarantee properties of the output that reflect a hardened system.
+In the absence of any transformation to implement such requirements, AGREE fails to prove the system meets these new requirements.
 
 The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components.
-These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
+These transformations not only change the architecture of the model by adding in new components, but they generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
 Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.
 
-Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}.
+The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, which is a verified implementation of ML \cite{cakeml}.
 This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
@@
 Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
-Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.
 
@@
 \begin{comment}
-The basic process proceeds as follows:
+Here is the basic process:
 \begin{enumerate}

### 2021-05-14 paper/paper.tex
Added Darren Cofer as co-author
Updates from Overleaf
Author names now all fit on one line

@@
   Synthesizing Verified Components for Cyber Assured Systems Engineering 
-  \thanks{{DISTRIBUTION STATEMENT A.  Approved for public release.}
+  \thanks{{This work was funded by DARPA contract HR00111890001. The
+views, opinions and/or findings expressed are those of the authors
+and should not be interpreted as representing the official views or
+policies of the Department of Defense or the U.S. Government.}
   }
@@
 \IEEEauthorblockN{
+Konrad Slind, \\
 Isaac Amundson and
-Konrad Slind}
+Darren Cofer}
 \IEEEauthorblockA{
@@
 \IEEEauthorblockN{
-Junaid Babar and 
-David Hardin}
+Junaid Babar, David Hardin}
 \IEEEauthorblockA{
@@
 
-\section*{Acknowledgments}
-This work was sponsored in part by the Defense Advanced Research
-Projects Agency (DARPA).
+%\section*{Acknowledgments}
+%This work was sponsored in part by the Defense Advanced Research
+%Projects Agency (DARPA).
 

### 2021-05-14 paper/synthesis.tex
Corrected the last word of the caption of Figure 8 from filter to monitor
Further wordsmithing on the Synthesis section

@@
 The verdict of a filter is made and performed within one thread
-invocation. Thus, in its given time slice, the filter must complete the
-following steps:
+invocation. Thus, in its given time slice, the
+following steps must be completed:
 
@@
 
-\item checks to see if there is any input available; if there is none
-then it yields control; otherwise,
+\item The filter checks to see if there is any input available.  If there is none
+then it yields control; otherwise:
 
-\item the input is read (and parsed if need be),
+\item The input is read (and parsed if need be);
 
-\item the wellformedness predicate is evaluated on the input,
+\item The wellformedness predicate is evaluated on the input;
 
-\item if the predicate returns \konst{true} then the input buffer
- is copied to the output; otherwise no action is taken, and
+\item If the predicate returns \konst{true} then the input buffer
+ is copied to the output, otherwise no action is taken; and
 
-\item the filter yields control.
+\item The filter yields control.
 \end{enumerate}
@@
 output, only one measurement will have been made. The specification
-must then explicitly state what the correct first output is.
+must then explicitly state the correct value for the first output.
 
@@
 \item Each input is parsed into data of the type specified by the port
-  type.
+  type;
 
@@
 
-\item Control is yielded.
+\item The monitor yields control.
 \end{enumerate}
@@
   \end{center}
-  \caption{Synthesized CakeML for the filter.}
+  \caption{Synthesized CakeML for the monitor.}
   \label{fig:monitor-cakeml}

### 2021-05-14 paper/case-study.tex

@@
-In this section, we apply the BriefCASE tool to the development of a UAV surveillance system.  The system includes  UAV receives commands from a ground station to conduct
-surveillance along a geographical feature such as a river.
-The on-board mission computer then generates a flight plan consisting ofa series of waypoints that the UAV must traverse to complete it smission.
-The UAV is also given a set of keep-in and keep-out zones that may constrain its flight path.
+In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature such as a river.
+The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
+The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.
 
-The initial software architecture
+We have modeled the system architecture of the UAV in AADL.
+It includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
+The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
+%
+The initial software architecture model (shown in Figure~\ref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
+The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.
+For this application, UxAS accepts three types of messages.
+The \textit{Operating Region} message defines where the UAV can and cannot fly.
+The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.
+The waypoints typically lie along some geographical feature of interest, such as a river or railway.
+Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.
+Anytime after receiving the Operating Region and Line Search Task messages, a ground station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
+UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.
+Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/sw-initial.png}
+	\caption{Initial software architecture.} 
+	\label{fig:sw-initial} 
+\end{figure}
+
+Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
+We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.
+For the initial version of our design, the verification passes.
+%
+Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.  
+In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
+The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design.
+To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
+%
+For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in Figure~\ref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be included in the component) will not impact other processes.
+
+In total, eight cyber requirements are generated and imported into our model.
+These include one \textit{sandboxing} requirement, four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the ground station software has not been tampered with.
+Requirements are imported into the model as goals in a Resolute assurance case.
+Because we can run Resolute at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
+
+The intent of the \textit{well-formedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
+In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
+By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
+The Filter transform is therefore applied for each well-formedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
+
+The filter behavior for each component is specified in AGREE.
+Not only does this enable formal verification within the modeling environment, but it also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.
+Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.
+When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying the architecture is correct.
+
+The AGREE filter policies for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
+For example, the Automation Response message filter, which drops message containing malformed flight plans, is defined as shown in Figure~\ref{fig:automation-response-filter}.
+Although Latitude, Longitude and Altitude are defined as 64-bit \textit{reals}, the filter ensures only messages containing waypoint values between (-90,90), (-180,180), and (0,15000), respectively.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/automation-response-filter.png}
+	\caption{Automation Response Filter specification.} 
+	\label{fig:automation-response-filter} 
+\end{figure}
+
+In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
+This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
+The Monitor transform is applied for this class of mitigation.
+In general, monitors observe a channel and compare its contents against a reference signal or constant.
+The monitor policy specifies acceptable comparisons, and if violated, the monitor sends out an alert.
+A monitor can choose to \textit{gate} the observed signal, in which case it also acts as a special kind of filter and drops the message if the policy is violated.
+The \textit{monitoring} requirements drive two transforms.
+The first adds a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+The second adds a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+The Geofence Monitor is a gated monitor; it prevents the observed Automation Response message from reaching the Waypoint Manager.
+
+Similar to the filters, the monitor policies are specified in AGREE.
+The Response Monitor specification is shown in Figure~\ref{fig:response-monitor}, and the Geofence Monitor specification is shown in Figure\ref{fig:geofence-monitor}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/response-monitor.png}
+	\caption{Response Monitor specification.} 
+	\label{fig:response-monitor} 
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/geofence-monitor.png}
+	\caption{Geofence Monitor specification.} 
+	\label{fig:geofence-monitor} 
+\end{figure}
+
+% Describe attestation?
+
+After transforming the model to address the cyber requirements, the software architecture now appears as shown in Figure~\ref{fig:hardened-sw}.
+The components in green were added to the model by way of an automated BriefCASE transform and are critical for mitigating cyber attacks.
+We formally verify the model with AGREE to show that all of the component contracts are satisfied, including the new contracts introduced during the model transformations.
+Because it is imperative that these high-assurance component implementations are correct,
+we run the SPLAT tool to produce provably-correct code.  
+The synthesized code is output to a directory in the build file system with the location specified for each component in the model.
+The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/hardened-sw.png}
+	\caption{Cyber-resilient software architecture.} 
+	\label{fig:hardened-sw} 
+\end{figure*}
+
+Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by SPLAT or other means, the system can be built and deployed.
+We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
+HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication and is consistent with the AADL specification computational model.
+
+% Something about seL4?

### 2021-05-14 paper/intro.tex

--- 2021-05-14 paper/intro.tex
@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is focused on developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
-  views expressed are those of the authors and do not reflect the
-  official policy or position of the Defense Advanced
-  Research Projects Agency (DARPA) or the U.S. Government.}
-The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE}, based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+%\footnote{The
+%  views expressed are those of the authors and do not reflect the
+%  official policy or position of the Defense Advanced
+%  Research Projects Agency (DARPA) or the U.S. Government.}
+The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
 BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

### 2021-05-14 paper/case-study.tex

@@
 The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.
+
 For this application, UxAS accepts three types of messages.
@@
 
-% Describe attestation?
+% Describe attestation? virtualization?
 
@@
 % Something about seL4?
+
+The UAV and Ground Station software were deployed on ODROID XU4 boards and communicated with each other over Ethernet.
+The AMASE flight simulator, representing the Flight Contol Computer was run on a Linux machine and connected to the UAV ODROID board via serial connection.
+The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
+In addition, some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior.
+
+We performed an initial set of tests on the initial un-hardened system (Figure~\ref{fig:sw-initial}) to verify the effectiveness of the malicious code.
+We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.

### 2021-05-14 paper/intro.tex
Updates from Overleaf

@@
 
-All the aforementioned tools are publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4, demonstration videos, etc.
-\cite{phase2, camkes} ---ADD LINK TO VIDEO.
+The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
+Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are available at \mbox{loonwerks.com/projects/case}.
 

### 2021-05-14 paper/case-study.tex

--- 2021-05-14 paper/case-study.tex
@@
 In total, eight cyber requirements are generated and imported into our model.
-These include one \textit{sandboxing} requirement, four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the ground station software has not been tampered with.
+These include one \textit{sandboxing} requirement, four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
 Requirements are imported into the model as goals in a Resolute assurance case.
@@
 
-Similar to the filters, the monitor policies are specified in AGREE.
-The Response Monitor specification is shown in Figure~\ref{fig:response-monitor}, and the Geofence Monitor specification is shown in Figure\ref{fig:geofence-monitor}.
+Similar to the filters, the monitor policies are specified in AGREE. For example, 
+%The Response Monitor specification is shown in Figure~\ref{fig:response-monitor}, and 
+the Geofence Monitor specification is shown in Figure\ref{fig:geofence-monitor}.
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1\columnwidth]{figs/response-monitor.png}
-	\caption{Response Monitor specification.} 
-	\label{fig:response-monitor} 
-\end{figure}
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=1\columnwidth]{figs/response-monitor.png}
+% 	\caption{Response Monitor specification.} 
+% 	\label{fig:response-monitor} 
+% \end{figure}
 
@@
 
-% Describe attestation?
+The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on board} the UAV.
+But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands.
+The final cyber-requirement is mitigated by the Attestation transform~\cite{attestation-copland}, which adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
+The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.
+Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in Figure~\ref{fig:attestation}.
+
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=1\columnwidth]{figs/attestation.png}
+% 	\caption{Atestation specification.} 
+% 	\label{fig:attestation} 
+% \end{figure}
 
@@
 
-\begin{figure*}[h]
+\begin{figure}[h]
 	\centering
@@
 	\label{fig:hardened-sw} 
-\end{figure*}
+\end{figure}
 
@@
 HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication and is consistent with the AADL specification computational model.
+%
+HAMR then compiles the software to execute on seL4~\cite{DBLP:journals/cacm/KleinAEHCDEEKNSTW10}, a
+verified capability-based microkernel (accompanied by formal proof of
+spatial isolation properties down to the binary level).
 
-% Something about seL4?
+The UAV and Ground Station software were deployed on the ODROID XU4 hardware and communicated with each other over Ethernet.
+The AMASE flight simulator, representing the Flight Contol Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
+The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
+Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior.
+In addition, we modified the Ground Station to simulate a breach for the attestation test.
+
+We performed an initial set of tests on the un-hardened system (Figure~\ref{fig:sw-initial}) to verify the effectiveness of the malicious code.  The following scenarios were exercised:
+
+\paragraph{Infected Ground Station} In this scenario, one of the UxAS files is modified on the Ground Station.
+When Ground Station files are modified in such an unauthorized manner, messages sent to the UAV will be rejected by the Attestation Manager.
+
+\paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.
+Such malformed messages could exploit vulnerabilities in the onboard software.
+This vulnerability is mitigated by inserting a well-formedness filter.
+The filter prevents the Line Search Task message from reaching UxAS.
+
+\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints will trigger a vulnerability that will crash UxAS.
+When this occurs, UxAS be prevented from generating an Automation Response.
+This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.
+For this scenario, we chose for the monitor to output a status message, which would then be received by the ground station and an appropriate action taken.
+
+\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS will attempt to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.
+However, the UAV will detect that it is being instructed to fly into a keep-out zone and instead return to Home Base.
+
+We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful, as shown in the status messages sent from the high-assurance components in Figure~\ref{fig:mitigation-output}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/mitigation-output.png}
+	\caption{Cyber-resilient system response.} 
+	\label{fig:mitigation-output} 
+\end{figure}

### 2021-05-14 paper/case-study.tex

@@
 
-In total, eight cyber requirements are generated and imported into our model.
-These include one \textit{sandboxing} requirement, four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
+In total, seven cyber requirements are generated and imported into our model.
+These include four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
 Requirements are imported into the model as goals in a Resolute assurance case.
@@
 For example, the Automation Response message filter, which drops message containing malformed flight plans, is defined as shown in Figure~\ref{fig:automation-response-filter}.
-Although Latitude, Longitude and Altitude are defined as 64-bit \textit{reals}, the filter ensures only messages containing waypoint values between (-90,90), (-180,180), and (0,15000), respectively.
+Although Latitude, Longitude and Altitude are defined as 64-bit \textit{reals}, the filter ensures only messages containing waypoint values between [-90,9], [-180,180], and [0,15000], respectively.
 
@@
 %
-HAMR then compiles the software to execute on seL4~\cite{DBLP:journals/cacm/KleinAEHCDEEKNSTW10}, a
-verified capability-based microkernel (accompanied by formal proof of
-spatial isolation properties down to the binary level).
+HAMR then compiles the software to execute on seL4~\cite{sel4-2009}, a verified capability-based microkernel (accompanied by formal proof of spatial isolation properties down to the binary level).
 
@@
 This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.
-For this scenario, we chose for the monitor to output a status message, which would then be received by the ground station and an appropriate action taken.
+For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.
 
@@
 
-We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful, as shown in the status messages sent from the high-assurance components in Figure~\ref{fig:mitigation-output}.
+We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful, as indicated by the status messages sent from the high-assurance components, shown in Figure~\ref{fig:mitigation-output}.
 
@@
 	\centering
-	\includegraphics[width=1\columnwidth]{figs/mitigation-output.png}
+	\includegraphics[width=0.8\columnwidth]{figs/mitigation-output.png}
 	\caption{Cyber-resilient system response.} 

### 2021-05-14 paper/agree.tex
Add Lustre reference

@@
 the high-assurance components. The AGREE specifications generated by
-the transforms for the high-assurance components are then
+the transforms for the high-assurance components are
 explained. The section ends with a concise formal statement of the
@@
 The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
-The AGREE specification language is a first-order predicate calculus.
-The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions.
+The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
+As with Lustre, the semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions.
 The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.

### 2021-05-14 paper/intro.tex
Add citation for CASE'

@@
 The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
-Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are available at \mbox{loonwerks.com/projects/case}.
-
-\begin{comment}
-The basic process proceeds as follows:
-\begin{enumerate}
-  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
-  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
-  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
-  \item Contiguity types include dependent types and the ability to define when a type is malformed
-  \item Define a contract language that makes assumptions on input and guarantees properties on output
-  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
-  \item Show how the contract can be expressed in terms of a step function 
-  \item The step-function is lifted to traces
-  \item A contract defines a language over traces
-  \item A step function recognizes membership in that language (prefix closed)
-  \item Follow the same pattern used in contiguity types
-  \item The CakeML comes for free once the HOL4 theories are shown
-\end{enumerate}
-
-Set the context for this work: define system level theorem that we intend to prove with the added components? Big picture of the overall goal of the project.
-
-Components (complexity (and expressiveness) increases along the way):
-* Filter: takes a stream of data and delivers a new stream of data where the property is enforced--reasons over an infinite stream of date.
-Connects to the system correctness over traces.
-Predicate on a single piece of data, paying no attention to previous history--no temporal awareness, deciting if it passes or not.
-Can be very secure and very efficient.
-* Monitor: Captures a relation over time on the data. Is able to reason about temporal properties. Supports multiple inputs and is able to do arbitrary computation for complex output over the inputs.
-
-The attestation gate is an aside to show how the tools we have are rich enough to create other complex things that blend the two: filter and monitor.
-
-This paper addresses only the verified components. Here is what we verify:
-
-* The resulting assembly code exactly implements the specification--preserves the meaning of the original specification in the actual assembly code.
-
-Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately.
-In the later section we give an example of a system that uses seL4 with a pacer to schedule over different domains to provide memory isolation.
-Scheduling atd the relation to the trace semantics is not part of this paper.
-
-Revised outline:
-
-Describe the system level context with the AGREE analysis that proves system level properties of streams.
-A system is a collection of components and connections.
-We assume a finite set of defined datatypes for the connections.
-A well-formed system is acyclic and provides some suitable set of input streams to consume.
-
-Every component is defined by its inputs, outputs, and a contract.
-The contract language is simplified AGREE: requires, guarantees, eq, and expressions that include \emph{pre}.
-Requires are pre-conditions assumed on the input.
-Guarantees are post-conditions the output must meet.
-A component consumes a stream of input and produces and stream of output.
-A well-formed component is also acyclic.
-
-For a component to by synthesized, its contract must be strong enough to define the meaning of every output (e.g., the value assumed by the output and under what conditions it assumes that value).
-
-The AGREE analysis reasons about the system with its component composition proving that all contracts are respected.
-Failures are counter-examples where some input requirement is violated.
-The analysis provides a system level guarantee.
-
-Component synthesis compiles the contract to imperative code that implements the contract.
-The claim is that the final compiled code preserves the meaning of the contract: it produces the same output stream for any input stream that conforms to its pre-condition.
-The synthesis is accomplished with the CakeML framework that provides the additional guarantee that the post-condition holds over an infinite stream (e.g., that it the component does not terminate and always satisfies the post-condition if the pre-condition holds---Hoare logic extended to statement that diverge).
-
-The semantics is synchronous data-flow. Computation takes zero time as does transport. The system must be well-formed.
-
-Motivate with some example: watchdog example on DNS monitoring for DOS attacks?
-Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.
-\end{comment}
+Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are available at \cite{case}.

### 2021-05-14 paper/conclusion.tex
Add the conclusion

Cyber-physical systems must be tolerant to cyber-attacks in the same way they are tolerant to faults. The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation through the design process to harden systems early in the design process. BriefCASE is the result of that program, and it provides cyber-analysis tools that add cyber requirements to the AGREE specification for the design and architectural transforms with AGREE specifications to satisfy the cyber requirements. BriefCASE is fully integrated into the OSATE AADL modeling environment for ease of use by system engineers.

The filter architectural transformation in BriefCASE prevents malformed data from being propagated downstream to other components while the monitor transformation enforces temporal properties to detect when untrusted components behave maliciously. The components are specified by corresponding auto-generated AGREE contracts that only require the system designer to state the policies to enforce.These can be adapted from the cyber requirements in the system design. The specifications are automatically synthesized to CakeML that can then be compiled to a wide array of backend targets. The synthesis to CakeML is done in a way that preserves the meaning of the specifications.

The BriefCASE tool suite demonstrated in action on a full-scale case study with the UxAS open source UAV AI planning system. The complex system requires several architectural transformations to meet cyber-requirements, and the entire system is synthesized and deployed on the seL4 platform running on ODRIOD using the synthesis tool described here and the BriefCASE HAMR tool to generate the communication fabric. The size and scale of the case study gives some evidence that BriefCASE, with its transformations, scales to the complexity demands often seen in real-world design.

Ongoing work is using BriefCASE on the Collin' Common Avionics Architecture System (CAAS) in the context of a CH47 rotor wing platform. The filter and monitor transformations are being employed to protect against \emph{Automatic Dependent Surveillance-Broadcast} (ADS-B) spoofing by malicious actors in the airspace. That work is expanding synthesis with uninterpreted functions to support non-linear and transcendental functions. Other ongoing work is mechanizing the synthesis proof in HOL4 and lifting the proof results given for finite streams to infinite streams.

### 2021-05-14 paper/conclusion.tex
Add conclusion
Shrink conclusions

@@
-Cyber-physical systems must be tolerant to cyber-attacks in the same way they are tolerant to faults.
-The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation through the design process to harden systems early in the design process.
-BriefCASE is the result of that program, and it provides cyber-analysis tools that add cyber requirements to the AGREE specification for the design and architectural transforms with AGREE specifications to satisfy the cyber requirements.
-BriefCASE is fully integrated into the OSATE AADL modeling environment for ease of use by system engineers.
+The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation.
+The resulting BriefCASE tool suite in OSATE includes analyses to add cyber requirements to systems and architectural transformations on systems to satisfy cyber requirements.
+The filter transformation prevents malformed data from being propagated downstream while the monitor transformation checks temporal properties to detect malicious behavior.
+The components are specified by corresponding auto-generated AGREE contracts that only require policy annotations taken from the cyber requirements.
+The specifications are automatically synthesized to CakeML and that synthesis preserves the meaning of the specifications.
 
-The filter architectural transformation in BriefCASE prevents malformed data from being propagated downstream to other components while the monitor transformation enforces temporal properties to detect when untrusted components behave maliciously.
-The components are specified by corresponding auto-generated AGREE contracts that only require the system designer to state the policies to enforce.These can be adapted from the cyber requirements in the system design.
-The specifications are automatically synthesized to CakeML that can then be compiled to a wide array of backend targets.
-The synthesis to CakeML is done in a way that preserves the meaning of the specifications.
+The BriefCASE analyses and transformations are applied to a full-scale case study on the Air Force Research Laboratory's OpenUxAS services requiring several filters and monitors to meet cyber-requirements.
+A video shows the synthesized system running on the ODROID platform with seL4.
+The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.
 
-The BriefCASE tool suite demonstrated in action on a full-scale case study with the UxAS open source UAV AI planning system.
-The complex system requires several architectural transformations to meet cyber-requirements, and the entire system is synthesized and deployed on the seL4 platform running on ODRIOD using the synthesis tool described here and the BriefCASE HAMR tool to generate the communication fabric.
-The size and scale of the case study gives some evidence that BriefCASE, with its transformations, scales to the complexity demands often seen in real-world design.
-
-Ongoing work is using BriefCASE on the Collin' Common Avionics Architecture System (CAAS) in the context of a CH47 rotor wing platform.
-The filter and monitor transformations are being employed to protect against \emph{Automatic Dependent Surveillance-Broadcast} (ADS-B) spoofing by malicious actors in the airspace.
-That work is expanding synthesis with uninterpreted functions to support non-linear and transcendental functions.
-Other ongoing work is mechanizing the synthesis proof in HOL4 and lifting the proof results given for finite streams to infinite streams.
+Ongoing work is applying BriefCASE in the Collin' Common Avionics Architecture System (CAAS), \cite{caas}, to protect against \emph{Automatic Dependent Surveillance-Broadcast} spoofing.
+That work adds support for uninterpreted functions.
+Other ongoing work is mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

### 2021-05-14 paper/agree.tex
Revise agree semantics

@@
 
-\subsection{Meaning of AGREE specifications}
+\subsection{Brief Semantics Definition}
 
-At this point it is important to state the meaning of a high-assurance
-component more formally. Assume that the inputs to an AGREE contract
-specification are raw strings (e.g., contiguous sequences of bytes)
-for each named port of the component, such as what might be sent over
-a wire by the communication fabric. An \emph{environment},
-$\theta: \lval \mapsto \konst{string}$, divides the raw string into
-its constituent parts as given by its data definition in AADL. As
-such, it binds \emph{L-values} to strings where an L-value (a concept
-from compilers) represents anything that can appear on the left-hand
-side of an assignment. In the context of this paper, an $\lval$ is any
-reference to a named port, field in a record, entry in an array, local
-value defined by an eq-statement etc. The environment $\theta$ maps
-any L-value in the scope of the AGREE contract specification to its
-associated raw representation as a subset of bytes in a contiguous
-string.\footnote{Might want to discuss why we want to stay down at the
-representation level.}
+Here the formal semantics of the AGREE contract specification are briefly
+presented to make clear the meaning of a high-assurance component. 
+These semantics are used in the next section to argue that the synthesis
+preserves the same input to output behavior as the AGREE specification.
 
+Assume that all data is in its raw form which is a contiguous sequence of bytes
+representing exactly what is sent over a wire by the communication fabric, so a
+datum is given by a \konst{string}. This assumption is important to the 
+correctness argument in the next section. An \emph{environment},
+$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to strings where 
+an L-value is anything that can appear on the left-hand side of an assignment
+such as a named port, a field in a record, a entry in an array, a local
+value defined by an eq-statement etc.
+
+Let $s$ be an AGREE contract specification for some high-assurance
+component. The notation $\theta_s$ is used to denote the environment that
+contains a binding for any L-value in the scope of $s$ and nothing else, and
+the notation $\Theta_s$ denotes the universe of all such environments. The
+semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
+are finite sequences of environments, $\pi \in \Theta_s^*$.
+
+\begin{comment}
 [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
@@
 least the bindings in $\theta$ needed to evaluate the spec?]
+\end{comment}
 
-Let $s$ be the AGREE contract specification for a high-assurance
-component. The function $\konst{eval}\; s\; \theta\dots$ evaluates $s$
-on the given stream of total environments, $\theta\dots$, and
-returns \konst{true} if the specification is invariant along the
-entire stream and \konst{false} otherwise. The meaning of invariance
-for the specification merits some discussion. A guarantee has its
-intuitive meaning: the truth value of the associated expression at
-every step of the stream must be preserved. An eq-statement, however,
-is treated as a guarantee rather than a computation that adds a new
-binding to the environment. It asserts an equivalence between the
-named value and its associated expression. That equivalence must be
-preserved at every step of the stream.
+The function $\konst{eval}\; s\; \pi$ is defined such that it evaluates $s$ on
+the stream, $\pi$, and returns \konst{true} if $s$ is invariant along the
+entire stream and \konst{false} otherwise. A guarantee in $s$ is invariant if
+its associated expression is true in the context of every step of the stream, while
+an eq-statement in $s$ is invariant if its binding in the context of every step
+is equivalent to the computed value of its associated expression in that same
+context. All guarantees and eq-statements must be invariant in the stream for
+the function to return \konst{true}.
 
-The set $\Theta_s$ is the universe of environment bindings for the
-given specification $s$.  The meaning of an AGREE contract
-specification is now defined as follows.
+The meaning of an AGREE contract
+specification is now defined as the set of environment streams on which it is invariant.
 %
 \[
-  \mathcal{L}(s) = \set{\theta\ldots \in \Theta_s^* \mid \konst{eval}\ s\ \theta\ldots}
+  \mathcal{L}(s) = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
 \]
 %
-The meaning is the prefix closed set of finite environment streams on
-which the specification is invariant. Each environment in the stream
-is total, meaning that it binds all L-values in the scope if the
-specification.
+\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
+the L-values in the eq-statements
+in a way that is consistent with their associated expressions and the guarantees are all
+true in that same step.
 
-We claim that a synthesized high-assurance component generates
-$\mathcal{L}(s)$ precisely. Let $\konst{synth}(s)$ be the CakeML
-synthesized component, and let $\theta^\prime
-= \konst{eval}\ \konst{synth}(s)\ \theta\ldots$ be a function that
-evaluates the synthesized component on the input stream to produce a
-corresponding output stream that is total in its bindings. Let
-$I_s(\theta\ldots)$ be a function that maps an environment stream to a
-new environment stream that only retains the input bindings of each
-environment in the stream. We then claim the following:
+We claim that a synthesized high-assurance component preserves the input to output
+relationship in the specification $s$ over every stream in $\mathcal{L}(s)$. 
+Let $\pi^\prime
+= \konst{SynthEval}\ s\ \pi$ denate a function that synthesizes $s$ and then
+evaluates that synthesized component on the stream $\pi$ to create a new
+stream $\pi^\prime$ containing added output and other bindings. We
+say that two streams, $\pi_a = \pi_b$ are equivalent in regards to a
+specification $s$ if and only if the two streams are the same length and
+agree on bindings for the input and output for $s$ at every step of the streams.
+We now formally state the correctness claim for synthesis.
 %
 \[
-  \forall \theta\ldots \in \mathcal{L}(s)\ (\konst{eval}\ \konst{synth}(s)\ I_s(\theta\ldots) = \theta\ldots) \\
+  \forall \pi \in \mathcal{L}(s)\ ((\konst{SynthEval}\ s\ I_s(\pi)) =_s \pi)\\
 \]
 %
+where $I_s(\pi)$ returns the corresponding stream that only retains bindings
+for inputs in each step and nothing else. The claim is that the synthesized component
+generates the same output stream as the specification for any stream
+belonging to the specification that is restricted to just input binding at each step.

### 2021-05-14 paper/agree.tex
Fix typos

@@
 Let $\pi^\prime
-= \konst{SynthEval}\ s\ \pi$ denate a function that synthesizes $s$ and then
+= \konst{SynthEval}\ s\ \pi$ denote a function that synthesizes $s$ and then
 evaluates that synthesized component on the stream $\pi$ to create a new
 stream $\pi^\prime$ containing added output and other bindings. We
-say that two streams, $\pi_a = \pi_b$ are equivalent in regards to a
-specification $s$ if and only if the two streams are the same length and
+say that two streams are equivalent in regards to a
+specification, denoted as $\pi =_s \pi^\prime$, if and only if the two streams are the same length and
 agree on bindings for the input and output for $s$ at every step of the streams.
@@
 generates the same output stream as the specification for any stream
-belonging to the specification that is restricted to just input binding at each step.
+belonging to the specification that is restricted to just input bindings at each step.

### 2021-05-14 paper/conclusion.tex

@@
 The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation.
-The resulting BriefCASE tool suite in OSATE includes analyses to add cyber requirements to systems and architectural transformations on systems to satisfy cyber requirements.
+The resulting BriefCASE tool suite includes analyses to add cyber requirements to systems and architectural transformations on systems to satisfy cyber requirements.
 The filter transformation prevents malformed data from being propagated downstream while the monitor transformation checks temporal properties to detect malicious behavior.
@@
 The BriefCASE analyses and transformations are applied to a full-scale case study on the Air Force Research Laboratory's OpenUxAS services requiring several filters and monitors to meet cyber-requirements.
-A video shows the synthesized system running on the ODROID platform with seL4.
+%A video shows the synthesized system running on the ODROID platform with seL4. 
 The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.
 
-Ongoing work is applying BriefCASE in the Collin' Common Avionics Architecture System (CAAS), \cite{caas}, to protect against \emph{Automatic Dependent Surveillance-Broadcast} spoofing.
+Ongoing work is applying BriefCASE in the Collins Common Avionics Architecture System (CAAS)~\cite{caas}, to protect against \emph{Automatic Dependent Surveillance-Broadcast} spoofing.
 That work adds support for uninterpreted functions.

### 2021-05-15 paper/abstract.tex
Merger from overleaf.

@@
-Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either recover or safely shut down as requirements dictate.
+Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either gracefully recover or safely shut down as requirements dictate.
 The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment.

### 2021-05-15 paper/case-study.tex
Merger from overleaf.

@@
-In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature such as a river.
+In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature.
 The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
@@
 %
-The initial software architecture model (shown in Figure~\ref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
+The initial software architecture model (shown in \figref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
 The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.
@@
 Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.
-Anytime after receiving the Operating Region and Line Search Task messages, a ground station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
+Anytime after receiving the Operating Region and Line Search Task messages, a Ground Station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
 UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.
@@
 %
-For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in Figure~\ref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be included in the component) will not impact other processes.
+For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be embedded in the component) will not impact other processes.
 
@@
 The AGREE filter policies for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
-For example, the Automation Response message filter, which drops message containing malformed flight plans, is defined as shown in Figure~\ref{fig:automation-response-filter}.
-Although Latitude, Longitude and Altitude are defined as 64-bit \textit{reals}, the filter ensures only messages containing waypoint values between [-90,9], [-180,180], and [0,15000], respectively.
+For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
+Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit \textit{reals}, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.
 
@@
 Similar to the filters, the monitor policies are specified in AGREE. For example, 
-%The Response Monitor specification is shown in Figure~\ref{fig:response-monitor}, and 
-the Geofence Monitor specification is shown in Figure\ref{fig:geofence-monitor}.
+%The Response Monitor specification is shown in \figref{fig:response-monitor}, and 
+the Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.
 
@@
 
-The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on board} the UAV.
+The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
 But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands.
@@
 The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.
-Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in Figure~\ref{fig:attestation}.
+Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in \figref{fig:attestation}.
 
@@
 
-After transforming the model to address the cyber requirements, the software architecture now appears as shown in Figure~\ref{fig:hardened-sw}.
+After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
 The components in green were added to the model by way of an automated BriefCASE transform and are critical for mitigating cyber attacks.
@@
 
-The UAV and Ground Station software were deployed on the ODROID XU4 hardware and communicated with each other over Ethernet.
-The AMASE flight simulator, representing the Flight Contol Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
+The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicate with each other over Ethernet.
+The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
 The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
 Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior.
-In addition, we modified the Ground Station to simulate a breach for the attestation test.
+In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
 
-We performed an initial set of tests on the un-hardened system (Figure~\ref{fig:sw-initial}) to verify the effectiveness of the malicious code.  The following scenarios were exercised:
+We first performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
+We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
+The following scenarios were exercised (status messages from the high-assurance components in the hardened system are shown in \figref{fig:mitigation-output}):
 
-\paragraph{Infected Ground Station} In this scenario, one of the UxAS files is modified on the Ground Station.
-When Ground Station files are modified in such an unauthorized manner, messages sent to the UAV will be rejected by the Attestation Manager.
+\paragraph{Infected Ground Station} In this scenario, one of the application files is modified on the Ground Station, which sends the UAV on a trajectory that violates mission requirements. 
+On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.
 
 \paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.
-Such malformed messages could exploit vulnerabilities in the onboard software.
-This vulnerability is mitigated by inserting a well-formedness filter.
-The filter prevents the Line Search Task message from reaching UxAS.
+Such malformed messages could exploit vulnerabilities in the on-board software.
+This vulnerability is mitigated by inserting a well-formedness filter.  The filter prevented the Line Search Task message from reaching UxAS.
 
-\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints will trigger a vulnerability that will crash UxAS.
-When this occurs, UxAS be prevented from generating an Automation Response.
+\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, which triggered a vulnerability that crashes UxAS.
+When this occurs, UxAS is prevented from generating an Automation Response.
 This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.
@@
 \paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS will attempt to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.
-However, the UAV will detect that it is being instructed to fly into a keep-out zone and instead return to Home Base.
-
-We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful, as indicated by the status messages sent from the high-assurance components, shown in Figure~\ref{fig:mitigation-output}.
+On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and instead returned the UAV to Home Base.
 

### 2021-05-15 paper/conclusion.tex
Merger from overleaf.

@@
-The DARPA CASE program is creating an MBSE environment for designers to integrate cyber-vulnerability analysis and mitigation.
-The resulting BriefCASE tool suite includes analyses to add cyber requirements to systems and architectural transformations on systems to satisfy cyber requirements.
-The filter transformation prevents malformed data from being propagated downstream while the monitor transformation checks temporal properties to detect malicious behavior.
-The components are specified by corresponding auto-generated AGREE contracts that only require policy annotations taken from the cyber requirements.
-The specifications are automatically synthesized to CakeML and that synthesis preserves the meaning of the specifications.
+The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation.
+The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code.
+Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified in the AGREE language.
+The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.
 
-The BriefCASE analyses and transformations are applied to a full-scale case study on the Air Force Research Laboratory's OpenUxAS services requiring several filters and monitors to meet cyber-requirements.
-%A video shows the synthesized system running on the ODROID platform with seL4. 
+BriefCASE was applied to a full-scale case study using the Air Force Research Laboratory's OpenUxAS software, exercising a range of built-in cyber resiliency mitigations to meet cyber-requirements. 
 The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.
 
-Ongoing work is applying BriefCASE in the Collins Common Avionics Architecture System (CAAS)~\cite{caas}, to protect against \emph{Automatic Dependent Surveillance-Broadcast} spoofing.
-That work adds support for uninterpreted functions.
-Other ongoing work is mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.
+We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47.
+Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

### 2021-05-15 paper/intro.tex
Merger from overleaf.

@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
 %\footnote{The
@@
 %  Research Projects Agency (DARPA) or the U.S. Government.}
-The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
-BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.
+We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
 
@@
 CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
-Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
+Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR-generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
 Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
@@
 The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}.
+Contract specification in the AGREE language is presented in Section~\ref{sec:agree}, followed by a description of the synthesis pathway in Section~\ref{sec:synthesis}.
 A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}.
 Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
-The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.
+The case study system is significantly more complex than the simple example and shows the viability of the modeling approach to a full-scale industrial design.
 
 The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
-Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are available at \cite{case}.
+Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are also available \cite{case}.

### 2021-05-15 paper/related-work.tex
Pass on related work

--- 2021-04-01 paper/related-work.tex
@@
 A monitor for a contract can be removed when it can be statically proved that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+
+Data-flow semantics are well studied \cite{10.1145/41625.41641,97300, 10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}.
+State machine semantics can be added to synchronous data-flow \cite{10.1145/1086228.1086261}.
+The idea is to translate imperative constructs into equivalent synchronous data flow constructs.
+The resulting Lustre can them be compiled to target a platform.
+The goal is to provide a seamless connection between pure data-flow and pure control design.
+
+There is a fully verified compiler that takes Lustre and turns it into a binary executable \cite{10.1145/3140587.3062358}.
+The first compilation pass turns synchronous data-flow in Lustre to assembly.
+The compilation is specified and verified in Coq.
+The key is in combining infinite sequences of data-flow models with incremental manipulation of memories akin to an imperative model.
+CompCert is used on the backend to create the final rendered executable.

### 2021-05-15 paper/conclusion.tex
Worsdmithing in the Conclusion section

@@
-The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation.
+The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation into their development workflow.
 The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code.
-Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified in the AGREE language.
+Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified using the AGREE language.
 The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.
@@
 
-We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47.
+We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the DARPA CASE program.
 Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

### 2021-05-15 paper/related-work.tex

@@
 Contracts are similar to assume-guarantee reasoning but are targeted to programming languages.
-Contracts are often more expressive than assume-guarantee reasoning and can not only be stateful but higher-order \cite{10.1145/583852.581484}.
+Contracts are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}.
 As contracts are often written in the target language, synthesis is not a problem for monitoring but comes with significant overhead \cite{10.1007/978-3-642-28869-2_11}.
-A monitor for a contract can be removed when it can be statically proved that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+A monitor for a contract can be removed when it can be statically proven that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
 
@@
 The idea is to translate imperative constructs into equivalent synchronous data flow constructs.
-The resulting Lustre can them be compiled to target a platform.
+The resulting Lustre can then be compiled.
 The goal is to provide a seamless connection between pure data-flow and pure control design.
 
-There is a fully verified compiler that takes Lustre and turns it into a binary executable \cite{10.1145/3140587.3062358}.
-The first compilation pass turns synchronous data-flow in Lustre to assembly.
-The compilation is specified and verified in Coq.
+There is a fully verified compiler that takes Lustre and turns it into a binary executable that is specified and verified in Coq \cite{10.1145/3140587.3062358}.
 The key is in combining infinite sequences of data-flow models with incremental manipulation of memories akin to an imperative model.

### 2021-05-15 paper/case-study.tex
Wordsmithing in the Case Study section

@@
 We have modeled the system architecture of the UAV in AADL.
-It includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
+The model includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
 The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
@@
 Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.
-When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying the architecture is correct.
+When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying that the architecture is correct.
 
@@
 For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
-Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit \textit{reals}, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.
+Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit floating-point values, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.
 
@@
 We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
-HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication and is consistent with the AADL specification computational model.
+HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication consistent with the AADL computational model.
 %
@@
 We first performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
-We repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
+A third party evaluator repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
 The following scenarios were exercised (status messages from the high-assurance components in the hardened system are shown in \figref{fig:mitigation-output}):
@@
 
-\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS will attempt to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.
+\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS attempts to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.
 On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and instead returned the UAV to Home Base.

### 2021-05-15 paper/case-study.tex
Merge

@@
-In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance along a geographical feature.
+In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance over a geographical region.
 The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.

### 2021-05-15 paper/agree.tex
add connection between stepFn and AGREE semantics; some other minor tweakage

@@
 \subsection{Brief Semantics Definition}
-
+\label{agree-semantics}
 Here the formal semantics of the AGREE contract specification are briefly
@@
 
-\begin{comment}
-[\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
-generate a contig type for the inputs plus state vars.} ]
+%% \begin{comment}
+%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
+%% generate a contig type for the inputs plus state vars.} ]
 
-[Question: what is a \emph{total} environment? One that supplies at
-least the bindings in $\theta$ needed to evaluate the spec?]
-\end{comment}
+%% [Question: what is a \emph{total} environment? One that supplies at
+%% least the bindings in $\theta$ needed to evaluate the spec?]
+%% \end{comment}
 
-The function $\konst{eval}\; s\; \pi$ is defined such that it evaluates $s$ on
-the stream, $\pi$, and returns \konst{true} if $s$ is invariant along the
-entire stream and \konst{false} otherwise. A guarantee in $s$ is invariant if
-its associated expression is true in the context of every step of the stream, while
-an eq-statement in $s$ is invariant if its binding in the context of every step
-is equivalent to the computed value of its associated expression in that same
-context. All guarantees and eq-statements must be invariant in the stream for
-the function to return \konst{true}.
+The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
+$\pi$, and returns \konst{true} if $s$ is invariant along the entire
+stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
+is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
+while an eq-statement in $s$ is invariant if its binding in the
+context of every step is equivalent to the computed value of its
+associated expression in that same context. All guarantees and
+eq-statements must be invariant in the stream for the function to
+return \konst{true}.
 
-The meaning of an AGREE contract
-specification is now defined as the set of environment streams on which it is invariant.
+The meaning of an AGREE contract specification is now defined as the
+set of environment streams on which it is invariant.
 %
 \[
-  \mathcal{L}(s) = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
+  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
 \]
@@
 \noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
-the L-values in the eq-statements
-in a way that is consistent with their associated expressions and the guarantees are all
-true in that same step.
+the L-values in the eq-statements in a way that is consistent with
+their associated expressions and the guarantees are all true in that
+same step.
 
-We claim that a synthesized high-assurance component preserves the input to output
-relationship in the specification $s$ over every stream in $\mathcal{L}(s)$. 
-Let $\pi^\prime
-= \konst{SynthEval}\ s\ \pi$ denote a function that synthesizes $s$ and then
-evaluates that synthesized component on the stream $\pi$ to create a new
-stream $\pi^\prime$ containing added output and other bindings. We
-say that two streams are equivalent in regards to a
-specification, denoted as $\pi =_s \pi^\prime$, if and only if the two streams are the same length and
-agree on bindings for the input and output for $s$ at every step of the streams.
-We now formally state the correctness claim for synthesis.
+We claim that a synthesized high-assurance component preserves the
+input to output relationship in the specification $s$ over every
+stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
+s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
+synthesized component on the stream $\pi$ to create a new stream
+$\pi^\prime$ containing added output and other bindings. We say that
+two streams are equivalent in regards to a specification, denoted as
+$\pi =_s \pi^\prime$, if and only if the two streams are the same
+length and agree on bindings for the input and output for $s$ at every
+step of the streams.  We now formally state the correctness claim for
+synthesis.
 %
 \[
-  \forall \pi \in \mathcal{L}(s)\ ((\konst{SynthEval}\ s\ I_s(\pi)) =_s \pi)\\
+  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
 \]
 %
-where $I_s(\pi)$ returns the corresponding stream that only retains bindings
-for inputs in each step and nothing else. The claim is that the synthesized component
-generates the same output stream as the specification for any stream
-belonging to the specification that is restricted to just input bindings at each step.
+where $I_s(\pi)$ returns the corresponding stream that only retains
+bindings for inputs in each step and nothing else. The claim is that
+the synthesized component generates the same output stream as the
+specification for any stream belonging to the specification that is
+restricted to just input bindings at each step.

### 2021-05-15 paper/conclusion.tex
add connection between stepFn and AGREE semantics; some other minor tweakage

@@
-The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation into their development workflow.
-The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code.
-Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified using the AGREE language.
-The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.
+The DARPA CASE program is creating tools for systems engineers to
+integrate cyber-vulnerability analysis and mitigation into their
+development workflow. The resulting BriefCASE tool suite includes
+analysis tools for generating cyber requirements, cyber resiliency
+tools for addressing the requirements, verification tools for ensuring
+design correctness, and synthesis tools for generating provably
+correct code. Several of the BriefCASE transforms (filter, monitor,
+gate) insert components into the model whose behavior can be formally
+specified using the AGREE language.  The SPLAT tool can then
+automatically generate CakeML implementations for these components,
+along with proofs of correctness for assurance that the implementation
+satisfies the specification.
 
-BriefCASE was applied to a full-scale case study using the Air Force Research Laboratory's OpenUxAS software, exercising a range of built-in cyber resiliency mitigations to meet cyber-requirements. 
-The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.
+BriefCASE was applied to a full-scale case study using the Air Force
+Research Laboratory's OpenUxAS software, exercising a range of
+built-in cyber resiliency mitigations to meet cyber-requirements.  The
+size and scale of the study suggests BriefCASE meets the complexity
+demands of real-world design.
 
-We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the DARPA CASE program.
-Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.
+We are currently in the process of applying BriefCASE to the design of
+an application using the Collins Common Avionics Architecture System
+(CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the
+DARPA CASE program.  Other ongoing work and future work includes
+adding support for uninterpreted functions, mechanizing the
+correctness-of-synthesis proofs in HOL4, and lifting the proof results
+to infinite streams.

### 2021-05-15 paper/synthesis.tex
Added example section reference to Synthesis section
add connection between stepFn and AGREE semantics; some other minor tweakage

@@
 
-The \konst{stepFn} for the monitor of the example is displayed in \figref{fig:monitor-cakeml}.
+The \konst{stepFn} for the monitor of the example described in
+Section~\ref{sec:example} is displayed in \figref{fig:monitor-cakeml}.
 
@@
             val policy = current or previous
-            alert = not policy
+            val alert = not policy
+            val () = (intStep := False)
         in (req,rsp,current,previous,policy,alert)
@@
             val policy = current or previous
-            alert = (is_latched and pre(alert)) or not(policy)
+            val alert = (is_latched and pre(alert)) or not(policy)
         in (req,rsp,current,previous,policy,alert)
@@
 
-Possibly some material to add here ...
+Intutively, for monitor specification $s$, \konst{stepFn} is the
+concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
+\ref{agree-semantics}. Its correctness amounts to showing that, given
+a sequence of inputs, and an initial state meeting the initialization
+constraints, iterating \konst{stepFn} produces a $\pi \in
+\Lang{s}$. Morevover, taking the union over all input sequences and
+initial states produces $\Lang{s}$ itself.

### 2021-05-15 paper/abstract.tex
Minor update to abstract

@@
 Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either gracefully recover or safely shut down as requirements dictate.
-The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment.
+The DARPA Cyber Assured Systems Engineering program is developing tools for design, analysis, and verification that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment.
 This paper describes automated model transformations that introduce high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively.
-A formal specification defines each component, and is used to verify that the component addresses system level cyber requirements.
-Implementations are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level.
+A formal specification defines each high-assurance component, and is used to verify that the component addresses system level cyber requirements.
+Implementations for these high-assurance components are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level.
 The model transformations are integrated into the Open Source AADL Tool Environment (OSATE).

### 2021-05-15 paper/intro.tex
Minor updates to intro

@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
 %\footnote{The
@@
 
-BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.
+BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models for potential cyber vulnerabilities and suggest cyber-security requirements to mitigate them.
 A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system.
 Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
-Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
-The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
+Formal verification that the transformed system model satisfies its cyber requirements is accomplished via the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
+%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output. 
+AGREE is a {\em compositional assume-guarantee} style model checker for AADL models that attempts to prove properties about one layer of the architecture using
+properties allocated to its subcomponents.
 Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
@@
 
-The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components.
+The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect (and alert) unexpected behaviors arising from untrusted components.
 These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
-Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.
+Those specifications are sufficient for model checking to prove that with the incorporation of these high-assurance components, the hardened system meets its cyber-resiliency requirements.
 
@@
 CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
-Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR-generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
+Assuming that the execution schedule of the deployed cyber-hardened system is as intended by the AADL model, and that the HAMR-generated communication fabric delivers messages between components as expected, the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
 Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

### 2021-05-15 paper/intro.tex
Restored Distribution Statement, other DARPA public approval boilerplate

@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.
-%\footnote{The
-%  views expressed are those of the authors and do not reflect the
-%  official policy or position of the Defense Advanced
-%  Research Projects Agency (DARPA) or the U.S. Government.}
+The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).
+The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S.
+Government.}
 We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.

### 2021-05-15 paper/paper.tex
Add clearpage for appendix
Merge overleaf
Restored Distribution Statement, other DARPA public approval boilerplate

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{flushend}
 
@@
   Synthesizing Verified Components for Cyber Assured Systems Engineering 
-  \thanks{{This work was funded by DARPA contract HR00111890001. The
-views, opinions and/or findings expressed are those of the authors
-and should not be interpreted as representing the official views or
-policies of the Department of Defense or the U.S. Government.}
-  }
+  \thanks{DISTRIBUTION STATEMENT A.  Approved for public release.}
 }
@@
 \IEEEauthorblockN{
-Konrad Slind, \\
-Isaac Amundson and
-Darren Cofer}
+Konrad Slind, Isaac Amundson, Darren Cofer}
 \IEEEauthorblockA{
@@
 
-\section{Introduction}
+\section{Introduction} \label{sec:intro}
 \input{intro}
 
-\section{Simple Example}
+\section{Simple Example} \label{sec:example}
 \input{example}
 
-\section{AGREE Contract Specification}
+\section{AGREE Contract Specification} \label{sec:agree}
 \input{agree}
 
-\section{Synthesis}
+\section{Synthesis} \label{sec:synthesis}
 \input{synthesis}
 
-\section{UxAS Case Study}
+\section{UxAS Case Study} \label{sec:case-study}
 \input{case-study}
 
-\section{Related Work}
+\section{Related Work} \label{sec:related-work}
 \input{related-work}
 
-\section{Conclusion}
+\section{Conclusion} \label{sec:conclusion}
 \input{conclusion}
@@
 % references section
+\clearpage
 \bibliographystyle{IEEEtran}

### 2021-05-15 paper/case-study.tex
Full pass done. Spell check. Minor edits. Paper fits.

@@
-In this section, we apply the BriefCASE tool to the development of a UAV surveillance system in which a UAV receives commands from a ground station to conduct surveillance over a geographical region.
-The on-board mission computer then generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
+In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
+%apply the BriefCASE tool towards the development of a UAV surveillance system in which 
+The UAV receives commands from a ground station to conduct surveillance over a geographical region.
+In response, the UAV's on-board mission computer generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
 The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.
@@
 In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
-The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design.
-To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
+The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design,
+and we need to satisfy these requirements by modifying the design.
+%  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
 %
-For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring unverified or malicious code (that could potentially be embedded in the component) will not impact other processes.
+For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) will not impact other processes.
 
 In total, seven cyber requirements are generated and imported into our model.
-These include four \textit{well-formedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
+These include four \textit{wellformedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
 Requirements are imported into the model as goals in a Resolute assurance case.
@@
 
-The intent of the \textit{well-formedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
+The intent of the \textit{wellformedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
 In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
 By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
-The Filter transform is therefore applied for each well-formedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
+The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
 
 The filter behavior for each component is specified in AGREE.
-Not only does this enable formal verification within the modeling environment, but it also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.
+This enables formal verification within the modeling environment, and also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.
 Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.
@@
 The Geofence Monitor is a gated monitor; it prevents the observed Automation Response message from reaching the Waypoint Manager.
-
+%
 Similar to the filters, the monitor policies are specified in AGREE. For example, 
@@
 
-The requirements that have been addressed so far mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
-But we also want to protect against a compromised Ground Station that could potentially transmit well-formed, but malicious commands.
-The final cyber-requirement is mitigated by the Attestation transform~\cite{attestation-copland}, which adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
+So far, we have addressed requirements that 
+%The requirements that have been addressed so far
+mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
+But we also want to protect against a compromised Ground Station transmitting wellformed, but malicious commands.
+The final cyber requirement is mitigated by the Attestation transform~\cite{attestation-copland} that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
 The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.
@@
 
-The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicate with each other over Ethernet.
+The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
 The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
@@
 In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
+%
+We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
+A third party evaluator repeated the tests on the hardened system and % were able to 
+demonstrated that our mitigations were successful.
+The following scenarios were exercised (status messages from the high-assurance components in the hardened system are in \figref{fig:mitigation-output}):
 
-We first performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.  
-A third party evaluator repeated the tests on the hardened system and were able to demonstrate that our mitigations were successful.
-The following scenarios were exercised (status messages from the high-assurance components in the hardened system are shown in \figref{fig:mitigation-output}):
-
-\paragraph{Infected Ground Station} In this scenario, one of the application files is modified on the Ground Station, which sends the UAV on a trajectory that violates mission requirements. 
+\paragraph{Infected Ground Station} An application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
 On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.
 
-\paragraph{Malformed Line Search Task message} In this scenario, the Line Search Task message contains a waypoint with a longitude value outside the permitted range.
-Such malformed messages could exploit vulnerabilities in the on-board software.
-This vulnerability is mitigated by inserting a well-formedness filter.  The filter prevented the Line Search Task message from reaching UxAS.
+\paragraph{Malformed Line Search Task message} The Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
+The wellformedness filter for Line Search Task messages prevented this message from reaching the UxAS.
 
-\paragraph{UxAS vulnerability exploit} In this scenario, Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, which triggered a vulnerability that crashes UxAS.
-When this occurs, UxAS is prevented from generating an Automation Response.
-This vulnerability is mitigated by inserting a Response Monitor, which checks to see that UxAS outputs an Automation Response message shortly after receiving an Automation Request.
+\paragraph{UxAS vulnerability exploit} Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
+When this occurs, UxAS is unable to generate an Automation Response.
+This vulnerability is mitigated by inserting a Response Monitor that checks to see if UxAS outputs an Automation Response message shortly after receiving an Automation Request.
 For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.
 
-\paragraph{UxAS trojan modifies flight plan} In this scenario, a trojan embedded in UxAS attempts to cause the UAV to fly into the specified keep-out zone by modifying the mission command waypoints in the Automation Response.
-On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and instead returned the UAV to Home Base.
+\paragraph{UxAS trojan modifies flight plan} A trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
+On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and returned the UAV to Home Base.
 

### 2021-05-15 paper/conclusion.tex
Full pass done. Spell check. Minor edits. Paper fits.

--- 2021-05-15 paper/conclusion.tex
@@
-The DARPA CASE program is creating tools for systems engineers to integrate cyber-vulnerability analysis and mitigation into their development workflow.
-The resulting BriefCASE tool suite includes analysis tools for generating cyber requirements, cyber resiliency tools for addressing the requirements, verification tools for ensuring design correctness, and synthesis tools for generating provably correct code.
-Several of the BriefCASE transforms (filter, monitor, gate) insert components into the model whose behavior can be formally specified using the AGREE language.
-The SPLAT tool can then automatically generate CakeML implementations for these components, along with proofs of correctness for assurance that the implementation satisfies the specification.
+The DARPA CASE program is creating tools for systems engineers to
+integrate cyber-vulnerability analysis and mitigation into their
+development workflow. The resulting BriefCASE tool suite includes
+analysis tools for generating cyber requirements, cyber resiliency
+tools for addressing the requirements, verification tools for ensuring
+design correctness, and synthesis tools for generating provably
+correct code. Several of the BriefCASE transforms (filter, monitor,
+gate) insert components into the model whose behavior can be formally
+specified using the AGREE language.  The SPLAT tool can then
+automatically generate CakeML implementations for these components,
+along with proofs of correctness for assurance that the implementation
+satisfies the specification.
+%
+BriefCASE was applied to a full-scale case study using the Air Force
+Research Laboratory's OpenUxAS software, exercising a range of
+built-in cyber resiliency mitigations to meet cyber-requirements.  The
+size and scale of the study suggests BriefCASE meets the complexity
+demands of real-world design.
 
-BriefCASE was applied to a full-scale case study using the Air Force Research Laboratory's OpenUxAS software, exercising a range of built-in cyber resiliency mitigations to meet cyber-requirements. 
-The size and scale of the study suggests BriefCASE meets the complexity demands of real-world design.
-
-We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the DARPA CASE program.
-Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.
+We are currently in the process of applying BriefCASE to the design of
+an application using the Collins Common Avionics Architecture System
+(CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the
+DARPA CASE program.  Other ongoing and future work includes
+adding support for uninterpreted functions, mechanizing the
+correctness-of-synthesis proofs in HOL4, and lifting the proof results
+to infinite streams.

### 2021-05-15 paper/example.tex
Full pass done. Spell check. Minor edits. Paper fits.

@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{example.png} \\
+      \includegraphics[scale=0.3]{example.png} \\
       (a) \\
-      \includegraphics[scale=0.4]{example-certificate.png} \\
+      \includegraphics[scale=0.3]{example-certificate.png} \\
       (b)
@@
 A cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
-In response, the developer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output.
+In response, the system designer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously
+(as an untrusted component)
+by removing any guarantees about its output.
 In other words, the AI output is unconstrained in the AGREE specification, allowing it to take on any value and allowing the AI component to send that value at any time.
@@
 For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner output is unconstrained.
-The systems designer is also responsible for defining the meaning of ``well-formed''; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.
+The systems designer is also responsible for defining the meaning of well-formed; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.
 
-The developer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.
+The designer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.
 The first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream).
@@
       (a) \\
-      \includegraphics[scale=0.4]{hardened-certificate.png} \\
+      \includegraphics[scale=0.3]{hardened-certificate.png} \\
       (b)

### 2021-05-15 paper/related-work.tex
Full pass done. Spell check. Minor edits. Paper fits.

@@
 Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}.
-Algorithms for actual component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
-These synthesis algorithms generate code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.
+Algorithms for %actual
+component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
+These algorithms synthesize code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.
 
-Contracts are similar to assume-guarantee reasoning but are targeted to programming languages.
-Contracts are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}.
-As contracts are often written in the target language, synthesis is not a problem for monitoring but comes with significant overhead \cite{10.1007/978-3-642-28869-2_11}.
+Contracts are similar to assume-guarantee reasoning except they target programming languages.
+They are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}.
+As contracts are often written in the target language, synthesizing monitors is not a problem but comes with %significant
+overhead \cite{10.1007/978-3-642-28869-2_11}.
 A monitor for a contract can be removed when it can be statically proven that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.

### 2021-05-15 paper/synthesis.tex
Full pass done. Spell check. Minor edits. Paper fits.

@@
 buffer into a data structure specified in AGREE and apply a
-user-defined \emph{well-formedness} property, also specified in AGREE,
-to the data. Arbitrarily complex well-formedness checks can be made in
+user-defined \emph{wellformedness} property, also specified in AGREE,
+to the data. Arbitrarily complex wellformedness checks can be made in
 this way. \figref{fig:filter-spec} shows a combination where the
@@
 message might need more space than has been allocated; another example
-could be if the timeslice provided by the scheduler is too short for
+could be if the time slice provided by the scheduler is too short for
 the wellformedness computation to finish. Thus resource bounds need to
@@
 
-The contiguity type specification and well-formedness predicate for
+The contiguity type specification and wellformedness predicate for
 the filter are shown in \figref{fig:filter-spec} and the synthesized
@@
 basic notion of a monitor is that it embodies a predicate over its
-input and output streams, able to access the value of a stream at any
+input and output streams, and is able to access the value of a stream at any
 earlier point in time, if necessary. Monitors commonly use state to
@@
 behavior. Until that threshold is attained, the monitor is essentially
-in a kind of \emph{initialization} phase. In order for correct code to
+in its \emph{initialization} phase. In order for correct code to
 be generated, monitor specifications need to spell out the values of
-output ports when in an initialization phase. For example, suppose a
+output ports when in their initialization phases. For example, suppose a
 monitor does some kind of differential assessment of inputs at
-adjacent timeslices, alerting when (say) the measured location of a
+adjacent time slices, alerting when (say) the measured location of a
 UAV at times $t$ and $t+1$ is such that the distance between the two
@@
 
-Intutively, for monitor specification $s$, \konst{stepFn} is the
+Intuitively, for monitor specification $s$, \konst{stepFn} is the
 concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
@@
 a sequence of inputs, and an initial state meeting the initialization
-constraints, iterating \konst{stepFn} produces a $\pi \in
-\Lang{s}$. Morevover, taking the union over all input sequences and
+constraints, iterating \konst{stepFn} produces a $\pi$ s.t. $\pi \in
+\Lang{s}$; and taking the union over all input sequences and
 initial states produces $\Lang{s}$ itself.

### 2021-08-01 paper/paper.tex
Added badges around paper title

@@
 \usepackage{flushend}
+\usepackage{hyperref}
+\usepackage[firstpage]{draftwatermark}
+
+\SetWatermarkAngle{0}
+\SetWatermarkFontSize{8pt}
+\SetWatermarkText{\raisebox{\textheight}{%
+\begin{minipage}{\textwidth}%
+\vspace{-18pt}\hfill%
+\href{http://doi.org/10.5281/zenodo.5129276}{% REPLACE WITH YOUR ARTIFACT DOI
+\includegraphics[width=0.6in, height=0.6in]{reusable}% IF REUSABLE BADGE GRANTED
+\includegraphics[width=0.6in, height=0.6in]{available}% IF AVAILABLE BADGE GRANTED
+}
+\end{minipage}%
+}}
 
@@
 \title{
-  Synthesizing Verified Components for Cyber Assured Systems Engineering 
+  Synthesizing Verified Components~~~\\
+  for Cyber Assured Systems Engineering~
   \thanks{DISTRIBUTION STATEMENT A.  Approved for public release.}

### 2021-08-01 paper/case-study.tex
Updated intro and case-study with more detailed scope

@@
 In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
+The motivation here was to showcase the toolchain on an industrial scale problem, to see if the technology works, i.e.
+provides engineers with early validation of their model; provides early information about cyber-vulnerabilities and in the effectiveness of related mitigations; to show that the tool can handle the UXAS message format (a complex, real-world message format); and to validate the utility and capability of the toolchain.
+
 %apply the BriefCASE tool towards the development of a UAV surveillance system in which 

### 2021-08-01 paper/intro.tex
Updated intro and case-study with more detailed scope

@@
 
+CASE was scoped to address the more common cyber-vulnerabilities that programmers introduce, such as overflow, lack of input validation, supply-chain issues, and identity verification; and not cyber-vulnerabilities such as side-channel attacks and denial of service.
+In this program, we have not invented a new type of filter or monitor in terms of capability, instead our contribution is in the automated synthesis of these components from a formal specification, and a means to show that the synthesis is correct.
+
 The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}.

### 2021-08-01 paper/case-study.tex
Updated case-study; removed figure to shorten it and added some more motivation for it

@@
 In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
-The motivation here was to showcase the toolchain on an industrial scale problem, to see if the technology works, i.e.
+The motivation was to showcase the toolchain on an industrial scale problem, to see if the technology works, i.e.
 provides engineers with early validation of their model; provides early information about cyber-vulnerabilities and in the effectiveness of related mitigations; to show that the tool can handle the UXAS message format (a complex, real-world message format); and to validate the utility and capability of the toolchain.
@@
 %apply the BriefCASE tool towards the development of a UAV surveillance system in which 
-The UAV receives commands from a ground station to conduct surveillance over a geographical region.
+In this surveillance system, the UAV receives commands from a ground station to conduct surveillance over a geographical region.
 In response, the UAV's on-board mission computer generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
@@
 demonstrated that our mitigations were successful.
-The following scenarios were exercised (status messages from the high-assurance components in the hardened system are in \figref{fig:mitigation-output}):
+%The following scenarios were exercised (status messages from the high-assurance components in the hardened system are in \figref{fig:mitigation-output}):
+The following scenarios were exercised:
 
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figs/mitigation-output.png}
-	\caption{Cyber-resilient system response.} 
-	\label{fig:mitigation-output} 
-\end{figure}
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.8\columnwidth]{figs/mitigation-output.png}
+% 	\caption{Cyber-resilient system response.} 
+% 	\label{fig:mitigation-output} 
+% \end{figure}

### 2021-08-01 paper/intro.tex
Slight wordsmithing

--- 2021-05-15 paper/intro.tex
@@
 Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
-Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
+Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
 
@@
 This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
-CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries.
+CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML source code is exactly preserved in the final binaries.
 Assuming that the execution schedule of the deployed cyber-hardened system is as intended by the AADL model, and that the HAMR-generated communication fabric delivers messages between components as expected, the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
 Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+
+CASE was scoped to address the more common cyber-vulnerabilities that programmers introduce, such as overflow, lack of input validation, supply-chain issues, and identity verification; but not other cyber-vulnerabilities such as side-channel attacks and denial of service.
+In this program, we have not invented a new type of filter or monitor in terms of capability, instead our contribution is in the automated synthesis of these components from a formal specification, and a means to show that the synthesis is correct.
 

### 2021-08-01 paper/case-study.tex
Slight wordsmithing

@@
 In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
-The motivation was to showcase the toolchain on an industrial scale problem, to see if the technology works, i.e.
-provides engineers with early validation of their model; provides early information about cyber-vulnerabilities and in the effectiveness of related mitigations; to show that the tool can handle the UXAS message format (a complex, real-world message format); and to validate the utility and capability of the toolchain.
+The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides engineers with early validation of their model; produces early information about cyber-vulnerabilities; provides effective mitigations; and handles a complex real-world message format, namely  the UxAS message format.
 

### 2021-10-13 paper/synthesis.tex
Update presentation

@@
      Check     : Assert
-      (~90.0 <= Latitude and Latitude <= 90.0 andp
+      (~90.0 <= Latitude and Latitude <= 90.0 and
        ~180.0 <= Longitude and Longitude <= 180.0 and

### 2022-01-31 paper/paper.tex
Update to SOSYM style and add summary new content to add

--- 2021-05-15 paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
+% TODO: add in the referee tag for the submission
+%\documentclass[global,twocolumn,referee]{svjour}
+\documentclass[global,twocolumn]{svjour}
 \usepackage{cite}
 \usepackage{amsmath}
-\usepackage{amsthm}
+% \usepackage{amsthm}
 \usepackage{amsfonts}
@@
 \usepackage{flushend}
+\usepackage{hyperref}
+\usepackage[firstpage]{draftwatermark}
 
@@
 
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
+% \theoremstyle{plain}
+% \newtheorem{theorem}{Theorem}
+% \newtheorem{lemma}{Lemma}
 
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
+% \theoremstyle{definition}
+% \newtheorem{definition}{Definition}
 
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
+% \theoremstyle{remark}
+% \newtheorem{remark}{Remark}
+
+\journalname{Software and Systems Modeling}
 
@@
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
 \author{
-\IEEEauthorblockN{Eric Mercer}
-\IEEEauthorblockA{Department of Computer Science\\
-Brigham Young University\\
-Provo, Utah}
-\and
-\IEEEauthorblockN{
-Konrad Slind, Isaac Amundson, Darren Cofer}
-\IEEEauthorblockA{
-Applied Research and Technology\\
-Collins Aerospace\\
-Minneapolis, Minnesota}
-\and
-\IEEEauthorblockN{
-Junaid Babar, David Hardin}
-\IEEEauthorblockA{
-Applied Research and Technology\\
-Collins Aerospace\\
-Cedar Rapids, Iowa}
+  Eric Mercer\inst{1}    \and 
+  Konrad Slind\inst{2}   \and
+  Isaac Amundson\inst{2} \and
+  Darren Cofer\inst{2}   \and
+  Junaid Babar\inst{3}   \and
+  David Hardin\inst{3}
 }
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
+
+\institute{
+  Brigham Young University        \\
+  Provo, Utah                     \and
+  Applied Research and Technology \\ 
+  Collins Aerospace               \\
+  Minneapolis, Minnesota          \and
+  Applied Research and Technology \\ 
+  Collins Aerospace               \\
+  Cedar Rapids, Iowa
+}
+
+\date{Received: date / Revised version: date}
 
@@
 
-\IEEEpeerreviewmaketitle
-
 \section{Introduction} \label{sec:intro}
 \input{intro}
+\subsection{SOSYM Journal Details}
+The submission must have 30\% new material over what was published in MODELS 2021.
+There is no page limit.
+The article will go through a full review cycle.
+The due date is \textbf{March} $1^\mathrm{st}$, \textbf{2022}.
 
+\noindent The summary of the discussion regarding how the MODELS publication might be extended is given below.
+\begin{compactitem}
+  \item Expand the case study with the TA6 models
+  \item Expand the case study with an actual full AGREE specification
+  \item AGREE formal semantics summary including the uninterpreted functions
+  \item Strengthen the assurance argument from AGREE to binary by way of CakeML
+  \item Expand related work (add recent work published after 2018). 
+  \item Respond more fully to the reviews from MODELS (see below).
+\end{compactitem}
+
+\noindent The summary of items from the MODELS reviews.
+\begin{compactitem}
+  \item Add recent work published after 2018
+  \item Add discussion on the limitation of the approach: what attacks can and cannot be covered, can valid messages be rejected, etc.
+  \item Expand the role of Briefcase in the case study with input examples to provide more proof of the proposed model's effectiveness on the hardened system (e.g., show a malformed line search task, it's filter, and why it works)
+  \item Clarify research questions answered by the case study and make the figures bigger
+  \item Discuss what efforts are optimized with the automations versus a fully manual process for creating the components--how does the automated transformation affect actual time for engineers
+  \item Without verifying the correctness of the transformations, can prevention of certain attacks be guaranteed?
+\end{compactitem}
 \section{Simple Example} \label{sec:example}
@@
 
-%\section*{Acknowledgments}
-%This work was sponsored in part by the Defense Advanced Research
-%Projects Agency (DARPA).
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
 \clearpage
-\bibliographystyle{IEEEtran}
+\bibliographystyle{plain}
 \bibliography{paper}

### 2022-01-31 paper/paper.tex
Add submission details

@@
 \end{compactitem}
+\noindent Go to \url{https://mc.manuscriptcentral.com/sosym/}.
+Create a new account (if you don't have one), login and go to the "Author"-section, select the link "Start New Submission" under Author Dashboard and select "Special Section Paper" as the paper type, and choose "MODELS 2021 Special Issue".
+Also, make sure to mention the MODELS 2021 special issue in the cover letter.
+
 \section{Simple Example} \label{sec:example}

### 2022-02-24 paper/agree.tex
Full pass done. Spell check. Minor edits. Paper fits.
snapshot

@@
 the high-assurance components. The AGREE specifications generated by
-the transforms for the high-assurance components are
+the transforms for the high-assurance components are also
 explained. The section ends with a concise formal statement of the
@@
 
-The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
-The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
-As with Lustre, the semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions.
-The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.
-Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
-The semantics do not model computation or communication delay.
-The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
+The AGREE specification for the SW component in the example of
+Section~\ref{sec:example} with the added cyber requirements is given
+in \figref{fig:sw}. The AGREE specification language is a first-order
+predicate calculus that uses stream concepts, and operators, from the
+Lustre language \cite{10.1145/41625.41641}. As with Lustre, the
+semantics are synchronous dataflow where the inputs, outputs, and
+expressions are data streams that comply with the input
+assumptions. Contracts are evaluated in dependency order with inputs
+being propagated to outputs through all contracts until they
+stabilize; as such, the contracts, and thereby the top-level model,
+must be acyclic.\footnote{An apparent syntactic cycle, where a
+component is linked back to itself, may be broken temporally by
+inserting delay elements.} Once the contracts have stabilized, the
+model takes a synchronous step to the next input data in the
+stream. The semantics do not model computation or communication
+delay. The output of one contract is seen at the input of any
+downstream contract in the same step of the input data stream.
 
-The AGREE model checker attempts to prove several properties of the top-level model being verified.
-The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input as well as the guarantees on the system output).
-These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
-The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself).
-These are the unexpanded results at the bottom of the figures.
+The AGREE model checker attempts to prove several properties of the
+top-level model being verified. The first is that the output
+guarantees of each component implementing the system are strong enough
+to validate the input assumptions of any downstream component as well
+as to satisfy the guarantees of the output of the top-level component
+being verified (\ie, the system composition meets input assumptions
+at each input as well as the guarantees on the system output). These
+properties are reported in the expanded lists
+in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set
+of properties prove that the contract specifications for each
+component are \emph{self-consistent} \ie, a contract does not contradict
+itself. These are the unexpanded results at the bottom of the
+figures.
 
-Returning back to the contract in \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
-For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
-In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
-An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
-The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.
+Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq}
+statements to define variables local to the contract
+specification. For example, the \texttt{req} variable is equivalent to
+the \texttt{event(AutomationRequest)} expression. In the AGREE
+semantics, there is an implicit \emph{event} input (or output)
+associated with every named event port in a component. The semantics
+used here do not buffer these events so the implicit input (or output)
+is only a boolean value. An \texttt{event} expression refers to that
+implicit input (or output) and is true when data is placed on the
+named port. The system contract here states assumptions on well-formed
+input, followed by guarantees on properties about the output.
 
-The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding.
-The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
-The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
+The \emph{Alert if start is not bounded relative to a request}
+guarantee is an invariant on the expression \texttt{policy or since},
+meaning that either the policy holds or the alert is
+sounding. The \texttt{policy} is defined by two local
+values: \texttt{current} and \texttt{previous}. The \texttt{current}
+value is asserted when in the current time step there is a request
+with a response, or there is no request and no response.
 
@@
 
-\subsection{Brief Semantics Definition}
+\subsection{AGREE Semantics}
 \label{agree-semantics}
-Here the formal semantics of the AGREE contract specification are briefly
-presented to make clear the meaning of a high-assurance component.
-These semantics are used in the next section to argue that the synthesis
-preserves the same input to output behavior as the AGREE specification.
 
-Assume that all data is in its raw form which is a contiguous sequence of bytes
-representing exactly what is sent over a wire by the communication fabric, so a
-datum is given by a \konst{string}. This assumption is important to the
-correctness argument in the next section. An \emph{environment},
-$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to strings where
-an L-value is anything that can appear on the left-hand side of an assignment
-such as a named port, a field in a record, a entry in an array, a local
-value defined by an eq-statement etc.
+Here the formal semantics of the AGREE contract specification are
+briefly presented to make clear the meaning of a high-assurance
+component.  These semantics are the basis of proofs showing that a
+synthesized component has input to output behavior specified by its
+contract.
+
+Assume that all data flowing between components is in the form of byte
+arrays, thus input and output is modelled at the level of what is
+passed ``over the wire'' by the communication fabric. An \emph{environment},
+$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
+strings where an L-value is anything that can appear on the left-hand
+side of an assignment such as a named port, a field in a record, a
+entry in an array, a local value defined by an eq-statement etc.
 

### 2022-03-09 paper/agree.tex
Structure paper for new semantics section

@@
 As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
-
-\subsection{AGREE Semantics}
-\label{agree-semantics}
-
-Here the formal semantics of the AGREE contract specification are
-briefly presented to make clear the meaning of a high-assurance
-component.  These semantics are the basis of proofs showing that a
-synthesized component has input to output behavior specified by its
-contract.
-
-Assume that all data flowing between components is in the form of byte
-arrays, thus input and output is modelled at the level of what is
-passed ``over the wire'' by the communication fabric. An \emph{environment},
-$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
-strings where an L-value is anything that can appear on the left-hand
-side of an assignment such as a named port, a field in a record, a
-entry in an array, a local value defined by an eq-statement etc.
-
-Let $s$ be an AGREE contract specification for some high-assurance
-component. The notation $\theta_s$ is used to denote the environment that
-contains a binding for any L-value in the scope of $s$ and nothing else, and
-the notation $\Theta_s$ denotes the universe of all such environments. The
-semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
-are finite sequences of environments, $\pi \in \Theta_s^*$.
-
-%% \begin{comment}
-%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
-%% generate a contig type for the inputs plus state vars.} ]
-
-%% [Question: what is a \emph{total} environment? One that supplies at
-%% least the bindings in $\theta$ needed to evaluate the spec?]
-%% \end{comment}
-
-The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
-$\pi$, and returns \konst{true} if $s$ is invariant along the entire
-stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
-is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
-while an eq-statement in $s$ is invariant if its binding in the
-context of every step is equivalent to the computed value of its
-associated expression in that same context. All guarantees and
-eq-statements must be invariant in the stream for the function to
-return \konst{true}.
-
-The meaning of an AGREE contract specification is now defined as the
-set of environment streams on which it is invariant.
-%
-\[
-  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
-\]
-%
-\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
-the L-values in the eq-statements in a way that is consistent with
-their associated expressions and the guarantees are all true in that
-same step.
-
-We claim that a synthesized high-assurance component preserves the
-input to output relationship in the specification $s$ over every
-stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
-s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
-synthesized component on the stream $\pi$ to create a new stream
-$\pi^\prime$ containing added output and other bindings. We say that
-two streams are equivalent in regards to a specification, denoted as
-$\pi =_s \pi^\prime$, if and only if the two streams are the same
-length and agree on bindings for the input and output for $s$ at every
-step of the streams.  We now formally state the correctness claim for
-synthesis.
-%
-\[
-  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
-\]
-%
-where $I_s(\pi)$ returns the corresponding stream that only retains
-bindings for inputs in each step and nothing else. The claim is that
-the synthesized component generates the same output stream as the
-specification for any stream belonging to the specification that is
-restricted to just input bindings at each step.

### 2022-03-09 paper/example.tex
Structure paper for new semantics section

@@
 If the contract model verification succeeds, then the meaning of those results hold for the deployed system, given that all other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.
+
+\subsection{Revisions}
+Add a background section for AGREE and how it works hierarchically to verify a system implements its contract specification.
+
+Move example to its own repository as a staging ground for archiving the files as an artifact
+
+Reformat text to put single sentence on each line.
+
+Revise example to state the system is initially designed assuming all components behave as expected: inputs and outputs are well-formed and timing is obeyed.
+The cyber analysis reveals the potential of the supple chain vulnerability with the AI.
+The designer encodes the vulnerability in contracts by making assumptions about input and output behavior for trusted components.
+Only input integrity is assumed for the untrusted component.
+All other behavior is unconstrained.
+
+Make clear the properties added:
+\begin{compactitem}
+  \item Assume at the system level that a new automation request is not sent unless the prior request, if any, is acknowledged
+  \item Assume at all levels input messages are well-formed
+  \item Guarantee at all levels that waypoint outputs coincide with air vehicle location messages
+  \item Guarantee that a start from the waypoint manager coincides with an automation response
+  \item Guarantee for all trusted components that outputs are well-formed.
+\end{compactitem}
+
+Add in the AGREE specification for the system.
+The figures should not change.
+The AGREE results should changes with new property names (make short enough to fit summary format in figure).

### 2022-03-09 paper/agree.tex
Format and mark text that needs to change

@@
-The goal of this section is to illustrate in more detail the process a
-system designer follows to add cyber requirements to the AGREE
-specifications and then automatically transform the system to insert
-the high-assurance components. The AGREE specifications generated by
-the transforms for the high-assurance components are also
-explained. The section ends with a concise formal statement of the
-meaning of an AGREE specification for a high-assurance component. This
-meaning is what must be preserved by the synthesis.
+The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
+The AGREE specifications generated by the transforms for the high-assurance components are also explained.
 
@@
 
-The AGREE specification for the SW component in the example of
-Section~\ref{sec:example} with the added cyber requirements is given
-in \figref{fig:sw}. The AGREE specification language is a first-order
-predicate calculus that uses stream concepts, and operators, from the
-Lustre language \cite{10.1145/41625.41641}. As with Lustre, the
-semantics are synchronous dataflow where the inputs, outputs, and
-expressions are data streams that comply with the input
-assumptions. Contracts are evaluated in dependency order with inputs
-being propagated to outputs through all contracts until they
-stabilize; as such, the contracts, and thereby the top-level model,
-must be acyclic.\footnote{An apparent syntactic cycle, where a
-component is linked back to itself, may be broken temporally by
-inserting delay elements.} Once the contracts have stabilized, the
-model takes a synchronous step to the next input data in the
-stream. The semantics do not model computation or communication
-delay. The output of one contract is seen at the input of any
-downstream contract in the same step of the input data stream.
+The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
+The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
+As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
+Contracts are evaluated in dependency order with inputs being propagated to outputs through all contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.\footnote{An apparent syntactic cycle, where a component is linked back to itself, may be broken temporally by inserting delay elements.}
+Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
+The semantics do not model computation or communication delay.
+The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
 
-The AGREE model checker attempts to prove several properties of the
-top-level model being verified. The first is that the output
-guarantees of each component implementing the system are strong enough
-to validate the input assumptions of any downstream component as well
-as to satisfy the guarantees of the output of the top-level component
-being verified (\ie, the system composition meets input assumptions
-at each input as well as the guarantees on the system output). These
-properties are reported in the expanded lists
-in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set
-of properties prove that the contract specifications for each
-component are \emph{self-consistent} \ie, a contract does not contradict
-itself. These are the unexpanded results at the bottom of the
-figures.
+The AGREE model checker attempts to prove several properties of the top-level model being verified.
+The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (\ie, the system composition meets input assumptions at each input as well as the guarantees on the system output).
+These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
+The next set of properties prove that the contract specifications for each component are \emph{self-consistent} \ie, a contract does not contradict itself.
+These are the unexpanded results at the bottom of the figures.
 
-Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq}
-statements to define variables local to the contract
-specification. For example, the \texttt{req} variable is equivalent to
-the \texttt{event(AutomationRequest)} expression. In the AGREE
-semantics, there is an implicit \emph{event} input (or output)
-associated with every named event port in a component. The semantics
-used here do not buffer these events so the implicit input (or output)
-is only a boolean value. An \texttt{event} expression refers to that
-implicit input (or output) and is true when data is placed on the
-named port. The system contract here states assumptions on well-formed
-input, followed by guarantees on properties about the output.
+Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
+For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
+In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
+The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
+An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
+The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.
 
-The \emph{Alert if start is not bounded relative to a request}
-guarantee is an invariant on the expression \texttt{policy or since},
-meaning that either the policy holds or the alert is
-sounding. The \texttt{policy} is defined by two local
-values: \texttt{current} and \texttt{previous}. The \texttt{current}
-value is asserted when in the current time step there is a request
-with a response, or there is no request and no response.
+REPLACE ALL: 
+The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding.
+The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
+The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
 
@@
 Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.
-Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.
-%Additionally, the filter can be attached to any requirement the vulnerability analysis.
-This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021].
-The \emph{OK} button creates the component and inserts it accordingly into the implementation
+REMOVE: Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.
 
@@
 The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
-%A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed.
 The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).

### 2022-03-14 paper/semantics.tex

Brief, high-level overview, of the AGREE semantics for component synthesis.
These semantics are those defined in HOL.
The component is defined by its state.
The state is its internal state and current output.
Input streams are constrained to obey the assumptions.
At each input, the \textbf{eq}-statements are evaluated, in order, to update the internal state and outputs.
After the \textbf{eq}-statements are evaluated, guarantees are checked.
The section should define clearly what it means for a component to be \emph{well-formed}. 
The section should conclude by stating the \emph{Component Correct} definition.

The correctness, and structure, of the synthesized CakeML code is based on these semantics.

\subsection{Old AGREE Semantics}
\label{agree-semantics}

Here the formal semantics of the AGREE contract specification are
briefly presented to make clear the meaning of a high-assurance
component.  These semantics are the basis of proofs showing that a
synthesized component has input to output behavior specified by its
contract.

Assume that all data flowing between components is in the form of byte
arrays, thus input and output is modelled at the level of what is
passed ``over the wire'' by the communication fabric. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
strings where an L-value is anything that can appear on the left-hand
side of an assignment such as a named port, a field in a record, a
entry in an array, a local value defined by an eq-statement etc.

Let $s$ be an AGREE contract specification for some high-assurance
component. The notation $\theta_s$ is used to denote the environment that
contains a binding for any L-value in the scope of $s$ and nothing else, and
the notation $\Theta_s$ denotes the universe of all such environments. The
semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
are finite sequences of environments, $\pi \in \Theta_s^*$.

%% \begin{comment}
%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
%% generate a contig type for the inputs plus state vars.} ]

%% [Question: what is a \emph{total} environment? One that supplies at
%% least the bindings in $\theta$ needed to evaluate the spec?]
%% \end{comment}

The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
$\pi$, and returns \konst{true} if $s$ is invariant along the entire
stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
while an eq-statement in $s$ is invariant if its binding in the
context of every step is equivalent to the computed value of its
associated expression in that same context. All guarantees and
eq-statements must be invariant in the stream for the function to
return \konst{true}.

The meaning of an AGREE contract specification is now defined as the
set of environment streams on which it is invariant.
%
\[
  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
\]
%
\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
the L-values in the eq-statements in a way that is consistent with
their associated expressions and the guarantees are all true in that
same step.

We claim that a synthesized high-assurance component preserves the
input to output relationship in the specification $s$ over every
stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
synthesized component on the stream $\pi$ to create a new stream
$\pi^\prime$ containing added output and other bindings. We say that
two streams are equivalent in regards to a specification, denoted as
$\pi =_s \pi^\prime$, if and only if the two streams are the same
length and agree on bindings for the input and output for $s$ at every
step of the streams.  We now formally state the correctness claim for
synthesis.
%
\[
  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
\]
%
where $I_s(\pi)$ returns the corresponding stream that only retains
bindings for inputs in each step and nothing else. The claim is that
the synthesized component generates the same output stream as the
specification for any stream belonging to the specification that is
restricted to just input bindings at each step.

### 2022-03-14 paper/agree.tex
minor textual stuff
Revise example and add TODOs in all sections

@@
 \begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree]
+\begin{lstlisting}[style=agree,numbers=left]
 eq req : bool = event(AutomationRequest);
@@
 eq wp : bool = event(Waypoint);
-eq rsp: bool = event(Start);
+eq strt: bool = event(Start);
 eq alrt : bool = event(Alert);
 
-assume "Automation request is well-formed" :
+assume "Automation requests are well-formed" :
   req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
-assume "Air vehicle location is well-formed" :
+assume "Air vehicle locations are well-formed" :
   avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
+assume "One automation request in flight at a time" :
+  true -> 
+  (req => pre(Historically(not req) or Since(not req, strt)));
 
-eq current : bool = (req = rsp);
-eq previous : bool = (req and not rsp) ->
-                      pre(req and not rsp) and (not req and rsp);
-eq policy : bool = current or previous;
-eq since : bool =  alrt or (alrt and (false -> pre(since)));
-
-guarantee "Start includes a waypoint" :
-    rsp => wp;
-guarantee "Locations required after the start waypoint" :
-    (wp and not rsp) => avl;
-guarantee "Waypoint is well-formed" :
+guarantee "Waypoints coincide with air vehicle locations":
+  wp => avl;
+guarantee "Starts include a new waypoint" :
+  strt => wp;
+guarantee "Waypoints are well-formed" : 
   wp => WELL_FORMED_WAYPOINT(Waypoint);
-guarantee "Alert if start is not bounded relative to a request" :
-    policy or since;
+guarantee "Starts within one cycle of requests if not alerting" :
+  (strt => ((not alrt) and req)) -> 
+  (strt => ((not alrt) and (req or pre(req))));
+guarantee "Alert if not started within one cycle of requests" :
+    true -> ((pre(req and not strt) and not strt) => alrt);
+guarantee "Once alerted always alerted" :
+  not alrt or (Once(alrt) and alrt);
 \end{lstlisting}
@@
   \begin{center}
-    \scalebox{0.60}{\usebox{\sw}}
+    \scalebox{0.62}{\usebox{\sw}}
   \end{center}
@@
 \begin{lrbox}{\flt}
-\begin{lstlisting}[style=agree]
+\begin{lstlisting}[style=agree,numbers=left]
 eq policy : bool = 
   WELL_FORMED_AUTOMATION_RESPONSE(Input);
-guarantee Filter_Output "Filter output is well-formed" :
+
+guarantee Filter_Output
+  "Filter output is well-formed" :
   if event(Input) and policy then 
@@
 \begin{lrbox}{\mntr}
-\begin{lstlisting}[style=agree]
-const is_latched : bool =
-  Get_Property(this, CASE_Properties::Monitor_Latched);
+\begin{lstlisting}[style=agree,numbers=left]
+const is_latched : bool = true;
+const MAX_LATENCY : int = 1;
+
 eq rsp : bool = event(Response);
 eq req : bool = event(Request);
-eq current : bool = (req = rsp);
-eq previous : bool = (req and not rsp) ->
-                      pre(req and not rsp) and (not req and rsp);
-eq policy : bool = current or previous;
-eq alert : bool = (not policy)
-                -> ((is_latched and pre(alert)) or not policy);
-guarantee Monitor_Alert
-  "Alert port tracks alert variable" :
+
+eq isPending : bool = Since(not rsp, req and not rsp);
+eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);
+
+eq policy : bool = (rsp => req) ->
+                   (    (isPending => latency < MAX_LATENCY)   
+                    and (rsp => (req or pre(isPending))));
+eq alert : bool = (not policy) -> 
+                  ((is_latched and pre(alert)) or not policy);
+
+assume "One outstanding request at a time" :
+  (true -> (req => not pre(isPending)));
+
+guarantee "Alert port tracks alert variable" :
   event(Alert) = alert;
-guarantee Monitor_Output
-  "Output if not alerted" :
-  if event(Alert) then (not event(Output)) else
-  if event(Response) then (event(Output) and (Output = Response))
-  else (not event(Output));
+guarantee "Output if not alerted" :
+  if (not(alert) and rsp) then
+    event(Output) and (Output = Response)
+  else
+    not (event(Output));    
 \end{lstlisting}
@@
     \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\flt}} \\
-      (a) \\
-      \scalebox{0.60}{\usebox{\mntr}} \\
-      (b)
+      \scalebox{0.62}{\usebox{\flt}} \\
     \end{tabular}
   \end{center}
-  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
-  \label{fig:assurance}
+  \caption{Contract specification for high-assurance filter.}
+  \label{fig:filter}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.62}{\usebox{\mntr}} \\
+    \end{tabular}
+  \end{center}
+  \caption{Contract specification for high-assurance monitor.}
+  \label{fig:monitor}
 \end{figure}
@@
 
-The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a).
+The AGREE contract specification generated by the transform is shown in \figref{fig:filter}.
 The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
-The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).
+The resulting AGREE specification for the monitor in this example is shown in \figref{fig:monitor}.
 The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.

### 2022-03-14 paper/example.tex
Revise example and add TODOs in all sections

--- 2021-05-14 paper/example.tex
@@
-\begin{figure*}
+\begin{figure*}[h]
   \begin{center}
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{example.png} \\
-      (a) \\
-      \includegraphics[scale=0.4]{example-certificate.png} \\
-      (b)
+      \includegraphics[scale=0.4]{example.png}
     \end{tabular}
   \end{center}
-\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
+\caption{Initial design for an automated UAV route planning system.}
 \label{fig:example}
@@
 
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.4]{example-certificate.png} \\
+    \end{tabular}
+  \end{center}
+\caption{AGREE failure certificate for initial design.}
+\label{fig:example-certificate}
+\end{figure}
+
 \figref{fig:example} is an AADL architectural model of a software system (SW) for route planning and automated control of a UAV.
-It is loosely based on the system in the case study introduced in Section~\ref{sec:case-study}.
+It is loosely based on one of the case studies in Section~\ref{sec:case-study}.
 The source for the entire model is found at \cite{repo}.
-The system receives an automation request that is forwarded to an untrusted third-party route planner (AI).
+
+The system receives an automation request that is forwarded to a third-party route planner (AI).
 The route planner decides the flight path of the UAV based on its current position and the requested task.
 The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
-The waypoint manager is an \emph{as is} legacy component.
+The waypoint manager is a repurposed legacy component that cannot be modified.
 
-The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contract specifications.
-The contracts constrain input and state properties of output for component models.
+The expected behavior of the SW system, and the components in its implementation, are specified (i.e., modeled) with AGREE contract specifications.
+The contracts constrain input with assumptions and state properties of output with guarantees.
 AGREE performs model checking on this assume-guarantee system to hierarchically prove that the composite system obeys all contract obligations under all possible finite input streams. 
+The AGREE contracts for the system and its implementing components are discussed in greater detail in \secref{sec:agree}.
 
-The initial AGREE contract specifications for the components and the overall system make no assumptions about the integrity of the inputs and outputs.
-A cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously.
-In response, the developer modifies the AGREE specification for the AI route planner to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output.
-In other words, the AI output is unconstrained in the AGREE specification, allowing it to take on any value and allowing the AI component to send that value at any time.
+The initial contract for the system in \figref{fig:example}, and its implementing components, assume and guarantee the absence of any malicious, or underspecified, component behavior.
+The system assumes that all inputs are \emph{well-formed} and there is never more than one automation request pending at a time.
+The meaning of well-formed is given by predicates for each data type.
+For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
+The system guarantees that a start includes a waypoint, a start is within one cycle of an automation request and if not, then it persistently alerts, new waypoints coincide with air vehicle location updates, and all output are well-formed.
 
-The contract specifications for the other components are also updated with the threat analysis information by adding in new requirements to mitigate the identified cyber-vulnerabilities.
-For example, a designer adds to the specification for the legacy waypoint manager assumptions about its input being \emph{well-formed} since that is no longer known \emph{a priori} as the AI route planner output is unconstrained.
-The systems designer is also responsible for defining the meaning of ``well-formed''; in this example, it is a predicate checking that the waypoints are within bounded value ranges for latitude, longitude, and altitude, with an additional integrity check on message IDs.
+The components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
+The AI further guarantees it only responds to automation requests and always in the same cycle.
+The legacy WM further guarantees a start from a response and always in the same cycle, start always comes with a waypoint, and further waypoints coincide with air vehicle location updates.
+This system implementation passes AGREE verification meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
 
-The developer adds two other requirements to the SW AGREE specification related to the cyber-vulnerability analysis.
-The first, \emph{Waypoint is well-formed} requires all the waypoints sent to the UAV flight controller to be well-formed (it detects if malformed waypoints are propagated downstream).
-The second, \emph{Alert if start is not bounded relative to a request}, requires an automation request to correspond with an automation response either in the same step or within one step (it detects if a mission is being started without a request or if the start is delayed more than one step after a request).
-The goal of these two requirements is to detect if the untrusted component is trying to prevent the UAV from flying its intended mission or to fly a wrong mission.
+A cyber-vulnerability analysis identifies the potential of a supply chain attack through the AI route planner provided by the third-party vendor without source code.
+The component is marked as \emph{untrusted}, and its contract is modified by removing all guarantees on its outputs to reflect its untrusted status.
+The AI output is now unconstrained in the assume-guarantee reasoning of AGREE and is able to generate any value on its output stream at anytime.
+AGREE must consider all possible output streams from the AI in proving the correctness of the component implementation of the automated UAV route planning system.
 
-AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber requirements, by way of model checking.
-The output from the model checker is shown in \figref{fig:example}(b).
-The red exclamation points designate properties that do not hold.
-Each of these failures comes with a counterexample.
-The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about the wellformedness of the input to the legacy waypoint manager.
-The counterexample for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request, a  clear malicious behavior from the untrusted component.
+The output from AGREE with the AI specification is shown in \figref{fig:example-certificate}.
+The red exclamation points designate component assumptions or system output guarantees that do not hold, and each failure comes with a corresponding counter-example.
+The results are not unexpected given the AI route planner's unconstrained behavior.
+
+For example, the first violation is that the automation response from the AI to the WM is no longer guaranteed to be well-formed.
+The consequence of that failing input assumption is that the WM outputs are no longer guaranteed; they are effectively unconstrained.
+These missing output guarantees from the WM lead to the rest of the failures in \figref{fig:example-certificate} because the WM provides the system level outputs.
 
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.3]{hardened.png} \\
-      (a) \\
-      \includegraphics[scale=0.4]{hardened-certificate.png} \\
-      (b)
+      \includegraphics[scale=0.3]{hardened.png}
     \end{tabular}
   \end{center}
-  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
+  \caption{Cyber-hardened design for an automated UAV route planning system}
   \label{fig:hardened}
@@
 
-The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a).
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.4]{hardened-certificate.png}
+    \end{tabular}
+  \end{center}
+  \caption{AGREE verification certificate for cyber-hardened design.}
+  \label{fig:hardened-certificate}
+\end{figure}
+
+The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}.
 A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input violates the filter invariant.
@@
 
-The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}(b).
-Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber vulnerabilities from the untrusted AI route planner.
+The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}.
+Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the untrusted AI route planner.
 
@@
 A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the generated code.
-In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.
+In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML code exactly match those defined by the high-assurance component's contract.
 
-Preserving the input/output relationship of streams between the two models lifts the contract verification results to the deployed system.
+Preserving the input to output relationship of streams between the two models lifts the contract verification results to the deployed system.
 If the contract model verification succeeds, then the meaning of those results hold for the deployed system, given that all other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.

### 2022-03-14 paper/paper.tex
Revise example and add TODOs in all sections

@@
 \usepackage[firstpage]{draftwatermark}
+\usepackage[lighttt]{lmodern}
+\usepackage[normalem]{ulem}
 
@@
   linewidth=2\linewidth,
-  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const}
+  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
 }
@@
 \newcommand{\figref}[1]{Fig.~\ref{#1}}
+\newcommand{\secref}[1]{Sec.~\ref{#1}}
 
@@
 \input{intro}
+
 \subsection{SOSYM Journal Details}
-The submission must have 30\% new material over what was published in MODELS 2021.
-There is no page limit.
-The article will go through a full review cycle.
-The due date is \textbf{March} $1^\mathrm{st}$, \textbf{2022}.
-
-\noindent The summary of the discussion regarding how the MODELS publication might be extended is given below.
-\begin{compactitem}
-  \item Expand the case study with the TA6 models
-  \item Expand the case study with an actual full AGREE specification
-  \item AGREE formal semantics summary including the uninterpreted functions
-  \item Strengthen the assurance argument from AGREE to binary by way of CakeML
-  \item Expand related work (add recent work published after 2018). 
-  \item Respond more fully to the reviews from MODELS (see below).
-\end{compactitem}
-
-\noindent The summary of items from the MODELS reviews.
-\begin{compactitem}
-  \item Add recent work published after 2018
-  \item Add discussion on the limitation of the approach: what attacks can and cannot be covered, can valid messages be rejected, etc.
-  \item Expand the role of Briefcase in the case study with input examples to provide more proof of the proposed model's effectiveness on the hardened system (e.g., show a malformed line search task, it's filter, and why it works)
-  \item Clarify research questions answered by the case study and make the figures bigger
-  \item Discuss what efforts are optimized with the automations versus a fully manual process for creating the components--how does the automated transformation affect actual time for engineers
-  \item Without verifying the correctness of the transformations, can prevention of certain attacks be guaranteed?
-\end{compactitem}
-\noindent Go to \url{https://mc.manuscriptcentral.com/sosym/}.
+The due date is \textbf{March} $22^\mathrm{nd}$, \textbf{2022}.
+Go to \url{https://mc.manuscriptcentral.com/sosym/}.
 Create a new account (if you don't have one), login and go to the "Author"-section, select the link "Start New Submission" under Author Dashboard and select "Special Section Paper" as the paper type, and choose "MODELS 2021 Special Issue".
@@
 
-\section{Simple Example} \label{sec:example}
+\subsection{Revisions}
+\begin{compactitem}
+  \item Add discussion on the limitation of the approach: what attacks can and cannot be covered, can valid messages be rejected, etc.
+\end{compactitem}
+
+\section{Illustrative Example} \label{sec:example}
 \input{example}
+
+\subsection{Revisions}
+\begin{compactitem}
+  \item \sout{Reformat text to put single sentence on each line.}
+  \item Move example to its own repository, archive, and give the DOI and URL for the example.
+  \item \sout{Update AGREE certificates with new model}
+  \item Separate example into separate files for initial design, cyber-components and tests, hardened design, and seL4 process design.
+  \item \sout{Make clear the initial properties for the assumed correct model:}
+    \begin{compactitem}
+      \item \sout{Assume at the system level that a new automation request is not sent unless the prior request, if any, is acknowledged}
+      \item \sout{Assume at all levels input messages are well-formed}
+      \item \sout{Guarantee at all levels that waypoint outputs coincide with air vehicle location messages}
+      \item \sout{Guarantee that a start from the waypoint manager coincides with an automation response}
+      \item \sout{Guarantee for all trusted components that outputs are well-formed}
+    \end{compactitem}
+\end{compactitem}
 
@@
 
-\section{Synthesis} \label{sec:synthesis}
+\subsection{Revisions}
+\begin{compactitem}
+  \item Add in the AGREE specification for the system
+  \item Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)
+\end{compactitem}
+
+\section{AGREE Semantics for Component Synthesis}
+\input{semantics}
+
+\subsection{Revisions}
+\begin{compactitem}
+  \item Add in intuitive definition or AGREE leaf-component semantics
+  \item State well-formed theorem
+  \item State correctness theorem or any key theorems to the synthesis proof
+\end{compactitem}
+
+\section{Synthesis to CakeML and Proof Sketch} \label{sec:synthesis}
 \input{synthesis}
 
-\section{UxAS Case Study} \label{sec:case-study}
+\subsection{Revisions}
+\begin{compactitem}
+  \item State theorem relating the step function to the meaning of the leaf-node semantics
+\end{compactitem}
+
+\section{Case Studies} \label{sec:case-study}
 \input{case-study}
+
+\subsection{Revisions}
+\begin{compactitem}
+  \item Clarify research questions answered by the case study
+  \item Make the figures bigger 
+  \item Rewrite to focus on synthesized components and the AGREE specifications
+  \item Add an interesting monitor from the TA6 model
+  \item Discuss what efforts are optimized with the automations versus a fully manual process for creating the components--how does the automated transformation affect actual time for engineers
+\end{compactitem}
 
@@
 \input{related-work}
+
+\subsection{Revisions}
+\begin{compactitem}
+  \item Add recent work published after 2018
+\end{compactitem}
 

### 2022-03-14 paper/agree.tex
Add AGREE verification conditions

@@
-The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components.
-The AGREE specifications generated by the transforms for the high-assurance components are also explained.
+This section introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
+It details the specification of that top-level system interface.
+It then discusses the specifications for the filter and the monitor added by the transforms.
 
+This section then formalizes AGREE's assume-guarantee reasoning by defining the set of verification conditions that AGREE must check.
+The verification conditions succinctly summarize what in means when AGREE proves that a system implementation satisfies it specification.
+These conditions are useful in understanding how the semantics in synthesis differ from those in the assume-guarantee reasoning in \secref{sec:semantics}.
+
+\subsection{Contract Specification}
 \newsavebox{\sw}
@@
 As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
+
+\subsection{Verification Conditions}
+\input{verification-conditions}

### 2022-03-15 paper/example.tex
Revise example section

@@
 
-The expected behavior of the SW system, and the components in its implementation, are specified (i.e., modeled) with AGREE contract specifications.
+The expected behavior of the SW system, and the components implementing the system, are modeled with AGREE contract specifications.
 The contracts constrain input with assumptions and state properties of output with guarantees.
 AGREE performs model checking on this assume-guarantee system to hierarchically prove that the composite system obeys all contract obligations under all possible finite input streams. 
-The AGREE contracts for the system and its implementing components are discussed in greater detail in \secref{sec:agree}.
+The AGREE analysis, contract language, and the contracts for this example system are discussed in greater detail in \secref{sec:agree}.
 
-The initial contract for the system in \figref{fig:example}, and its implementing components, assume and guarantee the absence of any malicious, or underspecified, component behavior.
-The system assumes that all inputs are \emph{well-formed} and there is never more than one automation request pending at a time.
+The initial AGREE contracts for the system in \figref{fig:example}, and its implementing components, assume and guarantee the absence of any malicious, or unspecified, component behavior.
+More precisely, the system assumes that all inputs are \emph{well-formed} and there is never more than one automation request pending at a time.
 The meaning of well-formed is given by predicates for each data type.
@@
 The legacy WM further guarantees a start from a response and always in the same cycle, start always comes with a waypoint, and further waypoints coincide with air vehicle location updates.
-This system implementation passes AGREE verification meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
+These initial specifications pass AGREE verification meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
 
-A cyber-vulnerability analysis identifies the potential of a supply chain attack through the AI route planner provided by the third-party vendor without source code.
-The component is marked as \emph{untrusted}, and its contract is modified by removing all guarantees on its outputs to reflect its untrusted status.
+A cyber-vulnerability analysis, however, identifies the potential of a supply chain attack through the AI route planner that is provided by a third-party vendor without source code.
+The component is marked as \emph{untrusted} after the analysis, and its contract is modified by removing all guarantees on its outputs to reflect its untrusted status.
 The AI output is now unconstrained in the assume-guarantee reasoning of AGREE and is able to generate any value on its output stream at anytime.
@@
 
-The output from AGREE with the AI specification is shown in \figref{fig:example-certificate}.
+The output from the AGREE analysis using the untrusted AI specification is shown in \figref{fig:example-certificate}.
 The red exclamation points designate component assumptions or system output guarantees that do not hold, and each failure comes with a corresponding counter-example.
@@
 The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further, that response must come with the request or in the next step after the request.
-As with the filter, the system designer provides the policy, and that policy is based on the existing AGREE specification in the SW system.
+As with the filter, the system designer provides the policy, and that policy is typically based on existing AGREE specifications in the SW system.
 
@@
 The high-assurance components are automatically synthesized by SPLAT from the AGREE specifications to equivalent models in the CakeML language.
-CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.
+The synthesis includes a proof that it captures the meaning of the AGREE specification in CakeML.
+In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML code exactly match those defined by the high-assurance component's guarantees.
+From here, CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.
 
-A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML.
-A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the generated code.
-In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML code exactly match those defined by the high-assurance component's contract.
-
-Preserving the input to output relationship of streams between the two models lifts the contract verification results to the deployed system.
-If the contract model verification succeeds, then the meaning of those results hold for the deployed system, given that all other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.
+Preserving the input to output relationship of streams between the AGREE contracts and CakeML lifts the AGREE contract verification results to the actual deployed system.
+If the contract model verification succeeds, then the meaning of those results hold for the deployed system.
+These results, however, are only valid under additional assumptions on the deployed system: contracts for non-synthesized components accurately model their deployed counterparts, an appropriate schedule exists to sequence component execution following the dependent data-flow in the design, and the communication fabric stitching components together works in harmony with the schedule.
+Support and automation for these aspects of the design process are discussed in other works.

### 2022-03-16 paper/agree.tex
Revise the AGREE semantics and reorder section

@@
-This section introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
-It details the specification of that top-level system interface.
-It then discusses the specifications for the filter and the monitor added by the transforms.
-
-This section then formalizes AGREE's assume-guarantee reasoning by defining the set of verification conditions that AGREE must check.
+This section formalizes AGREE's assume-guarantee reasoning by defining the set of verification conditions that AGREE must check.
 The verification conditions succinctly summarize what in means when AGREE proves that a system implementation satisfies it specification.
@@
 
-\subsection{Contract Specification}
+This section then introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
+It details the specification of that top-level system interface.
+It then discusses the specifications for the filter and the monitor.
+These two specifications are those that are synthesized to equivalent CakeML by SPLAT.
+
+\subsection{Verification Conditions}
+\input{verification-conditions}
+
+\subsection{Contract Language Specification}
 \newsavebox{\sw}
 \begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree,numbers=left]
-eq req : bool = event(AutomationRequest);
+\begin{lstlisting}[style=agree,numbers=left,escapeinside={*}{*}]
+eq req : bool = event(AutomationRequest);*\label{line:sw-event-def-start}*
 eq avl : bool = event(AirVehicleLocation);
@@
 eq strt: bool = event(Start);
-eq alrt : bool = event(Alert);
+eq alrt : bool = event(Alert);*\label{line:sw-event-def-end}*
 
-assume "Automation requests are well-formed" :
+assume "Automation requests are well-formed" : *\label{line:sw-assume-1}*
   req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
-assume "Air vehicle locations are well-formed" :
+assume "Air vehicle locations are well-formed" : *\label{line:sw-assume-2}*
   avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
-assume "One automation request in flight at a time" :
+assume "One automation request in flight at a time" : *\label{line:sw-assume-3}*
   true -> 
@@
 
-The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
-The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
+The AGREE specification language uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
 As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
@@
 The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
+The language is best introduced through example.
 
-The AGREE model checker attempts to prove several properties of the top-level model being verified.
-The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (\ie, the system composition meets input assumptions at each input as well as the guarantees on the system output).
-These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
-The next set of properties prove that the contract specifications for each component are \emph{self-consistent} \ie, a contract does not contradict itself.
-These are the unexpanded results at the bottom of the figures.
+The AGREE specification for the SW component in the example of Section~\ref{sec:example} is given in \figref{fig:sw}.
+The specification uses \texttt{eq} statements to define variables local to the contract specification.
+For example, \lineref{line:sw-event-def-start} defines the \texttt{req} variable to be equivalent to the \texttt{event(AutomationRequest)} expression.
 
-Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
-For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
-In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
-An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
-The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.
+All the named ports in the corresponding AADL component are in the scope of the specification, and there are additional implicit boolean \emph{event} inputs (or outputs) associated with event ports.
+An \texttt{event} expression refers to that implicit input (or output) boolean value and is true when data is placed on the named port and false otherwise.
+\linesref{line:sw-event-def-start}{line:sw-event-def-end} create local variables that are true when data is present on the corresponding event ports for the component.
+The local variables here are purely for convenience in writing the specification.
+
+Assumptions are best understood as being evaluated on the pre-state of a component while guarantees are evaluated on the post-state.
+Contract languages such as Eifle and Dafny make this pre-state assume evaluation and post-state guarantee evaluation part of the semantics, and it relies on operators such as \texttt{old} it indicate when a guarantee should use a pre-state value in evaluating a guarantee.
+AGREE differs from these languages in that it evaluates both the assumptions and guarantees in the post-state of the component.
+But like other languages it relies on the operator \texttt{pre} to refer to a pre-state evaluation in both the assumptions and guarantees.
+Since streams are associated with all expressions in AGREE meaning that they are defined through time, the \texttt{pre} operator simply returns the previous value of the enclosed expression (e.g., it looks back one step in time).
+
+The \texttt{assume} statement is a string description followed by a predicate.
+Those on \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2} are implications requiring that when data is present it is well-formed.
+The well-formed predicates themselves are defined elsewhere using AGREE functions.
+The assumption on \lineref{line:sw-assume-3} uses the \emph{followed-by} operator, \texttt{->}, and the \texttt{pre} operator to constrain when requests can arrive at the input.
+
+The followed-by expression defines what happens at the first instance in the stream, time 0, and then what follows after. 
+Here the assumption is \texttt{true} at time 0, and then takes on the truth value of the implication in all future instances.
+The implication uses the \texttt{pre} expression in its left operand. 
+In the example, it is guarded by the followed-by since expressions are undefined before time 0.
+So in the first instance at time 0, the assumption is \texttt{true}, and then at every instance after, it depends on when the request arrives.
+
+The stream semantics in AGREE mean that it is possible to use PLTL operators that look back in time.
 
@@
 As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
-
-\subsection{Verification Conditions}
-\input{verification-conditions}

### 2022-03-16 paper/example.tex
Add new tex files, reformat agree section, and revise that text

@@
 The AI output is now unconstrained in the assume-guarantee reasoning of AGREE and is able to generate any value on its output stream at anytime.
-AGREE must consider all possible output streams from the AI in proving the correctness of the component implementation of the automated UAV route planning system.
 
@@
 
-The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}.
+The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened-certificate}.
 Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the untrusted AI route planner.
+Having passed AGREE verification, the high-assurance components are ready to be synthesized.
 

### 2022-03-16 paper/high-assurance-contracts.tex
Add new tex files, reformat agree section, and revise that text

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \includegraphics[scale=0.3]{dialogue.png}
    \end{tabular}
  \end{center}
  \caption{BriefCASE dialogue for filter transformation.}
  \label{fig:dialogue}
\end{figure}

\newsavebox{\flt}
\begin{lrbox}{\flt}
  \begin{lstlisting}[style=agree,numbers=left]
    -- start user definitions
    eq policy : bool = 
    WELL_FORMED_AUTOMATION_RESPONSE(Input);
    -- stop user definitions

    guarantee Filter_Output
    "Filter output is well-formed" :
    if event(Input) and policy then 
        event(Output) and Output = Input
    else not event(Output);
  \end{lstlisting}
\end{lrbox}

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
  \begin{lstlisting}[style=agree,numbers=left]
    const is_latched : bool = true;

    -- start user definitions
    const MAX_LATENCY : int = 1;

    eq rsp : bool = event(Response);
    eq req : bool = event(Request);

    eq isPending : bool = Since(not rsp, req and not rsp);
    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);

    eq policy : bool = (rsp => req) ->
                        (    (isPending => latency < MAX_LATENCY)   
                        and (rsp => (req or pre(isPending))));
    -- stop user definitions

    eq alert : bool = (not policy) -> 
                    ((is_latched and pre(alert)) or not policy);

    assume "One outstanding request at a time" :
    (true -> (req => not pre(isPending)));

    guarantee "Alert port tracks alert variable" :
    event(Alert) = alert;
    guarantee "Output if not alerted" :
    if (not(alert) and rsp) then
        event(Output) and (Output = Response)
    else
        not (event(Output));    
  \end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

As noted previously, the system with the untrusted AI does not pass AGREE verification.
Transformations in BriefCASE are used to cyber-harden the system against malicious AI behavior by inserting a high-assurance filter and monitor between the AI and the WM as shown in \figref{fig:hardened}.
High-assurance means the behaviors of the components are verified.
Here the filter protects the WM from malformed data and the monitor to protects it from spontaneous or delayed responses.
These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear, and then choosing the appropriate transformation.

The system designer provides configuration parameters for the transformation in a dialogue box as shown in \figref{fig:dialogue} for the filter.
All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
A filter policy defines well-formed data while a monitor policy defines an invariant over inputs and outputs.
These policies can be stated directly in the wizard, or they can be left blank and added later to the AGREE contract generated by the transformation.

The BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
It also creates default AGREE specification that define everything about the component behavior but the policy itself unless it is stated in the dialogue.
The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
Here the policy \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)} was specified in the dialogue box.
The single guarantee completely defines the filter output under any input according to the truth value of the policy.
While such a string guarantee is not required for AGREE analysis, it is required for automated synthesis.

The resulting AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerable more complex than the filter.

The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
This behavior is one of the several options available in the dialogue.
The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

### 2022-03-16 paper/paper.tex
Add AGREE verification conditions
Revise example section
Revise the AGREE semantics and reorder section
Add new tex files, reformat agree section, and revise that text

@@
   linewidth=2\linewidth,
+  morecomment=[l]{--},
   morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
@@
 
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{op-tical net-works semi-conduc-tor Cake-ML}
 
@@
 \newcommand{\secref}[1]{Sec.~\ref{#1}}
+\newcommand{\lineref}[1]{Line~\ref{#1}}
+\newcommand{\linesref}[2]{Lines~\ref{#1}-\ref{#2}}
 
@@
 
-\section{Illustrative Example} \label{sec:example}
+\section{Illustrative Example}
+\label{sec:example}
 \input{example}
@@
     \end{compactitem}
+  \item Add citations to last paragraph for scheduled AGREE and HAMR.
 \end{compactitem}
 
-\section{AGREE Contract Specification} \label{sec:agree}
+\section{Assume-Guarantee Reasoning in AGREE}
+\label{sec:agree}
 \input{agree}
@@
 \section{AGREE Semantics for Component Synthesis}
+\label{sec:semantics}
 \input{semantics}
@@
 
-\section{Synthesis to CakeML and Proof Sketch} \label{sec:synthesis}
+\section{Synthesis to CakeML and Proof Sketch}
+\label{sec:synthesis}
 \input{synthesis}
@@
 
-\section{Case Studies} \label{sec:case-study}
+\section{Case Studies}
+\label{sec:case-study}
 \input{case-study}
@@
 
-\section{Related Work} \label{sec:related-work}
+\section{Related Work}
+\label{sec:related-work}
 \input{related-work}
@@
 
-\section{Conclusion} \label{sec:conclusion}
+\section{Conclusion}
+\label{sec:conclusion}
 \input{conclusion}

### 2022-03-16 paper/sw-contract.tex
Add new tex files, reformat agree section, and revise that text

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree,numbers=left,escapeinside={*}{*}]
eq req : bool = event(AutomationRequest);*\label{line:sw-event-def-start}*
eq avl : bool = event(AirVehicleLocation);
eq wp : bool = event(Waypoint);
eq strt: bool = event(Start);
eq alrt : bool = event(Alert);*\label{line:sw-event-def-end}*

assume "Automation requests are well-formed" : *\label{line:sw-assume-1}*
  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle locations are well-formed" : *\label{line:sw-assume-2}*
  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
assume "One automation request in flight at a time" : *\label{line:sw-assume-3}*
  true -> 
  (req => pre(Historically(not req) or Since(not req, strt)));

guarantee "Waypoints coincide with air vehicle locations": *\label{line:sw-guarantee-1}*
  wp => avl;
guarantee "Starts include a new waypoint" : *\label{line:sw-guarantee-2}*
  strt => wp;
guarantee "Waypoints are well-formed" :  *\label{line:sw-guarantee-3}*
  wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Starts within one cycle of requests if not alerting" : *\label{line:sw-guarantee-4}*
  (strt => ((not alrt) and req)) -> 
  (strt => ((not alrt) and (req or pre(req))));
guarantee "Alert if not started within one cycle of requests" : *\label{line:sw-guarantee-5}*
    true -> ((pre(req and not strt) and not strt) => alrt);
guarantee "Once alerted always alerted" : *\label{line:sw-guarantee-6}*
  not alrt or (Once(alrt) and alrt);
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.62}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

The AGREE specification language uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
Contracts are evaluated in dependency order with inputs being propagated to outputs through all contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.\footnote{An apparent syntactic cycle, where a component is linked back to itself, may be broken temporally by inserting delay elements.}

Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
The semantics do not model computation or communication delay.
The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
The language is best introduced through example.

The AGREE specification for the SW component in the example of Section~\ref{sec:example} is given in \figref{fig:sw}.
The specification uses \texttt{eq} statements to define variables local to the contract specification.
For example, \lineref{line:sw-event-def-start} defines the \texttt{req} variable to be equivalent to the \texttt{event} expression.

All the named ports in the corresponding AADL component are in the scope of the specification, and there are additional implicit boolean \emph{event} inputs (or outputs) associated with event ports.
An \texttt{event} expression refers to that implicit input (or output) boolean value and is true when data present the named port and false otherwise.
\linesref{line:sw-event-def-start}{line:sw-event-def-end} create local variables that are true when data is present on the corresponding event ports for the component.
The local variables here are purely for convenience in writing the specification.

The \texttt{assume} statement is a string description followed by a predicate.
Those on \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2} are implications requiring that when data is present it is well-formed.
The well-formed predicates themselves are defined elsewhere using AGREE functions.

The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
Streams are associated with every expression in AGREE, meaning that expressions are defined through time, and are therefore stateful.
When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input. AGREE provides the \texttt{pre} operator to make that distinction.

The \texttt{pre} operator returns the previous value of the enclosed expression by looking back one step in time on the expression's stream.
In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
Guarantees should use the \texttt{pre} operator anytime they need to reason about the current state relative to the previous state.

The assumption on \lineref{line:sw-assume-3} uses a \emph{followed-by} operator, \texttt{->}, to \emph{guard} the \texttt{pre} operator at time 0.
A followed-by expression defines what happens at the first instance of the stream, time 0, and then what follows after. 
Here the assumption is \texttt{true} at time 0, and then takes on the truth value of the implication in all future instances.
The followed-by guards the \texttt{pre} because expressions are undefined before time 0 and the followed-by prevents the \texttt{pre} from being evaluated before time 1.
So after the first instance, the assumption depends on the presence of a request and the pre-state of the component.

The stream semantics in AGREE mean that it is possible to use PLTL operators that look back in time.
The \texttt{Historically} operator is the same as $\historically$ defined previously; it is true if the enclosed expression has been true in all moments up to the present one. 
The \texttt{Since} operator is true if the first expression has been true in all moments up to the present one since the second expression most recently became true.
The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}.

The guarantees on \linesref{line:sw-guarantee-1}{line:sw-guarantee-3} coincide events and assert well-formed output.
The guarantees on \linesref{line:sw-guarantee-4}{line:sw-guarantee-6} define temporal properties of the component.
\lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request.
The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
The guarantee also does not force the start to take place, it only says that if it takes place, it is under the defined conditions.

\lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound.
Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
The final guarantee on \lineref{line:sw-guarantee-6} defines additional behavior for the alert.

The \texttt{Once} operator is true if the enclosed expression has ever been true before the current moment.
\lineref{line:sw-guarantee-6} ensures that once alerted always alerted because either there has never been an alert, \texttt{\textbf{not} alrt}, or there has been an alert sometime in the past and so it is alerting in now, \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
These six guarantees define the behavior of the SW component under the three assumptions.

### 2022-03-16 paper/high-assurance-contracts.tex
Complete the AGREE explanations in section 3

@@
     -- start user definitions
-    const MAX_LATENCY : int = 1;
+    const MAX_LATENCY : int = 1; *\label{line:mon-const}*
 
@@
 
-    eq isPending : bool = Since(not rsp, req and not rsp);
-    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);
+    eq isPending : bool = Since(not rsp, req and not rsp);*\label{line:mon-pending}*
+    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);*\label{line:mon-latency}*
 
-    eq policy : bool = (rsp => req) ->
+    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
                          (    (isPending => latency < MAX_LATENCY)   
                           and (rsp => (req or pre(isPending))));
+
+    assume "One outstanding request at a time" : *\label{line:mon-assume}*
+      (true -> (req => not pre(isPending))); 
     -- stop user definitions
 
-    eq alert : bool = (not policy) -> 
+    eq alert : bool = (not policy) -> *\label{line:mon-alert}*
                         ((is_latched and pre(alert)) or not policy);
-
-    assume "One outstanding request at a time" :
-    (true -> (req => not pre(isPending)));
 
@@
 High-assurance means the behaviors of the components are verified.
-Here the filter protects the WM from malformed data and the monitor to protects it from spontaneous or delayed responses.
-These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear, and then choosing the appropriate transformation.
+Here the filter protects the WM from malformed data and the monitor protects it from spontaneous or delayed responses.
+These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear and then choosing the appropriate transformation.
 
-The system designer provides configuration parameters for the transformation in a dialogue box as shown in \figref{fig:dialogue} for the filter.
+The system designer provides configuration parameters for the transformation in a dialogue box.
+The dialogue box for adding a filter is shown in \figref{fig:dialogue}.
 All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
@@
 
-The BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
-It also creates default AGREE specification that define everything about the component behavior but the policy itself unless it is stated in the dialogue.
+BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
+It also creates a default AGREE specification that defines everything about the component behavior but the policy itself unless it is provided in the dialogue.
 The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
-Here the policy \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)} was specified in the dialogue box.
 The single guarantee completely defines the filter output under any input according to the truth value of the policy.
-While such a string guarantee is not required for AGREE analysis, it is required for automated synthesis.
+While such a strong guarantee is not required for AGREE analysis, it is required for automated synthesis.
 
-The resulting AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerable more complex than the filter.
+The AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerably more complex than the filter.
+BriefCASE generates everything but the statements in \linesref{line:mon-const}{line:mon-assume}.
+As with the filter, the guarantees completely define the output, as required for synthesis, in terms of the policy and the alert.
+The \texttt{alert} variable defined on \lineref{line:mon-alert} depends not just on the policy but also on the indicated behavior from the BriefCASE dialogue box. 
+Marking \texttt{is\_latched} as true makes the alert persistent, meaning that once the alert is raised, it is always raised; otherwise, it is the complement of the policy value in the current step.
 
-The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
-This behavior is one of the several options available in the dialogue.
-The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
-As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
+The monitor policy is defined by marking when a request is not satisfied, \texttt{isPending} on \lineref{line:mon-pending}, and counting the number of steps between requests, \texttt{latency} on \lineref{line:mon-latency}.
+The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response--\emph{not response since request and not response}.
+The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.
+
+The latency bound for the monitor is defined by the constant \texttt{MAX\_LATENCY} on \lineref{line:mon-const}.
+Here the constant is set to one to be consistent with the system specification from the previous section.
+At time 0 the policy is trivial: a response requires an accompanying request.
+After time 0, if a request is pending in the current time step, then the latency must still be under the bound, and if there is a response, then it closes a request now or a pending request from earlier in time.
+
+The policy definition only works if there is never more than one outstanding request at a time.
+That requirement is encoded in the assumption on \lineref{line:mon-assume}.
+A request is only allowed if one is not already pending in the pre-state of the component.
+The policy can be written without needing the assumption, but it is much simpler to write with the assumption.
+
+As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
+That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
+Synthesis must generate precise implementations of the high-assurance component specifications for the AGREE verification result to apply to the deployed system.
+Precise in this context means that the implementations match the input to output relations defined in the specifications.

### 2022-03-16 paper/paper.tex
Complete the AGREE explanations in section 3

@@
   morecomment=[l]{--},
+  escapeinside={*}{*},
   morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
@@
 \newcommand{\lineref}[1]{Line~\ref{#1}}
-\newcommand{\linesref}[2]{Lines~\ref{#1}-\ref{#2}}
+\newcommand{\linesref}[2]{Lines~\ref{#1}--\ref{#2}}
 
@@
 \begin{compactitem}
-  \item Add in the AGREE specification for the system
-  \item Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)
+  \item \sout{Add in the AGREE specification for the system}
+  \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
 \end{compactitem}

### 2022-03-17 paper/agree.tex
prep for pull on paper

--- 2022-03-14 paper/agree.tex
@@
 \begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree,numbers=left]
+\begin{lstlisting}[style=agree]
 eq req : bool = event(AutomationRequest);
@@
 eq wp : bool = event(Waypoint);
-eq strt: bool = event(Start);
+eq rsp: bool = event(Start);
 eq alrt : bool = event(Alert);
 
-assume "Automation requests are well-formed" :
+assume "Automation request is well-formed" :
     req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
-assume "Air vehicle locations are well-formed" :
+assume "Air vehicle location is well-formed" :
     avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);
-assume "One automation request in flight at a time" :
-  true -> 
-  (req => pre(Historically(not req) or Since(not req, strt)));
 
-guarantee "Waypoints coincide with air vehicle locations":
-  wp => avl;
-guarantee "Starts include a new waypoint" :
-  strt => wp;
-guarantee "Waypoints are well-formed" : 
+eq current : bool = (req = rsp);
+eq previous : bool = (req and not rsp) ->
+                      pre(req and not rsp) and (not req and rsp);
+eq policy : bool = current or previous;
+eq since : bool =  alrt or (alrt and (false -> pre(since)));
+
+guarantee "Start includes a waypoint" :
+    rsp => wp;
+guarantee "Locations required after the start waypoint" :
+    (wp and not rsp) => avl;
+guarantee "Waypoint is well-formed" :
     wp => WELL_FORMED_WAYPOINT(Waypoint);
-guarantee "Starts within one cycle of requests if not alerting" :
-  (strt => ((not alrt) and req)) -> 
-  (strt => ((not alrt) and (req or pre(req))));
-guarantee "Alert if not started within one cycle of requests" :
-    true -> ((pre(req and not strt) and not strt) => alrt);
-guarantee "Once alerted always alerted" :
-  not alrt or (Once(alrt) and alrt);
+guarantee "Alert if start is not bounded relative to a request" :
+    policy or since;
 \end{lstlisting}
@@
   \begin{center}
-    \scalebox{0.62}{\usebox{\sw}}
+    \scalebox{0.60}{\usebox{\sw}}
   \end{center}
@@
 \begin{lrbox}{\flt}
-\begin{lstlisting}[style=agree,numbers=left]
+\begin{lstlisting}[style=agree]
 eq policy : bool =
   WELL_FORMED_AUTOMATION_RESPONSE(Input);
-
-guarantee Filter_Output
-  "Filter output is well-formed" :
+guarantee Filter_Output "Filter output is well-formed" :
   if event(Input) and policy then
@@
 \begin{lrbox}{\mntr}
-\begin{lstlisting}[style=agree,numbers=left]
-const is_latched : bool = true;
-const MAX_LATENCY : int = 1;
-
+\begin{lstlisting}[style=agree]
+const is_latched : bool =
+  Get_Property(this, CASE_Properties::Monitor_Latched);
 eq rsp : bool = event(Response);
 eq req : bool = event(Request);
-
-eq isPending : bool = Since(not rsp, req and not rsp);
-eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);
-
-eq policy : bool = (rsp => req) ->
-                   (    (isPending => latency < MAX_LATENCY)   
-                    and (rsp => (req or pre(isPending))));
-eq alert : bool = (not policy) -> 
-                  ((is_latched and pre(alert)) or not policy);
-
-assume "One outstanding request at a time" :
-  (true -> (req => not pre(isPending)));
-
-guarantee "Alert port tracks alert variable" :
+eq current : bool = (req = rsp);
+eq previous : bool = (req and not rsp) ->
+                      pre(req and not rsp) and (not req and rsp);
+eq policy : bool = current or previous;
+eq alert : bool = (not policy)
+                -> ((is_latched and pre(alert)) or not policy);
+guarantee Monitor_Alert
+  "Alert port tracks alert variable" :
   event(Alert) = alert;
-guarantee "Output if not alerted" :
-  if (not(alert) and rsp) then
-    event(Output) and (Output = Response)
-  else
-    not (event(Output));    
+guarantee Monitor_Output
+  "Output if not alerted" :
+  if event(Alert) then (not event(Output)) else
+  if event(Response) then (event(Output) and (Output = Response))
+  else (not event(Output));
 \end{lstlisting}
@@
     \begin{tabular}{c}
-      \scalebox{0.62}{\usebox{\flt}} \\
+      \scalebox{0.60}{\usebox{\flt}} \\
+      (a) \\
+      \scalebox{0.60}{\usebox{\mntr}} \\
+      (b)
     \end{tabular}
   \end{center}
-  \caption{Contract specification for high-assurance filter.}
-  \label{fig:filter}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.62}{\usebox{\mntr}} \\
-    \end{tabular}
-  \end{center}
-  \caption{Contract specification for high-assurance monitor.}
-  \label{fig:monitor}
+  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
+  \label{fig:assurance}
 \end{figure}
@@
 
-The AGREE contract specification generated by the transform is shown in \figref{fig:filter}.
+The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a).
 The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
-The resulting AGREE specification for the monitor in this example is shown in \figref{fig:monitor}.
+The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).
 The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.

### 2022-03-17 paper/agree.tex

--- 2022-03-14 paper/agree.tex
@@
 
-\subsection{Contract Specification}
-\newsavebox{\sw}
-\begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree,numbers=left]
-eq req : bool = event(AutomationRequest);
-eq avl : bool = event(AirVehicleLocation);
-eq wp : bool = event(Waypoint);
-eq strt: bool = event(Start);
-eq alrt : bool = event(Alert);
-
-assume "Automation requests are well-formed" :
-  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
-assume "Air vehicle locations are well-formed" :
-  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
-assume "One automation request in flight at a time" :
-  true -> 
-  (req => pre(Historically(not req) or Since(not req, strt)));
-
-guarantee "Waypoints coincide with air vehicle locations":
-  wp => avl;
-guarantee "Starts include a new waypoint" :
-  strt => wp;
-guarantee "Waypoints are well-formed" : 
-  wp => WELL_FORMED_WAYPOINT(Waypoint);
-guarantee "Starts within one cycle of requests if not alerting" :
-  (strt => ((not alrt) and req)) -> 
-  (strt => ((not alrt) and (req or pre(req))));
-guarantee "Alert if not started within one cycle of requests" :
-    true -> ((pre(req and not strt) and not strt) => alrt);
-guarantee "Once alerted always alerted" :
-  not alrt or (Once(alrt) and alrt);
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}
-  \begin{center}
-    \scalebox{0.62}{\usebox{\sw}}
-  \end{center}
-  \caption{The SW component contract.}
-  \label{fig:sw}
-\end{figure}
-
-The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}.
-The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
-As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
-Contracts are evaluated in dependency order with inputs being propagated to outputs through all contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.\footnote{An apparent syntactic cycle, where a component is linked back to itself, may be broken temporally by inserting delay elements.}
-Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
-The semantics do not model computation or communication delay.
-The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
-
-The AGREE model checker attempts to prove several properties of the top-level model being verified.
-The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (\ie, the system composition meets input assumptions at each input as well as the guarantees on the system output).
-These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).
-The next set of properties prove that the contract specifications for each component are \emph{self-consistent} \ie, a contract does not contradict itself.
-These are the unexpanded results at the bottom of the figures.
-
-Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification.
-For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression.
-In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component.
-The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value.
-An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port.
-The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.
-
-REPLACE ALL: 
-The \emph{Alert if start is not bounded relative to a request} guarantee is an invariant on the expression \texttt{policy or since}, meaning that either the policy holds or the alert is sounding.
-The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}.
-The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.
-
-The value of \texttt{previous} in the current time step relies on values from the previous time step.
-The \texttt{->} operator designates initialization, as the previous time step is undefined in the first step of the system.
-The left operand to the operator is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system.
-The right operand is the value of \texttt{previous} after the initial step.
-Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.
-
-The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step.
-The intuitive reading of the expression is that the alert has been true since the time when it first sounded.
-The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds.
-The \emph{Alert if start is not bounded relative to a request} guarantee defines one requirement of a cyber-hardened system implementation.
-Together with the other guarantees, the contract models the expected input and output of the system as a whole.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \includegraphics[scale=0.3]{dialogue.png}
-    \end{tabular}
-  \end{center}
-  \caption{Wizard for automatically transforming the model with a filter.}
-  \label{fig:dialogue}
-\end{figure}
-
-\newsavebox{\flt}
-\begin{lrbox}{\flt}
-\begin{lstlisting}[style=agree,numbers=left]
-eq policy : bool = 
-  WELL_FORMED_AUTOMATION_RESPONSE(Input);
-
-guarantee Filter_Output
-  "Filter output is well-formed" :
-  if event(Input) and policy then 
-    event(Output) and Output = Input
-  else not event(Output);
-\end{lstlisting}
-\end{lrbox}
-
-\newsavebox{\mntr}
-\begin{lrbox}{\mntr}
-\begin{lstlisting}[style=agree,numbers=left]
-const is_latched : bool = true;
-const MAX_LATENCY : int = 1;
-
-eq rsp : bool = event(Response);
-eq req : bool = event(Request);
-
-eq isPending : bool = Since(not rsp, req and not rsp);
-eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);
-
-eq policy : bool = (rsp => req) ->
-                   (    (isPending => latency < MAX_LATENCY)   
-                    and (rsp => (req or pre(isPending))));
-eq alert : bool = (not policy) -> 
-                  ((is_latched and pre(alert)) or not policy);
-
-assume "One outstanding request at a time" :
-  (true -> (req => not pre(isPending)));
-
-guarantee "Alert port tracks alert variable" :
-  event(Alert) = alert;
-guarantee "Output if not alerted" :
-  if (not(alert) and rsp) then
-    event(Output) and (Output = Response)
-  else
-    not (event(Output));    
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.62}{\usebox{\flt}} \\
-    \end{tabular}
-  \end{center}
-  \caption{Contract specification for high-assurance filter.}
-  \label{fig:filter}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \scalebox{0.62}{\usebox{\mntr}} \\
-    \end{tabular}
-  \end{center}
-  \caption{Contract specification for high-assurance monitor.}
-  \label{fig:monitor}
-\end{figure}
-
-As noted previously, the original system fails to guarantee the cyber requirements.
-BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor.
-The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation.
-The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}.
-The policy of the high-assurance component can be stated directly in the wizard, or it can be left blank.
-In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}.
-Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.
-REMOVE: Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run at any time to determine whether those artifacts are valid.
-
-The AGREE contract specification generated by the transform is shown in \figref{fig:filter}.
-The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
-The resulting AGREE specification for the monitor in this example is shown in \figref{fig:monitor}.
-The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
-This behavior is one of the several options available in the dialogue.
-The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
-As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
+This section then introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
+It details the specification of that top-level system interface.
+It then discusses the specifications for the filter and the monitor.
+These two specifications are those that are synthesized to equivalent CakeML by SPLAT.
 
@@
 \input{verification-conditions}
+
+\subsection{System Contract}
+\input{sw-contract}
+
+\subsection{Cyber-hardening the System Implementation}
+\input{high-assurance-contracts}

### 2022-03-17 paper/sw-contract.tex

@@
 \begin{lrbox}{\sw}
-\begin{lstlisting}[style=agree,numbers=left,escapeinside={*}{*}]
+\begin{lstlisting}[style=agree,numbers=left]
 eq req : bool = event(AutomationRequest);*\label{line:sw-event-def-start}*
@@
 All the named ports in the corresponding AADL component are in the scope of the specification, and there are additional implicit boolean \emph{event} inputs (or outputs) associated with event ports.
-An \texttt{event} expression refers to that implicit input (or output) boolean value and is true when data present the named port and false otherwise.
+An \texttt{event} expression refers to that implicit input (or output) boolean value and is true when data is present on the named port and false otherwise.
 \linesref{line:sw-event-def-start}{line:sw-event-def-end} create local variables that are true when data is present on the corresponding event ports for the component.
@@
 A followed-by expression defines what happens at the first instance of the stream, time 0, and then what follows after. 
-Here the assumption is \texttt{true} at time 0, and then takes on the truth value of the implication in all future instances.
+Here the assumption is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
 The followed-by guards the \texttt{pre} because expressions are undefined before time 0 and the followed-by prevents the \texttt{pre} from being evaluated before time 1.
-So after the first instance, the assumption depends on the presence of a request and the pre-state of the component.
+So after time 0, the assumption depends on the presence of a request and the pre-state of the component.
 
@@
 The \texttt{Since} operator is true if the first expression has been true in all moments up to the present one since the second expression most recently became true.
-The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}.
+The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}--\emph{not request since start}.
 
@@
 The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
-The guarantee also does not force the start to take place, it only says that if it takes place, it is under the defined conditions.
+It differentiates with the followed-by operator what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincides with the request or is a response to a request one step earlier.
+The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.
 
@@
 The \texttt{Once} operator is true if the enclosed expression has ever been true before the current moment.
-\lineref{line:sw-guarantee-6} ensures that once alerted always alerted because either there has never been an alert, \texttt{\textbf{not} alrt}, or there has been an alert sometime in the past and so it is alerting in now, \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
+\lineref{line:sw-guarantee-6} ensures that once alerted always alerted because either there has never been an alert, \texttt{\textbf{not} alrt}, or there has been an alert sometime in the past and so it is alerting now, \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
 These six guarantees define the behavior of the SW component under the three assumptions.

### 2022-03-17 paper/verification-conditions.tex

\newcommand{\globally}{\ensuremath{\mathbf{G}}}
\newcommand{\historically}{\ensuremath{\mathbf{H}}}
\newcommand{\assumes}{\ensuremath{A}}
\newcommand{\guarantees}{\ensuremath{P}}
\newcommand{\dispatch}{\ensuremath{\mathit{dispatch}}}
\newcommand{\complete}{\ensuremath{\mathit{complete}}}
\newcommand{\same}[1]{\ensuremath{\mathit{same}(#1)}}
\newcommand{\inputs}{\ensuremath{I}}
\newcommand{\outputs}{\ensuremath{O}}
\newcommand{\system}{\ensuremath{S}}
\newcommand{\components}{\ensuremath{C}}
\newcommand{\component}{\ensuremath{c}}
\newcommand{\schedule}{\ensuremath{\phi}}
\newcommand{\valid}{\ensuremath{\mathit{valid}}}
\newcommand{\dpred}{\ensuremath{\delta^\phi}}
\newcommand{\dispred}{\ensuremath{\mathbb{D}^\phi}}
\newcommand{\compred}{\ensuremath{\mathbb{C}^\phi}}
\newcommand{\dispredp}{\ensuremath{\mathbb{D}^{\phi\prime}}}
\newcommand{\compredp}{\ensuremath{\mathbb{C}^{\phi\prime}}}

AGREE generates, and then proves or disproves with model checking, a set of verification conditions to show that a system's component implementation is correct~\cite{agree2013}. 
The verification conditions are formalized in \emph{past-time linear temporal logic} (PLTL).
PLTL is a first-order predicate calculus with temporal operators to reason about the truth values of formulas through time.
Its semantics are defined relative to a finite trace of system states $\pi = s_0, s_1, \ldots$, and a moment of evaluation in the trace, $i$.

The two PLTL operators necessary for the AGREE generated verification conditions are $\globally$ (globally) that looks forward in time along the trace and $\historically$ (historically) that looks backward in time along the trace.
These are defined as
\begin{eqnarray*}
 (\pi, i) \models \globally(f) & \iff & \forall j \ge i, (\pi, j) \models f \\
(\pi, i) \models \historically(f) & \iff & \forall 0 \le j \le i, (\pi, j) \models f
\end{eqnarray*}
The $\models$-operator is read as \emph{satisfies}.
A trace at a moment in time satisfies $\globally(f)$ if and only if it satisfies $f$ in the current and all future states of $\pi$.
Globally is invariant from the current moment into the future and historically is invariant from the beginning of the trace to the current moment.

A system is a tuple, $\system = (\inputs, \outputs, \assumes, \guarantees)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, and $\guarantees$ is the set of guarantees.
A system is implemented by a set of components, $\components$, with their various connections.
A component $\component$ is defined by its own tuple in a manner similar to the system as $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component)$.
From the components and connections, $\mathbb{I}_\component$ is defined such that it is the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined such that it is the set of components that provide the output for the system.

An system is correct if and only if for all $\pi$ and for all $i \ge 0$ the following holds:
\[
\begin{array}{lll}
        & \forall \component \in \components &  
            \globally(\historically(\assumes \wedge 
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime}) 
            \implies \assumes_\component) \\
 \wedge &   & 
            \globally(\historically(\assumes \wedge 
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime}) 
            \implies \guarantees)
\end{array}
\]
The first condition verifies the input assumptions on each component under the system assumptions and upstream component guarantees. 
It checks if the component guarantees and system assumptions are strong enough to imply input assumptions on all immediate downstream components.
The second condition checks the output guarantees of the system under the system assumptions and component guarantees that provide the output. 
It checks if the guarantees on components providing primary outputs are strong enough to imply the system guarantees.

AGREE uses $k$-inductive model checking to prove or disprove each generated verification condition.
If all the verification conditions hold, then the system is said to be \emph{correct}, meaning that the system composition meets input assumptions at each input as well as the guarantees on the system output. A consequence of this result is that $\globally(\historically(\assumes) \implies \guarantees)$ holds for the system contract. As such, the system contract is a sound abstraction of the system implementation.

The expanded property lists in \figref{fig:example-certificate} and \figref{fig:hardened-certificate} are the results from disproving or proving out the above verification conditions.
The additional unexpanded results at the bottom of the figures prove \emph{self-consistency} in the contracts.
It is not uncommon to accidentally write contracts that are self-contradicting.
For example, a contract may guarantee an output be two different values in the same moment of time.
AGREE generates additional verification conditions that prove each component contract, and the composition of contracts, self-consistent.

### 2022-03-17 paper/intro.tex
work on the acronum paragraph so that it typesets better

@@
 
-BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models for potential cyber vulnerabilities and suggest cyber-security requirements to mitigate them.
-A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system.
-Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
-Formal verification that the transformed system model satisfies its cyber requirements is accomplished via the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.
+BriefCASE incorporates model-level cyber-analysis tools (presently
+GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
+examine AADL models for potential vulnerabilities and suggest
+cyber-security requirements to mitigate them.  A library of
+architectural transforms guides the system engineer through automated
+model transformations that modify the architecture to address these
+requirements, possibly inserting new high-assurance components into
+the system.  Implementations for the new components are synthesized
+from formal specifications using
+SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
+(Semantic Properties for Language and Automata Theory).  Formal
+verification that the transformed system model meets its requirements
+is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
+Environment).
 %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
-AGREE is a {\em compositional assume-guarantee} style model checker for AADL models that attempts to prove properties about one layer of the architecture using
-properties allocated to its subcomponents.
-Cyber-resilient code implementing the verified model is then automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.
-If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.
+AGREE is a {\em compositional assume-guarantee} style model checker
+for AADL models that attempts to prove properties about one layer of
+an architecture using properties allocated to its subcomponents.
+Cyber-resilient code implementing the verified model is then
+automatically generated using the High Assurance Modeling and Rapid
+Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
+desired, this code can be targeted to the formally verified seL4
+secure microkernel~\cite{sel4-2009}.
 

### 2022-03-17 paper/high-assurance-contracts.tex
Update example contracts

@@
     -- start user definitions
+    assume "One automation request in flight at a time" : *\label{line:mon-assume}*
+      true -> (req => pre(Historically(not req) or Since(not req, rsp)));
+
     const MAX_LATENCY : int = 1; *\label{line:mon-const}*
@@
                           and (rsp => (req or pre(isPending))));
-
-    assume "One outstanding request at a time" : *\label{line:mon-assume}*
-      (true -> (req => not pre(isPending))); 
     -- stop user definitions
@@
 The AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerably more complex than the filter.
-BriefCASE generates everything but the statements in \linesref{line:mon-const}{line:mon-assume}.
+BriefCASE generates everything but the statements in \linesref{line:mon-assume}{line:mon-policy}.
 As with the filter, the guarantees completely define the output, as required for synthesis, in terms of the policy and the alert.
@@
 
-The policy definition only works if there is never more than one outstanding request at a time.
-That requirement is encoded in the assumption on \lineref{line:mon-assume}.
-A request is only allowed if one is not already pending in the pre-state of the component.
+The policy definition only works if there is never more than one outstanding request at a time; otherwise, the latency counter resets incorrectly.
+That requirement is encoded in the assumption on \lineref{line:mon-assume} and is the same assumption used for the system input.
 The policy can be written without needing the assumption, but it is much simpler to write with the assumption.

### 2022-03-17 paper/sw-contract.tex
Update example contracts

@@
 guarantee "Once alerted always alerted" : *\label{line:sw-guarantee-6}*
-  not alrt or (Once(alrt) and alrt);
+  true -> (pre(alrt) => alrt);
 \end{lstlisting}
@@
 Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
-The final guarantee on \lineref{line:sw-guarantee-6} defines additional behavior for the alert.
-
-The \texttt{Once} operator is true if the enclosed expression has ever been true before the current moment.
-\lineref{line:sw-guarantee-6} ensures that once alerted always alerted because either there has never been an alert, \texttt{\textbf{not} alrt}, or there has been an alert sometime in the past and so it is alerting now, \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
+\lineref{line:sw-guarantee-6} ensures that once alerted always alerted since an alert in the pre-state requires an alert in the post-state., \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
 These six guarantees define the behavior of the SW component under the three assumptions.

### 2022-03-17 paper/intro.tex
more wordsmithing

@@
 
-The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect (and alert) unexpected behaviors arising from untrusted components.
-These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
-Those specifications are sufficient for model checking to prove that with the incorporation of these high-assurance components, the hardened system meets its cyber-resiliency requirements.
+The following security-enhancing transformations are discussed in the
+remainder of the paper: (1) the insertion of a filter to prevent
+malformed data from a malicious actor from being propagated to
+downstream components, and (2) the insertion of a monitor to detect
+(and alert) unexpected behaviors arising from untrusted
+components. These transformations not only change the architecture of
+the model by adding in new components; they also generate a formal
+specification for the behavior of the inserted high-assurance
+components in the AGREE language. Those specifications are sufficient
+for model checking to prove that---due to the incorporation of these
+high-assurance components---the hardened system meets its
+cyber-resiliency requirements.
 

### 2022-03-17 paper/sw-contract.tex
Update text in paper to match contract changes.

@@
 guarantee "Once alerted always alerted" : *\label{line:sw-guarantee-6}*
-  true -> (pre(alrt) => alrt);
+  Once(alrt) => alrt;
 \end{lstlisting}
@@
 Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
-\lineref{line:sw-guarantee-6} ensures that once alerted always alerted since an alert in the pre-state requires an alert in the post-state., \texttt{\textbf{Once}(alrt) \textbf{and} alrt}.
+The last guarantee ensures that alert is persistent.
+The \texttt{Once} operator is true if the enclosed expression has ever been true before the current moment.
+\lineref{line:sw-guarantee-6} ensures if alert has ever happened in the past, then it is happening in the present moment.
 These six guarantees define the behavior of the SW component under the three assumptions.

### 2022-03-18 paper/high-assurance-contracts.tex
Update text in paper to match contract changes.
Reformat filter to be consistent with other contracts
Expunging dialogue diagram

@@
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-    \includegraphics[scale=0.3]{dialogue.png}
-    \end{tabular}
-  \end{center}
-  \caption{BriefCASE dialogue for filter transformation.}
-  \label{fig:dialogue}
-\end{figure}
+%% \begin{figure}
+%%   \begin{center}
+%%     \begin{tabular}{c}
+%%     \includegraphics[scale=0.3]{dialogue.png}
+%%     \end{tabular}
+%%   \end{center}
+%%   \caption{BriefCASE dialogue for filter transformation.}
+%%   \label{fig:dialogue}
+%% \end{figure}
 
@@
 
-    guarantee Filter_Output
-    "Filter output is well-formed" :
+    guarantee "Filter output is well-formed" :
       if event(Input) and policy then
         event(Output) and Output = Input
-    else not event(Output);
+      else
+        not event(Output);
   \end{lstlisting}
@@
 
-    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
+    eq policy : bool = (rsp => req) ->
                          (    (isPending => latency < MAX_LATENCY)
                           and (rsp => (req or pre(isPending))));
+    *\label{line:mon-policy}*
     -- stop user definitions
@@
 
-As noted previously, the system with the untrusted AI does not pass AGREE verification.
-Transformations in BriefCASE are used to cyber-harden the system against malicious AI behavior by inserting a high-assurance filter and monitor between the AI and the WM as shown in \figref{fig:hardened}.
-High-assurance means the behaviors of the components are verified.
-Here the filter protects the WM from malformed data and the monitor protects it from spontaneous or delayed responses.
-These high-assurance components are added one at a time by selecting the connection in the model where the component is to appear and then choosing the appropriate transformation.
+As noted previously, the system with the untrusted AI does not pass
+AGREE verification.  Transformations in BriefCASE cyber-harden the
+system against malicious AI behavior by inserting a high-assurance
+filter and monitor between the AI and the WM as shown in
+\figref{fig:hardened}.  (In our context \emph{high-assurance} means
+that the behaviors of the components are verified to meet their
+contracts.)  In the example, the inserted filter protects the WM from
+malformed data and the monitor protects the WM from (possibly malicious)
+spontaneous or delayed responses.  These high-assurance components are
+added, one at a time, by
+\begin{itemize}
+  \item selecting the connection in the model where the
+    component is to appear,
+  \item choosing the appropriate transformation, and
+  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
+\end{itemize}
 
-The system designer provides configuration parameters for the transformation in a dialogue box.
-The dialogue box for adding a filter is shown in \figref{fig:dialogue}.
-All high-assurance components rely on a \emph{policy} to define behavior as seen in the last field of the dialogue box.
-A filter policy defines well-formed data while a monitor policy defines an invariant over inputs and outputs.
-These policies can be stated directly in the wizard, or they can be left blank and added later to the AGREE contract generated by the transformation.
+%% The system designer provides configuration parameters for the
+%% transformation in a dialogue box.  The dialogue box for adding a
+%% filter is shown in \figref{fig:dialogue}.  All high-assurance
+%% components rely on a \emph{policy} to define behavior as seen in the
+%% last field of the dialogue box.  A filter policy defines well-formed
+%% data while a monitor policy defines an invariant over inputs and
+%% outputs.  These policies can be stated directly in the wizard, or they
+%% can be left blank and added later to the AGREE contract generated by
+%% the transformation.
 
-BriefCASE creates all the needed AADL for the new high-assurance component and its connections in the system implementation as part of the transformation.
-It also creates a default AGREE specification that defines everything about the component behavior but the policy itself unless it is provided in the dialogue.
-The resulting AGREE specification from the transformation for the filter is shown in \figref{fig:filter}.
-The single guarantee completely defines the filter output under any input according to the truth value of the policy.
-While such a strong guarantee is not required for AGREE analysis, it is required for automated synthesis.
+BriefCASE creates all the needed AADL for the new high-assurance
+component and its connections in the system implementation as part of
+the transformation.  It also creates a default AGREE specification
+that defines everything about the component behavior but the policy
+itself, unless that is provided in the dialogue.  The resulting AGREE
+specification from the transformation for the filter is shown in
+\figref{fig:filter}.  The single guarantee completely defines the
+filter output under any input according to the truth value of the
+policy.  Such a strong guarantee is needed for synthesis (a weaker
+version suffices for AGREE analysis).
 
-The AGREE specification for the monitor is shown in \figref{fig:monitor} and is considerably more complex than the filter.
-BriefCASE generates everything but the statements in \linesref{line:mon-assume}{line:mon-policy}.
-As with the filter, the guarantees completely define the output, as required for synthesis, in terms of the policy and the alert.
-The \texttt{alert} variable defined on \lineref{line:mon-alert} depends not just on the policy but also on the indicated behavior from the BriefCASE dialogue box. 
-Marking \texttt{is\_latched} as true makes the alert persistent, meaning that once the alert is raised, it is always raised; otherwise, it is the complement of the policy value in the current step.
+The AGREE specification for the monitor is shown in
+\figref{fig:monitor}. As can be seen, it is considerably more complex
+than the filter specification.  BriefCASE generates everything but the
+statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
+the filter, the guarantees completely define the output, as required
+for synthesis, in terms of the policy and the alert.  The
+\texttt{alert} variable defined on \lineref{line:mon-alert} depends
+not just on the policy but also on configuration details supplied when
+the component is specified.  Setting \texttt{is\_latched} as true
+makes the alert persistent, meaning that once the alert is raised, it
+is always raised; otherwise, it is the complement of the policy value
+in the current step.
 
-The monitor policy is defined by marking when a request is not satisfied, \texttt{isPending} on \lineref{line:mon-pending}, and counting the number of steps between requests, \texttt{latency} on \lineref{line:mon-latency}.
-The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response--\emph{not response since request and not response}.
-The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.
+The monitor policy is defined by marking when a request is not
+satisfied, \texttt{isPending} on \lineref{line:mon-pending}, and
+counting the number of steps between requests, \texttt{latency} on
+\lineref{line:mon-latency}.  The \texttt{isPending} variable is true
+whenever there is a request that does not coincide with a
+response--\emph{not response since request and not response}.  The
+\texttt{latency} variable starts at zero, resets on every request, and
+otherwise increments by one.
 
-The latency bound for the monitor is defined by the constant \texttt{MAX\_LATENCY} on \lineref{line:mon-const}.
-Here the constant is set to one to be consistent with the system specification from the previous section.
-At time 0 the policy is trivial: a response requires an accompanying request.
-After time 0, if a request is pending in the current time step, then the latency must still be under the bound, and if there is a response, then it closes a request now or a pending request from earlier in time.
+The latency bound for the monitor is defined by the constant
+\texttt{MAX\_LATENCY} on \lineref{line:mon-const}.  Here the constant
+is set to one to be consistent with the system specification from the
+previous section.  At time 0 the policy is trivial: a response
+requires an accompanying request.  After time 0, if a request is
+pending in the current time step, then the latency must still be under
+the bound, and if there is a response, then it closes a request now or
+a pending request from earlier in time.
 
-The policy definition only works if there is never more than one outstanding request at a time; otherwise, the latency counter resets incorrectly.
-That requirement is encoded in the assumption on \lineref{line:mon-assume} and is the same assumption used for the system input.
-The policy can be written without needing the assumption, but it is much simpler to write with the assumption.
+The policy definition only works if there is never more than one
+outstanding request at a time; otherwise, the latency counter resets
+incorrectly.  That requirement is encoded in the assumption on
+\lineref{line:mon-assume} and is the same assumption used for the
+system input.  The policy can be written without needing the
+assumption, but it is much simpler to write with the assumption.
 
-As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
-That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
-Synthesis must generate precise implementations of the high-assurance component specifications for the AGREE verification result to apply to the deployed system.
-Precise in this context means that the implementations match the input to output relations defined in the specifications.
+As shown in \figref{fig:hardened-certificate}, the cyber-hardened
+system passes AGREE verification.  That means that the high-assurance
+components protect the WM from the untrusted AI generating malformed
+data, spontaneous responses, or delayed responses.  Synthesis must
+generate precise implementations of the high-assurance component
+specifications for the AGREE verification result to apply to the
+deployed system.  Precise in this context means that the
+implementations match the input to output relations defined in the
+specifications.

### 2022-03-19 paper/agree.tex
revised content added to semantics section. Checkpointing

@@
-This section formalizes AGREE's assume-guarantee reasoning by defining the set of verification conditions that AGREE must check.
-The verification conditions succinctly summarize what in means when AGREE proves that a system implementation satisfies it specification.
-These conditions are useful in understanding how the semantics in synthesis differ from those in the assume-guarantee reasoning in \secref{sec:semantics}.
+This section gives an overview of AGREE's assume-guarantee reasoning
+system, abstractly defining the verification conditions that AGREE
+checks in order to establish overall system correctness at the model
+level.  The discussion then becomes more concrete, showing how the
+AGREE contract specification language is used to specify the
+cyber-hardened system of \figref{fig:hardened}. It explains the the
+top-level specification of the example system, then shifts to the
+specifications for the filter and the monitor.  These two
+specifications are those that are synthesized to equivalent CakeML by
+SPLAT.
 
-This section then introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
-It details the specification of that top-level system interface.
-It then discusses the specifications for the filter and the monitor.
-These two specifications are those that are synthesized to equivalent CakeML by SPLAT.
+%% These conditions are useful in understanding how the semantics in
+%% synthesis differ from those in the assume-guarantee reasoning
+%% in \secref{sec:semantics}.
 

### 2022-03-19 paper/high-assurance-contracts.tex
revised content added to semantics section. Checkpointing

@@
 As noted previously, the system with the untrusted AI does not pass
-AGREE verification.  Transformations in BriefCASE cyber-harden the
-system against malicious AI behavior by inserting a high-assurance
-filter and monitor between the AI and the WM as shown in
+AGREE verification.  Transformations carried out in the BriefCASE
+interface cyber-harden the system against malicious AI behavior by
+inserting a high-assurance filter and monitor between the AI and the
+WM as shown in
 \figref{fig:hardened}.  (In our context \emph{high-assurance} means

### 2022-03-19 paper/semantics.tex
revised content added to semantics section. Checkpointing

@@
-Brief, high-level overview, of the AGREE semantics for component synthesis.
-These semantics are those defined in HOL.
-The component is defined by its state.
-The state is its internal state and current output.
-Input streams are constrained to obey the assumptions.
-At each input, the \textbf{eq}-statements are evaluated, in order, to update the internal state and outputs.
-After the \textbf{eq}-statements are evaluated, guarantees are checked.
-The section should define clearly what it means for a component to be \emph{well-formed}. 
-The section should conclude by stating the \emph{Component Correct} definition.
+In this section we give an overview of the AGREE semantics, a
+machine-checked formal definition of the meaning of AGREE expressions,
+formulas, and contracts. The semantics supports a formal connection
+between the results of the AGREE model checker and code generated from
+specifications of leaf level cyber-components. It also provides the root
+meaning at the base of a chain of translation steps moving from AGREE
+specifications to CakeML executables.
 
-The correctness, and structure, of the synthesized CakeML code is based on these semantics.
-
-\subsection{Old AGREE Semantics}
+\subsection{AGREE syntax and semantics}
 \label{agree-semantics}
 
-Here the formal semantics of the AGREE contract specification are
-briefly presented to make clear the meaning of a high-assurance
-component.  These semantics are the basis of proofs showing that a
-synthesized component has input to output behavior specified by its
-contract.
+The syntax of AGREE is essentially that of quantifier-free first order
+logic (FOL) supplemented with a few temporal operators. The terms
+(\emph{e}) are the usual arithmetic expressions built from variables
+and numeric literals, while formulas (\emph{b}) are built using the
+standard logical connectives from atomic formulas (\emph{a}) based on
+the usual comparison operators.
+\[
+\begin{array}{rcl}
+e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
+a & ::= & e\; \set{=,<,\leq}\; e \\
+b & ::= & v \mid c \mid a \mid \neg b
+            \mid b \; \set{\land,\lor,\imp,\iff}\; b
+\end{array}
+\]
 
-Assume that all data flowing between components is in the form of byte
-arrays, thus input and output is modelled at the level of what is
-passed ``over the wire'' by the communication fabric. An \emph{environment},
-$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
-strings where an L-value is anything that can appear on the left-hand
-side of an assignment such as a named port, a field in a record, a
-entry in an array, a local value defined by an eq-statement etc.
+There is a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and
+formulas. There are also temporal operators $\konst{pre}(-)$, delay
+$(-) \to (-)$, and $\konst{Historically}(-)$.
 
-Let $s$ be an AGREE contract specification for some high-assurance
-component. The notation $\theta_s$ is used to denote the environment that
-contains a binding for any L-value in the scope of $s$ and nothing else, and
-the notation $\Theta_s$ denotes the universe of all such environments. The
-semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
-are finite sequences of environments, $\pi \in \Theta_s^*$.
+The semantics of terms and formulas is in terms of \emph{streams of
+values}. Values encompass at least booleans and numbers, but can be
+readily extended to include records and arrays. A value stream is a
+total function from natural numbers to values:
+\[
+ \konst{stream} = \mathbb{N} \to \konst{value}
+\]
+Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
+binding variable names to value streams, the semantics $\sem{-}$ of terms and
+formulas, omitting temporal operators, is standard.
 
-%% \begin{comment}
-%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
-%% generate a contig type for the inputs plus state vars.} ]
-
-%% [Question: what is a \emph{total} environment? One that supplies at
-%% least the bindings in $\theta$ needed to evaluate the spec?]
-%% \end{comment}
-
-The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
-$\pi$, and returns \konst{true} if $s$ is invariant along the entire
-stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
-is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
-while an eq-statement in $s$ is invariant if its binding in the
-context of every step is equivalent to the computed value of its
-associated expression in that same context. All guarantees and
-eq-statements must be invariant in the stream for the function to
-return \konst{true}.
-
-The meaning of an AGREE contract specification is now defined as the
-set of environment streams on which it is invariant.
-%
 \[
-  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
+\begin{array}{rcl}
+\sem{v}_t & = & E(v)(t) \\
+\sem{c}_t & = & c \\
+\sem{e_1 + e_2}_t & = & \sem{e_1}_t + \sem{e_2}_t \\
+   & \cdots & \\
+\sem{b_1 \land b_2}_t & = & \sem{b_1}_t \land \sem{b_2}_t \\
+   & \cdots & \\
+\end{array}
 \]
-%
-\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
-the L-values in the eq-statements in a way that is consistent with
-their associated expressions and the guarantees are all true in that
-same step.
-
-We claim that a synthesized high-assurance component preserves the
-input to output relationship in the specification $s$ over every
-stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
-s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
-synthesized component on the stream $\pi$ to create a new stream
-$\pi^\prime$ containing added output and other bindings. We say that
-two streams are equivalent in regards to a specification, denoted as
-$\pi =_s \pi^\prime$, if and only if the two streams are the same
-length and agree on bindings for the input and output for $s$ at every
-step of the streams.  We now formally state the correctness claim for
-synthesis.
-%
-\[
-  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
-\]
-%
-where $I_s(\pi)$ returns the corresponding stream that only retains
-bindings for inputs in each step and nothing else. The claim is that
-the synthesized component generates the same output stream as the
-specification for any stream belonging to the specification that is
-restricted to just input bindings at each step.

### 2022-03-19 paper/intro.tex
wordsmithing on intro

--- 2022-03-17 paper/intro.tex
@@
 
-The DARPA Cyber Assured Systems Engineering (CASE) program is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).
+The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).
 The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S.
@@
 
-BriefCASE incorporates model-level cyber-analysis tools (presently
+BriefCASE incorporates model-level cyber analysis tools (presently
 GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
@@
 
-The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect (and alert) unexpected behaviors arising from untrusted components.
-These transformations not only change the architecture of the model by adding in new components; they also generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language.
-Those specifications are sufficient for model checking to prove that with the incorporation of these high-assurance components, the hardened system meets its cyber-resiliency requirements.
+Two security-enhancing transformations are discussed in the remainder
+of the paper: (1) the insertion of a filter to prevent malformed data
+from a malicious actor from being propagated to downstream components,
+and (2) the insertion of a monitor to detect (and alert) unexpected
+behaviors arising from untrusted components. These transformations not
+only change the architecture of the model by adding in new components;
+they also generate a formal specification for the behavior of the
+inserted high-assurance components in the AGREE language. Those
+specifications are sufficient for model checking to prove that---due
+to the newly included high-assurance components---the hardened
+system meets its cyber-resiliency requirements.
 
-Another novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, a verified compiler implementation for the functional programming language ML \cite{cakeml}.
-This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness.
-CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML source code is exactly preserved in the final binaries.
-Assuming that the execution schedule of the deployed cyber-hardened system is as intended by the AADL model, and that the HAMR-generated communication fabric delivers messages between components as expected, the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs.
-Work is ongoing to lift this result to infinite input traces as these systems are inherently reactive and intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+Another novel aspect of the approach is the synthesis of the AGREE
+specifications for the high-assurance components to CakeML, a verified
+compiler implementation for the functional programming language
+ML \cite{cakeml}. This paper describes in detail the synthesis path
+from specifications to CakeML code, providing a formal framework in
+which to argue correctness. CakeML then provides a verified
+compilation path to several different target binaries (and also
+proving that the meaning of the CakeML source code is exactly
+preserved in the final binaries). Assuming that the execution schedule
+of the deployed cyber-hardened system is as intended by the AADL
+model, and that the HAMR-generated communication fabric delivers
+messages between components as expected, the AGREE model checking
+results then hold for the deployed system, \ie, the system will detect
+and prevent the indicated cyber-vulnerabilities over all possible
+finite inputs. Preliminary work has shown how to lift this result to
+infinite input traces as these systems are inherently reactive and
+intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
 
-CASE was scoped to address the more common cyber-vulnerabilities that programmers introduce, such as overflow, lack of input validation, supply-chain issues, and identity verification; but not other cyber-vulnerabilities such as side-channel attacks and denial of service.
-In this program, we have not invented a new type of filter or monitor in terms of capability, instead our contribution is in the automated synthesis of these components from a formal specification, and a means to show that the synthesis is correct.
+CASE was scoped to address common cyber-vulnerabilities, such as
+overflow, lack of input validation, supply-chain issues, and identity
+verification; however, other cyber-vulnerabilities such as
+side-channel attacks and denial of service are yet dealt with in our
+work.  Research in CASE has not invented a new type of filter or
+monitor, in terms of capability; instead, our contribution is in the
+automated synthesis of these components from a formal specification,
+and a means to show that the synthesis is correct.
 
-The approach is motivated, and illustrated, in a simple example in Section~\ref{sec:example}.
-Contract specification in the AGREE language is presented in Section~\ref{sec:agree}, followed by a description of the synthesis pathway in Section~\ref{sec:synthesis}.
-A case study applying these transformations to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented in Section~\ref{sec:case-study}.
-Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
-The case study system is significantly more complex than the simple example and shows the viability of the modeling approach to a full-scale industrial design.
+The approach is illustrated by a simple example in
+Section~\ref{sec:example}. Contract specification in the AGREE
+language is presented in Section~\ref{sec:agree}, followed by a
+description of the synthesis pathway in Section~\ref{sec:synthesis}. A
+case study applying these transformations to an Unmanned Aerial
+Vehicle (UAV) system that uses the Air Force Research Laboratory's
+OpenUxAS services for route planning is presented in
+Section~\ref{sec:case-study}. Here the transforms add filters to guard
+against malformed input and monitors to guard against ground station
+spoofing and malicious flight plans from OpenUxAS. The case study
+system is significantly more complex than the simple example and shows
+the viability of the modeling approach to a full-scale industrial
+design.
 
-The BriefCASE tools are open source and publicly available \cite{fmide}, as is the motivating example \cite{repo}, and the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
-Videos demonstrating the use of the BriefCASE tools to build the UAV example presented in Section~\ref{sec:case-study} are also available \cite{case}.
+The BriefCASE tools are open source and publicly
+available \cite{fmide}, as is the motivating example \cite{repo}, and
+the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
+Videos demonstrating the use of the BriefCASE tools to build the UAV
+example presented in Section~\ref{sec:case-study} are also
+available \cite{case}.

### 2022-03-19 paper/intro.tex
checkpointing; outlined remainder of semantics section

@@
 
-Two security-enhancing transformations are discussed in the remainder
-of the paper: (1) the insertion of a filter to prevent malformed data
-from a malicious actor from being propagated to downstream components,
-and (2) the insertion of a monitor to detect (and alert) unexpected
-behaviors arising from untrusted components. These transformations not
-only change the architecture of the model by adding in new components;
-they also generate a formal specification for the behavior of the
-inserted high-assurance components in the AGREE language. Those
-specifications are sufficient for model checking to prove that---due
-to the newly included high-assurance components---the hardened
-system meets its cyber-resiliency requirements.
+Two security-enhancing transformations are discussed: (1) the
+insertion of a filter to prevent malformed data from a malicious actor
+from being propagated to downstream components, and (2) the insertion
+of a monitor to detect (and alert) unexpected behaviors arising from
+untrusted components. These transformations not only change the
+architecture of the model by adding in new components; they also
+generate a formal specification for the behavior of the inserted
+high-assurance components in the AGREE language. Those specifications
+are sufficient for model checking to prove that---due to the newly
+included high-assurance components---the hardened system meets its
+cyber-resiliency requirements.
 

### 2022-03-19 paper/semantics.tex
checkpointing; outlined remainder of semantics section

@@
 formulas. There are also temporal operators $\konst{pre}(-)$, delay
-$(-) \to (-)$, and $\konst{Historically}(-)$.
+$(-) \to (-)$, and $\konst{Hist}(-)$.
 
@@
 \]
+
+The temporal operators deal with time in more significant ways:
+\[
+\begin{array}{rcl}
+\sem{\konst{pre}(e)}_t & = & \sem{e}_{t-1}, \mathrm{when}\ t > 0 \\
+\sem{e_1 \to e_2}_t & = & \itelse{t=0}{\sem{e_1}_0}{\sem{e_2}_t} \\
+\sem{\konst{Hist}(b)}_t & = & \forall n \leq t.\; \sem{b}_n
+\end{array}
+\]
+
+\subsubsection{Extracting computation from specification}
+
+A subtle aspect of our approach is that AGREE does not explicitly
+present the computation that a component performs. This is entirely
+appropriate since AGREE is intended to reason about components solely
+at the specification level. However, specifications do support state
+variables, and updates to them; consequently---for inserted leaf-level
+security components---SPLAT interprets sequences of variable
+assignments as specifying the computation of the component.
+
+\begin{itemize}
+\item \textit{Point to discussion on output code guarantees}
+
+\item \textit{code contract representation}
+
+\item \textit{strmStep}
+
+\item \textit{Contract semantics}
+
+\item \textit{How example looks}
+\end{itemize}

### 2022-03-19 paper/sw-contract.tex
checkpointing; outlined remainder of semantics section

@@
 eq strt: bool = event(Start);
-eq alrt : bool = event(Alert);*\label{line:sw-event-def-end}*
+eq alrt: bool = event(Alert);*\label{line:sw-event-def-end}*
 

### 2022-03-19 paper/paper.tex
revised content added to semantics section. Checkpointing
checkpointing; outlined remainder of semantics section
Add section on testing

@@
   escapeinside={*}{*},
-  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event}
+  morekeywords={eq, bool, guarantee, assume, true, false, pre, not, and, or, property, const, Historically, Since, Once, event, annex, Implementation, process, subcomponents}
 }
@@
 \newcommand{\imp}{\Rightarrow}
-\newcommand{\lval}{\ensuremath{\mathit{lval}}}
 \newcommand{\set}[1]{\ensuremath{\{ {#1} \}}}
+\newcommand{\sem}[1]{\ensuremath{[\![ #1 ]\!]}}
 \newcommand{\kstar}[1]{\ensuremath{{#1}^{*}}}
 \newcommand{\Lang}[1]{\ensuremath{{\mathcal L}({#1})}}
-\newcommand{\LangTheta}[1]{\ensuremath{{\mathcal L}_{\theta}({#1})}}
 \newcommand{\itelse}[3]{\mbox{$\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}$}}
@@
 
-\section{Illustrative Example}
+\section{Illustrative example}
 \label{sec:example}
@@
 
-\section{Assume-Guarantee Reasoning in AGREE}
+\section{Assume-guarantee reasoning in AGREE}
 \label{sec:agree}
@@
 
-\section{AGREE Semantics for Component Synthesis}
+\section{AGREE semantics for component synthesis}
 \label{sec:semantics}
@@
 
-\section{Synthesis to CakeML and Proof Sketch}
+\section{Synthesis to CakeML}
 \label{sec:synthesis}
@@
 
-\section{Case Studies}
+\section{Testing Specifications with AGREE}
+\label{sec:testing}
+\input{testing}
+
+\section{Case studies}
 \label{sec:case-study}
@@
 
-\section{Related Work}
+\section{Related work}
 \label{sec:related-work}

### 2022-03-19 paper/testing.tex
Add section on testing

Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.
It is difficult to reason about the whole of the specification to understand the defined behavior of a component or system.
A natural consequence of this inherent complexity is that the AGREE analysis proves out, but that proof is about a flawed specification.
The AGREE analysis, and synthesized component, is only as good as the specifications.
As the saying goes, \emph{garbage in means garbage out}.

The case studies found four common mistakes in specifications: contradictions, vacuity, under-specification, and missed corner cases.
Contradictions arise from inconsistent contracts that state conflicting constraints in any given moment invalidating the verification results: a property can be proved true and it can be proved false.
Vacuity is a common pitfall with implication when the left hand side is always false making the implication true. 
Vacuity is referred to as the \emph{block of wood} problem because a system that does nothing passes verification with implications.

Under-specification is very challenging because verification is not able to prove anything (or it can prove everything in the case of existential properties) so nothing is actually known after verification except that the system is under-specified. 
The tell-tale symptom of underspecification is that AGREE verification is not able to prove anything about a property: it is not true and it is not false.
The property stated in the positive fails and the property stated in the negative fails.
Missed corner cases is a common problem to all disciplines.
It is especially hard here because there is no easy way break down, and check, specifications in a case by case input analysis.

AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification, and missed corner cases is always an issue.
The case studies navigated these challenges by writing a specification, synthesizing it to code with SPLAT, and then running unit 
tests on the synthesized code in a testing harness.
The synthesized code is known to precisely capture the meaning of the specification, so the tests directly reflect the behavior of the specification in much the same way as compiled code is assumed correct.
The designers are then able to iteratively test and refine the specifications into a final form.
The process is inefficient, requires an additional, unneeded, test harness, the code is proved correct by SPLAT after all, and steps outside the MBE environment using the implementation, rather than the model, to analyze and understand the system model.

There needs to be a bridge between the formal model in AGREE with its proof of system properties and more traditional reasoning with test that is a well established method for building assurance.
Classical black-box and white-box test coverage should be, and are, equally applicable to formal reasoning as traditional engineering in regards to developing, and eventually proving, the correctness of a specification.
AGREE is able to provide such a testing framework.

Unit tests are created in AGREE by writing a system specification that assumes the test input and then guarantees the test output.
The system is then implemented with the component under test and its specification.
The AGREE analysis then proves if the component guarantees, given the test inputs, are strong enough to prove the guaranteed test outputs.
Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
AGREE proves the component implements the test, and the Liskhov substitution principle holds.

\newsavebox{\tst}
\begin{lrbox}{\tst}
  \begin{lstlisting}[style=agree,numbers=left]
    process should_notAlertAndOutput_when_responseOneStepAfterRequest
      ... -- Omitted AADL features
      annex agree {**
        eq index : int = prev(index + 1, 0); *\label{line:tst-index}*

        assume "One Response one step after Request" : *\label{line:tst-assume-1}*
                ((index = 0) => not event(Response))
            and ((index = 1) => event(Response))
            and ((index >= 2) => not event(Response));
        assume "One Request" : *\label{line:tst-assume-2}*
            (event(Request) = true) ->
            (event(Request) = false);

        guarantee "Not Alert" : *\label{line:tst-guarantee-1}*
            not event(Alert);
        guarantee *\label{line:tst-guarantee-2}*
          "Output one step after Request at same time as Response" :
                ((index = 0) => not event(Output))
            and ((index = 1) => (event(Output) and Output = Response))
            and ((index >= 2) => not event(Output));
      **};
    end should_notAlertAndOutput_when_responseOneStepAfterRequest;

    process Implementation *\label{line:tst-imp-start}*
      should_notAlertAndOutput_when_responseOneStepAfterRequest.test
      subcomponents
        Monitor: thread CASE_Monitor_Thr.Impl; *\label{line:tst-imp-comp}*
      connections
        c00: port Response -> Monitor.Response;
        c01: port Request -> Monitor.Request;
        c02: port Monitor.Alert -> Alert;
        c03: port Monitor.Output -> Output;
    end should_notAlertAndOutput_when_responseOneStepAfterRequest.test; *\label{line:tst-imp-end}*
  \end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\tst}} \\
    \end{tabular}
  \end{center}
  \caption{AGREE Unit test for high-assurance monitor.}
  \label{fig:test}
\end{figure}

\figref{fig:test} is a unit test for the high-assurance monitor in \figref{fig:hardened}.
The test checks if the monitor output coincides with the response that comes one step after the request.
It also checks that the alarm does not fire.
\lineref{line:tst-index} creates an index to use to create the test input.
\lineref{line:tst-assume-1} defines when the response appears relative to the index.
\lineref{line:tst-assume-2} defines the request behavior.
\linesref{line:tst-guarantee-1}{line:tst-guarantee-2} define the expected test output.
The test additionally proves that nothing happens after the initial test input and output (e.g., no more input is provided and no more output is generated).
AGREE proves the test correct.
\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp})

This AGREE test method is effective for detecting vacuity, under-specified behavior, and missed corner cases.
It enables designers to iteratively develop specifications inside the MBE framework using AGREE, and it builds assurance that the specification reflects the desired behavior with black-box, white-box, and other test coverage metrics.
These same tests can be synthesized to the backend implementation for traditional testing.

Traditional testing is important for the implementations of components that are not high-assurance.
The behavior of regular components must be modeled in AGREE since they are part of any system implementation, and their specifications provide input to high-assurance components.
Like the high-assurance components, these specifications must be meaningful in modeling the representative component.
Testing is an effective means for validating such specifications, and AGREE can use the existing tests from the component implementation to check the specifications, and it can generate additional tests that probe the behavior of the component implementation when writing the specifications.

### 2022-03-19 paper/paper.tex
Add case study outline

@@
 
-\section{Testing Specifications with AGREE}
-\label{sec:testing}
-\input{testing}
-
 \section{Case studies}
@@
 \begin{compactitem}
+  \item Possible Outline: 
+  \begin{compactitem}
+    \item Two case studies: phase 2 is the proof on concept on a industrial size problem while phase 3 tests if AGREE is expressive enough for stating complex properties and if designers are able to use it.
+    \item Phase 2 write up focusses on the number of components and the relative ease for AGREE to prove out and for SPLAT to synthesize. It notes that the components were relatively simple.
+    \item Phase 3 write up focusses on the shear complexity of the components and in particular the gate and the transport monitor.
+Here it the need for testing is is most obvious.
+Also obvious is the need to make AGREE more expressive and to separate AGREE into a specification and implementation language as seen in Dafny.
+    \item Make the general observation that the transforms on the AADL model to add the components and connections doesn't really save much time when compared against the time it takes to write and test the specifications.
+The automations is nice, but not where time is saved.
+Where time is saved is in the assurance case because once a specification is done, it is synthesized and proved correct for ALL inputs and ALL outputs.
+Not just some.
+Not code inspections.
+Not additional test requirements.
+All that is saved.
+  \end{compactitem}
   \item Clarify research questions answered by the case study
@@
 \end{compactitem}
+
+\section{Testing Specifications with AGREE}
+\label{sec:testing}
+\input{testing}
 

### 2022-03-20 paper/agree.tex
checkpointing; outlined remainder of semantics section
added definition of code contract; other re-orgs and tweakage

@@
-This section gives an overview of AGREE's assume-guarantee reasoning
-system, abstractly defining the verification conditions that AGREE
-checks in order to establish overall system correctness at the model
-level.  The discussion then becomes more concrete, showing how the
-AGREE contract specification language is used to specify the
-cyber-hardened system of \figref{fig:hardened}. It explains the the
-top-level specification of the example system, then shifts to the
-specifications for the filter and the monitor.  These two
-specifications are those that are synthesized to equivalent CakeML by
-SPLAT.
+This section gives an abstract overview of AGREE's compositional
+reasoning system, defining the verification conditions that AGREE
+checks in order to modularly establish overall system correctness at
+the model level.  The discussion then becomes more concrete,
+showing---by example---how the AGREE contract language is used to
+specify the cyber-hardened system of \figref{fig:hardened}.  After the
+system-level contract, discussion shifts to the filter and monitor
+component specifications, which are structured to support code
+generation.
 
@@
 
-\subsection{Verification Conditions}
+\subsection{Model of computation}
+
+The AGREE specification language is based on stream concepts, and
+operators, from the Lustre language \cite{10.1145/41625.41641}.  Thus
+the setting is synchronous dataflow where the inputs, outputs, and
+expressions are data streams that comply with the input assumptions.
+Contracts are evaluated in dependency order with inputs being
+propagated to outputs through all contracts until they stabilize; as
+such, the contracts, and thereby the top-level model, must be
+acyclic. (An apparent syntactic cycle, where a component is linked
+back to itself, may be broken temporally by inserting delay
+elements.) \footnote{KLS: some undefined terms here: evaluation of
+contracts, stabilization, synchronous step.}
+
+Once the contracts have stabilized, the model takes a synchronous step
+to the next input data in the input stream(s).  The semantics do not
+model computation or communication delay.\footnote{KLS: how does a
+delay element (mentioned above) work then?}  The output of one
+contract is seen at the input of any downstream contract in the same
+step of the input data stream.
+\footnote{KLS: This subsection should give a helpful abstract description of what AG specifications are about, \ie, constraints on IO stream. Needs a rewrite, however.}
+
+\subsection{Verification conditions}
 \input{verification-conditions}
 
-\subsection{System Contract}
+\subsection{System contract}
 \input{sw-contract}
 
-\subsection{Cyber-hardening the System Implementation}
+\subsection{Cyber-hardening the system implementation}
 \input{high-assurance-contracts}

### 2022-03-20 paper/example.tex
added definition of code contract; other re-orgs and tweakage

@@
 The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further, that response must come with the request or in the next step after the request.
-As with the filter, the system designer provides the policy, and that policy is typically based on existing AGREE specifications in the SW system.
+As with the filter, the system designer provides the policy, and that policy is typically based on existing AADL/AGREE definitions in the SW system.
 
-The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened-certificate}.
-Here AGREE provides a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the untrusted AI route planner.
-Having passed AGREE verification, the high-assurance components are ready to be synthesized.
+The AGREE analysis of the cyber-hardened implementation is shown in
+\figref{fig:hardened-certificate}. AGREE has generated hihg-level
+evidence justifying the claim that the high-assurance components
+guarantee the correct behavior of the SW implementation in the
+presence of the untrusted AI route planner.  Having passed AGREE
+verification, the high-assurance components are ready to be
+synthesized.
 
-The high-assurance components are automatically synthesized by SPLAT from the AGREE specifications to equivalent models in the CakeML language.
-The synthesis includes a proof that it captures the meaning of the AGREE specification in CakeML.
-In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized CakeML code exactly match those defined by the high-assurance component's guarantees.
-From here, CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code \cite{cakeml}.
+High-assurance components are automatically synthesized by SPLAT from
+the AGREE specifications to equivalent programs in the CakeML
+language.  The synthesis toolchain generates a proof that equates the
+meaning of the AGREE specification to the behavior of the CakeML
+program.  In other words, for any set of input streams that meet the
+component's contract assumptions, the output streams produced from the
+synthesized CakeML code exactly match those defined by the
+high-assurance component's guarantees.  CakeML provides verified
+compilation to binaries for several different platforms, meaning that
+the resulting binaries exactly preserve the meaning of the original
+CakeML code \cite{cakeml}.
 
-Preserving the input to output relationship of streams between the AGREE contracts and CakeML lifts the AGREE contract verification results to the actual deployed system.
-If the contract model verification succeeds, then the meaning of those results hold for the deployed system.
-These results, however, are only valid under additional assumptions on the deployed system: contracts for non-synthesized components accurately model their deployed counterparts, an appropriate schedule exists to sequence component execution following the dependent data-flow in the design, and the communication fabric stitching components together works in harmony with the schedule.
-Support and automation for these aspects of the design process are discussed in other works.
+Preserving the input to output relationship of streams between the
+AGREE contracts and CakeML lifts the AGREE contract verification
+results to the actual deployed system.  If the contract model
+verification succeeds, then the meaning of those results hold for the
+deployed system.  These results, however, are only valid under
+additional assumptions on the deployed system:
+\begin{itemize}
+\item contracts for non-synthesized components accurately model their deployed
+counterparts;
+\item an appropriate schedule exists to sequence component
+  execution following the dependent data-flow in the design; and
+\item the communication fabric stitching components together works in harmony
+  with the schedule.
+\end{itemize}
+
+Support and automation for these aspects of the design process are
+discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}

### 2022-03-20 paper/high-assurance-contracts.tex
added definition of code contract; other re-orgs and tweakage

--- 2022-03-18 paper/high-assurance-contracts.tex
@@
   \end{center}
-  \caption{Contract specification for high-assurance filter.}
+  \caption{Code contract specification for high-assurance filter.}
   \label{fig:filter}
@@
   \end{center}
-  \caption{Contract specification for high-assurance monitor.}
+  \caption{Code contract specification for high-assurance monitor.}
   \label{fig:monitor}
@@
 
-As noted previously, the system with the untrusted AI does not pass
-AGREE verification.  Transformations in BriefCASE cyber-harden the
-system against malicious AI behavior by inserting a high-assurance
-filter and monitor between the AI and the WM as shown in
-\figref{fig:hardened}.  (In our context \emph{high-assurance} means
-that the behaviors of the components are verified to meet their
-contracts.)  In the example, the inserted filter protects the WM from
-malformed data and the monitor protects the WM from (possibly malicious)
-spontaneous or delayed responses.  These high-assurance components are
-added, one at a time, by
+As previously discussed, transformations carried out in the BriefCASE
+interface have created a cyber-hardened system, protecting against
+malicious AI behavior, by inserting a high-assurance filter and
+monitor between the AI and the WM (see \figref{fig:hardened}). The
+filter protects the WM from malformed data and the monitor protects
+the WM from (possibly malicious) spontaneous or delayed responses.
+These components were added, one at a time, by
 \begin{itemize}
@@
 BriefCASE creates all the needed AADL for the new high-assurance
-component and its connections in the system implementation as part of
-the transformation.  It also creates a default AGREE specification
-that defines everything about the component behavior but the policy
-itself, unless that is provided in the dialogue.  The resulting AGREE
-specification from the transformation for the filter is shown in
-\figref{fig:filter}.  The single guarantee completely defines the
-filter output under any input according to the truth value of the
-policy.  Such a strong guarantee is needed for synthesis (a weaker
-version suffices for AGREE analysis).
+component, and its connections in the system implementation, as part
+of the transformation.  It also creates a default AGREE specification
+that defines everything about the component behavior, leaving only the
+component behavior (policy) to be specified.
 
-The AGREE specification for the monitor is shown in
-\figref{fig:monitor}. As can be seen, it is considerably more complex
-than the filter specification.  BriefCASE generates everything but the
-statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
+\subsubsection{Code contracts}
+
+Generally, AGREE specifications \emph{do not} describe the computation
+that a component performs. This is entirely by design: AGREE is
+intended to reason about component behavior solely at the
+specification level. However, the syntax of AGREE specifications
+provides enough expressiveness to support the notion of an
+AGREE \emph{code contract}: a component contract from which SPLAT can
+extract an implementation.
+
+\begin{definition}[Output guarantee]
+An \emph{output guarantee} is a stylized guarantee that fully
+specifies the data written to an output port. There are three
+possibilities according to whether the output port $p$ is
+a \konst{data} port, an \konst{event} port, or an \konst{event data}
+port:
+\[
+\begin{array}{ll}
+\konst{data}: &  p = \mathit{e} \\
+\konst{event}: &  \konst{event} (p) = \mathit{b} \\
+\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
+\end{array}
+\]
+\end{definition}
+
+\begin{definition}[Code contract]
+A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
+if $E \cup G \subset P_c$ and $E$ is a non-empty list of \konst{eq}
+statements and $G$ is a set of code guarantees, one for each element
+of $O_c$
+\end{definition}
+
+The AGREE code contracts for the filter and monitor are shown in
+\figref{fig:filter} and \figref{fig:monitor}. In general, the
+\konst{eq}-statements of a code contract keep track of
+state and provide data values from which to compute values to place on
+the outputs.
+
+\subsubsection{Monitor specification}
+ For the monitor, BriefCASE generates everything but the
+\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
 the filter, the guarantees completely define the output, as required
@@
 not just on the policy but also on configuration details supplied when
-the component is specified.  Setting \texttt{is\_latched} as true
-makes the alert persistent, meaning that once the alert is raised, it
-is always raised; otherwise, it is the complement of the policy value
-in the current step.
+the component is specified.  At configuration
+time, \texttt{is\_latched} has been set to \konst{true}, thus making
+the alert persistent, meaning that once the alert is raised, it is
+always raised; otherwise, it is the complement of the policy value in
+the current step.
 

### 2022-03-20 paper/semantics.tex
added definition of code contract; other re-orgs and tweakage

--- 2022-03-19 paper/semantics.tex
@@
 \]
+
+The temporal operators deal with time in more significant ways:
+\[
+\begin{array}{rcl}
+\sem{\konst{pre}(e)}_t & = & \sem{e}_{t-1}, \mathrm{when}\ t > 0 \\
+\sem{e_1 \to e_2}_t & = & \itelse{t=0}{\sem{e_1}_0}{\sem{e_2}_t} \\
+\sem{\konst{Hist}(b)}_t & = & \forall n \leq t.\; \sem{b}_n
+\end{array}
+\]
+
+The temporal operators can be used to construct higher-level operators
+from PTLTL, such as \konst{Once} and \konst{Since}; they are also used
+to manipulate state.
+
+\begin{itemize}
+\item \textit{strmStep}.
+
+Given a code contract an implementation will execute the equations in
+order, and then compute the values on outputs.
+
+\item \textit{Contract semantics}
+
+\item \textit{How example looks}
+
+\end{itemize}

### 2022-03-20 paper/sw-contract.tex
added definition of code contract; other re-orgs and tweakage

@@
 
-The AGREE specification language uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}.
-As with Lustre, the semantics are synchronous dataflow where the inputs, outputs, and expressions are data streams that comply with the input assumptions.
-Contracts are evaluated in dependency order with inputs being propagated to outputs through all contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic.\footnote{An apparent syntactic cycle, where a component is linked back to itself, may be broken temporally by inserting delay elements.}
-
-Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream.
-The semantics do not model computation or communication delay.
-The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
-The language is best introduced through example.
-
-The AGREE specification for the SW component in the example of Section~\ref{sec:example} is given in \figref{fig:sw}.
-The specification uses \texttt{eq} statements to define variables local to the contract specification.
-For example, \lineref{line:sw-event-def-start} defines the \texttt{req} variable to be equivalent to the \texttt{event} expression.
+The AGREE specification for the SW component in the example of
+Section~\ref{sec:example} is given in \figref{fig:sw}.  The
+specification uses \texttt{eq} statements to define variables local to
+the contract specification.  For
+example, \lineref{line:sw-event-def-start} defines the \texttt{req}
+variable to be equivalent to the \texttt{event} expression.
 

### 2022-03-21 paper/high-assurance-contracts.tex
added definition of evaluation, plus related discussion

@@
 
-Generally, AGREE specifications \emph{do not} describe the computation
+Generally, AGREE specifications do not describe the computation
 that a component performs. This is entirely by design: AGREE is
@@
 specification level. However, the syntax of AGREE specifications
-provides enough expressiveness to support the notion of an
-AGREE \emph{code contract}: a component contract from which SPLAT can
-extract an implementation.
+provides enough expressiveness to support the notion of a
+\emph{code contract}: a component contract from which an implementation
+can be extracted.
 
@@
 A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
-if $E \cup G \subset P_c$ and $E$ is a non-empty list of \konst{eq}
-statements and $G$ is a set of code guarantees, one for each element
-of $O_c$
+if $\mathit{Eqs} \cup G \subseteq P_c$, where
+\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
+of \konst{eq} statements and $G$ is the set of output guarantees, one for
+each element of $O_c$. In the interpretation as code, the order of
+elements of $\mathit{Eqs}$ is important, and is simply taken to be the
+occurrence order of the \konst{eq} statements in the syntax. Thus we
+will work with the
+\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
 \end{definition}
 
-The AGREE code contracts for the filter and monitor are shown in
-\figref{fig:filter} and \figref{fig:monitor}. In general, the
+Code contracts for the filter and monitor are shown in
+\figref{fig:filter} and \figref{fig:monitor}.\footnote{KLS: do we discuss wellformedness anywhere?} In general, the
 \konst{eq}-statements of a code contract keep track of
@@
 \subsubsection{Monitor specification}
- For the monitor, BriefCASE generates everything but the
-\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
-the filter, the guarantees completely define the output, as required
-for synthesis, in terms of the policy and the alert.  The
+BriefCASE generates everything but the
+\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
+\figref{fig:monitor}.  As with the filter, the guarantees
+completely define the output, as required for synthesis, in terms of
+the policy and the alert.  The
 \texttt{alert} variable defined on \lineref{line:mon-alert} depends

### 2022-03-21 paper/intro.tex
added definition of evaluation, plus related discussion

@@
 
-CASE was scoped to address common cyber-vulnerabilities, such as
-overflow, lack of input validation, supply-chain issues, and identity
-verification; however, other cyber-vulnerabilities such as
-side-channel attacks and denial of service are yet dealt with in our
-work.  Research in CASE has not invented a new type of filter or
-monitor, in terms of capability; instead, our contribution is in the
-automated synthesis of these components from a formal specification,
-and a means to show that the synthesis is correct.
+Our approach currently applies to common cyber-vulnerabilities, such
+as overflow, lack of input validation, supply-chain issues, and
+identity verification; however, other cyber-vulnerabilities such as
+side-channel attacks and denial of service are not yet dealt with in
+our work.  Here, we do not report on the invention of a new type of
+filter or monitor, in terms of capability; instead, our contribution
+is in the automated synthesis of security-improving components from
+formal specifications, and a means to show that the synthesis is
+correct.
 

### 2022-03-21 paper/semantics.tex
added definition of evaluation, plus related discussion

--- 2022-03-19 paper/semantics.tex
@@
 e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
-a & ::= & e\; \set{=,<,\leq}\; e \\
+a & ::= & e\; \set{=,<}\; e \\
 b & ::= & v \mid c \mid a \mid \neg b
@@
 readily extended to include records and arrays. A value stream is a
-total function from natural numbers to values:
+total function from time (natural numbers) to values:
 \[
@@
 Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
-binding variable names to value streams, the semantics $\sem{-}$ of terms and
-formulas, omitting temporal operators, is standard.
+binding variable names to value streams, the semantics $\sem{-}^E_t$
+of terms and formulas, omitting the temporal operators, is
+conventional. The value of a variable $v$ at time $t$ is found by
+looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
+the value of $s$ at time $t$.
 
@@
 \begin{array}{rcl}
-\sem{v}_t & = & E(v)(t) \\
-\sem{c}_t & = & c \\
-\sem{e_1 + e_2}_t & = & \sem{e_1}_t + \sem{e_2}_t \\
+\sem{v}^E_t & = & E(v)(t) \\
+\sem{c}^E_t & = & c \\
+\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
    & \cdots & \\
-\sem{b_1 \land b_2}_t & = & \sem{b_1}_t \land \sem{b_2}_t \\
+\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
    & \cdots & \\
@@
 \]
+
+The temporal operators deal with time in more significant ways. For
+example, they are used to define higher-level operators from PTLTL,
+such as \konst{Once} and \konst{Since}. In code contracts, the
+temporal operators are essential for manipulating state variables
+between component invocations.
+
+\[
+\begin{array}{rcl}
+\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
+\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
+\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
+\end{array}
+\]
+
+Code generation is a multi-step transformation that starts with a code
+contract and ends with an executable. The first step maps the contract
+to a code-focused representation. Assume given code contract
+$(I,O,A,\mathit{Eqs} \cup G \cup P)$, environment $E$, and time
+$t$. The evaluation of $\mathit{Eqs}$ followed by the evaluation of
+$G$ results in a new environment $E'$ where the stream values for
+state variables and outputs have been computed for time $t$. The step
+from $E$ to $E'$ is one full cycle in the repeated evaluation of the
+component.
+
+\begin{definition}[Evaluation]
+We overload existing notation and write the evaluation of
+$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
+$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
+an environment transformation
+\[
+ \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
+\]
+modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
+{\Eqs} is evaluated by folding the transformation left-to-right
+through it. The transformation is also applied to the list of output
+guarantees $G$, computing the values on output ports. The details are
+omitted, being a bit messy because of the different kinds of output
+ports available.
+\end{definition}
+
+\begin{definition}[Code contract correctness]
+Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup P)$ and
+environment $E$.  The contract is \emph{correct}, if for all $t$,
+whenever the assumptions (historically) hold in $E$ and evaluation
+steps from $E$ to $E'$, then the guarantees hold in $E'$:
+\[
+\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
+\]
+\end{definition}
+\footnote
+{KLS: At this point we would like to assert that the AGREE ``consistency
+ check'' for a contract implies code contract correctness.}
+
+The evaluation semantics given above is based on the stream
+computation model, where values `deep in the past' can be accessed in
+computations. However, the conventional program languages we wish to
+target do not support such a view. Thus code generation is only
+defined for a class of \emph{well formed} equations.
+
+\begin{definition}[Wellformedness]
+
+\end{definition}
+
+\subsubsection{Example}

### 2022-03-21 paper/example.tex
Add todos and fixed spelling

@@
 The AGREE analysis of the cyber-hardened implementation is shown in
-\figref{fig:hardened-certificate}. AGREE has generated hihg-level
+\figref{fig:hardened-certificate}. AGREE has generated high-level
 evidence justifying the claim that the high-assurance components
@@
 \begin{itemize}
-\item contracts for non-synthesized components accurately model their deployed
+\item first, contracts for non-synthesized components accurately model their deployed
 counterparts;
-\item an appropriate schedule exists to sequence component
+\item second, an appropriate schedule exists to sequence component
   execution following the dependent data-flow in the design; and
-\item the communication fabric stitching components together works in harmony
+\item third the communication fabric stitching components together works in harmony
   with the schedule.
 \end{itemize}
-
-Support and automation for these aspects of the design process are
+\noindent Support and automation for these aspects of the design process are
 discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}

### 2022-03-21 paper/high-assurance-contracts.tex
Add todos and fixed spelling

@@
 \begin{itemize}
-  \item selecting the connection in the model where the
+  \item first, selecting the connection in the model where the
     component is to appear,
-  \item choosing the appropriate transformation, and
-  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
+  \item second, choosing the appropriate transformation, and
+  \item third, specifying the relevant filtering or monitoring \emph{policy} (behavior).
 \end{itemize}

### 2022-03-21 paper/paper.tex
added definition of evaluation, plus related discussion
Add todos and fixed spelling

@@
 \newcommand{\Lang}[1]{\ensuremath{{\mathcal L}({#1})}}
+\newcommand{\Eqs}{\ensuremath{\mathit{Eqs}}}
 \newcommand{\itelse}[3]{\mbox{$\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}$}}
@@
 \begin{compactitem}
+  \item Revise the end of the abstract regarding the case studies to be consistent with what ends up in this version of the paper.
+  \item Revise the end of the introduction that discusses the case studies to be consistent with what ends up in this version of the paper.
+  \item Fix the link to the motivating example to point to the new repository once it is created. Fix in the next section too.
   \item Add discussion on the limitation of the approach: what attacks can and cannot be covered, can valid messages be rejected, etc.
@@
     \end{compactitem}
-  \item Add citations to last paragraph for scheduled AGREE and HAMR.
+  \item Add citations, and remove KLS footnote to last paragraph---should cite HAMR, seL4, and Adventium.
 \end{compactitem}
@@
   \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
+  \item Add reference to the Liskhov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.
+  \item Add back in a few sentences on the contract for the filter since those have been removed with the addition of the formal definitions.
 \end{compactitem}

### 2022-03-21 paper/sw-contract.tex
Add todos and fixed spelling

@@
 The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
-It differentiates with the followed-by operator what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincides with the request or is a response to a request one step earlier.
+It differentiates with the followed-by operator what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
 The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.
@@
 The last guarantee ensures that alert is persistent.
-The \texttt{Once} operator is true if the enclosed expression has ever been true before the current moment.
-\lineref{line:sw-guarantee-6} ensures if alert has ever happened in the past, then it is happening in the present moment.
+The \texttt{Once} operator is true if the enclosed expression has ever been true before, or in, the current moment.
+\lineref{line:sw-guarantee-6} ensures if alert has ever happened, then it is happening in the present moment.
 These six guarantees define the behavior of the SW component under the three assumptions.

### 2022-03-21 paper/high-assurance-contracts.tex
moving things around; result is not entirely coherent

@@
 
-\subsubsection{Code contracts}
-
-Generally, AGREE specifications do not describe the computation
-that a component performs. This is entirely by design: AGREE is
-intended to reason about component behavior solely at the
-specification level. However, the syntax of AGREE specifications
-provides enough expressiveness to support the notion of a
-\emph{code contract}: a component contract from which an implementation
-can be extracted.
-
-\begin{definition}[Output guarantee]
-An \emph{output guarantee} is a stylized guarantee that fully
-specifies the data written to an output port. There are three
-possibilities according to whether the output port $p$ is
-a \konst{data} port, an \konst{event} port, or an \konst{event data}
-port:
-\[
-\begin{array}{ll}
-\konst{data}: &  p = \mathit{e} \\
-\konst{event}: &  \konst{event} (p) = \mathit{b} \\
-\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
-\end{array}
-\]
-\end{definition}
-
-\begin{definition}[Code contract]
-A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
-if $\mathit{Eqs} \cup G \subseteq P_c$, where
-\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
-of \konst{eq} statements and $G$ is the set of output guarantees, one for
-each element of $O_c$. In the interpretation as code, the order of
-elements of $\mathit{Eqs}$ is important, and is simply taken to be the
-occurrence order of the \konst{eq} statements in the syntax. Thus we
-will work with the
-\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
-\end{definition}
-
 Code contracts for the filter and monitor are shown in

### 2022-03-21 paper/semantics.tex
moving things around; result is not entirely coherent

@@
 specifications to CakeML executables.
-
-\subsection{AGREE syntax and semantics}
-\label{agree-semantics}
-
-The syntax of AGREE is essentially that of quantifier-free first order
-logic (FOL) supplemented with a few temporal operators. The terms
-(\emph{e}) are the usual arithmetic expressions built from variables
-and numeric literals, while formulas (\emph{b}) are built using the
-standard logical connectives from atomic formulas (\emph{a}) based on
-the usual comparison operators.
-\[
-\begin{array}{rcl}
-e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
-a & ::= & e\; \set{=,<}\; e \\
-b & ::= & v \mid c \mid a \mid \neg b
-            \mid b \; \set{\land,\lor,\imp,\iff}\; b
-\end{array}
-\]
-
-There is a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and
-formulas. There are also temporal operators $\konst{pre}(-)$, delay
-$(-) \to (-)$, and $\konst{Hist}(-)$.
-
-The semantics of terms and formulas is in terms of \emph{streams of
-values}. Values encompass at least booleans and numbers, but can be
-readily extended to include records and arrays. A value stream is a
-total function from time (natural numbers) to values:
-\[
- \konst{stream} = \mathbb{N} \to \konst{value}
-\]
-Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
-binding variable names to value streams, the semantics $\sem{-}^E_t$
-of terms and formulas, omitting the temporal operators, is
-conventional. The value of a variable $v$ at time $t$ is found by
-looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
-the value of $s$ at time $t$.
-
-\[
-\begin{array}{rcl}
-\sem{v}^E_t & = & E(v)(t) \\
-\sem{c}^E_t & = & c \\
-\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
-   & \cdots & \\
-\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
-   & \cdots & \\
-\end{array}
-\]
-
-The temporal operators deal with time in more significant ways. For
-example, they are used to define higher-level operators from PTLTL,
-such as \konst{Once} and \konst{Since}. In code contracts, the
-temporal operators are essential for manipulating state variables
-between component invocations.
-
-\[
-\begin{array}{rcl}
-\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
-\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
-\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
-\end{array}
-\]
 

### 2022-03-21 paper/verification-conditions.tex
moving things around; result is not entirely coherent

@@
 
-A system is a tuple, $\system = (\inputs, \outputs, \assumes, \guarantees)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, and $\guarantees$ is the set of guarantees.
-A system is implemented by a set of components, $\components$, with their various connections.
-A component $\component$ is defined by its own tuple in a manner similar to the system as $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component)$.
-From the components and connections, $\mathbb{I}_\component$ is defined such that it is the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined such that it is the set of components that provide the output for the system.
-
-An system is correct if and only if for all $\pi$ and for all $i \ge 0$ the following holds:
+A \emph{system} $\system = (\inputs, \outputs, \assumes,
+\guarantees)$, where $\inputs$ is the input set, $\outputs$ is the
+output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
+the set of guarantees.  A system is implemented by a set of
+\emph{components}, $\components$, together with with their
+inter-connections.  A component $\component$ is defined by its own
+tuple in a manner similar to the system as $(\inputs_\component,
+\outputs_\component, \assumes_\component, \guarantees_\component)$.
+From the components and connections, $\mathbb{I}_\component$ is
+defined to be the set of components providing input to some component
+$\component$ in the system, and $\mathbb{O}$ is defined to be the set
+of components that provide the output for the system.  A system is
+correct if and only if for all $\pi$ and for all $i \ge 0$ the
+following conditions hold:
 \[
@@
 \]
-The first condition verifies the input assumptions on each component under the system assumptions and upstream component guarantees. 
-It checks if the component guarantees and system assumptions are strong enough to imply input assumptions on all immediate downstream components.
-The second condition checks the output guarantees of the system under the system assumptions and component guarantees that provide the output. 
-It checks if the guarantees on components providing primary outputs are strong enough to imply the system guarantees.
+The first condition verifies the input assumptions on each component
+under the system assumptions and upstream component guarantees.  It
+checks if the component guarantees and system assumptions are strong
+enough to imply input assumptions on all immediate downstream
+components.  The second condition checks the output guarantees of the
+system under the system assumptions and component guarantees that
+provide the output.  It checks if the guarantees on components
+providing primary outputs are strong enough to imply the system
+guarantees.
 
@@
 
-The expanded property lists in \figref{fig:example-certificate} and \figref{fig:hardened-certificate} are the results from disproving or proving out the above verification conditions.
-The additional unexpanded results at the bottom of the figures prove \emph{self-consistency} in the contracts.
-It is not uncommon to accidentally write contracts that are self-contradicting.
-For example, a contract may guarantee an output be two different values in the same moment of time.
-AGREE generates additional verification conditions that prove each component contract, and the composition of contracts, self-consistent.
+The expanded property lists in \figref{fig:example-certificate} and
+\figref{fig:hardened-certificate} are the results from verifying or
+disproving the above verification conditions.  The additional
+unexpanded results at the bottom of the figures prove
+\emph{self-consistency} in the contracts.  It is not uncommon to
+accidentally write contracts that are self-contradicting.  For
+example, a contract may guarantee an output be two different values in
+the same moment of time.  AGREE generates additional verification
+conditions that prove each component contract, and the composition of
+contracts, self-consistent.\footnote{KLS:Is this the same as
+satisfiable?}
+
+\subsection{Syntax and semantics of AGREE}
+\label{agree-semantics}
+
+The syntax of AGREE is essentially that of quantifier-free first order
+logic (FOL) supplemented with a few temporal operators. The terms
+(\emph{e}) are the usual arithmetic expressions built from variables
+($v$) and numeric and boolean literals ($c$), while formulas
+(\emph{b}) are built using the standard logical connectives from
+atomic formulas (\emph{a}) based on the usual comparison operators.
+\[
+\begin{array}{rcl}
+e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
+a & ::= & e\; \set{=,<}\; e \\
+b & ::= & v \mid c \mid a \mid \neg b
+            \mid b \; \set{\land,\lor,\imp,\iff}\; b
+\end{array}
+\]
+
+Supplementing this basic syntax are is a conditional,
+$\itelse{b}{(-)}{(-)}$, for both terms and formulas. There are also
+temporal operators $\konst{pre}(-)$, delay $(-) \to (-)$, and
+$\konst{Hist}(-)$.
+
+The semantics of terms and formulas is in terms of \emph{streams of
+values}. Values encompass at least booleans and numbers, but can be
+readily extended to include records and arrays. A value stream is a
+total function from time (natural numbers) to values:
+\[
+ \konst{stream} = \mathbb{N} \to \konst{value}
+\]
+Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
+binding variable names to value streams, the semantics $\sem{-}^E_t$
+of terms and formulas, omitting the temporal operators, is
+conventional. The value of a variable $v$ at time $t$ is found by
+looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
+the value of $s$ at time $t$.
+
+\[
+\begin{array}{rcl}
+\sem{v}^E_t & = & E(v)(t) \\
+\sem{c}^E_t & = & c \\
+\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
+   & \cdots & \\
+\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
+   & \cdots & \\
+\end{array}
+\]
+
+The temporal operators deal with time in more significant ways. The
+value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
+$t-1$ (at time zero, \konst{pre} is undefined).  A delay $e_1 \to e_2$
+temporally ``shifts'' $e_2$ by means of prepending the first element
+of $e_1$ to it.
+
+\[
+\begin{array}{rcl}
+\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
+\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
+\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
+\end{array}
+\]
+
+Although basic, these definitions can be used to define higher-level
+operators from PTLTL, such as \konst{Once} and \konst{Since}.
+
+\subsection{Code contracts}
+
+Generally, AGREE specifications do not describe the computation
+that a component performs. This is entirely by design: AGREE is
+intended to reason about component behavior solely at the
+specification level. However, the syntax of AGREE specifications
+provides enough expressiveness to support the notion of a
+\emph{code contract}: a component contract from which an implementation
+can be extracted.
+
+\begin{definition}[Output guarantee]
+An \emph{output guarantee} is a stylized guarantee that fully
+specifies the data written to an output port. There are three
+possibilities according to whether the output port $p$ is
+a \konst{data} port, an \konst{event} port, or an \konst{event data}
+port:
+\[
+\begin{array}{ll}
+\konst{data}: &  p = \mathit{e} \\
+\konst{event}: &  \konst{event} (p) = \mathit{b} \\
+\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
+\end{array}
+\]
+\end{definition}
+
+\begin{definition}[Code contract]
+A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
+if $\mathit{Eqs} \cup G \subseteq P_c$, where
+\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
+of \konst{eq} statements and $G$ is the set of output guarantees, one for
+each element of $O_c$. In the interpretation as code, the order of
+elements of $\mathit{Eqs}$ is important, and is simply taken to be the
+occurrence order of the \konst{eq} statements in the syntax. Thus we
+will work with the
+\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
+\end{definition}

### 2022-03-22 paper/abstract.tex
Add todos and fixed spelling
fiddling with contracts discussion

@@
-Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either gracefully recover or safely shut down as requirements dictate.
-The DARPA Cyber Assured Systems Engineering program is developing tools for design, analysis, and verification that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment.
-This paper describes automated model transformations that introduce high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively.
-A formal specification defines each high-assurance component, and is used to verify that the component addresses system level cyber requirements.
-Implementations for these high-assurance components are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level.
-The model transformations are integrated into the Open Source AADL Tool Environment (OSATE).
-The paper further reports on a case study applying security-enhancing model transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
-In the case study, the model transformations add filters to guard against malformed input, as well as monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.
+Cyber-physical systems, such as avionics, must be tolerant to
+cyber-attacks in the same way they are tolerant to random faults: they
+either gracefully recover or safely shut down as requirements dictate.
+The DARPA Cyber Assured Systems Engineering program is developing
+tools for design, analysis, and verification that enable systems
+engineers to design-in cyber-resiliency in a Model-Based Systems
+Engineering environment.  This paper describes automated model
+transformations that introduce high-assurance cyber-resiliency
+components into a system, in particular filters and monitors that
+prevent malicious input and detect supply chain attacks, respectively.
+A formal specification defines each high-assurance component, and is
+used to verify that the component addresses system level cyber
+requirements.  Implementations for these high-assurance components are
+directly synthesized from their specifications, and are automatically
+proven to preserve the exact meaning of the specifications all the way
+down to the binary code level.  The model transformations are
+integrated into the Open Source AADL Tool Environment (OSATE).  The
+paper further reports on a case study applying cyber-resiliency model
+transformations to a UAV system that uses the Air Force Research
+Laboratory's OpenUxAS services for route planning.  In the case study,
+the model transformations add filters, to guard against malformed
+input, as well as monitors, to guard against ground station spoofing
+and malicious flight plans from OpenUxAS.

### 2022-03-22 paper/intro.tex
Add todos and fixed spelling
fiddling with contracts discussion

@@
 SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
-(Semantic Properties for Language and Automata Theory).  Formal
+(Semantic Properties for Language and Automata Theory).
+
+Formal
 verification that the transformed system model meets its requirements
@@
 
-Two security-enhancing transformations are discussed: (1) the
-insertion of a filter to prevent malformed data from a malicious actor
-from being propagated to downstream components, and (2) the insertion
-of a monitor to detect (and alert) unexpected behaviors arising from
+Two cyber-resiliency transformations are discussed: (1) the insertion
+of a filter to prevent malformed data from a malicious actor from
+being propagated to downstream components, and (2) the insertion of a
+monitor to detect (and alert) unexpected behaviors arising from
 untrusted components. These transformations not only change the

### 2022-03-22 paper/verification-conditions.tex
fiddling with contracts discussion

@@
 output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
-the set of guarantees.  A system is implemented by a set of
-\emph{components}, $\components$, together with with their
-inter-connections.  A component $\component$ is defined by its own
-tuple in a manner similar to the system as $(\inputs_\component,
-\outputs_\component, \assumes_\component, \guarantees_\component)$.
-From the components and connections, $\mathbb{I}_\component$ is
-defined to be the set of components providing input to some component
-$\component$ in the system, and $\mathbb{O}$ is defined to be the set
-of components that provide the output for the system.  A system is
-correct if and only if for all $\pi$ and for all $i \ge 0$ the
-following conditions hold:
-\[
-\begin{array}{lll}
-        & \forall \component \in \components &
+the set of guarantees.  A system is composed of a set of
+\emph{components}, $\components$, together with their
+inter-connections.  A component $\component$ is, hierarchically, also
+a system, and may be designated by its own tuple
+$(\inputs_\component, \outputs_\component, \assumes_\component,
+\guarantees_\component)$.  From the components and connections,
+$\mathbb{I}_\component$ is defined to be the set of components
+providing input to some component $\component$ in the system, and
+$\mathbb{O}$ is defined to be the set of components that provide the
+output for the system.  A system $S$ is \emph{correct} if and only if for all
+components $c \in C$:
+\begin{equation}
             \globally(\historically(\assumes \wedge
             \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
-            \implies \assumes_\component) \\
- \wedge &   &
+            \implies \assumes_\component)
+\end{equation}
+\begin{equation}
             \globally(\historically(\assumes \wedge
@@
             \implies \guarantees)
-\end{array}
-\]
-The first condition verifies the input assumptions on each component
-under the system assumptions and upstream component guarantees.  It
-checks if the component guarantees and system assumptions are strong
-enough to imply input assumptions on all immediate downstream
-components.  The second condition checks the output guarantees of the
-system under the system assumptions and component guarantees that
-provide the output.  It checks if the guarantees on components
-providing primary outputs are strong enough to imply the system
-guarantees.
+\end{equation}
+Condition (1) verifies the input assumptions on each component under
+the system assumptions and upstream component guarantees.  It checks
+if the component guarantees and system assumptions are strong enough
+to imply input assumptions on all immediate downstream components.
+Condition (2) checks the output guarantees of the system under the
+system assumptions and component guarantees that provide the output.
+It checks if the guarantees on components providing primary outputs
+are strong enough to imply the system guarantees.
 
-AGREE uses $k$-inductive model checking to prove or disprove each generated verification condition.
-If all the verification conditions hold, then the system is said to be \emph{correct}, meaning that the system composition meets input assumptions at each input as well as the guarantees on the system output.
-A consequence of this result is that $\globally(\historically(\assumes) \implies \guarantees)$ holds for the system contract.
-As such, the system contract is a sound abstraction of the system implementation.
+AGREE uses $k$-inductive model checking to prove or disprove each
+generated verification condition.  If all the verification conditions
+hold, then the system is said to be \emph{correct}, meaning that the
+system composition meets input assumptions at each input as well as
+the guarantees on the system output. A consequence of this result is
+that $\globally(\historically(\assumes) \implies \guarantees)$ holds
+for the system contract. As such, the system contract is a sound
+abstraction of the system implementation.
 

### 2022-03-22 paper/example.tex
checkpointing ... mostly tweakage to examples section

@@
 
-\figref{fig:example} is an AADL architectural model of a software system (SW) for route planning and automated control of a UAV.
-It is loosely based on one of the case studies in Section~\ref{sec:case-study}.
-The source for the entire model is found at \cite{repo}.
+\figref{fig:example} is an AADL architectural model of a software
+system (SW) for route planning and automated control of a UAV.  It is
+loosely based on one of the case studies in
+Section~\ref{sec:case-study}.  The source for the entire model is
+found at \cite{repo}.
 
-The system receives an automation request that is forwarded to a third-party route planner (AI).
-The route planner decides the flight path of the UAV based on its current position and the requested task.
-The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
-The waypoint manager is a repurposed legacy component that cannot be modified.
+The system receives an automation request that is forwarded to a
+third-party route planner (AI).  The route planner decides the flight
+path of the UAV based on its current position and the requested task.
+The waypoint manager (WM) receives the mission command as a set of
+waypoints from the planner and starts the UAV flying the mission,
+issuing waypoints to the UAV flight controller as the UAV location
+changes.  The waypoint manager is a legacy component that cannot be
+modified.
 
-The expected behavior of the SW system, and the components implementing the system, are modeled with AGREE contract specifications.
-The contracts constrain input with assumptions and state properties of output with guarantees.
-AGREE performs model checking on this assume-guarantee system to hierarchically prove that the composite system obeys all contract obligations under all possible finite input streams.
-The AGREE analysis, contract language, and the contracts for this example system are discussed in greater detail in \secref{sec:agree}.
+The expected behavior of the SW system, and the components
+implementing the system, are modeled with AGREE contract
+specifications.  The contracts constrain input with assumptions and
+state properties of output with guarantees.  AGREE performs model
+checking on this assume-guarantee system to hierarchically prove that
+SW obeys all contract obligations under all possible finite input
+streams.  The AGREE analysis, contract language, and contracts for SW
+are discussed in greater detail in \secref{sec:agree}.
 
-The initial AGREE contracts for the system in \figref{fig:example}, and its implementing components, assume and guarantee the absence of any malicious, or unspecified, component behavior.
-More precisely, the system assumes that all inputs are \emph{well-formed} and there is never more than one automation request pending at a time.
-The meaning of well-formed is given by predicates for each data type.
-For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
-The system guarantees that a start includes a waypoint, a start is within one cycle of an automation request and if not, then it persistently alerts, new waypoints coincide with air vehicle location updates, and all output are well-formed.
+The initial AGREE contracts for the system in \figref{fig:example},
+and its implementing components, both assume and guarantee the absence
+of any malicious, or unspecified, component behavior.  More precisely,
+the system contract assumes that all inputs are \emph{well-formed} and
+there is never more than one automation request pending at a time.
+The meaning of well-formed is given by user-written predicates for
+each data type.  For example, a waypoint is well-formed if it falls
+within bounds for latitude, longitude, and altitude.  The guarantees
+for SW ensure that
+\begin{itemize}
+\item a start includes a waypoint,\footnote{KLS: maybe describe the start port earlier.}
+\item a start is within one cycle of an automation request and if not, then it persistently alerts,
+\item new waypoints coincide with air vehicle location updates, and
+\item all outputs are well-formed.
+\end{itemize}
 
-The components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
-The AI further guarantees it only responds to automation requests and always in the same cycle.
-The legacy WM further guarantees a start from a response and always in the same cycle, start always comes with a waypoint, and further waypoints coincide with air vehicle location updates.
-These initial specifications pass AGREE verification meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
+The contracts for the sub-components assume their inputs are
+well-formed, and they guarantee their outputs are well-formed.  The AI
+contract guarantees it only responds to automation requests and always
+in the same cycle.  The legacy WM contract guarantees the following: a
+start from a response and always in the same cycle; start always comes
+with a waypoint; and, further waypoints coincide with air vehicle
+location updates.  These initial specifications pass AGREE
+verification meaning that the contract composition of the components
+with the system satisfies all component input assumptions and system
+output guarantees.
 
-A cyber-vulnerability analysis, however, identifies the potential of a supply chain attack through the AI route planner that is provided by a third-party vendor without source code.
-The component is marked as \emph{untrusted} after the analysis, and its contract is modified by removing all guarantees on its outputs to reflect its untrusted status.
-The AI output is now unconstrained in the assume-guarantee reasoning of AGREE and is able to generate any value on its output stream at anytime.
+\subsection{Detecting and addressing vulnerabilities}
+A cyber-vulnerability analysis, however, identifies the potential of a
+supply chain attack through the AI route planner (provided by a
+third-party vendor without source code).  BriefCASE marks the
+component as \emph{untrusted} after the analysis, and the contract for
+AI is modified by removing all guarantees on its outputs to reflect
+its untrusted status.  The AI output is now unconstrained in the
+assume-guarantee reasoning of AGREE and is able to generate any value
+on its output stream at anytime.
 
-The output from the AGREE analysis using the untrusted AI specification is shown in \figref{fig:example-certificate}.
-The red exclamation points designate component assumptions or system output guarantees that do not hold, and each failure comes with a corresponding counter-example.
-The results are not unexpected given the AI route planner's unconstrained behavior.
+The output from the AGREE analysis using the untrusted AI
+specification is shown in \figref{fig:example-certificate}.  The red
+exclamation points designate component assumptions or system output
+guarantees that do not hold, and each failure comes with a
+corresponding counter-example.  The results are not unexpected given
+the AI route planner's unconstrained behavior.
 
-For example, the first violation is that the automation response from the AI to the WM is no longer guaranteed to be well-formed.
-The consequence of that failing input assumption is that the WM outputs are no longer guaranteed; they are effectively unconstrained.
-These missing output guarantees from the WM lead to the rest of the failures in \figref{fig:example-certificate} because the WM provides the system level outputs.
+For example, the first violation is that the automation response from
+the AI to the WM is no longer guaranteed to be well-formed.  The
+consequence of that failing input assumption is that the WM outputs
+are no longer guaranteed; they are effectively unconstrained.  These
+missing output guarantees from the WM lead to the rest of the failures
+in \figref{fig:example-certificate} because the WM provides the system
+level outputs.
 
@@
 
-The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}.
-A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input violates the filter invariant.
-The auto-generated AGREE specification states that only well-formed inputs are passed to the output.
-The system developer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.
+The system designer now uses BriefCASE to cyber-harden SW by inserting
+high-assurance components in the form of a filter and a monitor, as
+shown in \figref{fig:hardened}.  A filter enforces an invariant over
+each datum in the data stream by not forwarding input to its output if
+that input violates the filter invariant.  The auto-generated AGREE
+specification states that only well-formed inputs are passed to the
+output.  The system developer must further define the filtering
+policy, but it is usually based on the existing assumptions made by
+downstream components that consume the filter output.
 
-A monitor captures a relation on input data over time and is thus able to reason about temporal properties of that input.
-A monitor raises an alert if the specified temporal properties are ever violated.
-The AGREE specification for the monitor in our example states that an automation response can only be generated in conjunction with an automation request; and further, that response must come with the request or in the next step after the request.
-As with the filter, the system designer provides the policy, and that policy is typically based on existing AADL/AGREE definitions in the SW system.
+A monitor captures a relation on input data over time and is thus able
+to reason about temporal properties of that input.  A monitor raises
+an alert if the specified temporal properties are ever violated.  The
+AGREE specification for the monitor in our example states that an
+automation response can only be generated in conjunction with an
+automation request; and further, that response must come with the
+request or in the next step after the request.  As with the filter,
+the system designer provides the policy, and that policy is typically
+based on existing AADL/AGREE definitions in the SW system.
 
@@
 \begin{itemize}
-\item first, contracts for non-synthesized components accurately model their deployed
+\item contracts for non-synthesized components accurately model their deployed
 counterparts;
-\item second, an appropriate schedule exists to sequence component
+\item an appropriate schedule exists to sequence component
   execution following the dependent data-flow in the design; and
-\item third the communication fabric stitching components together works in harmony
+\item the communication fabric stitching components together works in harmony
   with the schedule.

### 2022-03-22 paper/verification-conditions.tex
checkpointing ... mostly tweakage to examples section

@@
 logic (FOL) supplemented with a few temporal operators. The terms
-(\emph{e}) are the usual arithmetic expressions built from variables
-($v$) and numeric and boolean literals ($c$), while formulas
-(\emph{b}) are built using the standard logical connectives from
-atomic formulas (\emph{a}) based on the usual comparison operators.
+(\emph{e}) are arithmetic expressions built from variables ($v$) and
+numeric and boolean literals ($c$), while formulas (\emph{b}) are
+built using logical connectives from atomic formulas (\emph{a}) based
+on the familiar comparison operators.
 \[
@@
 
-Supplementing this basic syntax are is a conditional,
-$\itelse{b}{(-)}{(-)}$, for both terms and formulas. There are also
-temporal operators $\konst{pre}(-)$, delay $(-) \to (-)$, and
-$\konst{Hist}(-)$.
+There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
+and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
+delay $(-) \to (-)$, and $\konst{Hist}(-)$.
 

### 2022-03-23 paper/agree.tex
fiddling with contracts discussion
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

--- 2022-03-19 paper/agree.tex
@@
-This section gives an overview of AGREE's assume-guarantee reasoning
-system, abstractly defining the verification conditions that AGREE
-checks in order to establish overall system correctness at the model
-level.  The discussion then becomes more concrete, showing how the
-AGREE contract specification language is used to specify the
-cyber-hardened system of \figref{fig:hardened}. It explains the the
-top-level specification of the example system, then shifts to the
-specifications for the filter and the monitor.  These two
-specifications are those that are synthesized to equivalent CakeML by
-SPLAT.
+This section gives an abstract mathematical overview of AGREE's
+compositional reasoning system, defining assume-guarantee contracts
+and the verification conditions that AGREE checks in order to
+modularly establish overall system correctness at the model level.
+The syntax and semantics of AGREE are sketched in some detail.  The
+discussion then becomes more concrete, showing---by example---how the
+AGREE contract language is used to specify the cyber-hardened system
+of \figref{fig:hardened}. The discussion proceeds from the
+system-level contract to the filter and monitor components, which rely
+on the notion of a \emph{code contract} to support code generation.
 
-%% These conditions are useful in understanding how the semantics in
-%% synthesis differ from those in the assume-guarantee reasoning
-%% in \secref{sec:semantics}.
-
-\subsection{Verification Conditions}
+\subsection{Contracts and verification conditions}
 \input{verification-conditions}
-
-\subsection{System Contract}
-\input{sw-contract}
-
-\subsection{Cyber-hardening the System Implementation}
-\input{high-assurance-contracts}

### 2022-03-23 paper/example.tex
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

@@
 
+\subsection{Code synthesis}
 High-assurance components are automatically synthesized by SPLAT from

### 2022-03-23 paper/high-assurance-contracts.tex
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

@@
 These components were added, one at a time, by
-\begin{itemize}
-  \item first, selecting the connection in the model where the
+\begin{enumerate}
+  \item selecting the connection in the model where the
     component is to appear,
-  \item second, choosing the appropriate transformation, and
-  \item third, specifying the relevant filtering or monitoring \emph{policy} (behavior).
-\end{itemize}
+  \item choosing the appropriate transformation, and
+  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
+\end{enumerate}
 

### 2022-03-23 paper/paper.tex
moving things around; result is not entirely coherent
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

@@
 
-\section{Assume-guarantee reasoning in AGREE}
+\section{Compositional reasoning}
 \label{sec:agree}
@@
   \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
-  \item Add reference to the Liskhov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.
+  \item Add reference to the Liskov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.
   \item Add back in a few sentences on the contract for the filter since those have been removed with the addition of the formal definitions.
@@
 
-\section{AGREE semantics for component synthesis}
-\label{sec:semantics}
-\input{semantics}
+\section{Formal contracts for the example system}
+\input{sw-contract}
+
+\subsection{Code contracts for cyber components}
+\input{high-assurance-contracts}
+
+%% \section{Synthesis from code contracts}
+%% \label{sec:semantics}
+%% \input{semantics}
 

### 2022-03-23 paper/synthesis.tex
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

@@
+The semantics of Section \ref{agree-semantics} supports a formal
+connection between the original contract (and verification results of
+the AGREE model checker being run on it), and code generated from
+specifications of leaf level cyber-components. It also provides the
+root meaning at the base of a chain of translation steps moving from
+AGREE specifications to CakeML executables.
+
+The first step in the chain maps the contract to a code-focused
+representation. Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup
+P,\emptyset)$, environment $E$, and time $t$. The evaluation of $\mathit{Eqs}$
+followed by the evaluation of $G$ results in a new environment $E'$
+where the stream values for state variables and outputs have been
+computed for time $t$. The step from $E$ to $E'$ is one full cycle in
+the repeated evaluation of the component.
+
+\begin{definition}[Evaluation]
+We overload existing notation and write the evaluation of
+$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
+$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
+an environment transformation
+\[
+ \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
+\]
+modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
+{\Eqs} is evaluated by folding the transformation left-to-right
+through it. The transformation is also applied to the list of output
+guarantees $G$, computing the values on output ports. The details are
+omitted, being a bit messy because of the different kinds of output
+ports available.
+\end{definition}
+
+\begin{definition}[Code contract correctness]
+Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
+The contract is \emph{correct}, if for all $E$ and $t$, whenever the
+assumptions (historically) hold in $E$ and evaluation steps from $E$
+to $E'$, then the guarantees hold in $E'$:
+\[
+\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
+\]
+\end{definition}
+\footnote
+{KLS: At this point we would like to assert that the AGREE ``consistency
+ check'' for a contract implies code contract correctness.}
+
+How does this notion of correctness integrate with the verification
+conditions addressed by model checking? In \emph{contract
+correctness}, the model checker is proving the following property
+\[
+\konst{Hist}(A) \imp P
+\]
+and in \emph{code contract correctness} we essentially prove a Hoare
+triple of the form
+\[
+\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
+\]
+This accords with intution, namely that AGREE is a kind of
+`program-free program logic'; by pulling out a program $(\mathit{Eqs}
+\cdot G)$ from a code contract we find ourselves in the tepid
+bathwater of imperative program verification. In order to use the code
+contract result in system contract verifcation conditions, one has to
+show that the code contract result implies the system verification
+condtion.
+
+\subsubsection*{Wellformedness}
+The evaluation semantics given above is based on the stream
+computation model, where values `deep in the past' can be accessed in
+computations. However, the conventional program languages we target
+target do not support such a view. Thus code generation is only
+defined for a class of \emph{wellformed} equations.
+
+\begin{definition}[Wellformedness]
+\end{definition}
+
+\emph{Discussion on mapping program from stream environment to logic function
+and then to CakeML}
+
 Synthesis maps from model and specifications to code. The synthesis
@@
 \end{remark}
-
-%% \emph{Need some discussion of filters and their step-wise properties
-%%   in relation to their infinitary properties. Reference to Johannes'
-%%   work.}
 
@@
 an alert output when the relation is violated.
-
-% A \emph{system} is a collection of \emph{components}, \emph{connections}
-% between components, a \emph{scheduler} to order execution, and a
-% \emph{system environment} for primary inputs.
 
@@
 
-\subsection{Component Behavior}
+%% \subsection{Component Behavior}
 
-Intuitively, for monitor specification $s$, \konst{stepFn} is the
-concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
-\ref{agree-semantics}. Its correctness amounts to showing that, given
-a sequence of inputs, and an initial state meeting the initialization
-constraints, iterating \konst{stepFn} produces a $\pi$ s.t. $\pi \in
-\Lang{s}$; and taking the union over all input sequences and
-initial states produces $\Lang{s}$ itself.
+%% Intuitively, for monitor specification $s$, \konst{stepFn} is the
+%% concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
+%% \ref{agree-semantics}. Its correctness amounts to showing that, given
+%% a sequence of inputs, and an initial state meeting the initialization
+%% constraints, iterating \konst{stepFn} produces a $\pi$ s.t. $\pi \in
+%% \Lang{s}$; and taking the union over all input sequences and
+%% initial states produces $\Lang{s}$ itself.

### 2022-03-23 paper/verification-conditions.tex
checkpointing. Re-org is done (for now). Mostly done describing AGREE contracts. Synthesis.tex merged with semantics.tex

@@
 
-AGREE generates, and then proves or disproves with model checking, a set of verification conditions to show that a system's component implementation is correct~\cite{agree2013}.
-The verification conditions are formalized in \emph{past-time linear temporal logic} (PLTL).
-PLTL is a first-order predicate calculus with temporal operators to reason about the truth values of formulas through time.
-Its semantics are defined relative to a finite trace of system states $\pi = s_0, s_1, \ldots$, and a moment of evaluation in the trace, $i$.
+The AGREE specification language is based on stream concepts, and
+operators, from the Lustre language \cite{10.1145/41625.41641}. Thus
+the setting is synchronous dataflow where the inputs and outputs of
+components are streams, and contracts express relationships between
+input and output streams. When considering a system of components,
+data flows through the components in dependency order, with inputs
+being propagated to outputs through all contracts until they stabilize
+(can't propagate further). Therefore, the subcomponent contracts, and
+thus the top-level model, must be acyclic. (An apparent syntactic
+cycle, where a component is linked back to itself, may be broken
+temporally by inserting delay elements.)  Once the data propagation
+has stabilized, the model proceeds to the next input data in the input
+streams. The semantics do not model computation or communication
+delay. The output of one contract is seen at the input of any
+downstream contract in the same step of the input data stream.
 
-The two PLTL operators necessary for the AGREE generated verification conditions are $\globally$ (globally) that looks forward in time along the trace and $\historically$ (historically) that looks backward in time along the trace.
-These are defined as
+From the system and component contracts, AGREE generates a set of
+verification conditions to show that a system's component
+implementation is correct~\cite{agree2013}.  The AGREE model checker
+is then invoked to prove or disprove the verification
+conditions. Contracts and verification conditions are expressed in
+\emph{past-time linear temporal logic} (PLTL).\footnote{KLS: citation
+needed.}  PLTL is a logic enhanced with temporal operators able to
+reason about the truth values of formulas through time.  Its semantics
+are defined relative to a point in time $i$ and a finite trace of
+system states $\pi = s_0, s_1, \ldots, s_i$.
+
+The two PLTL operators necessary for the AGREE generated verification
+conditions are $\globally$ (globally) that looks forward in time along
+the trace and $\historically$ (historically) that looks backward in
+time along the trace.  These are defined as
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-The $\models$-operator is read as \emph{satisfies}.
-A trace at a moment in time satisfies $\globally(f)$ if and only if it satisfies $f$ in the current and all future states of $\pi$.
-Globally is invariant from the current moment into the future and historically is invariant from the beginning of the trace to the current moment.
+The $\models$-operator is read as \emph{satisfies}.  A trace at a
+moment in time satisfies $\globally(f)$ if and only if it satisfies
+$f$ in the current and all future states of $\pi$.  $\globally(f)$ is
+invariant from the current moment into the future and $\historically$ is
+invariant from the beginning of the trace to the current moment.
 
 A \emph{system} $\system = (\inputs, \outputs, \assumes,
-\guarantees)$, where $\inputs$ is the input set, $\outputs$ is the
-output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
-the set of guarantees.  A system is composed of a set of
-\emph{components}, $\components$, together with their
-inter-connections.  A component $\component$ is, hierarchically, also
-a system, and may be designated by its own tuple
-$(\inputs_\component, \outputs_\component, \assumes_\component,
-\guarantees_\component)$.  From the components and connections,
-$\mathbb{I}_\component$ is defined to be the set of components
-providing input to some component $\component$ in the system, and
-$\mathbb{O}$ is defined to be the set of components that provide the
-output for the system.  A system $S$ is \emph{correct} if and only if for all
-components $c \in C$:
+\guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the
+output set, $\assumes$ is the set of assumptions, $\guarantees$ is the
+set of guarantees, and $C$ are subcomponents.  A subcomponent
+$\component$ is, hierarchically, also a system, and may be designated
+by its own tuple $(\inputs_\component, \outputs_\component,
+\assumes_\component, \guarantees_\component, C_c)$.  From the
+components and their connections, $\mathbb{I}_\component$ is defined
+to be the set of components providing input to some component
+$\component$ in the system, and $\mathbb{O}$ is defined to be the set
+of components that provide the output for the system.  A system $S$ is
+\emph{correct} if and only if for all components $c \in C$ the
+following two verification conditions hold:
 \begin{equation}
@@
 
-AGREE uses $k$-inductive model checking to prove or disprove each
-generated verification condition.  If all the verification conditions
-hold, then the system is said to be \emph{correct}, meaning that the
-system composition meets input assumptions at each input as well as
-the guarantees on the system output. A consequence of this result is
-that $\globally(\historically(\assumes) \implies \guarantees)$ holds
-for the system contract. As such, the system contract is a sound
-abstraction of the system implementation.
+If all the verification conditions hold (AGREE uses $k$-inductive
+model checking to automatically prove or disprove each generated
+verification condition), then the system is said to be \emph{correct},
+meaning that the system composition meets input assumptions at each
+input as well as the guarantees on the system output. A consequence of
+this result is that $\globally(\historically(\assumes) \implies
+\guarantees)$ holds for the system contract.
 
@@
 conditions that prove each component contract, and the composition of
-contracts, self-consistent.\footnote{KLS:Is this the same as
-satisfiable?}
+contracts, self-consistent.
 
@@
 
-The syntax of AGREE is essentially that of quantifier-free first order
-logic (FOL) supplemented with a few temporal operators. The terms
-(\emph{e}) are arithmetic expressions built from variables ($v$) and
-numeric and boolean literals ($c$), while formulas (\emph{b}) are
-built using logical connectives from atomic formulas (\emph{a}) based
-on the familiar comparison operators.
+We now give an overview of a formal model for AGREE. This provides a
+setting in which we are able to relate the contract correctness
+results discussed above, obtained via model-checking, with the code
+generated from high-assurance contracts.  The syntax of AGREE is
+essentially that of quantifier-free first order predicate logic
+supplemented with a few temporal operators. The terms (\emph{e}) are
+arithmetic expressions built from variables ($v$) and numeric and
+boolean literals ($c$), while formulas (\emph{b}) are built using
+logical connectives from atomic formulas (\emph{a}) based on the
+familiar comparison operators.
 \[
@@
 and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
-delay $(-) \to (-)$, and $\konst{Hist}(-)$.
+\emph{delay} $(-) \to (-)$, and $\konst{Hist}(-)$.
 
@@
 binding variable names to value streams, the semantics $\sem{-}^E_t$
-of terms and formulas, omitting the temporal operators, is
-conventional. The value of a variable $v$ at time $t$ is found by
-looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
-the value of $s$ at time $t$.
-
+of terms and formulas defines the meaning of compound syntax in terms
+of the meaning of subexpressions. The value of a variable $v$ at time
+$t$ is found by looking up the stream bound to $v$ in $E$ (call it
+$s$) and returning $s_t$. Some clauses of the semantics follow,
+omitting the temporal operators:
 \[
@@
 \subsection{Code contracts}
+\label{code-contracts}
 
-Generally, AGREE specifications do not describe the computation
-that a component performs. This is entirely by design: AGREE is
-intended to reason about component behavior solely at the
-specification level. However, the syntax of AGREE specifications
-provides enough expressiveness to support the notion of a
-\emph{code contract}: a component contract from which an implementation
-can be extracted.
+Generally, AGREE specifications do not describe the computation that a
+component performs. This is entirely by design: AGREE is intended to
+reason about component behavior solely at the specification
+level. However, the syntax of AGREE specifications provides enough
+expressiveness to support the notion of a \emph{code contract}: a
+contract from which an implementation can be extracted. First we must
+discuss a class of guarantees---\emph{output guarantees}---which
+determine the values on all output ports of a component.
 
@@
 
-\begin{definition}[Code contract]
-A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
-if $\mathit{Eqs} \cup G \subseteq P_c$, where
+Informally, a code contract treats its \konst{eq} ``statements'' as
+defining a list of assignments to state variables, and its output
+guarantees as directives for producing output.
+
+\begin{definition}[Code contract] A
+  leaf component of the form $(I,O,A,P,\emptyset)$ is a
+  \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where
 \[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
 of \konst{eq} statements and $G$ is the set of output guarantees, one for
-each element of $O_c$. In the interpretation as code, the order of
+each element of $O$. In the interpretation as code, the order of
 elements of $\mathit{Eqs}$ is important, and is simply taken to be the

### 2022-03-23 paper/high-assurance-contracts.tex
re-re-org

@@
 %% \end{figure}
+
+\subsection{Code contracts}
+\label{code-contracts}
+
+Generally, AGREE specifications do not describe the computation that a
+component performs. This is entirely by design: AGREE is intended to
+reason about component behavior solely at the specification
+level. However, the syntax of AGREE specifications provides enough
+expressiveness to support the notion of a \emph{code contract}: a
+contract from which an implementation can be extracted. First we must
+discuss a class of guarantees---\emph{output guarantees}---which
+determine the values on all output ports of a component.
+
+\begin{definition}[Output guarantee]
+An \emph{output guarantee} is a stylized guarantee that fully
+specifies the data written to an output port. There are three
+possibilities according to whether the output port $p$ is
+a \konst{data} port, an \konst{event} port, or an \konst{event data}
+port:
+\[
+\begin{array}{ll}
+\konst{data}: &  p = \mathit{e} \\
+\konst{event}: &  \konst{event} (p) = \mathit{b} \\
+\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
+\end{array}
+\]
+\end{definition}
+
+Informally, a code contract treats its \konst{eq} ``statements'' as
+defining a list of assignments to state variables, and its output
+guarantees as directives for producing output.
+
+\begin{definition}[Code contract] A
+  leaf component of the form $(I,O,A,P,\emptyset)$ is a
+  \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where
+\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
+of \konst{eq} statements and $G$ is the set of output guarantees, one for
+each element of $O$. In the interpretation as code, the order of
+elements of $\mathit{Eqs}$ is important, and is simply taken to be the
+occurrence order of the \konst{eq} statements in the syntax. Thus we
+will work with the
+\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
+\end{definition}
+
+The semantics of Section \ref{agree-semantics} supports a formal
+connection between the original contract (and verification results of
+the AGREE model checker being run on it), and code generated from
+specifications of leaf level cyber-components. It also provides the
+root meaning at the base of a chain of translation steps moving from
+AGREE specifications to CakeML executables.
+
+The first step in the chain maps the contract to a code-focused
+representation. Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup
+P,\emptyset)$, environment $E$, and time $t$. The evaluation of $\mathit{Eqs}$
+followed by the evaluation of $G$ results in a new environment $E'$
+where the stream values for state variables and outputs have been
+computed for time $t$. The step from $E$ to $E'$ is one full cycle in
+the repeated evaluation of the component.
+
+\begin{definition}[Evaluation]
+We overload existing notation and write the evaluation of
+$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
+$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
+an environment transformation
+\[
+ \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
+\]
+modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
+{\Eqs} is evaluated by folding the transformation left-to-right
+through it. The transformation is also applied to the list of output
+guarantees $G$, computing the values on output ports. The details are
+omitted, being a bit messy because of the different kinds of output
+ports available.
+\end{definition}
+
+\begin{definition}[Code contract correctness]
+Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
+The contract is \emph{correct}, if for all $E$ and $t$, whenever the
+assumptions (historically) hold in $E$ and evaluation steps from $E$
+to $E'$, then the guarantees hold in $E'$:
+\[
+\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
+\]
+\end{definition}
+\footnote
+{KLS: At this point we would like to assert that the AGREE ``consistency
+ check'' for a contract implies code contract correctness.}
+
+How does this notion of correctness integrate with the verification
+conditions addressed by model checking? In \emph{contract
+correctness}, the model checker is proving the following property
+\[
+\konst{Hist}(A) \imp P
+\]
+and in \emph{code contract correctness} we essentially prove a Hoare
+triple of the form
+\[
+\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
+\]
+This accords with intution, namely that AGREE is a kind of
+`program-free program logic'; by pulling out a program $(\mathit{Eqs}
+\cdot G)$ from a code contract we find ourselves in the tepid
+bathwater of imperative program verification. In order to use the code
+contract result in system contract verifcation conditions, one has to
+show that the code contract result implies the system verification
+condtion.
+
+\subsubsection*{Wellformedness}
+The evaluation semantics given above is based on the stream
+computation model, where values `deep in the past' can be accessed in
+computations. However, the conventional program languages we target
+target do not support such a view. Thus code generation is only
+defined for a class of \emph{wellformed} equations.
+
+\begin{definition}[Wellformedness]
+\end{definition}
+
+\subsection{Code contracts for the hardened system}
 

### 2022-03-23 paper/synthesis.tex
re-re-org

--- 2021-10-13 paper/synthesis.tex
@@
+\emph{Discussion on mapping program from stream environment to logic
+function and then to CakeML}
+
 Synthesis maps from model and specifications to code. The synthesis

### 2022-03-23 paper/verification-conditions.tex
re-re-org

@@
 contracts, self-consistent.
-
-\subsection{Syntax and semantics of AGREE}
-\label{agree-semantics}
-
-We now give an overview of a formal model for AGREE. This provides a
-setting in which we are able to relate the contract correctness
-results discussed above, obtained via model-checking, with the code
-generated from high-assurance contracts.  The syntax of AGREE is
-essentially that of quantifier-free first order predicate logic
-supplemented with a few temporal operators. The terms (\emph{e}) are
-arithmetic expressions built from variables ($v$) and numeric and
-boolean literals ($c$), while formulas (\emph{b}) are built using
-logical connectives from atomic formulas (\emph{a}) based on the
-familiar comparison operators.
-\[
-\begin{array}{rcl}
-e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
-a & ::= & e\; \set{=,<}\; e \\
-b & ::= & v \mid c \mid a \mid \neg b
-            \mid b \; \set{\land,\lor,\imp,\iff}\; b
-\end{array}
-\]
-
-There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
-\emph{delay} $(-) \to (-)$, and $\konst{Hist}(-)$.
-
-The semantics of terms and formulas is in terms of \emph{streams of
-values}. Values encompass at least booleans and numbers, but can be
-readily extended to include records and arrays. A value stream is a
-total function from time (natural numbers) to values:
-\[
- \konst{stream} = \mathbb{N} \to \konst{value}
-\]
-Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
-binding variable names to value streams, the semantics $\sem{-}^E_t$
-of terms and formulas defines the meaning of compound syntax in terms
-of the meaning of subexpressions. The value of a variable $v$ at time
-$t$ is found by looking up the stream bound to $v$ in $E$ (call it
-$s$) and returning $s_t$. Some clauses of the semantics follow,
-omitting the temporal operators:
-\[
-\begin{array}{rcl}
-\sem{v}^E_t & = & E(v)(t) \\
-\sem{c}^E_t & = & c \\
-\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
-   & \cdots & \\
-\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
-   & \cdots & \\
-\end{array}
-\]
-
-The temporal operators deal with time in more significant ways. The
-value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
-$t-1$ (at time zero, \konst{pre} is undefined).  A delay $e_1 \to e_2$
-temporally ``shifts'' $e_2$ by means of prepending the first element
-of $e_1$ to it.
-
-\[
-\begin{array}{rcl}
-\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
-\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
-\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
-\end{array}
-\]
-
-Although basic, these definitions can be used to define higher-level
-operators from PTLTL, such as \konst{Once} and \konst{Since}.
-
-\subsection{Code contracts}
-\label{code-contracts}
-
-Generally, AGREE specifications do not describe the computation that a
-component performs. This is entirely by design: AGREE is intended to
-reason about component behavior solely at the specification
-level. However, the syntax of AGREE specifications provides enough
-expressiveness to support the notion of a \emph{code contract}: a
-contract from which an implementation can be extracted. First we must
-discuss a class of guarantees---\emph{output guarantees}---which
-determine the values on all output ports of a component.
-
-\begin{definition}[Output guarantee]
-An \emph{output guarantee} is a stylized guarantee that fully
-specifies the data written to an output port. There are three
-possibilities according to whether the output port $p$ is
-a \konst{data} port, an \konst{event} port, or an \konst{event data}
-port:
-\[
-\begin{array}{ll}
-\konst{data}: &  p = \mathit{e} \\
-\konst{event}: &  \konst{event} (p) = \mathit{b} \\
-\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
-\end{array}
-\]
-\end{definition}
-
-Informally, a code contract treats its \konst{eq} ``statements'' as
-defining a list of assignments to state variables, and its output
-guarantees as directives for producing output.
-
-\begin{definition}[Code contract] A
-  leaf component of the form $(I,O,A,P,\emptyset)$ is a
-  \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where
-\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
-of \konst{eq} statements and $G$ is the set of output guarantees, one for
-each element of $O$. In the interpretation as code, the order of
-elements of $\mathit{Eqs}$ is important, and is simply taken to be the
-occurrence order of the \konst{eq} statements in the syntax. Thus we
-will work with the
-\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
-\end{definition}

### 2022-03-23 paper/agree-semantics.tex
forgot

We now give an overview of a formal model for AGREE. This provides a
setting in which we are able to relate the contract correctness
results discussed above, obtained via model-checking, with the code
generated from high-assurance contracts.  The syntax of AGREE is
essentially that of quantifier-free first order predicate logic
supplemented with a few temporal operators. The terms (\emph{e}) are
arithmetic expressions built from variables ($v$) and numeric and
boolean literals ($c$), while formulas (\emph{b}) are built using
logical connectives from atomic formulas (\emph{a}) based on the
familiar comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
\emph{delay} $(-) \to (-)$, and $\konst{Hist}(-)$.

The semantics of terms and formulas is in terms of \emph{streams of
values}. Values encompass at least booleans and numbers, but can be
readily extended to include records and arrays. A value stream is a
total function from time (natural numbers) to values:
\[
 \konst{stream} = \mathbb{N} \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}^E_t$
of terms and formulas defines the meaning of compound syntax in terms
of the meaning of subexpressions. The value of a variable $v$ at time
$t$ is found by looking up the stream bound to $v$ in $E$ (call it
$s$) and returning $s_t$. Some clauses of the semantics follow,
omitting the temporal operators:
\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

The temporal operators deal with time in more significant ways. The
value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
$t-1$ (at time zero, \konst{pre} is undefined).  A delay $e_1 \to e_2$
temporally ``shifts'' $e_2$ by means of prepending the first element
of $e_1$ to it.

\[
\begin{array}{rcl}
\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
\end{array}
\]

Although basic, these definitions can be used to define higher-level
operators from PTLTL, such as \konst{Once} and \konst{Since}.

### 2022-03-23 paper/verification-conditions.tex
Unify the temporal operators and clarify PLTL

@@
-\newcommand{\globally}{\ensuremath{\mathbf{G}}}
-\newcommand{\historically}{\ensuremath{\mathbf{H}}}
+\newcommand{\globally}{\konst{Always}}
+\newcommand{\historically}{\konst{Hist}}
 \newcommand{\assumes}{\ensuremath{A}}
 \newcommand{\guarantees}{\ensuremath{P}}
-\newcommand{\dispatch}{\ensuremath{\mathit{dispatch}}}
-\newcommand{\complete}{\ensuremath{\mathit{complete}}}
-\newcommand{\same}[1]{\ensuremath{\mathit{same}(#1)}}
 \newcommand{\inputs}{\ensuremath{I}}
 \newcommand{\outputs}{\ensuremath{O}}
-\newcommand{\system}{\ensuremath{S}}
 \newcommand{\components}{\ensuremath{C}}
 \newcommand{\component}{\ensuremath{c}}
-\newcommand{\schedule}{\ensuremath{\phi}}
-\newcommand{\valid}{\ensuremath{\mathit{valid}}}
-\newcommand{\dpred}{\ensuremath{\delta^\phi}}
-\newcommand{\dispred}{\ensuremath{\mathbb{D}^\phi}}
-\newcommand{\compred}{\ensuremath{\mathbb{C}^\phi}}
-\newcommand{\dispredp}{\ensuremath{\mathbb{D}^{\phi\prime}}}
-\newcommand{\compredp}{\ensuremath{\mathbb{C}^{\phi\prime}}}
 
-The AGREE specification language is based on stream concepts, and
-operators, from the Lustre language \cite{10.1145/41625.41641}. Thus
-the setting is synchronous dataflow where the inputs and outputs of
-components are streams, and contracts express relationships between
-input and output streams. When considering a system of components,
+The AGREE specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
+Lustre reasons over a synchronous dataflow model where the inputs and outputs of
+components are streams.
+Informally, streams assign values to expressions at each point in time.
+AGREE contracts express relationships between input and output streams by assuming a set of allowed input streams and then guaranteeing a corresponding set of allowed output streams.
+Output for streams not included in the assumed input is undefined.
+
+When considering a system of components,
 data flows through the components in dependency order, with inputs
 being propagated to outputs through all contracts until they stabilize
-(can't propagate further). Therefore, the subcomponent contracts, and
-thus the top-level model, must be acyclic. (An apparent syntactic
-cycle, where a component is linked back to itself, may be broken
-temporally by inserting delay elements.)  Once the data propagation
+(i.e., they cannot propagate further). Therefore, the subcomponent contracts, and
+thus the top-level model, must be acyclic. 
+Cyclic systems are made acyclic by breaking cycles with delays.
+That is to say that input at the cut point reasons about previous input data rather than the current input data.
+Once the data propagation
 has stabilized, the model proceeds to the next input data in the input
-streams. The semantics do not model computation or communication
+streams. It is worth noting that the semantics do not model computation or communication
 delay. The output of one contract is seen at the input of any
@@
 is then invoked to prove or disprove the verification
-conditions. Contracts and verification conditions are expressed in
-\emph{past-time linear temporal logic} (PLTL).\footnote{KLS: citation
-needed.}  PLTL is a logic enhanced with temporal operators able to
-reason about the truth values of formulas through time.  Its semantics
+conditions. The contracts and verification conditions are expressed in
+\emph{past-time linear temporal logic} (PLTL) \cite{10.1093/jigpal/8.1.55}.
+
+PLTL is a logic enhanced with temporal operators able to
+reason about the truth values of formulas through time.
+It combines past-time operators that reason over the history of computation up to a point (PTLTL) and future-time operators that reason over the computation that follows that point (LTL).
+Its semantics
 are defined relative to a point in time $i$ and a finite trace of
@@
 The two PLTL operators necessary for the AGREE generated verification
-conditions are $\globally$ (globally) that looks forward in time along
+conditions are $\globally$ that looks forward in time along
 the trace and $\historically$ (historically) that looks backward in
@@
 
-A \emph{system} $\system = (\inputs, \outputs, \assumes,
+A \emph{system} is a tuple $(\inputs, \outputs, \assumes,
 \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the
@@
 $\component$ in the system, and $\mathbb{O}$ is defined to be the set
-of components that provide the output for the system.  A system $S$ is
+of components that provide the output for the system.
+
+A system $S$ is
 \emph{correct} if and only if for all components $c \in C$ the
 following two verification conditions hold:
-\begin{equation}
+\begin{equation}\label{eq:assumes}
             \globally(\historically(\assumes \wedge
@@
 \end{equation}
-\begin{equation}
+\begin{equation}\label{eq:guarantees}
             \globally(\historically(\assumes \wedge
@@
 \end{equation}
-Condition (1) verifies the input assumptions on each component under
-the system assumptions and upstream component guarantees.  It checks
+The conditions from \eqref{eq:assumes} verify the input assumptions on each component under
+the system assumptions and upstream component guarantees.  They check
 if the component guarantees and system assumptions are strong enough
 to imply input assumptions on all immediate downstream components.
-Condition (2) checks the output guarantees of the system under the
+The condition in \eqref{eq:guarantees} checks the output guarantees of the system under the
 system assumptions and component guarantees that provide the output.
@@
 \guarantees)$ holds for the system contract.
+In this way the Liskhov substitution principle applies meaning that the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
 

### 2022-03-24 paper/example.tex
Add citations to example section.

@@
 The waypoint manager (WM) receives the mission command as a set of
-waypoints from the planner and starts the UAV flying the mission,
+waypoints from the planner and starts the UAV flying the mission with an event on the \texttt{start} output port,
 issuing waypoints to the UAV flight controller as the UAV location
 changes.  The waypoint manager is a legacy component that cannot be
-modified.
+modified so it critically relies on assumptions about its input behavior to guarantee its intended output behavior.
 
@@
 specifications.  The contracts constrain input with assumptions and
-state properties of output with guarantees.  AGREE performs model
+state properties of output with guarantees if the assumptions are satisfied.  AGREE performs model
 checking on this assume-guarantee system to hierarchically prove that
-SW obeys all contract obligations under all possible finite input
+the SW component obeys all contract obligations under all possible finite input
 streams.  The AGREE analysis, contract language, and contracts for SW
@@
 for SW ensure that
-\begin{itemize}
-\item a start includes a waypoint,\footnote{KLS: maybe describe the start port earlier.}
-\item a start is within one cycle of an automation request and if not, then it persistently alerts,
-\item new waypoints coincide with air vehicle location updates, and
+\begin{compactitem}
+\item a start includes a waypoint;
+\item a start is within one cycle of an automation request and if not, then it persistently alerts;
+\item new waypoints coincide with air vehicle location updates; and
 \item all outputs are well-formed.
-\end{itemize}
+\end{compactitem}
 
@@
 contract guarantees it only responds to automation requests and always
-in the same cycle.  The legacy WM contract guarantees the following: a
-start from a response and always in the same cycle; start always comes
-with a waypoint; and, further waypoints coincide with air vehicle
-location updates.  These initial specifications pass AGREE
+in the same cycle.  The legacy WM contract guarantees the following if its input assumptions hold: 
+\begin{compactitem}
+  \item it generates a start from a response from the AI and always in the same cycle; 
+  \item the start always comes with a waypoint; and 
+  \item any further waypoints coincide with air vehicle location updates.  
+\end{compactitem}
+These initial specifications pass AGREE
 verification meaning that the contract composition of the components
@@
 high-assurance components in the form of a filter and a monitor, as
-shown in \figref{fig:hardened}.  A filter enforces an invariant over
+shown in \figref{fig:hardened}, to isolate the AI component and prevent it from violating the WM component input assumptions.
+A filter enforces an invariant over
 each datum in the data stream by not forwarding input to its output if
@@
 A monitor captures a relation on input data over time and is thus able
-to reason about temporal properties of that input.  A monitor raises
-an alert if the specified temporal properties are ever violated.  The
+to reason about the temporal, and other invariant, properties of that input.  A monitor raises
+an alert if the specified properties are ever violated.  The
 AGREE specification for the monitor in our example states that an
@@
 the system designer provides the policy, and that policy is typically
-based on existing AADL/AGREE definitions in the SW system.
+based on existing the assumptions and guarantees of the SW system and its implementing components.
 
@@
 additional assumptions on the deployed system:
-\begin{itemize}
+\begin{compactitem}
 \item contracts for non-synthesized components accurately model their deployed
@@
   with the schedule.
-\end{itemize}
+\end{compactitem}
 \noindent Support and automation for these aspects of the design process are
-discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}
+discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.

### 2022-03-24 paper/verification-conditions.tex
Add minor revisions to section 3

@@
 The AGREE specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
-Lustre reasons over a synchronous dataflow model where the inputs and outputs of
+Lustre reasons over a synchronous data-flow model where the inputs and outputs of
 components are streams.
@@
 Its semantics
-are defined relative to a point in time $i$ and a finite trace of
-system states $\pi = s_0, s_1, \ldots, s_i$.
+are defined relative to a point in time $i$, a finite trace of
+system states $\pi = s_0, s_1, \ldots, s_i$, and some logic formula $f$.
 
@@
 output set, $\assumes$ is the set of assumptions, $\guarantees$ is the
-set of guarantees, and $C$ are subcomponents.  A subcomponent
+set of guarantees, and $C$ are subcomponents.  Assumptions and guarantees only use the past-time operators from PLTL. A subcomponent
 $\component$ is, hierarchically, also a system, and may be designated
@@
 
-A system $S$ is
+A system is
 \emph{correct} if and only if for all components $c \in C$ the
@@
 \guarantees)$ holds for the system contract.
-In this way the Liskhov substitution principle applies meaning that the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
+Therefore the Liskhov substitution principle applies, and as such, the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
 

### 2022-03-24 paper/high-assurance-contracts.tex
forgot unwritten file
wellformedness discussion (partial)

@@
 
-\subsection{Code contracts}
-\label{code-contracts}
-
-Generally, AGREE specifications do not describe the computation that a
+Generally, AGREE contracts do not describe the computation that a
 component performs. This is entirely by design: AGREE is intended to
 reason about component behavior solely at the specification
-level. However, the syntax of AGREE specifications provides enough
-expressiveness to support the notion of a \emph{code contract}: a
-contract from which an implementation can be extracted. First we must
-discuss a class of guarantees---\emph{output guarantees}---which
-determine the values on all output ports of a component.
+level. However, the syntax of AGREE provides enough expressiveness to
+support the notion of a \emph{code contract}: a contract from which an
+implementation can be extracted. First we must discuss a class of
+guarantees---\emph{output guarantees}---which determine the values on
+all output ports of a component.
 
@@
 The semantics of Section \ref{agree-semantics} supports a formal
-connection between the original contract (and verification results of
-the AGREE model checker being run on it), and code generated from
+connection between the original contract---and verification results of
+the AGREE model checker being run on it---and code generated from
 specifications of leaf level cyber-components. It also provides the
 root meaning at the base of a chain of translation steps moving from
-AGREE specifications to CakeML executables.
+an AGREE contract to a CakeML executables.
 
 The first step in the chain maps the contract to a code-focused
-representation. Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup
-P,\emptyset)$, environment $E$, and time $t$. The evaluation of $\mathit{Eqs}$
-followed by the evaluation of $G$ results in a new environment $E'$
-where the stream values for state variables and outputs have been
-computed for time $t$. The step from $E$ to $E'$ is one full cycle in
-the repeated evaluation of the component.
+representation. Assume given a code contract
+$(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$, environment $E$, and
+time $t$. The \emph{evaluation} of $\mathit{Eqs}$ followed by the
+evaluation of $G$ results in a new environment $E'$ where the stream
+values for state variables and outputs have been computed for time
+$t$. The step from $E$ to $E'$ is one full cycle in the repeated
+evaluation of the component.
 
@@
 $\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
-$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
-an environment transformation
+$\sem{\mathit{Eqs}\cdot G}^E_t$. The evaluation of $v = e \in {\Eqs}$
+is an environment transformation
 \[
@@
 modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
-{\Eqs} is evaluated by folding the transformation left-to-right
-through it. The transformation is also applied to the list of output
+The list {\Eqs} is evaluated by folding the transformation left-to-right
+through it. The transformation is similarly applied to the list of output
 guarantees $G$, computing the values on output ports. The details are
@@
 \]
-This accords with intution, namely that AGREE is a kind of
+This accords with intuition, namely that AGREE is a kind of
 `program-free program logic'; by pulling out a program $(\mathit{Eqs}
 \cdot G)$ from a code contract we find ourselves in the tepid
-bathwater of imperative program verification. In order to use the code
-contract result in system contract verifcation conditions, one has to
-show that the code contract result implies the system verification
-condtion.
+bathwater of imperative program verification. For example, at this
+point one could apply a verification condition generator to help break
+down the proof obligation. In order to use the code contract result in
+system contract verifcation conditions, one has to show that the code
+contract result implies the system verification condtion.
 
@@
 The evaluation semantics given above is based on the stream
-computation model, where values `deep in the past' can be accessed in
-computations. However, the conventional program languages we target
-target do not support such a view. Thus code generation is only
-defined for a class of \emph{wellformed} equations.
+computation model, where values arbitrarily `deep in the past' can be
+accessed in computations by use of $\konst{pre}(-)$. However, the
+conventional program languages we target do not support such a
+view. We are currently examining transformations that compile such
+deep temporal accesses away by introducing intermediate
+variables. Avoiding discussion of the algorithm, we focus on the
+notion of \emph{wellformed} equations that needs to hold. The guiding
+intuition is that an \emph{implementable} {\Eqs} can allow access at
+most one step in the past, and accesses of a past value for a variable
+are only possible until the variable is updated. In lieu of a formal
+definition, we give a few examples.
 
 \begin{definition}[Wellformedness]
+
 \end{definition}

### 2022-03-25 paper/high-assurance-contracts.tex
finalized wellformedness discussion

@@
 \end{definition}
-\footnote
-{KLS: At this point we would like to assert that the AGREE ``consistency
- check'' for a contract implies code contract correctness.}
 
 How does this notion of correctness integrate with the verification
-conditions addressed by model checking? In \emph{contract
-correctness}, the model checker is proving the following property
+conditions generated by AGREE? In \emph{contract correctness}, the
+model checker is proving the following property
 \[
@@
 \subsubsection*{Wellformedness}
-The evaluation semantics given above is based on the stream
-computation model, where values arbitrarily `deep in the past' can be
-accessed in computations by use of $\konst{pre}(-)$. However, the
-conventional program languages we target do not support such a
-view. We are currently examining transformations that compile such
-deep temporal accesses away by introducing intermediate
-variables. Avoiding discussion of the algorithm, we focus on the
+Stream-based evaluation allows values arbitrarily `deep in the past'
+to be accessed by use of $\konst{pre}(-)$. However, the programming
+languages we target do not directly support streams.  We are currently
+developing a means of translating away such deep temporal accesses by
+the introduction of intermediate variables. For now, we focus on the
 notion of \emph{wellformed} equations that needs to hold. The guiding
-intuition is that an \emph{implementable} {\Eqs} can allow access at
-most one step in the past, and accesses of a past value for a variable
-are only possible until the variable is updated. In lieu of a formal
-definition, we give a few examples.
-
-\begin{definition}[Wellformedness]
-
-\end{definition}
+intuition is that a wellformed {\Eqs} allows access at most one step
+in the past, and accesses of a past value for a variable are only
+possible until the variable is updated. In informal terms, this means
+that ${\Eqs}\cdot G$ in a code contract must be such that:
+\begin{itemize}
+\item \konst{fby} only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
+\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
+\item $\konst{pre}(v)$ is allowed before the defining equation $v = e$, and may occur in $e$.
+\end{itemize}
+For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
+defined, albeit in a non-wellformed manner, by
+{\small
+\begin{verbatim}
+  Fib = 1 -> pre(1 -> Fib + pre Fib)
+\end{verbatim}
+}
+\noindent or by (again non-wellformed)
+{\small
+\begin{verbatim}
+  N = 0 -> 1 + pre(N)
+  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
+\end{verbatim}
+}
+\noindent or by (wellformed):
+{\small
+\begin{verbatim}
+  N = 0 -> 1 + pre(N)
+  A = 42 -> pre(Fib)
+  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
+\end{verbatim}
+}
+A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
+conventional programming languages, as we will see in
+Section \ref{sec:synthesis}.
 
@@
 \begin{lrbox}{\flt}
-  \begin{lstlisting}[style=agree,numbers=left]
-    -- start user definitions
-    eq policy : bool =
-      WELL_FORMED_AUTOMATION_RESPONSE(Input);
-    -- stop user definitions
+  \begin{lstlisting}[style=agree,numbers=left] -- start user
+  definitions eq policy : bool =
+  WELL_FORMED_AUTOMATION_RESPONSE(Input); -- stop user definitions
 

### 2022-03-25 paper/agree-semantics.tex
Minor edits in section 4

@@
-We now give an overview of a formal model for AGREE. This provides a
-setting in which we are able to relate the contract correctness
-results discussed above, obtained via model-checking, with the code
-generated from high-assurance contracts.  The syntax of AGREE is
+\begin{comment}
+This section gives an abstract mathematical overview of AGREE's
+compositional reasoning system, defining assume-guarantee contracts
+and the verification conditions that AGREE checks in order to
+modularly establish overall system correctness at the model level.
+The syntax and semantics of AGREE are sketched in some detail.  The
+discussion then becomes more concrete, showing---by example---how the
+AGREE contract language is used to specify the cyber-hardened system
+of \figref{fig:hardened}. The discussion proceeds from the
+system-level contract to the filter and monitor components, which rely
+on the notion of a \emph{code contract} to support code generation.
+\end{comment}
+
+We now give an overview of a formal model for AGREE syntax and semantics.
+The discussion then becomes more concrete, showing by example, how the AGREE language is used to specify the properties of the SW system introduced in \secref{sec:example}.
+Any implementation of the SW system must guarantee the specification output under the system assumptions.
+
+The syntax of AGREE is
 essentially that of quantifier-free first order predicate logic
@@
 \]
+There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
+and formulas. Lastly, there are temporal operators $\konst{pre}(-)$ (i.e., \emph{yesterday} in PLTL),
+\emph{followed-by} $(-) \to (-)$, and $\konst{Hist}(-)$.
 
-There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
-\emph{delay} $(-) \to (-)$, and $\konst{Hist}(-)$.
-
-The semantics of terms and formulas is in terms of \emph{streams of
+The semantics of terms and formulas, as mentioned in the previous section, is in terms of \emph{streams of
 values}. Values encompass at least booleans and numbers, but can be
 readily extended to include records and arrays. A value stream is a
-total function from time (natural numbers) to values:
+total function from time (natural numbers including 0) to values:
 \[
- \konst{stream} = \mathbb{N} \to \konst{value}
+ \konst{stream} = \mathbb{N}_0 \to \konst{value}
 \]
+
 Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
@@
 value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
-$t-1$ (at time zero, \konst{pre} is undefined).  A delay $e_1 \to e_2$
-temporally ``shifts'' $e_2$ by means of prepending the first element
-of $e_1$ to it.
-
+$t-1$ (at time zero, \konst{pre} is undefined).  
+A followed-by, $e_1 \to e_2$, evaluates to $e_1$ at time 0; otherwise it is $e_2$.
+Followed-by is often used to give a value to a stream in the first step so that subsequent steps are able to use $\konst{pre}(e)$.
 \[
@@
 operators from PTLTL, such as \konst{Once} and \konst{Since}. 
+\[
+   \begin{array}{rcl}
+      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(\konst{Once}(b)))}^E_t
+   \end{array}
+\]

### 2022-03-25 paper/paper.tex
re-re-org
Unify the temporal operators and clarify PLTL
Minor edits in section 4

@@
 \newcommand{\linesref}[2]{Lines~\ref{#1}--\ref{#2}}
+% \newcommand{\eqref}[1]{Eq.~\ref{#1}}
 
@@
 \journalname{Software and Systems Modeling}
+
+\newif\ifREVISIONS
+\REVISIONSfalse
 
@@
 
+\ifREVISIONS
 \subsection{SOSYM Journal Details}
@@
 Also, make sure to mention the MODELS 2021 special issue in the cover letter.
+\fi
 
+\ifREVISIONS
 \subsection{Revisions}
@@
 \end{compactitem}
+\fi
 
@@
 
-\subsection{Revisions}
-\begin{compactitem}
-  \item \sout{Reformat text to put single sentence on each line.}
-  \item Move example to its own repository, archive, and give the DOI and URL for the example.
-  \item \sout{Update AGREE certificates with new model}
-  \item Separate example into separate files for initial design, cyber-components and tests, hardened design, and seL4 process design.
-  \item \sout{Make clear the initial properties for the assumed correct model:}
-    \begin{compactitem}
-      \item \sout{Assume at the system level that a new automation request is not sent unless the prior request, if any, is acknowledged}
-      \item \sout{Assume at all levels input messages are well-formed}
-      \item \sout{Guarantee at all levels that waypoint outputs coincide with air vehicle location messages}
-      \item \sout{Guarantee that a start from the waypoint manager coincides with an automation response}
-      \item \sout{Guarantee for all trusted components that outputs are well-formed}
-    \end{compactitem}
-  \item Add citations, and remove KLS footnote to last paragraph---should cite HAMR, seL4, and Adventium.
-\end{compactitem}
-
 \section{Compositional reasoning}
 \label{sec:agree}
-\input{agree}
+\input{verification-conditions}
 
+\section{Syntax and semantics of AGREE}
+\label{agree-semantics}
+\input{agree-semantics}
+
+\subsection{Formal contract for the example system}
+\input{sw-contract}
+
+\ifREVISIONS
 \subsection{Revisions}
@@
   \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
-  \item Add reference to the Liskov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.
+  \item \sotu{Add reference to the Liskov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.}
   \item Add back in a few sentences on the contract for the filter since those have been removed with the addition of the formal definitions.
 \end{compactitem}
+\fi
 
-\section{Formal contracts for the example system}
-\input{sw-contract}
+\section{Code contracts for cyber components}
+\label{code-contracts}
+\input{high-assurance-contracts}
 
-\subsection{Code contracts for cyber components}
-\input{high-assurance-contracts}
+\section{Testing Specifications with AGREE}
+\label{sec:testing}
+\input{testing}
 
@@
 
+\ifREVISIONS
 \subsection{Revisions}
@@
 \end{compactitem}
-
-\section{Testing Specifications with AGREE}
-\label{sec:testing}
-\input{testing}
+\fi
 

### 2022-03-25 paper/high-assurance-contracts.tex
checkpointing

@@
 to be accessed by use of $\konst{pre}(-)$. However, the programming
-languages we target do not directly support streams.  We are currently
-developing a means of translating away such deep temporal accesses by
-the introduction of intermediate variables. For now, we focus on the
-notion of \emph{wellformed} equations that needs to hold. The guiding
-intuition is that a wellformed {\Eqs} allows access at most one step
-in the past, and accesses of a past value for a variable are only
-possible until the variable is updated. In informal terms, this means
-that ${\Eqs}\cdot G$ in a code contract must be such that:
+languages being targeted by SPLAT do not directly support streams.  We
+are currently developing a means of translating away such deep
+temporal accesses by the introduction of intermediate variables. For
+now, we focus on the notion of \emph{wellformed} equations that needs
+to hold. The guiding intuition is that a wellformed {\Eqs} allows
+access at most one step in the past, and accesses of a past value for
+a variable are only possible until the variable is updated. In
+informal terms, this means that ${\Eqs}\cdot G$ in a code contract
+must be such that:
 \begin{itemize}
@@
 \item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
-\item $\konst{pre}(v)$ is allowed before the defining equation $v = e$, and may occur in $e$.
+\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
 \end{itemize}
@@
   \end{center}
-  \caption{Code contract specification for high-assurance filter.}
+  \caption{Code contract for high-assurance filter.}
   \label{fig:filter}
@@
   \end{center}
-  \caption{Code contract specification for high-assurance monitor.}
+  \caption{Code contract for high-assurance monitor.}
   \label{fig:monitor}

### 2022-03-25 paper/synthesis.tex
checkpointing

@@
-\emph{Discussion on mapping program from stream environment to logic
-function and then to CakeML}
+Synthesis maps from model and specifications to code. The SPLAT tool
+traverses the system architecture looking for occurrences of high
+assurance components specified by code contracts; for each such
+occurrence it generates a CakeML program. Starting from the semantics
+of the code contract described in Section \ref{sec:code-contracts}
+a sequence of semantics-preserving steps are taken:
+\begin{itemize}
+\item $\sem{{\Eqs}\cdot G}$ yields a function on arbitrary depth streams
+\item transform to function over depth-1 streams (wellformedness)
+\item transform to logic function
+\item attach buffer handling to inputs and outputs (contig type parsing)
+\item translate to CakeML
+\item run CakeML compiler
+\end{itemize}
 
-Synthesis maps from model and specifications to code. The synthesis
-algorithm traverses the system architecture looking for occurrences of
-filter and monitor specifications;  for each such occurrence it
-generates a CakeML program. In the following, we examine both filter
-and monitor synthesis. The latter is typically much more involved, and
-we will therefore devote more attention to it.
+In the following, we examine further details in both filter and
+monitor synthesis.
 

### 2022-03-25 paper/agree-semantics.tex
Add minor revisions to section 4

@@
 Although basic, these definitions can be used to define higher-level
-operators from PTLTL, such as \konst{Once} and \konst{Since}. 
+past-time operators from PLTL such as \konst{Once} and \konst{Since}.
+These are defined relative to their corresponding bound output stream $o$.
 \[
    \begin{array}{rcl}
-      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(\konst{Once}(b)))}^E_t
+      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(o))}^E_t \\
+      \sem{\konst{Since}(a, b)}^E_t & = & \sem{b \vee (a and (\mathit{false} \to \konst{pre}(o)))}^E_t \\
    \end{array}
 \]
+$\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the current moment.
+$\konst{Since}(a, b)$ is true if $a$ has been true in all moments up to the present one since $b$ most recently became true.
+These both refer the their bound output stream $o$ to know if they were true or false in their previous step.
+All other past-time operators of PLTL can be defined similarly.

### 2022-03-25 paper/sw-contract.tex
Add minor revisions to section 4

@@
 
+The actual surface syntax for AGREE that is embedded in OSATE is slightly different than the formal semantics in the previous section.
+The mapping between the two is straightforward.
 The AGREE specification for the SW component in the example of
@@
 The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
-Streams are associated with every expression in AGREE, meaning that expressions are defined through time, and are therefore stateful.
+The state of the contract is captured by stream semantics of AGREE.
 When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input.
-AGREE provides the \texttt{pre} operator to make that distinction.
+The \texttt{pre} naturally makes that distinction.
+In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
+Guarantees should use \texttt{pre} anytime they need to reason about the current state relative to the previous state.
 
-The \texttt{pre} operator returns the previous value of the enclosed expression by looking back one step in time on the expression's stream.
-In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
-Guarantees should use the \texttt{pre} operator anytime they need to reason about the current state relative to the previous state.
-
-The assumption on \lineref{line:sw-assume-3} uses a \emph{followed-by} operator, \texttt{->}, to \emph{guard} the \texttt{pre} operator at time 0.
-A followed-by expression defines what happens at the first instance of the stream, time 0, and then what follows after.
-Here the assumption is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
-The followed-by guards the \texttt{pre} because expressions are undefined before time 0 and the followed-by prevents the \texttt{pre} from being evaluated before time 1.
-So after time 0, the assumption depends on the presence of a request and the pre-state of the component.
-
-The stream semantics in AGREE mean that it is possible to use PLTL operators that look back in time.
-The \texttt{Historically} operator is the same as $\historically$ defined previously; it is true if the enclosed expression has been true in all moments up to the present one.
-The \texttt{Since} operator is true if the first expression has been true in all moments up to the present one since the second expression most recently became true.
+The assumption on \lineref{line:sw-assume-3} uses a followed-by expression so it is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
+The followed-by guards the \texttt{pre} as mentioned previously;
+thus after time 0, the assumption depends on the presence of a request and the pre-state of the component.
 The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}--\emph{not request since start}.
@@
 The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
-It differentiates with the followed-by operator what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
+It differentiates with the followed-by what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
 The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.
@@
 Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
-The last guarantee ensures that alert is persistent.
-The \texttt{Once} operator is true if the enclosed expression has ever been true before, or in, the current moment.
 \lineref{line:sw-guarantee-6} ensures if alert has ever happened, then it is happening in the present moment.
-These six guarantees define the behavior of the SW component under the three assumptions.
+These six guarantees define the behavior of the SW component under the three assumptions and correspond to the informal descriptions given in \secref{sec:example}.

### 2022-03-26 paper/high-assurance-contracts.tex
Add minor revisions mostly to section 5

@@
 root meaning at the base of a chain of translation steps moving from
-an AGREE contract to a CakeML executables.
+an AGREE contract to a CakeML executable.
 
 The first step in the chain maps the contract to a code-focused
-representation. Assume given a code contract
+representation. Assume a code contract
 $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$, environment $E$, and
@@
 `program-free program logic'; by pulling out a program $(\mathit{Eqs}
-\cdot G)$ from a code contract we find ourselves in the tepid
-bathwater of imperative program verification. For example, at this
+\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification. For example, at this
 point one could apply a verification condition generator to help break
 down the proof obligation. In order to use the code contract result in
-system contract verifcation conditions, one has to show that the code
-contract result implies the system verification condtion.
+system contract verification conditions, one has to show that the code
+contract result implies the system verification condition.
 
@@
 \begin{itemize}
-\item \konst{fby} only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
+\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
 \item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
@@
   \begin{lstlisting}[style=agree,numbers=left] -- start user
-  definitions eq policy : bool =
-  WELL_FORMED_AUTOMATION_RESPONSE(Input); -- stop user definitions
+    -- start user definitions 
+    eq policy : bool =
+      WELL_FORMED_AUTOMATION_RESPONSE(Input); 
+    -- stop user definitions
 
@@
 Code contracts for the filter and monitor are shown in
-\figref{fig:filter} and \figref{fig:monitor}.\footnote{KLS: do we discuss wellformedness anywhere?} In general, the
-\konst{eq}-statements of a code contract keep track of
-state and provide data values from which to compute values to place on
-the outputs.
+\figref{fig:filter} and \figref{fig:monitor}.
+The filter is direct: it defines the policy according to the well-formed requirement for data.
+We now discuss in some detail the monitor.
 
@@
 \konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
-\figref{fig:monitor}.  As with the filter, the guarantees
-completely define the output, as required for synthesis, in terms of
-the policy and the alert.  The
+\figref{fig:monitor}. 
+The
 \texttt{alert} variable defined on \lineref{line:mon-alert} depends
@@
 is set to one to be consistent with the system specification from the
-previous section.  At time 0 the policy is trivial: a response
-requires an accompanying request.  After time 0, if a request is
+previous section.  At time zero the policy is trivial: a response
+requires an accompanying request.  After time zero, if a request is
 pending in the current time step, then the latency must still be under

### 2022-03-26 paper/sw-contract.tex
Add minor revisions mostly to section 5

@@
 
-The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
-The state of the contract is captured by stream semantics of AGREE.
+The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract defined be the associated streams.
 When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input.
@@
 In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
-Guarantees should use \texttt{pre} anytime they need to reason about the current state relative to the previous state.
+Additionally, guarantees should use \texttt{pre} anytime they need to reason about the current state relative to the previous state.
 
-The assumption on \lineref{line:sw-assume-3} uses a followed-by expression so it is \texttt{true} at time 0 and then it is the truth value of the implication in all future instances.
+The assumption on \lineref{line:sw-assume-3} uses a followed-by expression so it is \texttt{true} at time zero and then it is the truth value of the implication in all future instances.
 The followed-by guards the \texttt{pre} as mentioned previously;
-thus after time 0, the assumption depends on the presence of a request and the pre-state of the component.
-The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}--\emph{not request since start}.
+thus after time zero, the assumption depends on the presence of a request and the pre-state of the component.
+The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---\emph{not request since start}.
 
@@
 The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
-It differentiates with the followed-by what is required at time 0, the start must coincide with the request, with what is required after time 0, the start must coincide with the request or is a response to a request one step earlier.
+It differentiates with the followed-by what is required at time zero, the start must coincide with the request, with what is required after time zero, the start must coincide with the request or is a response to a request one step earlier.
 The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.

### 2022-03-26 paper/testing.tex
Revise the section on testing

@@
+Before discussing the final steps in synthesis, we take a moment to discuss how code-contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
 Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
 The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.
-It is difficult to reason about the whole of the specification to understand the defined behavior of a component or system.
-A natural consequence of this inherent complexity is that the AGREE analysis proves out, but that proof is about a flawed specification.
-The AGREE analysis, and synthesized component, is only as good as the specifications.
-As the saying goes, \emph{garbage in means garbage out}.
 
-The case studies found four common mistakes in specifications: contradictions, vacuity, under-specification, and missed corner cases.
-Contradictions arise from inconsistent contracts that state conflicting constraints in any given moment invalidating the verification results: a property can be proved true and it can be proved false.
+Four common mistakes in specifications have been observed: contradictions, vacuity, under-specification, and missed corner cases.
+Contradictions arise from inconsistent contracts that state conflicting constraints in any given moment invalidating the verification results.
 Vacuity is a common pitfall with implication when the left hand side is always false making the implication true. 
-Vacuity is referred to as the \emph{block of wood} problem because a system that does nothing passes verification with implications.
+Under-specification is very challenging because verification is not able to prove anything (or it can prove everything in the case of existential properties) so nothing is actually known after verification except that the system is under-specified.
+Missed corner cases is a common problem to all disciplines.
+It is especially hard here because there is no direct way to provide test input and check test output.
 
-Under-specification is very challenging because verification is not able to prove anything (or it can prove everything in the case of existential properties) so nothing is actually known after verification except that the system is under-specified.
-The tell-tale symptom of underspecification is that AGREE verification is not able to prove anything about a property: it is not true and it is not false.
-The property stated in the positive fails and the property stated in the negative fails.
-Missed corner cases is a common problem to all disciplines.
-It is especially hard here because there is no easy way break down, and check, specifications in a case by case input analysis.
-
-AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification, and missed corner cases is always an issue.
-The case studies navigated these challenges by writing a specification, synthesizing it to code with SPLAT, and then running unit 
-tests on the synthesized code in a testing harness.
-The synthesized code is known to precisely capture the meaning of the specification, so the tests directly reflect the behavior of the specification in much the same way as compiled code is assumed correct.
-The designers are then able to iteratively test and refine the specifications into a final form.
-The process is inefficient, requires an additional, unneeded, test harness, the code is proved correct by SPLAT after all, and steps outside the MBE environment using the implementation, rather than the model, to analyze and understand the system model.
-
-There needs to be a bridge between the formal model in AGREE with its proof of system properties and more traditional reasoning with test that is a well established method for building assurance.
-Classical black-box and white-box test coverage should be, and are, equally applicable to formal reasoning as traditional engineering in regards to developing, and eventually proving, the correctness of a specification.
-AGREE is able to provide such a testing framework.
+AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification.
+The tempting approach to check these issues is to use the synthesize path and then apply traditional testing.
+Such an approach adds extra steps that are not needed as it is possible to bridge the formal model in AGREE with its proof system to traditional reasoning with test.
 
@@
 These same tests can be synthesized to the backend implementation for traditional testing.
-
-Traditional testing is important for the implementations of components that are not high-assurance.
-The behavior of regular components must be modeled in AGREE since they are part of any system implementation, and their specifications provide input to high-assurance components.
-Like the high-assurance components, these specifications must be meaningful in modeling the representative component.
-Testing is an effective means for validating such specifications, and AGREE can use the existing tests from the component implementation to check the specifications, and it can generate additional tests that probe the behavior of the component implementation when writing the specifications.

### 2022-03-26 paper/paper.tex
checkpointing
Bibliography fixes
Add minor revisions mostly to section 5

@@
 %\documentclass[global,twocolumn,referee]{svjour}
-\documentclass[global,twocolumn]{svjour}
+\documentclass[global,twocolumn,keeplastbox]{svjour}
 \usepackage{cite}
@@
 
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item \sout{Add in the AGREE specification for the system}
-  \item \sout{Add in the AGREE verification conditions (e.g., similar to those in the Scheduled Components NFM submission)}
-  \item \sotu{Add reference to the Liskov principle of safe-substitution when stating that the system contract is a sound abstraction of the implementation.}
-  \item Add back in a few sentences on the contract for the filter since those have been removed with the addition of the formal definitions.
-\end{compactitem}
-\fi
-
 \section{Code contracts for cyber components}
-\label{code-contracts}
+\label{sec:code-contracts}
 \input{high-assurance-contracts}
@@
 \input{testing}
-
-%% \section{Synthesis from code contracts}
-%% \label{sec:semantics}
-%% \input{semantics}
-
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item Add in intuitive definition or AGREE leaf-component semantics
-  \item State well-formed theorem
-  \item State correctness theorem or any key theorems to the synthesis proof
-\end{compactitem}
-\fi
 

### 2022-03-26 paper/high-assurance-contracts.tex
supposedly well-formed Fib example was not; fixed

@@
   N = 0 -> 1 + pre(N)
-  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
+  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
 \end{verbatim}
@@
   N = 0 -> 1 + pre(N)
-  A = 42 -> pre(Fib)
-  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
+  F2 = 42 -> pre(F1)
+  F1 = 42 -> pre(Fib)
+  Fib = if N <= 1 then 1 else F1 + F2
 \end{verbatim}

### 2022-03-26 paper/verification-conditions.tex

@@
 \guarantees)$ holds for the system contract.
-Therefore the Liskhov substitution principle applies, and as such, the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
+Therefore the Liskov substitution principle applies, and as such, the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
 

### 2022-03-26 paper/agree-semantics.tex
declutter defs for Once and Since; minot yidying

@@
 $t-1$ (at time zero, \konst{pre} is undefined).
-A followed-by, $e_1 \to e_2$, evaluates to $e_1$ at time 0; otherwise it is $e_2$.
+A followed-by, $e_1 \to e_2$, evaluates to $e_1$ at time zero; otherwise it is $e_2$.
 Followed-by is often used to give a value to a stream in the first step so that subsequent steps are able to use $\konst{pre}(e)$.
@@
 past-time operators from PLTL such as \konst{Once} and \konst{Since}.
-These are defined relative to their corresponding bound output stream $o$.
+These can be expressed in a recursive style as follows:
+%% \[
+%%    \begin{array}{rcl}
+%%       \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\konst{false} \to \konst{pre}(o))}^E_t \\
+%%       \sem{\konst{Since}(a, b)}^E_t & = & \sem{b \vee (a and (\konst{false} \to \konst{pre}(o)))}^E_t \\
+%%    \end{array}
+%% \]
 \[
 \begin{array}{rcl}
-      \sem{\konst{Once}(b)}^E_t & = & \sem{b \vee (\mathit{false} \to \konst{pre}(o))}^E_t \\
-      \sem{\konst{Since}(a, b)}^E_t & = & \sem{b \vee (a and (\mathit{false} \to \konst{pre}(o)))}^E_t \\
+ \konst{Once}(b) & = & b \vee \konst{pre}(\konst{Once}(b)) \\
+ \konst{Since}(a, b) & = & b \lor (a \land \konst{pre}(\konst{Since}(a, b))
    \end{array}
 \]
-$\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the current moment.
-$\konst{Since}(a, b)$ is true if $a$ has been true in all moments up to the present one since $b$ most recently became true.
-These both refer the their bound output stream $o$ to know if they were true or false in their previous step.
+
+$\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the
+current moment.  $\konst{Since}(a, b)$ is true if $a$ has been true in
+all moments up to the present one since $b$ most recently became true.
 All other past-time operators of PLTL can be defined similarly.

### 2022-03-26 paper/related-work.tex
fixed bugs in Once/Since defs; added details to comparison with the INRIA work in Related Work

@@
 
-Data-flow semantics are well studied \cite{10.1145/41625.41641,97300, 10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}.
-State machine semantics can be added to synchronous data-flow \cite{10.1145/1086228.1086261}.
-The idea is to translate imperative constructs into equivalent synchronous data flow constructs.
-The resulting Lustre can then be compiled.
-The goal is to provide a seamless connection between pure data-flow and pure control design.
+Data-flow semantics are well studied \cite{10.1145/41625.41641,97300,
+10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}. State
+machine semantics can be added to synchronous
+data-flow \cite{10.1145/1086228.1086261}. The idea is to translate
+imperative constructs into equivalent synchronous data flow
+constructs. The resulting Lustre can then be compiled. The goal is to
+provide a seamless connection between pure data-flow and pure control
+design.
 
-There is a fully verified compiler that takes Lustre and turns it into a binary executable that is specified and verified in Coq \cite{10.1145/3140587.3062358}.
-The key is in combining infinite sequences of data-flow models with incremental manipulation of memories akin to an imperative model.
-CompCert is used on the backend to create the final rendered executable.
+There is a fully verified compiler that takes Lustre and turns it into
+a binary executable that is specified and verified in
+Coq \cite{10.1145/3140587.3062358}. The key is in combining infinite
+sequences of data-flow models with incremental manipulation of
+memories akin to an imperative model. CompCert is used on the backend
+to create the final rendered executable. The work is quite similar to
+ours in providing a verified compilation chain from Lustre. However,
+there are significant differences: first, they tackle a more
+comprehensive subset of Lustre than our work. We have been able to
+work effectively with a relatively simple core language, \eg, without
+\emph{clocking} features, and this aligns with our contract language, so that, as well as generating behaviorally equivalent code, we also can show that the generated code meets a logical specification that is a leaf-level claim in a system-wide correctness argument conducted by compositional reasoning in AGREE.

### 2022-03-26 paper/high-assurance-contracts.tex
declutter defs for Once and Since; minot yidying
type setting Fib example to use lstlisting

--- 2022-03-26 paper/high-assurance-contracts.tex
@@
 The first step in the chain maps the contract to a code-focused
-representation. Assume a code contract
+representation. Assume given a code contract
 $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$, environment $E$, and
@@
 {\small
-\begin{verbatim}
+\begin{lstlisting}[style=agree]
   Fib = 1 -> pre(1 -> Fib + pre Fib)
-\end{verbatim}
+\end{lstlisting}
 }
@@
 {\small
-\begin{verbatim}
+\begin{lstlisting}[style=agree]
   N = 0 -> 1 + pre(N)
-  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
-\end{verbatim}
+  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
+\end{lstlisting}
 }
@@
 {\small
-\begin{verbatim}
+\begin{lstlisting}[style=agree]
   N = 0 -> 1 + pre(N)
-  A = 42 -> pre(Fib)
-  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
-\end{verbatim}
+  F2 = 42 -> pre(F1)
+  F1 = 42 -> pre(Fib)
+  Fib = if N <= 1 then 1 else F1 + F2
+\end{lstlisting}
 }

### 2022-03-26 paper/sw-contract.tex
Add test output figure and fix typo.

@@
 
-The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract defined be the associated streams.
+The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract.
 When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input.
-The \texttt{pre} naturally makes that distinction.
+The \texttt{pre} naturally makes that distinction by looking back one step in the stream.
 In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.

### 2022-03-26 paper/testing.tex
Corrected spelling of Liskov, a few more reference cleanups
Add test output figure and fix typo.

@@
 Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
-AGREE proves the component implements the test, and the Liskhov substitution principle holds.
+AGREE proves the component implements the test, and the Liskov substitution principle holds.
 
@@
   \end{center}
-  \caption{AGREE Unit test for high-assurance monitor.}
+  \caption{AGREE unit test for high-assurance monitor.}
   \label{fig:test}
@@
 AGREE proves the test correct.
-\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp})
+\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp}).
+\figref{fig:test-output} is the AGREE output for the test is showing that it passes.
 
@@
 These same tests can be synthesized to the backend implementation for traditional testing.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.38]{agree-test-output.png} \\
+    \end{tabular}
+  \end{center}
+  \caption{AGREE output for unit test.} 
+  \label{fig:test-output}
+\end{figure}

### 2022-03-26 paper/sw-contract.tex

--- 2022-03-26 paper/sw-contract.tex
@@
 
-The assumption on \lineref{line:sw-assume-3} constrains when a request can arrive by reasoning about the \emph{state} of the contract defined be the associated streams.
-When writing assumptions and guarantees, it is important to differentiate pre-state, before a contract updates its state, and post-state, after a contract updates its state, in response to the current input.
-The \texttt{pre} naturally makes that distinction.
-In general, assumptions regarding state should be evaluated in the pre-state of the component with the current inputs, and guarantees regarding state should be evaluated on the post-state of the component given the current inputs.
-Additionally, guarantees should use \texttt{pre} anytime they need to reason about the current state relative to the previous state.
+The assumption on \lineref{line:sw-assume-3} constrains when a request
+can arrive by reasoning about the \emph{state} of the contract defined
+by the associated streams.  When writing assumptions and guarantees,
+it is important to differentiate pre-state, before a contract updates
+its state, and post-state, after a contract updates its state, in
+response to the current input.  The \texttt{pre} naturally makes that
+distinction.  In general, assumptions regarding state should be
+evaluated in the pre-state of the component with the current inputs,
+and guarantees regarding state should be evaluated on the post-state
+of the component given the current inputs.  Additionally, guarantees
+should use \texttt{pre} anytime they need to reason about the current
+state relative to the previous state.
 

### 2022-03-26 paper/testing.tex
Revise abstract to match new case study

@@
 This AGREE test method is effective for detecting vacuity, under-specified behavior, and missed corner cases.
-It enables designers to iteratively develop specifications inside the MBE framework using AGREE, and it builds assurance that the specification reflects the desired behavior with black-box, white-box, and other test coverage metrics.
+It enables designers to iteratively develop specifications inside the MBSE framework using AGREE, and it builds assurance that the specification reflects the desired behavior with black-box, white-box, and other test coverage metrics.
 These same tests can be synthesized to the backend implementation for traditional testing.

### 2022-03-26 paper/verification-conditions.tex
Add definitions for the AADL ports

@@
 The AGREE specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
-Lustre reasons over a synchronous data-flow model where the inputs and outputs of
+Lustre semantics are synchronous data-flow where the inputs and outputs of
 components are streams.
@@
 
+A component is defined by its ports in AADL.
+There are three types of AADL ports: data ports, event data ports, and event ports.
+A data port is a sampled port: the data can be read at anytime.
+An event data port indicates when data is present with an event.
+An event port has no data and only signals the presence or absence of the event.
+Event based ports are not buffered in our application here.
+An AGREE contracts defines properties of the input to output behavior of a component's ports.
+
+A system implementation is given by is implementing components and connections between their ports.
 When considering a system of components,

### 2022-03-26 paper/agree-semantics.tex
fixed bugs in Once/Since defs; added details to comparison with the INRIA work in Related Work
Add reference to PLTL monitor work by Hauvland and Rocu

@@
 There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. Lastly, there are temporal operators $\konst{pre}(-)$ (i.e., \emph{yesterday} in PLTL),
-\emph{followed-by} $(-) \to (-)$, and $\konst{Hist}(-)$.
+and formulas. Lastly, there are temporal operators \emph{previously} $\konst{pre}(-)$,
+\emph{followed-by} $(-) \to (-)$, and historically $\konst{Hist}(-)$.
+These temporal operators are sufficient to define all other past-time operators \cite{monitor}.
 
@@
 \begin{array}{rcl}
- \konst{Once}(b) & = & b \vee \konst{pre}(\konst{Once}(b)) \\
- \konst{Since}(a, b) & = & b \lor (a \land \konst{pre}(\konst{Since}(a, b))
+\konst{Once}(b) & = & b \to b \vee \konst{pre}(\konst{Once}(b)) \\
+\konst{Since}(a, b) & = & b \to b \lor (a \land \konst{pre}(\konst{Since}(a, b))
 \end{array}
 \]
-
-$\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the
+\noindent $\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the
 current moment.  $\konst{Since}(a, b)$ is true if $a$ has been true in
 all moments up to the present one since $b$ most recently became true.
-All other past-time operators of PLTL can be defined similarly.
+All other past-time operators of PLTL can be defined similarly \cite{monitor}.

### 2022-03-27 paper/high-assurance-contracts.tex
replace wellformed by Pascalish. Doesn't sound quite right but will do for now

@@
 This accords with intuition, namely that AGREE is a kind of
-`program-free program logic'; by pulling out a program $(\mathit{Eqs}
-\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification. For example, at this
+`program-free program logic'; by pulling out a program
+$(\mathit{Eqs}\cdot G)$ from a code contract we find ourselves in the
+domain of imperative program verification. For example, at this
 point one could apply a verification condition generator to help break
@@
 
-\subsubsection*{Wellformedness}
+\subsubsection*{Imperative code contracts}
 Stream-based evaluation allows values arbitrarily `deep in the past'
 to be accessed by use of $\konst{pre}(-)$. However, the programming
-languages being targeted by SPLAT do not directly support streams.  We
-are currently developing a means of translating away such deep
-temporal accesses by the introduction of intermediate variables. For
-now, we focus on the notion of \emph{wellformed} equations that needs
-to hold. The guiding intuition is that a wellformed {\Eqs} allows
-access at most one step in the past, and accesses of a past value for
-a variable are only possible until the variable is updated. In
-informal terms, this means that ${\Eqs}\cdot G$ in a code contract
-must be such that:
+languages targeted by SPLAT do not directly support streams.  We are
+developing a translation to eliminate such deep temporal accesses by
+the introduction of intermediate variables. The essential property the
+translation needs to satisfy is a syntactic restriction ---call
+it \emph{Pascalish}---that supports easy translation to standard
+imperative languages.  The guiding intuition is that a Pascalish
+{\Eqs} allows access at most one step in the past, and accesses to a
+past value of a variable are only possible until the variable is
+updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
+contract has the following properties:
 \begin{itemize}
@@
 For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
-defined, albeit in a non-wellformed manner, by
+defined, albeit in a non-Pascalish manner, by
 {\small
@@
 }
-\noindent or by (again non-wellformed)
+\noindent or by (again non-Pascalish)
 {\small
@@
 }
-\noindent or by (wellformed):
+\noindent or by (Pascalish):
 {\small
@@
 \end{lstlisting}
-}
-A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
+} A Pascalish ${\Eqs} \cdot G$ is straightforward to implement in
 conventional programming languages, as we will see in

### 2022-03-27 paper/abstract.tex

@@
 prevent malicious input and detect supply chain attacks, respectively.
-A formal specification defines each high-assurance component, and is
+A formal specification in the form of a code-contract defines each high-assurance component and is
 used to verify that the component addresses system level cyber
 requirements.  Implementations for these high-assurance components are
-directly synthesized from their specifications, and are automatically
-proven to preserve the exact meaning of the specifications all the way
+directly synthesized from their code-contracts, and are automatically
+proven to preserve the exact meaning of the code-contract all the way
 down to the binary code level.  The model transformations are
 integrated into the Open Source AADL Tool Environment (OSATE).  The
-paper further reports on a case study applying cyber-resiliency model
-transformations to a UAV system that uses the Air Force Research
-Laboratory's OpenUxAS services for route planning.  In the case study,
-the model transformations add filters, to guard against malformed
-input, as well as monitors, to guard against ground station spoofing
-and malicious flight plans from OpenUxAS.
+paper further reports on two case studies, one by formal method experts 
+and the other by non-experts,
+applying cyber-resiliency model
+transformations to two different systems. 
+One is a UAV system that uses the Air Force Research
+Laboratory's OpenUxAS services for route planning.
+The other is a tablet used for inflight communication between pilots in an Apache helicopter.
+In each case study, the model transformations add filters, to guard against malformed
+input, as well as monitors, to guard against spoofing and other malicious behavior.
+The case studies show the MBSE is feasible in real-world systems engineering.

### 2022-03-27 paper/high-assurance-contracts.tex
Light proofreading

@@
 The filter is direct: it defines the policy according to the well-formed requirement for data.
-We now discuss in some detail the monitor.
+We now discuss the monitor in some detail.
 
@@
 \lineref{line:mon-assume} and is the same assumption used for the
-system input.  The policy can be written without needing the
-assumption, but it is much simpler to write with the assumption.
+system input.  The policy can be written without the assumption, 
+but the policy is much simpler to write with the assumption.
 
@@
 deployed system.  Precise in this context means that the
-implementations match the input to output relations defined in the
+implementations match the input-to-output relations defined in the
 specifications.

### 2022-03-27 paper/testing.tex
Light proofreading

@@
-Before discussing the final steps in synthesis, we take a moment to discuss how code-contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
+Before discussing the final steps in synthesis, we take a moment to
+discuss how code contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
 Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
@@
 AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification.
-The tempting approach to check these issues is to use the synthesize path and then apply traditional testing.
-Such an approach adds extra steps that are not needed as it is possible to bridge the formal model in AGREE with its proof system to traditional reasoning with test.
+The tempting approach to check these issues is to use the synthesis path and then apply traditional testing.
+Such an approach adds extra steps that are not needed, as it is
+possible to bridge the formal model in AGREE, with its proof system, to
+traditional validation using tests.
 
@@
 Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
-AGREE proves the component implements the test, and the Liskov substitution principle holds.
+AGREE proves that the component implements the test, and the Liskov substitution principle holds.
 
@@
 \linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp}).
-\figref{fig:test-output} is the AGREE output for the test is showing that it passes.
+\figref{fig:test-output} is the AGREE output for the test, showing that it passes.
 

### 2022-03-27 paper/high-assurance-contracts.tex
minor futzing

--- 2022-03-27 paper/high-assurance-contracts.tex
@@
 imperative languages.  The guiding intuition is that a Pascalish
-{\Eqs} allows access at most one step in the past, and accesses to a
+{\Eqs} accesses at most one step in the past; also, accesses to a
 past value of a variable are only possible until the variable is

### 2022-03-27 paper/agree-semantics.tex
wordsmithing on example; discussion of event ports tidied

@@
 
-We now give an overview of a formal model for AGREE syntax and semantics.
-The discussion then becomes more concrete, showing by example, how the AGREE language is used to specify the properties of the SW system introduced in \secref{sec:example}.
-Any implementation of the SW system must guarantee the specification output under the system assumptions.
+We now give an overview of a formal model for AGREE syntax and
+semantics.  The discussion then becomes more concrete, showing by
+example, how the AGREE language is used to specify the properties of
+the SW system introduced in \secref{sec:example}.  Any implementation
+of the SW system must guarantee the specification output under the
+system assumptions.
 
-The syntax of AGREE is
-essentially that of quantifier-free first order predicate logic
-supplemented with a few temporal operators. The terms (\emph{e}) are
-arithmetic expressions built from variables ($v$) and numeric and
-boolean literals ($c$), while formulas (\emph{b}) are built using
-logical connectives from atomic formulas (\emph{a}) based on the
-familiar comparison operators.
+The syntax of AGREE is essentially that of quantifier-free first order
+predicate logic supplemented with a few temporal operators. The terms
+(\emph{e}) are arithmetic expressions built from variables ($v$) and
+numeric and boolean literals ($c$), while formulas (\emph{b}) are
+built using logical connectives from atomic formulas (\emph{a}) based
+on the familiar comparison operators.
 \[
@@
 There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. Lastly, there are temporal operators \emph{previously} $\konst{pre}(-)$,
-\emph{followed-by} $(-) \to (-)$, and historically $\konst{Hist}(-)$.
-These temporal operators are sufficient to define all other past-time operators \cite{monitor}.
+and formulas. To support port handling, there is a predicate
+$\konst{event}(p)$, which tests if an event (or event-data) port $p$ is
+signaled. Lastly, there are temporal operators \emph{previously}
+$\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically
+$\konst{Hist}(-)$.
 
-The semantics of terms and formulas, as mentioned in the previous section, is in terms of \emph{streams of
-values}. Values encompass at least booleans and numbers, but can be
-readily extended to include records and arrays. A value stream is a
-total function from time (natural numbers including 0) to values:
+The semantics of terms and formulas, as mentioned in the previous
+section, is in terms of \emph{streams of values}. Values encompass at
+least booleans and numbers, but can be readily extended to include
+records and arrays. A value stream is a total function from time
+(natural numbers including 0) to values:
 \[
@@
 \]
-
 Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$

### 2022-03-27 paper/example.tex
wordsmithing on example; discussion of event ports tidied

@@
 
-\figref{fig:example} is an AADL architectural model of a software
-system (SW) for route planning and automated control of a UAV.  It is
-loosely based on one of the case studies in
-Section~\ref{sec:case-study}.  The source for the entire model is
-found at \cite{repo}.
+We now describe a simple software system, named SW, for route planning
+and automated control of a UAV. A picture of the AADL architectural
+model for SW is in \figref{fig:example}. SW is loosely based on one of
+the case studies in Section~\ref{sec:case-study}.  The source for the
+entire model is found at \cite{repo}.
 
-The system receives an automation request that is forwarded to a
-third-party route planner (AI).  The route planner decides the flight
-path of the UAV based on its current position and the requested task.
-The waypoint manager (WM) receives the mission command as a set of
-waypoints from the planner and starts the UAV flying the mission with an event on the \texttt{start} output port,
-issuing waypoints to the UAV flight controller as the UAV location
-changes.  The waypoint manager is a legacy component that cannot be
-modified so it critically relies on assumptions about its input behavior to guarantee its intended output behavior.
+From time to time, SW receives an \texttt{AutomationRequest} message,
+forwarding it to a route planner (AI). The function of AI is to
+compute the flight path (a list of waypoints) for the UAV, outputting
+the resulting \emph{mission command} on \texttt{AutomationResponse}.
+The waypoint manager (WM) receives the mission command from AI and
+starts the UAV flying the mission by putting an \emph{event} on the
+\texttt{Start} output port, continuing to issue waypoints to the UAV
+flight controller on the \texttt{Waypoint} port as the UAV location
+changes.\footnote{\konst{KLS}:What about Alert? When raised, is it
+persistent, where is it computed?}  The AI component is third-party
+software and WM is a legacy component.
+
+%% that cannot be modified so it critically relies on
+%% assumptions about its input behavior to guarantee its intended output
+%% behavior.
 
 The expected behavior of the SW system, and the components
-implementing the system, are modeled with AGREE contract
+implementing the system, is modeled with AGREE contract
 specifications.  The contracts constrain input with assumptions and
-state properties of output with guarantees if the assumptions are satisfied.  AGREE performs model
+state properties of output with guarantees.  AGREE performs model
 checking on this assume-guarantee system to hierarchically prove that
-the SW component obeys all contract obligations under all possible finite input
-streams.  The AGREE analysis, contract language, and contracts for SW
-are discussed in greater detail in \secref{sec:agree}.
+the SW component obeys all contract obligations under all possible
+finite input streams.  The AGREE analysis, contract language, and
+contracts for SW are discussed in greater detail in
+\secref{sec:agree}.
 
-The initial AGREE contracts for the system in \figref{fig:example},
+The AGREE contracts for the initial system in \figref{fig:example},
 and its implementing components, both assume and guarantee the absence
 of any malicious, or unspecified, component behavior.  More precisely,
-the system contract assumes that all inputs are \emph{well-formed} and
+the contract for SW assumes that its inputs are well-formed and that
 there is never more than one automation request pending at a time.
-The meaning of well-formed is given by user-written predicates for
-each data type.  For example, a waypoint is well-formed if it falls
-within bounds for latitude, longitude, and altitude.  The guarantees
+(\emph{Well-formed} generally refers to a syntactic restriction on
+data at a port. For example, a waypoint is well-formed if it falls
+within bounds for latitude, longitude, and altitude.)  The guarantees
 for SW ensure that
 \begin{compactitem}
-\item a start includes a waypoint;
+\item a start includes a waypoint;\footnote{\konst{KLS}: needs clarification; Start is just an event port.}
 \item a start is within one cycle of an automation request and if not, then it persistently alerts;
@@
 contract guarantees it only responds to automation requests and always
-in the same cycle.  The legacy WM contract guarantees the following if its input assumptions hold: 
+in the same cycle.  The legacy WM contract guarantees the following if
+its input assumptions hold:
 \begin{compactitem}
@@
 \end{compactitem}
-These initial specifications pass AGREE
-verification meaning that the contract composition of the components
-with the system satisfies all component input assumptions and system
-output guarantees.
+These initial specifications pass AGREE verification, meaning that the
+contract composition of the components with the system satisfies all
+component input assumptions and system output guarantees.
 
@@
 the AGREE specifications to equivalent programs in the CakeML
-language.  The synthesis toolchain generates a proof that equates the
-meaning of the AGREE specification to the behavior of the CakeML
-program.  In other words, for any set of input streams that meet the
-component's contract assumptions, the output streams produced from the
-synthesized CakeML code exactly match those defined by the
-high-assurance component's guarantees.  CakeML provides verified
-compilation to binaries for several different platforms, meaning that
-the resulting binaries exactly preserve the meaning of the original
-CakeML code \cite{cakeml}.
+language.
+%% Too strong at the moment!
+%The synthesis toolchain generates a proof that equates the
+%% meaning of the AGREE specification to the behavior of the CakeML
+%% program.
+ In other words, for any set of input streams that meet the
+ component's contract assumptions, the output streams produced from
+ the synthesized CakeML code exactly match those defined by the
+ high-assurance component's guarantees.  The CakeML compiler rovides
+ verified compilation to binaries for several different platforms,
+ meaning that the resulting binaries exactly preserve the meaning of
+ the original CakeML code \cite{cakeml}.
 

### 2022-03-27 paper/verification-conditions.tex
Remove sentence on cut point in cycle
wordsmithing on example; discussion of event ports tidied

--- 2022-03-26 paper/verification-conditions.tex
@@
 
-The AGREE specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
-Lustre reasons over a synchronous data-flow model where the inputs and outputs of
-components are streams.
-Informally, streams assign values to expressions at each point in time.
-AGREE contracts express relationships between input and output streams by assuming a set of allowed input streams and then guaranteeing a corresponding set of allowed output streams.
-Output for streams not included in the assumed input is undefined.
+The AGREE specification language, and semantics, are inspired by the
+Lustre language \cite{10.1145/41625.41641}. Lustre semantics are
+synchronous data-flow where the inputs and outputs of components are
+streams.  Informally, streams assign values to expressions at each
+point in time.  AGREE contracts express relationships between input
+and output streams by assuming a set of allowed input streams and then
+guaranteeing a corresponding set of allowed output streams.
 
-When considering a system of components,
-data flows through the components in dependency order, with inputs
-being propagated to outputs through all contracts until they stabilize
-(i.e., they cannot propagate further). Therefore, the subcomponent contracts, and
-thus the top-level model, must be acyclic. 
-Cyclic systems are made acyclic by breaking cycles with delays.
-That is to say that input at the cut point reasons about previous input data rather than the current input data.
-Once the data propagation
-has stabilized, the model proceeds to the next input data in the input
-streams. It is worth noting that the semantics do not model computation or communication
-delay. The output of one contract is seen at the input of any
-downstream contract in the same step of the input data stream.
+\subsection{Components}
+A system implementation is described by its subcomponents plus
+connections between their ports.  When considering a system of
+components, data flows through the components in dependency order,
+with inputs being propagated to outputs through all contracts until
+they stabilize (i.e., they cannot propagate further). Therefore, the
+subcomponent contracts, and thus the top-level model, must be acyclic.
+Cyclic systems are made acyclic by breaking cycles with delays.  Once
+the data propagation has stabilized, the model proceeds to the next
+input data in the input streams. It is worth noting that the semantics
+do not model computation or communication delay. The output of one
+contract is seen at the input of any downstream contract in the same
+step of the input data stream.
+
+An component in AGREE is described relative to its \emph{ports}, a
+concept inherited from AADL.  There are three types of AADL ports:
+\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
+An event port has no data and only signals occurrence (or not) of the
+event. A data port is a sampled port: the data can be read at anytime.
+An event data port indicates when data is present with an event.
+
+\subsection{Contracts and verification conditions}
 
@@
 
-PLTL is a logic enhanced with temporal operators able to
-reason about the truth values of formulas through time.
-It combines past-time operators that reason over the history of computation up to a point (PTLTL) and future-time operators that reason over the computation that follows that point (LTL).
-Its semantics
-are defined relative to a point in time $i$, a finite trace of
-system states $\pi = s_0, s_1, \ldots, s_i$, and some logic formula $f$.
+PLTL is a logic enhanced with temporal operators able to reason about
+the truth values of formulas through time.  It combines past-time
+operators that reason over the history of computation up to a point
+(PTLTL) and future-time operators that reason over the computation
+that follows that point (LTL).  Its semantics are defined relative to
+a point in time $i$, a finite trace of system states $\pi = s_0, s_1,
+\ldots, s_i$, and some logic formula $f$.
 
@@
 \end{equation}
-The conditions from \eqref{eq:assumes} verify the input assumptions on each component under
-the system assumptions and upstream component guarantees.  They check
-if the component guarantees and system assumptions are strong enough
-to imply input assumptions on all immediate downstream components.
-The condition in \eqref{eq:guarantees} checks the output guarantees of the system under the
-system assumptions and component guarantees that provide the output.
-It checks if the guarantees on components providing primary outputs
-are strong enough to imply the system guarantees.
+The conditions from \eqref{eq:assumes} verify the input assumptions on
+each component under the system assumptions and upstream component
+guarantees.  They check if the component guarantees and system
+assumptions are strong enough to imply input assumptions on all
+immediate downstream components.  The condition in
+\eqref{eq:guarantees} checks the output guarantees of the system under
+the system assumptions and component guarantees that provide the
+output.  It checks if the guarantees on components providing primary
+outputs are strong enough to imply the system guarantees.
 

### 2022-03-27 paper/abstract.tex
Wordsmithing in Abstract, also cleaning up spillover into the gutter between columns

--- 2022-03-22 paper/abstract.tex
@@
 engineers to design-in cyber-resiliency in a Model-Based Systems
-Engineering environment.  This paper describes automated model
+Engineering (MBSE) environment.  This paper describes automated model
 transformations that introduce high-assurance cyber-resiliency
-components into a system, in particular filters and monitors that
-prevent malicious input and detect supply chain attacks, respectively.
-A formal specification defines each high-assurance component, and is
-used to verify that the component addresses system level cyber
-requirements.  Implementations for these high-assurance components are
-directly synthesized from their specifications, and are automatically
-proven to preserve the exact meaning of the specifications all the way
-down to the binary code level.  The model transformations are
-integrated into the Open Source AADL Tool Environment (OSATE).  The
-paper further reports on a case study applying cyber-resiliency model
-transformations to a UAV system that uses the Air Force Research
-Laboratory's OpenUxAS services for route planning.  In the case study,
-the model transformations add filters, to guard against malformed
-input, as well as monitors, to guard against ground station spoofing
-and malicious flight plans from OpenUxAS.
+components into a system; in particular, filters that guard against
+malformed input, as well as monitors that guard against
+spoofing and other malicious behavior.  A formal specification in the
+form of a code contract defines each high-assurance component and is
+used to verify that the component addresses system-level cyber
+requirements.  Implementations for these high-assurance components
+are directly synthesized from their code contracts, and are
+automatically proven to preserve the exact meaning of the code
+contract all the way down to the binary code level.  The model
+transformations are integrated into the Open Source AADL 
+Tool Environment (OSATE).  We further report on two case
+studies, one by formal method experts and the other by non-experts,
+applying model transformations to two different systems to improve
+cyber resiliency using high-assurance filters and monitors.  One is a
+UAV system utilizing the Air Force Research Laboratory's
+OpenUxAS services for route planning.  The other is a tablet used for
+situational awareness in a CH-47F Chinook helicopter application.
+The case studies demonstrate that our MBSE approach is feasible in
+real-world systems engineering.

### 2022-03-28 paper/agree-semantics.tex
Condensing Sec 4.2; added table mapping math defs to AGREE. Some entries look redundant

@@
 
+\subsection{Semantics}
 The semantics of terms and formulas, as mentioned in the previous
@@
 value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
-$t-1$ (at time zero, \konst{pre} is undefined).
-A followed-by, $e_1 \to e_2$, evaluates to $e_1$ at time zero; otherwise it is $e_2$.
-Followed-by is often used to give a value to a stream in the first step so that subsequent steps are able to use $\konst{pre}(e)$.
+$t-1$ (at time zero, \konst{pre} is undefined).  A followed-by, $e_1
+\to e_2$, evaluates to $e_1$ at time zero; otherwise it is $e_2$.
+Followed-by is typically used to give a value to a stream in the first
+step so that subsequent steps are able to use $\konst{pre}(e)$.
 \[
@@
 All other past-time operators of PLTL can be defined similarly \cite{monitor}.
+
+\paragraph{AGREE notation.}
+AGREE uses ASCII notation for the mathematical syntax we have
+defined. The following table spells out the mapping.
+\[
+\begin{tabular}{l|l|l}
+ Operation & Math. notation & AGREE \\ \hline
+ followed-by &  $e \to e$          & e \verb+->+ e  \\
+ logical implication &  $e \imp e$ & e \verb+=>+ e \\
+ if-and-only-if &  $e \iff e$      & e \verb+<=>+ e \\
+ variable definition & $v = e$     & \konst{eq} v\ e \\
+ Historically  & $\konst{Hist}(b)$ & \konst{Historically}(b) \\
+\end{tabular}
+\]

### 2022-03-28 paper/sw-contract.tex
Condensing Sec 4.2; added table mapping math defs to AGREE. Some entries look redundant

@@
 
-The actual surface syntax for AGREE that is embedded in OSATE is slightly different than the formal semantics in the previous section.
-The mapping between the two is straightforward.
-The AGREE specification for the SW component in the example of
-Section~\ref{sec:example} is given in \figref{fig:sw}.  The
-specification uses \texttt{eq} statements to define variables local to
-the contract specification.  For
-example, \lineref{line:sw-event-def-start} defines the \texttt{req}
-variable to be equivalent to the \texttt{event} expression.
+%% The actual surface syntax for AGREE that is embedded in OSATE is
+%% slightly different than the formal semantics in the previous section.
+%% The mapping between the two is straightforward.
 
-All the named ports in the corresponding AADL component are in the scope of the specification, and there are additional implicit boolean \emph{event} inputs (or outputs) associated with event ports.
-An \texttt{event} expression refers to that implicit input (or output) boolean value and is true when data is present on the named port and false otherwise.
-\linesref{line:sw-event-def-start}{line:sw-event-def-end} create local variables that are true when data is present on the corresponding event ports for the component.
-The local variables here are purely for convenience in writing the specification.
+%% The
+%% specification uses \texttt{eq} statements to define variables local to
+%% the contract specification.  For
+%% example, \lineref{line:sw-event-def-start} defines the \texttt{req}
+%% variable to be equivalent to the \texttt{event} expression.
 
-The \texttt{assume} statement is a string description followed by a predicate.
-Those on \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2} are implications requiring that when data is present it is well-formed.
-The well-formed predicates themselves are defined elsewhere using AGREE functions.
+%% All the named
+%% ports in the corresponding AADL component are in the scope of the
+%% specification, and there are additional implicit boolean \emph{event}
+%% inputs (or outputs) associated with event ports.  An \texttt{event}
+%% expression refers to that implicit input (or output) boolean value and
+%% is true when data is present on the named port and false otherwise.
 
-The assumption on \lineref{line:sw-assume-3} constrains when a request
-can arrive by reasoning about the \emph{state} of the contract defined
-by the associated streams.  When writing assumptions and guarantees,
-it is important to differentiate pre-state, before a contract updates
-its state, and post-state, after a contract updates its state, in
-response to the current input.  The \texttt{pre} naturally makes that
-distinction.  In general, assumptions regarding state should be
-evaluated in the pre-state of the component with the current inputs,
-and guarantees regarding state should be evaluated on the post-state
-of the component given the current inputs.  Additionally, guarantees
-should use \texttt{pre} anytime they need to reason about the current
-state relative to the previous state.
+%% An \texttt{assume} statement is a string description followed by a
+%% predicate.
 
-The assumption on \lineref{line:sw-assume-3} uses a followed-by expression so it is \texttt{true} at time zero and then it is the truth value of the implication in all future instances.
-The followed-by guards the \texttt{pre} as mentioned previously;
-thus after time zero, the assumption depends on the presence of a request and the pre-state of the component.
-The assumption on \lineref{line:sw-assume-3} is that if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it is after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---\emph{not request since start}.
+%% When writing assumptions and guarantees,
+%% it is important to differentiate pre-state, before a contract updates
+%% its state, and post-state, after a contract updates its state, in
+%% response to the current input.  The \texttt{pre} naturally makes that
+%% distinction.  In general, assumptions regarding state should be
+%% evaluated in the pre-state of the component with the current inputs,
+%% and guarantees regarding state should be evaluated on the post-state
+%% of the component given the current inputs.  Additionally, guarantees
+%% should use \texttt{pre} anytime they need to reason about the current
+%% state relative to the previous state.
 
-The guarantees on \linesref{line:sw-guarantee-1}{line:sw-guarantee-3} coincide events and assert well-formed output.
-The guarantees on \linesref{line:sw-guarantee-4}{line:sw-guarantee-6} define temporal properties of the component.
-\lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request.
-The guarantee uses the assumption on \lineref{line:sw-assume-3} and does not check for two requests in a row without a start as the assumption precludes that input behavior.
-It differentiates with the followed-by what is required at time zero, the start must coincide with the request, with what is required after time zero, the start must coincide with the request or is a response to a request one step earlier.
-The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.
+%% This condition constrains when a request
+%% can arrive by reasoning about the \emph{state} of the contract defined
+%% by the associated streams. The assumption uses a followed-by
+%% expression so it is \texttt{true} at time zero and then it is the
+%% truth value of the implication in all future instances.  The
+%% followed-by guards the \texttt{pre} as mentioned previously; thus
+%% after time zero, the assumption depends on the presence of a request
+%% and the pre-state of the component.  The assumption
+%% on \lineref{line:sw-assume-3} is that
 
-\lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound.
-Together with \lineref{line:sw-guarantee-4} the contract model allows for non-alerting and alerting behavior.
-\lineref{line:sw-guarantee-6} ensures if alert has ever happened, then it is happening in the present moment.
-These six guarantees define the behavior of the SW component under the three assumptions and correspond to the informal descriptions given in \secref{sec:example}.
+The AGREE specification for the SW component in
+Section~\ref{sec:example} is given in \figref{fig:sw}.
+\linesref{line:sw-event-def-start}{line:sw-event-def-end} define
+local variables that are true when data is present on the
+corresponding event ports for the component.  These variables are
+purely for notational convenience. The contract has three assumptions:
+\begin{itemize}
+\item \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
+When input is present it is well-formed. The well-formed predicates
+are defined elsewhere using AGREE functions.
+
+\item \lineref{line:sw-assume-3}. If there is an incoming request,
+it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)},
+or it arrives after the component has output a start event in response to a previous
+request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
+request since start}.
+\end{itemize}
+
+The contract has six guarantees:
+\begin{itemize}
+\item \linesref{line:sw-guarantee-1}{line:sw-guarantee-3}
+coincide events and assert well-formed output.\footnote{\konst{KLS}:I am confused about the ``coincide events''.}
+\item \lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request. At time zero the start must coincide
+with the request; after that, the start must coincide with the request
+or else is a response to a request one step earlier.  The guarantee
+works with the assumption on \lineref{line:sw-assume-3} and thus does
+not have to check for two requests in a row without a start, since the
+assumption precludes that input behavior. The guarantee also does not
+force the start to always happen, it only says that if it does happen,
+it is under the defined conditions.
+
+\item \lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound. Together with \lineref{line:sw-guarantee-4} this allows
+for non-alerting and alerting behavior.
+\item \lineref{line:sw-guarantee-6} guarantees that, if an alert has ever happened,
+it continues to happen.
+\end{itemize}
+These six guarantees define the behavior of the SW component under the
+three assumptions and correspond to the informal descriptions given
+in \secref{sec:example}.

### 2022-03-28 paper/high-assurance-contracts.tex
Minor upgrades

@@
 \end{lstlisting}
-} A Pascalish ${\Eqs} \cdot G$ is straightforward to implement in
-conventional programming languages, as we will see in
+}
+(Note that the initial values of \verb+F1+ and \verb+F2+ are never
+accessed, so 42 could be replaced by any number in their defining
+equations.) A Pascalish ${\Eqs} \cdot G$ is straightforward to
+implement in conventional programming languages, as we will see in
 Section \ref{sec:synthesis}.

### 2022-03-28 paper/verification-conditions.tex
Minor upgrades

@@
 \subsection{Components}
-A system implementation is described by its subcomponents plus
-connections between their ports.  When considering a system of
-components, data flows through the components in dependency order,
+A system implementation is described by its ports, its subcomponents,
+and the connections between ports on different subcomponents.  Ports
+in AGREE are inherited from AADL.  There are three kinds of port:
+\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
+An event port carries no data and only signals occurrence (or not) of the
+event. A data port is a sampled port: the data can be read at anytime.
+An event data port indicates when data is present with an event.
+
+Data flows through the subcomponents of a system in dependency order,
 with inputs being propagated to outputs through all contracts until
-they stabilize (i.e., they cannot propagate further). Therefore, the
-subcomponent contracts, and thus the top-level model, must be acyclic.
-Cyclic systems are made acyclic by breaking cycles with delays.  Once
-the data propagation has stabilized, the model proceeds to the next
-input data in the input streams. It is worth noting that the semantics
-do not model computation or communication delay. The output of one
+they stabilize \ie, cannot propagate further. Therefore, subcomponent
+contracts, and thus the top-level model, must be acyclic.  Cyclic
+systems are made acyclic by breaking cycles with delays.  Once the
+data propagation has stabilized, the model proceeds to the next input
+data in the input streams. It is worth noting that the semantics do
+not model computation or communication delay: the output of one
 contract is seen at the input of any downstream contract in the same
 step of the input data stream.
-
-An component in AGREE is described relative to its \emph{ports}, a
-concept inherited from AADL.  There are three types of AADL ports:
-\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
-An event port has no data and only signals occurrence (or not) of the
-event. A data port is a sampled port: the data can be read at anytime.
-An event data port indicates when data is present with an event.
 
@@
 this result is that $\globally(\historically(\assumes) \implies
-\guarantees)$ holds for the system contract.
-Therefore the Liskov substitution principle applies, and as such, the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
+\guarantees)$ holds for the system contract.  Therefore the Liskov
+substitution principle applies and the implementation is a safe
+substitution for the system contract \cite{10.1145/62139.62141}.
 
@@
 unexpanded results at the bottom of the figures prove
-\emph{self-consistency} in the contracts.  It is not uncommon to
+\emph{self-consistency} in the contracts.\emph{Needs a definition?}  It is not uncommon to
 accidentally write contracts that are self-contradicting.  For

### 2022-03-28 paper/high-assurance-contracts.tex
checkpointing before tackling Section 7

@@
 
-\subsubsection*{Imperative code contracts}
+\subsubsection*{Imperative code contracts.}
 Stream-based evaluation allows values arbitrarily `deep in the past'
@@
 interface have created a cyber-hardened system, protecting against
-malicious AI behavior, by inserting a high-assurance filter and
+malicious AI behavior by the insertion of a high-assurance filter and
 monitor between the AI and the WM (see \figref{fig:hardened}). The
@@
 \figref{fig:filter} and \figref{fig:monitor}.
-The filter is direct: it defines the policy according to the well-formed requirement for data.
-We now discuss the monitor in some detail.
+The filter is direct: it defines the policy according to the
+well-formed requirement for data.  We now discuss the monitor in some
+detail.
 
@@
 \figref{fig:monitor}.
-The
-\texttt{alert} variable defined on \lineref{line:mon-alert} depends
-not just on the policy but also on configuration details supplied when
-the component is specified.  At configuration
-time, \texttt{is\_latched} has been set to \konst{true}, thus making
-the alert persistent, meaning that once the alert is raised, it is
-always raised; otherwise, it is the complement of the policy value in
-the current step.
+
+The \texttt{alert} variable defined on \lineref{line:mon-alert}
+(distinct from the \texttt{Alert} port) depends not just on the policy
+but also on configuration details supplied when the component is
+specified.  At configuration time, \texttt{is\_latched} has been set
+to \konst{true}, thus making the alert persistent, meaning that once
+the alert is raised, it is always raised; otherwise, it is the
+complement of the policy value in the current step.
 
@@
 data, spontaneous responses, or delayed responses.  Synthesis must
-generate precise implementations of the high-assurance component
+generate faithful implementations of the high-assurance component
 specifications for the AGREE verification result to apply to the
-deployed system.  Precise in this context means that the
-implementations match the input-to-output relations defined in the
-specifications.
+deployed system.
+
+%%   Precise in this context means that the
+%% implementations match the input-to-output relations defined in the
+%% specifications.

### 2022-03-28 paper/synthesis.tex
latest version of Section 7. Details remain to be checked. Listings may not be up-to-date I fear

@@
 assurance components specified by code contracts; for each such
-occurrence it generates a CakeML program. Starting from the semantics
-of the code contract described in Section \ref{sec:code-contracts}
-a sequence of semantics-preserving steps are taken:
-\begin{itemize}
-\item $\sem{{\Eqs}\cdot G}$ yields a function on arbitrary depth streams
-\item transform to function over depth-1 streams (wellformedness)
-\item transform to logic function
-\item attach buffer handling to inputs and outputs (contig type parsing)
-\item translate to CakeML
-\item run CakeML compiler
-\end{itemize}
+occurrence it generates a CakeML program. More generally, SPLAT supports
+three kinds of behavioral specification from which to generate
+code. In increasing order of expressiveness, these are:
+
+\paragraph{Regular expressions.} Often the well-formedness of data on connections
+can be enforced by matching against a regular expression. In other
+words, the data can be characterized by a \emph{regular language}. As
+is well known, regular expressions can be translated to efficient
+deterministic finite state machine (DFA)
+implementations. In \cite{formal-filter-synth-langsec,case-verified-filter}
+we report on the creation and application of filters from a verified
+regular expression-to-DFA compiler.
+
+\paragraph{Contiguity types\cite{contiguity-types}.}
+A contiguity type, like regular expressions, is effective at
+describing data at the network message level, \eg, as flat
+strings. The notation is however more expressive, being able to
+declare many kinds of so-called \emph{self-describing} messages, \ie,
+those where message structure is dependent on data held elsewhere in
+the message. Figure \ref{fig:filter-spec} gives a contiguity type
+specification for the welll-formed waypoints of our example system. We
+have formalized and verified a contiguity type matcher. The matcher
+code has been used in our case studies.
+
+\paragraph{AGREE code contracts.} When full computational power is needed, we can use the
+expressive power of AGREE code contracts to write arbitrarily complex
+filter or monitor operations, as we have already seen. Our formal
+model of AGREE syntax and semantics is applied to extract code
+contracts and map them to equivalent logic functions.
+
+\noindent For each of these three approaches, we take the following steps to
+arrive at an executable.
+
+\begin{enumerate}
+\item Map from specification to computable function in logic, producing
+a proof of equivalence.
+\item Add parsers and printers (specified by contiguity types) for input and output.
+\item Map to CakeML and add calls to input/output library routines via
+      the CakeML foreign function interface.
+\item Invoke the CakeML compiler.
+\end{enumerate}
+
+%% Starting from the semantics of the code contract described in
+%% Section \ref{sec:code-contracts} a sequence of semantics-preserving
+%% steps are taken:
+%% \begin{itemize}
+%% \item $\sem{{\Eqs}\cdot G}$ yields a function on arbitrary depth streams
+%% \item transform to Pascalish
+%% \item transform to logic function
+%% \item attach buffer handling to inputs and outputs (contig type parsing)
+%% \item translate to CakeML
+%% \item run CakeML compiler
+%% \end{itemize}
 
@@
 suffices to enforce the policy. For example, we have used the
-expressive power of Contiguity Types \cite{contiguity-types} to
-enforce \emph{lightweight} bounds constraints on GPS coordinates in
-UxAS messages. On the other hand, a filter may need to parse the input
-buffer into a data structure specified in AGREE and apply a
-user-defined \emph{wellformedness} property, also specified in AGREE,
-to the data. Arbitrarily complex wellformedness checks can be made in
-this way. \figref{fig:filter-spec} shows a combination where the
-checking specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+}
-depends on an underlying check specified by the contiguity type
-checking bounds on waypoints.
+expressive power of regular expressions and Contiguity
+Types \cite{contiguity-types} to enforce \emph{lightweight} bounds
+constraints on GPS coordinates in UxAS messages. On the other hand, a
+filter may need to parse the input buffer into a data structure
+specified in AGREE and apply a user-defined \emph{wellformedness}
+property, also specified in AGREE, to the data. Arbitrarily complex
+wellformedness checks can be made in this
+way. \figref{fig:filter-spec} shows a combination where the checking
+specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+} depends on
+an underlying check specified by the contiguity type checking bounds
+on waypoints.
 
@@
 the wellformedness computation to finish. Thus resource bounds need to
-be included in the correctness argument.
+be included in the correctness argument. Some preliminary work on this
+has been done in CakeML \cite{cakeml-space-cost}.
 
@@
 Monitors are intended to track and analyze the externally visible
-behavior of system components through time. Therefore, they require
-more extensive computational ability than filters. In particular, our
-basic notion of a monitor is that it embodies a predicate over its
-input and output streams, and is able to access the value of a stream at any
-earlier point in time, if necessary. Monitors commonly use state to
-keep track of earlier values, unlike filters which, for us, are
-typically stateless components. (However, there is nothing in our
-approach that forbids stateful filters: they can be realized by
-monitors.) A monitor specification is mapped by code generation to a
-state transformation function of the following abstract type:
+behavior of system components through time. Therefore, they tend to
+require more extensive computational ability than filters. In
+particular, our basic notion of a monitor is that it embodies a
+relation over its input and output streams, and is able to access the
+value of a stream at any earlier point in time, if necessary. Monitors
+commonly use state to keep track of earlier values, unlike filters
+which, for us, are typically stateless components. (However, there is
+nothing in our approach that forbids stateful filters: they can be
+realized by monitors.) A monitor specification is mapped by code
+generation to a state transformation function of the following
+abstract type:
 \[

### 2022-03-28 paper/example.tex
Add minor edits to example

@@
 
-From time to time, SW receives an \texttt{AutomationRequest} message,
+From time to time, SW receives a message on \texttt{AutomationRequest},
 forwarding it to a route planner (AI). The function of AI is to 
@@
 flight controller on the \texttt{Waypoint} port as the UAV location
-changes.\footnote{\konst{KLS}:What about Alert? When raised, is it
-persistent, where is it computed?}  The AI component is third-party
+updates with messages on \texttt{AirVehicleLocation}.
+An event is repeatedly generated on \texttt{Alert} if there is no
+mission command from a request. The AI component is third-party
 software and WM is a legacy component.
@@
 of any malicious, or unspecified, component behavior.  More precisely,
-the contract for SW assumes that its inputs are well-formed and that
+the contract for SW assumes that its inputs are \emph{well-formed} and that
 there is never more than one automation request pending at a time.
-(\emph{Well-formed} generally refers to a syntactic restriction on
+Well-formed generally refers to a syntactic restriction on
 data at a port. For example, a waypoint is well-formed if it falls
@@
 \begin{compactitem}
-\item a start includes a waypoint;\footnote{\konst{KLS}: needs clarification; Start is just an event port.}
+\item a start coincides with a new waypoint message being output;
 \item a start is within one cycle of an automation request and if not, then it persistently alerts;
@@
   \item it generates a start from a response from the AI and always in the same cycle;
-  \item the start always comes with a waypoint; and
+  \item the start always coincides with a waypoint being output; and
   \item any further waypoints coincide with air vehicle location updates.
@@
 assume-guarantee reasoning of AGREE and is able to generate any value
-on its output stream at anytime.
+on its output at anytime.
 
@@
 
-For example, the first violation is that the automation response from
+For example, the first violation is that the mission command on \texttt{AutomationResponse} from
 the AI to the WM is no longer guaranteed to be well-formed.  The
@@
 that input violates the filter invariant.  The auto-generated AGREE
-specification states that only well-formed inputs are passed to the
+code-contract specification states that only well-formed inputs are passed to the
 output.  The system developer must further define the filtering
@@
 an alert if the specified properties are ever violated.  The
-AGREE specification for the monitor in our example states that an
+AGREE code-contract specification for the monitor in our example states that an
 automation response can only be generated in conjunction with an
@@
 based on existing the assumptions and guarantees of the SW system and its implementing components.
+AGREE code-contract specifications are discussed in detail in \secref{sec:code-contracts}.
 
@@
 High-assurance components are automatically synthesized by SPLAT from
-the AGREE specifications to equivalent programs in the CakeML
+the AGREE code-contract specifications to equivalent programs in the CakeML
 language.
@@
  the synthesized CakeML code exactly match those defined by the
- high-assurance component's guarantees.  The CakeML compiler rovides
+ high-assurance component's guarantees.  The CakeML compiler provides
  verified compilation to binaries for several different platforms,

### 2022-03-28 paper/example.tex
Fix over-full column

@@
 
-From time to time, SW receives a message on \texttt{AutomationRequest},
+From time to time, SW receives a message on the \texttt{AutomationRequest} input,
 forwarding it to a route planner (AI). The function of AI is to 
@@
 \end{compactitem}
-These initial specifications pass AGREE verification, meaning that the
+These initial specifications pass verification, meaning that the
 contract composition of the components with the system satisfies all

### 2022-03-28 paper/sw-contract.tex
Minor upgrades
Answer Konrad's questions is section 4 and 5

@@
 
-The contract has six guarantees:
+\noindent The contract has six guarantees:
 \begin{itemize}
-\item \linesref{line:sw-guarantee-1}{line:sw-guarantee-3}
-coincide events and assert well-formed output.\footnote{\konst{KLS}:I am confused about the ``coincide events''.}
-\item \lineref{line:sw-guarantee-4} insists that a start happens with a request or one step after a request. At time zero the start must coincide
+\item \lineref{line:sw-guarantee-1}, \lineref{line:sw-guarantee-2}, and \lineref{line:sw-guarantee-3}.
+A waypoint coincides with a update on the UAV location, a start coincides with a new waypoint, and waypoints are well-formed.
+\item \lineref{line:sw-guarantee-4}. A start happens with a request or one step after a request. At time zero the start must coincide
 with the request; after that, the start must coincide with the request
@@
 
-\item \lineref{line:sw-guarantee-5} forces the alert to sound if the start does not arrive within the one-step bound. Together with \lineref{line:sw-guarantee-4} this allows
+\item \lineref{line:sw-guarantee-5}. The alert to sounds if the start does not arrive within the one-step bound. Together with \lineref{line:sw-guarantee-4} this allows
 for non-alerting and alerting behavior.
-\item \lineref{line:sw-guarantee-6} guarantees that, if an alert has ever happened,
+\item \lineref{line:sw-guarantee-6}. If an alert has ever happened,
 it continues to happen.

### 2022-03-28 paper/verification-conditions.tex
Answer Konrad's questions is section 4 and 5

@@
 unexpanded results at the bottom of the figures prove
-\emph{self-consistency} in the contracts.\emph{Needs a definition?}  It is not uncommon to
-accidentally write contracts that are self-contradicting.  For
-example, a contract may guarantee an output be two different values in
-the same moment of time.  AGREE generates additional verification
+\emph{self-consistency} in the contracts.
+A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
+AGREE generates additional verification
 conditions that prove each component contract, and the composition of

### 2022-03-28 paper/example.tex
Answer Konrad's questions is section 4 and 5
Remove hyphen on code contract

@@
 contracts for SW are discussed in greater detail in
-\secref{sec:agree}.
+\secref{sec:agree} and \secref{sec:agree-semantics}.
 
@@
 that input violates the filter invariant.  The auto-generated AGREE
-code-contract specification states that only well-formed inputs are passed to the
+code contract specification states that only well-formed inputs are passed to the
 output.  The system developer must further define the filtering
@@
 an alert if the specified properties are ever violated.  The
-AGREE code-contract specification for the monitor in our example states that an
+AGREE code contract for the monitor in our example states that an
 automation response can only be generated in conjunction with an
@@
 based on existing the assumptions and guarantees of the SW system and its implementing components.
-AGREE code-contract specifications are discussed in detail in \secref{sec:code-contracts}.
+AGREE code contracts are discussed in detail in \secref{sec:code-contracts}.
 
@@
 High-assurance components are automatically synthesized by SPLAT from
-the AGREE code-contract specifications to equivalent programs in the CakeML
+the AGREE code contracts to equivalent programs in the CakeML
 language.

### 2022-03-28 paper/high-assurance-contracts.tex
Add minor edits no section 5

@@
 component, and its connections in the system implementation, as part
-of the transformation.  It also creates a default AGREE specification
-that defines everything about the component behavior, leaving only the
-component behavior (policy) to be specified.
+of the transformation.  It also creates a default code contract with
+appropriate output guarantees leaving only the component behavior 
+(policy) to be specified.
 
@@
 \subsubsection{Monitor specification}
-BriefCASE generates everything but the
-\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
+The monitor policy is given by the \konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
 \figref{fig:monitor}.
-
-The \texttt{alert} variable defined on \lineref{line:mon-alert}
+\lineref{line:mon-alert}. The \texttt{alert} variable
 (distinct from the \texttt{Alert} port) depends not just on the policy
@@
 
+\lineref{line:mon-pending} and \lineref{line:mon-latency}.
 The monitor policy is defined by marking when a request is not
-satisfied, \texttt{isPending} on \lineref{line:mon-pending}, and
-counting the number of steps between requests, \texttt{latency} on
-\lineref{line:mon-latency}.  The \texttt{isPending} variable is true
+satisfied, \texttt{isPending}, and
+counting the number of steps between requests, \texttt{latency}.  
+The \texttt{isPending} variable is true
 whenever there is a request that does not coincide with a
-response--\emph{not response since request and not response}.  The
+response--to be read as \emph{there is no response since the most recent request without a corresponding response}.  The
 \texttt{latency} variable starts at zero, resets on every request, and
@@
 
-The latency bound for the monitor is defined by the constant
-\texttt{MAX\_LATENCY} on \lineref{line:mon-const}.  Here the constant
+\lineref{line:mon-const}. The latency bound for the monitor is defined by the constant
+\texttt{MAX\_LATENCY}.  Here the constant
 is set to one to be consistent with the system specification from the
@@
 
-The policy definition only works if there is never more than one
+\lineref{line:mon-assume}. The policy definition only works if there is never more than one
 outstanding request at a time; otherwise, the latency counter resets
-incorrectly.  That requirement is encoded in the assumption on
-\lineref{line:mon-assume} and is the same assumption used for the
+incorrectly.  That requirement is encoded in the assumption and is the same assumption used for the
 system input.  The policy can be written without the assumption,
@@
 data, spontaneous responses, or delayed responses.  Synthesis must
-generate faithful implementations of the high-assurance component
-specifications for the AGREE verification result to apply to the
+generate faithful implementations of the code contracts for the AGREE verification result to apply to the
 deployed system.

### 2022-03-28 paper/testing.tex
Revise test section

@@
 
-Four common mistakes in specifications have been observed: contradictions, vacuity, under-specification, and missed corner cases.
-Contradictions arise from inconsistent contracts that state conflicting constraints in any given moment invalidating the verification results.
+There are three common mistakes in specifications, aside from self-consistency that is already checked, that have been observed: vacuity, under-specification, and missed corner cases.
 Vacuity is a common pitfall with implication when the left hand side is always false making the implication true. 
@@
 
-AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification.
-The tempting approach to check these issues is to use the synthesis path and then apply traditional testing.
+The tempting approach is to use the synthesis path to check these issues with traditional testing.
 Such an approach adds extra steps that are not needed, as it is
@@
 traditional validation using tests.
+The key is to consider the component as implementing a test specification.
 
@@
 It also checks that the alarm does not fire.
-\lineref{line:tst-index} creates an index to use to create the test input.
-\lineref{line:tst-assume-1} defines when the response appears relative to the index.
-\lineref{line:tst-assume-2} defines the request behavior.
-\linesref{line:tst-guarantee-1}{line:tst-guarantee-2} define the expected test output.
-The test additionally proves that nothing happens after the initial test input and output (e.g., no more input is provided and no more output is generated).
-AGREE proves the test correct.
-\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp}).
+\begin{compactitem}
+\item \lineref{line:tst-index}. An index to use to define the test input and expected test output through time.
+\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}. The test input. These define when the request and response message events should appear relative to the index.
+\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}. The expected test output. 
+The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
+\item \linesref{line:tst-imp-start}{line:tst-imp-end}. The implementation with the component under test (\lineref{line:tst-imp-comp}).
+\end{compactitem}
 \figref{fig:test-output} is the AGREE output for the test, showing that it passes.

### 2022-03-28 paper/intro.tex
Add code contract to intro and fix the outline

@@
 architecture of the model by adding in new components; they also
-generate a formal specification for the behavior of the inserted
-high-assurance components in the AGREE language. Those specifications
+generate a formal specification in the form of a \emph{code contract} in the AGREE language for their intended behavior. 
+A code contract describes the computation of a high-assurance component and not just what it computes.
+These code contracts
 are sufficient for model checking to prove that---due to the newly
@@
 Another novel aspect of the approach is the synthesis of the AGREE
-specifications for the high-assurance components to CakeML, a verified
+code contracts for the high-assurance components to CakeML, a verified
 compiler implementation for the functional programming language
 ML \cite{cakeml}. This paper describes in detail the synthesis path
-from specifications to CakeML code, providing a formal framework in
+from code contracts to CakeML code, providing a formal framework in
 which to argue correctness. CakeML then provides a verified
@@
 The approach is illustrated by a simple example in
-Section~\ref{sec:example}. Contract specification in the AGREE
-language is presented in Section~\ref{sec:agree}, followed by a
-description of the synthesis pathway in Section~\ref{sec:synthesis}. A
+\secref{sec:example}. AGREE verification and specification in the AGREE
+language are presented in \secref{sec:agree} and \secref{agree-semantics}.
+Code contracts are defined in \secref{sec:code-contracts} with a method
+for testing code contracts in the AGREE framework before synthesis given
+\secref{sec:testing}.
+The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
 case study applying these transformations to an Unmanned Aerial

### 2022-03-28 cover-letter/letter.tex
Add cover-letter

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

My name is Dr. Eric Mercer, and I am one of the corresponding authors on a recently published paper in ACM/IEEE 24th International Conference on Model Driven Engineering Languages and Systems (MoDELS 2021).
On behalf of my coauthors, I would like to submit the enclosed manuscript entitled "Synthesizing Verified Components for Cyber Assured Systems Engineering" for consideration for publication in the MoDELS 2021 special issue of the Journal of Software and Systems Modeling (SOSYM). 
I confirm that the enclosed manuscript is an original work and has not been published nor has it been submitted simultaneously elsewhere. 
Also, all authors have checked the manuscript and have agreed to the submission.

The published manuscript in MODELS 2021 describes a Model Based Systems Engineering (MBSE) environment developed as part of the the DARPA Cyber Assured Systems Engineering program.
The environment enables systems engineers to design-in cyber-resiliency through automated model transformations that introduce high-assurance components into a system; in particular, filters that guard against
malformed input, as well as monitors that guard against spoofing and other malicious behavior.  
A formal specification defines each high-assurance component and is used to verify that the component addresses system-level cyber requirements.
Implementations for these high-assurance components are directly synthesized from their specifications.  
The model transformations are integrated into the Open Source AADL Tool Environment (OSATE).
The manuscript discusses a real-world case study by verification experts and reports on their experience working in the environment.

The new manuscript submitted here for consideration expands on the MODELS 2021 manuscript with the following original contributions:
\begin{itemize}
    \item A new, heavily revised to be more intuitive, example to motivate and explain the MBSE environment, the need for cyber-assurance, and the role of formal verification is synthesizing high-assurance components to cyber-harden a system.
    \item A section that formally defines the assume-guarantee reasoning used to prove cyber-requirements of a system design with its requisite verification conditions that are dispatched by a model checker.
    \item A new section that formally defines the syntax and semantics of the specification language to describe contract specifications on components and systems.
    \item A new section that formally defines a code contract that is a specialized specification for high-assurance components, such as filters and monitors, that are sufficient for proving cyber-properties of the system and for synthesis to an actual implementation.
    \item A heavily revised section on the synthesis process that starts with a code contract and ends with CakeML code describing is some considerable detail how the synthesis works and is proved correct.
    \item A new real-world case study with system developers that are not expert with assume-guarantee reasoning to show feasibility with such developers. 
\end{itemize} 
The authors feel this submitted manuscript represents a new and original contribution over the prior publication.

I hope that you find this manuscript of interest to your readers. Thank you for your consideration and your generosity in working with me, and my coauthors, in extending the submission deadline.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2022-03-28 paper/high-assurance-contracts.tex
Fix label on the monitor contract

@@
 
-    eq policy : bool = (rsp => req) ->
+    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
                          (    (isPending => latency < MAX_LATENCY)
                           and (rsp => (req or pre(isPending))));
-    *\label{line:mon-policy}*
     -- stop user definitions

### 2022-03-28 paper/high-assurance-contracts.tex
Revise to match pascalish version

@@
 
+    eq preIsPending : bool = pre(isPending);*\label{line:mon-pre-pending}*
     eq isPending : bool = Since(not rsp, req and not rsp);*\label{line:mon-pending}*
@@
                          (    (isPending => latency < MAX_LATENCY)
-                          and (rsp => (req or pre(isPending))));
+                          and (rsp => (req or preIsPending)));
     -- stop user definitions
@@
 
-\lineref{line:mon-pending} and \lineref{line:mon-latency}.
+\lineref{line:mon-pre-pending}, \lineref{line:mon-pending}, and \lineref{line:mon-latency}.
 The monitor policy is defined by marking when a request is not
@@
 otherwise increments by one. 
+The \texttt{preIsPending} is needed to make the contract Pascalish since since the current and previous value of \texttt{isPending} are used in the definition of \texttt{policy}.
+
+\lineref{line:mon-policy}. 
+The policy makes sure the latency is bounded whenever there is a pending request, and it it makes sure that a response is a result of a request.
+The initial step differentiates from later steps in only needing to be sure a response is a result of a request.
 

### 2022-03-28 paper/high-assurance-contracts.tex
Slight wordsmithing

@@
 \figref{fig:monitor}.
-\lineref{line:mon-alert}. The \texttt{alert} variable
+\lineref{line:mon-alert}: The \texttt{alert} variable
 (distinct from the \texttt{Alert} port) depends not just on the policy
@@
 
-\lineref{line:mon-pre-pending}, \lineref{line:mon-pending}, and \lineref{line:mon-latency}.
+\lineref{line:mon-pre-pending}, \lineref{line:mon-pending}, and \lineref{line:mon-latency}:
 The monitor policy is defined by marking when a request is not
@@
 
-\lineref{line:mon-policy}. 
+\lineref{line:mon-policy}: 
 The policy makes sure the latency is bounded whenever there is a pending request, and it it makes sure that a response is a result of a request.
@@
 
-\lineref{line:mon-const}. The latency bound for the monitor is defined by the constant
+\lineref{line:mon-const}: The latency bound for the monitor is defined by the constant
 \texttt{MAX\_LATENCY}.  Here the constant
@@
 
-\lineref{line:mon-assume}. The policy definition only works if there is never more than one
+\lineref{line:mon-assume}: The policy definition only works if there is never more than one
 outstanding request at a time; otherwise, the latency counter resets

### 2022-03-28 paper/testing.tex
Slight wordsmithing

@@
 \begin{compactitem}
-\item \lineref{line:tst-index}. An index to use to define the test input and expected test output through time.
-\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}. The test input. These define when the request and response message events should appear relative to the index.
-\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}. The expected test output. 
+\item \lineref{line:tst-index}: An index to use to define the test input and expected test output through time.
+\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: The test input. These define when the request and response message events should appear relative to the index.
+\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: The expected test output. 
 The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
-\item \linesref{line:tst-imp-start}{line:tst-imp-end}. The implementation with the component under test (\lineref{line:tst-imp-comp}).
+\item \linesref{line:tst-imp-start}{line:tst-imp-end}: The implementation with the component under test (\lineref{line:tst-imp-comp}).
 \end{compactitem}

### 2022-03-28 paper/testing.tex
Wordsmithing in 'mistakes' paragraph

@@
 
-There are three common mistakes in specifications, aside from self-consistency that is already checked, that have been observed: vacuity, under-specification, and missed corner cases.
-Vacuity is a common pitfall with implication when the left hand side is always false making the implication true. 
-Under-specification is very challenging because verification is not able to prove anything (or it can prove everything in the case of existential properties) so nothing is actually known after verification except that the system is under-specified.
+We have observed three common types of mistakes in specifications (aside from self-consistency, which is
+already checked): vacuity, under-specification, and missed corner cases.
+Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true. 
+Under-specification is very challenging, as a verifier in the face of under-specification is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the system is under-specified.
 Missed corner cases is a common problem to all disciplines.
-It is especially hard here because there is no direct way to provide test input and check test output.
+It is an especially difficult challenge here because there is no direct way to provide test input and check test output.
 

### 2022-03-29 paper/high-assurance-contracts.tex
Add guard on pre in contract

@@
 
-    eq preIsPending : bool = pre(isPending);*\label{line:mon-pre-pending}*
+    eq preIsPending : bool = false -> pre(isPending);*\label{line:mon-pre-pending}*
     eq isPending : bool = Since(not rsp, req and not rsp);*\label{line:mon-pending}*

### 2022-03-29 paper/related-work.tex
Wordsmithing in Lustre compiler paragraph

--- 2021-05-15 paper/related-work.tex
@@
 
-Data-flow semantics are well studied \cite{10.1145/41625.41641,97300, 10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}.
-State machine semantics can be added to synchronous data-flow \cite{10.1145/1086228.1086261}.
-The idea is to translate imperative constructs into equivalent synchronous data flow constructs.
-The resulting Lustre can then be compiled.
-The goal is to provide a seamless connection between pure data-flow and pure control design.
+Data-flow semantics are well studied \cite{10.1145/41625.41641,97300,
+10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}. State
+machine semantics can be added to synchronous
+data-flow \cite{10.1145/1086228.1086261}. The idea is to translate
+imperative constructs into equivalent synchronous data flow
+constructs. The resulting Lustre can then be compiled. The goal is to
+provide a seamless connection between pure data-flow and pure control
+design.
 
-There is a fully verified compiler that takes Lustre and turns it into a binary executable that is specified and verified in Coq \cite{10.1145/3140587.3062358}.
-The key is in combining infinite sequences of data-flow models with incremental manipulation of memories akin to an imperative model.
-CompCert is used on the backend to create the final rendered executable.
+A fully verified compiler has been developed that takes Lustre and compiles
+it into assembly code, specified and verified in Coq \cite{10.1145/3140587.3062358}.
+The key is in combining infinite sequences of data-flow models with incremental
+manipulation of memories akin to an imperative model. CompCert is used on the
+backend to create the final rendered executable. The work is quite similar to
+ours in providing a verified compilation chain from Lustre. However,
+there are significant differences: notably, they tackle a more
+comprehensive subset of Lustre than our work. We have been able to
+work effectively with a relatively simple core language, \eg, without
+\emph{clocking} features.  This subset aligns with our contract language, and
+allows us to generate behaviorally equivalent code so we can show that the
+generated code meets its logical specification.  This, in turn,
+can be used as a leaf-level claim in a system-wide correctness argument
+conducted by compositional reasoning in AGREE.

### 2022-03-29 paper/testing.tex
More wordsmithing in beginning of Testing section

@@
 discuss how code contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
-Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
-The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.
+Writing and proving properties about AGREE specifications is difficult not just because it is a formal process but because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
+Case studies have shown that engineers, as well as formal method practitioners, make mistakes writing these specifications, especially if the computation or temporal reasoning is complex.
 

### 2022-03-29 paper/abstract.tex
delete mention of CASE from abstract

@@
 cyber-attacks in the same way they are tolerant to random faults: they
-either gracefully recover or safely shut down as requirements dictate.
-The DARPA Cyber Assured Systems Engineering program is developing
-tools for design, analysis, and verification that enable systems
-engineers to design-in cyber-resiliency in a Model-Based Systems
-Engineering (MBSE) environment.  This paper describes automated model
-transformations that introduce high-assurance cyber-resiliency
-components into a system; in particular, filters that guard against
-malformed input, as well as monitors that guard against
-spoofing and other malicious behavior.  A formal specification in the
-form of a code contract defines each high-assurance component and is
-used to verify that the component addresses system-level cyber
-requirements.  Implementations for these high-assurance components
-are directly synthesized from their code contracts, and are
-automatically proven to preserve the exact meaning of the code
-contract all the way down to the binary code level.  The model
-transformations are integrated into the Open Source AADL 
-Tool Environment (OSATE).  We further report on two case
-studies, one by formal method experts and the other by non-experts,
-applying model transformations to two different systems to improve
-cyber resiliency using high-assurance filters and monitors.  One is a
-UAV system utilizing the Air Force Research Laboratory's
-OpenUxAS services for route planning.  The other is a tablet used for
-situational awareness in a CH-47F Chinook helicopter application.
-The case studies demonstrate that our MBSE approach is feasible in
-real-world systems engineering.
+should either gracefully recover or safely shut down as requirements dictate.
+%
+%% The DARPA Cyber Assured Systems Engineering program is developing
+%% tools for design, analysis, and verification that enable systems
+%% engineers to design-in cyber-resiliency in a Model-Based Systems
+%% Engineering (MBSE) environment.
+%
+This paper describes automated model transformations that introduce
+high-assurance cyber-resiliency components into a system; in
+particular, filters that guard against malformed input, as well as
+monitors that guard against spoofing and other malicious behavior. A
+formal specification in the form of a code contract defines each
+high-assurance component and is used to verify that the component
+addresses system-level cyber requirements. Implementations for these
+high-assurance components are directly synthesized from their code
+contracts, and are backed up by proofs showing that high-level
+specifications map in a menaing-preserving way to code. The model
+transformations are integrated into the Open Source AADL Tool
+Environment (OSATE).  We further report on two case studies, one by
+formal method experts and the other by non-experts, applying model
+transformations to two different systems to improve cyber resiliency
+using high-assurance filters and monitors.  One is a UAV system
+utilizing the Air Force Research Laboratory's OpenUxAS services for
+route planning.  The other is a tablet used for situational awareness
+in a CH-47F Chinook helicopter application.  The case studies
+demonstrate that our MBSE approach is feasible in real-world systems
+engineering.

### 2022-03-29 paper/agree-semantics.tex

@@
 \]
-There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. To support port handling, there is a predicate
+AGREE includes a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
+and formulas. To support port handling, AGREE defines a predicate
 $\konst{event}(p)$, which tests if an event (or event-data) port $p$ is
-signaled. Lastly, there are temporal operators \emph{previously}
+signaled. Lastly, AGREE supports temporal operators \emph{previously}
 $\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically
@@
  Operation & Math. notation & AGREE \\ \hline
- followed-by &  $e \to e$          & e \verb+->+ e  \\
- logical implication &  $e \imp e$ & e \verb+=>+ e \\
- if-and-only-if &  $e \iff e$      & e \verb+<=>+ e \\
- variable definition & $v = e$     & \konst{eq} v\ e \\
- Historically  & $\konst{Hist}(b)$ & \konst{Historically}(b) \\
+ followed-by &  $e \to e$          & \verb+e -> e+  \\
+ logical implication &  $e \imp e$ & \verb+e => e+ \\
+ if-and-only-if &  $e \iff e$      & \verb+e <=> e+ \\
+ logical and &  $e \land e$      & \verb+e and e+ \\
+ logical or &  $e \lor e$      & \verb+e or e+ \\
+ logical not & $\neg b$      & \verb+not b+ \\
+ variable definition & $v = e$     & \verb+eq v e+ \\
+ Previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
+ Historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
 \end{tabular}

### 2022-03-29 paper/high-assurance-contracts.tex
clarifying exposition of behavior of hardened monitor

@@
       WELL_FORMED_AUTOMATION_RESPONSE(Input);
-    -- stop user definitions
+    -- end user definitions
 
@@
                           and (rsp => (req or preIsPending)));
-    -- stop user definitions
+    -- end user definitions
 
@@
 \subsubsection{Monitor specification}
-The monitor policy is given by the \konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
-\figref{fig:monitor}.
-\lineref{line:mon-alert}: The \texttt{alert} variable
-(distinct from the \texttt{Alert} port) depends not just on the policy
-but also on configuration details supplied when the component is
-specified.  At configuration time, \texttt{is\_latched} has been set
-to \konst{true}, thus making the alert persistent, meaning that once
-the alert is raised, it is always raised; otherwise, it is the
-complement of the policy value in the current step.
 
-\lineref{line:mon-pre-pending}, \lineref{line:mon-pending}, and \lineref{line:mon-latency}:
-The monitor policy is defined by marking when a request is not
-satisfied, \texttt{isPending}, and
-counting the number of steps between requests, \texttt{latency}.  
-The \texttt{isPending} variable is true
-whenever there is a request that does not coincide with a
-response--to be read as \emph{there is no response since the most recent request without a corresponding response}.  The
+The monitor informally behaves as follows: if the \emph{policy} is
+obeyed in the current step, and the alert has not yet been raised,
+then the \texttt{Response} input is sent out on \texttt{Output}, and
+the \texttt{Alert} output stays
+low \lineref{line:mon-alert}. Otherwise \texttt{Alert} goes high and
+stays high, and there is nothing sent henceforth on \texttt{Output}.
+
+The \texttt{alert} variable (distinct from the \texttt{Alert} port)
+depends not just on the policy but also on the
+\texttt{is\_latched} constant, which has been set at configuration time.
+Since \texttt{is\_latched} is \konst{true}, the alert is persistent,
+meaning that once the alert is raised, it is always raised; otherwise,
+it is the complement of the policy value in the current step.
+
+The monitor policy is defined by marking, with variable \texttt{isPending},
+when a request is made but not yet satisfied, and by counting the
+number of steps (\texttt{latency}) between
+requests. See \linesref{line:mon-pre-pending}{line:mon-latency}:
+The \texttt{isPending} variable is true whenever there is a request
+that does not coincide with a response, \ie, \texttt{isPending} holds
+when \emph{there is no response since the most recent request without
+a corresponding response}.  The
 \texttt{latency} variable starts at zero, resets on every request, and
-otherwise increments by one. 
-The \texttt{preIsPending} is needed to make the contract Pascalish since since the current and previous value of \texttt{isPending} are used in the definition of \texttt{policy}.
+otherwise increments by one.  Note that the \texttt{preIsPending}
+variable is a temporary needed to make the contract \emph{Pascalish};
+otherwise, the current and previous value of \texttt{isPending} would
+be simulataneously accessed in the definition of \texttt{policy}.
 
-\lineref{line:mon-policy}: 
-The policy makes sure the latency is bounded whenever there is a pending request, and it it makes sure that a response is a result of a request.
-The initial step differentiates from later steps in only needing to be sure a response is a result of a request.
+The policy (\lineref{line:mon-policy}) makes sure the latency is
+bounded whenever there is a pending request, and it ensures that a
+response is a result of a request.  The initial step differentiates
+from later steps in only needing to be sure a response is a result of
+a request, both happening in the initial step.
 
-\lineref{line:mon-const}: The latency bound for the monitor is defined by the constant
-\texttt{MAX\_LATENCY}.  Here the constant
-is set to one to be consistent with the system specification from the
+The latency bound for the monitor is defined by the constant
+\texttt{MAX\_LATENCY} (\lineref{line:mon-const}).  In this example the
+constant is set to one to be consistent with the system specification from the
 previous section.  At time zero the policy is trivial: a response
@@
 
-\lineref{line:mon-assume}: The policy definition only works if there is never more than one
+The policy definition only works if there is never more than one
 outstanding request at a time; otherwise, the latency counter resets
-incorrectly.  That requirement is encoded in the assumption and is the same assumption used for the
-system input.  The policy can be written without the assumption,
-but the policy is much simpler to write with the assumption.
+incorrectly.  That requirement is encoded in the assumption
+(\lineref{line:mon-assume}) and is the same assumption used for the
+system input.  In fact, the policy can be written without the assumption, but
+the policy is much simpler to write with the assumption.
 
@@
 data, spontaneous responses, or delayed responses.  Synthesis must
-generate faithful implementations of the code contracts for the AGREE verification result to apply to the
-deployed system.
-
-%%   Precise in this context means that the
-%% implementations match the input-to-output relations defined in the
-%% specifications.
+generate faithful implementations of the code contracts for the AGREE
+verification result to apply to the deployed system.

### 2022-03-29 paper/synthesis.tex
clarifying exposition of behavior of hardened monitor

@@
 \begin{lstlisting}[style=myML]
-stepFn (Request,Response)
-       (req,rsp,current,previous,policy,alert) =
+fun stepFn (response,request)
+    (rsp,req,preIsPending,isPending,latency,policy,alert)
 let val stateVars' =
       if !initStep then
-        let val req = event(Request)
-            val rsp = event(Response)
-            val current = (req = rsp)
-            val previous = req and not(rsp)
-            val policy = current or previous
-            val alert = not policy
-            val () = (intStep := False)
-        in (req,rsp,current,previous,policy,alert)
+      let val rsp = Option.isSome response
+          val req = Option.isSome request
+          val preIsPending = False
+          val isPending = req andalso not rsp
+          val latency = 0
+          val policy = (if rsp then req else True)
+          val alert = not(policy)
+          val () = (initStep := False)
+      in
+        (rsp,req,preIsPending,isPending,latency,policy,alert)
       end
       else
-        let val req = event(Request)
-            val rsp = event(Response)
-            val current = (req = rsp)
-            val previous = pre(req and not rsp) and (not req and rsp)
-            val policy = current or previous
-            val alert = (is_latched and pre(alert)) or not(policy)
-        in (req,rsp,current,previous,policy,alert)
+      let ...
+          val rsp = Option.isSome response
+          val req = Option.isSome request
+          val preIsPending = pre isPending
+          val isPending = (req andalso not rsp) orelse
+                          ((not rsp) andalso pre isPending)
+          val latency = (if req then 0 else (pre latency + 1)
+          val policy =
+            ((if isPending then
+                (latency < Defs.max_latency) else True)
+              andalso
+              (if rsp then
+                 (req orelse preIsPending) else True))
+          val alert = (Defs.is_latched andalso (pre alert))
+                      orelse not(policy)
+      in
+        (rsp,req,preIsPending,isPending,latency,policy,alert)
       end
-    val (_,rsp',_,_,_,alert') = stateVars'
-    val Alert = if alert' then Some () else None
-    val Output =
+   val (rsp',_,_,_,_,_,alert') = stateVars'
+   val alertPort = if alert' then Some () else None
+   val outputPort =
      if alert' then None else
-       if rsp'   then Some Response
+     if rsp' then Some response
      else None
 in
-   (stateVars', (Alert,Output))
+   (stateVars', (alertPort,outputPort))
 end

### 2022-03-29 paper/conclusion.tex
Changed tense on most recent CASE work

@@
-The DARPA CASE program is creating tools for systems engineers to
+The DARPA CASE program has created tools for systems engineers to
 integrate cyber-vulnerability analysis and mitigation into their
@@
 
-We are currently in the process of applying BriefCASE to the design of
+Most recently, we have applied BriefCASE to the design of
 an application using the Collins Common Avionics Architecture System

### 2022-03-29 paper/synthesis.tex
checkpointing; Section 7 proof story fleshed out

@@
 assurance components specified by code contracts; for each such
-occurrence it generates a CakeML program. More generally, SPLAT supports
-three kinds of behavioral specification from which to generate
-code. In increasing order of expressiveness, these are:
+occurrence it generates a CakeML program. SPLAT supports three kinds
+of behavioral specification from which to generate code. In increasing
+order of expressiveness, these are: regular expressions, contiguity
+types, and AGREE specifications. In earlier work we have described how
+regular expressions and contiguity types are used as specification
+languages for verified filter implementations. In the present paper,
+we extend this \emph{specification oriented} approach to the
+Turing-complete specifications of AGREE.  We give a short review of
+that previous work before going into detail on AGREE.
 
@@
 implementations. In \cite{formal-filter-synth-langsec,case-verified-filter}
-we report on the creation and application of filters from a verified
-regular expression-to-DFA compiler.
+we report on the creation and application of table-based DFA filters
+from a verified regular expression-to-DFA compiler. The latter paper
+offers a fully-worked out example synthesizing a regular expression
+filter and formally proving that it meets an infinitary behavioral
+specification.
 
 \paragraph{Contiguity types\cite{contiguity-types}.}
-A contiguity type, like regular expressions, is effective at
+Contiguity types, similar to regular expressions, are effective at
 describing data at the network message level, \eg, as flat
-strings. The notation is however more expressive, being able to
+strings. The notation is however much more expressive, being able to
 declare many kinds of so-called \emph{self-describing} messages, \ie,
@@
 specification for the welll-formed waypoints of our example system. We
-have formalized and verified a contiguity type matcher. The matcher
-code has been used in our case studies.
+have formalized and verified a contiguity type matcher, which has been
+extensively applied throughout our case studies.
 
-\paragraph{AGREE code contracts.} When full computational power is needed, we can use the
-expressive power of AGREE code contracts to write arbitrarily complex
-filter or monitor operations, as we have already seen. Our formal
-model of AGREE syntax and semantics is applied to extract code
-contracts and map them to equivalent logic functions.
+\subsection{Compiling from code contracts}
 
-\noindent For each of these three approaches, we take the following steps to
+When full computational power is needed, we can use the expressive
+power of AGREE code contracts to write arbitrarily complex filter or
+monitor operations, as we have already seen. Our formal model of AGREE
+syntax and semantics is applied to extract code contracts and map them
+to equivalent logic functions. In particular, given a code contract with
+functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state
+variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics
+given in Section \ref{sec:code-contracts} can be applied to create a
+function
+\[
+ \konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env} = \sem{{\Eqs}\cdot G}
+\]
+Assume ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially
+imperative. Then we can direcctly create a corresponding logic function
+\[
+\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}
+\]
+that takes input and state as parameters and returns an updated state
+and output. Thus the environment $\mathit{Env}$ holding all previous
+values of each variable can be replaced by \emph{parameters}, each of
+which is able to hold only one data element. Assuming
+$\mathit{Env}' = \konst{strmStep}\;t\; \mathit{Env}$, we can formally show that
+\[
+\vdash (\mathit{Env}' \vert_{\mathbb{S}},
+ \mathit{Env}' \vert_{\mathbb{O}}) =
+\konst{stateStep} (\mathit{Env}\vert_{\mathbb{I}(t)},
+                   \mathit{Env}\vert_{\mathbb{S}(t)})
+\]
+In other words, projecting out the state and output components from
+$\konst{strmStep}\;t\; \mathit{Env}$ equals the result of
+applying \konst{stateStep} on the input and state components of
+$\mathit{Env}$ at time $t$. This theorem connects the stream semantics
+of the code contract to a pure logic function over the values
+appearing in the stream. From this point we can use tools in the
+CakeML eco-system to jump from the setting of higher order logic to
+CakeML abstract syntax \cite{cakeml-translator}. Thus we have moved
+from stream semantics to logic function to CakeML evaluation semantics
+in a property-preserving way. Once in CakeML, further necessary
+features can be added, namely imperative variables for storing state
+between component invocations, also invocations of the (verified) foreign
+function interface of CakeML. Finally, the CakeML compiler can be
+applied to the program. It delivers an executable which can be shown
+to preserve the original contract.
+
+%% where the $\mathit{state}$ parameters to the function are just the
+%% variables defined in ${\Eqs}$, obtained by projecting out the
+
+%%  = \mathit{Env}\vert_{\mathbb{I}}$ $\mathit{input}
+%%  = \mathit{Env}\vert_{\mathbb{I}}$ $\mathit{input}
+%%  = \mathit{Env}\vert_{\mathbb{I}}$
+
+\subsection{Generating CakeML}
+For any of the three specification languages, we take the following general steps to
 arrive at an executable.
@@
 The verdict of a filter is made and performed within one thread
-invocation. Thus, in its given time slice, the
-following steps must be completed:
+invocation. Thus, in its given time slice, the following steps must be
+completed:
 
@@
 the wellformedness computation to finish. Thus resource bounds need to
-be included in the correctness argument. Some preliminary work on this
-has been done in CakeML \cite{cakeml-space-cost}.
+be included in the correctness argument. Some preliminary work that
+could support formal proofs of such bounds has been done for
+CakeML \cite{cakeml-space-cost}.
 
@@
 abstract type:
+%
 \[
@@
 \]
-
-The system scheduler \emph{activates} components in some order. It is
-an obligation on the system that the scheduler follows some sensible
-partial order of component activation and allows each component
-sufficient time for its computation.  Activating a monitor component
-takes the form of the following pseudo-code, in which the monitor
-evaluates the \konst{stepFn} on its current inputs and the current
-values of the state variables, returning the new state and the output
-values.
-\[
-\begin{array}{ll}
- \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
- (v_1,\ldots) & = \konst{readState}() ; \\
- ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
- \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
- \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
-\end{array}
-\]
-
-\subsubsection{Initialization}
-
-A monitor may need to accumulate a certain minimum number of
-observations before being able to make a meaningful assessment of
-behavior. Until that threshold is attained, the monitor is essentially
-in its \emph{initialization} phase. In order for correct code to
-be generated, monitor specifications need to spell out the values of
-output ports when in their initialization phases. For example, suppose a
-monitor does some kind of differential assessment of inputs at
-adjacent time slices, alerting when (say) the measured location of a
-UAV at times $t$ and $t+1$ is such that the distance between the two
-locations is unusually large. Such a monitor needs two measurements
-before making its first judgement, but at the time of its first
-output, only one measurement will have been made. The specification
-must then explicitly state the correct value for the first output.
-
-\subsubsection{Step function}
-
 The \konst{stepFn} works as follows:
-
 \begin{enumerate}
@@
 
+\subsection{Incorporation into system execution}
+
+The system scheduler \emph{activates} components in some order. It is
+an obligation on the system that the scheduler follows some sensible
+partial order of component activation and allows each component
+sufficient time for its computation.  Activating a monitor component
+takes the form of the following pseudo-code, in which the monitor
+evaluates the \konst{stepFn} on its current inputs and the current
+values of the state variables, returning the new state and the output
+values.
+\[
+\begin{array}{ll}
+ \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
+ (v_1,\ldots) & = \konst{readState}() ; \\
+ ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
+ \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
+ \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
+\end{array}
+\]
+
 %% \subsection{Component Behavior}
@@
 %% initial states produces $\Lang{s}$ itself.
+
+%% \subsubsection{Initialization}
+
+%% A monitor may need to accumulate a certain minimum number of
+%% observations before being able to make a meaningful assessment of
+%% behavior. Until that threshold is attained, the monitor is essentially
+%% in its \emph{initialization} phase. In order for correct code to
+%% be generated, monitor specifications need to spell out the values of
+%% output ports when in their initialization phases. For example, suppose a
+%% monitor does some kind of differential assessment of inputs at
+%% adjacent time slices, alerting when (say) the measured location of a
+%% UAV at times $t$ and $t+1$ is such that the distance between the two
+%% locations is unusually large. Such a monitor needs two measurements
+%% before making its first judgement, but at the time of its first
+%% output, only one measurement will have been made. The specification
+%% must then explicitly state the correct value for the first output.
+
+% \subsubsection{Step function}

### 2022-03-29 paper/synthesis.tex
Done Section 7

@@
 which is able to hold only one data element. Assuming
-$\mathit{Env}' = \konst{strmStep}\;t\; \mathit{Env}$, we can formally show that
+$\mathit{Env}' = \konst{strmStep}\;t\; \mathit{Env}$, we can formally prove that
 \[
@@
 between component invocations, also invocations of the (verified) foreign
-function interface of CakeML. Finally, the CakeML compiler can be
+function interface of CakeML \cite{cakeml-monadic}. Finally, the CakeML compiler can be
 applied to the program. It delivers an executable which can be shown
 to preserve the original contract.
-
-%% where the $\mathit{state}$ parameters to the function are just the
-%% variables defined in ${\Eqs}$, obtained by projecting out the
-
-%%  = \mathit{Env}\vert_{\mathbb{I}}$ $\mathit{input}
-%%  = \mathit{Env}\vert_{\mathbb{I}}$ $\mathit{input}
-%%  = \mathit{Env}\vert_{\mathbb{I}}$
 
@@
 \end{enumerate}
-
-\begin{remark}[Partiality]
-
-Partiality is an important consideration: steps 2 and 3 above can
-fail; the data might not be parseable or the wellformedness
-computation could be badly written and fail at runtime. In such cases,
-the filter should recover and yield control without passing the input
-onwards. In these cases, the filter is behaving as it should, but we
-must also guard against situations in which a \emph{correctly
-specified} filter fails at runtime. This kind of defect arises when
-the filter \emph{ought} to accept a message, but lack of resources
-results in the filter failing to do so. For example, the parse of a
-message might need more space than has been allocated; another example
-could be if the time slice provided by the scheduler is too short for
-the wellformedness computation to finish. Thus resource bounds need to
-be included in the correctness argument. Some preliminary work that
-could support formal proofs of such bounds has been done for
-CakeML \cite{cakeml-space-cost}.
-
-\end{remark}
 
@@
 
-Monitors are intended to track and analyze the externally visible
-behavior of system components through time. Therefore, they tend to
-require more extensive computational ability than filters. In
-particular, our basic notion of a monitor is that it embodies a
-relation over its input and output streams, and is able to access the
-value of a stream at any earlier point in time, if necessary. Monitors
-commonly use state to keep track of earlier values, unlike filters
-which, for us, are typically stateless components. (However, there is
-nothing in our approach that forbids stateful filters: they can be
-realized by monitors.) A monitor specification is mapped by code
-generation to a state transformation function of the following
-abstract type:
+%% Monitors are intended to track and analyze the externally visible
+%% behavior of system components through time. Therefore, they tend to
+%% require more extensive computational ability than filters. In
+%% particular, our basic notion of a monitor is that it embodies a
+%% relation over its input and output streams, and is able to access the
+%% value of a stream at any earlier point in time, if necessary. Monitors
+%% commonly use state to keep track of earlier values, unlike filters
+%% which, for us, are typically stateless components. (However, there is
+%% nothing in our approach that forbids stateful filters: they can be
+%% realized by monitors.)
+As discussed above, a monitor specification is
+mapped to a state transformation function
 %
 \[
-\konst{stepFn} : \mathit{input} \times \mathit{stateVars} \to \mathit{stateVars} \times \mathit{output}
+\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}.
 \]
-The \konst{stepFn} works as follows:
+which computes as follows:
 \begin{enumerate}
 
-\item Each input is parsed into data of the type specified by the port
+\item Each input buffer is parsed into data of the type specified by the port
   type;
@@
 \item New values for the state variables are computed, in dependency
-  order. The discussion above on initialization now comes into
-  play. Suppose the variable declarations have the following form:
+  order. At this point variables defined by \emph{followd-by} are
+  dealt with. Suppose the variable definitions have the following
+  form:
 \[
@@
 \]
-In the generated code, for the first invocation of \konst{stepFn} only,
+In the generated function, for the first invocation of \konst{stateStep} only,
 the initializations are executed in order:
 \[
-\begin{array}{l}
-  v_1 = i_1; \\
-  \cdots \\
-  v_n = i_n; \\
-\end{array}
+  v_1 = i_1; \cdots ;  v_n = i_n;
 \]
-In all subsequent steps, the \emph{non-initialization} assignments are performed:
+In all later invocations, the \emph{non-initialization} assignments are performed, also in order:
 \[
-\begin{array}{l}
-  v_1 = e_1; \\
-  \cdots \\
-  v_n = e_n; \\
-\end{array}
+  v_1 = e_1; \cdots ; v_n = e_n; \\
 \]
@@
 \begin{lstlisting}[style=myML]
-fun stepFn (response,request)
+fun stateStep (response,request)
     (rsp,req,preIsPending,isPending,latency,policy,alert)
@@
 takes the form of the following pseudo-code, in which the monitor
-evaluates the \konst{stepFn} on its current inputs and the current
+evaluates the \konst{stateStep} on its current inputs and the current
 values of the state variables, returning the new state and the output
@@
  (v_1,\ldots) & = \konst{readState}() ; \\
- ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stepFn} ((i_1,\ldots),(v_1,\ldots)) ; \\
+ ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stateStep} ((i_1,\ldots),(v_1,\ldots)) ; \\
  \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
@@
 \]
+
+\paragraph{Partiality.} Partiality is an important consideration for both filters and
+monitors: the input data might not be parseable or the wellformedness
+computation of a filter could be badly written and fail at runtime. In
+such cases, the filter should recover and yield control without
+passing the input onwards. For these failure cases, the filter is
+behaving as it should, but we must also guard against situations in
+which a \emph{correctly specified} filter or monitor fails at
+runtime. This kind of defect arises when a filter, for
+example, \emph{ought} to accept a message, but lack of resources
+results in the filter failing to do so. For example, the parse of a
+message might need more space than has been allocated; another example
+could be if the time slice provided by the scheduler is too short for
+the wellformedness computation to finish. Thus resource bounds need to
+be included in the correctness argument. Some preliminary work that
+could support formal proofs of such bounds has been done for
+CakeML \cite{cakeml-space-cost}.
 

### 2022-03-29 paper/case-study.tex
checkpointing; Section 7 proof story fleshed out
Minor edits to case study

@@
 In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
-The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design,
-and we need to satisfy these requirements by modifying the design.
+The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the model,
+and we need to satisfy these requirements by modifying the model.
 %  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
 %
-For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) will not impact other processes.
+For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}),
+the cyber analysis tools generate requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) will not impact other processes.
 
@@
 These include four \textit{wellformedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
-Requirements are imported into the model as goals in a Resolute assurance case.
+Requirements are imported into the model as goals in a Resolute assurance case \cite{resolute-destion}.
 Because we can run Resolute at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
 
-The intent of the \textit{wellformedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
-In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
-By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
-The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
+The intent of the \textit{wellformedness} requirements is to prevent
+malformed messages from causing a buffer overrun or code injection
+attack.  In the UAV design, such messages are most likely to originate
+from a remote source or the uncontrolled UxAS component.  By placing
+filters on the connections upstream of mission-critical components,
+such attacks could be mitigated.  The Filter transform is therefore
+applied for each wellformedness requirement, inserting filter
+components on the incoming and outgoing UxAS connections.
 

### 2022-03-30 paper/abstract.tex
Remove mention of 2nd case study

@@
 contracts, and are backed up by proofs showing that high-level
-specifications map in a menaing-preserving way to code. The model
+specifications map in a meaning-preserving way to code. The model
 transformations are integrated into the Open Source AADL Tool
-Environment (OSATE).  We further report on two case studies, one by
-formal method experts and the other by non-experts, applying model
-transformations to two different systems to improve cyber resiliency
-using high-assurance filters and monitors.  One is a UAV system
+Environment (OSATE). 
+We further report a case study applying the model
+transformations to a UAV system
 utilizing the Air Force Research Laboratory's OpenUxAS services for
-route planning.  The other is a tablet used for situational awareness
-in a CH-47F Chinook helicopter application.  The case studies
-demonstrate that our MBSE approach is feasible in real-world systems
+route planning.  The case study
+demonstrates that our MBSE approach is feasible in real-world systems
 engineering.
+
+% We further report on two case studies, one by
+% formal method experts and the other by non-experts, applying model
+% transformations to two different systems to improve cyber resiliency
+% using high-assurance filters and monitors.  One is a UAV system
+% utilizing the Air Force Research Laboratory's OpenUxAS services for
+% route planning.  The other is a tablet used for situational awareness
+% in a CH-47F Chinook helicopter application.  The case studies
+% demonstrate that our MBSE approach is feasible in real-world systems
+% engineering.

### 2022-08-17 paper/overview.tex
Add BriefCASE overview for big picture

% Copied from DESTION paper
The BriefCASE environment provides systems engineers with a workflow and tool support for developing
products with inherent cyber-resiliency.

The  workflow starts with the development of a baseline AADL model of the system architecture
focusing on the desired functionality. This model can be analyzed using any of the existing AADL 
tools (e.g., resource usage, information flow, latency) to determine whether it is acceptable.
BriefCASE integrates additional tools that analyze the architecture model for cybersecurity vulnerabilities and
generate requirements that, when addressed, will mitigate those vulnerabilities.
These requirements are imported into the model and may be addressed using a 
collection of automated model transforms. As requirements are addressed in the design, an assurance case is updated with
corresponding evidence, computed directly from the model or by supporting analysis tools.  
Code implementing new high-assurance components as well as communication and execution infrastructure
is generated from the model along with associated assurance evidence.

The BriefCASE tools and their interactions are shown in \figref{fig:tool-arch}.   
The following sections describe each step of the workflow in more detail, corresponding
to the tools and artifacts shown in green in the figure.

\begin{figure*}
	\begin{center}
	  \includegraphics[width=\textwidth]{./figs/tool-arch.jpg}
  	\end{center}
	\caption{BriefCASE tool architecture.  Tools and code/artifacts discussed in the BriefCASE workflow are shown in green.} 
	\label{fig:tool-arch} 
\end{figure*}

\subsection{Requirements Analysis}

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and
DCRYPPS~\cite{dcrypps2019}) that examine AADL models, report potential cyber vulnerabilities,
and suggest requirements for mitigation.
%
A BriefCASE project contains a repository for requirements. Imported requirements are represented 
as assurance case goals to be satisfied. 
The goal is marked {\em undeveloped} initially.
Evidential statements are added to the goal as the design is updated to address this requirement.
Some requirements are formalized as assume-guarantee contracts added to the AADL model.
Formal verification provides the evidence for the satisfaction of these requirements.

\subsection{Cyber Transforms}

To address the new cyber-resiliency requirement, the architecture will need to be transformed in such a way as to harden the design against the vulnerability.
BriefCASE provides an extendable library of model transformations for addressing common cyber vulnerabilities.  Currently, the following transformations are supported:

\begin{compactitem}
	\item Filter -- Blocks messages that do not conform to the given specification
	\item Monitor -- Detects violations of a given run-time condition and generates an alert
	\item Switch -- Used with a Monitor to block messages when an alert is generated (also referred to as a \textit{gate})
	\item Attestation -- Performs a measurement on non-local software to assess its trustworthiness
	\item Virtualization -- Isolates software component(s) in a virtual machine
	\item Proxy -- Inserts a pair of components to allow inspection of %https
		HTTPS message payloads
	\item seL4 -- Transforms the model to comply with seL4 component properties
\end{compactitem}

The transformations are automated by the BriefCASE tool to create a hardened model that is correct-by-construction.  
These transformations include updates to the assurance case with new evidential statements indicating that the associated goal has been satisfied including the strategy used and links to context and associated evidence.

\subsection{Compositional Analysis}

The Assume Guarantee Reasoning Environment (AGREE) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree}. 
AGREE attempts to prove properties about one layer of the architecture using properties allocated to subcomponents.
The composition is performed in terms of \emph{assumptions} and \emph{guarantees} that are provided for each component and from a \emph{contract} for the component.
Assumptions describe the expectations the component has on the environment while guarantees describe bounds on the behavior of the component.

The AGREE model checker verifies the \emph{consistency} of the composition of the component contracts.
\begin{compactenum}
\item Component interfaces -- The output guarantees of each component must be strong enough to
satisfy the input assumptions of downstream components. 
\item Correctness of implementations -- The input assumptions of a system along with the 
output guarantees of its \emph{sub}-components must be strong enough to satisfy its output guarantees.
\end{compactenum}
If the real-time schedule of the components for the deployed system is available, then AGREE is additionally able to generate an enhanced model that verifies the composition consistent under the schedule \cite{nfm:agree}.

\subsection{High Assurance Component Synthesis}

The correctness of the filter and monitors inserted by the BriefCASE transformations means that each such component must meet its AGREE contract.
This obligation is addressed by \emph{formal synthesis} using the Semantic Properties of Language
and Automata Theory (SPLAT) tool and is the topic of this paper.
SPLAT generates code to implement the AGREE contract and then proves that its implementation exactly preserves the meaning of the contract all the way down to the binary for the target platform \cite{case-models-2021}.

\subsection{Remote Attestation}

Semantic remote attestation is a technique for establishing trust in software running on a non-local computer.  
An appraiser requests an attestation from a target, receives evidence in response to the request, and appraises
the evidence to determine trust~\cite{Coker::Principles-of-R}.
Because remote attestation does not require modification of its measurement target, we utilize it to establish trust in legacy software that cannot otherwise be verified.
We construct a verified remote attestation infrastructure around the legacy target that generates run-time and boot-time evidence.

\subsection{Information Flow Analysis}

As systems become more complex, it is essential to have trustworthy methods to provide a common understanding of dependencies in the system and the respective responsibilities of the developers who may be from separate organizations.  
%
The Awas~\cite{awas} AADL information flow analyzer and visualizer has been applied to enable developers and auditors to understand, reason about, explore, and visualize system dependencies and information flows at scale across components and subsystems.
Awas processes the AADL system architecture model, specifically its inter-component connection descriptions and intra-component flow specifications, to provide formal system-wide impact and flow analyses.
Such flows include component data/control flows, security-oriented information flows, and fault/error propagation specified using the AADL Error Modeling Annex (EMv2).
Awas also provides a user-friendly Domain Specific Language (DSL) to query, check, and visualize custom safety or security system properties.

\subsection{Secure Microkernel}

The seL4 microkernel~\cite{sel4-sosp09} is a lightweight, fast, and secure operating system kernel.
Its implementation is fully formally verified, from high-level security properties down to the binary level.
It was the first OS kernel with this degree of formal verification, and after more than a decade of further research and engineering is still not only the leading formally verified OS kernel, but also the fastest OS kernel on the Arm architecture \cite{sel4-tocs14}.

The true power of seL4 lies in its ability to scale formal analysis and verification to the much larger code bases that make up entire systems.
It does so by providing strong isolation between user-level components~\cite{sel4-cacm18}.
This isolation means that components can be analyzed separately from each other and be composed safely --- in this way seL4 provides the foundation that the soundness of the highly automated analysis tools such as AGREE depend on.
It makes it possible to run entire untrusted virtual machines and securely monitor their behavior on the
same hardware.
It makes it possible to provide filter and monitor components and prove that these components cannot be tampered with by the components they protect.
And it makes it possible to guarantee that the limited communication channels that the analysis tools assume to exist in the AADL model are the \emph{only} communication channels that are available to the
components in the system.
The combination of these enables automated high-level analysis with high assurance.

\subsection{Real-time Scheduling}

In mixed-criticality mission systems execution threads often have strict confidentiality, integrity, and availability requirements. 
\emph{Temporal isolation} is a secure technique to restrict timing channels and reduce temporal interference between software threads executing on the same platform. 
BriefCASE achieves temporal isolation through static cyclic scheduling using the sel4 domain scheduler.

BriefCASE generates a start-of-frame synchronization signal for each thread using a special thread called the \emph{Pacer}.
The Pacer sends periodic signals to each component thread.
Each component thread blocks until it receives its signal from the Pacer.
The thread then runs to completion and blocks again on the Pacer signal for the next iteration. 
Each thread subsequently executes exactly once during its statically scheduled time slice.

The new sel4 Mixed-Criticality Systems (MCS) variant provides additional capability that can support temporally isolated real-time systems. 
BriefCASE is also able to generate a static cyclic scheduler for MCS.
The MCS scheduler includes a start-of-frame signal so it eliminates the need for the Pacer component.
It also includes kernel level support for flexible dynamic scheduling.

\subsection{Infrastructure Code Generation}

The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) is a multi-platform, multi-language AADL code generation framework \cite{hamr}. 
Using seL4 as a foundation, HAMR enables AADL to be used as a model-based development and systems engineering framework for seL4-based applications.
Its primary objectives is to support system builds that leverage seL4 separation and information flow guarantees to achieve the AADL-specified component isolation and inter-component communication needed for cyber-resiliency.
HAMR ensures that seL4 is configured to permit the exact inter-component information flows analyzed and visualized by Awas at the model level.

For each AADL thread component, HAMR generates a thread code skeleton and application programming interfaces (APIs) for communicating over the ports declared on the component.
For components that are implemented manually, the developer fills out the thread skeleton with application code.
HAMR supports coding component application logic in either C, Slang~\cite{slang} (a high-assurance subset of Scala that can be translated to C), or CakeML.

HAMR generates component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, and port-based communication.
For port communication, shared memory communication (AADL data ports), buffered messaging (AADL event data ports), and buffered notification (AADL event ports) are supported.
HAMR code generation is staged using a translation architecture that facilitates adding new backends for different target platforms.

The seL4 deployment uses the component architecture for microkernel-based embedded systems (CAmkES) code-generation framework to configure the microkernel.
The HAMR generated CAmkES directly encodes the AADL model's component and communication topology and includes the AADL run-time infrastructure with its thread scheduling.
HAMR leverages the existing seL4 domain scheduler to enforce time partitioning and provide static cyclic scheduling.

HAMR also supports Linux-based virtual machine components in the seL4 deployment and the ability to run the entire system on the QEMU emulator.
HAMR automatically configures virtual machine based components, and this feature is used to sandbox the untrusted legacy code for the Mission Planner in the example UAV system.
The QEMU emulator support facilitates rapid prototyping for test, debug, and analysis, and it enables automated regression testing.

As part of its code generation process, HAMR produces flow graphs reflecting the inter-component information flow at both the AADL architecture level and the CAmkES level for the seL4 deployment.  
Visual representations are provided for manual inspection, and SMT-based representations are generated for formal reasoning.
The SMT-based representations are used to prove the following properties:
\begin{compactenum}
\item All AADL modeled flows are in the CAmkES configuration. 
\item No extraneous flows have been added to the CAmkES configuration.
\end{compactenum}
These properties focus on cyber-resiliency, but other key semantic properties can be verified as well.

\subsection{Assurance Case}

Assurance activities for high-integrity systems in the aerospace domain are currently driven by industry and government standards such as DO-178C and MIL-HDBK-516C.
The use of assurance cases to show conformance to standards (or provide alternate means of assurance) is being pursued in separate research programs.  
In addition, we have found that in assuring the cyber-resiliency properties of aircraft designs we need to integrate different kinds of evidence with varying levels of formality.
This has been our motivation for incorporating assurance case methods in BriefCASE.

In previous work, we developed the {\em Resolute} language and tool~\cite{resolute-destion} as a way to help engineers create an assurance argument describing the steps taken during the design process to make the system safe and secure.
The Resolute syntax supports construction of assurance cases that comply with the Goal Structuring
Notation (GSN) v2 standard.
Claims are expressed as \textit{goals} and \textit{strategies}, and can contain attributes such as \textit{context}, \textit{assumptions}, and \textit{justification}.
Claims can be marked \textit{undeveloped}, which Resolute interprets as an unsupported claim or with a \textit{solution}, which is an explicit assertion that the claim is supported.
Rather than being a separate document, a Resolute assurance case is part of the architecture model and can refer to elements within the model.
Since it is not a static representation, it can ensure that the assurance argument remains consistent with the evolving
design.

BriefCASE includes a library of Resolute assurance strategies, or \emph{patterns}.
The patterns are instantiated with context from the AADL model and specify the evidence required to support the cyber-resiliency goals of the system.
For example, the \texttt{add\_filter} strategy is automatically inserted into the assurance case when the
\textit{Filter} transformation is performed, and it includes logical rules that Resolute uses to
determine whether its claim is supported by evidence.
\begin{compactitem} 
\item Component property implemented -- The filter has been implemented correctly 
to meet its AGREE specification as shown by the proof produced by SPLAT.
\item Filter is connected -- The filter component is still present at the correct location in the model and has not been altered or deleted by subsequent design changes. 
\item Filter cannot be bypassed -- There is no alternate information flow in the model that 
would allow the filter to be bypassed and therefore not perform its function.   
\item AGREE properties are valid -- The filter specification has been verified by AGREE to meet its
intended purpose in the system.   
\end{compactitem}

### 2022-09-12 cover-letter/letter.tex
Revise the overview

@@
     \item A new section that formally defines a code contract that is a specialized specification for high-assurance components, such as filters and monitors, that are sufficient for proving cyber-properties of the system and for synthesis to an actual implementation.
-    \item A heavily revised section on the synthesis process that starts with a code contract and ends with CakeML code describing is some considerable detail how the synthesis works and is proved correct.
-    \item A new real-world case study with system developers that are not expert with assume-guarantee reasoning to show feasibility with such developers. 
+    \item A heavily revised section on the synthesis process that starts with a code contract and ends with CakeML code describing in some considerable detail how the synthesis works and is proved correct.
 \end{itemize} 

### 2022-09-12 paper/example.tex
Revise the overview

@@
+\begin{figure*}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[scale=0.4]{flowchart.png} \\
+    \end{tabular}
+  \end{center}
+\caption{AGREE failure certificate for initial design.}
+\label{fig:example-certificate}
+\end{figure*}
+
 \begin{figure*}[h]
@@
 \end{figure}
+
+\egm{
+  Merge the AGREE overview here into this section.
+  Condense it down to just the opening two paragraphes that currently exist in the overview section.
+  Walk through the first part of the example up to where AGREE fails because of Cyber Requirements.
+  Introduce the new figure showing its workflow.
+  Finish the example walking through the new figure. 
+  Add the tests to occur before final AGREE check.
+  Send to SPLAT to get verified CakeML code. 
+}
 

### 2022-09-12 paper/high-assurance-contracts.tex
Revise the overview

@@
 \subsubsection*{Imperative code contracts.}
+\egm{Remove.
+Replace with the informal presentation of temporal squashing \emph{or} remove entirely.}
+
 Stream-based evaluation allows values arbitrarily `deep in the past'

### 2022-09-12 paper/intro.tex
Revise the overview

@@
+\egm{Rewrite to make clear the contribution here: code contracts, tests, and synthesis through SPLAT. 
+Present the \emph{unverified} path as is done in the introduction.
+See if it is possible to present the verified path (ping Konrad) to state the various correctness guarantees.}
+
 In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
@@
 The approach is illustrated by a simple example in
-\secref{sec:example}. AGREE verification and specification in the AGREE
-language are presented in \secref{sec:agree} and \secref{agree-semantics}.
+\secref{sec:example}. AGREE verification and specification are
+presented in \secref{sec:agree} and \secref{agree-semantics}.
 Code contracts are defined in \secref{sec:code-contracts} with a method

### 2022-09-12 paper/overview.tex
Revise the overview

@@
-% Copied from DESTION paper
-The BriefCASE environment provides systems engineers with a workflow and tool support for developing
-products with inherent cyber-resiliency.
-
-The  workflow starts with the development of a baseline AADL model of the system architecture
-focusing on the desired functionality. This model can be analyzed using any of the existing AADL 
-tools (e.g., resource usage, information flow, latency) to determine whether it is acceptable.
-BriefCASE integrates additional tools that analyze the architecture model for cybersecurity vulnerabilities and
-generate requirements that, when addressed, will mitigate those vulnerabilities.
-These requirements are imported into the model and may be addressed using a 
-collection of automated model transforms. As requirements are addressed in the design, an assurance case is updated with
-corresponding evidence, computed directly from the model or by supporting analysis tools.  
-Code implementing new high-assurance components as well as communication and execution infrastructure
-is generated from the model along with associated assurance evidence.
-
-The BriefCASE tools and their interactions are shown in \figref{fig:tool-arch}.   
-The following sections describe each step of the workflow in more detail, corresponding
-to the tools and artifacts shown in green in the figure.
-
 \begin{figure*}
@@
   	\end{center}
-	\caption{BriefCASE tool architecture.  Tools and code/artifacts discussed in the BriefCASE workflow are shown in green.} 
+	\caption{\brfcs\ integrated tool architecture with key components in the general workflow shown in green.} 
 	\label{fig:tool-arch} 
@@
 
-\subsection{Requirements Analysis}
+The BriefCASE OSATE environment, \figref{fig:tool-arch}, provides designers with workflow and tool support for developing products with inherent cyber-resiliency.
+It starts with the development of a baseline AADL model of the system architecture that defines the desired functionality.
+This baseline model can be evaluated using existing tools such as resource usage, information flow, latency analysis, etc. for validation.
 
-BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and
-DCRYPPS~\cite{dcrypps2019}) that examine AADL models, report potential cyber vulnerabilities,
-and suggest requirements for mitigation.
-%
-A BriefCASE project contains a repository for requirements. Imported requirements are represented 
-as assurance case goals to be satisfied. 
-The goal is marked {\em undeveloped} initially.
-Evidential statements are added to the goal as the design is updated to address this requirement.
-Some requirements are formalized as assume-guarantee contracts added to the AADL model.
-Formal verification provides the evidence for the satisfaction of these requirements.
+\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline AADL model (top of \figref{fig:tool-arch}).
+\brfcs\ provides access to analysis tools that examine AADL models, report potential cyber vulnerabilities,
+and suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
+Designers add these requirements to the \brfcs\ project requirements repository. 
+A key part of the requirements management in \brfcs\ is a corresponding assurance case that provides evidence, computed directly from the model or by supporting analysis tools, showing how each requirement is satisfied.
+Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project.
 
-\subsection{Cyber Transforms}
+\brfcs\ provides an extendable library of model transformations to address requirements for common cyber vulnerabilities:
+\begin{compactitem}
+	\item \textbf{Filter}: block messages that do not conform to a given specification
+	\item \textbf{Monitor}: detect and alert violations at runtime
+	\item \textbf{Gate}: block messages in the presence of an alert
+	\item \textbf{Attestation}: measure non-local software to assess trustworthiness
+	\item \textbf{Virtualization}: isolate components with virtualization
+	\item \textbf{Proxy}: enable HTTPS message inspection 
+	\item \textbf{seL4}: transform component definitions for the seL4 secure micro kernel
+\end{compactitem}
+Transformations update the AADL model architecture with new components and the assurance case with new evidential statements for the associated goal.
+Filters, monitors, and gates are considered high assurance components that require formal reasoning in building the assurance case.
+This manuscript discusses how these high assurance components are specified and synthesized with the requisite evidence for the assurance case.
 
-To address the new cyber-resiliency requirement, the architecture will need to be transformed in such a way as to harden the design against the vulnerability.
-BriefCASE provides an extendable library of model transformations for addressing common cyber vulnerabilities.  Currently, the following transformations are supported:
+When \brfcs\ inserts a filter, monitor, or gate, high assurance component, it requires evidence that the component is 
+\begin{compactenum}
+	\item connected in the model, 
+	\item cannot be bypassed, 
+	\item meets its intended purpose, and
+	\item is properly implemented.
+\end{compactenum}
+Information flow analysis provides evidence for the first two properties while synthesis provide evidence for the last two properties.
 
-\begin{compactitem}
-	\item Filter -- Blocks messages that do not conform to the given specification
-	\item Monitor -- Detects violations of a given run-time condition and generates an alert
-	\item Switch -- Used with a Monitor to block messages when an alert is generated (also referred to as a \textit{gate})
-	\item Attestation -- Performs a measurement on non-local software to assess its trustworthiness
-	\item Virtualization -- Isolates software component(s) in a virtual machine
-	\item Proxy -- Inserts a pair of components to allow inspection of %https
-		HTTPS message payloads
-	\item seL4 -- Transforms the model to comply with seL4 component properties
-\end{compactitem}
+The Awas~\cite{awas} AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model.
+The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
+Along with the code generation, HAMR constructs a proof that all AADL modeled flows are in the implementation and that no extraneous flows have been added to the implementation.
+HAMR relies on the seL4 microkernel for its proofs.
+The seL4 microkernel, among other security properties, provides strong isolation between components ensuring that there are no communication channels outside of what is defined \cite{sel4-sosp09, sel4-tocs14, sel4-cacm18}.
 
-The transformations are automated by the BriefCASE tool to create a hardened model that is correct-by-construction.  
-These transformations include updates to the assurance case with new evidential statements indicating that the associated goal has been satisfied including the strategy used and links to context and associated evidence.
+This manuscript discusses the aspects of \brfcs\ concerning evidence that the component meets its intended purpose and is properly implemented.
+The Assume Guarantee Reasoning Environment (AGREE) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
+AGREE attempts to prove properties about one layer of the architecture using properties allocated to subcomponents.
+The composition is performed in terms of \emph{contracts} provided for each component that state the component's \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
+AGREE verifies the \emph{consistency} of the composition of the component contracts: output guarantees must be strong enough to satisfy input assumptions of downstream components and input assumptions of a system along with the output guarantees of its \emph{sub}-components must be strong enough to satisfy its output guarantees.
 
-\subsection{Compositional Analysis}
-
-The Assume Guarantee Reasoning Environment (AGREE) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree}. 
-AGREE attempts to prove properties about one layer of the architecture using properties allocated to subcomponents.
-The composition is performed in terms of \emph{assumptions} and \emph{guarantees} that are provided for each component and from a \emph{contract} for the component.
-Assumptions describe the expectations the component has on the environment while guarantees describe bounds on the behavior of the component.
-
-The AGREE model checker verifies the \emph{consistency} of the composition of the component contracts.
-\begin{compactenum}
-\item Component interfaces -- The output guarantees of each component must be strong enough to
-satisfy the input assumptions of downstream components. 
-\item Correctness of implementations -- The input assumptions of a system along with the 
-output guarantees of its \emph{sub}-components must be strong enough to satisfy its output guarantees.
-\end{compactenum}
-If the real-time schedule of the components for the deployed system is available, then AGREE is additionally able to generate an enhanced model that verifies the composition consistent under the schedule \cite{nfm:agree}.
-
-\subsection{High Assurance Component Synthesis}
-
-The correctness of the filter and monitors inserted by the BriefCASE transformations means that each such component must meet its AGREE contract.
-This obligation is addressed by \emph{formal synthesis} using the Semantic Properties of Language
-and Automata Theory (SPLAT) tool and is the topic of this paper.
-SPLAT generates code to implement the AGREE contract and then proves that its implementation exactly preserves the meaning of the contract all the way down to the binary for the target platform \cite{case-models-2021}.
-
-\subsection{Remote Attestation}
-
-Semantic remote attestation is a technique for establishing trust in software running on a non-local computer.  
-An appraiser requests an attestation from a target, receives evidence in response to the request, and appraises
-the evidence to determine trust~\cite{Coker::Principles-of-R}.
-Because remote attestation does not require modification of its measurement target, we utilize it to establish trust in legacy software that cannot otherwise be verified.
-We construct a verified remote attestation infrastructure around the legacy target that generates run-time and boot-time evidence.
-
-\subsection{Information Flow Analysis}
-
-As systems become more complex, it is essential to have trustworthy methods to provide a common understanding of dependencies in the system and the respective responsibilities of the developers who may be from separate organizations.
-%
-The Awas~\cite{awas} AADL information flow analyzer and visualizer has been applied to enable developers and auditors to understand, reason about, explore, and visualize system dependencies and information flows at scale across components and subsystems.
-Awas processes the AADL system architecture model, specifically its inter-component connection descriptions and intra-component flow specifications, to provide formal system-wide impact and flow analyses.
-Such flows include component data/control flows, security-oriented information flows, and fault/error propagation specified using the AADL Error Modeling Annex (EMv2).
-Awas also provides a user-friendly Domain Specific Language (DSL) to query, check, and visualize custom safety or security system properties.
-
-\subsection{Secure Microkernel}
-
-The seL4 microkernel~\cite{sel4-sosp09} is a lightweight, fast, and secure operating system kernel.
-Its implementation is fully formally verified, from high-level security properties down to the binary level.
-It was the first OS kernel with this degree of formal verification, and after more than a decade of further research and engineering is still not only the leading formally verified OS kernel, but also the fastest OS kernel on the Arm architecture \cite{sel4-tocs14}.
-
-The true power of seL4 lies in its ability to scale formal analysis and verification to the much larger code bases that make up entire systems.
-It does so by providing strong isolation between user-level components~\cite{sel4-cacm18}.
-This isolation means that components can be analyzed separately from each other and be composed safely --- in this way seL4 provides the foundation that the soundness of the highly automated analysis tools such as AGREE depend on.
-It makes it possible to run entire untrusted virtual machines and securely monitor their behavior on the
-same hardware.
-It makes it possible to provide filter and monitor components and prove that these components cannot be tampered with by the components they protect.
-And it makes it possible to guarantee that the limited communication channels that the analysis tools assume to exist in the AADL model are the \emph{only} communication channels that are available to the
-components in the system.
-The combination of these enables automated high-level analysis with high assurance.
-
-\subsection{Real-time Scheduling}
-
-In mixed-criticality mission systems execution threads often have strict confidentiality, integrity, and availability requirements. 
-\emph{Temporal isolation} is a secure technique to restrict timing channels and reduce temporal interference between software threads executing on the same platform. 
-BriefCASE achieves temporal isolation through static cyclic scheduling using the sel4 domain scheduler.
-
-BriefCASE generates a start-of-frame synchronization signal for each thread using a special thread called the \emph{Pacer}.
-The Pacer sends periodic signals to each component thread.
-Each component thread blocks until it receives its signal from the Pacer.
-The thread then runs to completion and blocks again on the Pacer signal for the next iteration. 
-Each thread subsequently executes exactly once during its statically scheduled time slice.
-
-The new sel4 Mixed-Criticality Systems (MCS) variant provides additional capability that can support temporally isolated real-time systems. 
-BriefCASE is also able to generate a static cyclic scheduler for MCS.
-The MCS scheduler includes a start-of-frame signal so it eliminates the need for the Pacer component.
-It also includes kernel level support for flexible dynamic scheduling.
-
-\subsection{Infrastructure Code Generation}
-
-The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) is a multi-platform, multi-language AADL code generation framework \cite{hamr}. 
-Using seL4 as a foundation, HAMR enables AADL to be used as a model-based development and systems engineering framework for seL4-based applications.
-Its primary objectives is to support system builds that leverage seL4 separation and information flow guarantees to achieve the AADL-specified component isolation and inter-component communication needed for cyber-resiliency.
-HAMR ensures that seL4 is configured to permit the exact inter-component information flows analyzed and visualized by Awas at the model level.
-
-For each AADL thread component, HAMR generates a thread code skeleton and application programming interfaces (APIs) for communicating over the ports declared on the component.
-For components that are implemented manually, the developer fills out the thread skeleton with application code.
-HAMR supports coding component application logic in either C, Slang~\cite{slang} (a high-assurance subset of Scala that can be translated to C), or CakeML.
-
-HAMR generates component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, and port-based communication.
-For port communication, shared memory communication (AADL data ports), buffered messaging (AADL event data ports), and buffered notification (AADL event ports) are supported.
-HAMR code generation is staged using a translation architecture that facilitates adding new backends for different target platforms.
-
-The seL4 deployment uses the component architecture for microkernel-based embedded systems (CAmkES) code-generation framework to configure the microkernel.
-The HAMR generated CAmkES directly encodes the AADL model's component and communication topology and includes the AADL run-time infrastructure with its thread scheduling.
-HAMR leverages the existing seL4 domain scheduler to enforce time partitioning and provide static cyclic scheduling.
-
-HAMR also supports Linux-based virtual machine components in the seL4 deployment and the ability to run the entire system on the QEMU emulator.
-HAMR automatically configures virtual machine based components, and this feature is used to sandbox the untrusted legacy code for the Mission Planner in the example UAV system.
-The QEMU emulator support facilitates rapid prototyping for test, debug, and analysis, and it enables automated regression testing.
-
-As part of its code generation process, HAMR produces flow graphs reflecting the inter-component information flow at both the AADL architecture level and the CAmkES level for the seL4 deployment.  
-Visual representations are provided for manual inspection, and SMT-based representations are generated for formal reasoning.
-The SMT-based representations are used to prove the following properties:
-\begin{compactenum}
-\item All AADL modeled flows are in the CAmkES configuration. 
-\item No extraneous flows have been added to the CAmkES configuration.
-\end{compactenum}
-These properties focus on cyber-resiliency, but other key semantic properties can be verified as well.
-
-\subsection{Assurance Case}
-
-Assurance activities for high-integrity systems in the aerospace domain are currently driven by industry and government standards such as DO-178C and MIL-HDBK-516C.
-The use of assurance cases to show conformance to standards (or provide alternate means of assurance) is being pursued in separate research programs.  
-In addition, we have found that in assuring the cyber-resiliency properties of aircraft designs we need to integrate different kinds of evidence with varying levels of formality.
-This has been our motivation for incorporating assurance case methods in BriefCASE.
-
-In previous work, we developed the {\em Resolute} language and tool~\cite{resolute-destion} as a way to help engineers create an assurance argument describing the steps taken during the design process to make the system safe and secure.
-The Resolute syntax supports construction of assurance cases that comply with the Goal Structuring
-Notation (GSN) v2 standard.
-Claims are expressed as \textit{goals} and \textit{strategies}, and can contain attributes such as \textit{context}, \textit{assumptions}, and \textit{justification}.
-Claims can be marked \textit{undeveloped}, which Resolute interprets as an unsupported claim or with a \textit{solution}, which is an explicit assertion that the claim is supported.
-Rather than being a separate document, a Resolute assurance case is part of the architecture model and can refer to elements within the model.
-Since it is not a static representation, it can ensure that the assurance argument remains consistent with the evolving
-design.
-
-BriefCASE includes a library of Resolute assurance strategies, or \emph{patterns}.
-The patterns are instantiated with context from the AADL model and specify the evidence required to support the cyber-resiliency goals of the system.
-For example, the \texttt{add\_filter} strategy is automatically inserted into the assurance case when the
-\textit{Filter} transformation is performed, and it includes logical rules that Resolute uses to
-determine whether its claim is supported by evidence.
-\begin{compactitem} 
-\item Component property implemented -- The filter has been implemented correctly 
-to meet its AGREE specification as shown by the proof produced by SPLAT.
-\item Filter is connected -- The filter component is still present at the correct location in the model and has not been altered or deleted by subsequent design changes. 
-\item Filter cannot be bypassed -- There is no alternate information flow in the model that 
-would allow the filter to be bypassed and therefore not perform its function.   
-\item AGREE properties are valid -- The filter specification has been verified by AGREE to meet its
-intended purpose in the system.   
-\end{compactitem}
+The behavior of a high assurance component is specified by the designer in a \emph{code contract}.
+A code contract is a subset of the AGREE language with formally defined semantics needed for code synthesis as apposed to only compositional reasoning.
+AGREE proves that the code contract meets its intended purpose in the system composition by satisfying the associated cyber-requirements.
+Then the Semantic Properties of Language and Automata Theory (SPLAT) tool generates \emph{CakeML} code to implement the code contract.
+The CakeML compiler is verified to generate binary implementations that are equivalent in semantics to the original input program \cite{cakeml}.
+It proves that the resulting binary exactly preserves the meaning of the original contract to assure that the component is properly implemented.

### 2022-09-13 paper/overview.tex
Add figs and revise overview

@@
 \end{compactitem}
-Transformations update the AADL model architecture with new components and the assurance case with new evidential statements for the associated goal.
+Transformations update the AADL model architecture with new components and the assurance case with new required evidential statements.
+
 Filters, monitors, and gates are considered high assurance components that require formal reasoning in building the assurance case.
-This manuscript discusses how these high assurance components are specified and synthesized with the requisite evidence for the assurance case.
-
-When \brfcs\ inserts a filter, monitor, or gate, high assurance component, it requires evidence that the component is 
+When added to a system, \brfcs\ requires evidence that the component is 
 \begin{compactenum}
@@
 \end{compactenum}
-Information flow analysis provides evidence for the first two properties while synthesis provide evidence for the last two properties.
+Information flow analysis provides evidence for the first two properties while synthesis provides evidence for the last two properties.
 
@@
 
-This manuscript discusses the aspects of \brfcs\ concerning evidence that the component meets its intended purpose and is properly implemented.
 The Assume Guarantee Reasoning Environment (AGREE) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
-AGREE attempts to prove properties about one layer of the architecture using properties allocated to subcomponents.
-The composition is performed in terms of \emph{contracts} provided for each component that state the component's \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
-AGREE verifies the \emph{consistency} of the composition of the component contracts: output guarantees must be strong enough to satisfy input assumptions of downstream components and input assumptions of a system along with the output guarantees of its \emph{sub}-components must be strong enough to satisfy its output guarantees.
+AGREE attempts to prove properties about one layer of the architecture using properties allocated to sub-components.
+The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
+AGREE verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy destination constraints.
 
-The behavior of a high assurance component is specified by the designer in a \emph{code contract}.
-A code contract is a subset of the AGREE language with formally defined semantics needed for code synthesis as apposed to only compositional reasoning.
-AGREE proves that the code contract meets its intended purpose in the system composition by satisfying the associated cyber-requirements.
-Then the Semantic Properties of Language and Automata Theory (SPLAT) tool generates \emph{CakeML} code to implement the code contract.
+The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
+Regarding intended purpose, this manuscript defines a \emph{code contract} language that is a subset of the AGREE language with sequential evaluation semantics as opposed to the existing parallel semantics of AGREE.
+The sequential semantics are requisite to code synthesis and are defined such that AGREE is able to still prove consistency in the composition. 
+As such, the behavior of a high assurance component is defined by the designer in a code contract, and AGREE proves that the code contract meets its intended purpose in the system composition by satisfying the associated cyber-requirements.
+
+Showing that a componest is properly implemented is accomplished through formal synthesis from the code contract.
+The Semantic Properties of Language and Automata Theory (SPLAT) tool generates \emph{CakeML} code to implement the code contract.
 The CakeML compiler is verified to generate binary implementations that are equivalent in semantics to the original input program \cite{cakeml}.

### 2022-09-13 paper/overview.tex
Udpate bib entries

@@
 \brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline AADL model (top of \figref{fig:tool-arch}).
-\brfcs\ provides access to analysis tools that examine AADL models, report potential cyber vulnerabilities,
+\brfcs\ provides access to analysis tools that examine the model, report potential cyber vulnerabilities,
 and suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
@@
 
-Showing that a componest is properly implemented is accomplished through formal synthesis from the code contract.
+Showing that a component is properly implemented is accomplished through formal synthesis from the code contract.
 The Semantic Properties of Language and Automata Theory (SPLAT) tool generates \emph{CakeML} code to implement the code contract.

### 2022-09-13 paper/overview.tex
Revise overview

@@
 It starts with the development of a baseline AADL model of the system architecture that defines the desired functionality.
-This baseline model can be evaluated using existing tools such as resource usage, information flow, latency analysis, etc. for validation.
+This model can be evaluated using existing tools such as resource usage, information flow, latency analysis, etc. for validation.
 
@@
 Filters, monitors, and gates are considered high assurance components that require formal reasoning in building the assurance case.
-When added to a system, \brfcs\ requires evidence that the component is 
+When added to a system, \brfcs\ requires evidence that such components are 
 \begin{compactenum}
@@
 	\item cannot be bypassed, 
-	\item meets its intended purpose, and
-	\item is properly implemented.
+	\item meet their intended purposes, and
+	\item are properly implemented.
 \end{compactenum}
-Information flow analysis provides evidence for the first two properties while synthesis provides evidence for the last two properties.
+Information flow analysis provides evidence for the first two properties while model checking and code synthesis provide evidence for the last two properties.
 
-The Awas~\cite{awas} AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model.
-The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
-Along with the code generation, HAMR constructs a proof that all AADL modeled flows are in the implementation and that no extraneous flows have been added to the implementation.
+The Awas AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model \cite{awas}.
+The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
+Along with that code generation, HAMR constructs a proof that all flows in the AADL model are in the implementation and that no extraneous flows have been added to the implementation.
 HAMR relies on the seL4 microkernel for its proofs.
@@
 The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
-AGREE verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy destination constraints.
+AGREE verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.
 
 The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
-Regarding intended purpose, this manuscript defines a \emph{code contract} language that is a subset of the AGREE language with sequential evaluation semantics as opposed to the existing parallel semantics of AGREE.
-The sequential semantics are requisite to code synthesis and are defined such that AGREE is able to still prove consistency in the composition. 
-As such, the behavior of a high assurance component is defined by the designer in a code contract, and AGREE proves that the code contract meets its intended purpose in the system composition by satisfying the associated cyber-requirements.
+This manuscript defines a \emph{code contract} language that is a subset of the AGREE language with sequential evaluation semantics as opposed to the existing parallel semantics of AGREE.
+The sequential semantics are requisite to code synthesis and are defined such that AGREE is able to still prove consistency in the contract composition of the model. 
+As such, the behavior of a high assurance component is defined by the designer in a code contract, and AGREE proves that the code contract meets its intended purpose in two ways: first, its input to output definition is satisfied, and second, in the system composition, it satisfies associated cyber-requirements.
 

### 2022-09-13 paper/example.tex
Revise example to match overview and flow

@@
 \caption{AGREE failure certificate for initial design.}
-\label{fig:example-certificate}
+\label{fig:flowchart}
 \end{figure*}
@@
 
-We now describe a simple software system, named SW, for route planning
+\figref{fig:flowchart} illustrates a simplified \brfcs\ workflow with the shaded boxes representing the contributions discussed in this manuscript: test contracts, code contracts, and SPLAT synthesis.
+The unshaded boxes are the existing tools that generate evidence for the assurance case.
+The rest of this section follows an illustrative example through the workflow.
+
+The baseline AADL model is a simple software system, named SW, for route planning
 and automated control of a UAV. A picture of the AADL architectural
@@
 The expected behavior of the SW system, and the components
-implementing the system, is modeled with AGREE contract
-specifications.  The contracts constrain input with assumptions and
-state properties of output with guarantees.  AGREE performs model
-checking on this assume-guarantee system to hierarchically prove that
-the SW component obeys all contract obligations under all possible
-finite input streams.  The AGREE analysis, contract language, and
-contracts for SW are discussed in greater detail in
-\secref{sec:agree} and \secref{sec:agree-semantics}.
-
-The AGREE contracts for the initial system in \figref{fig:example},
-and its implementing components, both assume and guarantee the absence
+implementing the system, are modeled with AGREE contract
+specifications.  
+These contracts assume and guarantee the absence
 of any malicious, or unspecified, component behavior.  More precisely,
@@
 component input assumptions and system output guarantees.
+AGREE background and the baseline contracts are discussed in detail in \secref{sec:agree} and \secref{sec:agree-semantics}.
 
-\subsection{Detecting and addressing vulnerabilities}
-A cyber-vulnerability analysis, however, identifies the potential of a
+\subsection{Adding cyber requirements}
+
+A cyber-vulnerability analysis identifies the potential of a
 supply chain attack through the AI route planner (provided by a
-third-party vendor without source code).  BriefCASE marks the
-component as \emph{untrusted} after the analysis, and the contract for
-AI is modified by removing all guarantees on its outputs to reflect
-its untrusted status.  The AI output is now unconstrained in the
+third-party vendor without source code).
+Based on the analysis, the designer adds a requirement that the system must guard agaist malicious AI behavior, marks the component as \emph{untrusted}, and removes all output guarantees from its contract.
+The AI output is now unconstrained in the
 assume-guarantee reasoning of AGREE and is able to generate any value
@@
 guarantees that do not hold, and each failure comes with a
-corresponding counter-example.  The results are not unexpected given
-the AI route planner's unconstrained behavior.
+corresponding counter-example.
 
-For example, the first violation is that the mission command on \texttt{AutomationResponse} from
+The first violation is that the mission command on \texttt{AutomationResponse} from
 the AI to the WM is no longer guaranteed to be well-formed.  The
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{hardened-certificate.png}
+      \includegraphics[scale=0.38]{agree-test-output.png} \\
+      (a) \\ \\
+      \includegraphics[scale=0.4]{hardened-certificate.png} \\
+      (b)    
     \end{tabular}
   \end{center}
-  \caption{AGREE verification certificate for cyber-hardened design.}
+  \caption{AGREE verification certificates. (a) Test contract verification results. (b) Cyber-hardened design verification results.}
   \label{fig:hardened-certificate}
@@
 
-The system designer now uses BriefCASE to cyber-harden SW by inserting
+\emph{Adding high assurance components}
+
+The system designer now uses \brfcs\ to cyber-harden SW by inserting
 high-assurance components in the form of a filter and a monitor, as
-shown in \figref{fig:hardened}, to isolate the AI component and prevent it from violating the WM component input assumptions.
+shown in \figref{fig:hardened}.
+These isolate the AI component and prevent it from violating the WM component input assumptions.
 A filter enforces an invariant over
-each datum in the data stream by not forwarding input to its output if
-that input violates the filter invariant.  The auto-generated AGREE
-code contract specification states that only well-formed inputs are passed to the
-output.  The system developer must further define the filtering
-policy, but it is usually based on the existing assumptions made by
+each datum in the data stream by not forwarding input to its output if the invariant does not hold.
+\brfcs\ generates the initial code contract for the filter which the designer completes by writing the invariant property.
+That property it is usually based on the existing assumptions made by
 downstream components that consume the filter output.
@@
 to reason about the temporal, and other invariant, properties of that input.  A monitor raises
-an alert if the specified properties are ever violated.  The
-AGREE code contract for the monitor in our example states that an
+an alert if the specified properties are ever violated.  
+As with the filter, \brfcs\ generates a template code contract that is completed by the designer with the desired behavior that is typically defined by the SW system and its implementing component contracts.
+The code contract in our example states that an
 automation response can only be generated in conjunction with an
 automation request; and further, that response must come with the
-request or in the next step after the request.  As with the filter,
-the system designer provides the policy, and that policy is typically
-based on existing the assumptions and guarantees of the SW system and its implementing components.
-AGREE code contracts are discussed in detail in \secref{sec:code-contracts}.
+request or in the next step after the request.  
+Code contracts are discussed in detail in \secref{sec:code-contracts}.
+
+Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
+Test contracts allow the designer to validate the behavior, and general input to output properties, of a code contract through unit testing.
+For example, one of the test contracts for the monitor is that it should not alert and pass its input when the monitor sees the \texttt{AutomationReponse} one step after it sees the \texttt{AutomationRequest}.
+Test contracts are discussed in detail in \secref{sec:test-contracts}.
 
 The AGREE analysis of the cyber-hardened implementation is shown in
-\figref{fig:hardened-certificate}. AGREE has generated high-level
-evidence justifying the claim that the high-assurance components
-guarantee the correct behavior of the SW implementation in the
-presence of the untrusted AI route planner.  Having passed AGREE
-verification, the high-assurance components are ready to be
+\figref{fig:hardened-certificate}.
+\figref{fig:hardened-certificate}(a) is the test contract results and \figref{fig:hardened-certificate}(b) is the hardened system composition.
+Here AGREE has generated high-level
+evidence justifying the claim that the high-assurance components meet their intended purposes.
+Having passed AGREE verification, the high-assurance components are ready to be
 synthesized.
 
-\subsection{Code synthesis}
+\subsection{Synthesizing code from code contracts}
+
 High-assurance components are automatically synthesized by SPLAT from
-the AGREE code contracts to equivalent programs in the CakeML
+the code contracts to equivalent programs in the CakeML
 language.
@@
 discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.
+Synthesis from code contracts is discussed in detail in \secref{sec:sythesis}.

### 2022-09-13 paper/overview.tex
Revise example to match overview and flow

@@
 The sequential semantics are requisite to code synthesis and are defined such that AGREE is able to still prove consistency in the contract composition of the model. 
-As such, the behavior of a high assurance component is defined by the designer in a code contract, and AGREE proves that the code contract meets its intended purpose in two ways: first, its input to output definition is satisfied, and second, in the system composition, it satisfies associated cyber-requirements.
+This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
+As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing AGREE model checker, test contracts prove the desired code contract behavior, and once the designer is satisfied that is correct, the system composition is verified to conclude that the contract meets its intended purpose.
+dressing associated cyber-requirements.
 

### 2022-09-13 paper/intro.tex
Revise intro to use macros

@@
 verification that the transformed system model meets its requirements
-is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
+is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
 Environment).
 %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
-AGREE is a {\em compositional assume-guarantee} style model checker
+\agr\ is a {\em compositional assume-guarantee} style model checker
 for AADL models that attempts to prove properties about one layer of
@@
 architecture of the model by adding in new components; they also
-generate a formal specification in the form of a \emph{code contract} in the AGREE language for their intended behavior. 
+generate a formal specification in the form of a \emph{code contract} in the \agr\ language for their intended behavior. 
 A code contract describes the computation of a high-assurance component and not just what it computes.
@@
 
-Another novel aspect of the approach is the synthesis of the AGREE
-code contracts for the high-assurance components to CakeML, a verified
+Another novel aspect of the approach is the synthesis of the \agr\ 
+code contracts for the high-assurance components to \ckml, a verified
 compiler implementation for the functional programming language
 ML \cite{cakeml}. This paper describes in detail the synthesis path
-from code contracts to CakeML code, providing a formal framework in
-which to argue correctness. CakeML then provides a verified
+from code contracts to \ckml\ code, providing a formal framework in
+which to argue correctness. \ckml\ then provides a verified
 compilation path to several different target binaries (and also
-proving that the meaning of the CakeML source code is exactly
+proving that the meaning of the \ckml\ source code is exactly
 preserved in the final binaries). Assuming that the execution schedule
@@
 model, and that the HAMR-generated communication fabric delivers
-messages between components as expected, the AGREE model checking
+messages between components as expected, the \agr\ model checking
 results then hold for the deployed system, \ie, the system will detect
@@
 The approach is illustrated by a simple example in
-\secref{sec:example}. AGREE verification and specification are
+\secref{sec:example}. \agr\ verification and specification are
 presented in \secref{sec:agree} and \secref{agree-semantics}.
 Code contracts are defined in \secref{sec:code-contracts} with a method
-for testing code contracts in the AGREE framework before synthesis given
+for testing code contracts in the \agr\ framework before synthesis given
 \secref{sec:testing}.
@@
 
-The BriefCASE tools are open source and publicly
+The \brfcs\ tools are open source and publicly
 available \cite{fmide}, as is the motivating example \cite{repo}, and
 the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
-Videos demonstrating the use of the BriefCASE tools to build the UAV
+Videos demonstrating the use of the \brfcs\ tools to build the UAV
 example presented in Section~\ref{sec:case-study} are also

### 2022-09-14 paper/paper.tex
Add BriefCASE overview for big picture
Revise the overview
Revise example to match overview and flow
Revise intro to use macros
Remove revision notes

@@
 
+% Comment macros
+\newcommand{\egm}[1]{\textcolor{orange}{egm: #1}}
+
 \newcommand{\konst}[1]{\ensuremath{\mathsf{#1}}}
@@
 \newcommand{\linesref}[2]{Lines~\ref{#1}--\ref{#2}}
+
+% Acronym macros
+\newcommand{\brfcs}{BriefCASE}
+\newcommand{\agr}{AGREE}
+\newcommand{\splt}{SPLAT}
+\newcommand{\ckml}{CakeML}
+
 % \newcommand{\eqref}[1]{Eq.~\ref{#1}}
@@
 
-\ifREVISIONS
-\subsection{SOSYM Journal Details}
-The due date is \textbf{March} $22^\mathrm{nd}$, \textbf{2022}.
-Go to \url{https://mc.manuscriptcentral.com/sosym/}.
-Create a new account (if you don't have one), login and go to the "Author"-section, select the link "Start New Submission" under Author Dashboard and select "Special Section Paper" as the paper type, and choose "MODELS 2021 Special Issue".
-Also, make sure to mention the MODELS 2021 special issue in the cover letter.
-\fi
-
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item Revise the end of the abstract regarding the case studies to be consistent with what ends up in this version of the paper.
-  \item Revise the end of the introduction that discusses the case studies to be consistent with what ends up in this version of the paper.
-  \item Fix the link to the motivating example to point to the new repository once it is created. Fix in the next section too.
-  \item Add discussion on the limitation of the approach: what attacks can and cannot be covered, can valid messages be rejected, etc.
-\end{compactitem}
-\fi
+\section{Overview of BriefCASE}
+\label{sec:overview}
+\input{overview}
 
@@
 
-\section{Code contracts for cyber components}
+\section{Code contracts for high assurance components}
 \label{sec:code-contracts}
@@
 
-\section{Testing Specifications with AGREE}
+\section{Test contracts}
 \label{sec:testing}
@@
 
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item State theorem relating the step function to the meaning of the leaf-node semantics
-\end{compactitem}
-\fi
-
-\section{Case studies}
+\section{Case study}
 \label{sec:case-study}
 \input{case-study}
-
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item Possible Outline:
-  \begin{compactitem}
-    \item Two case studies: phase 2 is the proof on concept on a industrial size problem while phase 3 tests if AGREE is expressive enough for stating complex properties and if designers are able to use it.
-    \item Phase 2 write up focusses on the number of components and the relative ease for AGREE to prove out and for SPLAT to synthesize. It notes that the components were relatively simple.
-    \item Phase 3 write up focusses on the shear complexity of the components and in particular the gate and the transport monitor.
-Here it the need for testing is is most obvious.
-Also obvious is the need to make AGREE more expressive and to separate AGREE into a specification and implementation language as seen in Dafny.
-    \item Make the general observation that the transforms on the AADL model to add the components and connections doesn't really save much time when compared against the time it takes to write and test the specifications.
-The automations is nice, but not where time is saved.
-Where time is saved is in the assurance case because once a specification is done, it is synthesized and proved correct for ALL inputs and ALL outputs.
-Not just some.
-Not code inspections.
-Not additional test requirements.
-All that is saved.
-  \end{compactitem}
-  \item Clarify research questions answered by the case study
-  \item Make the figures bigger
-  \item Rewrite to focus on synthesized components and the AGREE specifications
-  \item Add an interesting monitor from the TA6 model
-  \item Discuss what efforts are optimized with the automations versus a fully manual process for creating the components--how does the automated transformation affect actual time for engineers
-\end{compactitem}
-\fi
 

### 2022-09-14 paper/overview.tex
Revise introduction

@@
 
-The BriefCASE OSATE environment, \figref{fig:tool-arch}, provides designers with workflow and tool support for developing products with inherent cyber-resiliency.
+The \brfcs\ OSATE environment, \figref{fig:tool-arch}, provides designers with workflow and tool support for developing products with inherent cyber-resiliency.
 It starts with the development of a baseline AADL model of the system architecture that defines the desired functionality.
@@
 The Awas AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model \cite{awas}.
-The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model component infrastructure and integration code implementing the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
-Along with that code generation, HAMR constructs a proof that all flows in the AADL model are in the implementation and that no extraneous flows have been added to the implementation.
+The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model the component infrastructure and integration code that implements the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
+Along with that code generation, HAMR constructs a proof that all flows in the model are in the implementation and that no extraneous flows have been added to the implementation.
 HAMR relies on the seL4 microkernel for its proofs.
@@
 
-The Assume Guarantee Reasoning Environment (AGREE) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
-AGREE attempts to prove properties about one layer of the architecture using properties allocated to sub-components.
+The Assume Guarantee Reasoning Environment (\agr) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
+\agr\ attempts to prove properties about one layer of the model using properties allocated to sub-components.
 The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
-AGREE verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.
+\agr\ verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.
 
 The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
-This manuscript defines a \emph{code contract} language that is a subset of the AGREE language with sequential evaluation semantics as opposed to the existing parallel semantics of AGREE.
-The sequential semantics are requisite to code synthesis and are defined such that AGREE is able to still prove consistency in the contract composition of the model. 
+This manuscript defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
+The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model. 
 This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
-As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing AGREE model checker, test contracts prove the desired code contract behavior, and once the designer is satisfied that is correct, the system composition is verified to conclude that the contract meets its intended purpose.
-dressing associated cyber-requirements.
+As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing \agr\ model checker, test contracts prove the desired code contract behavior.
+Once the designer is satisfied that the code contract is correct, the system composition is verified to conclude that the contract meets its intended purpose.
 
-Showing that a component is properly implemented is accomplished through formal synthesis from the code contract.
-The Semantic Properties of Language and Automata Theory (SPLAT) tool generates \emph{CakeML} code to implement the code contract.
-The CakeML compiler is verified to generate binary implementations that are equivalent in semantics to the original input program \cite{cakeml}.
-It proves that the resulting binary exactly preserves the meaning of the original contract to assure that the component is properly implemented.
+Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
+The Semantic Properties of Language and Automata Theory (\splt) tool generates \emph{\ckml} code to implement the code contract.
+\ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.
+The \ckml\ compiler itself is verified to generate binary implementations that are equivalent in semantics to the original input program.
+That proof ensures that the resulting binary exactly preserves the meaning of the original code contract and assures that the component is properly implemented.

### 2022-09-14 paper/example.tex
Revise example

@@
   \end{center}
-\caption{AGREE failure certificate for initial design.}
+\caption{Simplified \brfcs\ workflow with the contributions detailed in this manuscript shaded.}
 \label{fig:flowchart}
@@
   \end{center}
-\caption{AGREE failure certificate for initial design.}
+\caption{\agr\ failure certificate for initial design.}
 \label{fig:example-certificate}
 \end{figure}
-
-\egm{
-  Merge the AGREE overview here into this section.
-  Condense it down to just the opening two paragraphes that currently exist in the overview section.
-  Walk through the first part of the example up to where AGREE fails because of Cyber Requirements.
-  Introduce the new figure showing its workflow.
-  Finish the example walking through the new figure. 
-  Add the tests to occur before final AGREE check.
-  Send to SPLAT to get verified CakeML code. 
-}
 
@@
 The expected behavior of the SW system, and the components
-implementing the system, are modeled with AGREE contract
+implementing the system, are modeled with \agr\ contract
 specifications.  
@@
 data at a port. For example, a waypoint is well-formed if it falls
-within bounds for latitude, longitude, and altitude.)  The guarantees
+within bounds for latitude, longitude, and altitude.  The guarantees
 for SW ensure that
@@
 component input assumptions and system output guarantees.
-AGREE background and the baseline contracts are discussed in detail in \secref{sec:agree} and \secref{sec:agree-semantics}.
+The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree} and \secref{sec:agree-semantics}.
 
@@
 A cyber-vulnerability analysis identifies the potential of a
-supply chain attack through the AI route planner (provided by a
-third-party vendor without source code).
-Based on the analysis, the designer adds a requirement that the system must guard agaist malicious AI behavior, marks the component as \emph{untrusted}, and removes all output guarantees from its contract.
-The AI output is now unconstrained in the
-assume-guarantee reasoning of AGREE and is able to generate any value
-on its output at anytime.
+supply chain attack through the AI route planner since its third-party.
+Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior, marks the component as \emph{untrusted}, and removes all output guarantees from its contract.
+The behavior of AI is now unconstrained.
 
-The output from the AGREE analysis using the untrusted AI
-specification is shown in \figref{fig:example-certificate}.  The red
+The output from the \agr\ analysis is shown in \figref{fig:example-certificate}.  The red
 exclamation points designate component assumptions or system output
 guarantees that do not hold, and each failure comes with a
-corresponding counter-example.
-
-The first violation is that the mission command on \texttt{AutomationResponse} from
+corresponding counter-example.  The first violation is that the mission command on \texttt{AutomationResponse} from
 the AI to the WM is no longer guaranteed to be well-formed.  The
 consequence of that failing input assumption is that the WM outputs
-are no longer guaranteed; they are effectively unconstrained.  These
-missing output guarantees from the WM lead to the rest of the failures
-in \figref{fig:example-certificate} because the WM provides the system
+are no longer guaranteed.  These
+missing guarantees lead to the rest of the failures
+in \figref{fig:example-certificate} as the WM provides the system
 level outputs.
@@
   \end{center}
-  \caption{AGREE verification certificates. (a) Test contract verification results. (b) Cyber-hardened design verification results.}
+  \caption{\agr\ verification certificates. (a) Test contract verification results. (b) Cyber-hardened design verification results.}
   \label{fig:hardened-certificate}
@@
 
-\emph{Adding high assurance components}
+\subsection{Adding high assurance components}
 
+Here is where the shaded regions in \figref{fig:flowchart} come into play.
 The system designer now uses \brfcs\ to cyber-harden SW by inserting
@@
 shown in \figref{fig:hardened}.
-These isolate the AI component and prevent it from violating the WM component input assumptions.
+These are intended to mitigate and report any observed malicious behavior by AI.
 A filter enforces an invariant over
 each datum in the data stream by not forwarding input to its output if the invariant does not hold.
-\brfcs\ generates the initial code contract for the filter which the designer completes by writing the invariant property.
-That property it is usually based on the existing assumptions made by
+\brfcs\ generates an initial code contract for the filter which the designer completes by writing the invariant.
+That invariant is usually based on the existing assumptions made by
 downstream components that consume the filter output.
+In this example, it is the well-formed property for the automation request.
 
-A monitor captures a relation on input data over time and is thus able
-to reason about the temporal, and other invariant, properties of that input.  A monitor raises
+A monitor captures a relation on data over time and is thus able
+to reason about the temporal, and other invariant, properties of that data.  It raises
 an alert if the specified properties are ever violated.  
-As with the filter, \brfcs\ generates a template code contract that is completed by the designer with the desired behavior that is typically defined by the SW system and its implementing component contracts.
-The code contract in our example states that an
+As with the filter, \brfcs\ generates an initial code contract that is completed by the designer. 
+In this example, that contract states that an
 automation response can only be generated in conjunction with an
@@
 Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
-Test contracts allow the designer to validate the behavior, and general input to output properties, of a code contract through unit testing.
-For example, one of the test contracts for the monitor is that it should not alert and pass its input when the monitor sees the \texttt{AutomationReponse} one step after it sees the \texttt{AutomationRequest}.
+Test contracts allow the designer to validate their behavior with unit testing.
+For example, one of the test contracts for the monitor is that when it sees an automation response one step after a request, then it should not alert, and it should pass the response downstream.
 Test contracts are discussed in detail in \secref{sec:test-contracts}.
 
-The AGREE analysis of the cyber-hardened implementation is shown in
+The \agr\ analysis of the cyber-hardened implementation is shown in
 \figref{fig:hardened-certificate}.
 \figref{fig:hardened-certificate}(a) is the test contract results and \figref{fig:hardened-certificate}(b) is the hardened system composition.
-Here AGREE has generated high-level
+Here \agr\ has generated high-level
 evidence justifying the claim that the high-assurance components meet their intended purposes.
-Having passed AGREE verification, the high-assurance components are ready to be
+Having passed \agr\ verification, the high-assurance components are ready to be
 synthesized.
@@
 
-High-assurance components are automatically synthesized by SPLAT from
-the code contracts to equivalent programs in the CakeML
+High-assurance components are automatically synthesized by \splt\ from
+the code contracts to equivalent programs in the \ckml\
 language.
-%% Too strong at the moment!
-%The synthesis toolchain generates a proof that equates the
-%% meaning of the AGREE specification to the behavior of the CakeML
-%% program.
 In other words, for any set of input streams that meet the
 component's contract assumptions, the output streams produced from
- the synthesized CakeML code exactly match those defined by the
- high-assurance component's guarantees.  The CakeML compiler provides
- verified compilation to binaries for several different platforms,
- meaning that the resulting binaries exactly preserve the meaning of
- the original CakeML code \cite{cakeml}.
+the synthesized \ckml\ code exactly match those defined by the contract guarantees.
+As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the resulting binaries from the \ckml\ compiler.
+This result does not extend to the whole system composition.
+Synthesis from code contracts is discussed in detail in \secref{sec:sythesis}.
 
-Preserving the input to output relationship of streams between the
-AGREE contracts and CakeML lifts the AGREE contract verification
-results to the actual deployed system.  If the contract model
-verification succeeds, then the meaning of those results hold for the
-deployed system.  These results, however, are only valid under
-additional assumptions on the deployed system:
-\begin{compactitem}
-\item contracts for non-synthesized components accurately model their deployed
-counterparts;
-\item an appropriate schedule exists to sequence component
-  execution following the dependent data-flow in the design; and
-\item the communication fabric stitching components together works in harmony
-  with the schedule.
-\end{compactitem}
-\noindent Support and automation for these aspects of the design process are
-discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.
-Synthesis from code contracts is discussed in detail in \secref{sec:sythesis}.
+Extending the \agr\ verification results to the system composition requires additional guarantees not discussed in this manuscript.
+For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, and the HAMR generated scheduler follows the dependent data-flow in the design.
+These additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.

### 2022-09-15 paper/agree-semantics.tex
Revision to make paper flow with overview

--- 2022-03-28 paper/agree-semantics.tex
@@
-\begin{comment}
-This section gives an abstract mathematical overview of AGREE's
-compositional reasoning system, defining assume-guarantee contracts
-and the verification conditions that AGREE checks in order to
-modularly establish overall system correctness at the model level.
-The syntax and semantics of AGREE are sketched in some detail.  The
-discussion then becomes more concrete, showing---by example---how the
-AGREE contract language is used to specify the cyber-hardened system
-of \figref{fig:hardened}. The discussion proceeds from the
-system-level contract to the filter and monitor components, which rely
-on the notion of a \emph{code contract} to support code generation.
-\end{comment}
-
-We now give an overview of a formal model for AGREE syntax and
-semantics.  The discussion then becomes more concrete, showing by
-example, how the AGREE language is used to specify the properties of
-the SW system introduced in \secref{sec:example}.  Any implementation
-of the SW system must guarantee the specification output under the
-system assumptions.
-
-The syntax of AGREE is essentially that of quantifier-free first order
+The syntax of \agr\ is essentially that of quantifier-free first order
 predicate logic supplemented with a few temporal operators. The terms
@@
 \]
-There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
-and formulas. To support port handling, there is a predicate
+\agr\ includes a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
+and formulas. To support port handling, \agr\ defines a predicate
 $\konst{event}(p)$, which tests if an event (or event-data) port $p$ is
-signaled. Lastly, there are temporal operators \emph{previously}
+signaled. Lastly, \agr\ supports temporal operators \emph{previously}
 $\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically
@@
 \subsection{Semantics}
-The semantics of terms and formulas, as mentioned in the previous
-section, is in terms of \emph{streams of values}. Values encompass at
-least booleans and numbers, but can be readily extended to include
+The semantics of terms and formulas is defined over \emph{streams of values}. Values encompass at
+least booleans and numbers but can be readily extended to include
 records and arrays. A value stream is a total function from time
@@
 
-\paragraph{AGREE notation.}
-AGREE uses ASCII notation for the mathematical syntax we have
+\agr\ uses ASCII notation for the mathematical syntax we have
 defined. The following table spells out the mapping.
@@
 \begin{tabular}{l|l|l}
- Operation & Math. notation & AGREE \\ \hline
- followed-by &  $e \to e$          & e \verb+->+ e  \\
- logical implication &  $e \imp e$ & e \verb+=>+ e \\
- if-and-only-if &  $e \iff e$      & e \verb+<=>+ e \\
- variable definition & $v = e$     & \konst{eq} v\ e \\
- Historically  & $\konst{Hist}(b)$ & \konst{Historically}(b) \\
+ Operation & Math. notation & \agr \\ \hline
+ followed-by &  $e \to e$          & \verb+e -> e+  \\
+ logical implication &  $e \imp e$ & \verb+e => e+ \\
+ if-and-only-if &  $e \iff e$      & \verb+e <=> e+ \\
+ logical and &  $e \land e$      & \verb+e and e+ \\
+ logical or &  $e \lor e$      & \verb+e or e+ \\
+ logical not & $\neg b$      & \verb+not b+ \\
+ variable definition & $v = e$     & \verb+eq v e+ \\
+ Previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
+ Historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
 \end{tabular}

### 2022-09-15 paper/example.tex
Revision to make paper flow with overview

@@
 The baseline AADL model is a simple software system, named SW, for route planning
-and automated control of a UAV. A picture of the AADL architectural
+and automated control of a UAV. A picture of the architectural
 model for SW is in \figref{fig:example}. SW is loosely based on one of
@@
 updates with messages on \texttt{AirVehicleLocation}.
-An event is repeatedly generated on \texttt{Alert} if there is no
-mission command from a request. The AI component is third-party
+%%An event is repeatedly generated on \texttt{Alert} if there is no
+%%mission command from a request. 
+The AI component is third-party
 software and WM is a legacy component.
@@
 \begin{compactitem}
-\item a start coincides with a new waypoint message being output;
+\item a start coincides with a new waypoint being output;
 \item a start is within one cycle of an automation request and if not, then it persistently alerts;
-\item new waypoints coincide with air vehicle location updates; and
+\item new waypoints coincide with location updates; and
 \item all outputs are well-formed.
@@
 contract guarantees it only responds to automation requests and always
-in the same cycle.  The legacy WM contract guarantees the following if
-its input assumptions hold:
+in the same cycle.  The legacy WM contract guarantees that
 \begin{compactitem}
   \item it generates a start from a response from the AI and always in the same cycle;
-  \item the start always coincides with a waypoint being output; and
-  \item any further waypoints coincide with air vehicle location updates.
+  \item the start coincides with a waypoint; and
+  \item any further waypoints coincide with location updates.
 \end{compactitem}
+
 These initial specifications pass verification, meaning that the
@@
 component input assumptions and system output guarantees.
-The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree} and \secref{sec:agree-semantics}.
+The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}.
 
@@
 A cyber-vulnerability analysis identifies the potential of a
-supply chain attack through the AI route planner since its third-party.
-Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior, marks the component as \emph{untrusted}, and removes all output guarantees from its contract.
-The behavior of AI is now unconstrained.
+supply chain attack through the AI route planner since its third-party.  Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior.
+They additionally remove all output guarantees from its contract since it is untrusted.
 
@@
 These are intended to mitigate and report any observed malicious behavior by AI.
+
 A filter enforces an invariant over
@@
 Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
-Test contracts allow the designer to validate their behavior with unit testing.
+Test contracts allow the designer to validate code contract behavior with unit testing.
 For example, one of the test contracts for the monitor is that when it sees an automation response one step after a request, then it should not alert, and it should pass the response downstream.
-Test contracts are discussed in detail in \secref{sec:test-contracts}.
+Test contracts are discussed in detail in \secref{sec:testing}.
 
@@
 the synthesized \ckml\ code exactly match those defined by the contract guarantees.
-As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the resulting binaries from the \ckml\ compiler.
-This result does not extend to the whole system composition.
-Synthesis from code contracts is discussed in detail in \secref{sec:sythesis}.
+As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the results equally apply to the resulting binaries from the \ckml\ compiler.
+Synthesis from code contracts is discussed in detail in \secref{sec:synthesis}.
 
 Extending the \agr\ verification results to the system composition requires additional guarantees not discussed in this manuscript.
-For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, and the HAMR generated scheduler follows the dependent data-flow in the design.
-These additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.
+For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, the HAMR generated scheduler must be certified to follow the dependent data-flow in the design, scheduling windows must be certified to cover worst-case execution times, etc.
+Such additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.

### 2022-09-15 paper/high-assurance-contracts.tex
Revision to make paper flow with overview

@@
 
-The semantics of Section \ref{agree-semantics} supports a formal
+The semantics of \secref{sec:agree-semantics} supports a formal
 connection between the original contract---and verification results of

### 2022-09-15 paper/overview.tex
Revision to make paper flow with overview

@@
 
-\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline AADL model (top of \figref{fig:tool-arch}).
+\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline model (top of \figref{fig:tool-arch}).
 \brfcs\ provides access to analysis tools that examine the model, report potential cyber vulnerabilities,
@@
 	\item \textbf{Attestation}: measure non-local software to assess trustworthiness
-	\item \textbf{Virtualization}: isolate components with virtualization
+	\item \textbf{Virtualization}: isolate components
 	\item \textbf{Proxy}: enable HTTPS message inspection 
@@
 \end{compactitem}
-Transformations update the AADL model architecture with new components and the assurance case with new required evidential statements.
+Transformations update the model's architecture with new components and the assurance case with new required evidential statements.
 
@@
 The Awas AADL information flow analyzer and visualizer provides evidence that high assurance components are connected and not bypassed in the model \cite{awas}.
-The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model the component infrastructure and integration code that implements the semantics of AADL-compliant thread scheduling, thread dispatching, port-based communication, and real-time scheduling \cite{hamr}.
-Along with that code generation, HAMR constructs a proof that all flows in the model are in the implementation and that no extraneous flows have been added to the implementation.
-HAMR relies on the seL4 microkernel for its proofs.
-The seL4 microkernel, among other security properties, provides strong isolation between components ensuring that there are no communication channels outside of what is defined \cite{sel4-sosp09, sel4-tocs14, sel4-cacm18}.
+The High Assurance Modeling and Rapid engineering for embedded systems tool (HAMR) generates from the model the infrastructure and integration code that implements the semantics of AADL-compliant scheduling, dispatching, and port-based communication \cite{hamr}.
+HAMR additionally constructs a proof that all flows in the model are in the implementation and that no extraneous flows have been added.
+HAMR relies on properties of the seL4 microkernel for some aspects of these proofs.
+The seL4 microkernel, among other things, provides strong isolation between components ensuring that there are no communication channels outside of what is defined \cite{sel4-sosp09, sel4-tocs14, sel4-cacm18}.
 
-The Assume Guarantee Reasoning Environment (\agr) is a compositional, assume-guarantee-style, model checker for AADL models \cite{compositional-analysis-agree,nfm:agree}. 
+Important to the contributions presented in this paper is the Assume Guarantee Reasoning Environment (\agr) \cite{compositional-analysis-agree,nfm:agree}.
+It is a compositional, assume-guarantee-style, model checker for AADL models.
 \agr\ attempts to prove properties about one layer of the model using properties allocated to sub-components.
@@
 The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model.
+
 This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
@@
 Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
-The Semantic Properties of Language and Automata Theory (\splt) tool generates \emph{\ckml} code to implement the code contract.
+This manuscript presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \emph{\ckml} code to implement the code contract.
 \ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.

### 2022-09-15 paper/testing.tex
Revision to make paper flow with overview

--- 2022-03-28 paper/testing.tex
@@
-Before discussing the final steps in synthesis, we take a moment to
-discuss how code contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
-Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
-The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.
+Writing and proving properties about code contracts is difficult because it is a formal process, and because it reasons about constraints that define the entire input space, and the entire output space, of a high-assurance component at once.
+Case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
 
-We have observed three common types of mistakes in specifications (aside from self-consistency, which is
+We have observed three common types of mistakes in code contracts (aside from self-consistency, which is
 already checked): vacuity, under-specification, and missed corner cases.
 Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true. 
-Under-specification is very challenging, as a verifier in the face of under-specification is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the system is under-specified.
+Under-specification is very challenging, as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
 Missed corner cases is a common problem to all disciplines.
-It is an especially difficult challenge here because there is no direct way to provide test input and check test output.
+It is an especially difficult challenge here because there is no direct to exercise code contracts (e.g., there is no runtime \emph{per se}).
 
-The tempting approach is to use the synthesis path to check these issues with traditional testing.
-Such an approach adds extra steps that are not needed, as it is
-possible to bridge the formal model in AGREE, with its proof system, to
+A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with with traditional testing.
+Such an approach adds extra steps, including test harneses, that are not needed, as it is
+possible to bridge the formal model in \agr, with its proof system, to
 traditional validation using tests.
-The key is to consider the component as implementing a test specification.
+The intuition for that bridge comes from the compositional reasoning in AADL and \agr. 
+Effectively, a code contract for a high assurance component should be a valid implementation of a test scenario.
+In other words, the test scenario, as expressed in a test contract, is the desired behavior of the system, and the code contract implements that behavior.
 
-Unit tests are created in AGREE by writing a system specification that assumes the test input and then guarantees the test output.
-The system is then implemented with the component under test and its specification.
-The AGREE analysis then proves if the component guarantees, given the test inputs, are strong enough to prove the guaranteed test outputs.
-Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
-AGREE proves that the component implements the test, and the Liskov substitution principle holds.
+A test contract defines the test scenario for the code contract to implement.
+It does this by assuming the test input and guarantying the test output akin to a unit test. 
+The code contract implementation (e.g., the subject under test).
+The \agr\ analysis then proves if the code contract, given the test inputs, is strong enough to prove the test contract outputs.
+Here the test contract assumptions strengthen the code contract assumptions to a single input, and the test contract guarantees weaken, or leave unchanged, the code contract output for that test input, depending on the goal of the test.
+In other words, according to Liskov substitution, \agr\ proves if the code contract is a safe substitution for the test contract.
 
@@
     \begin{tabular}{c}
-    \scalebox{0.62}{\usebox{\tst}} \\
+    \scalebox{0.60}{\usebox{\tst}} \\
     \end{tabular}
   \end{center}
-  \caption{AGREE unit test for high-assurance monitor.}
+  \caption{A unit test for the monitor with its test contract.}
   \label{fig:test}
@@
 
-\figref{fig:test} is a unit test for the high-assurance monitor in \figref{fig:hardened}.
-The test checks if the monitor output coincides with the response that comes one step after the request.
-It also checks that the alarm does not fire.
+\figref{fig:test} is a unit test, with its test contract, for the monitor in \figref{fig:hardened}.
+The test contract checks if the monitor output coincides with the response that comes one step after the request.
+It also checks that the alarm is not output.
 \begin{compactitem}
@@
 \end{compactitem}
-\figref{fig:test-output} is the AGREE output for the test, showing that it passes.
+\figref{fig:test-output} is the \agr\ output for the test, showing that it passes.
 
-This AGREE test method is effective for detecting vacuity, under-specified behavior, and missed corner cases.
-It enables designers to iteratively develop specifications inside the MBSE framework using AGREE, and it builds assurance that the specification reflects the desired behavior with black-box, white-box, and other test coverage metrics.
-These same tests can be synthesized to the backend implementation for traditional testing.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c}
-      \includegraphics[scale=0.38]{agree-test-output.png} \\
-    \end{tabular}
-  \end{center}
-  \caption{AGREE output for unit test.} 
-  \label{fig:test-output}
-\end{figure}
+Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
+They enables designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract reflects the desired behavior with black-box, white-box, and other test coverage metrics.
+These same tests can be synthesized to the backend implementation for traditional testing if for example, the code contract is a model of an existing legacy binary.

### 2022-09-15 paper/verification-conditions.tex
Revision to make paper flow with overview

@@
 
-The AGREE specification language, and semantics, are inspired by the
+The \agr\ specification language, and semantics, are inspired by the
 Lustre language \cite{10.1145/41625.41641}. Lustre semantics are
@@
 streams.  Informally, streams assign values to expressions at each
-point in time.  AGREE contracts express relationships between input
+point in time.  \agr\ contracts express relationships between input
 and output streams by assuming a set of allowed input streams and then
@@
 and the connections between ports on different subcomponents.  Ports
-in AGREE are inherited from AADL.  There are three kinds of port:
+in \agr\ are inherited from AADL.  There are three kinds of port:
 \emph{event} ports, \emph{data} ports, and \emph{event data} ports.
@@
 
-From the system and component contracts, AGREE generates a set of
+From the system and component contracts, \agr\ generates a set of
 verification conditions to show that a system's component
-implementation is correct~\cite{agree2013}.  The AGREE model checker
+implementation is correct~\cite{agree2013}.  The \agr\ model checker
 is then invoked to prove or disprove the verification
@@
 
-The two PLTL operators necessary for the AGREE generated verification
+The two PLTL operators necessary for the \agr\ generated verification
 conditions are $\globally$ that looks forward in time along
@@
 
-If all the verification conditions hold (AGREE uses $k$-inductive
+If all the verification conditions hold (\agr\ uses $k$-inductive
 model checking to automatically prove or disprove each generated
@@
 A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
-AGREE generates additional verification
+\agr\ generates additional verification
 conditions that prove each component contract, and the composition of

### 2022-09-15 paper/abstract.tex
Revise the introduction

--- 2022-03-27 paper/abstract.tex
@@
 Cyber-physical systems, such as avionics, must be tolerant to
-cyber-attacks in the same way they are tolerant to random faults: they
-either gracefully recover or safely shut down as requirements dictate.
-The DARPA Cyber Assured Systems Engineering program is developing
-tools for design, analysis, and verification that enable systems
-engineers to design-in cyber-resiliency in a Model-Based Systems
-Engineering (MBSE) environment.  This paper describes automated model
-transformations that introduce high-assurance cyber-resiliency
-components into a system; in particular, filters that guard against
-malformed input, as well as monitors that guard against
-spoofing and other malicious behavior.  A formal specification in the
-form of a code contract defines each high-assurance component and is
-used to verify that the component addresses system-level cyber
-requirements.  Implementations for these high-assurance components
-are directly synthesized from their code contracts, and are
-automatically proven to preserve the exact meaning of the code
-contract all the way down to the binary code level.  The model
-transformations are integrated into the Open Source AADL 
-Tool Environment (OSATE).  We further report on two case
-studies, one by formal method experts and the other by non-experts,
-applying model transformations to two different systems to improve
-cyber resiliency using high-assurance filters and monitors.  One is a
-UAV system utilizing the Air Force Research Laboratory's
-OpenUxAS services for route planning.  The other is a tablet used for
-situational awareness in a CH-47F Chinook helicopter application.
-The case studies demonstrate that our MBSE approach is feasible in
-real-world systems engineering.
+cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
+\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
+This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
+Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
+A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
+Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
+Implementations for these
+high-assurance components are directly synthesized from their code
+contracts and are backed up by proofs showing that high-level
+specifications map in a meaning-preserving way to code.
+The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
+route planning.  The case study
+demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
+engineering.

### 2022-09-15 paper/intro.tex
Revise the introduction

@@
-\egm{Rewrite to make clear the contribution here: code contracts, tests, and synthesis through SPLAT. 
-Present the \emph{unverified} path as is done in the introduction.
-See if it is possible to present the verified path (ping Konrad) to state the various correctness guarantees.}
-
 In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
@@
 
-Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.
+Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
+They instead rely on process-oriented checklists and guidelines.
 Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
@@
 Government.}
-We have developed a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
+One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
 BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
 
+This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
+These are referred to as \emph{transforms} by \brfcs.
+Two cyber-resiliency transformations are discussed: (1) the insertion
+of a filter to prevent malformed data from a malicious actor from
+being propagated to downstream components, and (2) the insertion of a
+monitor to detect (and alert) unexpected behaviors arising from
+untrusted components.
+\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.
+
+A code contract is a formal specification in the Assume Guarantee Reasoning
+Environment (\agr) language.
+\agr is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
+The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
+These contracts are unit tested for correctness with \emph{test contracts}.
+Test contracts define test scenarios to be implemented by the code contract under test. 
+\agr\ proves whether or not that code contract implementation is correct enabling to designer to iteratively test the component behavior inside the \brfcs\ environment.
+Once the behavior of the code contract is verified, \agr proves that---due to the newly
+included high-assurance components---the hardened system meets its
+cyber-resiliency requirements.
+
+Another novel aspect of the approach discussed in this manuscript is the synthesis of the code contracts for the high-assurance components to \ckml, a verified
+compiler implementation for the functional programming language
+ML \cite{cakeml}. This manuscript describes in detail the synthesis path
+from code contracts to \ckml\ code, providing a formal framework in
+which to argue correctness. \ckml\ then provides a verified
+compilation path to several different target binaries (and also
+proving that the meaning of the \ckml\ source code is exactly
+preserved in the final binaries). 
+A key contribution here is that the code contract semantics are defined in such a way that the \agr\ verification 
+results for the code contract hold for the deployed component, \ie, the component will detect
+and prevent the indicated cyber-vulnerabilities over all possible
+finite inputs. Preliminary work has shown how to lift this result to
+infinite input traces as these systems are inherently reactive and
+intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+
+The manuscript further details a 
+case study applying these transformations with \brfcs\ to an Unmanned Aerial
+Vehicle (UAV) system that uses the Air Force Research Laboratory's
+OpenUxAS services for route planning.
+OpenUxAS, as an open source product, is considered \emph{untrusted}.
+The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
+Here the transforms add filters to guard
+against malformed input and monitors to guard against malicious flight plans from OpenUxAS. The case study
+system is complex and shows
+the viability of the approach in potential full-scale industrial
+design.
+
+The \brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
+Our approach currently applies to common cyber-vulnerabilities, such
+as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
+side-channel attacks and denial of service are not yet dealt with in
+our work.  Here, we do not report on the invention of a new type of
+high-assurance component in terms of capability; instead, our contribution
+is in the automated synthesis of security-improving components from
+formal specifications and a means to show that the synthesis is
+correct.
+
+The approach is illustrated by a simple example in
+\secref{sec:example}. \agr\ verification and specification are
+presented in \secref{sec:agree} and \secref{agree-semantics}.
+Code contracts are defined in \secref{sec:code-contracts} with a method
+for testing code contracts in the \agr\ framework before synthesis given
+\secref{sec:testing}.
+The synthesis pathway is covered in Section~\ref{sec:synthesis}.
+
+\begin{comment}
   BriefCASE incorporates model-level cyber analysis tools (presently
@@
   secure microkernel~\cite{sel4-2009}.
-
-%% DSH  This paragraph seems mostly duplicative of the paragraph above; attempted to merge the two.
-%A novel aspect of the BriefCASE approach is the ability to automatically transform the model to insert high-assurance %components to satisfy requirements added to mitigate cyber vulnerabilities.
-In this work, the behavior of the system is %modeled by contracts on components expressed in the AGREE modeling language.
-These contracts are modified as a result %of the cyber-vulnerability analysis with the new requirements for mitigation.
-The requirements assume properties on the %input and guarantee properties of the output that reflect a hardened system.
-In the absence of any transformation to %implement such requirements, AGREE fails to prove the system meets these new requirements.
-
-Two cyber-resiliency transformations are discussed: (1) the insertion
-of a filter to prevent malformed data from a malicious actor from
-being propagated to downstream components, and (2) the insertion of a
-monitor to detect (and alert) unexpected behaviors arising from
-untrusted components. These transformations not only change the
-architecture of the model by adding in new components; they also
-generate a formal specification in the form of a \emph{code contract} in the \agr\ language for their intended behavior. 
-A code contract describes the computation of a high-assurance component and not just what it computes.
-These code contracts
-are sufficient for model checking to prove that---due to the newly
-included high-assurance components---the hardened system meets its
-cyber-resiliency requirements.
-
-Another novel aspect of the approach is the synthesis of the \agr\ 
-code contracts for the high-assurance components to \ckml, a verified
-compiler implementation for the functional programming language
-ML \cite{cakeml}. This paper describes in detail the synthesis path
-from code contracts to \ckml\ code, providing a formal framework in
-which to argue correctness. \ckml\ then provides a verified
-compilation path to several different target binaries (and also
-proving that the meaning of the \ckml\ source code is exactly
-preserved in the final binaries). Assuming that the execution schedule
-of the deployed cyber-hardened system is as intended by the AADL
-model, and that the HAMR-generated communication fabric delivers
-messages between components as expected, the \agr\ model checking
-results then hold for the deployed system, \ie, the system will detect
-and prevent the indicated cyber-vulnerabilities over all possible
-finite inputs. Preliminary work has shown how to lift this result to
-infinite input traces as these systems are inherently reactive and
-intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
-
-Our approach currently applies to common cyber-vulnerabilities, such
-as overflow, lack of input validation, supply-chain issues, and
-identity verification; however, other cyber-vulnerabilities such as
-side-channel attacks and denial of service are not yet dealt with in
-our work.  Here, we do not report on the invention of a new type of
-filter or monitor, in terms of capability; instead, our contribution
-is in the automated synthesis of security-improving components from
-formal specifications, and a means to show that the synthesis is
-correct.
-
-The approach is illustrated by a simple example in
-\secref{sec:example}. \agr\ verification and specification are
-presented in \secref{sec:agree} and \secref{agree-semantics}.
-Code contracts are defined in \secref{sec:code-contracts} with a method
-for testing code contracts in the \agr\ framework before synthesis given
-\secref{sec:testing}.
-The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
-case study applying these transformations to an Unmanned Aerial
-Vehicle (UAV) system that uses the Air Force Research Laboratory's
-OpenUxAS services for route planning is presented in
-Section~\ref{sec:case-study}. Here the transforms add filters to guard
-against malformed input and monitors to guard against ground station
-spoofing and malicious flight plans from OpenUxAS. The case study
-system is significantly more complex than the simple example and shows
-the viability of the modeling approach to a full-scale industrial
-design.
-
-The \brfcs\ tools are open source and publicly
-available \cite{fmide}, as is the motivating example \cite{repo}, and
-the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
-Videos demonstrating the use of the \brfcs\ tools to build the UAV
-example presented in Section~\ref{sec:case-study} are also
-available \cite{case}.
+\end{comment}

### 2022-09-15 paper/abstract.tex
Add notes from Junaid

@@
 Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
-A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
+A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts for testing the specified behavior.
 Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.

### 2022-09-15 paper/intro.tex
Add notes from Junaid

@@
+\egm{Add to introduction and related work, that what is done at Galois called Copilot.
+Its a stream processing langague that generates code from the specification.
+Copilot to C.
+Effectively CodeGen from Lustre.
+Bounds how far needed to look in past for any value.}
 In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
@@
 They instead rely on process-oriented checklists and guidelines.
-Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation.
-This is not a sustainable development model.
+Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely
+expensive and time-consuming remediation. This is not a sustainable development model.
 
+\egm{Remove DARPA junk.}
 The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).
@@
 This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
-These are referred to as \emph{transforms} by \brfcs.
+These \egm{make clear what these refers to} are referred to as \emph{transforms} by \brfcs.
 Two cyber-resiliency transformations are discussed: (1) the insertion
@@
 Environment (\agr) language.
-\agr is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
+\agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
 The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
@@
 Test contracts define test scenarios to be implemented by the code contract under test. 
-\agr\ proves whether or not that code contract implementation is correct enabling to designer to iteratively test the component behavior inside the \brfcs\ environment.
-Once the behavior of the code contract is verified, \agr proves that---due to the newly
+\agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
+Once the behavior of the code contract is verified, \agr\ proves that---due to the newly
 included high-assurance components---the hardened system meets its
@@
 compiler implementation for the functional programming language
-ML \cite{cakeml}. This manuscript describes in detail the synthesis path
+ML \cite{cakeml}. 
+This manuscript describes in detail the synthesis path
 from code contracts to \ckml\ code, providing a formal framework in
@@
 
-The \brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
+\brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
 Our approach currently applies to common cyber-vulnerabilities, such
@@
 
-The approach is illustrated by a simple example in
-\secref{sec:example}. \agr\ verification and specification are
-presented in \secref{sec:agree} and \secref{agree-semantics}.
-Code contracts are defined in \secref{sec:code-contracts} with a method
-for testing code contracts in the \agr\ framework before synthesis given
-\secref{sec:testing}.
+We now give a summary of the contributions detailed in this manuscript.
+\begin{compactitem}
+  \item The language and semantics of code contracts to specify the behavior of high assurance components.
+  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
+  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
+  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
+  \item A case study that used the implementation of the approach in \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
+\end{compactitem}
+The rest of this manuscript is organized as follows.
+\secref{sec:overview} is an overview of the \brfcs\ environment and key tools in that environment relative to the contributions of this work. The approach is illustrated by a simple example in
+\secref{sec:example}. Essential background on \agr\ verification and specification are
+presented in \secref{sec:agree}.
+The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are defined in \secref{sec:testing}.
 The synthesis pathway is covered in Section~\ref{sec:synthesis}. 
+\secref{sec:case-study} discusses the case study.
+This is followed by related work in \secref{sec:related-work}. 
+The conclusions and future work are in \secref{sec:conclusion}.
 

### 2022-09-16 paper/conclusion.tex
Revise conclusion based on reviews

@@
-The DARPA CASE program has created tools for systems engineers to
-integrate cyber-vulnerability analysis and mitigation into their
-development workflow. The resulting BriefCASE tool suite includes
-analysis tools for generating cyber requirements, cyber resiliency
-tools for addressing the requirements, verification tools for ensuring
-design correctness, and synthesis tools for generating provably
-correct code. Several of the BriefCASE transforms (filter, monitor,
-gate) insert components into the model whose behavior can be formally
-specified using the AGREE language.  The SPLAT tool can then
-automatically generate CakeML implementations for these components,
-along with proofs of correctness for assurance that the implementation
-satisfies the specification.
-%
-BriefCASE was applied to a full-scale case study using the Air Force
-Research Laboratory's OpenUxAS software, exercising a range of
-built-in cyber resiliency mitigations to meet cyber-requirements.  The
-size and scale of the study suggests BriefCASE meets the complexity
-demands of real-world design.
+Safety critical systems, such as those found in avionics, must be cyber-resilient in the same way that they are fault tolerant. 
+Cyber-resiliency means that they are tolerant to cyber-attacks.
+The \brfcs\ environment integrated into OSATE provides systems designers with tools to analyze and address cyber-vulnerabilities in the AADL architecture design of a system.
+We discussed here transformations provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
 
-Most recently, we have applied BriefCASE to the design of
+A filter enforces invariant properties on data such that data that violate the invariant are not forwarded.
+A monitor detects and alerts malicious behavior in the system. 
+The actual behavior of such filters, monitors, and other high-assurance components are formally specified by code contracts.
+A code contract is a dialect of the \agr\ language with formal semantics embedded in HOL for formal reasoning.
+A code contract states assumptions on the inputs and then provides output guarantees that completely define the output in terms of the inputs and the current state of the component.
+The code contract language itself is Turing complete.
+
+Code contracts are validated with unit tests that are verified with the \agr\ verification engine.
+A unit test in expressed in a test contract in the \agr\ language.
+A test contract is a stylized contract that strengthens the input assumptions of a code contract to a single input and either weakens or leaves unchanged the code contract output guarantees.
+\agr\ then proves whether or not the computation expressed in the code contract implements the test contract.
+In this way, a designer is able to test the behavior of the code contract.
+Once the code contract is tested, the \agr\ verification engine proves whether or note the system, with the added high-assurance components, meets its cyber requirements.
+
+Code contracts are automatically synthesized to \ckml.
+We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transitions function that only considers the current input and state of the component.
+That state transition function is then rendered in \ckml, and the \ckml\ compiler compiles it to the target binary providing the proof of equivalence.
+
+We further demonstrated the viability of the approach in real-world industrial design by reporting a case study using the Air Force Research Laboratory's OpenUxAS software.
+The study required a range of filters, monitors, and other mitigation components to meet cyber-requirements.
+The case study demonstrated the viability of the approach to address complex message formats and non-trivial cyber-requirements.
+
+The reported case study was carried out by those that created \brfcs\ with its several tools. 
+Future work is to conduct another case study involving system designers.
+We have started such a study with the design of
 an application using the Collins Common Avionics Architecture System
-(CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the
-DARPA CASE program.  Other ongoing and future work includes
-adding support for uninterpreted functions, mechanizing the
-correctness-of-synthesis proofs in HOL4, and lifting the proof results
-to infinite streams.
+(CAAS)~\cite{caas} on the CH-47F Chinook helicopter program.
+Early results are encouraging and are revealing ways to expand the expressiveness of the code contract language to make it easier to express iterative computation with arrays.
+
+Aside from adding expressiveness to the code contract language, future work is to formally verify every transformation step from the code contract to the final \ckml\ function that is compiled to the target binary.
+The HOL embedding of the code contract semantics we discussed here is the starting point.
+The key steps are as follows:
+\begin{compactitem}
+  \item Automatically add state to the code contract to remove $\konst{pre}(-)$ and prove the initial and modified contracts are equivalent for all input streams
+  \item Create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version.
+  \item Create a pure HOL4 function, suitable to \ckml\, from the step version of the code contract and prove it equivalent.
+\end{compactitem}
+We have automated in HOL4 the transformation to remove $\konst{pre}(-)$ in the first step in what we call \emph{temporal squashing}.
+We have additionally manually written proofs for every step for a set of example code contracts.
+From the manual proofs, we have started work on proofs for the first two steps that are universal in that they reason over any input code contract.
+The third step must be generated for each input contract, and the entire process still needs to be mechanized.
+Other future work includes lifting the proof results to infinite streams.

### 2022-09-16 paper/conclusion.tex
Revise conclusion

@@
 Aside from adding expressiveness to the code contract language, future work is to formally verify every transformation step from the code contract to the final \ckml\ function that is compiled to the target binary.
-The HOL embedding of the code contract semantics we discussed here is the starting point.
+The HOL4 embedding of the code contract semantics we discussed here is the starting point.
 The key steps are as follows:
@@
   \item Automatically add state to the code contract to remove $\konst{pre}(-)$ and prove the initial and modified contracts are equivalent for all input streams
-  \item Create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version.
-  \item Create a pure HOL4 function, suitable to \ckml\, from the step version of the code contract and prove it equivalent.
+  \item Create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version
+  \item Create a pure HOL4 function, suitable to \ckml, from the step version of the code contract and prove it equivalent
 \end{compactitem}

### 2022-09-16 paper/related-work.tex
Revise related work as per reviewer requests

@@
-Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, agree2013, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}.
-Automated proofs of realizability for assume-guarantee reasoning are useful for engineers implementing components in the system \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}.
-Algorithms for %actual
-component synthesis for Lustre models using k-induction or IC3/PDR provide an automated path from the assume-guarantee reasoning to an actual satisfying node implementation \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
-These algorithms synthesize code in the Lustre modeling language but do not provide a path to a low-level implementation that could be fielded.
+%% DO NOT REMOVE --- BUT FEEL FREE TO REVISE
+%% Specifically requested by the reviewers to be included in this paper!!!
+Another model based approach to secure system architectures follows a pattern similar to \brfcs\ by inserting pre-verified components into the architecture to give certain guarantees \cite{6209212}.
+In that work, the data connections are annotated with security properties as are the components.
+A secondary analysis determines if the components meet the security requirements of the data connections, and if not, inserts new security enhancing components in the connection.
+The insertion algorithm considers time cost when inserting components to preserve worst-case execution times mandated by the system.
+\brfcs\ has the ability to insert pre-verified components based on cyber-requirements, but does not currently have the ability to annotate connections with security properties or automatically infer where secure components need to be inserted.
 
-Contracts are similar to assume-guarantee reasoning except they target programming languages.
-They are often more expressive than assume-guarantee reasoning being stateful and higher-order \cite{10.1145/583852.581484}.
-As contracts are often written in the target language, synthesizing monitors is not a problem but comes with %significant
-overhead \cite{10.1007/978-3-642-28869-2_11}.
-A monitor for a contract can be removed when it can be statically proven that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}.
+The first steps toward synthesis of leaf-component contracts determined if leaf-component contracts are realizable \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}.
+These led to algorithms for component synthesis from Lustre models based on realizability \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
+The algorithms discover a transition relation that preserve the contract guarantees under all allowed inputs and can be directly transformed into the language intended for the system's implementation.
+Two such algorithms are implemented in the \agr\ verification engine \cite{jkind}.
+These synthesis algorithms take in a declarative specification that states \emph{what is to be computed}, and then they have to infer \emph{how to compute it} for the implementation.
+Code contracts are imperative, by design, so the designer specifies directly how the computation of the high-assurance component takes place thereby avoiding the question of realizability entirely.
+That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishe that.
 
 Data-flow semantics are well studied \cite{10.1145/41625.41641,97300,
-10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}. State
-machine semantics can be added to synchronous
+10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}.
+State machine semantics can be added to synchronous
 data-flow \cite{10.1145/1086228.1086261}. The idea is to translate
@@
 design.
+These approaches are similar to the code contracts in this paper in that they embed imperative computation into the synchronous data-flow semantics of Lustre which are then passed to compilers to render the final target binaries.
+These compilers however are not verified as is the \ckml\ compiler targeted in this work.
 
@@
 manipulation of memories akin to an imperative model. CompCert is used on the
-backend to create the final rendered executable. The work is quite similar to
+backend to create the final rendered executable \cite{compcert}. 
+This work is quite similar to
 ours in providing a verified compilation chain from Lustre. However,
@@
 conducted by compositional reasoning in AGREE.
+
+CoPilot is a domain specific language for writing runtime monitors for hard real-time distributed systems embedded in Haskel and requires working Haskel knowledge to write \cite{10.1007/s11334-013-0223-x}.
+The language semantics are defined over streams, there is an interpreter for testing the behavior of the runtime monitors, and there are two backend compilers to generate C code.
+The generated C code is validated by generating random CoPilot programs, generating random input for each program, and checking consensus on output between the interpreter and C programs.
+Further assurance is accomplished with bounded model checking by showing equivalence between the code from the two compiler backends over a user defined number of steps.
+Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high assurance components.
+
+As seen in this paper, contracts can be stateful and even higher-order \cite{10.1145/583852.581484}. 
+Verification aware languages compile contracts to runtime monitors \cite{10.1007/978-3-642-28869-2_11}, or omit them when they can statically prove that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+Dafny is a notable verification aware language that provides a contract language, an implementation language, a proof engine to show if an implementation adheres to the contract, and backend compilers for several different target platforms \cite{dafny}.
+The Dafny contracts provide the verification conditions to prove out the implementation code and are not compiled.
+Compared to the work discussed in this paper, the test contracts are more similar to the Dafny contracts as they are also not compiled and only serve to prove properties of the code contracts.
+In that regard, the code contracts are the implementation, but unlike Dafny, \ckml\ provide a verified compilation path to binaries.

### 2022-09-16 paper/case-study.tex
Add text regarding who adds requirements and remove redundancy.

--- 2021-08-01 paper/case-study.tex
@@
-In this section, we outline the application of the BriefCASE tool towards the development of a UAV surveillance system as a part of the DARPA CASE program.
-The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides engineers with early validation of their model; produces early information about cyber-vulnerabilities; provides effective mitigations; and handles a complex real-world message format, namely  the UxAS message format.
+In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
+The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides designers with early validation of their model; produces early information about cyber-vulnerabilities; provides effective mitigations; and handles a complex real-world message format, namely  the UxAS message format.
 
@@
 Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
-We perform a formal analysis using the AGREE tool (which is integrated with the BriefCASE environment) to verify that the model satisfies the contracts.
+We perform a formal analysis using the \agr\ to verify that the model satisfies the contracts.
 For the initial version of our design, the verification passes.
@@
 Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.
-In BriefCASE, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
-The tools generate a list of new requirements corresponding to cyber vulnerabilities found in the design,
-and we need to satisfy these requirements by modifying the design.
+In \brfcs, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
+As the designers of the system, we reviewed the list of potential vulnerabilities, and then added new requirements to the system model.
 %  To satisfy these requirements we need to mitigate the vulnerabilities discovered by the analysis by modifying the design.
 %
-For example, because we annotated the open-source UxAS component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}), the cyber analysis tools generate requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) will not impact other processes.
+For example, the vulnerbility analysis noted that UxAS as an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
 
-In total, seven cyber requirements are generated and imported into our model.
+In total, we added seven cyber requirements from the vulnerability analysis into our model.
 These include four \textit{wellformedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
-Requirements are imported into the model as goals in a Resolute assurance case.
-Because we can run Resolute at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
+Requirements are imported into the model as goals in th assurance case \cite{resolute-destion}.
+Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
 
-The intent of the \textit{wellformedness} requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
-In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
-By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
-The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
+The intent of the \textit{wellformedness} requirements is to prevent
+malformed messages from causing a buffer overrun or code injection
+attack.  In the UAV design, such messages are most likely to originate
+from a remote source or the uncontrolled UxAS component.  By placing
+filters on the connections upstream of mission-critical components,
+such attacks could be mitigated.  The Filter transform is therefore
+applied for each wellformedness requirement, inserting filter
+components on the incoming and outgoing UxAS connections.
 
-The filter behavior for each component is specified in AGREE.
-This enables formal verification within the modeling environment, and also provides a means for synthesizing the component implementation in a provably correct manner using the SPLAT tool.
-Because SPLAT is integrated with BriefCASE, the proof it emits when synthesizing component code is used as evidence in the Resolute goal for the corresponding mitigation.
-When Resolute evaluates whether such a goal is supported by evidence, it checks for the existence of the synthesis proof in addition to verifying that the architecture is correct.
-
-The AGREE filter policies for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
+The filter code contracts for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
 For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
@@
 The Monitor transform is applied for this class of mitigation.
-In general, monitors observe a channel and compare its contents against a reference signal or constant.
-The monitor policy specifies acceptable comparisons, and if violated, the monitor sends out an alert.
-A monitor can choose to \textit{gate} the observed signal, in which case it also acts as a special kind of filter and drops the message if the policy is violated.
-The \textit{monitoring} requirements drive two transforms.
-The first adds a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
-The second adds a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
-The Geofence Monitor is a gated monitor; it prevents the observed Automation Response message from reaching the Waypoint Manager.
-%
-Similar to the filters, the monitor policies are specified in AGREE. For example, 
-%The Response Monitor specification is shown in \figref{fig:response-monitor}, and 
-the Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.
+We add a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+And we add a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+The Geofence Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
+The Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.
 
@@
 mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
-But we also want to protect against a compromised Ground Station transmitting wellformed, but malicious commands.
-The final cyber requirement is mitigated by the Attestation transform~\cite{attestation-copland} that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
-The Attestation Manager is implemented in CakeML and automatically inserted into the application code base by BriefCASE.
-Because the Attestation Gate acts as a filter, the transform automatically generates its complete AGREE specification.%, as shown in \figref{fig:attestation}.
-
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=1\columnwidth]{figs/attestation.png}
-% 	\caption{Atestation specification.} 
-% 	\label{fig:attestation} 
-% \end{figure}
+But we also want to protect against a compromised ground station transmitting well-formed, but malicious, commands.
+The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager~\cite{attestation-copland}.
+The Attestation Manager \ckml\ implementation is automatically inserted into the application code base by \brfcs.
+The Attestation Gate does require a code contract that is completely generated by the transform since it has a very specialized behavior as defined by the transform.
 
 After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
-The components in green were added to the model by way of an automated BriefCASE transform and are critical for mitigating cyber attacks.
-We formally verify the model with AGREE to show that all of the component contracts are satisfied, including the new contracts introduced during the model transformations.
-Because it is imperative that these high-assurance component implementations are correct,
-we run the SPLAT tool to produce provably-correct code.  
+The green shaded components were added by the transform with the exception of the UxAS component.
+The added components are critical for mitigating cyber attacks.
+After unit testing the monitors, we formally verify the model with \agr\ to show that all of the component contracts are satisfied.
+We then run \splt\ to produce provably-correct code.
 The synthesized code is output to a directory in the build file system with the location specified for each component in the model.
@@
 
-Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by SPLAT or other means, the system can be built and deployed.
+Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by \splt\ or other means, the system can be built and deployed.
 We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
-HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication consistent with the AADL computational model.
+% HAMR translates the AADL system model to code that implements threading infrastructure and inter-component communication consistent with the AADL computational model.
 %
-HAMR then compiles the software to execute on seL4~\cite{sel4-2009}, a verified capability-based microkernel (accompanied by formal proof of spatial isolation properties down to the binary level).
+HAMR then compiles the software to execute on seL4.
 
@@
 A third party evaluator repeated the tests on the hardened system and % were able to
-demonstrated that our mitigations were successful.
+demonstrated that our mitigation strategy was successful.
 %The following scenarios were exercised (status messages from the high-assurance components in the hardened system are in \figref{fig:mitigation-output}):

### 2022-09-17 cover-letter/resubmit-letter.tex
Add letter summarizing review responses

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the first version of our paper.
We have carefully studied that feedback, made appropriate revisions, and are hopeful to have responded to the reviewer comments in a satisfactory way.
We feel that those comments have resulted in a much improved submission.

To aid in reviewing this new manuscript, I am including here a detailed summary of the changes in this new submission.
\begin{itemize}
  \item We revised the introduction to make clear the problem being solved relative to existing work and state the novelty of our contribution. It has also been expanded to address upfront the other requests, highlighted by the editor, in the prior review to (1) describe the dependency on the chosen language and tools; (2) define the limitations of the approach, and (3) discuss how this approach could be extended to other domains.
  \item We added a new section summarizing BriefCASE, with its many tools, and its general workflow. That section additionally makes clear the parts of BriefCASE that relate to our contributions in the paper.
  \item We added a new workflow diagram to the section discussing the illustrative example to show the inputs, outputs, existing adopted tools, and new tools that are part of the contribution of this work. We also removed from the example anything in BriefCASE that is not directly related to the contributions discussed in this submission.
  \item We have grouped into a section titled "Background" everything relating to the AGREE tool as it is an existing tool that we leverage as part of our solution to the synthesis problem.
  \item We have revised the case study to identify roles in identifying and adding requirements.
  \item We have completely revised the related work to contrast our solution to synthesis relative to the existing solutions with an effort to highlight that we have solved the problems identified in the introduction. We have also added new related work as suggested by the reviewers and other related work that we have became aware of since submitting our original paper.
  \item We have revised the conclusion to be specific in the planned and ongoing future work as requested by the reviewers. In particular, we have detailed how we plan to fully verify the transformation from a code contract to CakeML and reported on some early results in that regard. We have also referenced ongoing useability studies with Collins designers.
  \item We have made several edits through the document for consistency and to address other specific reviewer comments.
\end{itemize}

I hope that in reviewing this new submission that satisfied with how we have addressed the feedback in the prior reviews.
We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2022-09-17 paper/abstract.tex
start on revising first part of paper

--- 2022-09-15 paper/abstract.tex
@@
 Cyber-physical systems, such as avionics, must be tolerant to
-cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
-\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
-This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
-Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
-A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
-Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
-Implementations for these
-high-assurance components are directly synthesized from their code
-contracts and are backed up by proofs showing that high-level
-specifications map in a meaning-preserving way to code.
-The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
-route planning.  The case study
-demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
-engineering.
+cyber-attacks in the same way they are tolerant to random faults: they
+must gracefully recover, or safely shut down, as requirements dictate.
+We have developed a workflow for creating, and inserting,
+high-assurance components implementing cyber-resiliency into a
+model-based systems engineering environment.  Example high-assurance
+components are filters, which guard against malformed input, and
+runtime monitors, which guard against spoofing and other malicious
+behavior. A formal specification in the form of a \emph{code contract}
+defines each high-assurance component and is developed with the
+support of \emph{test contracts} for testing the specified behavior.
+Once tested, model checking used to verify that the added
+high-assurance component indeed addresses system-level cyber
+requirements.  Implementations for these high-assurance components are
+directly synthesized from their code contracts and are backed up by
+proofs showing that high-level specifications map in a
+semantics-preserving way to code generated by a verified compiler.  We
+report on a case study that cyber-hardened a UAV system by inserting
+high assurance components to harden the open source Air Force Research
+Laboratory's OpenUxAS services for route planning.  The case study
+demonstrates that synthesizing correct implementations from code
+contracts is feasible in real-world systems engineering.

### 2022-09-17 paper/intro.tex
start on revising first part of paper

@@
-\egm{Add to introduction and related work, that what is done at Galois called Copilot.
-Its a stream processing langague that generates code from the specification.
-Copilot to C.
-Effectively CodeGen from Lustre.
-Bounds how far needed to look in past for any value.}
-In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
-Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
-Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
+\egm{Add to introduction and related work, that what is done at
+Galois called Copilot. Its a stream processing langague that generates
+code from the specification. Copilot to C. Effectively CodeGen from
+Lustre. Bounds how far needed to look in past for any value.}
 
-Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
-They instead rely on process-oriented checklists and guidelines.
-Cyber vulnerabilities are often discovered during penetration testing late in the development process; or worse yet, they may be discovered only after the product has been fielded, necessitating extremely
-expensive and time-consuming remediation. This is not a sustainable development model.
+In recent years, aerospace stakeholders have realized that avionics
+systems are subject to possible cyber-attacks just like other
+cyber-physical systems.
+%
+\footnote{This work was funded in part by the
+Defense Advanced Research Projects Agency (DARPA).  The views
+expressed are those of the authors and do not reflect the official
+policy or position of DARPA or the U.S. Government.}
+%
+Thus, in addition to being fault-tolerant, safety-critical avionics
+systems must also be {\em cyber-resilient}.  Cyber-resiliency means
+that the system is tolerant to cyber-attacks just as safety-critical
+systems are tolerant to random faults: they recover and continue to
+execute their mission function, or safely shut down, as requirements
+dictate.
 
-\egm{Remove DARPA junk.}
-The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).
-The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S.
-Government.}
-One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.
-BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
+Unfortunately, systems engineers are currently given few development
+tools to help answer even basic questions about potential
+vulnerabilities and ways to mitigate vulnerabilities.  They instead
+rely on process-oriented checklists and guidelines.  Cyber
+vulnerabilities are often discovered during penetration testing late
+in the development process; or worse yet, they may be discovered only
+after the product has been fielded, necessitating extremely expensive
+and time-consuming remediation. This is not a sustainable development
+model.
 
-This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
-These \egm{make clear what these refers to} are referred to as \emph{transforms} by \brfcs.
-Two cyber-resiliency transformations are discussed: (1) the insertion
-of a filter to prevent malformed data from a malicious actor from
-being propagated to downstream components, and (2) the insertion of a
-monitor to detect (and alert) unexpected behaviors arising from
-untrusted components.
-\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.
+%% The DARPA Cyber Assured Systems Engineering (CASE) project is targeted
+%% at developing tools for design, analysis, and verification that enable
+%% systems engineers to {\em design-in} cyber-resiliency for complex
+%% cyber-physical systems.
 
-A code contract is a formal specification in the Assume Guarantee Reasoning
-Environment (\agr) language.
+We have been developing the {\em BriefCASE} toolsuite to address this need.
+\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
+built in the Open Source AADL\footnote{AADL is the acronym for
+Architecture Analysis and Design Language~\cite{aadl}.}  Tool
+Environment (OSATE) to add new design, analysis, and code generation
+capabilities for building cyber-resilient systems.
+
+In this paper we describe how \brfcs\ facilitates inserting,
+specifying, testing, and synthesizing high assurance components into a
+system to improve its cyber-resiliency.  The main organizing concept
+is that of an architecture-to-architecture \emph{security-improving
+transform}, achieved via the insertion of a new architectural
+component aimed at mitigating a cyber-vulnerability.  We describe two
+cyber-resiliency transformations in this paper: (1) the insertion of a
+filter to prevent malformed data from a malicious actor being
+propagated to downstream components, and (2) the insertion of a
+runtime monitor to detect (and alert) unexpected behaviors arising
+from untrusted components.
+
+A code contract is a formal specification in the Assume Guarantee
+Reasoning Environment (\agr) language.
 \agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
-The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
-These contracts are unit tested for correctness with \emph{test contracts}.
-Test contracts define test scenarios to be implemented by the code contract under test. 
+The code contract language is Turing complete allowing the designer to
+specify arbitrarily complex behavior.  These contracts are unit tested
+for correctness with \emph{test contracts}.  Test contracts define
+test scenarios to be implemented by the code contract under test.
 \agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
@@
 
-Another novel aspect of the approach discussed in this manuscript is the synthesis of the code contracts for the high-assurance components to \ckml, a verified
-compiler implementation for the functional programming language
-ML \cite{cakeml}. 
-This manuscript describes in detail the synthesis path
-from code contracts to \ckml\ code, providing a formal framework in
-which to argue correctness. \ckml\ then provides a verified
-compilation path to several different target binaries (and also
-proving that the meaning of the \ckml\ source code is exactly
-preserved in the final binaries). 
-A key contribution here is that the code contract semantics are defined in such a way that the \agr\ verification 
-results for the code contract hold for the deployed component, \ie, the component will detect
-and prevent the indicated cyber-vulnerabilities over all possible
-finite inputs. Preliminary work has shown how to lift this result to
-infinite input traces as these systems are inherently reactive and
-intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+Another novel aspect of our workthe approach discussed in this manuscript is
+the synthesis of the code contracts for the high-assurance components
+to \ckml, a verified compiler implementation for the functional
+programming language ML \cite{cakeml}.  This manuscript describes in
+detail the synthesis path from code contracts to \ckml\ code,
+providing a formal framework in which to argue correctness. \ckml\
+then provides a verified compilation path to several different target
+binaries (and also proving that the meaning of the \ckml\ source code
+is exactly preserved in the final binaries).  A key contribution here
+is that the code contract semantics are defined in such a way that
+the \agr\ verification results for the code contract hold for the
+deployed component, \ie, the component will detect and prevent the
+indicated cyber-vulnerabilities over all possible finite
+inputs. Preliminary work has shown how to lift this result to infinite
+input traces as these systems are inherently reactive and intended to
+run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
 
-The manuscript further details a 
-case study applying these transformations with \brfcs\ to an Unmanned Aerial
-Vehicle (UAV) system that uses the Air Force Research Laboratory's
-OpenUxAS services for route planning.
-OpenUxAS, as an open source product, is considered \emph{untrusted}.
-The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
-Here the transforms add filters to guard
-against malformed input and monitors to guard against malicious flight plans from OpenUxAS. The case study
-system is complex and shows
-the viability of the approach in potential full-scale industrial
-design.
+The manuscript further details a case study applying these
+transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV)
+system that uses the Air Force Research Laboratory's OpenUxAS services
+for route planning.  OpenUxAS, as an open source product, is
+considered \emph{untrusted}.  The UAV system is thus transformed to be
+resilient to malicious behavior that may arise in the untrusted
+component.  Here the transforms add filters to guard against malformed
+input and monitors to guard against malicious flight plans from
+OpenUxAS. The case study system is complex and shows the viability of
+the approach in potential full-scale industrial design.
 
@@
 Our approach currently applies to common cyber-vulnerabilities, such
-as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
-side-channel attacks and denial of service are not yet dealt with in
-our work.  Here, we do not report on the invention of a new type of
-high-assurance component in terms of capability; instead, our contribution
-is in the automated synthesis of security-improving components from
-formal specifications and a means to show that the synthesis is
-correct.
+as overflow, lack of input validation, and supply-chain issues;
+however, other cyber-vulnerabilities such as side-channel attacks and
+denial of service are not yet dealt with in our work.  Here, we do not
+report on the invention of a new type of high-assurance component in
+terms of capability; instead, our contribution is in the automated
+synthesis of security-improving components from formal specifications
+and a means to show that the synthesis is correct.
 
-We now give a summary of the contributions detailed in this manuscript.
+We now give a summary of the contributions detailed in this
+manuscript.
 \begin{compactitem}

### 2022-09-17 paper/case-study.tex
Fix minor typos

@@
 Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
-We perform a formal analysis using the \agr\ to verify that the model satisfies the contracts.
+We perform a formal analysis with \agr\ to verify that the model satisfies the contracts.
 For the initial version of our design, the verification passes.
@@
 Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.
+
 In \brfcs, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
@@
 %
-For example, the vulnerbility analysis noted that UxAS as an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
+For example, the vulnerbility analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
 
 In total, we added seven cyber requirements from the vulnerability analysis into our model.
-These include four \textit{wellformedness} requirements, two requirements for \textit{monitoring} the behavior of the open-source UxAS component, and an \textit{attestation} requirement for ensuring the Ground Station software has not been tampered with.
-Requirements are imported into the model as goals in th assurance case \cite{resolute-destion}.
+These include four wellformedness requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the Ground Station software has not been tampered with.
+Requirements are imported into the model as goals in the assurance case.
 Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
 
-The intent of the \textit{wellformedness} requirements is to prevent
+The intent of the wellformedness requirements is to prevent
 malformed messages from causing a buffer overrun or code injection
@@
 The Geofence Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
-The Geofence Monitor specification is shown in \figref{fig:geofence-monitor}.
+The Geofence Monitor code contract is shown in \figref{fig:geofence-monitor}.
 
@@
 But we also want to protect against a compromised ground station transmitting well-formed, but malicious, commands.
-The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager~\cite{attestation-copland}.
+The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
 The Attestation Manager \ckml\ implementation is automatically inserted into the application code base by \brfcs.
-The Attestation Gate does require a code contract that is completely generated by the transform since it has a very specialized behavior as defined by the transform.
+The code contract for the Attestation Gate is completely generated by the transform since its behavior is specific to attestation.
 
 After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
-The green shaded components were added by the transform with the exception of the UxAS component.
-The added components are critical for mitigating cyber attacks.
+The shaded components (green) were added by the transform with the exception of the UxAS component (red).
+
 After unit testing the monitors, we formally verify the model with \agr\ to show that all of the component contracts are satisfied.
@@
 When this occurs, UxAS is unable to generate an Automation Response.
-This vulnerability is mitigated by inserting a Response Monitor that checks to see if UxAS outputs an Automation Response message shortly after receiving an Automation Request.
-For this scenario, we chose for the monitor to output a status message, which would then be received by the Ground Station and an appropriate action taken.
+The response monitor detected and signaled the Ground Station when UxAS failed to respond to the request.
 

### 2022-09-17 paper/high-assurance-contracts.tex
Fix minor typos

@@
 Generally, AGREE contracts do not describe the computation that a
-component performs. This is entirely by design: AGREE is intended to
+component performs. This is entirely by design: AGREE is intended to be declarative and
 reason about component behavior solely at the specification
-level. However, the syntax of AGREE provides enough expressiveness to
+level. It defines what a component computes and not how it computes it. However, the syntax of AGREE provides enough expressiveness to
 support the notion of a \emph{code contract}: a contract from which an
@@
 \subsubsection*{Imperative code contracts.}
-\egm{Remove.
-Replace with the informal presentation of temporal squashing \emph{or} remove entirely.}
 
@@
 
-As previously discussed, transformations carried out in the BriefCASE
-interface have created a cyber-hardened system, protecting against
-malicious AI behavior by the insertion of a high-assurance filter and
-monitor between the AI and the WM (see \figref{fig:hardened}). The
-filter protects the WM from malformed data and the monitor protects
-the WM from (possibly malicious) spontaneous or delayed responses.
-These components were added, one at a time, by
-\begin{enumerate}
-  \item selecting the connection in the model where the
-    component is to appear,
-  \item choosing the appropriate transformation, and
-  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
-\end{enumerate}
-
-%% The system designer provides configuration parameters for the
-%% transformation in a dialogue box.  The dialogue box for adding a
-%% filter is shown in \figref{fig:dialogue}.  All high-assurance
-%% components rely on a \emph{policy} to define behavior as seen in the
-%% last field of the dialogue box.  A filter policy defines well-formed
-%% data while a monitor policy defines an invariant over inputs and
-%% outputs.  These policies can be stated directly in the wizard, or they
-%% can be left blank and added later to the AGREE contract generated by
-%% the transformation.
-
-BriefCASE creates all the needed AADL for the new high-assurance
-component, and its connections in the system implementation, as part
-of the transformation.  It also creates a default code contract with
-appropriate output guarantees leaving only the component behavior
-(policy) to be specified.
-
 Code contracts for the filter and monitor are shown in
@@
 detail.
-
-\subsubsection{Monitor specification}
 

### 2022-09-17 paper/overview.tex
Fix minor typos

@@
 A key part of the requirements management in \brfcs\ is a corresponding assurance case that provides evidence, computed directly from the model or by supporting analysis tools, showing how each requirement is satisfied.
-Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project.
+Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project \cite{resolute-destion}.
 
@@
 	\item \textbf{Gate}: block messages in the presence of an alert
-	\item \textbf{Attestation}: measure non-local software to assess trustworthiness
+	\item \textbf{Attestation}: measure non-local software to assess trustworthiness \cite{attestation-copland}
 	\item \textbf{Virtualization}: isolate components
@@
 
-The contributions discussed in this manuscript relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
-This manuscript defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
+The contributions discussed in this paper relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
+This paper defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
 The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model.
 
-This manuscript further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
+This paper further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
 As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing \agr\ model checker, test contracts prove the desired code contract behavior.
@@
 Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
-This manuscript presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \emph{\ckml} code to implement the code contract.
+This paper presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \emph{\ckml} code to implement the code contract.
 \ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.

### 2022-09-17 paper/related-work.tex
Fix minor typos

@@
 %% Specifically requested by the reviewers to be included in this paper!!!
-Another model based approach to secure system architectures follows a pattern similar to \brfcs\ by inserting pre-verified components into the architecture to give certain guarantees \cite{6209212}.
+Another model based approach to secure system architectures follows a pattern similar to \brfcs\ with attestation by inserting pre-verified components into the architecture to give certain guarantees \cite{6209212}.
 In that work, the data connections are annotated with security properties as are the components.
@@
 These led to algorithms for component synthesis from Lustre models based on realizability \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
-The algorithms discover a transition relation that preserve the contract guarantees under all allowed inputs and can be directly transformed into the language intended for the system's implementation.
+The algorithms discover a transition relation that preserves the contract guarantees under all allowed inputs and can be directly transformed into the language intended for the system's implementation.
 Two such algorithms are implemented in the \agr\ verification engine \cite{jkind}.

### 2022-09-17 paper/synthesis.tex
Remove filter alerts sentence
Fix minor typos

@@
 order of expressiveness, these are: regular expressions, contiguity
-types, and AGREE specifications. In earlier work we have described how
+types, and code contracts. In earlier work we have described how
 regular expressions and contiguity types are used as specification
@@
 we extend this \emph{specification oriented} approach to the
-Turing-complete specifications of AGREE.  We give a short review of
-that previous work before going into detail on AGREE.
+Turing-complete specifications of code contracts in \agr.  We give a short review of
+that previous work before going into detail on the synthesis.
 
@@
 assignments in a way that matches the truth value of the predicate in
-the filter guarantee.  The auto-generated AGREE specification raises
-an alert output when the relation is violated.
+the filter guarantee.
 

### 2022-09-17 paper/testing.tex
Fix minor typos

@@
 Missed corner cases is a common problem to all disciplines.
-It is an especially difficult challenge here because there is no direct to exercise code contracts (e.g., there is no runtime \emph{per se}).
+It is an especially difficult challenge here because there is no direct way to exercise code contracts, \eg, there is no runtime \emph{per se}.
 
-A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with with traditional testing.
+A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with traditional testing.
 Such an approach adds extra steps, including test harneses, that are not needed, as it is
@@
 It does this by assuming the test input and guarantying the test output akin to a unit test. 
-The code contract implementation (e.g., the subject under test).
+The code contract is used as the implementation, \eg, it is the test subject.
 The \agr\ analysis then proves if the code contract, given the test inputs, is strong enough to prove the test contract outputs.
@@
 \end{compactitem}
-\figref{fig:test-output} is the \agr\ output for the test, showing that it passes.
+\figref{fig:hardened-certificate}(a) is the \agr\ output for the test showing that it passes.
 
 Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
-They enables designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract reflects the desired behavior with black-box, white-box, and other test coverage metrics.
-These same tests can be synthesized to the backend implementation for traditional testing if for example, the code contract is a model of an existing legacy binary.
+They enable designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
+These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.

### 2022-09-17 paper/conclusion.tex
Fix typo in conclusion

@@
 The actual behavior of such filters, monitors, and other high-assurance components are formally specified by code contracts.
-A code contract is a dialect of the \agr\ language with formal semantics embedded in HOL for formal reasoning.
+A code contract is a dialect of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
 A code contract states assumptions on the inputs and then provides output guarantees that completely define the output in terms of the inputs and the current state of the component.
@@
 In this way, a designer is able to test the behavior of the code contract.
-Once the code contract is tested, the \agr\ verification engine proves whether or note the system, with the added high-assurance components, meets its cyber requirements.
+Once the code contract is tested, the \agr\ verification engine proves whether or not the system, with the added high-assurance components, meets its cyber requirements.
 
 Code contracts are automatically synthesized to \ckml.
-We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transitions function that only considers the current input and state of the component.
+We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transition function that only considers the current input and state of the component.
 That state transition function is then rendered in \ckml, and the \ckml\ compiler compiles it to the target binary providing the proof of equivalence.

### 2022-09-18 paper/conclusion.tex
Updated conclusion with minor edits.

@@
-Safety critical systems, such as those found in avionics, must be cyber-resilient in the same way that they are fault tolerant. 
-Cyber-resiliency means that they are tolerant to cyber-attacks.
-The \brfcs\ environment integrated into OSATE provides systems designers with tools to analyze and address cyber-vulnerabilities in the AADL architecture design of a system.
-We discussed here transformations provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
+Safety critical systems, such as those found in avionics, must be cyber-resilient, i.e. tolerant to cyber-attacks, in the same way that they are fault tolerant. 
+The \brfcs\ environment integrated into OSATE provides systems designers with tools to analyze and address cyber-vulnerabilities in the architectural design of a system modelled using AADL.
+We have discussed transformations provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
 
@@
 The actual behavior of such filters, monitors, and other high-assurance components are formally specified by code contracts.
-A code contract is a dialect of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
-A code contract states assumptions on the inputs and then provides output guarantees that completely define the output in terms of the inputs and the current state of the component.
+A code contract is a (TODO: replace dialect with subset?) dialect of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
+A component's code contract states assumptions on the component inputs and provides guarantees that define the component outputs in terms of the component's inputs and current state.
 The code contract language itself is Turing complete.
@@
 A test contract is a stylized contract that strengthens the input assumptions of a code contract to a single input and either weakens or leaves unchanged the code contract output guarantees.
-\agr\ then proves whether or not the computation expressed in the code contract implements the test contract.
+\agr\ then proves whether or not the computation expressed in the code contract (TODO: satisfies instead of implements?) implements the test contract.
 In this way, a designer is able to test the behavior of the code contract.
@@
 We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transition function that only considers the current input and state of the component.
-That state transition function is then rendered in \ckml, and the \ckml\ compiler compiles it to the target binary providing the proof of equivalence.
+That state transition function is then rendered in \ckml, and the \ckml\ compiler compiles it to the target binary while also generating the proof of equivalence.
 
-We further demonstrated the viability of the approach in real-world industrial design by reporting a case study using the Air Force Research Laboratory's OpenUxAS software.
+We demonstrated the viability of the approach in real-world industrial design by reporting a case study using the Air Force Research Laboratory's OpenUxAS software.
 The study required a range of filters, monitors, and other mitigation components to meet cyber-requirements.
@@
 The reported case study was carried out by those that created \brfcs\ with its several tools. 
-Future work is to conduct another case study involving system designers.
+Future work is to conduct another case study involving system designers unfamiliar with the \brfcs\ toolchain.
 We have started such a study with the design of
@@
 Aside from adding expressiveness to the code contract language, future work is to formally verify every transformation step from the code contract to the final \ckml\ function that is compiled to the target binary.
-The HOL4 embedding of the code contract semantics we discussed here is the starting point.
-The key steps are as follows:
+The HOL4 embedding of the code contract semantics we have discussed is the starting point.
+The key remaining steps are:
 \begin{compactitem}
@@
 We have automated in HOL4 the transformation to remove $\konst{pre}(-)$ in the first step in what we call \emph{temporal squashing}.
-We have additionally manually written proofs for every step for a set of example code contracts.
-From the manual proofs, we have started work on proofs for the first two steps that are universal in that they reason over any input code contract.
-The third step must be generated for each input contract, and the entire process still needs to be mechanized.
+We have additionally written proofs (manually) for every step for a set of example code contracts.
+From the manual proofs, we have started work on proofs for the first two steps.
+Since these two steps are universal to all input code contracts their proofs need to be generated just once (as opposed to generating them for every component contract).
+%are universal in that they reason over any input code contract.
+The proof for the third step must be generated for each input contract. Finally, the entire process needs to be mechanized.
 Other future work includes lifting the proof results to infinite streams.

### 2022-09-18 paper/conclusion.tex
Resolve Junaid's TODOs in conclusion

@@
 The actual behavior of such filters, monitors, and other high-assurance components are formally specified by code contracts.
-A code contract is a (TODO: replace dialect with subset?) dialect of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
+A code contract is a subset of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
 A component's code contract states assumptions on the component inputs and provides guarantees that define the component outputs in terms of the component's inputs and current state.
@@
 A test contract is a stylized contract that strengthens the input assumptions of a code contract to a single input and either weakens or leaves unchanged the code contract output guarantees.
-\agr\ then proves whether or not the computation expressed in the code contract (TODO: satisfies instead of implements?) implements the test contract.
+\agr\ then proves whether or not the computation expressed in the code contract satisfies the test contract.
 In this way, a designer is able to test the behavior of the code contract.

### 2022-09-19 paper/abstract.tex
done with abstract and intro now

@@
-Cyber-physical systems, such as avionics, must be tolerant to
-cyber-attacks in the same way they are tolerant to random faults: they
-must gracefully recover, or safely shut down, as requirements dictate.
-We have developed a workflow for creating, and inserting,
-high-assurance components implementing cyber-resiliency into a
+%% Cyber-physical systems, such as avionics, must be tolerant to
+%% cyber-attacks in the same way they are tolerant to random faults: they
+%% must gracefully recover, or safely shut down, as requirements dictate.
+
+We have developed a workflow for creating and inserting high-assurance
+components that enforce cyber-resiliency requirements into a
 model-based systems engineering environment.  Example high-assurance
@@
 support of \emph{test contracts} for testing the specified behavior.
-Once tested, model checking used to verify that the added
-high-assurance component indeed addresses system-level cyber
-requirements.  Implementations for these high-assurance components are
+Model checking is used to verify that an inserted high-assurance
+component indeed addresses system-level cyber
+requirements. Implementations for high-assurance components are
 directly synthesized from their code contracts and are backed up by

### 2022-09-19 paper/intro.tex
done with abstract and intro now

@@
-\egm{Add to introduction and related work, that what is done at
-Galois called Copilot. Its a stream processing langague that generates
-code from the specification. Copilot to C. Effectively CodeGen from
-Lustre. Bounds how far needed to look in past for any value.}
+%% \egm{Add to introduction and related work, that what is done at
+%% Galois called Copilot. Its a stream processing langague that generates
+%% code from the specification. Copilot to C. Effectively CodeGen from
+%% Lustre. Bounds how far needed to look in past for any value.}
 
@@
 Architecture Analysis and Design Language~\cite{aadl}.}  Tool
-Environment (OSATE) to add new design, analysis, and code generation
-capabilities for building cyber-resilient systems.
+Environment (OSATE) to add new capabilities for building
+cyber-resilient systems.
 
-In this paper we describe how \brfcs\ facilitates inserting,
-specifying, testing, and synthesizing high assurance components into a
+In this paper we describe how \brfcs\ facilitates specifying, testing,
+verifying, and synthesizing high assurance components inserted into a
 system to improve its cyber-resiliency.  The main organizing concept
@@
 component aimed at mitigating a cyber-vulnerability.  We describe two
-cyber-resiliency transformations in this paper: (1) the insertion of a
+such transformations in this paper: (1) the insertion of a
 filter to prevent malformed data from a malicious actor being
@@
 
-A code contract is a formal specification in the Assume Guarantee
-Reasoning Environment (\agr) language.
-\agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
-The code contract language is Turing complete allowing the designer to
-specify arbitrarily complex behavior.  These contracts are unit tested
-for correctness with \emph{test contracts}.  Test contracts define
-test scenarios to be implemented by the code contract under test.
-\agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
-Once the behavior of the code contract is verified, \agr\ proves that---due to the newly
-included high-assurance components---the hardened system meets its
+{\agr}
+\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment} \cite{agree2013}.}
+provides the formal setting for our work.
+{\agr} is implemented in OSATE as an \emph{annex}; hence AGREE
+specifications inherit AADL constructs for describing system
+structure, interconnections, and data. {\agr} uses \emph{contracts}
+expressed in a variant of Past-Time Linear Temporal Logic to formally
+specify input and output properties of components and systems. In this
+paper we introduce the notion of a \emph{code contract}: a class of
+contracts that support the synthesis of implementations. The code
+contract language is Turing complete, allowing the designer to specify
+arbitrarily complex behavior.  These contracts are unit tested for
+correctness with \emph{test contracts}.  Test contracts define test
+scenarios to be implemented by the code contract under test.  {\agr}
+proves whether or not that code contract implementation is correct,
+enabling the designer to iteratively test the component behavior
+inside the \brfcs\ environment.  Once the behavior of the code
+contract is verified, \agr\ proves that---due to the newly included
+high-assurance components---the hardened system meets its
 cyber-resiliency requirements.
 
-Another novel aspect of our workthe approach discussed in this manuscript is
-the synthesis of the code contracts for the high-assurance components
-to \ckml, a verified compiler implementation for the functional
-programming language ML \cite{cakeml}.  This manuscript describes in
-detail the synthesis path from code contracts to \ckml\ code,
-providing a formal framework in which to argue correctness. \ckml\
-then provides a verified compilation path to several different target
-binaries (and also proving that the meaning of the \ckml\ source code
-is exactly preserved in the final binaries).  A key contribution here
-is that the code contract semantics are defined in such a way that
-the \agr\ verification results for the code contract hold for the
-deployed component, \ie, the component will detect and prevent the
-indicated cyber-vulnerabilities over all possible finite
-inputs. Preliminary work has shown how to lift this result to infinite
-input traces as these systems are inherently reactive and intended to
-run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+A novel aspect of our work is the synthesis of {\ckml} code from the
+code contracts for high-assurance components. {\ckml} is a dialect of
+the ML programming language possessing a verified
+compiler \cite{cakeml}.  We give a detailed outline of the synthesis
+path from code contracts to \ckml\ code, providing a formal framework
+in which to argue correctness. \ckml\ then provides a verified
+compilation path to several different target binaries (and also
+proving that the meaning of the \ckml\ source code is exactly
+preserved in the final binaries).  A key contribution here is that the
+code contract semantics are defined in such a way that the \agr\
+verification results for the code contract hold for the deployed
+component, \ie, the component will detect and prevent the indicated
+cyber-vulnerabilities over all possible finite inputs. Preliminary
+work has shown how to extend our work to handle infinite input traces
+as these systems are inherently reactive and intended to run
+forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
 
-The manuscript further details a case study applying these
-transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV)
-system that uses the Air Force Research Laboratory's OpenUxAS services
-for route planning.  OpenUxAS, as an open source product, is
-considered \emph{untrusted}.  The UAV system is thus transformed to be
-resilient to malicious behavior that may arise in the untrusted
-component.  Here the transforms add filters to guard against malformed
-input and monitors to guard against malicious flight plans from
-OpenUxAS. The case study system is complex and shows the viability of
-the approach in potential full-scale industrial design.
+We report on a case study applying these transformations with \brfcs\
+to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force
+Research Laboratory's OpenUxAS services for route planning.  OpenUxAS,
+as an open source product, is considered \emph{untrusted}.  The UAV
+system is thus transformed to be resilient to malicious behavior that
+may arise in the untrusted component.  In particular, filters are
+added to guard against malformed input and a monitor is created to
+guard against malicious flight plans generated from OpenUxAS. The case
+study system is complex and shows the viability of the approach in
+potential full-scale industrial design.
 
-\brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
+{\brfcs} is open source and publicly available \cite{fmide}, as are the
+examples and case study we discussed in this paper \cite{repo,
+phase2, camkes, case}.
+
+\subsubsection*{Contributions}
 Our approach currently applies to common cyber-vulnerabilities, such
@@
 synthesis of security-improving components from formal specifications
-and a means to show that the synthesis is correct.
+and a means to show that the synthesis is correct. We see the
+following as the contributions of this paper:
 
-We now give a summary of the contributions detailed in this
-manuscript.
 \begin{compactitem}
@@
   \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
-  \item A case study that used the implementation of the approach in \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
+  \item A case study that uses \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
 \end{compactitem}
-The rest of this manuscript is organized as follows.
-\secref{sec:overview} is an overview of the \brfcs\ environment and key tools in that environment relative to the contributions of this work. The approach is illustrated by a simple example in
-\secref{sec:example}. Essential background on \agr\ verification and specification are
-presented in \secref{sec:agree}.
-The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are defined in \secref{sec:testing}.
-The synthesis pathway is covered in Section~\ref{sec:synthesis}.
+
+The rest of the paper is organized as follows.
+\secref{sec:overview} is an overview of the {\brfcs} environment.
+Our approach is illustrated by a simple example in
+\secref{sec:example}. Essential background on {\agr} verification and specification are
+presented in \secref{sec:agree}.  The language and semantics of code
+contracts are defined in \secref{sec:code-contracts}, and test
+contracts are described in \secref{sec:testing}.  The synthesis pathway
+is covered in Section~\ref{sec:synthesis}.
 \secref{sec:case-study} discusses the case study.

### 2022-09-19 paper/example.tex
Fix minor typos
Slight cleanup in overview-of-BriefCASE section

@@
   \end{center}
-\caption{Simplified \brfcs\ workflow with the contributions detailed in this manuscript shaded.}
+\caption{Simplified \brfcs\ workflow. Shaded portions are the focus of the present discussion.}
 \label{fig:flowchart}
@@
 
-\figref{fig:flowchart} illustrates a simplified \brfcs\ workflow with the shaded boxes representing the contributions discussed in this manuscript: test contracts, code contracts, and SPLAT synthesis.
+\figref{fig:flowchart} illustrates a simplified \brfcs\ workflow with the shaded boxes representing our focus: test contracts, code contracts, and SPLAT synthesis.
 The unshaded boxes are the existing tools that generate evidence for the assurance case.
@@
 and automated control of a UAV. A picture of the architectural
-model for SW is in \figref{fig:example}. SW is loosely based on one of
-the case studies in Section~\ref{sec:case-study}.  The source for the
+model for SW is in \figref{fig:example}. SW is loosely based on
+the case study in \secref{sec:case-study}.  The source for the
 entire model is found at \cite{repo}.
@@
 High-assurance components are automatically synthesized by \splt\ from
-the code contracts to equivalent programs in the \ckml\
-language.
-In other words, for any set of input streams that meet the
-component's contract assumptions, the output streams produced from
-the synthesized \ckml\ code exactly match those defined by the contract guarantees.
-As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the results equally apply to the resulting binaries from the \ckml\ compiler.
-Synthesis from code contracts is discussed in detail in \secref{sec:synthesis}.
+the code contracts to equivalent programs in the \ckml\ language.  In
+other words, for any set of input streams that meet the component's
+contract assumptions, the output streams produced from the synthesized
+\ckml\ code exactly match those defined by the contract guarantees.
+As such, the \agr\ verification results regarding the code contract
+apply equally to the synthesized \ckml, and thus, the results equally
+apply to the resulting binaries from the \ckml\ compiler.  Synthesis
+from code contracts is discussed in detail in \secref{sec:synthesis}.
 
-Extending the \agr\ verification results to the system composition requires additional guarantees not discussed in this manuscript.
-For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, the HAMR generated scheduler must be certified to follow the dependent data-flow in the design, scheduling windows must be certified to cover worst-case execution times, etc.
-Such additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, scheduled-agree, 9734792}.
+Extending the \agr\ verification results to the system composition
+requires additional guarantees not discussed here.  For
+example, among other things, contracts for non-synthesized components
+must be certified to accurately model their deployed counterparts, the
+HAMR generated scheduler must be certified to follow the dependent
+data-flow in the design, scheduling windows must be certified to cover
+worst-case execution times, \etc.  Such additional guarantees are
+discussed in other works \cite{gearcase2020, dcrypps2019,
+  10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17,
+  sel4-2009, scheduled-agree, 9734792}.

### 2022-09-19 paper/overview.tex
Slight cleanup in overview-of-BriefCASE section

@@
 
-The \brfcs\ OSATE environment, \figref{fig:tool-arch}, provides designers with workflow and tool support for developing products with inherent cyber-resiliency.
-It starts with the development of a baseline AADL model of the system architecture that defines the desired functionality.
-This model can be evaluated using existing tools such as resource usage, information flow, latency analysis, etc. for validation.
+The \brfcs\ OSATE environment, \figref{fig:tool-arch}, provides
+designers with workflow and tool support for developing products with
+inherent cyber-resiliency.  It starts with the development of a
+baseline AADL model of the system architecture that defines the
+desired functionality.  This model can be evaluated on aspects such as
+resource usage, information flow, latency analysis, {\etc} using
+existing tools.
 
-\brfcs\ integrates additional tools into OSATE to analyze, identify, and address cybersecurity vulnerabilities in the baseline model (top of \figref{fig:tool-arch}).
-\brfcs\ provides access to analysis tools that examine the model, report potential cyber vulnerabilities,
-and suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
-Designers add these requirements to the \brfcs\ project requirements repository. 
-A key part of the requirements management in \brfcs\ is a corresponding assurance case that provides evidence, computed directly from the model or by supporting analysis tools, showing how each requirement is satisfied.
-Evidential statements must exist to address the added cyber requirements in order for \brfcs\ to build the assurance case for the project \cite{resolute-destion}.
+\brfcs\ integrates additional tools into OSATE to analyze, identify,
+and address cybersecurity vulnerabilities in the baseline model (top
+of \figref{fig:tool-arch}).  \brfcs\ provides access to analysis tools
+that examine the model, report potential cyber vulnerabilities, and
+suggest requirements for mitigation \cite{dcrypps2019,gearcase2020}.
+Designers add these requirements to the \brfcs\ project requirements
+repository.  A key part of the requirements management in \brfcs\ is a
+corresponding assurance case that provides evidence, computed directly
+from the model or by supporting analysis tools, showing how each
+requirement is satisfied.  Evidential statements must exist to address
+the added cyber requirements in order for \brfcs\ to build the
+assurance case for the project \cite{resolute-destion}.
 
@@
 
-Important to the contributions presented in this paper is the Assume Guarantee Reasoning Environment (\agr) \cite{compositional-analysis-agree,nfm:agree}.
-It is a compositional, assume-guarantee-style, model checker for AADL models.
-\agr\ attempts to prove properties about one layer of the model using properties allocated to sub-components.
-The composition is performed in terms of \emph{contracts} provided for each component and sub-component that state \emph{assumptions} on inputs and \emph{guarantees} on outputs under the assumptions.
-\agr\ verifies the \emph{consistency} of the composition of the component and sub-component contracts meaning that in every connection the source constraints satisfy the destination constraints.
+As mentioned earlier,
+\agr\ \cite{compositional-analysis-agree,nfm:agree} provides the
+formal platform for our work. It bundles together a formal contract
+language with a compositional, assume-guarantee-style, model checker
+for AADL models.  \agr\ attempts to prove properties about one layer
+of the model using properties allocated to sub-components.  The
+composition is performed in terms of \emph{contracts} provided for
+each component and sub-component that state \emph{assumptions} on
+inputs and \emph{guarantees} on outputs under the assumptions.
+\agr\ verifies the \emph{consistency} of the composition of the
+component and sub-component contracts meaning that in every connection
+the source constraints satisfy the destination constraints.
 
-The contributions discussed in this paper relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented.
-This paper defines a \emph{code contract} language that is a subset of the \agr\ language with sequential evaluation semantics as opposed to the existing parallel semantics of \agr.
-The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove consistency in the contract composition of the model.
+The contributions of this paper relate to the aspects of
+\brfcs\ concerning evidence that a high-assurance component meets its
+intended purpose and is properly implemented.  This paper defines a
+\emph{code contract} language that is a subset of the \agr\ language
+with sequential evaluation semantics as opposed to the existing
+parallel semantics of \agr.  The sequential semantics are requisite to
+code synthesis and are defined such that \agr\ is able to still prove
+consistency in the contract composition of the model.
 
-This paper further defines a stylized \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
-As such, the behavior of a high assurance component is defined by the designer in a code contract, and then using the existing \agr\ model checker, test contracts prove the desired code contract behavior.
-Once the designer is satisfied that the code contract is correct, the system composition is verified to conclude that the contract meets its intended purpose.
+This paper further defines a notion of \emph{test contract} for unit
+testing code contracts outside of the system composition to verify
+input to output behavior as well as general properties of the code
+contract.  In practice, the behavior of a high assurance component is
+defined by the designer in a code contract, and then using the
+existing \agr\ model checker, test contracts prove the desired code
+contract behavior.  Once the designer is satisfied that the code
+contract is correct, the system composition is verified to conclude
+that the contract meets the original cyber-requirement.
 
-Showing that a component is properly implemented is accomplished through formal synthesis from the code contract (bottom of \figref{fig:tool-arch}).
-This paper presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \emph{\ckml} code to implement the code contract.
-\ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.
-The \ckml\ compiler itself is verified to generate binary implementations that are equivalent in semantics to the original input program.
-That proof ensures that the resulting binary exactly preserves the meaning of the original code contract and assures that the component is properly implemented.
+Showing that a component is properly implemented is accomplished
+through formal synthesis from the code contract (bottom of
+\figref{fig:tool-arch}).  This paper presents the Semantic Properties
+of Language and Automata Theory (\splt) tool that generates
+\emph{\ckml} code to implement the code contract.  \ckml\ is a
+functional language that includes additional proofs and tools built
+around it \cite{cakeml}.  The \ckml\ compiler itself is verified to
+generate binary implementations that are equivalent in semantics to
+the original input program.  That proof ensures that the resulting
+binary exactly preserves the meaning of the original code contract and
+assures that the component is faithfully implemented.

### 2022-09-19 paper/intro.tex
tidied up

@@
 
-\subsubsection*{Contributions}
-Our approach currently applies to common cyber-vulnerabilities, such
-as overflow, lack of input validation, and supply-chain issues;
-however, other cyber-vulnerabilities such as side-channel attacks and
-denial of service are not yet dealt with in our work.  Here, we do not
-report on the invention of a new type of high-assurance component in
-terms of capability; instead, our contribution is in the automated
-synthesis of security-improving components from formal specifications
-and a means to show that the synthesis is correct. We see the
-following as the contributions of this paper:
+\subsubsection*{Contributions and limitations}
+
+Our work is embedded in a substantial collection of tools, so it is
+important to clarify the research contribution. We see the following
+as the contributions of this paper:
 
 \begin{compactitem}
-  \item The language and semantics of code contracts to specify the behavior of high assurance components.
-  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
-  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
-  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
-  \item A case study that uses \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
+  \item The language and semantics of \emph{code contracts} provide a means
+  to join system-level contract analysis results (arising from model
+  checking) with component implementation correctness results (arising
+  from property-preserving code generation).
+
+\item \emph{Test contracts} as a
+  stylized mechanism to unit test code contracts inside the \brfcs\
+  framework with \agr.
+
+\item A synthesis path from code contracts to \ckml, supported by a formal
+   framework to argue correctness.
+
+\item A formal argument that \agr\ verification of code contracts carry over
+  to the resulting binaries from the \ckml\ compiler.
+
+\item A case
+  study that uses \brfcs\ to add cyber-resiliency to a non-trivial UAV
+  system.
 \end{compactitem}
 
+Our approach currently applies to common cyber-vulnerabilities, such
+as buffer overflow, lack of input validation, and supply-chain issues;
+however, other cyber-vulnerabilities such as side-channel attacks and
+denial of service have not yet been addressed. Moreover, this work
+does not introduce a new type of high-assurance component in terms of
+capability; instead, our contribution is in the automated synthesis of
+security-improving components from formal cyber-resilience
+specifications and a means to show that the synthesis is correct. We
+have not yet tried to apply {\brfcs} to problems outside the realm of
+cyber-resilience; however, the notions of code and test contracts,
+along with the system analysis and verified code generation facilities
+promise to be useful in a wide variety of settings.
+
+\subsubsection*{Organization}
 The rest of the paper is organized as follows.

### 2022-09-19 paper/related-work.tex
Minor edits to related work.

@@
 \brfcs\ has the ability to insert pre-verified components based on cyber-requirements, but does not currently have the ability to annotate connections with security properties or automatically infer where secure components need to be inserted.
+\TODO{not clear where our work has benefits compared to this.}
 
@@
 Code contracts are imperative, by design, so the designer specifies directly how the computation of the high-assurance component takes place thereby avoiding the question of realizability entirely.
-That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishe that.
+That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishes that.
 
@@
 conducted by compositional reasoning in AGREE.
+\TODO{not clear where the CompCert work falls short compared to ours.}
 
-CoPilot is a domain specific language for writing runtime monitors for hard real-time distributed systems embedded in Haskel and requires working Haskel knowledge to write \cite{10.1007/s11334-013-0223-x}.
-The language semantics are defined over streams, there is an interpreter for testing the behavior of the runtime monitors, and there are two backend compilers to generate C code.
-The generated C code is validated by generating random CoPilot programs, generating random input for each program, and checking consensus on output between the interpreter and C programs.
-Further assurance is accomplished with bounded model checking by showing equivalence between the code from the two compiler backends over a user defined number of steps.
-Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high assurance components.
+CoPilot is a domain specific language for writing runtime monitors for
+hard real-time distributed systems embedded in Haskell and requires
+working Haskell knowledge to write \cite{10.1007/s11334-013-0223-x}.
+The language semantics are defined over streams, there is an
+interpreter for testing the behavior of the runtime monitors, and
+there are two backend compilers to generate C code.  The generated C
+code is validated by generating random CoPilot programs, generating
+random input for each program, and checking consensus on output
+between the interpreter and C programs.  Further assurance is
+accomplished with bounded model checking by showing equivalence
+between the code from the two compiler backends over a user defined
+number of steps.  Aside from the HOL4 embedding of the code contract
+semantics and not requiring any HOL4 working knowledge, the work in
+this paper differs in that it is targeting a fully verified path from
+the code contract all the way to the binaries leveraging the \ckml\
+compiler specifically to provide stronger guarantees on the high
+assurance components.
 
-As seen in this paper, contracts can be stateful and even higher-order \cite{10.1145/583852.581484}. 
-Verification aware languages compile contracts to runtime monitors \cite{10.1007/978-3-642-28869-2_11}, or omit them when they can statically prove that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
-Dafny is a notable verification aware language that provides a contract language, an implementation language, a proof engine to show if an implementation adheres to the contract, and backend compilers for several different target platforms \cite{dafny}.
-The Dafny contracts provide the verification conditions to prove out the implementation code and are not compiled.
-Compared to the work discussed in this paper, the test contracts are more similar to the Dafny contracts as they are also not compiled and only serve to prove properties of the code contracts.
-In that regard, the code contracts are the implementation, but unlike Dafny, \ckml\ provide a verified compilation path to binaries.
+As seen in this paper, contracts can be stateful and even
+higher-order \cite{10.1145/583852.581484}.  Verification aware
+languages compile contracts to runtime
+monitors \cite{10.1007/978-3-642-28869-2_11}, or omit them when they
+can statically prove that the code preserves the contract under all
+possible inputs and executions \cite{10.1145/3158139}. Dafny is a
+notable verification aware language that provides a contract language,
+an implementation language, a proof engine to show if an
+implementation adheres to the contract, and backend compilers for
+several different target platforms \cite{dafny}.  Dafny contracts
+provide the verification conditions to prove out the implementation
+code and are not compiled.  Compared to the work discussed in this
+paper, our test contracts are more similar to the Dafny contracts as
+they are also not compiled and only serve to prove properties of the
+code contracts.  In that regard, the code contracts are the
+implementation, but unlike Dafny, \ckml\ provide a verified
+compilation path to binaries.

### 2022-09-19 paper/intro.tex
latest revision

@@
 Our work is embedded in a substantial collection of tools, so it is
-important to clarify the research contribution. We see the following
-as the contributions of this paper:
+important to clarify the research impact. We see the following
+as the distinguishing contributions of this paper:
 
 \begin{compactitem}
-  \item The language and semantics of \emph{code contracts} provide a means
-  to join system-level contract analysis results (arising from model
-  checking) with component implementation correctness results (arising
-  from property-preserving code generation).
+  \item The language and semantics of \emph{code contracts} provide a
+  means to join system-level contract analysis results (arising from
+  model checking) with component implementation correctness results
+  (arising from property-preserving code generation). This provides a
+  strong formal path connecting (A) top-level system cyber-resilience
+  requirements, (B) specifications for high-assurance components inserted
+  to mitigate against specific cyber-resilience problems, and (C) synthesis
+  to a programming language with a verified compiler.
 

### 2022-09-19 paper/related-work.tex
latest revision

@@
 \brfcs\ has the ability to insert pre-verified components based on cyber-requirements, but does not currently have the ability to annotate connections with security properties or automatically infer where secure components need to be inserted.
-\TODO{not clear where our work has benefits compared to this.}
 
@@
 Code contracts are imperative, by design, so the designer specifies directly how the computation of the high-assurance component takes place thereby avoiding the question of realizability entirely.
-That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishes that.
+That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishe that.
 
@@
 conducted by compositional reasoning in AGREE.
-\TODO{not clear where the CompCert work falls short compared to ours.}
 

### 2022-09-19 paper/synthesis.tex
latest revision

@@
 syntax and semantics is applied to extract code contracts and map them
-to equivalent logic functions. In particular, given a code contract with
-functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state
+to equivalent logic functions. In particular, given a code contract
+with functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state
 variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics
-given in Section \ref{sec:code-contracts} can be applied to create a
-function
+given in Section \ref{sec:code-contracts} can be applied to define a
+stream-oriented ``step'' function
+$\konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env}$ defined as
 \[
- \konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env} = \sem{{\Eqs}\cdot G}
+ \konst{strmStep}\;t \; E = \sem{\mathit{Eqs} \cdot G}^E_t
 \]
-Assume ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially
-imperative. Then we can direcctly create a corresponding logic function
+Now suppose ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially
+imperative. Then we can direcctly create a logic function
 \[
@@
 \]
-that takes input and state as parameters and returns an updated state
-and output. Thus the environment $\mathit{Env}$ holding all previous
-values of each variable can be replaced by \emph{parameters}, each of
-which is able to hold only one data element. Assuming
-$\mathit{Env}' = \konst{strmStep}\;t\; \mathit{Env}$, we can formally prove that
+corresponding to \konst{strmStep} that takes input and state as
+parameters and returns an updated state and output. Thus the
+environment $\mathit{Env}$ holding all previous values of each
+variable can be replaced by \emph{parameters}, each of which is able
+to hold only one data element. Assuming $\mathit{E}'
+= \konst{strmStep}\;t\; \mathit{E}$, we can formally prove that
 \[
-\vdash (\mathit{Env}' \vert_{\mathbb{S}},
- \mathit{Env}' \vert_{\mathbb{O}}) =
-\konst{stateStep} (\mathit{Env}\vert_{\mathbb{I}(t)},
-                   \mathit{Env}\vert_{\mathbb{S}(t)})
+\vdash (\mathit{E}' \vert_{\mathbb{S}},
+ \mathit{E}' \vert_{\mathbb{O}}) =
+\konst{stateStep} (\mathit{E}\vert_{\mathbb{I}(t)},
+                   \mathit{E}\vert_{\mathbb{S}(t)})
 \]
 In other words, projecting out the state and output components from
-$\konst{strmStep}\;t\; \mathit{Env}$ equals the result of
+the result of $\konst{strmStep}\;t\; \mathit{E}$ equals the result of
 applying \konst{stateStep} on the input and state components of
-$\mathit{Env}$ at time $t$. This theorem connects the stream semantics
+$\mathit{E}$ at time $t$. This theorem connects the stream semantics
 of the code contract to a pure logic function over the values
@@
 CakeML eco-system to jump from the setting of higher order logic to
-CakeML abstract syntax \cite{cakeml-translator}. Thus we have moved
-from stream semantics to logic function to CakeML evaluation semantics
-in a property-preserving way. Once in CakeML, further necessary
-features can be added, namely imperative variables for storing state
-between component invocations, also invocations of the (verified) foreign
-function interface of CakeML \cite{cakeml-monadic}. Finally, the CakeML compiler can be
+CakeML abstract syntax \cite{cakeml-translator}. In this way we have
+moved from stream semantics to logic function to CakeML evaluation
+semantics in a property-preserving way. Once in CakeML, further
+necessary features can be added, namely imperative variables for
+storing state between component invocations, also invocations of the
+(verified) foreign function interface of
+CakeML \cite{cakeml-monadic}. Finally, the CakeML compiler can be
 applied to the program. It delivers an executable which can be shown
@@
 \item New values for the state variables are computed, in dependency
-  order. At this point variables defined by \emph{followd-by} are
+  order. At this point variables defined by \emph{followed-by} are
   dealt with. Suppose the variable definitions have the following

### 2022-09-19 paper/intro.tex
adding problem definition to intro

--- 2022-09-19 paper/intro.tex
@@
 
-Our work is embedded in a substantial collection of tools, so it is
-important to clarify the research contribution. We see the following
-as the contributions of this paper:
+As indicated in the preceding discussion, {\brfcs} is aimed at
+addressing the lack of tool support for the the problem of enforcing
+and implementing the class of \emph{cyber-resilience} system
+requirements.  Our approach to the problem is embedded in a
+substantial collection of tools, so it is important to clarify the
+research impact. We see the following as the distinguishing
+contributions of this paper:
 
 \begin{compactitem}
-  \item The language and semantics of \emph{code contracts} provide a means
-  to join system-level contract analysis results (arising from model
-  checking) with component implementation correctness results (arising
-  from property-preserving code generation).
+  \item The language and semantics of \emph{code contracts} provide a
+  means to join system-level contract analysis results (arising from
+  model checking) with component implementation correctness results
+  (arising from property-preserving code generation).
+
+\item We provide a strong formal path connecting (A) top-level system cyber-resilience
+  requirements, (B) specifications for high-assurance components
+  inserted to mitigate against specific cyber-resilience problems, and
+  (C) synthesis to a programming language with a verified compiler.
 
@@
 
-\item A synthesis path from code contracts to \ckml, supported by a formal
-   framework to argue correctness.
-
-\item A formal argument that \agr\ verification of code contracts carry over
-  to the resulting binaries from the \ckml\ compiler.
-
-\item A case
-  study that uses \brfcs\ to add cyber-resiliency to a non-trivial UAV
-  system.
+\item A case study that demonstrates the use of \brfcs\ to add cyber-resiliency
+ to a non-trivial UAV system.
 \end{compactitem}
@@
 does not introduce a new type of high-assurance component in terms of
-capability; instead, our contribution is in the automated synthesis of
+capability. Instead, our contribution is in the automated synthesis of
 security-improving components from formal cyber-resilience
-specifications and a means to show that the synthesis is correct. We
-have not yet tried to apply {\brfcs} to problems outside the realm of
-cyber-resilience; however, the notions of code and test contracts,
-along with the system analysis and verified code generation facilities
-promise to be useful in a wide variety of settings.
+specifications, paired with a verification showing that the
+synthesized implementation is correct. We have not yet tried to apply
+{\brfcs} to problems outside the realm of cyber-resilience; however,
+the notions of code and test contracts, along with the system analysis
+and verified code generation facilities promise to be useful in a wide
+variety of settings.
 

### 2022-09-19 paper/related-work.tex
Resolve TODOs in conclusion

--- 2022-09-19 paper/related-work.tex
@@
 The insertion algorithm considers time cost when inserting components to preserve worst-case execution times mandated by the system.
-\brfcs\ has the ability to insert pre-verified components based on cyber-requirements, but does not currently have the ability to annotate connections with security properties or automatically infer where secure components need to be inserted.
-\TODO{not clear where our work has benefits compared to this.}
+In the \brfcs\ tool chain there are tools that can analyze the model annotated with security properties and identify where security vulnerabilities may be located, and there are other tools to transform the model to address those vulnerabilities such as those discussed in this paper.
+
+has the ability to insert pre-verified components based on cyber-requirements, and it does have the ability to automatically infer where secure components need to be inserted based on the security properties of the components and annotations on the data connections.
 
@@
 conducted by compositional reasoning in AGREE.
-\TODO{not clear where the CompCert work falls short compared to ours.}
+It is important to note that we target a wholistic systems approach to cyber-security that is integrated into the AGREE reasoning environment whereas the other work targets more specifically a compiler for Lustre.
 

### 2022-09-19 cover-letter/resubmit-letter.tex
Fix minor typos

@@
   \item We have completely revised the related work to contrast our solution to synthesis relative to the existing solutions with an effort to highlight that we have solved the problems identified in the introduction.
-We have also added new related work as suggested by the reviewers and other related work that we have became aware of since submitting our original paper.
+We have also added new related work as suggested by the reviewers and other related work that we have become aware of since submitting our original paper.
   \item We have revised the conclusion to be specific in the planned and ongoing future work as requested by the reviewers.

### 2022-09-19 paper/intro.tex
Add CASE in footnote
Fix minor typos

@@
 \footnote{This work was funded in part by the
-Defense Advanced Research Projects Agency (DARPA).  The views
+Defense Advanced Research Projects Agency (DARPA) CASE program.  The views
 expressed are those of the authors and do not reflect the official
@@
 {\brfcs} is open source and publicly available \cite{fmide}, as are the
-examples and case study we discussed in this paper \cite{repo,
+examples and case study we discuss in this paper \cite{repo,
 phase2, camkes, case}.
@@
 As indicated in the preceding discussion, {\brfcs} is aimed at
-addressing the lack of tool support for the the problem of enforcing
+addressing the lack of tool support for the problem of enforcing
 and implementing the class of \emph{cyber-resilience} system
 requirements.  Our approach to the problem is embedded in a
-substantial collection of tools, so it is important to clarify the
+substantial collection of tools to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
 research impact. We see the following as the distinguishing
@@
 specifications, paired with a verification showing that the
-synthesized implementation is correct. We have not yet tried to apply
+synthesized implementation is correct in the context of the whole system all the way down to the binary implementation.
+
+We have not yet tried to apply
 {\brfcs} to problems outside the realm of cyber-resilience; however,
@@
 variety of settings.
+Writing code and test contracts does require a working knowledge of propositional logic but not to a degree that would exceed what would be expected in a designer.
+The general synthesis approach in this paper could be used outside of the \brfcs\ environment, but it would only then provide guarantees relative to the complied artifacts and not say anything about the correctness of the system composition.
 

### 2022-12-03 cover-letter/refactor-letter.tex
Revise abstract as per reviewer feedback

\documentclass[10pt]{byu-cs}

\begin{document}
\pagestyle{headings}

\signature{Eric G Mercer\\
           Associate Professor}

\begin{letter}{MoDELS 2021 Special Issue Resubmission}

\date{\today}

\opening{Shiva Nejati and D\'{a}niel Varr\'{o},}

On behalf of my co-authors, I want to thank the reviewers for the very thoughtful feedback on the second version of our paper.
We have carefully studied that feedback, again made appropriate revisions, and are hopeful to have responded to the reviewers' feedback in a satisfactory way.
We feel that the feedback has improved the submission.

To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect.
\begin{itemize}
  \item We rewrote the abstract to be active present tense, define the problem, approach, contributions, and results.
\end{itemize}

I hope this new submission satisfies the reviewers' requests.

We thank you for your consideration and the valuable feedback.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2022-12-03 paper/abstract.tex
Revise abstract as per reviewer feedback

@@
-%% Cyber-physical systems, such as avionics, must be tolerant to
-%% cyber-attacks in the same way they are tolerant to random faults: they
-%% must gracefully recover, or safely shut down, as requirements dictate.
+\newjunk{%
+Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
+%
+As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
+%
+Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
+%
+Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
+%
+We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
+%
+A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
+%
+The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
+%
+We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
+%
+The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
+%
+We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
+%
+Our case study demonstrates that synthesizing correct implementations from code
+contracts is feasible in real-world systems engineering.
+}
 
+\oldjunk{%
 We have developed a workflow for creating and inserting high-assurance
@@
 contracts is feasible in real-world systems engineering.
+}

### 2022-12-05 paper/abstract.tex
Revise the introduction as per reviewer comments

@@
 %
-Our case study demonstrates that synthesizing correct implementations from code
-contracts is feasible in real-world systems engineering.
+Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.<
 }

### 2022-12-05 paper/intro.tex
Revise the introduction as per reviewer comments

@@
 
-We have been developing the {\em BriefCASE} toolsuite to address this need.
+We have been developing the {\em BriefCASE} \oldjunk{toolsuite} toolkit to address this need.
 \brfcs\ is a Model-Based Systems Engineering (MBSE) environment
@@
 cyber-resilient systems. 
-
-In this paper we describe how \brfcs\ facilitates specifying, testing,
+%
+In this paper we describe how it facilitates specifying, testing,
 verifying, and synthesizing high assurance components inserted into a
@@
 is that of an architecture-to-architecture \emph{security-improving
-transform}, achieved via the insertion of a new architectural
-component aimed at mitigating a cyber-vulnerability.  We describe two
+transform}, achieved via the insertion of \oldjunk{a} new architectural
+component\newjunk{s} aimed at mitigating \oldjunk{a} cyber-vulnerabilit\oldjunk{y}\newjunk{ies}.  We describe two
 such transformations in this paper: (1) the insertion of a
@@
 
+\oldjunk{%
 {\agr}
-\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment} \cite{agree2013}.}
 provides the formal setting for our work.
@@
 cyber-resiliency requirements.
+}
 
+\newjunk{%
+The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented. 
+%
+We leverage the existing {\agr}\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,cakeml}.
+%
+Specifically, designers write \emph{contracts} to formally specify input and output properties of the system and its components, and \agr\ proves via model checking whether or not the system composition preserves contract requirements at component inputs and guarantees contract obligations at system outputs.
+%
+In this way, \agr\ provides the formal framework we use to describe the behavior of an inserted high-assurance component, such as a filter or monitor, and then prove if adding such a component to the system improves
+cyber-resiliency.
+}
+
+\newjunk{
+We then use semantics preserving transformations to rewrite contracts into \ckml\ where the \ckml\ compiler then turns such contracts into binaries for the intended target platforms.
+%
+The verified compiler, with the semantics preserving transforms, ensures that the high-assurance component adheres to its contract specification.
+%
+This assurance fulfills the underlying assumption of the \agr\ model checker's proof: a component implementation adheres to its contract.
+}
+
+\newjunk{
+As such, this paper defines a
+\emph{code contract} language that is a subset of the \agr\ language
+but with sequential evaluation semantics as opposed to \agr's existing
+parallel semantics.  The sequential semantics are requisite to
+code synthesis and are defined such that \agr\ is able to still prove the contract composition of the system works.
+This paper further defines a notion of \emph{test contract} for unit
+testing code contracts outside of the system composition to verify
+input to output behavior as well as general properties of the code
+contract.  
+%
+Intuitively, the behavior of a high assurance component is
+defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contracts input assumptions.
+%
+Once the designer is satisfied that the code
+contract is correct, the system composition, with its various contracts, is verified to conclude
+that the system meets its cyber-requirements.
+}
+
+\newjunk{%
+Showing that a component is properly implemented is accomplished
+through formal synthesis from the code contract.  This paper presents the Semantic Properties
+of Language and Automata Theory (\splt) tool that generates
+\ckml\ code to implement the code contract.  \ckml\ is a
+functional language that includes additional proofs and tools built
+around it \cite{cakeml}.  The \ckml\ compiler itself is verified to
+generate binary implementations that are equivalent in semantics to
+the original input program.  That proof ensures that the resulting
+binary exactly preserves the meaning of the original code contract and
+assures that the component is faithfully implemented over all possible finite inputs. 
+Preliminary
+work has shown how to extend our work to handle infinite input
+since most systems of interest are inherently reactive and intended to run
+forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+}
+
+\oldjunk{%
 A novel aspect of our work is the synthesis of {\ckml} code from the
@@
 the ML programming language possessing a verified
-compiler \cite{cakeml}.  We give a detailed outline of the synthesis
+compiler.  We give a detailed outline of the synthesis
 path from code contracts to \ckml\ code, providing a formal framework
@@
 component, \ie, the component will detect and prevent the indicated
-cyber-vulnerabilities over all possible finite inputs. Preliminary
-work has shown how to extend our work to handle infinite input traces
-as these systems are inherently reactive and intended to run
-forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
+cyber-vulnerabilities over all possible finite inputs.
+}
 
@@
 addressing the lack of tool support for the problem of enforcing
-and implementing the class of \emph{cyber-resilience} system
+and implementing the class of cyber-resilience system
 requirements.  Our approach to the problem is embedded in a
@@
 \begin{compactitem}
-  \item The language and semantics of \emph{code contracts} provide a
+  \item The language and semantics of \emph{code contracts} \oldjunk{provide} \newjunk{as} a
   means to join system-level contract analysis results (arising from
@@
 
-\item We provide a strong formal path connecting (A) top-level system cyber-resilience
+  \item \emph{Test contracts} as a
+  stylized mechanism to unit test code contracts inside the \brfcs\
+  framework with \agr.
+
+  \item A strong formal path connecting (A) top-level system cyber-resilience
   requirements, (B) specifications for high-assurance components
@@
   (C) synthesis to a programming language with a verified compiler.
-
-\item \emph{Test contracts} as a
-  stylized mechanism to unit test code contracts inside the \brfcs\
-  framework with \agr.
 
@@
 The conclusions and future work are in \secref{sec:conclusion}.
-
-\begin{comment}
-  BriefCASE incorporates model-level cyber analysis tools (presently
-  GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
-  examine AADL models for potential vulnerabilities and suggest
-  cyber-security requirements to mitigate them.  A library of
-  architectural transforms guides the system engineer through automated
-  model transformations that modify the architecture to address these
-  requirements, possibly inserting new high-assurance components into
-  the system.  Implementations for the new components are synthesized
-  from formal specifications using
-  SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
-  (Semantic Properties for Language and Automata Theory).
-
-  Formal
-  verification that the transformed system model meets its requirements
-  is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
-  Environment).
-  %The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
-  \agr\ is a {\em compositional assume-guarantee} style model checker
-  for AADL models that attempts to prove properties about one layer of
-  an architecture using properties allocated to its subcomponents.
-  Cyber-resilient code implementing the verified model is then
-  automatically generated using the High Assurance Modeling and Rapid
-  Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
-  desired, this code can be targeted to the formally verified seL4
-  secure microkernel~\cite{sel4-2009}.
-\end{comment}

### 2022-12-08 paper/abstract.tex
Revise abstract and intro and sout moved paragraphs in section 2

--- 2022-12-03 paper/abstract.tex
@@
 %
-Our case study demonstrates that synthesizing correct implementations from code
-contracts is feasible in real-world systems engineering.
+Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.
 }

### 2022-12-08 paper/overview.tex
Revise abstract and intro and sout moved paragraphs in section 2

@@
 
+\oldjunk{
 The contributions of this paper relate to the aspects of
@@
 consistency in the contract composition of the model.
+}
 
+\oldjunk{
 This paper further defines a notion of \emph{test contract} for unit
@@
 that the contract meets the original cyber-requirement.
+}
 
+\oldjunk{
 Showing that a component is properly implemented is accomplished
@@
 assures that the component is faithfully implemented.
+}

### 2022-12-08 paper/overview.tex
Add new BriefCASE workflow figure

@@
 	\begin{center}
-	  \includegraphics[width=\textwidth]{./figs/tool-arch.jpg}
+	  \includegraphics[width=\textwidth]{./figs/workflow.png}
   	\end{center}
-	\caption{\brfcs\ integrated tool architecture with key components in the general workflow shown in green.}
+	\caption{The general workflow for \brfcs.}
 	\label{fig:tool-arch}
@@
 functional language that includes additional proofs and tools built
-around it \cite{cakeml}.  The \ckml\ compiler itself is verified to
+around it.  The \ckml\ compiler itself is verified to
 generate binary implementations that are equivalent in semantics to

### 2022-12-09 cover-letter/refactor-letter.tex
Revise the introduction as per reviewer comments
Add new BriefCASE workflow figure
Revise case study with discussion and larger figures

@@
   \item We rewrote the abstract to be active present tense, define the problem, approach, contributions, and results.
+  \item We revised the introduction to pull in the last three paragraphs of section two, and to better funnel readers from the big picture, intuitive overview, to specific details on our tools and contributions.
+  \item We added is discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
 \end{itemize}

### 2022-12-09 paper/case-study.tex
Revise case study with discussion and larger figures

@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
-	\includegraphics[width=1\columnwidth]{figs/sw-initial.png}
+	\includegraphics[width=2\columnwidth]{figs/sw-initial.png}
 	\caption{Initial software architecture.}
 	\label{fig:sw-initial}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
-	\includegraphics[width=1\columnwidth]{figs/hardened-sw.png}
+	\includegraphics[width=2\columnwidth]{figs/hardened-sw.png}
 	\caption{Cyber-resilient software architecture.}
 	\label{fig:hardened-sw}
-\end{figure}
+\end{figure*}
 
@@
 
+\subsection{Discussion}
+
+\newjunk{
+The case study demonstrated the viability of writing code contracts and synthesizing implementations from code contracts in a real-world system.
+It additionally provided insights for improvement.
+We discovered very early that we needed a mechanism to test contracts.
+We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the OSATE environment and an external test platform.
+Test contracts are a direct result of our experience, and they obviate the back and forth by integrating test into the modeling and analysis environment.
+}
+
+\newjunk{
+Writing code contracts for the more complex monitors revealed that although the language is very effective at declarative specifications that define what is being calculated, it is non-intuitive, and sometimes cumbersome, for defining how the computation needs to take place.
+The primary reason for its unease in describing how computation is performed is that it lacks any type of iterative looping structure for arrays.
+Instead, it relies on folding and mapping to describe such computation which is non-intuitive for many types of array updates.
+Growing the expressiveness of the language arround arrays is something to investigate in future work.
+}
+
+\newjunk{
+	There are several, practical, low-level concerns that need to be addressed in the high-level model before synthesis takes place that in our process were only discovered after running the actual system on the designated hardware and software.
+We specifically found it difficult to know the memory bounds for heaps, stacks, and buffers and time bounds for the scheduler \emph{a priori} to code synthesis.
+As a result, we would have to employ a costly iterative approach to set bounds, synthesize code, and then check if anything failed.
+Integrating better memory and running time analysis at the modeling level is something to investigate in future work.
+}
 % \begin{figure}[h]

### 2022-12-19 cover-letter/refactor-letter.tex
Add transition paragraph to the AGREE background.

@@
   \item We revised the introduction to pull in the last three paragraphs of section two, and to better funnel readers from the big picture, intuitive overview, to specific details on our tools and contributions.
-  \item We added is discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
+  \item We added a discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
+  \item We added a paragraph to the end of the example section explaining that our discussion of AGREE early in the paper, of necessity, is a birdseye view of its role in the BriefCASE workflow for the reader that perhaps doesn't want all the detail in the following sections.
+The actual AGREE section of the paper then presents notation, semantics, and usage so that the reader is able to contextualize our primary contributions.
+We also added in that paragraph direct references to the sections in the paper where the detailed contracts for the example system are discussed.
 \end{itemize}

### 2022-12-19 paper/example.tex
Make figures bigger
Add transition paragraph to the AGREE background.

@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{flowchart.png} \\
+      \includegraphics[width=0.8\textwidth]{flowchart.png} \\
     \end{tabular}
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{example.png}
+      \includegraphics[width=\textwidth]{example.png}
     \end{tabular}
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.4]{example-certificate.png} \\
+      \includegraphics[width=1\columnwidth]{example-certificate.png} \\
     \end{tabular}
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.3]{hardened.png}
+      \includegraphics[width=2\columnwidth]{hardened.png}
     \end{tabular}
@@
     \begin{tabular}{c}
-      \includegraphics[scale=0.38]{agree-test-output.png} \\
+      \includegraphics[width=\columnwidth]{agree-test-output.png} \\
       (a) \\ \\
-      \includegraphics[scale=0.4]{hardened-certificate.png} \\
+      \includegraphics[width=\columnwidth]{hardened-certificate.png} \\
       (b)
@@
   sel4-2009, scheduled-agree, 9734792}.
+
+\newjunk{
+Our discussion so far regarding the role of \agr\ in the \brfcs\ workflow in \figref{fig:flowchart} has been high-level and abstract.
+Indeed, we have only given a birdseye view of how \agr\ supports the verification of the system, code contracts, and test contracts, as well as how it provides the proof results that underpin our claim that the synthesized high-assurance components guarantee the \agr\ verification results.
+In the following sections we delve into the details of \agr, including its notation, semantics, and usage, to set the stage for our primary contributions in \secref{sec:code-contracts}, \secref{sec:testing}, and \secref{sec:synthesis}.
+For reference to where we discuss the actual contracts for the example system, \secref{sec:sw-contract} is the exposition of the \agr\ contract for the SW component, \secref{sec:ha-contracts} is the exposition of the code contracts for the high-assurance filter and monitor shown in \figref{fig:hardened}, and \secref{sec:testing} includes an exposition of a test contract for the monitor.
+}

### 2022-12-19 cover-letter/refactor-letter.tex
Revise the example and overview with reviewer comments

@@
   \item We added a discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
+  \item We removed the overview section, created a new figure showing the workflow through the briefcase tool, and integrated that into the example section.
+We added to the example section references to the workflow figure.
+We moved to the beginning of the section a better overview of the example UAV system, and we added text explaining why the system must be cyber-resilient.
+We did not remove sections 3.1 - 3.3 as suggested by the reviewer as these sections discuss the paper contributions at a high-level in the context of our illustrative example.
+We did verify that we had appropriate references to the example, and its various figures, in those sections as suggested by the reviewer.
+  \item We verified that sections 4 through 7 (now sections 3 through 6) have appropriate references to the example system.
   \item We added a paragraph to the end of the example section explaining that our discussion of AGREE early in the paper, of necessity, is a birdseye view of its role in the BriefCASE workflow for the reader that perhaps doesn't want all the detail in the following sections.

### 2022-12-19 paper/intro.tex
Revise the example and overview with reviewer comments

@@
 %
-We leverage the existing {\agr}\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,cakeml}.
+We leverage the existing {\agr}\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,compositional-analysis-agree,nfm:agree,cakeml}.
 %
@@
 Intuitively, the behavior of a high assurance component is
-defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contracts input assumptions.
+defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contract's input assumptions.
 %

### 2022-12-20 paper/intro.tex
Revise based on reviewer input and read

@@
 from untrusted components. 
+\newjunk{
+We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.
+}
 
@@
 \newjunk{%
-The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that a high-assurance component meets its intended purpose and is properly implemented. 
+The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such high-assurance components meet their intended purpose and are properly implemented. 
 %

### 2022-12-20 cover-letter/refactor-letter.tex
Revise the cover-letter

--- 2022-12-09 cover-letter/refactor-letter.tex
@@
 
-To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect.
+To aid in reviewing this new refactored manuscript, I am including here a detailed summary of the changes, and where possible, the submission itself indicates the changes directly to make it easier for the reviewers to inspect by striking out old text and indicating new text in red.
+These annotations are easily removed for the camera version should no further edits be required by the reviewers.
 \begin{itemize}
@@
   \item We revised the introduction to pull in the last three paragraphs of section two, and to better funnel readers from the big picture, intuitive overview, to specific details on our tools and contributions.
-  \item We added is discussion section to the case study in section 8 to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
+  \item We added a discussion section to the case study in section 8 (now section 7) to address lessons learned from our experience. We also enlarged significantly figures 14 and 17.
+  \item We removed the overview section, created a new figure showing the workflow through the BriefCASE tool, and integrated that into the example section.
+We added to the example section references to the workflow figure.
+We moved to the beginning of the section a better overview of the example UAV system, and we added text explaining why the system must be cyber-resilient.
+We did not remove sections 3.1 - 3.3 (now sections 2.2 to 2.4) as suggested by one of our reviewers as these sections discuss our contributions at a high-level in the context of our illustrative example.
+We did verify that we had appropriate references to the example, and its various figures, in those sections as suggested by the same reviewer.
+  \item We verified that sections 4 through 7 (now sections 3 through 6) have appropriate references to the example system.
+  \item We added a paragraph to the end of the example section explaining that our discussion of AGREE early in the paper, of necessity, is a birdseye view of its role in the BriefCASE workflow.
+The actual AGREE section of the paper then presents notation, semantics, and usage so that the reader is able to contextualize our primary contributions.
+We also added in that paragraph, as well as in the presentation of the example in section 2, direct references to the sections, and figures, in the paper where the detailed contracts for the example system are discussed.
+  \item We made various other stylistic changes to improve the flow of the paper.
 \end{itemize}

### 2022-12-21 paper/revised-example.tex
Add source file for revised example

--- 2022-12-19 paper/example.tex
@@
   \end{center}
-\caption{Simplified \brfcs\ workflow. Shaded portions are the focus of the present discussion.}
+\caption{A simplified illustration of the \brfcs\ toolkit with shaded portions being the contributions presented here.}
 \label{fig:flowchart}
 \end{figure*}
+
+\begin{figure*}
+	\begin{center}
+	  \includegraphics[width=\textwidth]{./figs/workflow.png}
+  	\end{center}
+	\caption{The general workflow for \brfcs.}
+	\label{fig:workflow}
+\end{figure*}
+
+The \brfcs\ OSATE environment provides designers with workflow and tool support for developing products with
+inherent cyber-resiliency.
+%
+\figref{fig:flowchart} is a simplified illustration of the toolkit that omits tooling outside of our contributions discussed here (see \cite{dcrypps2019,gearcase2020,resolute-destion,attestation-copland,awas,hamr,sel4-sosp09, sel4-tocs14, sel4-cacm18} for other aspects of \brfcs).
+%
+\newjunk{
+The shaded boxes represent our contribution: test contracts, code contracts, and SPLAT synthesis.
+The unshaded boxes are the existing tools that provide the foundation for our contributions.}
+
+\newjunk{
+The intended \brfcs\ workflow is in \figref{fig:workflow}.
+%
+The workflow starts with a behavioral description of the system and its components on the left.
+%
+These descriptions are contracts that assume properties on inputs and guarantee properties on outputs when the input assumptions hold.
+Once \agr\ verifies the system contract composition, the workflow moves right to the threat analysis, and the contracts of the system and components are modified to reflect that analysis.
+%
+If \agr\ verifies the system contract composition with the changes from the threat analysis, then the workflow moves to the far right for code generation.
+%
+Otherwise it moves immediate right to transforms. 
+Here the designer inserts high-assurance components such as filters and monitors, with their accompanying code contracts, until \agr\ verifies the new system composition to be cyber-resilient to the indicated threats.
+%
+Once \agr\ verifies the new system composition, the workflow moves to code generation, where \splt\ synthesizes implementations for the inserted high-assurance components from their corresponding code contracts.
+}
+
+\newjunk{
+We use this section to explain both figures in more detail through an example.
+%
+The example is a UAV system for remote monitoring, and it is loosely based on
+the case study in \secref{sec:case-study}.  
+%
+The source for the example is at \cite{repo}.
+%
+The intent of the UAV system is to receive a mission from the ground station, plan a path to fly that mission from its current location, and then fly the planned path updating the ground station along the way.
+%
+The system must by cyber-resilient and actively protect itself from malicious actors to ensure it is sending accurate monitoring information back to the ground station and that it actually flies the intended mission.
+}
+
+\subsection{System and Component Contract Implementation}
 
@@
 
-\figref{fig:flowchart} illustrates a simplified \brfcs\ workflow with the shaded boxes representing our focus: test contracts, code contracts, and SPLAT synthesis.
-The unshaded boxes are the existing tools that generate evidence for the assurance case.
-The rest of this section follows an illustrative example through the workflow.
-
-The baseline AADL model is a simple software system, named SW, for route planning
-and automated control of a UAV. A picture of the architectural
-model for SW is in \figref{fig:example}. SW is loosely based on
-the case study in \secref{sec:case-study}.  The source for the
-entire model is found at \cite{repo}.
+\newjunk{
+We focus on the software system for the UAV in this example.
+%
+The AADL architecture model for that system consists of two components as shown in \figref{fig:example}.
+The first is the component for the automatic path planning for a mission (AI). 
+And the second is the component to meter the flight path waypoints to the actual UAV flight controller (WM).
+}
 
 From time to time, SW receives a message on the \texttt{AutomationRequest} input,
-forwarding it to a route planner (AI). The function of AI is to
+forwarding it to the AI. The function of AI is to
 compute the flight path (a list of waypoints) for the UAV, outputting
 the resulting \emph{mission command} on \texttt{AutomationResponse}.
-The waypoint manager (WM) receives the mission command from AI and
+The WM receives the mission command from AI and
 starts the UAV flying the mission by putting an \emph{event} on the
@@
 implementing the system, are modeled with \agr\ contract
-specifications.
+specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
 These contracts assume and guarantee the absence
@@
 
-These initial specifications pass verification, meaning that the
+\newjunk{
+\figref{fig:workflow} shows the intent of this behavioral contract definition of the UAV software system on the left-side of the figure at the arrow marked start.
+The \agr\ verification engine tries to prove if the component composition of the software system is correct. 
+It is an iterative process where the designer updates contracts in the composition until \agr\ proves out the composition.
+}
+For our example, the initial specifications described above pass verification, meaning that the
 contract composition of the components with the system satisfies all
@@
 The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}. 
+\newjunk{\figref{fig:sw} in that section is the \agr\ contract for SW.}
 
@@
 
+\newjunk{
+Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the threat analysis stage.
+The UAV must actively protect itself from malicious actors, but to this point, the designers have only been focusing on its core functional behavior with little consideration for cyber-resilience.
+}
+%
 A cyber-vulnerability analysis identifies the potential of a
 supply chain attack through the AI route planner since its third-party.  Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior.
-They additionally remove all output guarantees from its contract since it is untrusted.
+This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered untrusted.
+Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.
 
-The output from the \agr\ analysis is shown in \figref{fig:example-certificate}.  The red
+The output from that analysis is shown in \figref{fig:example-certificate}.  The red
 exclamation points designate component assumptions or system output
@@
 corresponding counter-example.  The first violation is that the mission command on \texttt{AutomationResponse} from
-the AI to the WM is no longer guaranteed to be well-formed.  The
+the AI to the WM is no longer guaranteed to be well-formed: \newjunk{the red exclamation point on \emph{WM assume: Automation responses are well-formed}.  The
 consequence of that failing input assumption is that the WM outputs
-are no longer guaranteed.  These
-missing guarantees lead to the rest of the failures
+are no longer guaranteed. For a contract, when an input assumption does not hold, then the output may be arbitrary.}
+The now arbitrary output from the WM due to input assumption violations lead to the rest of the failures
 in \figref{fig:example-certificate} as the WM provides the system
@@
 
+\newjunk{As \agr\ proves the system does not meet its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the transforms box.}
 Here is where the shaded regions in \figref{fig:flowchart} come into play.
@@
 Code contracts are discussed in detail in \secref{sec:code-contracts}.
+\newjunk{
+\figref{fig:filter} and \figref{fig:monitor} are the code contracts from that section for the high-assurance filter and monitor inserted into the original system as shown in \figref{fig:hardened}.}
 
@@
 Test contracts are discussed in detail in \secref{sec:testing}.
+\newjunk{
+An example test from that section the the monitor in our example here is shown in \figref{fig:test}.
+}
 
-The \agr\ analysis of the cyber-hardened implementation is shown in
+\newjunk{
+As shown in \figref{fig:workflow}, making the system cyber-resilient is an iterative process of adding high-assurance components, writing their code contracts, testing those contracts with test contracts, and then seeing if they are sufficient to prove out the system.
+}
+Having completed such an iterative for our example system, 
+the \agr\ analysis of the cyber-hardened implementation is shown in
 \figref{fig:hardened-certificate}.
@@
 
+\newjunk{
+As \agr\ now proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
+}
 High-assurance components are automatically synthesized by \splt\ from
@@
 \ckml\ code exactly match those defined by the contract guarantees.
+
+\newjunk{
+The synthesis algorithm is a sequence of semantics preserving transforms that refine the code contract to a step function over the inputs and current component state.
+The function computes the next state and outputs for the component at each scheduling step.
+}
 As such, the \agr\ verification results regarding the code contract
@@
 from code contracts is discussed in detail in \secref{sec:synthesis}.
+\newjunk{
+\figref{fig:filter-cakeml} and \figref{fig:monitor-cakeml} are the implementations from that section for the high-assurance filter and monitor from our example here.
+}
 
@@
 must be certified to accurately model their deployed counterparts, the
-HAMR generated scheduler must be certified to follow the dependent
+generated scheduler must be certified to follow the dependent
 data-flow in the design, scheduling windows must be certified to cover
@@
 \newjunk{
-Our discussion so far regarding the role of \agr\ in the \brfcs\ workflow in \figref{fig:flowchart} has been high-level and abstract.
+Our discussion so far regarding the role of \agr\ in \brfcs\ in \figref{fig:flowchart}, and its expected workflow in \figref{fig:workflow}, has been high-level and abstract.
 Indeed, we have only given a birdseye view of how \agr\ supports the verification of the system, code contracts, and test contracts, as well as how it provides the proof results that underpin our claim that the synthesized high-assurance components guarantee the \agr\ verification results.

### 2022-12-21 paper/revised-example.tex
Correct bibtex entry

@@
   10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17,
-  sel4-2009, scheduled-agree, 9734792}.
+  sel4-2009, nfm:agree, 9734792}.
 

### 2023-02-14 paper/paper.tex
Revise example
Revision to make paper flow with overview
Slight cleanup in overview-of-BriefCASE section
Revise abstract as per reviewer feedback
Add transition paragraph to the AGREE background.
Revise the example and overview with reviewer comments
Inline abstract and introduction

@@
-% TODO: add in the referee tag for the submission
-%\documentclass[global,twocolumn,referee]{svjour}
-\documentclass[global,twocolumn,keeplastbox]{svjour}
+\documentclass[global,twocolumn]{svjour}
 \usepackage{cite}
 \usepackage{amsmath}
-% \usepackage{amsthm}
 \usepackage{amsfonts}
@@
 % Comment macros
-\newcommand{\egm}[1]{\textcolor{orange}{egm: #1}}
+\newcommand{\newjunk}[1]{\textcolor{red}{#1}}
+\newcommand{\oldjunk}[1]{\sout{#1}}
 
@@
 
-% \newcommand{\eqref}[1]{Eq.~\ref{#1}}
-
-% Latex trickery for infix div operator, from stackexchange
-
-\makeatletter
-\newcommand*{\bdiv}{%
-  \nonscript\mskip-\medmuskip\mkern5mu%
-  \mathbin{\operator@font div}\penalty900\mkern5mu%
-  \nonscript\mskip-\medmuskip
-}
-\makeatother
-
 \newcommand{\ie}{\textit{i.e.}}
@@
 
-% \theoremstyle{plain}
-% \newtheorem{theorem}{Theorem}
-% \newtheorem{lemma}{Lemma}
-
-% \theoremstyle{definition}
-% \newtheorem{definition}{Definition}
-
-% \theoremstyle{remark}
-% \newtheorem{remark}{Remark}
-
 \journalname{Software and Systems Modeling}
-
-\newif\ifREVISIONS
-\REVISIONSfalse
 
@@
 \begin{abstract}
-\input{abstract}
+Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
+%
+As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
+%
+Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
+%
+Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
+%
+We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
+%
+A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
+%
+The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
+%
+We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
+%
+The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
+%
+We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
+%
+Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
-\input{intro}
+In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
+%
+\footnote{This work was funded in part by the
+Defense Advanced Research Projects Agency (DARPA) CASE program.
+%
+The views expressed are those of the authors and do not reflect the official
+policy or position of DARPA or the U.S. Government.}
+%
+Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
+%
+Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-criticalsystems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
 
-\section{Overview of BriefCASE}
-\label{sec:overview}
-\input{overview}
+Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
+They instead rely on process-oriented checklists and guidelines.
+%
+Cyber vulnerabilities are often discovered during penetration testing late in the development process;
+%
+or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.
+
+We have been developing the {\em BriefCASE} \oldjunk{toolsuite} toolkit to address this need.
+%
+\brfcs\ is a Model-Based Systems Engineering (MBSE) environment built in the Open Source AADL Tool Environment (OSATE) to add new capabilities for building cyber-resilient systems.
+%
+\footnote{AADL is the acronym for Architecture Analysis and Design Language~\cite{aadl}.}
+%
+In this paper we describe how it facilitates specifying, testing, verifying, and synthesizing high assurance components inserted into a system to improve its cyber-resiliency.
+%
+The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
+%
+We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from untrusted components.
+%
+We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.
+
+The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such high-assurance components meet their intended purpose and are properly implemented.
+%
+We leverage the existing {\agr} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,compositional-analysis-agree,nfm:agree,cakeml}.
+%
+\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.}
+%
+Specifically, designers write \emph{contracts} to formally specify input and output properties of the system and its components, and \agr\ proves via model checking whether or not the system composition preserves contract requirements at component inputs and guarantees contract obligations at system outputs.
+%
+In this way, \agr\ provides the formal framework we use to describe the behavior of an inserted high-assurance component, such as a filter or monitor, and then prove if adding such a component to the system improves cyber-resiliency.
+
+We then use semantics preserving transformations to rewrite contracts into \ckml\ where the \ckml\ compiler then turns such contracts into binaries for the intended target platforms.
+%
+The verified compiler, with the semantics preserving transforms, ensures that the high-assurance component adheres to its contract specification.
+%
+This assurance fulfills the underlying assumption of the \agr\ model checker's proof: a component implementation adheres to its contract.
+
+As such, this paper defines a \emph{code contract} language that is a subset of the \agr\ language but with sequential evaluation semantics as opposed to \agr's existing parallel semantics.
+%
+The sequential semantics are requisite to code synthesis and are defined such that \agr\ is able to still prove the contract composition of the system works.
+%
+This paper further defines a notion of \emph{test contract} for unit testing code contracts outside of the system composition to verify input to output behavior as well as general properties of the code contract.
+%
+Intuitively, the behavior of a high assurance component is defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contract's input assumptions.
+%
+Once the designer is satisfied that the code contract is correct, the system composition, with its various contracts, is verified to conclude that the system meets its cyber-requirements.
+
+Showing that a component is properly implemented is accomplished through formal synthesis from the code contract.
+This paper presents the Semantic Properties of Language and Automata Theory (\splt) tool that generates \ckml\ code to implement the code contract.
+%
+\ckml\ is a functional language that includes additional proofs and tools built around it \cite{cakeml}.
+%
+The \ckml\ compiler itself is verified to generate binary implementations that are equivalent in semantics to the original input program.
+%
+That proof ensures that the resulting binary exactly preserves the meaning of the original code contract and assures that the component is faithfully implemented over all possible finite inputs.
+%
+Preliminary work has shown how to extend our work to handle infinite input since most systems of interest are inherently reactive and intended to run forever~\cite{case-verified-filter,cakeml-space-cost}.
+
+We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
+%
+OpenUxAS, as an open source product, is considered \emph{untrusted}.
+%
+The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
+%
+In particular, filters are added to guard against malformed input and a monitor is created to guard against malicious flight plans generated from OpenUxAS.
+%
+The case study system is complex and shows the viability of the approach in potential full-scale industrial design.
+
+{\brfcs} is open source and publicly available \cite{fmide}, as are the examples and case study we discuss in this paper \cite{repo,
+phase2, camkes, case}.
+
+\subsubsection*{Contributions and limitations}
+
+As indicated in the preceding discussion, {\brfcs} is aimed at addressing the lack of tool support for the problem of enforcing and implementing the class of cyber-resilience system requirements.
+%
+Our approach to the problem is embedded in a substantial collection of tools to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
+research impact.
+%
+We see the following as the distinguishing contributions of this paper:
+\begin{compactitem}
+  \item The language and semantics of \emph{code contracts} as a means to join system-level contract analysis results (arising from model checking) with component implementation correctness results (arising from property-preserving code generation).
+  \item \emph{Test contracts} as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
+  \item A strong formal path connecting (A) top-level system cyber-resilience requirements, (B) specifications for high-assurance components inserted to mitigate against specific cyber-resilience problems, and (C) synthesis to a programming language with a verified compiler.
+  \item A case study that demonstrates the use of \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
+\end{compactitem}
+
+Our approach currently applies to common cyber-vulnerabilities, such as buffer overflow, lack of input validation, and supply-chain issues;
+%
+however, other cyber-vulnerabilities such as side-channel attacks and denial of service have not yet been addressed.
+Moreover, this work does not introduce a new type of high-assurance component in terms of capability.
+%
+Instead, our contribution is in the automated synthesis of security-improving components from formal cyber-resilience specifications, paired with a verification showing that the synthesized implementation is correct in the context of the whole system all the way down to the binary implementation.
+
+We have not yet tried to apply {\brfcs} to problems outside the realm of cyber-resilience;
+%
+however, the notions of code and test contracts, along with the system analysis and verified code generation facilities promise to be useful in a wide variety of settings.
+%
+Writing code and test contracts does require a working knowledge of propositional logic but not to a degree that would exceed what would be expected in a designer.
+%
+The general synthesis approach in this paper could be used outside of the \brfcs\ environment, but it would only then provide guarantees relative to the complied artifacts and not say anything about the correctness of the system composition.
+
+\subsubsection*{Organization}
+The rest of the paper is organized as follows.
+%
+Our approach is illustrated by a simple example in \secref{sec:example}.
+%
+Essential background on {\agr} verification and specification are presented in \secref{sec:agree}.
+%
+The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are described in \secref{sec:testing}.
+%
+The synthesis pathway is covered in Section~\ref{sec:synthesis}.
+%
+\secref{sec:case-study} discusses the case study.
+%
+This is followed by related work in \secref{sec:related-work}.
+%
+The conclusions and future work are in \secref{sec:conclusion}.
 
@@
 \label{sec:example}
-\input{example}
+\input{revised-example}
 
-\section{Compositional reasoning}
+\section{\agr}
 \label{sec:agree}
@@
 
-\section{Syntax and semantics of AGREE}
-\label{agree-semantics}
+\subsection{Syntax}
+\label{sec:agree-semantics}
 \input{agree-semantics}
@@
 \subsection{Formal contract for the example system}
+\label{sec:sw-contract}
 \input{sw-contract}
@@
 
-\ifREVISIONS
-\subsection{Revisions}
-\begin{compactitem}
-  \item Add recent work published after 2018
-\end{compactitem}
-\fi
-
 \section{Conclusion}
@@
 
-% \appendix
-% \section{Contiguity Types}
-% \input{contig}
-
-% \section{System Model and Semantics}
-% \input{system}
-
 \end{document}

### 2023-02-14 paper/paper.tex
Inline the revised example

@@
 
+%%
+% INTRODUCTION
+%%
+
 \section{Introduction} \label{sec:intro}
@@
 
+%%
+% CONTRIBUTIONS AND LIMITATIONS
+%%
+
 \subsubsection*{Contributions and limitations}
@@
 
+%%
+% ORGANIZATION
+%%
+
 \subsubsection*{Organization}
@@
 
+%%
+% ILLUSTRATIVE EXAMPLE
+%%
+
 \section{Illustrative example}
 \label{sec:example}
-\input{revised-example}
+
+\begin{figure*}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[width=0.8\textwidth]{flowchart.png} \\
+    \end{tabular}
+  \end{center}
+\caption{A simplified illustration of the \brfcs\ toolkit with shaded portions being the contributions presented here.}
+\label{fig:flowchart}
+\end{figure*}
+
+\begin{figure*}
+	\begin{center}
+	  \includegraphics[width=\textwidth]{./figs/workflow.png}
+  	\end{center}
+	\caption{The general workflow for \brfcs.}
+	\label{fig:workflow}
+\end{figure*}
+
+The \brfcs\ OSATE environment provides designers with workflow and tool support for developing products with
+inherent cyber-resiliency.
+%
+\figref{fig:flowchart} is a simplified illustration of the toolkit that omits tooling outside of our contributions discussed here (see \cite{dcrypps2019,gearcase2020,resolute-destion,attestation-copland,awas,hamr,sel4-sosp09, sel4-tocs14, sel4-cacm18} for other aspects of \brfcs).
+%
+The shaded boxes represent our contribution: test contracts, code contracts, and SPLAT synthesis.
+The unshaded boxes are the existing tools that provide the foundation for our contributions.
+
+The intended \brfcs\ workflow is in \figref{fig:workflow}.
+%
+The workflow starts with a behavioral description of the system and its components on the left.
+%
+These descriptions are contracts that assume properties on inputs and guarantee properties on outputs when the input assumptions hold.
+Once \agr\ verifies the system contract composition, the workflow moves right to the threat analysis, and the contracts of the system and components are modified to reflect that analysis.
+%
+If \agr\ verifies the system contract composition with the changes from the threat analysis, then the workflow moves to the far right for code generation.
+%
+Otherwise it moves immediate right to transforms.
+Here the designer inserts high-assurance components such as filters and monitors, with their accompanying code contracts, until \agr\ verifies the new system composition to be cyber-resilient to the indicated threats.
+%
+Once \agr\ verifies the new system composition, the workflow moves to code generation, where \splt\ synthesizes implementations for the inserted high-assurance components from their corresponding code contracts.
+
+We use this section to explain both figures in more detail through an example.
+%
+The example is a UAV system for remote monitoring, and it is loosely based on
+the case study in \secref{sec:case-study}.
+%
+The source for the example is at \cite{repo}.
+%
+The intent of the UAV system is to receive a mission from the ground station, plan a path to fly that mission from its current location, and then fly the planned path updating the ground station along the way.
+%
+The system must by cyber-resilient and actively protect itself from malicious actors to ensure it is sending accurate monitoring information back to the ground station and that it actually flies the intended mission.
+
+%%
+% SYSTEM AND COMPONENT CONTRACT IMPLEMENTATION
+%%
+
+\subsection{System and Component Contract Implementation}
+
+\begin{figure*}[h]
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[width=\textwidth]{example.png}
+    \end{tabular}
+  \end{center}
+\caption{Initial design for an automated UAV route planning system.}
+\label{fig:example}
+\end{figure*}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[width=1\columnwidth]{example-certificate.png} \\
+    \end{tabular}
+  \end{center}
+\caption{\agr\ failure certificate for initial design.}
+\label{fig:example-certificate}
+\end{figure}
+
+We focus on the software system for the UAV in this example.
+%
+The AADL architecture model for that system consists of two components as shown in \figref{fig:example}.
+%
+The first is the component for the automatic path planning for a mission (AI).
+%
+And the second is the component to meter the flight path waypoints to the actual UAV flight controller (WM).
+
+From time to time, SW receives a message on the \texttt{AutomationRequest} input, forwarding it to the AI.
+%
+The function of AI is to compute the flight path (a list of waypoints) for the UAV, outputting the resulting \emph{mission command} on \texttt{AutomationResponse}.
+%
+The WM receives the mission command from AI and starts the UAV flying the mission by putting an \emph{event} on the \texttt{Start} output port, continuing to issue waypoints to the UAV flight controller on the \texttt{Waypoint} port as the UAV location updates with messages on \texttt{AirVehicleLocation}.
+%
+The AI component is third-party software and WM is a legacy component.
+
+The expected behavior of the SW system, and the components implementing the system, are modeled with \agr\ contract specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
+%
+These contracts assume and guarantee the absence of any malicious, or unspecified, component behavior.
+%
+More precisely, the contract for SW assumes that its inputs are \emph{well-formed} and that there is never more than one automation request pending at a time.
+%
+Well-formed generally refers to a syntactic restriction on data at a port. For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
+%
+The guarantees for SW ensure that
+\begin{compactitem}
+\item a start coincides with a new waypoint being output;
+\item a start is within one cycle of an automation request and if not, then it persistently alerts;
+\item new waypoints coincide with location updates; and
+\item all outputs are well-formed.
+\end{compactitem}
+
+The contracts for the sub-components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
+%
+The AI contract guarantees it only responds to automation requests and always in the same cycle.
+%
+The legacy WM contract guarantees that
+\begin{compactitem}
+  \item it generates a start from a response from the AI and always in the same cycle;
+  \item the start coincides with a waypoint; and
+  \item any further waypoints coincide with location updates.
+\end{compactitem}
+
+\figref{fig:workflow} shows the intent of this behavioral contract definition of the UAV software system on the left-side of the figure at the arrow marked start.
+%
+The \agr\ verification engine tries to prove if the component composition of the software system is correct.
+%
+It is an iterative process where the designer updates contracts in the composition until \agr\ proves out the composition.
+%
+For our example, the initial specifications described above pass verification, meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
+%
+The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}.
+%
+\figref{fig:sw} in that section is the \agr\ contract for SW.
+
+%%
+% ADDING CYBER REQUIREMENTS
+%%
+
+\subsection{Adding cyber requirements}
+
+Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the threat analysis stage.
+%
+The UAV must actively protect itself from malicious actors, but to this point, the designers have only been focusing on its core functional behavior with little consideration for cyber-resilience.
+%
+A cyber-vulnerability analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
+%
+Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior.
+%
+This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered untrusted.
+%
+Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.
+
+The output from that analysis is shown in \figref{fig:example-certificate}.
+%
+The red exclamation points designate component assumptions or system output guarantees that do not hold, and each failure comes with a corresponding counter-example.
+%
+The first violation is that the mission command on \texttt{AutomationResponse} from the AI to the WM is no longer guaranteed to be well-formed: the red exclamation point on \emph{WM assume: Automation responses are well-formed}.
+%
+The consequence of that failing input assumption is that the WM outputs are no longer guaranteed.
+%
+For a contract, when an input assumption does not hold, then the output may be arbitrary.
+%
+The now arbitrary output from the WM due to input assumption violations lead to the rest of the failures in \figref{fig:example-certificate} as the WM provides the system level outputs.
+
+\begin{figure*}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[width=2\columnwidth]{hardened.png}
+    \end{tabular}
+  \end{center}
+  \caption{Cyber-hardened design for an automated UAV route planning system}
+  \label{fig:hardened}
+\end{figure*}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \includegraphics[width=\columnwidth]{agree-test-output.png} \\
+      (a) \\ \\
+      \includegraphics[width=\columnwidth]{hardened-certificate.png} \\
+      (b)
+    \end{tabular}
+  \end{center}
+  \caption{\agr\ verification certificates. (a) Test contract verification results. (b) Cyber-hardened design verification results.}
+  \label{fig:hardened-certificate}
+\end{figure}
+
+%%
+% ADDING HIGH ASSURANCE COMPONENTS
+%%
+
+\subsection{Adding high assurance components}
+
+As \agr\ proves the system does not meet its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the transforms box.
+%
+Here is where the shaded regions in \figref{fig:flowchart} come into play.
+%
+The system designer now uses \brfcs\ to cyber-harden SW by inserting high-assurance components in the form of a filter and a monitor, as shown in \figref{fig:hardened}.
+%
+These are intended to mitigate and report any observed malicious behavior by AI.
+
+A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if the invariant does not hold.
+%
+\brfcs\ generates an initial code contract for the filter which the designer completes by writing the invariant.
+%
+That invariant is usually based on the existing assumptions made by downstream components that consume the filter output.
+%
+In this example, it is the well-formed property for the automation request.
+
+A monitor captures a relation on data over time and is thus able to reason about the temporal, and other invariant, properties of that data.
+%
+It raises an alert if the specified properties are ever violated.
+%
+As with the filter, \brfcs\ generates an initial code contract that is completed by the designer.
+%
+In this example, that contract states that an automation response can only be generated in conjunction with an automation request;
+%
+and further, that response must come with the request or in the next step after the request.
+%
+Code contracts are discussed in detail in \secref{sec:code-contracts}.
+%
+\figref{fig:filter} and \figref{fig:monitor} are the code contracts from that section for the high-assurance filter and monitor inserted into the original system as shown in \figref{fig:hardened}.
+
+Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
+%
+Test contracts allow the designer to validate code contract behavior with unit testing.
+For example, one of the test contracts for the monitor is that when it sees an automation response one step after a request, then it should not alert, and it should pass the response downstream.
+%
+Test contracts are discussed in detail in \secref{sec:testing}.
+%
+An example test from that section the the monitor in our example here is shown in \figref{fig:test}.
+
+As shown in \figref{fig:workflow}, making the system cyber-resilient is an iterative process of adding high-assurance components, writing their code contracts, testing those contracts with test contracts, and then seeing if they are sufficient to prove out the system.
+%
+Having completed such an iterative for our example system, the \agr\ analysis of the cyber-hardened implementation is shown in \figref{fig:hardened-certificate}.
+%
+\figref{fig:hardened-certificate}(a) is the test contract results and \figref{fig:hardened-certificate}(b) is the hardened system composition.
+%
+Here \agr\ has generated high-level evidence justifying the claim that the high-assurance components meet their intended purposes.
+%
+Having passed \agr\ verification, the high-assurance components are ready to be synthesized.
+
+%%
+% SYNTHESIZING CODE FROM CODE CONTRACTS
+%
+\subsection{Synthesizing code from code contracts}
+
+As \agr\ now proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
+%
+High-assurance components are automatically synthesized by \splt\ from the code contracts to equivalent programs in the \ckml\ language.
+%
+In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized \ckml\ code exactly match those defined by the contract guarantees.
+
+The synthesis algorithm is a sequence of semantics preserving transforms that refine the code contract to a step function over the inputs and current component state.
+The function computes the next state and outputs for the component at each scheduling step.
+%
+As such, the \agr\ verification results regarding the code contract apply equally to the synthesized \ckml, and thus, the results equally apply to the resulting binaries from the \ckml\ compiler.
+%
+Synthesis from code contracts is discussed in detail in \secref{sec:synthesis}.
+%
+\figref{fig:filter-cakeml} and \figref{fig:monitor-cakeml} are the implementations from that section for the high-assurance filter and monitor from our example here.
+
+Extending the \agr\ verification results to the system composition requires additional guarantees not discussed here.
+%
+For example, among other things, contracts for non-synthesized components must be certified to accurately model their deployed counterparts, the generated scheduler must be certified to follow the dependent data-flow in the design, scheduling windows must be certified to cover worst-case execution times, \etc.
+%
+Such additional guarantees are discussed in other works \cite{gearcase2020, dcrypps2019, 10.1007/978-3-030-89159-6_18, 10.1007/978-3-030-89159-6_17, sel4-2009, nfm:agree, 9734792}.
+
+Our discussion so far regarding the role of \agr\ in \brfcs\ in \figref{fig:flowchart}, and its expected workflow in \figref{fig:workflow}, has been high-level and abstract.
+%
+Indeed, we have only given a birdseye view of how \agr\ supports the verification of the system, code contracts, and test contracts, as well as how it provides the proof results that underpin our claim that the synthesized high-assurance components guarantee the \agr\ verification results.
+%
+In the following sections we delve into the details of \agr, including its notation, semantics, and usage, to set the stage for our primary contributions in \secref{sec:code-contracts}, \secref{sec:testing}, and \secref{sec:synthesis}.
+%
+For reference to where we discuss the actual contracts for the example system, \secref{sec:sw-contract} is the exposition of the \agr\ contract for the SW component, \secref{sec:ha-contracts} is the exposition of the code contracts for the high-assurance filter and monitor shown in \figref{fig:hardened}, and \secref{sec:testing} includes an exposition of a test contract for the monitor.
+
+%%
+% AGREE
+%%
 

### 2023-02-14 paper/paper.tex
Inline verification conditions

@@
 \label{sec:agree}
-\input{verification-conditions}
+
+\newcommand{\globally}{\konst{Always}}
+\newcommand{\historically}{\konst{Hist}}
+\newcommand{\assumes}{\ensuremath{A}}
+\newcommand{\guarantees}{\ensuremath{P}}
+\newcommand{\inputs}{\ensuremath{I}}
+\newcommand{\outputs}{\ensuremath{O}}
+\newcommand{\components}{\ensuremath{C}}
+\newcommand{\component}{\ensuremath{c}}
+
+The \agr\ specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
+%
+Lustre semantics are synchronous data-flow where the inputs and outputs of components are streams.
+%
+Informally, streams assign values to expressions at each point in time.
+%
+\agr\ contracts express relationships between input and output streams by assuming a set of allowed input streams and then guaranteeing a corresponding set of allowed output streams.
+
+%%
+% COMPONENTS
+%%
+
+\subsection{Components}
+
+A system implementation is described by its ports, its subcomponents, and the connections between ports on different subcomponents.
+%
+Ports in \agr\ are inherited from AADL.
+%
+There are three kinds of port: \emph{event} ports, \emph{data} ports, and \emph{event data} ports.
+%
+An event port carries no data and only signals occurrence (or not) of the event.
+%
+A data port is a sampled port: the data can be read at anytime.
+%
+An event data port indicates when data is present with an event.
+
+Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize \ie, cannot propagate further.
+%
+Therefore, subcomponent contracts, and thus the top-level model, must be acyclic.
+%
+Cyclic systems are made acyclic by breaking cycles with delays.
+%
+Once the data propagation has stabilized, the model proceeds to the next input data in the input streams.
+%
+It is worth noting that the semantics do not model computation or communication delay: the output of one contract is seen at the input of any downstream contract in the same step of the input data stream.
+
+%%
+% CONTRACTS AND VERIFICATION CONDITIONS
+%%
+
+\subsection{Contracts and verification conditions}
+
+From the system and component contracts, \agr\ generates a set of verification conditions to show that a system's component implementation is correct~\cite{agree2013}.
+%
+The \agr\ model checker is then invoked to prove or disprove the verification conditions.
+%
+The contracts and verification conditions are expressed in \emph{past-time linear temporal logic} (PLTL) \cite{10.1093/jigpal/8.1.55}.
+
+PLTL is a logic enhanced with temporal operators able to reason about the truth values of formulas through time.
+%
+It combines past-time operators that reason over the history of computation up to a point (PTLTL) and future-time operators that reason over the computation that follows that point (LTL).
+%
+Its semantics are defined relative to a point in time $i$, a finite trace of system states $\pi = s_0, s_1, \ldots, s_i$, and some logic formula $f$.
+
+The two PLTL operators necessary for the \agr\ generated verification conditions are $\globally$ that looks forward in time along the trace and $\historically$ (historically) that looks backward in time along the trace.
+%
+These are defined as
+%
+\begin{eqnarray*}
+ (\pi, i) \models \globally(f) & \iff & \forall j \ge i, (\pi, j) \models f \\
+(\pi, i) \models \historically(f) & \iff & \forall 0 \le j \le i, (\pi, j) \models f
+\end{eqnarray*}
+%
+The $\models$-operator is read as \emph{satisfies}.
+%
+A trace at a moment in time satisfies $\globally(f)$ if and only if it satisfies $f$ in the current and all future states of $\pi$.
+%
+$\globally(f)$ is invariant from the current moment into the future and $\historically$ is invariant from the beginning of the trace to the current moment.
+
+A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are subcomponents.
+%
+Assumptions and guarantees only use the past-time operators from PLTL.
+%
+A subcomponent $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
+%
+From the components and their connections, $\mathbb{I}_\component$ is defined to be the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined to be the set of components that provide the output for the system.
+
+A system is \emph{correct} if and only if for all components $c \in C$ the following two verification conditions hold:
+%
+\begin{equation}\label{eq:assumes}
+            \globally(\historically(\assumes \wedge
+            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
+            \implies \assumes_\component)
+\end{equation}
+%
+\begin{equation}\label{eq:guarantees}
+            \globally(\historically(\assumes \wedge
+            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime})
+            \implies \guarantees)
+\end{equation}
+%
+The conditions from \eqref{eq:assumes} verify the input assumptions on each component under the system assumptions and upstream component guarantees.
+%
+They check if the component guarantees and system assumptions are strong enough to imply input assumptions on all immediate downstream components.
+%
+The condition in \eqref{eq:guarantees} checks the output guarantees of the system under the system assumptions and component guarantees that provide the output.
+%
+It checks if the guarantees on components providing primary outputs are strong enough to imply the system guarantees.
+
+If all the verification conditions hold (\agr\ uses $k$-inductive model checking to automatically prove or disprove each generated verification condition), then the system is said to be \emph{correct}, meaning that the system composition meets input assumptions at each input as well as the guarantees on the system output.
+%
+A consequence of this result is that $\globally(\historically(\assumes) \implies \guarantees)$ holds for the system contract.
+%
+Therefore the Liskov substitution principle applies and the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.
+
+The expanded property lists in \figref{fig:example-certificate} and \figref{fig:hardened-certificate} are the results from verifying or disproving the above verification conditions.
+%
+The additional unexpanded results at the bottom of the figures prove \emph{self-consistency} in the contracts.
+%
+A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
+%
+\agr\ generates additional verification conditions that prove each component contract, and the composition of contracts, self-consistent.
+
+%%
+% SYNTAX
+%%
 

### 2023-02-14 paper/paper.tex
Inline semantics and example contract

@@
 %
-The synthesis pathway is covered in Section~\ref{sec:synthesis}.
+The synthesis pathway is covered in \secref{sec:synthesis}.
 %
@@
 \label{sec:agree-semantics}
-\input{agree-semantics}
+
+The syntax of \agr\ is essentially that of quantifier-free first order predicate logic supplemented with a few temporal operators.
+%
+The terms (\emph{e}) are arithmetic expressions built from variables ($v$) and numeric and boolean literals ($c$), while formulas (\emph{b}) are built using logical connectives from atomic formulas (\emph{a}) based on the familiar comparison operators.
+%
+\[
+\begin{array}{rcl}
+e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
+a & ::= & e\; \set{=,<}\; e \\
+b & ::= & v \mid c \mid a \mid \neg b
+            \mid b \; \set{\land,\lor,\imp,\iff}\; b
+\end{array}
+\]
+%
+\agr\ includes a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and formulas.
+To support port handling, \agr\ defines a predicate $\konst{event}(p)$, which tests if an event (or event-data) port $p$ is signaled.
+%
+Lastly, \agr\ supports temporal operators \emph{previously} $\konst{pre}(-)$, \emph{followed-by} $(-) \to (-)$, and historically $\konst{Hist}(-)$.
+
+%%
+% SEMANTICS
+%%
+
+\subsection{Semantics}
+
+The semantics of terms and formulas is defined over \emph{streams of values}.
+%
+Values encompass at least booleans and numbers but can be readily extended to include records and arrays.
+%
+A value stream is a total function from time (natural numbers including 0) to values:
+%
+\[
+ \konst{stream} = \mathbb{N}_0 \to \konst{value}
+\]
+%
+Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$ binding variable names to value streams, the semantics $\sem{-}^E_t$ of terms and formulas defines the meaning of compound syntax in terms of the meaning of subexpressions.
+%
+The value of a variable $v$ at time $t$ is found by looking up the stream bound to $v$ in $E$ (call it $s$) and returning $s_t$. Some clauses of the semantics follow, omitting the temporal operators:
+%
+\[
+\begin{array}{rcl}
+\sem{v}^E_t & = & E(v)(t) \\
+\sem{c}^E_t & = & c \\
+\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
+   & \cdots & \\
+\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
+   & \cdots & \\
+\end{array}
+\]
+
+The temporal operators deal with time in more significant ways.
+%
+The value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time $t-1$ (at time zero, \konst{pre} is undefined).
+%
+A followed-by, $e_1 \to e_2$, evaluates to $e_1$ at time zero; otherwise it is $e_2$.
+%
+Followed-by is typically used to give a value to a stream in the first step so that subsequent steps are able to use $\konst{pre}(e)$.
+%
+\[
+\begin{array}{rcl}
+\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
+\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
+\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
+\end{array}
+\]
+
+Although basic, these definitions can be used to define higher-level past-time operators from PLTL such as \konst{Once} and \konst{Since}.
+%
+These can be expressed in a recursive style as follows:
+%
+\[
+\begin{array}{rcl}
+\konst{Once}(b) & = & b \to b \vee \konst{pre}(\konst{Once}(b)) \\
+\konst{Since}(a, b) & = & b \to b \lor (a \land \konst{pre}(\konst{Since}(a, b))
+\end{array}
+\]
+%
+\noindent $\konst{Once}(b)$ is true if $b$ has ever been true before, or in, the current moment.
+%
+$\konst{Since}(a, b)$ is true if $a$ has been true in all moments up to the present one since $b$ most recently became true.
+%
+All other past-time operators of PLTL can be defined similarly \cite{monitor}.
+
+\agr\ uses ASCII notation for the mathematical syntax we have defined.
+%
+The following table spells out the mapping.
+%
+\[
+\begin{tabular}{l|l|l}
+ Operation & Math. notation & \agr \\ \hline
+ followed-by &  $e \to e$          & \verb+e -> e+  \\
+ logical implication &  $e \imp e$ & \verb+e => e+ \\
+ if-and-only-if &  $e \iff e$      & \verb+e <=> e+ \\
+ logical and &  $e \land e$      & \verb+e and e+ \\
+ logical or &  $e \lor e$      & \verb+e or e+ \\
+ logical not & $\neg b$      & \verb+not b+ \\
+ variable definition & $v = e$     & \verb+eq v e+ \\
+ Previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
+ Historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
+\end{tabular}
+\]
+
+%%
+% FORMAL CONTRACT FOR THE EXAMPLE SYSTEM
+%%
 
@@
 \label{sec:sw-contract}
-\input{sw-contract}
+
+\newsavebox{\sw}
+\begin{lrbox}{\sw}
+\begin{lstlisting}[style=agree,numbers=left]
+eq req : bool = event(AutomationRequest);*\label{line:sw-event-def-start}*
+eq avl : bool = event(AirVehicleLocation);
+eq wp  : bool = event(Waypoint);
+eq strt: bool = event(Start);
+eq alrt: bool = event(Alert);*\label{line:sw-event-def-end}*
+
+assume "Automation requests are well-formed" : *\label{line:sw-assume-1}*
+  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
+assume "Air vehicle locations are well-formed" : *\label{line:sw-assume-2}*
+  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);
+assume "One automation request in flight at a time" : *\label{line:sw-assume-3}*
+  true ->
+  (req => pre(Historically(not req) or Since(not req, strt)));
+
+guarantee "Waypoints coincide with air vehicle locations": *\label{line:sw-guarantee-1}*
+  wp => avl;
+guarantee "Starts include a new waypoint" : *\label{line:sw-guarantee-2}*
+  strt => wp;
+guarantee "Waypoints are well-formed" :  *\label{line:sw-guarantee-3}*
+  wp => WELL_FORMED_WAYPOINT(Waypoint);
+guarantee "Starts within one cycle of requests if not alerting" : *\label{line:sw-guarantee-4}*
+  (strt => ((not alrt) and req)) ->
+  (strt => ((not alrt) and (req or pre(req))));
+guarantee "Alert if not started within one cycle of requests" : *\label{line:sw-guarantee-5}*
+  true -> ((pre(req and not strt) and not strt) => alrt);
+guarantee "Once alerted always alerted" : *\label{line:sw-guarantee-6}*
+  Once(alrt) => alrt;
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \scalebox{0.62}{\usebox{\sw}}
+  \end{center}
+  \caption{The SW component contract.}
+  \label{fig:sw}
+\end{figure}
+
+The AGREE specification for the SW component in \secref{sec:example} is given in \figref{fig:sw}.
+%
+\linesref{line:sw-event-def-start}{line:sw-event-def-end} define local variables that are true when data is present on the corresponding event ports for the component.  These variables are
+purely for notational convenience.
+%
+The contract has three assumptions:
+%
+\begin{itemize}
+\item \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
+When input is present it is well-formed. The well-formed predicates
+are defined elsewhere using AGREE functions.
+\item \lineref{line:sw-assume-3}. If there is an incoming request,
+it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)},
+or it arrives after the component has output a start event in response to a previous
+request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
+request since start}.
+\end{itemize}
+
+\noindent The contract has six guarantees:
+%
+\begin{itemize}
+\item
+\lineref{line:sw-guarantee-1}, \lineref{line:sw-guarantee-2}, and \lineref{line:sw-guarantee-3}.
+%
+A waypoint coincides with a update on the UAV location, a start coincides with a new waypoint, and waypoints are well-formed.
+
+\item
+\lineref{line:sw-guarantee-4}.
+%
+A start happens with a request or one step after a request.
+%
+At time zero the start must coincide with the request;
+%
+after that, the start must coincide with the request or else is a response to a request one step earlier.
+%
+The guarantee works with the assumption on \lineref{line:sw-assume-3} and thus does not have to check for two requests in a row without a start, since the assumption precludes that input behavior.
+%
+The guarantee also does not force the start to always happen, it only says that if it does happen, it is under the defined conditions.
+
+\item
+\lineref{line:sw-guarantee-5}.
+%
+The alert is signaled if the start does not arrive within the one-step bound.
+%
+Together with \lineref{line:sw-guarantee-4} this allows for non-alerting and alerting behavior.
+
+\item
+%
+\lineref{line:sw-guarantee-6}.
+%
+If an alert has ever happened, it continues to happen.
+\end{itemize}
+%
+These six guarantees define the behavior of the SW component under the three assumptions and correspond to the informal descriptions given in \secref{sec:example}.
 

### 2023-02-14 paper/paper.tex
Inline the high assurance contracts sections

@@
 \begin{itemize}
-\item \lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
-When input is present it is well-formed. The well-formed predicates
-are defined elsewhere using AGREE functions.
-\item \lineref{line:sw-assume-3}. If there is an incoming request,
-it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)},
-or it arrives after the component has output a start event in response to a previous
-request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
+\item
+\lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
+%
+When input is present it is well-formed.
+%
+The well-formed predicates are defined elsewhere using AGREE functions.
+
+\item
+\lineref{line:sw-assume-3}.
+%
+If there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it arrives after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
 request since start}.
@@
 \label{sec:code-contracts}
-\input{high-assurance-contracts}
+
+Generally, AGREE contracts do not describe the computation that a component performs.
+%
+This is entirely by design: AGREE is intended to be declarative and reasoning about component behavior takes place solely at the specification level.
+%
+It defines what a component computes and not how it computes it.
+%
+However, the syntax of AGREE provides enough expressiveness to support the notion of a \emph{code contract}: a contract from which an implementation can be extracted.
+%
+First we must discuss a class of guarantees, \emph{output guarantees}, which determine the values on all output ports of a component.
+
+\begin{definition}[Output guarantee]
+An \emph{output guarantee} is a stylized guarantee that fully specifies the data written to an output port.
+%
+There are three possibilities according to whether the output port $p$ is a \konst{data} port, an \konst{event} port, or an \konst{event data} port:
+%
+\[
+\begin{array}{ll}
+\konst{data}: &  p = \mathit{e} \\
+\konst{event}: &  \konst{event} (p) = \mathit{b} \\
+\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
+\end{array}
+\]
+%
+\end{definition}
+
+Informally, a code contract treats its \konst{eq} ``statements'' as defining a list of assignments to state variables, and its output guarantees as directives for producing output.
+
+\begin{definition}[Code contract]
+A leaf component of the form $(I,O,A,P,\emptyset)$ is a \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where \[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set of \konst{eq} statements and $G$ is the set of output guarantees, one for each element of $O$.
+%
+In the interpretation as code, the order of elements of $\mathit{Eqs}$ is important, and is simply taken to be the occurrence order of the \konst{eq} statements in the syntax.
+%
+Thus we will work with the \emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
+\end{definition}
+
+The semantics of \secref{sec:agree-semantics} supports a formal connection between the original contract, and verification results of the AGREE model checker being run on it, and code generated from specifications of leaf level cyber-components.
+%
+It also provides the root meaning at the base of a chain of translation steps moving from an AGREE contract to a CakeML executable.
+
+The first step in the chain maps the contract to a code-focused representation.
+Assume given a code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$, environment $E$, and time $t$.
+%
+The \emph{evaluation} of $\mathit{Eqs}$ followed by the evaluation of $G$ results in a new environment $E'$ where the stream values for state variables and outputs have been computed for time $t$.
+%
+The step from $E$ to $E'$ is one full cycle in the repeated evaluation of the component.
+
+\begin{definition}[Evaluation]
+We overload existing notation and write the evaluation of $\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as $\sem{\mathit{Eqs}\cdot G}^E_t$.
+%
+The evaluation of $v = e \in {\Eqs}$ is an environment transformation
+%
+\[
+ \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
+\]
+%
+modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
+%
+The list {\Eqs} is evaluated by folding the transformation left-to-right through it.
+%
+The transformation is similarly applied to the list of output guarantees $G$, computing the values on output ports.
+%
+The details are omitted, being a bit messy because of the different kinds of output ports available.
+\end{definition}
+
+\begin{definition}[Code contract correctness]
+Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
+%
+The contract is \emph{correct}, if for all $E$ and $t$, whenever the assumptions (historically) hold in $E$ and evaluation steps from $E$ to $E'$, then the guarantees hold in $E'$:
+%
+\[
+\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
+\]
+%
+\end{definition}
+
+How does this notion of correctness integrate with the verification conditions generated by AGREE?
+%
+In \emph{contract correctness}, the model checker is proving the following property
+%
+\[
+\konst{Hist}(A) \imp P
+\]
+%
+and in \emph{code contract correctness} we essentially prove a Hoare triple of the form
+%
+\[
+\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
+\]
+%
+This accords with intuition, namely that AGREE is a kind of \emph{program-free program logic};
+%
+by pulling out a program $(\mathit{Eqs}\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification.
+%
+For example, at this point one could apply a verification condition generator to help break down the proof obligation.
+%
+In order to use the code contract result in system contract verification conditions, one has to show that the code contract result implies the system verification condition.
+
+%%
+% IMPERATIVE CODE CONTRACTS
+%%
+
+\subsubsection*{Imperative code contracts}
+
+Stream-based evaluation allows values arbitrarily \emph{deep in the past} to be accessed by use of $\konst{pre}(-)$.
+%
+However, the programming languages targeted by SPLAT do not directly support streams.
+%
+We are developing a translation to eliminate such deep temporal accesses by the introduction of intermediate variables.
+%
+The essential property the translation needs to satisfy is a syntactic restriction, call it \emph{Pascal-like}, that supports easy translation to standard imperative languages.
+%
+The guiding intuition is that a Pascal-like {\Eqs} accesses at most one step in the past;
+%
+also, accesses to a past value of a variable are only possible until the variable is updated.
+%
+In informal terms, this means that ${\Eqs}\cdot G$ in a code contract has the following properties:
+%
+\begin{itemize}
+\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
+\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
+\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
+\end{itemize}
+%
+For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be defined, albeit in a manner that is not Pascal-like, by
+%
+{\small
+\begin{lstlisting}[style=agree]
+  Fib = 1 -> pre(1 -> Fib + pre Fib)
+\end{lstlisting}
+}
+%
+\noindent or by (again this is not Pascal-like)
+%
+{\small
+\begin{lstlisting}[style=agree]
+  N = 0 -> 1 + pre(N)
+  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
+\end{lstlisting}
+}
+%
+\noindent or by (this time Pascal-like):
+%
+{\small
+\begin{lstlisting}[style=agree]
+  N = 0 -> 1 + pre(N)
+  F2 = 42 -> pre(F1)
+  F1 = 42 -> pre(Fib)
+  Fib = if N <= 1 then 1 else F1 + F2
+\end{lstlisting}
+}
+%
+Please note that the initial values of \verb+F1+ and \verb+F2+ are never accessed, so 42 could be replaced by any number in their defining equations.
+%
+A Pascal-like ${\Eqs} \cdot G$ is straightforward to implement in conventional programming languages, as we will see in \secref{sec:synthesis}.
+
+\subsection{Code contracts for the hardened system}
+\label{sec:ha-contracts}
+
+\newsavebox{\flt}
+\begin{lrbox}{\flt}
+  \begin{lstlisting}[style=agree,numbers=left] -- start user
+    -- start user definitions
+    eq policy : bool =
+      WELL_FORMED_AUTOMATION_RESPONSE(Input);
+    -- end user definitions
+
+    guarantee "Filter output is well-formed" :
+      if event(Input) and policy then
+        event(Output) and Output = Input
+      else
+        not event(Output);
+  \end{lstlisting}
+\end{lrbox}
+
+\newsavebox{\mntr}
+\begin{lrbox}{\mntr}
+  \begin{lstlisting}[style=agree,numbers=left]
+    const is_latched : bool = true;
+
+    -- start user definitions
+    assume "One automation request in flight at a time" : *\label{line:mon-assume}*
+      true -> (req => pre(Historically(not req) or Since(not req, rsp)));
+
+    const MAX_LATENCY : int = 1; *\label{line:mon-const}*
+
+    eq rsp : bool = event(Response);
+    eq req : bool = event(Request);
+
+    eq preIsPending : bool = false -> pre(isPending);*\label{line:mon-pre-pending}*
+    eq isPending : bool = Since(not rsp, req and not rsp);*\label{line:mon-pending}*
+    eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);*\label{line:mon-latency}*
+
+    eq policy : bool = (rsp => req) -> *\label{line:mon-policy}*
+                         (    (isPending => latency < MAX_LATENCY)
+                          and (rsp => (req or preIsPending)));
+    -- end user definitions
+
+    eq alert : bool = (not policy) -> *\label{line:mon-alert}*
+                        ((is_latched and pre(alert)) or not policy);
+
+    guarantee "Alert port tracks alert variable" :
+      event(Alert) = alert;
+    guarantee "Output if not alerted" :
+      if (not(alert) and rsp) then
+          event(Output) and (Output = Response)
+      else
+          not (event(Output));
+  \end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.62}{\usebox{\flt}} \\
+    \end{tabular}
+  \end{center}
+  \caption{Code contract for high-assurance filter.}
+  \label{fig:filter}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+    \scalebox{0.62}{\usebox{\mntr}} \\
+    \end{tabular}
+  \end{center}
+  \caption{Code contract for high-assurance monitor.}
+  \label{fig:monitor}
+\end{figure}
+
+Code contracts for the filter and monitor are shown in \figref{fig:filter} and \figref{fig:monitor}.
+%
+The filter is direct: it defines the policy according to the well-formed requirement for data.
+%
+We now discuss the monitor in some detail.
+
+The monitor informally behaves as follows: if the \emph{policy} is obeyed in the current step, and the alert has not yet been raised, then the \texttt{Response} input is sent out on \texttt{Output}, and the \texttt{Alert} output stays low \lineref{line:mon-alert}.
+%
+Otherwise \texttt{Alert} goes high and stays high, and there is nothing sent henceforth on \texttt{Output}.
+
+The \texttt{alert} variable (distinct from the \texttt{Alert} port) depends not just on the policy but also on the \texttt{is\_latched} constant, which has been set at configuration time.
+%
+Since \texttt{is\_latched} is \konst{true}, the alert is persistent, meaning that once the alert is raised, it is always raised;
+%
+otherwise, it is the complement of the policy value in the current step.
+
+The monitor policy is defined by marking, with variable \texttt{isPending}, when a request is made but not yet satisfied, and by counting the number of steps (\texttt{latency}) between requests.
+%
+See \linesref{line:mon-pre-pending}{line:mon-latency}:
+%
+The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response, \ie, \texttt{isPending} holds when \emph{there is no response since the most recent request without a corresponding response}.
+%
+The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.
+Note that the \texttt{preIsPending} variable is a temporary needed to make the contract Pascal-like;
+%
+otherwise, the current and previous value of \texttt{isPending} would be simulataneously accessed in the definition of \texttt{policy}.
+
+The policy (\lineref{line:mon-policy}) makes sure the latency is bounded whenever there is a pending request, and it ensures that a response is a result of a request.
+%
+The initial step differentiates from later steps in only needing to be sure a response is a result of a request, both happening in the initial step.
+
+The latency bound for the monitor is defined by the constant \texttt{MAX\_LATENCY} (\lineref{line:mon-const}).
+%
+In this example the constant is set to one to be consistent with the system specification from the previous section.
+%
+At time zero the policy is trivial: a response requires an accompanying request.
+%
+After time zero, if a request is pending in the current time step, then the latency must still be under the bound, and if there is a response, then it closes a request now or a pending request from earlier in time.
+
+The policy definition only works if there is never more than one outstanding request at a time;
+%
+otherwise, the latency counter resets incorrectly.
+%
+That requirement is encoded in the assumption (\lineref{line:mon-assume}) and is the same assumption used for the system input.
+%
+In fact, the policy can be written without the assumption, but the policy is much simpler to write with the assumption.
+
+As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
+%
+That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
+%
+Synthesis must generate faithful implementations of the code contracts for the AGREE verification result to apply to the deployed system.
 

### 2023-02-14 paper/paper.tex
Inline the testing section

@@
 
+%%
+% TEST CONTRACTS
+%%
+
 \section{Test contracts}
 \label{sec:testing}
-\input{testing}
+
+Writing and proving properties about code contracts is difficult because it is a formal process, and because it reasons about constraints that define the entire input space, and the entire output space, of a high-assurance component at once.
+%
+Case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
+
+We have observed three common types of mistakes in code contracts (aside from self-consistency, which is already checked): vacuity, under-specification, and missed corner cases.
+%
+Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true.
+%
+Under-specification is very challenging, as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
+%
+Missed corner cases is a common problem to all disciplines.
+%
+It is an especially difficult challenge here because there is no direct way to exercise code contracts, \eg, there is no runtime \emph{per se}.
+
+A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with traditional testing.
+%
+Such an approach adds extra steps, including test harneses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional validation using tests.
+%
+The intuition for that bridge comes from the compositional reasoning in AADL and \agr.
+%
+Effectively, a code contract for a high assurance component should be a valid implementation of a test scenario.
+%
+In other words, the test scenario, as expressed in a test contract, is the desired behavior of the system, and the code contract implements that behavior.
+
+A test contract defines the test scenario for the code contract to implement.
+%
+It does this by assuming the test input and guarantying the test output akin to a unit test.
+%
+The code contract is used as the implementation, \eg, it is the test subject.
+%
+The \agr\ analysis then proves if the code contract, given the test inputs, is strong enough to prove the test contract outputs.
+%
+Here the test contract assumptions strengthen the code contract assumptions to a single input, and the test contract guarantees weaken, or leave unchanged, the code contract output for that test input, depending on the goal of the test.
+%
+In other words, according to Liskov substitution, \agr\ proves if the code contract is a safe substitution for the test contract.
+
+\newsavebox{\tst}
+\begin{lrbox}{\tst}
+  \begin{lstlisting}[style=agree,numbers=left]
+    process should_notAlertAndOutput_when_responseOneStepAfterRequest
+      ... -- Omitted AADL features
+      annex agree {**
+        eq index : int = prev(index + 1, 0); *\label{line:tst-index}*
+
+        assume "One Response one step after Request" : *\label{line:tst-assume-1}*
+                ((index = 0) => not event(Response))
+            and ((index = 1) => event(Response))
+            and ((index >= 2) => not event(Response));
+        assume "One Request" : *\label{line:tst-assume-2}*
+            (event(Request) = true) ->
+            (event(Request) = false);
+
+        guarantee "Not Alert" : *\label{line:tst-guarantee-1}*
+            not event(Alert);
+        guarantee *\label{line:tst-guarantee-2}*
+          "Output one step after Request at same time as Response" :
+                ((index = 0) => not event(Output))
+            and ((index = 1) => (event(Output) and Output = Response))
+            and ((index >= 2) => not event(Output));
+      **};
+    end should_notAlertAndOutput_when_responseOneStepAfterRequest;
+
+    process Implementation *\label{line:tst-imp-start}*
+      should_notAlertAndOutput_when_responseOneStepAfterRequest.test
+      subcomponents
+        Monitor: thread CASE_Monitor_Thr.Impl; *\label{line:tst-imp-comp}*
+      connections
+        c00: port Response -> Monitor.Response;
+        c01: port Request -> Monitor.Request;
+        c02: port Monitor.Alert -> Alert;
+        c03: port Monitor.Output -> Output;
+    end should_notAlertAndOutput_when_responseOneStepAfterRequest.test; *\label{line:tst-imp-end}*
+  \end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+    \scalebox{0.60}{\usebox{\tst}} \\
+    \end{tabular}
+  \end{center}
+  \caption{A unit test for the monitor with its test contract.}
+  \label{fig:test}
+\end{figure}
+
+\figref{fig:test} is a unit test, with its test contract, for the monitor in \figref{fig:hardened}.
+%
+The test contract checks if the monitor output coincides with the response that comes one step after the request.
+%
+It also checks that the alarm is not output.
+%
+\begin{compactitem}
+\item \lineref{line:tst-index}: An index to use to define the test input and expected test output through time.
+\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: The test input. These define when the request and response message events should appear relative to the index.
+\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: The expected test output.
+The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
+\item \linesref{line:tst-imp-start}{line:tst-imp-end}: The implementation with the component under test (\lineref{line:tst-imp-comp}).
+\end{compactitem}
+%
+\figref{fig:hardened-certificate}(a) is the \agr\ output for the test showing that it passes.
+
+Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
+%
+They enable designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
+%
+These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.
 

### 2023-02-14 paper/paper.tex
Inline synthesis

@@
 
+%%
+% SYNTHESIS TO CAKEML
+%%
+
 \section{Synthesis to CakeML}
 \label{sec:synthesis}
-\input{synthesis}
+
+Synthesis maps from model and specifications to code.
+%
+The \splt\ tool traverses the system architecture looking for occurrences of high assurance components specified by code contracts;
+%
+for each such occurrence it generates a CakeML program.
+%
+\splt\ supports three kinds of behavioral specification from which to generate code.
+%
+In increasing order of expressiveness, these are: regular expressions, contiguity types, and code contracts.
+%
+In earlier work we have described how regular expressions and contiguity types are used as specification languages for verified filter implementations.
+%
+In this paper, we extend this \emph{specification oriented} approach to the Turing-complete specifications of code contracts in \agr.
+%
+We give a short review of that previous work before going into detail on the synthesis.
+
+\paragraph{Regular expressions:\/} often the well-formedness of data on connections can be enforced by matching against a regular expression.
+%
+In other words, the data can be characterized by a \emph{regular language}.
+%
+As is well known, regular expressions can be translated to efficient deterministic finite state machine (DFA) implementations.
+%
+In \cite{formal-filter-synth-langsec,case-verified-filter} we report on the creation and application of table-based DFA filters from a verified regular expression-to-DFA compiler.
+%
+The latter paper offers a fully-worked out example synthesizing a regular expression filter and formally proving that it meets an non-terminating behavioral specification.
+
+\paragraph{Contiguity types\cite{contiguity-types}:\/} contiguity types, similar to regular expressions, are effective at describing data at the network message level, \eg, as flat strings.
+%
+The notation is however much more expressive, being able to declare many kinds of so-called \emph{self-describing} messages, \ie, those where message structure is dependent on data held elsewhere in the message.
+%
+\figref{fig:filter-spec} gives a contiguity type specification for the well-formed waypoints of our example system.
+%
+We have formalized and verified a contiguity type matcher, which has been extensively applied throughout our case studies.
+
+%%
+% COMPILING FROM CODE CONTRACTS
+%%
+
+\subsection{Compiling from code contracts}
+
+When full computational power is needed, we can use the expressive power of AGREE code contracts to write arbitrarily complex filter or monitor operations, as we have already seen.
+%
+Our formal model of AGREE syntax and semantics is applied to extract code contracts and map them to equivalent logic functions.
+%
+In particular, given a code contract with functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics given in \secref{sec:code-contracts} can be applied to define a stream-oriented \emph{step} function $\konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env}$ defined as
+%
+\[
+ \konst{strmStep}\;t \; E = \sem{\mathit{Eqs} \cdot G}^E_t
+\]
+%
+Now suppose ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially imperative.
+%
+Then we can direcctly create a logic function
+%
+\[
+\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}
+\]
+%
+corresponding to \konst{strmStep} that takes input and state as parameters and returns an updated state and output.
+%
+Thus the environment $\mathit{Env}$ holding all previous values of each variable can be replaced by \emph{parameters}, each of which is able to hold only one data element.
+%
+Assuming $\mathit{E}' = \konst{strmStep}\;t\; \mathit{E}$, we can formally prove that
+%
+\[
+\vdash (\mathit{E}' \vert_{\mathbb{S}},
+ \mathit{E}' \vert_{\mathbb{O}}) =
+\konst{stateStep} (\mathit{E}\vert_{\mathbb{I}(t)},
+                   \mathit{E}\vert_{\mathbb{S}(t)})
+\]
+%
+In other words, projecting out the state and output components from the result of $\konst{strmStep}\;t\; \mathit{E}$ equals the result of applying \konst{stateStep} on the input and state components of $\mathit{E}$ at time $t$.
+%
+This theorem connects the stream semantics of the code contract to a pure logic function over the values appearing in the stream.
+%
+From this point we can use tools in the \ckml\ eco-system to jump from the setting of higher order logic to \ckml\ abstract syntax \cite{cakeml-translator}.
+%
+In this way we have moved from stream semantics to logic function to CakeML evaluation semantics in a property-preserving way.
+%
+Once in \ckml, further necessary features can be added, namely imperative variables for storing state between component invocations, also invocations of the (verified) foreign function interface of \ckml\ \cite{cakeml-monadic}.
+%
+Finally, the CakeML compiler can be applied to the program.
+%
+It delivers an executable which can be shown to preserve the original contract.
+
+%%
+% GENERATING CAKEML
+%%
+
+\subsection{Generating \ckml}
+For any of the three specification languages, we take the following general steps to arrive at an executable.
+%
+\begin{compactenum}
+\item Map from specification to computable function in logic, producing a proof of equivalence.
+\item Add parsers and printers (specified by contiguity types) for input and output.
+\item Map to CakeML and add calls to input/output library routines via the CakeML foreign function interface.
+\item Invoke the CakeML compiler.
+\end{compactenum}
+%
+In the following, we examine further details in both filter and monitor synthesis.
+
+%%
+% FILTER GENERATION
+%%
+
+\subsection{Filter Generation}
+
+A filter is intended to be simple, although it may make deep semantic checks.
+%
+A filter has one input port and one output; messages on the input that the filter policy admits pass unchanged to the output port;
+%
+all others are dropped (not passed on).
+%
+We have investigated two kinds of filter.
+%
+In the first, a relatively shallow scan of the input suffices to enforce the policy.
+%
+For example, we have used the expressive power of regular expressions and Contiguity Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints on GPS coordinates in UxAS messages.
+%
+On the other hand, a filter may need to parse the input buffer into a data structure specified in AGREE and apply a user-defined \emph{wellformedness} property, also specified in AGREE, to the data.
+%
+Arbitrarily complex wellformedness checks can be made in this way.
+%
+\figref{fig:filter-spec} shows a combination where the checking specified by {\small\verb+WELL_FORMED_AUTOMATION_RESPONSE+} depends on an underlying check specified by the contiguity type checking bounds on waypoints.
+
+The verdict of a filter is made and performed within one thread invocation.
+%
+Thus, in its given time slice, the following steps must be completed:
+\begin{compactenum}
+\item
+The filter checks to see if there is any input available.
+%
+If there is none then it yields control; otherwise:
+
+\item The input is read (and parsed if need be);
+
+\item The wellformedness predicate is evaluated on the input;
+
+\item If the predicate returns \konst{true} then the input buffer
+ is copied to the output, otherwise no action is taken; and
+
+\item The filter yields control.
+\end{compactenum}
+
+\newsavebox{\contig}
+\begin{lrbox}{\contig}
+\begin{lstlisting}[style=myML]
+  Waypoint =
+    {Latitude  : f64
+     Longitude : f64
+     Altitude  : f32
+     Check     : Assert
+      (~90.0 <= Latitude and Latitude <= 90.0 and
+       ~180.0 <= Longitude and Longitude <= 180.0 and
+       1000.0 <= Altitude and Altitude <= 15000.0)}
+
+  AutomationResponse =
+    {TaskID : i64
+     Length : u8
+     Waypoints : Waypoint [3]}
+
+ fun WELL_FORMED_AUTOMATION_RESPONSE(aresp) =
+   (forall wpt in aresp.Waypoints, WELL_FORMED_WAYPOINT(wpt))
+   and ... ;
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\contig}}
+    \end{tabular}
+  \end{center}
+  \caption{Filter specification.}
+  \label{fig:filter-spec}
+\end{figure}
+
+\newsavebox{\cml}
+\begin{lrbox}{\cml}
+\begin{lstlisting}[style=myML]
+fun filter_step () =
+ let val () = Utils.clear_buf buffer
+     val () = API.callFFI "get_input" "" buffer
+ in
+    if WELL_FORMED_AUTOMATION_RESPONSE buffer
+    then
+      API.callFFI "put_output" buffer Utils.emptybuf
+    else print"Filter rejects message.\n"
+end
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\cml}}
+    \end{tabular}
+  \end{center}
+  \caption{Synthesized CakeML for the filter.}
+  \label{fig:filter-cakeml}
+\end{figure}
+
+The contiguity type specification and wellformedness predicate for the filter are shown in \figref{fig:filter-spec} and the synthesized \ckml\ code is in \figref{fig:filter-cakeml}.
+%
+The code is called at dispatch by the scheduler.
+%
+The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
+%
+The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
+
+%%
+% MONITOR GENERATION
+%%
+
+\subsection{Monitor Generation}
+
+As discussed above, a monitor specification is mapped to a state transformation function
+%
+\[
+\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}.
+\]
+%
+which computes as follows:
+%
+\begin{compactenum}
+
+\item Each input buffer is parsed into data of the type specified by the port type;
+
+\item
+New values for the state variables are computed, in dependency order.
+%
+At this point variables defined by \emph{followed-by} are dealt with.
+%
+Suppose the variable definitions have the following form:
+%
+\[
+\begin{array}{l}
+  v_1 = i_1 \longrightarrow e_1 \\
+  \cdots \\
+  v_n = i_n \longrightarrow e_n \\
+\end{array}
+\]
+%
+In the generated function, for the first invocation of \konst{stateStep} only, the initializations are executed in order:
+%
+\[
+  v_1 = i_1; \cdots ;  v_n = i_n;
+\]
+%
+In all later invocations, the \emph{non-initialization} assignments are performed, also in order:
+%
+\[
+  v_1 = e_1; \cdots ; v_n = e_n; \\
+\]
+
+\item Values of the outputs are computed;
+
+\item Outputs are written and the new state is written;
+
+\item The monitor yields control.
+\end{compactenum}
+
+The \konst{stepFn} for the monitor of the example described in \secref{sec:example} is displayed in \figref{fig:monitor-cakeml}.
+
+\newsavebox{\monFn}
+\begin{lrbox}{\monFn}
+\begin{lstlisting}[style=myML]
+fun stateStep (response,request)
+    (rsp,req,preIsPending,isPending,latency,policy,alert)
+let val stateVars' =
+      if !initStep then
+      let val rsp = Option.isSome response
+          val req = Option.isSome request
+          val preIsPending = False
+          val isPending = req andalso not rsp
+          val latency = 0
+          val policy = (if rsp then req else True)
+          val alert = not(policy)
+          val () = (initStep := False)
+      in
+        (rsp,req,preIsPending,isPending,latency,policy,alert)
+      end
+      else
+      let ...
+          val rsp = Option.isSome response
+          val req = Option.isSome request
+          val preIsPending = pre isPending
+          val isPending = (req andalso not rsp) orelse
+                          ((not rsp) andalso pre isPending)
+          val latency = (if req then 0 else (pre latency + 1)
+          val policy =
+            ((if isPending then
+                (latency < Defs.max_latency) else True)
+              andalso
+              (if rsp then
+                 (req orelse preIsPending) else True))
+          val alert = (Defs.is_latched andalso (pre alert))
+                      orelse not(policy)
+      in
+        (rsp,req,preIsPending,isPending,latency,policy,alert)
+      end
+   val (rsp',_,_,_,_,_,alert') = stateVars'
+   val alertPort = if alert' then Some () else None
+   val outputPort =
+     if alert' then None else
+     if rsp' then Some response
+     else None
+in
+   (stateVars', (alertPort,outputPort))
+end
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c}
+      \scalebox{0.60}{\usebox{\monFn}}
+    \end{tabular}
+  \end{center}
+  \caption{Synthesized CakeML for the monitor.}
+  \label{fig:monitor-cakeml}
+\end{figure}
+
+\subsection{Incorporation into system execution}
+
+The system scheduler \emph{activates} components in some order.
+%
+It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+%
+Activating a monitor component takes the form of the following pseudo-code, in which the monitor evaluates the \konst{stateStep} on its current inputs and the current values of the state variables, returning the new state and the output values.
+%
+\[
+\begin{array}{ll}
+ \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
+ (v_1,\ldots) & = \konst{readState}() ; \\
+ ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stateStep} ((i_1,\ldots),(v_1,\ldots)) ; \\
+ \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
+ \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
+\end{array}
+\]
+
+\paragraph{Partiality:\/} partiality is an important consideration for both filters and monitors: the input data might not parse or the computation of a filter could be badly written and fail at runtime.
+%
+In such cases, the filter should recover and yield control without passing the input onwards.
+%
+For these failure cases, the filter is behaving as it should, but we must also guard against situations in which a \emph{correctly specified} filter or monitor fails in the runtime.
+%
+This kind of defect arises when a filter, for example, \emph{ought} to accept a message, but lack of resources results in the filter failing to do so.
+%
+For example, the parse of a message might need more space than has been allocated;
+%
+another example could be if the time slice provided by the scheduler is too short for the computation to finish.
+%
+Thus resource bounds need to be included in the correctness argument.
+%
+Some preliminary work that could support formal proofs of such bounds has been done for \ckml\ \cite{cakeml-space-cost}.
 

### 2023-02-14 paper/paper.tex
Inline the case study

@@
 
+%%
+% CASE STUDY
+%%
+
 \section{Case study}
 \label{sec:case-study}
-\input{case-study}
+
+In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
+%
+The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides designers with early validation of their model;
+%
+produces early information about cyber-vulnerabilities;
+%
+provides effective mitigations;
+%
+and handles a complex real-world message format, namely the UxAS message format.
+
+In this surveillance system, the UAV receives commands from a ground station to conduct surveillance over a geographical region.
+%
+In response, the UAV's on-board mission computer generates a flight plan consisting of a series of waypoints that the UAV must traverse to complete its mission.
+%
+The UAV is also given a set of \textit{keep-in} and \textit{keep-out} zones that may constrain its flight path.
+
+We have modeled the system architecture of the UAV in AADL.
+%
+The model includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
+%
+The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
+%
+The initial software architecture model (shown in \figref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
+%
+The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.
+
+For this application, UxAS accepts three types of messages.
+%
+The \textit{Operating Region} message defines where the UAV can and cannot fly.
+%
+The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.
+%
+The waypoints typically lie along some geographical feature of interest, such as a river or railway.
+%
+Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.
+%
+Anytime after receiving the Operating Region and Line Search Task messages, a Ground Station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
+%
+UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.
+%
+Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=2\columnwidth]{figs/sw-initial.png}
+	\caption{Initial software architecture.}
+	\label{fig:sw-initial}
+\end{figure*}
+
+Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
+%
+We perform a formal analysis with \agr\ to verify that the model satisfies the contracts.
+%
+For the initial version of our design, the verification passes.
+%
+Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.
+
+In \brfcs, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
+%
+As the designers of the system, we reviewed the list of potential vulnerabilities, and then added new requirements to the system model.
+%
+For example, the vulnerbility analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
+
+In total, we added seven cyber requirements from the vulnerability analysis into our model.
+%
+These include four wellformedness requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
+%
+Requirements are imported into the model as goals in the assurance case.
+%
+Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.
+
+The intent of the wellformedness requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
+%
+In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
+%
+By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
+%
+The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
+
+The filter code contracts for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
+%
+For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
+%
+Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit floating-point values, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/automation-response-filter.png}
+	\caption{Automation Response Filter specification.}
+	\label{fig:automation-response-filter}
+\end{figure}
+
+In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
+%
+This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
+%
+The Monitor transform is applied for this class of mitigation.
+%
+We add a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+%
+And we add a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+%
+The Geofence Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
+%
+The Geofence Monitor code contract is shown in \figref{fig:geofence-monitor}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\columnwidth]{figs/geofence-monitor.png}
+	\caption{Geofence Monitor specification.}
+	\label{fig:geofence-monitor}
+\end{figure}
+
+So far, we have addressed requirements that mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
+%
+But we also want to protect against a compromised ground station transmitting well-formed, but malicious, commands.
+%
+The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
+%
+The Attestation Manager \ckml\ implementation is automatically inserted into the application code base by \brfcs.
+%
+The code contract for the Attestation Gate is completely generated by the transform since its behavior is specific to attestation.
+
+After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
+%
+The shaded components (green) were added by the transform with the exception of the UxAS component (red).
+
+After unit testing the monitors, we formally verify the model with \agr\ to show that all of the component contracts are satisfied.
+%
+We then run \splt\ to produce provably-correct code.
+%
+The synthesized code is output to a directory in the build file system with the location specified for each component in the model.
+%
+The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=2\columnwidth]{figs/hardened-sw.png}
+	\caption{Cyber-resilient software architecture.}
+	\label{fig:hardened-sw}
+\end{figure*}
+
+Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by \splt\ or other means, the system can be built and deployed.
+%
+We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
+%
+HAMR then compiles the software to execute on seL4.
+
+The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
+%
+The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
+%
+The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
+%
+Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior.
+In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
+%
+We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.
+A third party evaluator repeated the tests on the hardened system and demonstrated that our mitigation strategy was successful.
+%
+The following scenarios were exercised:
+
+\paragraph{Infected Ground Station:\/} An application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
+%
+On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.
+
+\paragraph{Malformed Line Search Task message:\/}
+%
+The Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
+%
+The wellformedness filter for Line Search Task messages prevented this message from reaching the UxAS.
+
+\paragraph{UxAS vulnerability exploit:\/} Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
+%
+When this occurs, UxAS is unable to generate an Automation Response.
+%
+The response monitor detected and signaled the Ground Station when UxAS failed to respond to the request.
+
+\paragraph{UxAS trojan modifies flight plan:\/} A trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
+%
+On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and returned the UAV to Home Base.
+
+%%
+% DISSCUSSION
+%%
+
+\subsection{Discussion}
+
+The case study demonstrated the viability of writing code contracts and synthesizing implementations from code contracts in a real-world system.
+%
+It additionally provided insights for improvement.
+%
+We discovered very early that we needed a mechanism to test contracts.
+%
+We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the OSATE environment and an external test platform.
+%
+Test contracts are a direct result of our experience, and they obviate the back and forth by integrating test into the modeling and analysis environment.
+
+Writing code contracts for the more complex monitors revealed that although the language is very effective at declarative specifications that define what is being calculated, it is non-intuitive, and sometimes cumbersome, for defining how the computation needs to take place.
+%
+The primary reason for its unease in describing how computation is performed is that it lacks any type of iterative looping structure for arrays.
+%
+Instead, it relies on folding and mapping to describe iterative computation which is non-intuitive for many types of array updates.
+%
+Growing the expressiveness of the language arround arrays is something to investigate in future work.
+
+There are several, practical, low-level concerns that need to be addressed in the high-level model before synthesis takes place that in our process were only discovered after running the actual system on the designated hardware and software.
+%
+We specifically found it difficult to know the memory bounds for heaps, stacks, and buffers and time bounds for the scheduler \emph{a priori} to code synthesis.
+%
+As a result, we would have to employ a costly iterative approach to set bounds, synthesize code, and then check if anything failed.
+%
+Integrating better memory and running time analysis at the modeling level is something to investigate in future work.
 

### 2023-02-14 paper/paper.tex
Inline related work

@@
 \label{sec:related-work}
-\input{related-work}
+
+Another model based approach to secure system architectures follows a pattern similar to \brfcs\ with attestation by inserting pre-verified components into the architecture to give certain guarantees \cite{6209212}.
+%
+In that work, the data connections are annotated with security properties as are the components.
+%
+A secondary analysis determines if the components meet the security requirements of the data connections, and if not, inserts new security enhancing components in the connection.
+%
+The insertion algorithm considers time cost when inserting components to preserve worst-case execution times mandated by the system.
+%
+In the \brfcs\ tool chain there are tools that can analyze the model annotated with security properties and identify where security vulnerabilities may be located, and there are other tools to transform the model to address those vulnerabilities such as those discussed in this paper.
+
+Assume-guarantee reasoning for compositional verification in reactive systems is well-studied \cite{10.1007/978-3-642-28891-3_13, 10.1145/2658982.2527272, 10.1007/978-3-319-17524-9_7}.
+%
+The first steps toward synthesis of leaf-component contracts determined if leaf-component contracts are realizable \cite{10.1007/978-3-319-17524-9_13, 10.1007/978-3-319-29613-5_7}.
+%
+These led to algorithms for component synthesis from Lustre models based on realizability \cite{katis2017synthesis, 10.1007/978-3-319-89963-3_10}.
+%
+The algorithms discover a transition relation that preserves the contract guarantees under all allowed inputs and can be directly transformed into the language intended for the system's implementation.
+%
+Two such algorithms are implemented in the \agr\ verification engine \cite{jkind}.
+%
+These synthesis algorithms take in a declarative specification that states \emph{what is to be computed}, and then they have to infer \emph{how to compute it} for the implementation.
+%
+Code contracts are imperative, by design, so the designer specifies directly how the computation of the high-assurance component takes place thereby avoiding the question of realizability entirely.
+%
+That said, the test contracts in this paper can be declarative, stating what is to be computed, if desired, and \agr\ can then prove if the code contracts accomplishes that.
+
+Data-flow semantics are well studied \cite{10.1145/41625.41641,97300, 10.1145/1379023.1375674,10.1145/2345141.2248426,10.1007/978-3-540-45212-6_10}.
+%
+State machine semantics can be added to synchronous data-flow \cite{10.1145/1086228.1086261}.
+%
+The idea is to translate imperative constructs into equivalent synchronous data flow constructs.
+%
+The resulting Lustre can then be compiled. The goal is to provide a seamless connection between pure data-flow and pure control design.
+%
+These approaches are similar to the code contracts in this paper in that they embed imperative computation into the synchronous data-flow semantics of Lustre which are then passed to compilers to render the final target binaries.
+%
+These compilers however are not verified as is the \ckml\ compiler targeted in this work.
+
+A fully verified compiler has been developed that takes Lustre and compiles it into assembly code, specified and verified in Coq \cite{10.1145/3140587.3062358}.
+%
+The key is in combining infinite sequences of data-flow models with incremental manipulation of memories akin to an imperative model.
+%
+CompCert is used on the backend to create the final rendered executable \cite{compcert}.
+%
+This work is quite similar to ours in providing a verified compilation chain from Lustre.
+%
+However, there are significant differences: notably, they tackle a more comprehensive subset of Lustre than our work.
+%
+We have been able to work effectively with a relatively simple core language, \eg, without \emph{clocking} features.
+%
+This subset aligns with our contract language, and allows us to generate behaviorally equivalent code so we can show that the generated code meets its logical specification.
+%
+This, in turn, can be used as a leaf-level claim in a system-wide correctness argument conducted by compositional reasoning in \agr.
+%
+It is important to note that we target a wholistic systems approach to cyber-security that is integrated into the \agr\ reasoning environment whereas the other work targets more specifically a compiler for Lustre.
+
+CoPilot is a domain specific language for writing runtime monitors for hard real-time distributed systems embedded in Haskell and requires working Haskell knowledge to write \cite{10.1007/s11334-013-0223-x}.
+%
+The language semantics are defined over streams, there is an interpreter for testing the behavior of the runtime monitors, and there are two backend compilers to generate C code.
+%
+The generated C code is validated by generating random CoPilot programs, generating random input for each program, and checking consensus on output between the interpreter and C programs.
+%
+Further assurance is accomplished with bounded model checking by showing equivalence between the code from the two compiler backends over a user defined number of steps.
+%
+Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high assurance components.
+
+As seen in this paper, contracts can be stateful and even higher-order \cite{10.1145/583852.581484}.
+%
+Verification aware languages compile contracts to runtime monitors \cite{10.1007/978-3-642-28869-2_11}, or omit them when they can statically prove that the code preserves the contract under all possible inputs and executions \cite{10.1145/3158139}.
+%
+Dafny is a notable verification aware language that provides a contract language, an implementation language, a proof engine to show if an implementation adheres to the contract, and backend compilers for several different target platforms \cite{dafny}.
+%
+Dafny contracts provide the verification conditions to prove out the implementation code and are not compiled.
+%
+Compared to the work discussed in this paper, our test contracts are more similar to the Dafny contracts as they are also not compiled and only serve to prove properties of the code contracts.
+%
+In that regard, the code contracts are the implementation, but unlike Dafny, \ckml\ provide a verified compilation path to binaries.
 

### 2023-02-14 paper/paper.tex
Inline the conclusion

@@
 
+%%
+% CONCLUSION
+%%
+
 \section{Conclusion}
 \label{sec:conclusion}
-\input{conclusion}
+
+Safety critical systems, such as those found in avionics, must be cyber-resilient, i.e. tolerant to cyber-attacks, in the same way that they are fault tolerant.
+%
+The \brfcs\ environment integrated into OSATE provides systems designers with tools to analyze and address cyber-vulnerabilities in the architectural design of a system modelled using AADL.
+%
+We have discussed transformations provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
+
+A filter enforces invariant properties on data such that data that violate the invariant are not forwarded.
+%
+A monitor detects and alerts malicious behavior in the system.
+%
+The actual behavior of such filters, monitors, and other high-assurance components are formally specified by code contracts.
+%
+A code contract is a subset of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
+%
+A component's code contract states assumptions on the component inputs and provides guarantees that define the component outputs in terms of the component's inputs and current state.
+%
+The code contract language itself is Turing complete.
+
+Code contracts are validated with unit tests that are verified with the \agr\ verification engine.
+%
+A unit test in expressed in a test contract in the \agr\ language.
+%
+A test contract is a stylized contract that strengthens the input assumptions of a code contract to a single input and either weakens or leaves unchanged the code contract output guarantees.
+%
+\agr\ then proves whether or not the computation expressed in the code contract satisfies the test contract.
+%
+In this way, a designer is able to test the behavior of the code contract.
+%
+Once the code contract is tested, the \agr\ verification engine proves whether or not the system, with the added high-assurance components, meets its cyber requirements.
+
+Code contracts are automatically synthesized to \ckml.
+%
+We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transition function that only considers the current input and state of the component.
+%
+That state transition function is then rendered in \ckml, and the \ckml\ compiler compiles it to the target binary while also generating the proof of equivalence.
+
+We demonstrated the viability of the approach in real-world industrial design by reporting a case study using the Air Force Research Laboratory's OpenUxAS software.
+%
+The study required a range of filters, monitors, and other mitigation components to meet cyber-requirements.
+%
+The case study demonstrated the viability of the approach to address complex message formats and non-trivial cyber-requirements.
+
+The reported case study was carried out by those that created \brfcs\ with its several tools.
+
+Future work is to conduct another case study involving system designers unfamiliar with the \brfcs\ toolchain.
+%
+We have started such a study with the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter program.
+%
+Early results are encouraging and are revealing ways to expand the expressiveness of the code contract language to make it easier to express iterative computation with arrays.
+
+%%
+% FUTURE WORK
+%%
+
+\subsubsection*{Future work}
+
+Aside from adding expressiveness to the code contract language, future work is to formally verify every transformation step from the code contract to the final \ckml\ function that is compiled to the target binary.
+%
+The HOL4 embedding of the code contract semantics we have discussed is the starting point.  The key remaining steps are:
+%
+\begin{compactitem}
+  \item Automatically add state to the code contract to remove $\konst{pre}(-)$ and prove the initial and modified contracts are equivalent for all input streams
+  \item Create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version
+  \item Create a pure HOL4 function, suitable to \ckml, from the step version of the code contract and prove it equivalent
+\end{compactitem}
+%
+We have automated in HOL4 the transformation to remove $\konst{pre}(-)$ in the first step in what we call \emph{temporal squashing}.
+%
+We have additionally written proofs (manually) for every step for a set of example code contracts.
+%
+From the manual proofs, we have started work on proofs for the first two steps.
+%
+Since these two steps are universal to all input code contracts their proofs need to be generated just once (as opposed to generating them for every component contract).
+%
+The proof for the third step must be generated for each input contract.
+%
+Finally, the entire process needs to be mechanized.
+%
+Other future work includes lifting the proof results to infinite streams.
 

### 2023-02-17 paper/paper.tex
Remove unused figures
Add corrections to intro

@@
 %
-Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-criticalsystems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
-
-Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
+Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.
+
+Unfortunately, systems designers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
+%
 They instead rely on process-oriented checklists and guidelines.
@@
 
-We have been developing the {\em BriefCASE} \oldjunk{toolsuite} toolkit to address this need.
-%
-\brfcs\ is a Model-Based Systems Engineering (MBSE) environment built in the Open Source AADL Tool Environment (OSATE) to add new capabilities for building cyber-resilient systems.
+We have been developing the {\em BriefCASE} toolkit to address this need.
+%
+\brfcs\ is a Model-Based Systems Engineering (MBSE) environment that is integrated into the Open Source AADL Tool Environment (OSATE).
 %
@@
 %
+It adds new capabilities for building cyber-resilient systems.
+%
 In this paper we describe how it facilitates specifying, testing, verifying, and synthesizing high assurance components inserted into a system to improve its cyber-resiliency.
-%
+
 The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
@@
 
-As indicated in the preceding discussion, {\brfcs} is aimed at addressing the lack of tool support for the problem of enforcing and implementing the class of cyber-resilience system requirements.
-%
-Our approach to the problem is embedded in a substantial collection of tools to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
+As indicated in the preceding discussion, {\brfcs} is aimed at addressing the lack of tool support for the problem of enforcing and implementing cyber-resilience system requirements.
+%
+Our approach to the problem is embedded in a collection of tools that together are intended to be wholistic in addressing cyber-resilience at the system level, so it is important to clarify the
 research impact.
@@
 
-Our approach currently applies to common cyber-vulnerabilities, such as buffer overflow, lack of input validation, and supply-chain issues;
-%
-however, other cyber-vulnerabilities such as side-channel attacks and denial of service have not yet been addressed.
+Our approach currently applies to common cyber-vulnerabilities, such as buffer overflow, lack of input validation, and supply-chain issues.
+%
+Other cyber-vulnerabilities such as side-channel attacks and denial of service have not yet been addressed.
 Moreover, this work does not introduce a new type of high-assurance component in terms of capability.
@@
 %
-however, the notions of code and test contracts, along with the system analysis and verified code generation facilities promise to be useful in a wide variety of settings.
+however, the notions of code and test contracts, along with the system analysis and verified code generation facilities should be useful in a wide variety of settings.
 %
@@
 	\centering
-	\includegraphics[width=2\columnwidth]{figs/sw-initial.png}
+	\includegraphics[width=2\columnwidth]{figs/sw-initial.PNG}
 	\caption{Initial software architecture.}

### 2023-02-17 paper/paper.tex
Add corrections to section 2

@@
 The shaded boxes represent our contribution: test contracts, code contracts, and SPLAT synthesis.
-The unshaded boxes are the existing tools that provide the foundation for our contributions.
+The unshaded boxes are the existing elements that provide the foundation for our contributions.
 
@@
 
-\figref{fig:workflow} shows the intent of this behavioral contract definition of the UAV software system on the left-side of the figure at the arrow marked start.
+This behavioral contract definition of the UAV software system is the starting point on the left-side of \figref{fig:workflow}.
 %
@@
 
-\subsection{Adding cyber requirements}
-
-Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the threat analysis stage.
-%
-The UAV must actively protect itself from malicious actors, but to this point, the designers have only been focusing on its core functional behavior with little consideration for cyber-resilience.
-%
-A cyber-vulnerability analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
-%
-Based on the analysis, the designer adds a requirement that the system must guard against malicious AI behavior.
+\subsection{Adding Cyber Requirements}
+
+Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
+%
+Threat analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
+%
+To mitigate the identified supply chain attack, the designer adds a requirement that the system must guard against malicious AI behavior.
 %
@@
 %
-The now arbitrary output from the WM due to input assumption violations lead to the rest of the failures in \figref{fig:example-certificate} as the WM provides the system level outputs.
+The rest of the failures in \figref{fig:example-certificate} are the result of the WM outputs now being arbitrary.
 
@@
 
-\subsection{Adding high assurance components}
+\subsection{Adding High-assurance Components}
 
@@
 %
-An example test from that section the the monitor in our example here is shown in \figref{fig:test}.
+\figref{fig:test} in that section is an example test contract for the added monitor in our example.
 
@@
 %
-Having completed such an iterative for our example system, the \agr\ analysis of the cyber-hardened implementation is shown in \figref{fig:hardened-certificate}.
+The \agr\ analysis of the cyber-hardened implementation at the end of that iterative process is shown in \figref{fig:hardened-certificate}.
 %
@@
 Here \agr\ has generated high-level evidence justifying the claim that the high-assurance components meet their intended purposes.
-%
-Having passed \agr\ verification, the high-assurance components are ready to be synthesized.
 
@@
 
-As \agr\ now proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
+As \agr\ proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
 %

### 2023-02-22 paper/paper.tex
Revise section 3

@@
 %
-In this paper we describe how it facilitates specifying, testing, verifying, and synthesizing high assurance components inserted into a system to improve its cyber-resiliency.
+In this paper we describe how it facilitates specifying, testing, verifying, and synthesizing high-assurance components inserted into a system to improve its cyber-resiliency.
 
@@
 %
-Intuitively, the behavior of a high assurance component is defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contract's input assumptions.
+Intuitively, the behavior of a high-assurance component is defined by the designer in a code contract, the designer then writes test contracts to validate that behavior by asking \agr\ to prove if the code contract ensures the output obligations in the test contract under the test contract's input assumptions.
 %
@@
 
-Code contracts for high assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
+Code contracts for high-assurance components can be arbitrarily complex since they can have persistent state that evolves over time.
 %
@@
 
-Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize \ie, cannot propagate further.
+Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize (\ie, they cannot propagate further).
 %
@@
 
-\subsection{Contracts and verification conditions}
+\subsection{Contracts and Verification Conditions}
 
@@
 
-\agr\ uses ASCII notation for the mathematical syntax we have defined.
-%
-The following table spells out the mapping.
+\agr\ uses ASCII notation for the mathematical syntax we have defined as shown in the following table.
 %
@@
 \begin{tabular}{l|l|l}
- Operation & Math. notation & \agr \\ \hline
+ Operation & Math & \agr \\ \hline
  followed-by &  $e \to e$          & \verb+e -> e+  \\
@@
  variable definition & $v = e$     & \verb+eq v e+ \\
- Previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
- Historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
+ previously  & $\konst{pre}(b)$ & \verb+pre(b)+ \\
+ historically  & $\konst{Hist}(b)$ & \verb+Historically(b)+ \\
 \end{tabular}
@@
 
-\subsection{Formal contract for the example system}
+\subsection{Formal Contract for the Example System}
 \label{sec:sw-contract}
@@
   \begin{center}
-    \scalebox{0.62}{\usebox{\sw}}
+    \resizebox{\columnwidth}{!}{\usebox{\sw}}
   \end{center}
@@
 %
-\begin{itemize}
+\begin{compactitem}
 \item
-\lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}.
-%
-When input is present it is well-formed.
+\lineref{line:sw-assume-1} and \lineref{line:sw-assume-2}:
+%
+when input is present it is well-formed.
 %
@@
 \item
-\lineref{line:sw-assume-3}.
-%
-If there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it arrives after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
+\lineref{line:sw-assume-3}:
+%
+if there is an incoming request, it is either the very first one, \texttt{\textbf{Historically}(\textbf{not} req)}, or it arrives after the component has output a start event in response to a previous request, \texttt{\textbf{Since}(\textbf{not} req, strt)}---to be read as \emph{not
 request since start}.
-\end{itemize}
+\end{compactitem}
 
@@
 %
-\begin{itemize}
+\begin{compactitem}
 \item
-\lineref{line:sw-guarantee-1}, \lineref{line:sw-guarantee-2}, and \lineref{line:sw-guarantee-3}.
-%
-A waypoint coincides with a update on the UAV location, a start coincides with a new waypoint, and waypoints are well-formed.
+\lineref{line:sw-guarantee-1}, \lineref{line:sw-guarantee-2}, and \lineref{line:sw-guarantee-3}:
+%
+a waypoint coincides with a update on the UAV location, a start coincides with a new waypoint, and waypoints are well-formed.
 
 \item
-\lineref{line:sw-guarantee-4}.
-%
-A start happens with a request or one step after a request.
+\lineref{line:sw-guarantee-4}:
+%
+a start happens with a request or one step after a request.
 %
@@
 \item
-\lineref{line:sw-guarantee-5}.
-%
-The alert is signaled if the start does not arrive within the one-step bound.
+\lineref{line:sw-guarantee-5}:
+%
+the alert is signaled if the start does not arrive within the one-step bound.
 %
@@
 %
-\lineref{line:sw-guarantee-6}.
-%
-If an alert has ever happened, it continues to happen.
-\end{itemize}
+\lineref{line:sw-guarantee-6}:
+%
+if an alert has ever happened, it continues to happen.
+\end{compactitem}
 %
@@
 
-\section{Code contracts for high assurance components}
+\section{Code Contracts for High-assurance Components}
 \label{sec:code-contracts}
@@
 %
-\begin{itemize}
+\begin{compactitem}
 \item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
@@
 \item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
-\end{itemize}
+\end{compactitem}
 %
@@
     \begin{tabular}{c}
-      \scalebox{0.62}{\usebox{\flt}} \\
+      \resizebox{\columnwidth}{!}{\usebox{\flt}} \\
     \end{tabular}
@@
     \begin{tabular}{c}
-    \scalebox{0.62}{\usebox{\mntr}} \\
+      \resizebox{\columnwidth}{!}{\usebox{\mntr}} \\
     \end{tabular}
@@
 %
-Such an approach adds extra steps, including test harneses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional validation using tests.
+Such an approach adds extra steps, including test harnesses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional validation using tests.
 %
@@
 %
-Effectively, a code contract for a high assurance component should be a valid implementation of a test scenario.
+Effectively, a code contract for a high-assurance component should be a valid implementation of a test scenario.
 %
@@
     \begin{tabular}{c}
-    \scalebox{0.60}{\usebox{\tst}} \\
+      \resizebox{\columnwidth}{!}{\usebox{\tst}} \\
     \end{tabular}
@@
 %
-The \splt\ tool traverses the system architecture looking for occurrences of high assurance components specified by code contracts;
+The \splt\ tool traverses the system architecture looking for occurrences of high-assurance components specified by code contracts;
 %
@@
     \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\contig}}
+      \resizebox{\columnwidth}{!}{\usebox{\contig}}
     \end{tabular}
@@
     \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\cml}}
+      \resizebox{\columnwidth}{!}{\usebox{\cml}}
     \end{tabular}
@@
     \begin{tabular}{c}
-      \scalebox{0.60}{\usebox{\monFn}}
+      \resizebox{\columnwidth}{!}{\usebox{\monFn}}
     \end{tabular}
@@
 %
-Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high assurance components.
+Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high-assurance components.
 

### 2023-02-23 paper/paper.tex
Revise section 4

@@
 \newcommand{\lineref}[1]{Line~\ref{#1}}
+\newcommand{\defref}[1]{Def.~\ref{#1}}
 \newcommand{\linesref}[2]{Lines~\ref{#1}--\ref{#2}}
@@
 %
-\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment}.}
+\footnote{\agr\ is the acronym for \emph{Assume Guarantee Reasoning Environment}.}
 %
@@
 
-The AGREE specification for the SW component in \secref{sec:example} is given in \figref{fig:sw}.
+The \agr\ specification for the SW component in \secref{sec:example} is given in \figref{fig:sw}.
 %
@@
 %
-The well-formed predicates are defined elsewhere using AGREE functions.
+The well-formed predicates are defined elsewhere using \agr\ functions.
 
@@
 
+%%
+% Code contracts for High-assurance Components
+%%
+
 \section{Code Contracts for High-assurance Components}
@@
 
-Generally, AGREE contracts do not describe the computation that a component performs.
-%
-This is entirely by design: AGREE is intended to be declarative and reasoning about component behavior takes place solely at the specification level.
+\agr\ contracts do not describe the computation that a component performs by design:
+%
+it is intended to be declarative so it reasons about component behavior at the specification level.
 %
@@
 %
-However, the syntax of AGREE provides enough expressiveness to support the notion of a \emph{code contract}: a contract from which an implementation can be extracted.
+However, the syntax of \agr\ provides enough expressiveness to support the notion of a \emph{code contract}: a contract from which an implementation can be extracted.
 %
@@
 \begin{definition}[Code contract]
-A leaf component of the form $(I,O,A,P,\emptyset)$ is a \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where \[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set of \konst{eq} statements and $G$ is the set of output guarantees, one for each element of $O$.
-%
-In the interpretation as code, the order of elements of $\mathit{Eqs}$ is important, and is simply taken to be the occurrence order of the \konst{eq} statements in the syntax.
-%
-Thus we will work with the \emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
+A leaf component of the form $(I,O,A,P,\emptyset)$ is a \emph{code contract} if $(\mathit{Eqs} \cup G) \subseteq P$, where \[\mathit{Eqs} = [v_1 = e_1, \cdots , v_n = e_n] \] is a non-empty list of \konst{eq} statements and $G$ is the set of output guarantees, one for each element of $O$.
+%
+In the interpretation as code, the order of elements of $\mathit{Eqs}$ is important and is simply taken to be the occurrence order of the \konst{eq} statements in the syntax.
 \end{definition}
 
-The semantics of \secref{sec:agree-semantics} supports a formal connection between the original contract, and verification results of the AGREE model checker being run on it, and code generated from specifications of leaf level cyber-components.
-%
-It also provides the root meaning at the base of a chain of translation steps moving from an AGREE contract to a CakeML executable.
+The semantics of \secref{sec:agree-semantics} give a formal connection between the system contract, the verification results of the \agr\ model checker, and the code synthesized for leaf level components.
+%
+It thus provides the root meaning at the base of a chain of translation steps moving from a code contract to a \ckml\ executable.
 
@@
 \[
-\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
+(\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t) \imp \sem{P}^{E'}_t
 \]
 %
-\end{definition}
-
-How does this notion of correctness integrate with the verification conditions generated by AGREE?
-%
-In \emph{contract correctness}, the model checker is proving the following property
+\end{definition}\label{def:correct}
+
+How does this notion of correctness integrate with the verification conditions generated by \agr?
+%
+In \agr, the model checker is proving the following property
 %
@@
 %
-and in \emph{code contract correctness} we essentially prove a Hoare triple of the form
+and in \defref{def:correct} we essentially prove a Hoare triple of the form
 %
@@
 %
-This accords with intuition, namely that AGREE is a kind of \emph{program-free program logic};
+This accords with intuition, namely that \agr\ is a kind of \emph{program-free program logic};
 %
@@
 %
-In order to use the code contract result in system contract verification conditions, one has to show that the code contract result implies the system verification condition.
-
-%%
-% IMPERATIVE CODE CONTRACTS
-%%
-
-\subsubsection*{Imperative code contracts}
-
-Stream-based evaluation allows values arbitrarily \emph{deep in the past} to be accessed by use of $\konst{pre}(-)$.
-%
-However, the programming languages targeted by SPLAT do not directly support streams.
-%
-We are developing a translation to eliminate such deep temporal accesses by the introduction of intermediate variables.
-%
-The essential property the translation needs to satisfy is a syntactic restriction, call it \emph{Pascal-like}, that supports easy translation to standard imperative languages.
-%
-The guiding intuition is that a Pascal-like {\Eqs} accesses at most one step in the past;
-%
-also, accesses to a past value of a variable are only possible until the variable is updated.
-%
-In informal terms, this means that ${\Eqs}\cdot G$ in a code contract has the following properties:
+In order to use a code contract's computed result in verification, we have to show it's defined computation implies the system verification condition as stated by \defref{def:correct}.
+
+\paragraph{Rewriting code contracts to be imperative:\/} stream evaluation allows values arbitrarily deep in the past to be accessed by use of $\konst{pre}(-)$, but streams, and accesses to past variable values, are not part of \ckml\ and most other programming languages.
+%
+As such, we automatically rewrite code contracts to eliminate deep temporal accesses by the introduction of intermediate variables.
+%
+The essential property of the rewrite is a syntactic restriction that supports easy translation to standard imperative languages.
+%
+The guiding intuition is that accesses to values by expressions in {\Eqs} are at most one step in the past and accesses to such past values are only possible until the associated variable is updated.
+%
+In informal terms, this means that ${\Eqs} \cdot G$ in a code contract has the following properties:
 %
 \begin{compactitem}
-\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
+\item the followed-by operator only occurs at the top level of a variable definition, \eg, $v = e_1 \to e_2$;
 \item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
@@
 \end{compactitem}
-%
-For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be defined, albeit in a manner that is not Pascal-like, by
+
+For example, the Fibonacci sequence is computed by
 %
@@
 %
-\noindent or by (again this is not Pascal-like)
+\noindent and with a temporary variable can be rewritten to
 %
@@
   N = 0 -> 1 + pre(N)
-  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
+  Fib = if N <= 1 then 1
+         else pre(Fib) + pre(pre(Fib))
 \end{lstlisting}
@@
 %
-\noindent or by (this time Pascal-like):
+\noindent and finally, with a few more temporary variables, can be rewritten to
 %
@@
 %
-Please note that the initial values of \verb+F1+ and \verb+F2+ are never accessed, so 42 could be replaced by any number in their defining equations.
-%
-A Pascal-like ${\Eqs} \cdot G$ is straightforward to implement in conventional programming languages, as we will see in \secref{sec:synthesis}.
-
-\subsection{Code contracts for the hardened system}
+\noindent This final version of the computation is imperative as the instances of $\konst{pre}$ can be resolved with the current values of the referenced variables:
+%
+the instances appear before the associated variables are updated when following a sequential order of evaluation.
+%
+Please note that the initial values of \verb+F1+ and \verb+F2+ are never accessed: 42 could be replaced by any number in their defining equations.
+%
+An imperative ${\Eqs} \cdot G$ is straightforward to implement in conventional programming languages, as we will see in \secref{sec:synthesis}.
+
+\subsection{Code Contracts for the Hardened System}
 \label{sec:ha-contracts}
@@
 
-The monitor informally behaves as follows: if the \emph{policy} is obeyed in the current step, and the alert has not yet been raised, then the \texttt{Response} input is sent out on \texttt{Output}, and the \texttt{Alert} output stays low \lineref{line:mon-alert}.
-%
-Otherwise \texttt{Alert} goes high and stays high, and there is nothing sent henceforth on \texttt{Output}.
-
-The \texttt{alert} variable (distinct from the \texttt{Alert} port) depends not just on the policy but also on the \texttt{is\_latched} constant, which has been set at configuration time.
+The monitor informally behaves as follows: if the \emph{policy} is obeyed in the current step, and the alert has not yet been raised, then the \texttt{Response} input is sent out on \texttt{Output}, and the \texttt{Alert} output stays low \lineref{line:mon-alert};
+%
+otherwise, \texttt{Alert} goes high and stays high, and there is nothing sent henceforth on \texttt{Output}.
+
+The \texttt{alert} variable (distinct from the \texttt{Alert} port) depends not just on the policy but also on the constant \texttt{is\_latched}, which has been set at configuration time.
 %
@@
 %
-The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response, \ie, \texttt{isPending} holds when \emph{there is no response since the most recent request without a corresponding response}.
-%
-The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.
-Note that the \texttt{preIsPending} variable is a temporary needed to make the contract Pascal-like;
-%
-otherwise, the current and previous value of \texttt{isPending} would be simulataneously accessed in the definition of \texttt{policy}.
-
-The policy (\lineref{line:mon-policy}) makes sure the latency is bounded whenever there is a pending request, and it ensures that a response is a result of a request.
-%
-The initial step differentiates from later steps in only needing to be sure a response is a result of a request, both happening in the initial step.
+The \texttt{isPending} variable is true whenever there is a request that does not coincide with a response, \ie, \texttt{isPending} holds when there is no response since the most recent request without a corresponding response.
+%
+The \texttt{latency} variable starts at zero, resets on every request, and otherwise increments by one.  Note that the \texttt{preIsPending} variable is a temporary that makes the contract imperative;
+%
+otherwise, the current and previous value of \texttt{isPending} would be simultaneously accessed in the definition of \texttt{policy}.
+
+The policy (\lineref{line:mon-policy}) ensures the latency is bounded whenever there is a pending request and that a response is a result of a request.
+%
+The initial step differentiates from later steps in only needing to be sure that a response is a result of a request with both happening in that initial step.
 
@@
 %
-In fact, the policy can be written without the assumption, but the policy is much simpler to write with the assumption.
+Although the policy can be written without the assumption, it is much simpler to write with the assumption.
 
@@
 %
-Synthesis must generate faithful implementations of the code contracts for the AGREE verification result to apply to the deployed system.
+Synthesis must generate faithful implementations of the code contracts for the \agr\ verification result to apply to the deployed system.
 
@@
 
-\section{Synthesis to CakeML}
+\section{Synthesis to \ckml}
 \label{sec:synthesis}
@@
 %
-for each such occurrence it generates a CakeML program.
+for each such occurrence it generates a \ckml\ program.
 %
@@
 
-\paragraph{Regular expressions:\/} often the well-formedness of data on connections can be enforced by matching against a regular expression.
+\paragraph{Regular expressions:\/} often the well-formed property of data on connections can be enforced by matching against a regular expression.
 %
@@
 
-When full computational power is needed, we can use the expressive power of AGREE code contracts to write arbitrarily complex filter or monitor operations, as we have already seen.
-%
-Our formal model of AGREE syntax and semantics is applied to extract code contracts and map them to equivalent logic functions.
+When full computational power is needed, we can use the expressive power of \agr\ code contracts to write arbitrarily complex filter or monitor operations, as we have already seen.
+%
+Our formal model of \agr\ syntax and semantics is applied to extract code contracts and map them to equivalent logic functions.
 %
@@
 %
-Finally, the CakeML compiler can be applied to the program.
+Finally, the \ckml compiler can be applied to the program.
 %
@@
 \item Add parsers and printers (specified by contiguity types) for input and output.
-\item Map to CakeML and add calls to input/output library routines via the CakeML foreign function interface.
-\item Invoke the CakeML compiler.
+\item Map to \ckml\ and add calls to input/output library routines via the \ckml\ foreign function interface.
+\item Invoke the \ckml\ compiler.
 \end{compactenum}
@@
 %
-On the other hand, a filter may need to parse the input buffer into a data structure specified in AGREE and apply a user-defined \emph{wellformedness} property, also specified in AGREE, to the data.
+On the other hand, a filter may need to parse the input buffer into a data structure specified in \agr\ and apply a user-defined \emph{wellformedness} property, also specified in \agr, to the data.
 %
@@
   \end{center}
-  \caption{Synthesized CakeML for the filter.}
+  \caption{Synthesized \ckml\ for the filter.}
   \label{fig:filter-cakeml}
@@
   \end{center}
-  \caption{Synthesized CakeML for the monitor.}
+  \caption{Synthesized \ckml\ for the monitor.}
   \label{fig:monitor-cakeml}
@@
 %
-Although we are satisfied with the results of the formal verification using AGREE, we have not yet analyzed the design for cyber-vulnerabilities.
+Although we are satisfied with the results of the formal verification using \agr, we have not yet analyzed the design for cyber-vulnerabilities.
 

### 2023-02-23 paper/paper.tex
Revise section 6

@@
 %
-Case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
+Our case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
 
@@
 %
-Here the test contract assumptions strengthen the code contract assumptions to a single input, and the test contract guarantees weaken, or leave unchanged, the code contract output for that test input, depending on the goal of the test.
+Here the test contract assumptions strengthen the code contract assumptions to a single input, and the test contract guarantees weaken, or leave unchanged, the code contract output guarantees for that test input, depending on the goal of the test.
 %
@@
 \begin{compactitem}
-\item \lineref{line:tst-index}: An index to use to define the test input and expected test output through time.
-\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: The test input. These define when the request and response message events should appear relative to the index.
-\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: The expected test output.
+\item \lineref{line:tst-index}: an index to use to define the test input and expected test output through time.
+\item \lineref{line:tst-assume-1} and \lineref{line:tst-assume-2}: the test input. These define when the request and response message events should appear relative to the index.
+\item \lineref{line:tst-guarantee-1} and \lineref{line:tst-guarantee-2}: the expected test output.
 The test additionally proves that nothing happens after the initial test input and expected output, \ie, no more input is provided and no more output is generated.
-\item \linesref{line:tst-imp-start}{line:tst-imp-end}: The implementation with the component under test (\lineref{line:tst-imp-comp}).
+\item \linesref{line:tst-imp-start}{line:tst-imp-end}: the implementation with the component under test (\lineref{line:tst-imp-comp}).
 \end{compactitem}
@@
 %
-They enable designers to iteratively develop code contracts inside OSATE with \agr, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
+They enable designers to iteratively develop code contracts with \agr\ in the modeling environment, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
 %
@@
 %
-We give a short review of that previous work before going into detail on the synthesis.
+We give a short review of our previous work before going into detail on the synthesis.
 
@@
 
-\paragraph{Contiguity types\cite{contiguity-types}:\/} contiguity types, similar to regular expressions, are effective at describing data at the network message level, \eg, as flat strings.
-%
-The notation is however much more expressive, being able to declare many kinds of so-called \emph{self-describing} messages, \ie, those where message structure is dependent on data held elsewhere in the message.
+\paragraph{Contiguity types:\/} contiguity types, \cite{contiguity-types}, similar to regular expressions, are effective at describing data at the network message level, \eg, as flat strings.
+%
+The notation is however much more expressive than regular expressions, being able to declare many kinds of so-called \emph{self-describing} messages, \ie, those where message structure is dependent on data held elsewhere in the message.
 %
@@
 %
-We have formalized and verified a contiguity type matcher, which has been extensively applied throughout our case studies.
+We have formalized and verified a contiguity type matcher, and that matcher has been extensively applied throughout our case studies.
 
@@
 %
-Now suppose ${\Eqs}\cdot G$ is \emph{Pascalish}, \ie, sequentially imperative.
+Now suppose ${\Eqs}\cdot G$ is sequentially imperative.
 %
@@
 %
-Finally, the \ckml compiler can be applied to the program.
+Finally, the \ckml\ compiler can be applied to the program.
 %
@@
 \item Map from specification to computable function in logic, producing a proof of equivalence.
-\item Add parsers and printers (specified by contiguity types) for input and output.
+\item Add parsers and printers (specified by contiguity types) for input/output.
 \item Map to \ckml\ and add calls to input/output library routines via the \ckml\ foreign function interface.
@@
 %
-A filter has one input port and one output; messages on the input that the filter policy admits pass unchanged to the output port;
+A filter has one input and one output; messages on the input that the filter policy admits pass unchanged to the output;
 %
@@
 %
-For example, we have used the expressive power of regular expressions and Contiguity Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints on GPS coordinates in UxAS messages.
-%
-On the other hand, a filter may need to parse the input buffer into a data structure specified in \agr\ and apply a user-defined \emph{wellformedness} property, also specified in \agr, to the data.
-%
-Arbitrarily complex wellformedness checks can be made in this way.
+For example, we have used regular expressions and Contiguity Types to enforce bounds on GPS coordinates in UxAS messages.
+%
+On the other hand, a filter may need to parse the input buffer into a data structure specified in \agr\ and apply a user-defined property, also specified in \agr, to the data.
+%
+Arbitrarily complex checks can be made in this way.
 %
@@
 \item
-The filter checks to see if there is any input available.
+the filter checks to see if there is any input available.
 %
@@
 
-\item The input is read (and parsed if need be);
-
-\item The wellformedness predicate is evaluated on the input;
-
-\item If the predicate returns \konst{true} then the input buffer
+\item the input is read (and parsed if need be);
+
+\item the wellformedness predicate is evaluated on the input;
+
+\item if the predicate returns \konst{true} then the input buffer
  is copied to the output, otherwise no action is taken; and
 
-\item The filter yields control.
+\item the filter yields control.
 \end{compactenum}
@@
 
-The contiguity type specification and wellformedness predicate for the filter are shown in \figref{fig:filter-spec} and the synthesized \ckml\ code is in \figref{fig:filter-cakeml}.
+The synthesized \ckml\ code for \figref{fig:filter-spec} is in \figref{fig:filter-cakeml}.
 %
@@
 
-As discussed above, a monitor specification is mapped to a state transformation function
+As discussed previously, the code contract specification for the monitor is mapped to a state transformation function
 %
@@
 
-\item Each input buffer is parsed into data of the type specified by the port type;
+\item each input buffer is parsed into data of the type specified by the port type;
 
 \item
-New values for the state variables are computed, in dependency order.
+new values for the state variables are computed, in dependency order.
 %
@@
 
-\item Values of the outputs are computed;
-
-\item Outputs are written and the new state is written;
-
-\item The monitor yields control.
+\item values of the outputs are computed;
+
+\item outputs are written and the new state is written;
+
+\item the monitor yields control.
 \end{compactenum}
-
-The \konst{stepFn} for the monitor of the example described in \secref{sec:example} is displayed in \figref{fig:monitor-cakeml}.
+\noindent The \konst{stepFn} for the monitor in the example described in \secref{sec:example} is displayed in \figref{fig:monitor-cakeml}.
 
@@
 
-\subsection{Incorporation into system execution}
+\subsection{Incorporation into System Execution}
 
@@
 
-\paragraph{Partiality:\/} partiality is an important consideration for both filters and monitors: the input data might not parse or the computation of a filter could be badly written and fail at runtime.
+\paragraph{Partiality of the Correctess Argument:\/} partiality is an important consideration for both filters and monitors as the input data might not parse or the computation of a filter could be badly written and fail at runtime.
 %
@@
 %
-For these failure cases, the filter is behaving as it should, but we must also guard against situations in which a \emph{correctly specified} filter or monitor fails in the runtime.
-%
-This kind of defect arises when a filter, for example, \emph{ought} to accept a message, but lack of resources results in the filter failing to do so.
+For these failure cases, the filter is behaving as it should, but we must also guard against situations in which a correctly specified filter or monitor fails in the runtime.
+%
+This kind of defect arises when a filter, for example, ought to accept a message, but lack of resources results in the filter failing to do so.
 %

### 2023-02-23 paper/paper.tex
Revise section 7

@@
 %
-We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from untrusted components.
+We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from uncontrolled components.
 %
@@
 %
-OpenUxAS, as an open source product, is considered \emph{untrusted}.
-%
-The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
+OpenUxAS, as an open source product, is considered \emph{uncontrolled}.
+%
+The UAV system is thus transformed to be resilient to malicious behavior that may arise in the uncontrolled component.
 %
@@
 %
-This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered untrusted.
+This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered uncontrolled.
 %
@@
 %
-That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
+That means that the high-assurance components protect the WM from the uncontrolled AI generating malformed data, spontaneous responses, or delayed responses.
 %
@@
 
-\item the wellformedness predicate is evaluated on the input;
+\item the predicate is evaluated on the input;
 
@@
 %
-The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides designers with early validation of their model;
-%
-produces early information about cyber-vulnerabilities;
-%
-provides effective mitigations;
+The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
+%
+Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
+%
+produces early information about cyber-vulnerabilities,
+%
+provides effective mitigations,
 %
@@
 %
-Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.
+Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to constraints specified in the Operating Region message.
 %
@@
 %
-For example, the vulnerbility analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
+For example, the vulnerability analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.
 
@@
 %
-These include four wellformedness requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
+These include four integrity requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
 %
@@
 
-The intent of the wellformedness requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
+The intent of the data integrity requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
 %
@@
 %
-The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.
-
-The filter code contracts for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
+We therefore inserted filter components on the incoming and outgoing UxAS connections.
+
+The filter code contracts for the four UxAS connections are similar in that they check that record values contained in the messages are within appropriate ranges.
 %
@@
 %
-The Monitor transform is applied for this class of mitigation.
-%
-We add a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
-%
-And we add a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
-%
-The Geofence Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
-%
-The Geofence Monitor code contract is shown in \figref{fig:geofence-monitor}.
+We therefore add a response monitor (Resp\_Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+%
+And we add a geo-fence monitor (Geofence\_Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+%
+The Geofence\_Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
+%
+Its code contract is shown in \figref{fig:geofence-monitor}.
 
@@
 
-So far, we have addressed requirements that mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
+So far, we have addressed requirements that mitigate vulnerabilities related to malformed messages and malicious behavior on-board the UAV.
 %
@@
 %
-The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
+As such we added two components to the UAV software: the Attestation Manager for evaluating remote systems like the Ground Station, and the Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
 %
@@
 %
-The code contract for the Attestation Gate is completely generated by the transform since its behavior is specific to attestation.
+The code contract for the Attestation Gate is generated by \brfcs\ since its behavior is specific to attestation.
 
@@
 %
-The shaded components (green) were added by the transform with the exception of the UxAS component (red).
-
-After unit testing the monitors, we formally verify the model with \agr\ to show that all of the component contracts are satisfied.
-%
-We then run \splt\ to produce provably-correct code.
-%
-The synthesized code is output to a directory in the build file system with the location specified for each component in the model.
-%
-The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.
+The shaded components (green) were added by \brfcs\ to harden the system with the exception of the UxAS component (red).
+
+After unit testing the monitors, we formally verified the model with \agr\ to show that all of the component contracts were satisfied.
+%
+We then ran \splt\ to produce provably-correct code.
 
@@
 
-Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by \splt\ or other means, the system can be built and deployed.
-%
-We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
-%
-HAMR then compiles the software to execute on seL4.
-
+The demonstration system was then assembled for testing.
+%
+\brfcs\ includes the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
+%
+HAMR was used here to target the verified micro-kernel seL4.
+%
 The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
@@
 The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
-%
-The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
+
+The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the constraints in the Operating Region message.
 %
 Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior.
-In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
-%
-We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.
+In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation.
+
+We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigation in order to verify the effectiveness of the malicious code.
 A third party evaluator repeated the tests on the hardened system and demonstrated that our mitigation strategy was successful.
-%
 The following scenarios were exercised:
 
-\paragraph{Infected Ground Station:\/} An application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
+\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
 %
@@
 %
-The Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
-%
-The wellformedness filter for Line Search Task messages prevented this message from reaching the UxAS.
-
-\paragraph{UxAS vulnerability exploit:\/} Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
+the Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
+%
+The filter for Line Search Task messages prevented this message from reaching the UxAS.
+
+\paragraph{UxAS vulnerability exploit:\/} tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
 %
@@
 
-\paragraph{UxAS trojan modifies flight plan:\/} A trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
+\paragraph{UxAS trojan modifies flight plan:\/} a trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
 %
@@
 %
-We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the OSATE environment and an external test platform.
+We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the \brfcs\ environment and an external test platform.
 %

### 2023-02-23 paper/paper.tex
Revise section 8

@@
 %
-We therefore add a response monitor (Resp\_Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
-%
-And we add a geo-fence monitor (Geofence\_Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+We therefore add a response monitor (Resp Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+%
+And we add a geo-fence monitor (Geofence Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
 %
@@
 %
-Growing the expressiveness of the language arround arrays is something to investigate in future work.
+Growing the expressiveness of the language around arrays is something to investigate in future work.
 
@@
 %
-We have been able to work effectively with a relatively simple core language, \eg, without \emph{clocking} features.
+We have been able to work effectively with a relatively simple core language, \eg, without clocking features.
 %
@@
 %
-In that regard, the code contracts are the implementation, but unlike Dafny, \ckml\ provide a verified compilation path to binaries.
+In that regard, the code contracts are the implementation, but unlike Dafny, \ckml\ provides a verified compilation path to binaries.
 

### 2023-02-23 paper/paper.tex
Revise section 9

@@
 %
-Aside from the HOL4 embedding of the code contract semantics and not requiring any HOL4 working knowledge, the work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high-assurance components.
+The work in this paper differs in that it is targeting a fully verified path from the code contract all the way to the binaries leveraging the \ckml\ compiler specifically to provide stronger guarantees on the high-assurance components.
 
@@
 %
-We have discussed transformations provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
+We have discussed tools provided by \brfcs\ to insert, create, and synthesize high-assurance components such as filters and monitors to improve the cyber-resiliency of the system design.
 
@@
 %
-A code contract is a subset of the \agr\ language with formal semantics embedded in HOL4 for formal reasoning.
+A code contract is a subset of the \agr\ language with formal semantics for formal reasoning.
 %
@@
 %
-A unit test in expressed in a test contract in the \agr\ language.
+A unit test is expressed in a test contract in the \agr\ language.
 %
@@
 
-Code contracts are automatically synthesized to \ckml.
-%
 We discussed the \splt\ synthesis engine that iteratively transforms the code contract defined over streams of data into a state transition function that only considers the current input and state of the component.
@@
 The case study demonstrated the viability of the approach to address complex message formats and non-trivial cyber-requirements.
-
-The reported case study was carried out by those that created \brfcs\ with its several tools.
-
-Future work is to conduct another case study involving system designers unfamiliar with the \brfcs\ toolchain.
-%
-We have started such a study with the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter program.
+%
+This reported case study was carried out by those that created \brfcs\ with its several tools.
+
+%
+
+%%
+% FUTURE WORK
+%%
+
+\paragraph{Future work:\/}
+
+we are conducting another case study involving system designers unfamiliar with the \brfcs\ toolchain.
+%
+This study is with the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter program.
 %
@@
 
-%%
-% FUTURE WORK
-%%
-
-\subsubsection*{Future work}
-
 Aside from adding expressiveness to the code contract language, future work is to formally verify every transformation step from the code contract to the final \ckml\ function that is compiled to the target binary.
 %
-The HOL4 embedding of the code contract semantics we have discussed is the starting point.  The key remaining steps are:
-%
-\begin{compactitem}
-  \item Automatically add state to the code contract to remove $\konst{pre}(-)$ and prove the initial and modified contracts are equivalent for all input streams
-  \item Create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version
-  \item Create a pure HOL4 function, suitable to \ckml, from the step version of the code contract and prove it equivalent
-\end{compactitem}
+The code contract semantics are embedded in HOL4 already.  The key remaining steps are:
+%
+\begin{compactenum}
+  \item automatically add state to the code contract to remove $\konst{pre}(-)$ and prove the initial and modified contracts are equivalent for all input streams;
+  \item create a step based version of the modified code contract that does not use streams and prove it equivalent to the stream version; and
+  \item create a pure HOL4 function, suitable to \ckml, from the step version of the code contract and prove it equivalent.
+\end{compactenum}
 %
@@
 %
-We have additionally written proofs (manually) for every step for a set of example code contracts.
-%
-From the manual proofs, we have started work on proofs for the first two steps.
-%
-Since these two steps are universal to all input code contracts their proofs need to be generated just once (as opposed to generating them for every component contract).
-%
-The proof for the third step must be generated for each input contract.
-%
-Finally, the entire process needs to be mechanized.
+We have additionally written HOL4 proofs for every step above for a set of examples.
+%
+From these proofs, we have started work on automatically generating proofs for any input.
+%
+The first two steps above are universal to all input code contracts so their proofs need to be generated just once (as opposed to generating them for every component contract).
+%
+The proof for the third step must be automatically generated for each input contract.
 %

### 2023-02-23 paper/paper.tex
Fix overful boxes

@@
 \begin{abstract}
-Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
+Safety critical systems such as avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
 %
@@
 \section{Introduction} \label{sec:intro}
-In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
+In recent years, stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
 %
@@
 
-Unfortunately, systems designers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
+Unfortunately, systems designers are given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
 %
@@
 
-The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
+The main concept is that of \emph{security-improving transform} through the insertion of new components aimed at mitigating cyber-vulnerabilities.
 %
@@
 
-The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such high-assurance components meet their intended purpose and are properly implemented.
+The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such components meet their intended purpose and are properly implemented.
 %
@@
 
-We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
+We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system.
+%
+That system uses the Air Force Research Laboratory's OpenUxAS services for route planning.
 %
@@
 
-Our approach currently applies to common cyber-vulnerabilities, such as buffer overflow, lack of input validation, and supply-chain issues.
+Our approach currently applies to common vulnerabilities such as buffer overflow, lack of input validation, and supply-chain issues.
 %
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
   \begin{center}
@@
 
-The expected behavior of the SW system, and the components implementing the system, are modeled with \agr\ contract specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
+The behavior of the SW system, and the components implementing the system, are modeled with \agr\ contract specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
 %
@@
   \item the start coincides with a waypoint; and
-  \item any further waypoints coincide with location updates.
+  \item further waypoints coincide with location updates.
 \end{compactitem}
 
-This behavioral contract definition of the UAV software system is the starting point on the left-side of \figref{fig:workflow}.
+This behavioral definition of the UAV software system is the starting point on the left-side of \figref{fig:workflow}.
 %
@@
 
-Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
+After \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
 %
@@
 
-As \agr\ proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
+As \agr\ verifies the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
 %
@@
 
-A system implementation is described by its ports, its subcomponents, and the connections between ports on different subcomponents.
+A system implementation is described by its ports, its sub-components, and the connections between ports on different sub-components.
 %
@@
 
-Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize (\ie, they cannot propagate further).
-%
-Therefore, subcomponent contracts, and thus the top-level model, must be acyclic.
+Data flows through the sub-components of a system in dependency order: the inputs propagate to outputs through the contracts until everything stabilizes.
+%
+Therefore, sub-component contracts, and thus the top-level model, must be acyclic.
 %
@@
 
-A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are subcomponents.
+A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are sub-components.
 %
@@
 %
-A subcomponent $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
+A sub-component $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
 %
@@
 %
-Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$ binding variable names to value streams, the semantics $\sem{-}^E_t$ of terms and formulas defines the meaning of compound syntax in terms of the meaning of subexpressions.
+Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$ binding variable names to value streams, the semantics $\sem{-}^E_t$ of terms and formulas defines the meaning of compound syntax in terms of the meaning of sub-expressions.
 %
@@
 %
-the alert is signaled if the start does not arrive within the one-step bound.
+the alert is signaled if the start does not arrive within one-step.
 %
@@
 
-How does this notion of correctness integrate with the verification conditions generated by \agr?
+How does this definition of code contract correctness integrate with the verification conditions generated by \agr?
 %
@@
 \begin{lrbox}{\flt}
-  \begin{lstlisting}[style=agree,numbers=left] -- start user
+  \begin{lstlisting}[style=agree,numbers=left]
     -- start user definitions
@@
 
-As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
+\figref{fig:hardened-certificate} shows the hardened system to pass AGREE verification.
 %
@@
 
-Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
+Test contracts are a tool to help detecting vacuity, under-specification, and missed corner cases.
 %
@@
 
-\paragraph{Regular expressions:\/} often the well-formed property of data on connections can be enforced by matching against a regular expression.
+\paragraph{Regular expressions:\/} often the well-formed property of data can be enforced by matching against a regular expression.
 %
@@
 %
-Then we can direcctly create a logic function
+Then we can directly create a logic function
 %
@@
 \item Map from specification to computable function in logic, producing a proof of equivalence.
-\item Add parsers and printers (specified by contiguity types) for input/output.
+\item Add needed parsers and printers (specified by contiguity types) for input/output.
 \item Map to \ckml\ and add calls to input/output library routines via the \ckml\ foreign function interface.
@@
 
-The synthesized \ckml\ code for \figref{fig:filter-spec} is in \figref{fig:filter-cakeml}.
+The synthesized code for \figref{fig:filter-spec} is shown in \figref{fig:filter-cakeml}.
 %
@@
 %
-provides effective mitigations,
+provides effective mitigation,
 %
@@
 
-We have modeled the system architecture of the UAV in AADL.
+We modeled the system architecture of the UAV in AADL.
 %
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
 	\centering
@@
 
-The intent of the data integrity requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
+The data integrity requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
 %
@@
 
-The filter code contracts for the four UxAS connections are similar in that they check that record values contained in the messages are within appropriate ranges.
+The code contracts for these filters are similar in that they check that values contained in the messages are within appropriate ranges.
 %
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 %%
-% DISSCUSSION
+% DISCUSSION
 %%
@@
 
-Code contracts are validated with unit tests that are verified with the \agr\ verification engine.
+Code contracts are validated with unit tests that are verified by \agr.
 %

### 2023-02-24 paper/paper.tex
Add minor corrections

@@
 %
-The verified compiler, with the semantics preserving transforms, ensures that the high-assurance component adheres to its contract specification.
+The verified compiler, with the semantics preserving transforms, ensure that the high-assurance component adheres to its contract specification.
 %
@@
 %
-That requirement is encoded in the assumption (\lineref{line:mon-assume}) and is the same assumption used for the system input.
+That requirement is encoded in an assumption (\lineref{line:mon-assume}) and is the same assumption used for the system input.
 %
@@
 %
-Our case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
+Our case studies have shown that designers, as well as formal methods practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.
 
@@
 %
-Under-specification is very challenging, as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
+Under-specification is very challenging as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
 %
@@
 
-A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with traditional testing.
-%
-Such an approach adds extra steps, including test harnesses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional validation using tests.
+A tempting approach to exercising code contracts directly is to use synthesis to create the target binary and then use the target runtime with traditional testing.
+%
+Such an approach adds extra steps, including test harnesses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional test validation.
 %
@@
 
-Test contracts are a tool to help detecting vacuity, under-specification, and missed corner cases.
+Test contracts are a tool to help in detecting vacuity, under-specification, and missed corner cases.
 %
@@
 %
-In this paper, we extend this \emph{specification oriented} approach to the Turing-complete specifications of code contracts in \agr.
+In this paper, we extend this specification oriented approach to the Turing-complete specifications of code contracts in \agr.
 %
@@
 
-\subsection{Compiling from code contracts}
+\subsection{Compiling from Code Contracts}
 
@@
 %
-It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
+It is an obligation on the system that the scheduler follows some sensible total order of component activation and allows each component sufficient time for its computation.
 %
@@
 
-\paragraph{Partiality of the Correctess Argument:\/} partiality is an important consideration for both filters and monitors as the input data might not parse or the computation of a filter could be badly written and fail at runtime.
-%
-In such cases, the filter should recover and yield control without passing the input onwards.
-%
-For these failure cases, the filter is behaving as it should, but we must also guard against situations in which a correctly specified filter or monitor fails in the runtime.
+\paragraph{Partiality of the Correctess Argument:\/} partiality is an important consideration for both filters and monitors as the input data might not parse or the computation of a component could be badly written and fail at runtime.
+%
+In such cases, the component should recover and yield control without passing the input onwards.
+%
+For these failure cases, the component is behaving as it should, but we must also guard against situations in which a correctly specified filter or monitor fails in the runtime.
 %
@@
 %
-Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
+Another motivation was to determine whether or not the technology provides designers with early validation of their models,
 %
@@
 %
-We perform a formal analysis with \agr\ to verify that the model satisfies the contracts.
-%
-For the initial version of our design, the verification passes.
+We performed a formal analysis with \agr\ to verify that the model satisfies the contracts.
+%
+For the initial version of our design, the verification passed.
 %
@@
 %
-These include four integrity requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
-%
-Requirements are imported into the model as goals in the assurance case.
+These include four well-formed requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
+%
+Requirements were imported into the model as goals in the assurance case.
 %
@@
 
-The data integrity requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
+The well-formed data requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
 %
@@
 %
-By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
+By placing filters on the connections upstream of mission-critical components, such attacks are mitigated.
 %
@@
 %
-This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
-%
-We therefore add a response monitor (Resp Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
-%
-And we add a geo-fence monitor (Geofence Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
+This required adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
+%
+We therefore added a response monitor (Resp Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
+%
+And we added a geo-fence monitor (Geofence Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
 %
@@
 %
-\brfcs\ includes the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
-%
-HAMR was used here to target the verified micro-kernel seL4.
+\brfcs\ includes the tools to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
+%
+These tools were used here to target the verified micro-kernel seL4.
 %
@@
 
-\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
+\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station to send the UAV on a mission-violating trajectory.
 %
